### 2016-06-15 inst/doc/Exploration.Rmd
v0.1.9011 Doc reorganization

---
title: "Uppangala trees functionnal traits : Exploration"
author: "Sylvain Schmitt"
date: '`r Sys.Date()`'
output: html_document
fig_caption: yes
vignette: |
  %\VignetteIndexEntry{Uppangala PFT} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

# Traits

## Data cleaning

### Load data

Data loading from gen functions, that open data from raw tables in `inst/extdata`

```{r Load data}
rm(list = ls()); invisible(gc())
library(Uppangala)
PFT <- genPFT()
captions <- genCaptions()
```

#### Uncorrectable errors

Errors in PFT values :

* Wood density > 1, being Gummi gutta sample P5.122 to small to measure a correct wood density
* Thickness outliers due to veins but correct entered data

### Outlier research

```{r variances}
par(mfrow = c(1,1))
bymedian <- with(PFT, reorder(Sp_Code, -Thick,  mean, na.rm = T))
boxplot(Thick ~ bymedian, data = PFT, las = 2, horizontal = F, main = captions['Thick'])
bymedian <- with(PFT, reorder(Sp_Code, -LA,  mean, na.rm = T))
boxplot(log(LA) ~ bymedian, data = PFT, las = 2, horizontal = F, main = captions['log(LA)'])
bymedian <- with(PFT, reorder(Sp_Code, -LDMC,  mean, na.rm = T))
boxplot(LDMC ~ bymedian, data = PFT, las = 2, horizontal = F, main = captions['LDMC'])
bymedian <- with(PFT, reorder(Sp_Code, -SLA,  mean, na.rm = T))
boxplot(log(SLA) ~ bymedian, data = PFT, las = 2, horizontal = F, main = captions['log(SLA)'])
bymedian <- with(PFT, reorder(Sp_Code, -log(SHA),  mean, na.rm = T))
boxplot(log(SHA) ~ bymedian, data = PFT, las = 2, horizontal = F, main = captions['log(SHA)'])
bymedian <- with(PFT, reorder(Sp_Code, -WD,  mean, na.rm = T))
boxplot(WD ~ bymedian, data = PFT, las = 2, horizontal = F, main = captions['WD'])
bymedian <- with(PFT, reorder(Sp_Code, -LTD,  median, na.rm = T))
boxplot(LTD ~ bymedian, data = PFT, las = 2, horizontal = F, main = captions['LTD'])
rm(bymedian)
```

### Leaf Principal component analysis

```{r Leaf PCA}
# pca <- princomp(~ LA + WD + Thick + LDMC + LTD + SLA + SHA, data = PFT, cor = T)
pca <- princomp(~ LA + WD + Thick + LDMC + SLA + SHA, data = PFT, cor = T)
par(mfrow = c(1,1))
col = rainbow(93)[as.numeric(PFT[row.names(pca$scores),3])]
plot(pca$scores[,1:2], pch = 16, cex = 0.6, col = col)
rm(col, pca)
```

## Traits informations

#### Distributions

```{r distributions}
par(mfrow = c(1,1))
hist(PFT$Thick, prob = T)
lines(density(PFT$Thick, na.rm = T), col = "red")
hist(log(PFT$LA), prob = T)
lines(density(log(PFT$LA), na.rm = T), col = "red")
hist(PFT$LDMC, prob = T)
lines(density(PFT$LDMC, na.rm = T), col = "red")
# hist(PFT$LTD, prob = T)
# lines(density(PFT$LTD, na.rm = T), col = "red")
hist(log(PFT$SLA), prob = T)
lines(density(log(PFT$SLA), na.rm = T), col = "red")
hist(log(PFT$SHA), prob = T)
lines(density(log(PFT$SHA), na.rm = T), col = "red")
hist(PFT$WD, prob = T)
lines(density(PFT$WD, na.rm = T), col = "red")
```

### Individual aggregation and coefficient of variation

```{r Individual aggragetion, echo = T}
PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
row.names(PFT_ind) <- PFT_ind$Group.1
SpT <- unique(PFT[1:2])
Sp <- SpT$SP
names(Sp) <- SpT$Tree
SpCodeT <- unique(PFT[2:3])
SpCode <- SpCodeT$Sp_Code
names(SpCode) <- SpCodeT$SP
CET <- unique(PFT[c(1,4)])
CE <- CET$CE
names(CE) <- CET$Tree
PFT_ind$SP <- Sp[row.names(PFT_ind)]
PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
PFT_ind$CE <- CE[row.names(PFT_ind)]
PFT_ind <- PFT_ind[-c(1:2)]
rm(SpT, Sp, SpCodeT, CET, CE)
```

### Coefficient of variation
```{r CV}
cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
cv$Individual <- unlist(lapply(
  as.list(row.names(cv)), 
  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Species <- unlist(lapply(
  as.list(row.names(cv)), 
  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Community <- unlist(
  lapply(as.list(row.names(cv)), 
         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))                 
cv <- round(cv, 2)
cv['WD','Individual'] <- ''
knitr::kable(cv)
rm(cv, PFT)
```

### Tree Principal component analysis

```{r Tree PCA}
library(vegan)
# pca <- princomp(~ LA + WD + Thick + LDMC + LTD + SLA, data = PFT_ind, cor = T)
pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_ind, cor = T)
par(mfrow = c(1,1))
pca.var = apply(pca$score, 2, var)
barplot((pca.var / sum(pca.var)), las = 2, 
        main = 'StdInertia', ylim = c(0,range(bstick(7))[2]))
lines(bstick(6), col = 'red')
legend('bottomleft', c('PCA', 'broken stick'), fill = c('grey', 'red'))
biplot(pca, col = c('grey', 'red'), choices = c(1,2))
PCAscores <- as.data.frame(pca$scores)[1:3] # Adding PCA axis in PFT values
names(PCAscores) <- c('PCA1', 'PCA2', 'PCA3')
PCAscores$ID <- row.names(PCAscores)
PFT_ind$ID <- row.names(PFT_ind)
PFT_ind <- merge(PFT_ind, PCAscores, all = T)
row.names(PFT_ind) <- PFT_ind$ID
PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
rm(pca.var, PCAscores)
```

### Traits relations

```{r Traits relations}
library(smatr)
par(mfrow = c(2,2))
PFTrel("LDMC", "Thick", PFT_ind, captions, cex = 0.7)
PFTrel("LDMC", "WD", PFT_ind, captions)
PFTrel("LDMC", "log(SLA)",  PFT_ind, captions)
PFTrel("Thick", "log(SLA)",  PFT_ind, captions)
PFTrel('log(LA)', "LDMC", PFT_ind, captions)
PFTrel('log(LA)', "WD", PFT_ind, captions)
PFTrel('log(LA)', "Thick", PFT_ind, captions)
X = PFT_ind$SLA*PFT_ind$LDMC
Y = PFT_ind$Thick
plot(Y ~ X, xlab = paste(captions['SLA'], 'x', captions['LDMC']), ylab = captions['Thick'], main = 'Thick ~ (SLA*LDMC)')
abline(sma(Y ~ X), col = 'red')
mtext(paste('Standardized Major Axis : R\u00b2 =',round(sma(Y ~ X)$r2[[1]], 2)), 1, col = 'red')
c <- cor.test(X, Y, method = 'spearman')
mtext(paste('Spearman\'s correlation : cor =', round(c$estimate, 2), stars(c$p.value)) , 3, col = 'blue')
rm(c, X, Y)
```

### Crown exposure effect

```{r CE, eval=FALSE, include=FALSE}
library(lme4)
CEtest <- data.frame(pca1 = pca$scores[,1],
                     CEs = PFT_ind[row.names(pca$scores),which(names(PFT) == 'CEs')],
                     Sp = PFT_ind[row.names(pca$scores),3])
CEtest <- CEtest[-which(is.na(CEtest$CEs)),]
m0 <- lmer(pca1 ~ CEs + (1 | CEs), CEtest)
m1 <- lmer(pca1 ~ CEs + (Sp | CEs), CEtest)
anova(m0, m1)
rm(CEtest)
```

# Environment

## Load data

```{r Load env}
library(raster)
Env <- genEnv()
Trees <- genTrees()
Species <- genSpecies()
hillshade <- DEMderiv(Env$DEM, c("slope", "aspect", "hillshade"))$Hillshade # For plotting purposes
```

## Trees
```{r Trees plot}
par(mfrow = c(1,1))
plot(hillshade, col = grey(0:100/100), main = 'Trees ~ Girth + Sp')
points(Y ~ X, Trees, col = SpCode, cex = `2013_girth`/100, pch = 20)
legend("bottomleft", legend = 'Species', fill = 'lightgrey')
```

## Traits

```{r Species aggregation}
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
Trees$SpCode <- tolower(Trees$SpCode)
rm(PFT_ind)
setdiff(PFT_sp$Sp_Code, Trees$SpCode)
setdiff(Trees$SpCode, PFT_sp$Sp_Code)
```

Difference between species

* Agrosystachis indica (agin) is not yet registered in the database
* Toona ciliata (toci) is not yet removed from the database
* We didn't collect 25 rare species from Uppangala plots

```{r Merging Trees and traits, echo = T}
Trees$ID <- row.names(Trees)
l <- length(row.names(Trees))
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
l - length(row.names(Trees))
rm(l)
```

There is 91 trees in missing species of our PFT dataset.

```{r Plot WD}
plot(hillshade, col=grey(0:100/100), main = 'Trees ~ Girth + WD')
col = heat.colors(100)[as.numeric(cut(Trees$WD, breaks = 100))]
points(Y ~ X, Trees, col = "black", cex = `2013_girth`/100, pch = 20)
points(Y ~ X, Trees, cex = `2013_girth`/100, pch = 20, col = col)
legend("bottomleft", legend = captions['WD'], fill = 'orange')
rm(col)
```

# Null Models

## Communities

Community definition by quadrat

```{r Communities, echo = T}
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
```
```{r Plot com}
plot(hillshade, col=grey(0:100/100), main = 'Communities')
plot(as(com, "SpatialPolygons"), add = T)
col =  sample(rainbow(length(levels(as.factor(Trees$com)))))[as.numeric(as.factor(Trees$com))]
points(Y ~ X, Trees, cex = `2013_girth`/100, pch = 20, col = col)
legend("bottomleft", legend = 'community', fill = 'grey')
rm(col, XY, Trees_bin)
```

Some species distribtuion for SLA and WD by community

```{r dist com}
library(cati)
par(mfrow = c(4,4), cex = 0.5)
rows = which(Trees$com %in% sample(levels(as.factor(Trees$com)), 8))
plotDistri(as.matrix(Trees$SLA[rows]), Trees$com[rows], Trees$SpCode[rows],
           ylim.cex = 3, plot.ask = F, multipanel = F, leg = F, xlim = c(0, 35))
plotDistri(as.matrix(Trees$WD[rows]), Trees$com[rows], Trees$SpCode[rows],
           ylim.cex = 3, plot.ask = F, multipanel = F, leg = F, xlim = c(0.3, 0.85))
rm(rows)
```

Adding topo to communities

```{r com topo}
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$SouthWesterness <- sin((com@data$Aspect + 225) / 180 * pi)
com@data$NorthEasterness <- cos((com@data$Aspect + 225) / 180 * pi)
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
rm(Deriv)
```

PCA on topgraphical variables
```{r topo pca}
pca <- princomp(~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SouthWesterness, data = com@data, cor = T)
lambda <- pca$sdev * sqrt(nrow(pca$scores))
pca.var = apply(pca$score, 2, var)
plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
     xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
Hmisc::subplot(function(){
  b <- barplot((pca.var / sum(pca.var)), las = 2,
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(pca.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.05,0.35), from='npc'),
  y=grconvertY(c(0.025,0.30), from='npc')
)
par(new=T)
Rot <- t(t(pca$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, lwd = 2)
lines(c(0,0), XLIM, lty = 2, lwd = 2)
col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2,col=col)
legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
       text.col = col, cex = 1.2)
text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 1), rownames(Rot),cex=1.2,col=col)
axis (3)
axis (4)
PCAscores <- as.data.frame(pca$scores)[1:3] # Adding PCA axis in PFT values
names(PCAscores) <- c('PCA1', 'PCA2', 'PCA3')
PCAscores$ID <- row.names(PCAscores)
```

Topographical classes assignation and plot

```{r Plot topo}
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
par(mfrow = c(1,1))
col <- c("#008B00", "#00CD00", "#009ACD", "#00BFFF", "#CD9B1D", "#FFC125")
# col <- terrain.colors(7)[c(1,2,5,6,3,4)]
plot(raster(com, 'topo'), col = col, main = 'Topographical classes', legend = F)
contour(Env$DEM, add = T)
legend('bottomleft', fill = col, legend = levels(com@data$topo), cex = 0.7)
rm(col)
```

Adding CWM to communities
```{r CWM}
CWM <- aggregate(Trees, by = list(Trees$com), mean, na.rm = T) #CWM abundance
names(CWM)[1] <- 'id'
CWM <- CWM[which(names(CWM) %in% c('id', 'Thick', 'LA', 'LDMC', 'SLA', 'WD', 'PCA1', 'PCA2', 'PCA3'))]
com_abund <- merge(com, CWM)
Trees_bin <- Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),] #CWM binary
CWM <- aggregate(Trees_bin, by = list(Trees_bin$com), mean, na.rm = T)
names(CWM)[1] <- 'id'
CWM <- CWM[which(names(CWM) %in% c('id', 'Thick', 'LA', 'LDMC', 'SLA', 'WD', 'PCA1', 'PCA2', 'PCA3'))]
com_bin <- merge(com, CWM)
library(plyr)
TreesBA <- Trees[-which(is.na(Trees$`2013_girth`)),] #CWM ba
CWM <- ddply(TreesBA, .(com), function(x){
  weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})
CWM <- data.frame(id = CWM$com,
                  Thick = CWM$V1,
                  LA = ddply(TreesBA, .(com), function(x){weighted.mean(x$LA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  LDMC = ddply(TreesBA, .(com), function(x){weighted.mean(x$LDMC, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  SLA = ddply(TreesBA, .(com), function(x){weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  WD = ddply(TreesBA, .(com), function(x){weighted.mean(x$WD, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  PCA1 = ddply(TreesBA, .(com), function(x){weighted.mean(x$PCA1, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  PCA2 = ddply(TreesBA, .(com), function(x){weighted.mean(x$PCA2, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  PCA3 = ddply(TreesBA, .(com), function(x){weighted.mean(x$PCA3, x$`2013_girth`/2*pi, na.rm = T)})[,2])
com_ba <- merge(com, CWM)
plot(gstat::variogram(WD~1, as(com_abund, 'SpatialPointsDataFrame'))) # 1 variogram
```

## Filters

Linear regression & Null models on all variables

```{r lm}
reg <- '~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SouthWesterness'
LM <- lapply(list(com_abund, com_bin, com_ba), 
             function(x) {with(x@data, list(PCA1 = lm(formula(paste('PCA1', reg))),
                                            PCA2 = lm(formula(paste('PCA2', reg))),
                                            PCA3 = lm(formula(paste('PCA3', reg))),
                                            logSLA = lm(formula(paste('log(SLA)', reg))),
                                            LDMC = lm(formula(paste('LDMC', reg))),
                                            # LTD = lm(formula(paste('LTD', reg))),
                                            logLA = lm(formula(paste('log(LA)', reg))),
                                            Thick = lm(formula(paste('Thick', reg))),
                                            WD = lm(formula(paste('WD', reg)))))})
names(LM) <- c('Abundance', 'Presence / Absence', 'Basal area')
cat('Shapiro test p-values :\n') ; unlist(lapply(LM, function(y){lapply(y, function(x){stars(shapiro.test(x$residuals)$p.value)})})) # Homoscedasticity
LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
# lapply(LM, function(x){lapply(x, summary)}) # No need with summary table
```

Null models

```{r Model, eval=FALSE, include=FALSE}
TreesL <- list(Trees, Trees_bin, TreesBA)
comL <- list(com_abund, com_bin, com_ba)
baL <- list(F, F, T)
model <- mapply(function(lm, trees, com, ba){
  null_model(formulas = lapply(lm, function(x){as.character(x$call)[2]}), 
             ba = ba, test = 'lm deviance', PFT_sp = PFT_sp, Trees = trees, 
             com = com, SpCode = SpCode, n = 1000, plot = F, count = F, time = T)},
  LM, TreesL, comL, baL, SIMPLIFY = F)
# lapply(model, function(x){
#   par(mfrow = c(2,4))
#   for(i in 1:8){
#     plot_model(x, index = i)
#   }
# })
res <- regression_table(LM, model, reg)
View(res)
knitr::kable(res)
```

Anova with topographical classes

```{r anova, eval=FALSE, include=FALSE}
ANOVA <- with(com@data, list(PCA1 = aov(PCA1 ~ topo),
                             PCA2 = aov(PCA2 ~ topo),
                             PCA3 = aov(PCA3 ~ topo),
                             SLA = aov(SLA ~ topo),
                             LDMC = aov(LDMC ~ topo),
                             # LTD = aov(LTD ~ topo),
                             LA = aov(LA ~ topo),
                             Thick = aov(Thick ~ topo),
                             WD = aov(WD ~ topo)))
cat('ANOVA p-values :\n') ; unlist(lapply(ANOVA, function(x){stars(summary(x)[[1]][['Pr(>F)']][1])}))
ANOVA <- ANOVA[which(lapply(ANOVA, function(x){summary(x)[[1]][['Pr(>F)']][1]}) < 0.05)]
lapply(ANOVA, TukeyHSD)
rm(ANOVA)
```

### 2016-06-15 inst/doc/Genus.Rmd
v0.1.9011 Doc reorganization

---
title: "Genus"
author: "Sylvain Schmitt"
date: "May 26, 2016"
output: html_document
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 10, fig.width = 10,
  cache = T, cache.lazy = F)
```

# Data
```{r Data, echo=TRUE}
rm(list = ls()); invisible(gc())
library(Uppangala)
library(raster)
Env <- genEnv()
Trees <- genTrees()
Species <- genSpecies()
Trees$SpCode <- as.factor(tolower(as.character(Trees$SpCode))) # Should be treated in gen functions
row.names(Species) <- tolower(row.names(Species))
# Inconsistancies
sort(Species$SpCode)[which(!(sort(Species$SpCode) %in% sort(levels(Trees$SpCode))))]
sort(levels(Trees$SpCode))[which(!(sort(levels(Trees$SpCode)) %in% (sort(Species$SpCode))))]
```

# Species plots
```{r Plot sp, echo=FALSE}
SpCode <- sort(Species$SpCode)[which(sort(Species$SpCode) %in% sort(levels(Trees$SpCode)))] # Using only common SpCode
Genus <- unlist(lapply(strsplit(Species[SpCode,2], ' ', fixed = T), function(x){x[[1]]}))
for(Gen in unique(Genus)){
  if(length(which(Genus == Gen)) > 1){
    main <- paste(Gen, Species[which(Genus == Gen)[1],3], sep = ', ')
    plot(Env$DEM, main = main, col = grey(100:-1:0/100), alpha = 0.8)
    contour(Env$DEM, add = T)
    col <- rainbow(length(which(Genus == Gen)))
    j <-  1
    Sp.l <- c()
    for(i in which(Genus == Gen)){
      Sp <-  SpCode[i]
      Sp.l <- c(Sp.l, Sp)
      points(Y ~ X, Trees[Trees$SpCode == Sp,], pch = 20, col = col[j])
      j <-  j + 1
    }
    species.v <- unlist(lapply(strsplit(Species[Sp.l,2], ' ', fixed = T), function(x){x[[2]]}))
    legend('bottomright', species.v, fill = col)
  }
}
```

### 2016-06-15 inst/doc/Habitat.Rmd
v0.1.9011 Doc reorganization

---
title: "TopoSpecies"
author: "Sylvain Schmitt"
date: "May 26, 2016"
output: html_document
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = T, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

# Data
```{r Data}
rm(list = ls()); invisible(gc())
library(Uppangala)
library(raster)
Env <- genEnv()
Trees <- genTrees()
Species <- genSpecies()
Trees$SpCode <- as.factor(tolower(as.character(Trees$SpCode))) # Should be treated in gen functions
row.names(Species) <- tolower(row.names(Species))
Trees <- Trees[-which(is.na(Trees$X)),]
```

# Community definition
```{r Community}
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
```

# Adding topo to communities

```{r com topo}
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 2), c('slope', 'curvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
```

# Classes assignation
```{r Plot pop}
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
Trees$topo <- (XY %over% com)[,6]
```

# Species habitat association table
```{r Sp habitat table, echo = F}
habitat_association <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$SpCode, Trees$topo), sum))
habitat_association <- round(habitat_association / rowSums(habitat_association, na.rm = T) * 100)
habitat_association[is.na(habitat_association)] <- 0
knitr::kable(habitat_association)
```

# Species - Habitats plots
```{r Plot sp habitat, echo = F}
SpCode <- levels(Trees$SpCode)
col <- c("#008B00", "#00CD00", "#009ACD", "#00BFFF", "#CD9B1D", "#FFC125")
# col <- terrain.colors(7)[c(1,2,5,6,3,4)]
for(Sp in SpCode){
  main <- Sp
  if(Sp %in% row.names(Species)){
    main <- paste(Species[as.character(Sp),2], Species[as.character(Sp),3], sep = ', ')
  }
  plot(raster(com, 'topo'), col = col, legend = F, main = main)
  contour(Env$DEM, add = T)
  legend('bottomleft', fill = col, legend = levels(com@data$topo), cex = 0.7)
  points(Y ~ X, Trees[Trees$SpCode == Sp,], pch = 20)
}
```

### 2016-06-15 inst/doc/PreliminaryResults.Rmd
v0.1.9011 Doc reorganization

---
title: 'Uppangala tree functionnal traits: Preliminray Results'
author: "Sylvain Schmitt"
date: '`r Sys.Date()`'
output: pdf_document
bibliography: refs.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

# Traits

## Measured traits

Table 1: Functional traits measured. *Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development.*

```{r Traits, eval=T}
rm(list = ls()); invisible(gc())
library(Uppangala)
PFT <- genPFT()
captions <- genCaptions()
table <- data.frame(
  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
                         round(sd(PFT$Thick, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
                         round(sd(PFT$LA, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
                         round(sd(PFT$LDMC, na.rm = T)), ')'),
                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
                         round(sd(PFT$SLA, na.rm = T),2), ')'),
                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
                         round(sd(PFT$WD, na.rm = T),2), ')')),
  unit = c('um', 'mm2', 'mg.g-1', 'm2.kg-1', 'g.cm-3'),
  strategy = c('Leaf defense vs. investment',
               'Leaf ressource capture vs.investment',
               'Leaf defense vs. investment',
               'Leaf ressource acquisiton vs. defense',
               'Stem transport, structure and defense')
  )
kable(table,  label="traits", row.names = F)
```

Uppangala measured functional traits vary in the ranges of tropical species values that we can find in literature. We can notice in table 1 that wood density has a relatively high mean as we could expect in an old untouched forest. Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010], and can be as well explained by environment condition. For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011]. We could also have added a sixth trait, called leaf tissue density (LTD) with the formula: $LTD=Leaf dry weight/Thickness*Leaf area$. LTD is highly correlated to LDMC [@hodgson_is_2011], but can still bring extra information **To be discussed**.

## Traits distribution and variation

Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific. *Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species. See table 1 for abbreviation.*

```{r CV}
PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
row.names(PFT_ind) <- PFT_ind$Group.1
SpT <- unique(PFT[1:2])
Sp <- SpT$SP
names(Sp) <- SpT$Tree
SpCodeT <- unique(PFT[2:3])
SpCode <- SpCodeT$Sp_Code
names(SpCode) <- SpCodeT$SP
CET <- unique(PFT[c(1,4)])
CE <- CET$CE
names(CE) <- CET$Tree
PFT_ind$SP <- Sp[row.names(PFT_ind)]
PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
PFT_ind$CE <- CE[row.names(PFT_ind)]
PFT_ind <- PFT_ind[-c(1:2)]
cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
cv$Individual <- unlist(lapply(
  as.list(row.names(cv)),
  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Species <- unlist(lapply(
  as.list(row.names(cv)),
  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Community <- unlist(
  lapply(as.list(row.names(cv)),
         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
cv <- round(cv, 2)
cv['WD','Individual'] <- ''
kable(cv)
```

Some outliers are still to be corrected in individual leaf trait value, but thickness outliers and extreme values are due to too close secondary veins. Table 2 shows a intra-individual variation lower than intra-specific and inter-specific variation, indicating that leaves are even in morphology inside an individual but can start to show some variation between trees of the same species. More specifically, whereas traits like thickness and wood density seems to stay stable in intra-individual and intra-specific, leaf area shows a high coefficient of variation for both intra-individual and intra-specific values.

## Covariations in traits

```{r PCA PFT}
library(vegan)
pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_ind, cor = T)
lambda <- pca$sdev * sqrt(nrow(pca$scores))
pca.var = apply(pca$score, 2, var)
PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
par(new=T)
Rot <- t(t(pca$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
       text.col = c("green", "firebrick"))

arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
Hmisc::subplot(function(){
  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(pca.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.75,1), from='npc'),
  y=grconvertY(c(0.05,0.30), from='npc')
)
axis (3)
axis (4)
PFT_ind$ID <- row.names(PFT_ind)
PFT_ind <- merge(PFT_ind, PCAscores, all = T)
row.names(PFT_ind) <- PFT_ind$ID
PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
```

Figure 1: Principal component analysis (PCA) performed on individual trait means. *Left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line). See table 1 for abbreviation.*

We realized a principal component analysis on individual trait means (see figure 1), where first axis (PCA1) covary mainly with leaf thickness, dry matter content and wood density. Whereas second axis (PCA2) covary mainly with specific leaf area representative of the leaf economic spectrum [LES, @wright_worldwide_2004]. PCA1 might be associated with the wood density covariation to the wood economic spectrum [WES, @chave_towards_2009]. Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us). Our results integrate less traits (5/14), and less species (92/758), but still capture the same strategy axes.

We can see that as wood density ans specific leaf area are orthogonal following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$). Leaf area is almost opposed to wood density and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007. This relationship could be explained by an equilibrium in tree hydraulics (Coomes et al, 2008). Finally, we found the positive relation suggested by @diaz_global_2015 between leaf dry matter content and wood density (Spearman's $rho = 0.24$, $p-value < 0.001$).

# Environmental filters

## Environment and trees

We collected trait information on 92 over 117 (79%) species focusing in priority on abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates (*Warning : one of our 92 species is not referenced yet among currently known Uppangala species, Agrosystachis indica, one of 117 species is not removed yet, Toona ciliata.*). Thus we generated data for 98.8% of the tree individuals present in Uppangala plot.

```{r env-com}
library(raster)
Env <- genEnv()
Env$Rivers <- readAll(Env$Rivers)
Trees <- genTrees()
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$SouthWesterness <- sin((com@data$Aspect + 225) / 180 * pi)
com@data$NorthEasterness <- cos((com@data$Aspect + 225) / 180 * pi)
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
rm(Deriv, XY)
```

To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (aspect in direction of south-west), and distance to the nearest potential river. Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect in degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008]. Distance to the nearest potential river was caluclated under ArcGis (cite).

In order to avoid multi-colinearity in future regressions, we reduced our explanatory variables via a principal component analysis, see figure 2. Curvature, profile curvature, plan curvature, and distance to potential rivers covary together and with the principal component analysis first axis (PCA1). Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables. As a conclusion, we choose to reduce our variable to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness to represent monsoon winds [@bunyan_effect_2015] and slope.

```{r PCA Env}
pca.env <- princomp(~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SouthWesterness, data = com@data, cor = T)
lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
pca.env.var = apply(pca.env$score, 2, var)
plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
Hmisc::subplot(function(){
  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(pca.env.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.05,0.35), from='npc'),
  y=grconvertY(c(0.025,0.30), from='npc')
)
par(new=T)
Rot <- t(t(pca.env$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, lwd = 2)
lines(c(0,0), XLIM, lty = 2, lwd = 2)
col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
        Rot[,1],Rot[,2],lwd=2,col=col)
legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
       text.col = col, cex = 1.2)
text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 1), rownames(Rot),cex=1.2,col=col)
axis (3)
axis (4)
```

Figure 2: Principal component analysis (PCA) performed on topgraphical variables. *Left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).*

Among all topographical variables, slope, curvature and south-westerness were chosen to define 6 topographical classes. We rejected elevation because our plot is situated in a mountain slope and doesn't include valley. Consequently elevation mainly inform on the position in the mountain general slope. And because the elevation range between the top of the plot and the bottom is low (325-490), we assumed it won't affect directly tree ecology.

Slope was separated in three categories: plane surface (below 20 degrees), mean slope (between 20 and 30degrees) and steep slope (above 30 degrees). Plane surface were then discriminated on curvature: concave (positive curvature) and convex (negative curvature). Whereas slopes where discriminated by their south-west or north east aspect. Finally, 6 topographical classes were defined: concave plane surface (CvP), convex plane surface (CxP), north-east mean slope (NEMS), north-east steep slope (NESS), south-west mean slope (SWMS), and south-west steep slope (SWSS), see figure 3. The predominance of north-east slope (62%) in our study site is noticeable and should be took into account.

```{r topo classes}
col <- c("#008B00", "#00CD00", "#009ACD", "#00BFFF", "#CD9B1D", "#FFC125")
plot(raster(com, 'topo'), col = col, legend = F)
contour(Env$DEM, add = T)
legend('bottomleft', fill = col, legend = levels(com@data$topo), bg = 'white')
```

Figure 3: Uppangala 10 hectares plot topographical classes.*6 classes were defined one slope, curvature, and south-westerness: concave plane surface (CvP) in dark green, convexe plane surface (CxP) in light green, north-east mean slope (NEMS) in dark blue, north-east steep slope (NESS) in light blue, south-west mean slope (SWMS) in dark orange, and south-west steep slope (SWSS) in ligh orange.*

To study abiotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density (StD) *(we may add growth in the future)*. Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$. Both basal area and stem density are calculated with 2013 Uppangala plots girth inventory only on stem with a diameter above 10 centimeters at breast height ($DBH>10$). Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$). So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).

For each species the potential strata was defined. Because species were defined in several strata we kept the lowest strata as the species strata. We tested the strata effect on each trait with an anova combined to a tukey test, but no trait show a significant variation (all F values weren't significanlty different from 0, $p-values>0.1$).

```{r strata effect, eval=FALSE, include=FALSE, cache=F}
AOV <- list(SLA = aov(PFT_sp$SLA ~ Species[PFT_sp$Sp_Code,]$Strata), 
            WD = aov(PFT_sp$WD ~ Species[PFT_sp$Sp_Code,]$Strata),
            LDMC = aov(PFT_sp$LDMC ~ Species[PFT_sp$Sp_Code,]$Strata),
            LA = aov(PFT_sp$LA ~ Species[PFT_sp$Sp_Code,]$Strata),
            Thick = aov(PFT_sp$Thick ~ Species[PFT_sp$Sp_Code,]$Strata))
lapply(AOV, summary)
# lapply(AOV, TukeyHSD)
```

## Communities

```{r Com}
CWM <- aggregate(Trees, by = list(Trees$com), mean, na.rm = T) #CWM abundance
names(CWM)[1] <- 'id'
CWM <- CWM[which(names(CWM) %in% c('id', 'LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'))]
com_abund <- merge(com, CWM)
Trees_bin <- Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),] #CWM binary
CWM <- aggregate(Trees_bin, by = list(Trees_bin$com), mean, na.rm = T)
names(CWM)[1] <- 'id'
CWM <- CWM[which(names(CWM) %in% c('id', 'LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'))]
com_bin <- merge(com, CWM)
library(plyr)
TreesBA <- Trees[-which(is.na(Trees$`2013_girth`)),] #CWM ba
CWM <- ddply(TreesBA, .(com), function(x){
  weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})
CWM <- data.frame(id = CWM$com,
                  Thick = CWM$V1,
                  LA = ddply(TreesBA, .(com), function(x){weighted.mean(x$LA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  LDMC = ddply(TreesBA, .(com), function(x){weighted.mean(x$LDMC, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  SLA = ddply(TreesBA, .(com), function(x){weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  WD = ddply(TreesBA, .(com), function(x){weighted.mean(x$WD, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  WES = ddply(TreesBA, .(com), function(x){weighted.mean(x$WES, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  LES = ddply(TreesBA, .(com), function(x){weighted.mean(x$LES, x$`2013_girth`/2*pi, na.rm = T)})[,2])
com_ba <- merge(com, CWM)
```

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. For all those community, community weighted means [CWM, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES). Weighted mean were realized by abundance, presence-absence, or basal area.

Spatial variograms of community weighted means didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$.

```{r Variograms}
library(gstat)
dist <- variogram(WD~1, as(com_abund, 'SpatialPointsDataFrame'))$dist
plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', xlab = 'distance (m)', ylab = 'relative variance (%)')
com.L <- list(Abundance = com_abund,
              'Presence-absence' = com_bin,
              'Basal area' = com_ba)
for(j in 1:length(com.L)){
  var <- data.frame(lapply(lapply(as.list(c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')), function(x){variogram(formula(paste(x, '~ 1')), as(com.L[[j]], 'SpatialPointsDataFrame'))}), function(x){x[3]}))
  var <- data.frame(apply(var, 2, function(x){x/max(x)}))
  names(var) <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
  colramp <- rainbow(length(var))
  for(i in 1:length(var)){
    lines(var[,i] ~ dist, col = colramp[i], lty = j + 1)
  }
}
legend('bottomright', names(var), fill = colramp)
legend('bottomleft', names(com.L), lty = c(2:4))
```

Figure 4: Trait variograms for all communities.*Variogram was initiated on* $20*20 m^2$ *communities with the wood density weighted mean, and seems to reach a plateau around* $70 m$.

Links between community weighted means and environment, for both abiotic and biotic factors, were studied by linear regression on continuous variables and anova with factors. Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables (slope, curvature, wetness and southwesterness for abiotic factors, and canopy height and basal area for biotic one). A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix). Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance. Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 3 and 4 (respectively for abiotic and biotic factors).

## Abiotic filters

```{r topo results}
reg <- '~ Slope + Curvature + Wetness + SouthWesterness'
LM <- lapply(list(com_abund, com_bin, com_ba),
             function(x) {with(x@data, list(LES = lm(formula(paste('LES', reg))),
                                            logSLA = lm(formula(paste('log(SLA)', reg))),
                                            WES = lm(formula(paste('WES', reg))),
                                            WD = lm(formula(paste('WD', reg))),
                                            LDMC = lm(formula(paste('LDMC', reg))),
                                            logLA = lm(formula(paste('log(LA)', reg))),
                                            Thick = lm(formula(paste('Thick', reg)))))})
names(LM) <- c('Abundance', 'Presence/Absence', 'Basal area')
LM <- lapply(LM, function(y){lapply(y, function(x){
  step(x, direction = 'both', trace = 0)})})
TreesL <- list(Trees, Trees_bin, TreesBA)
comL <- list(com_abund, com_bin, com_ba)
baL <- list(F, F, T)
model <- mapply(function(lm, trees, com, ba){
  null_model(formulas = lapply(lm, function(x){as.character(x$call)[2]}),
             ba = ba, test = 'lm deviance', PFT_sp = PFT_sp, Trees = trees,
             com = com, SpCode = SpCode, n = 1000, plot = F, count = F, time = F)},
  LM, TreesL, comL, baL, SIMPLIFY = F)

kable(regression_table(LM, model, reg))
```

Table 3: Step mesure on abiotic factors and null models summary for traits community weighted means. *Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW). Model deviance was compared to 999 null models deviance in which traits were randomized among species.*

First we can notice that, whereas the leaf economic spectrum shows significantly non-random links with abiotic factors, most of weighted means associated to the wood economic spectrum on our principal component analysis (see figure 1) didn't differ from null models. A spatial structure due to local individual history and dispersion is responsible for the observed links between abiotic factors and weighted mean associated to the wood economic spectrum. Outside from the leaf economic spectrum (LES), topography doesn't localy afect Uppangala plot trees as an environmental filter.

Leaf economic spectrum (LES) and specific leaf area (SLA) is associated non-randomly with convex and dry surfaces, as the thickness for basal area community weighted mean. Trees will tend to increase their surface compare to their carbon investment in areas where the water will flow but not accumulate nor stay, and where the soil will tend to accumulate and be stabilized. In favourable water and soil conditions, trees will tend to invest more in leaf ressource capture at the expense of leaf defense, adopting the acquisitive [@sterck_functional_2011] our fast growing strategy [@reich_world-wide_2014].

Wood density (WD) in presence absence shows a non-random association with southwest steep slope and concave surfaces with a relatively important explanation level ($R^2=0.23$). Trees will increase their wood density in thin soil and water and soil departure conditions under potential monsoon winds *(We need to decide or not if regarding the plot condition we can accept or not the southwest orientation as being linked to monsoon winds)*. In water stress and mechanical hazard conditions, trees will compensate by increasing their water transport safety and mechanical resistance and resilience with an increased wood density [@chave_towards_2009].

## Biotic filters

```{r biotic results}
reg <- '~ Canopy + BA'
LM <- lapply(list(com_abund, com_bin, com_ba),
             function(x) {with(x@data, list(LES = lm(formula(paste('LES', reg))),
                                            logSLA = lm(formula(paste('log(SLA)', reg))),
                                            WES = lm(formula(paste('WES', reg))),
                                            WD = lm(formula(paste('WD', reg))),
                                            LDMC = lm(formula(paste('LDMC', reg))),
                                            logLA = lm(formula(paste('log(LA)', reg))),
                                            Thick = lm(formula(paste('Thick', reg)))))})
names(LM) <- c('Abundance', 'Presence/Absence', 'Basal area')
LM <- lapply(LM, function(y){lapply(y, function(x){
  step(x, direction = 'both', trace = 0)})})
TreesL <- list(Trees, Trees_bin, TreesBA)
comL <- list(com_abund, com_bin, com_ba)
baL <- list(F, F, T)
model <- mapply(function(lm, trees, com, ba){
  null_model(formulas = lapply(lm, function(x){as.character(x$call)[2]}),
             ba = ba, test = 'lm deviance', PFT_sp = PFT_sp, Trees = trees,
             com = com, SpCode = SpCode, n = 1000, plot = F, count = F, time = F)},
  LM, TreesL, comL, baL, SIMPLIFY = F)

kable(regression_table(LM, model, reg))
```

Table 4: Step mesure on abiotic factors and null models summary for traits community weighted means. *Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with canopy height (canopy), basal area (BA), and stem density (StD). Model deviance was compared to 999 null models deviance in which traits were randomized among species.*

Most of community weighted means don't show association with biotic factors, and most of the rare association are not significantly different from null models randomly generated. Still leaf area (LA) in basal area is associated with high trees. Voluminous individuals tend to increase their leaf surface in heigher canopy condition to compensate understory shadow.

# *Future directions*

**As a conclusion, some question need to be discussed:**

- **Inclusion or not of the Leaf Tissue Density (LTD) trait**
- **Interpretation of our PCA axes**
- **The study of covariogram between community weighted mean and topographical variables**
- **And more generally, the way of building the analysis (is the present trajectory okay)**

# References

### 2016-06-15 inst/doc/Presentation.Rmd
v0.1.9011 Doc reorganization

---
title: Study of spatial structure of taxonomic and functional diversity in a tropical
  forest tree community
author: "Sylvain Schmitt"
date: "June 9, 2016"
output:
  ioslides_presentation: default
  beamer_presentation:
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    fonttheme: professionalfonts
  slidy_presentation: default
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = F, message = F, warning = F, cache = T, cache.lazy = F)
```

## Sampling

- 92 species
- 272 trees
- 1362 dried leaves
- 272 fresh leaves
- 272 fresh branch wood sample

## Direct measurements

- Tree crown exposure
- Leaf fresh weight
- Leaf thickness
- Leaf area
- Leaf dry weight
- Wood volume
- Wood dry weight

## Traits {.smaller}

```{r Traits, echo=FALSE, message=FALSE, warning=FALSE}
rm(list = ls()); invisible(gc())
library(Uppangala)
PFT <- genPFT()
captions <- genCaptions()
table <- data.frame(
  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
                         round(sd(PFT$Thick, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
                         round(sd(PFT$LA, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
                         round(sd(PFT$LDMC, na.rm = T)), ')'),
                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
                         round(sd(PFT$SLA, na.rm = T),2), ')'),
                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
                         round(sd(PFT$WD, na.rm = T),2), ')')),
  unit = c('um', 'mm2', 'mg.g-1', 'm2.kg-1', 'g.cm-3'),
  strategy = c('Leaf defense vs. investment',
               'Leaf ressource capture vs.investment',
               'Leaf defense vs. investment',
               'Leaf ressource acquisiton vs. defense',
               'Stem transport, structure and defense')
  )
names(table)[3] <- 'mean (sd)'
library(knitr)
kable(table)
```

Table 1: Functionnal traits measured

## Traits variation {.smaller}

```{r Coefficients of variation, echo=FALSE, message=FALSE, warning=FALSE}
PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
row.names(PFT_ind) <- PFT_ind$Group.1
SpT <- unique(PFT[1:2])
Sp <- SpT$SP
names(Sp) <- SpT$Tree
SpCodeT <- unique(PFT[2:3])
SpCode <- SpCodeT$Sp_Code
names(SpCode) <- SpCodeT$SP
CET <- unique(PFT[c(1,4)])
CE <- CET$CE
names(CE) <- CET$Tree
PFT_ind$SP <- Sp[row.names(PFT_ind)]
PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
PFT_ind$CE <- CE[row.names(PFT_ind)]
PFT_ind <- PFT_ind[-c(1:2)]
cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
cv$Individual <- unlist(lapply(
  as.list(row.names(cv)), 
  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Species <- unlist(lapply(
  as.list(row.names(cv)), 
  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Community <- unlist(
  lapply(as.list(row.names(cv)), 
         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))                 
cv <- round(cv, 2)
cv['WD','Individual'] <- ''
knitr::kable(cv)
```

Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific

## Pricnipal component analysis {.smaller}

```{r PCA, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Principal component analysis (PCA) performed on individual trait means.', fig.height=5}
library(vegan)
pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_ind, cor = T)
lambda <- pca$sdev * sqrt(nrow(pca$scores))
plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
     xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
par(new=T)
Rot <- t(t(pca$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
par(mfrow = c(1,1))
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
       text.col = c("green", "firebrick"))

arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
pca.var = apply(pca$score, 2, var)
Hmisc::subplot(function(){
  barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(bstick(6), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.75,1), from='npc'),
  y=grconvertY(c(0.05,0.30), from='npc')
)
axis (3) ; axis (4)
PCAscores <- as.data.frame(pca$scores)[1:3] # Adding PCA axis in PFT values
names(PCAscores) <- c('PCA1', 'PCA2', 'PCA3')
PCAscores$ID <- row.names(PCAscores)
PFT_ind$ID <- row.names(PFT_ind)
PFT_ind <- merge(PFT_ind, PCAscores, all = T)
row.names(PFT_ind) <- PFT_ind$ID
PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
```

##  Traits relations {.smaller}

```{r Traits relations, echo=FALSE, message=FALSE, warning=FALSE}
library(smatr)
par(mfrow = c(2,2))
PFTrel("WD", "log(SLA)", PFT_ind, captions)
PFTrel("log(LA)", "WD", PFT_ind, captions)
PFTrel("LDMC", "WD",  PFT_ind, captions)
X = PFT_ind$SLA*PFT_ind$LDMC
Y = PFT_ind$Thick
plot(Y ~ X, xlab = paste(captions['SLA'], 'x', captions['LDMC']), ylab = captions['Thick'], main = 'Thick ~ (SLA*LDMC)')
abline(sma(Y ~ X), col = 'red')
mtext(paste('Standardized Major Axis : R\u00b2 =',round(sma(Y ~ X)$r2[[1]], 2)), 1, col = 'red')
c <- cor.test(X, Y, method = 'spearman')
mtext(paste('Spearman\'s correlation : cor =', round(c$estimate, 2), stars(c$p.value)) , 3, col = 'blue')

```

## Uppangala plot trees {.smaller}

```{r Traits and Trees, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Mean species WD with all Uppangala individuals'}
library(raster)
Env <- genEnv()
Trees <- genTrees()
Species <- genSpecies()
hillshade <- DEMderiv(Env$DEM, c("slope", "aspect", "hillshade"))$Hillshade
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
par(mfrow = c(1,1))
plot(hillshade, col=grey(0:100/100))
col = heat.colors(100)[as.numeric(cut(Trees$WD, breaks = 100))]
points(Y ~ X, Trees, col = "black", cex = `2013_girth`/100, pch = 20)
points(Y ~ X, Trees, cex = `2013_girth`/100, pch = 20, col = col)
legend("bottomleft", legend = captions['WD'], fill = 'orange')
```

- 92 over 117 species (79%) & 98.8% tree individuals

## Communities definition {.smaller}

```{r Communities definition, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='20*20m2 communities'}
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
plot(hillshade, col=grey(0:100/100))
plot(as(com, "SpatialPolygons"), add = T)
col =  sample(rainbow(length(levels(as.factor(Trees$com)))))[as.numeric(as.factor(Trees$com))]
points(Y ~ X, Trees, cex = `2013_girth`/100, pch = 20, col = col)
legend("bottomleft", legend = 'community', fill = 'grey')
```

## Communities distribution {.smaller}

```{r Communities distribution, echo=FALSE, fig.cap='2 examples of community trait distribution for SLA and WD', message=FALSE, warning=FALSE}
library(cati)
par(mfrow = c(2,2), cex = 0.5)
rows = which(Trees$com %in% sample(levels(as.factor(Trees$com)), 2))
plotDistri(as.matrix(Trees$SLA[rows]), Trees$com[rows], Trees$SpCode[rows],
           ylim.cex = 3, plot.ask = F, multipanel = F, leg = F, xlim = c(0, 35))
plotDistri(as.matrix(Trees$WD[rows]), Trees$com[rows], Trees$SpCode[rows],
           ylim.cex = 3, plot.ask = F, multipanel = F, leg = F, xlim = c(0.3, 0.85))
```

## Topography - Conitnuous variable {.smaller}

```{r Topography continous, echo=FALSE, fig.cap='Community topography', message=FALSE, warning=FALSE}
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$SouthWesterness <- sin((com@data$Aspect + 225) / 180 * pi)
com@data$NorthEasterness <- cos((com@data$Aspect + 225) / 180 * pi)
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
par(mfrow = c(1,1), cex = 1)
plot(stack(com)[[-1]])
```

## Topography - Classes {.smaller}

```{r Topgraphy clases, echo=FALSE, fig.cap='Topographical classes', message=FALSE, warning=FALSE}
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
col <- c("#008B00", "#00CD00", "#009ACD", "#00BFFF", "#CD9B1D", "#FFC125")
plot(raster(com, 'topo'), col = col, legend = F)
contour(Env$DEM, add = T)
legend('bottomleft', fill = col, legend = levels(com@data$topo), cex = 0.7)
```

## CWM - Abundances {.smaller}

```{r CWM ab, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Community weighted mean by abundance'}
CWM <- aggregate(Trees, by = list(Trees$com), mean, na.rm = T) #CWM abundance
names(CWM)[1] <- 'id'
CWM <- CWM[which(names(CWM) %in% c('id', 'Thick', 'LA', 'LDMC', 'SLA', 'WD', 'PCA1', 'PCA2'))]
com_abund <- merge(com, CWM)
plot(stack(com_abund[13:19]))
```

## CWM - Presence-Absence {.smaller}

```{r CWM bin, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Community weighted mean by presence-absence'}
Trees_bin <- Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),] #CWM binary
CWM <- aggregate(Trees_bin, by = list(Trees_bin$com), mean, na.rm = T)
names(CWM)[1] <- 'id'
CWM <- CWM[which(names(CWM) %in% c('id', 'Thick', 'LA', 'LDMC', 'SLA', 'WD', 'PCA1', 'PCA2', 'PCA3'))]
com_bin <- merge(com, CWM)
plot(stack(com_bin[13:19]))
```

## CWM - Basal Area {.smaller}

```{r CWM ba, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Community weighted mean by basal area'}
library(plyr)
TreesBA <- Trees[-which(is.na(Trees$`2013_girth`)),] #CWM ba
CWM <- ddply(TreesBA, .(com), function(x){
  weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})
CWM <- data.frame(id = CWM$com,
                  Thick = CWM$V1,
                  LA = ddply(TreesBA, .(com), function(x){weighted.mean(x$LA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  LDMC = ddply(TreesBA, .(com), function(x){weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  SLA = ddply(TreesBA, .(com), function(x){weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  WD = ddply(TreesBA, .(com), function(x){weighted.mean(x$WD, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  PCA1 = ddply(TreesBA, .(com), function(x){weighted.mean(x$PCA1, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  PCA2 = ddply(TreesBA, .(com), function(x){weighted.mean(x$PCA2, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  PCA3 = ddply(TreesBA, .(com), function(x){weighted.mean(x$PCA3, x$`2013_girth`/2*pi, na.rm = T)})[,2])
com_ba <- merge(com, CWM)
plot(stack(com_ba[13:19]))
```

## Spatial organization {.smaller}

```{r Variogram, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Variogram example'}
plot(gstat::variogram(WD~1, as(com_abund, 'SpatialPointsDataFrame')))
```

### 2016-06-15 R/draft.R

@@ -77,3 +77,96 @@
 #                            SLA = aov(SLA ~ Strata),
 #                            WD = aov(WD ~ Strata)))
 # lapply(AOV, summary)
+
+# # Data
+# library(Uppangala)
+# library(raster)
+# library(ade4)
+# Env <- genEnv()
+# Deriv <- DEMderiv(Env$DEM, c('slope', 'curvature', 'aspect', 'wetness'))
+# Deriv$SW <- cos(Deriv$Aspect + 220)
+# Deriv <- Deriv[[-3]]
+#
+# # Individuals
+# TreesXY <- Trees[-which(is.na(Trees$X)),]
+# TreesXY <- TreesXY[-which(is.na(TreesXY$`2013_girth`)),]
+# TreesXY <- TreesXY[-which(is.na(extract(Env$DEM, TreesXY[2:3], df = T)[2])),]
+# TreesXY <- TreesXY[-which(TreesXY$SpCode %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
+#
+# # Matrices
+# Lm <- TreesXY[1]
+# Lm <- data.frame(tapply(rep(1, length(Lm[,1])), list(row.names(Lm), Lm$SpCode), sum))
+# Lm[is.na(Lm)] <- 0
+# Qm <- PFT_sp[which(names(PFT_sp) %in% c('SLA', 'LDMC', 'WD', 'Thick', 'LA'))]
+# row.names(Qm) <- PFT_sp$Sp_Code
+# Qm <- Qm[-which(is.na(Qm$WD)),]
+# Qm <- Qm[names(Lm),]
+# Rm <- cbind(extract(Deriv, TreesXY[2:3], df = T)[-1], TreesXY$`2013_girth`)
+# names(Rm)[5] <- 'Girth'
+# row.names(Rm) <- row.names(Lm)
+#
+# # RLQ
+# L <- dudi.coa(Lm, scannf = FALSE)
+# R <- dudi.pca(Rm, scannf = FALSE, row.w = L$lw)
+# Q <- dudi.pca(Qm, scannf = FALSE, row.w = L$cw)
+# rlq <- rlq(R, L, Q, scannf = FALSE)
+# rd <- randtest(rlq)
+# fq <- fourthcorner.rlq(rlq, type = "Q.axes")
+# fr <- fourthcorner.rlq(rlq, type = "R.axes")
+# fqr <- fourthcorner(Rm, Lm, Qm)
[…]

### 2016-06-15 R/genEnv.R

@@ -1,5 +1,5 @@
 #' @include genTrees.R
-#' @importFrom raster shapefile raster crop resample hillShade stack
+#' @importFrom raster shapefile raster crop resample hillShade stack extent
 NULL
 
 #' Environmental data generation

### 2016-06-15 R/topo_deriv.R

@@ -1,5 +1,5 @@
 #' @importFrom RSAGA rsaga.env rsaga.slope rsaga.curvature rsaga.plan.curvature rsaga.profile.curvature rsaga.aspect rsaga.fill.sinks rsaga.wetness.index
-#' @importFrom raster writeRaster raster hillShade
+#' @importFrom raster writeRaster raster hillShade readAll
 NULL
 
 #' Create DEM derivative

### 2016-06-20 inst/doc/PreliminaryResults.Rmd
v0.1.9012 CWT.R, null_model.R, RLQ, SAC, PhyloS, CWV, SEM

@@
 
+# *News*
+Add from the june 20th version :
+
+- RLQ to investigate individuals filters
+- Spatial autocorrelation for CWM and models residuals with Moran's I
+- Phylosignal
+- CWV
+- Structural Equation Modelling (SEM)
+
 # Traits
@@
 
-```{r Traits, eval=T}
+```{r 1 Traits}
 rm(list = ls()); invisible(gc())
@@
 
-```{r CV}
+```{r 2 CV}
 PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
@@
 
-```{r PCA PFT}
+```{r 3 PCA PFT}
 library(vegan)
@@
 
-```{r env-com}
+```{r 4 env-com}
 library(raster)
@@
 
-To study abiotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density (StD) *(we may add growth in the future)*.
+To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density (StD) *(we may add tree height and growth in the future)*.
 Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$.
@@
 For each species the potential strata was defined.
-Because species were defined in several strata we kept the lowest strata as the species strata.
+Because species were defined in several strata, we kept the lowest strata as the species strata.
 We tested the strata effect on each trait with an anova combined to a tukey test, but no trait show a significant variation (all F values weren't significanlty different from 0, $p-values>0.1$).
 
-```{r strata effect, eval=FALSE, include=FALSE, cache=F}
+```{r strata effect}
 AOV <- list(SLA = aov(PFT_sp$SLA ~ Species[PFT_sp$Sp_Code,]$Strata), 
@@
             Thick = aov(PFT_sp$Thick ~ Species[PFT_sp$Sp_Code,]$Strata))
-lapply(AOV, summary)
+# lapply(AOV, summary)
 # lapply(AOV, TukeyHSD)
@@
 
-```{r Com}
-CWM <- aggregate(Trees, by = list(Trees$com), mean, na.rm = T) #CWM abundance
-names(CWM)[1] <- 'id'
-CWM <- CWM[which(names(CWM) %in% c('id', 'LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'))]
-com_abund <- merge(com, CWM)
-Trees_bin <- Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),] #CWM binary
-CWM <- aggregate(Trees_bin, by = list(Trees_bin$com), mean, na.rm = T)
-names(CWM)[1] <- 'id'
-CWM <- CWM[which(names(CWM) %in% c('id', 'LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'))]
-com_bin <- merge(com, CWM)
-library(plyr)
-TreesBA <- Trees[-which(is.na(Trees$`2013_girth`)),] #CWM ba
-CWM <- ddply(TreesBA, .(com), function(x){
-  weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})
-CWM <- data.frame(id = CWM$com,
-                  Thick = CWM$V1,
-                  LA = ddply(TreesBA, .(com), function(x){weighted.mean(x$LA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
-                  LDMC = ddply(TreesBA, .(com), function(x){weighted.mean(x$LDMC, x$`2013_girth`/2*pi, na.rm = T)})[,2],
-                  SLA = ddply(TreesBA, .(com), function(x){weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
-                  WD = ddply(TreesBA, .(com), function(x){weighted.mean(x$WD, x$`2013_girth`/2*pi, na.rm = T)})[,2],
-                  WES = ddply(TreesBA, .(com), function(x){weighted.mean(x$WES, x$`2013_girth`/2*pi, na.rm = T)})[,2],
-                  LES = ddply(TreesBA, .(com), function(x){weighted.mean(x$LES, x$`2013_girth`/2*pi, na.rm = T)})[,2])
-com_ba <- merge(com, CWM)
+```{r 5 Com}
+com <- CWT(Trees, com)
 ```
@@
 Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
-For all those community, community weighted means [CWM, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES).
-Weighted mean were realized by abundance, presence-absence, or basal area.
+For all those community, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES).
+Weighted mean and variances were realized by abundance, presence-absence, or basal area.
 
-Spatial variograms of community weighted means didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$.
+Spatial variograms of community weighted means and variances didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$.
 
@@
 library(gstat)
-dist <- variogram(WD~1, as(com_abund, 'SpatialPointsDataFrame'))$dist
-plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', xlab = 'distance (m)', ylab = 'relative variance (%)')
-com.L <- list(Abundance = com_abund,
-              'Presence-absence' = com_bin,
-              'Basal area' = com_ba)
-for(j in 1:length(com.L)){
-  var <- data.frame(lapply(lapply(as.list(c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')), function(x){variogram(formula(paste(x, '~ 1')), as(com.L[[j]], 'SpatialPointsDataFrame'))}), function(x){x[3]}))
+dist <- variogram(WD.cwm~1, as(com$abundance, 'SpatialPointsDataFrame'))$dist
+par(mfrow = c(2,1))
+for(i in 1:2){
+  traits <- paste0(c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES'), c('.cwm','.cwv')[i])
+  plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', main = c('(a)', '(b)')[i], xlab = 'distance (m)', ylab = 'relative variance (%)')
+  for(j in 1:length(com)){
+    var <- data.frame(lapply(lapply(as.list(traits), function(x){variogram(formula(paste(x, '~ 1')), as(com[[j]], 'SpatialPointsDataFrame'))}), function(x){x[3]}))
     var <- data.frame(apply(var, 2, function(x){x/max(x)}))
@@
   }
-legend('bottomright', names(var), fill = colramp)
-legend('bottomleft', names(com.L), lty = c(2:4))
+legend('bottomright', names(var), fill = colramp, cex = 0.7)
+legend('bottomleft', names(com), lty = c(2:4), cex = 0.7)
+}
 ```
 
-Figure 4: Trait variograms for all communities.*Variogram was initiated on* $20*20 m^2$ *communities with the wood density weighted mean, and seems to reach a plateau around* $70 m$.
+Figure 4: Trait variograms for all communities.*Variograms were initiated on* $20*20 m^2$ *communities for all traits with both community weighted means* **(a)** *and variances* **(b)** *(by abundance, presence-absence and basal area).
+Some variaograms seem to reach a plateau around* $70 m$.
 
-Links between community weighted means and environment, for both abiotic and biotic factors, were studied by linear regression on continuous variables and anova with factors.
+In order to see if environmental biotic factors were influencing functionnal traits or if environmental biotic factors were the results of functionnal traits we used structural equation modelling on community weighted means by abundance.
+Three latent variables were defined: *Abiotic environment* defined with our 4 selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
+Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
+Retained model was the first one (**m1**) with a significantly lower Aikaike's Criterion Indice on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)
+
+```{r SEM}
+library(lavaan)
+m1 <- '
+Abiotic =~ Slope+Curvature+Wetness+SouthWesterness
+Traits =~ WD.cwm+Thick.cwm+LA.cwm+LDMC.cwm+SLA.cwm
+Biotic =~ BA+Canopy
+Traits ~ Abiotic+Biotic
+'
+m2 <- '
+Abiotic =~ Slope+Curvature+Wetness+SouthWesterness
+Traits =~ WD.cwm+Thick.cwm+LA.cwm+LDMC.cwm+SLA.cwm
+Biotic =~ BA+Canopy
+Biotic ~ Traits +Abiotic
+'
+fit1 <- sem(m1, data = com$abundance@data)
+fit2 <- sem(m2, data = com$abundance@data)
+semPlot::semPaths(fit1,what='path',whatLabels='std',intercept=FALSE,residuals=FALSE,thresholds=FALSE,nCharNodes=0,nCharEdges=0,groups='manlat',pastel=TRUE,title=TRUE,curveAdjacent=TRUE,ask=FALSE)
+```
+
+Figure 5: Path diagram for structural equation modelling. *Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic*.
+
+Links between both community weighted means and variances with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
 Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables (slope, curvature, wetness and southwesterness for abiotic factors, and canopy height and basal area for biotic one).
@@
 Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance.
-Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 3 and 4 (respectively for abiotic and biotic factors).
+Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 4 and 5 (respectively for community weighted means and variances).
+
+An assessment of spatial autocorrelation (SAC) of the community weighted means was conducted using Moran's I statistic.
+Most of community weighed means, especially in abundance, were significantly positively spatially autocorrelated (see table 3).
+But half of presence-absence and basal area community weighted means spatial autocrorrelation didn't significantly differ from 0 and the other were all low (Moran's I from 0.01 to 0.03).
+Specific leaf area and wood density are slightly spatially correlated suggesting a spatial filter due to environment or local history pressure (as dispersion limitation).
+Leaf dry mater content and more generally the wood economic spectrum possess a stochastic spatial distribution implying that they're not shaped by factors with a spatial organization.
+Finally the leaf economic spectrum, thickness and area are spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter.
+
+```{r SAC}
+library(ape)
+traits <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
+traits <- c(paste0(traits, '.cwm'), paste0(traits, '.cwv'))
+weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
+diag(weights) <- 0
+SAC <- lapply(com, function(x){lapply(as.list(x@data[traits]), Moran.I, weights)})
+SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
+SAC <- data.frame(lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))}))
+names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
+SAC$Trait <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA')
+SAC$Type <- ''
+SAC$Type[c(1,8)] <- c('CWM', 'CWV')
+SAC <- SAC[c('Type', 'Trait', 'Abundance', 'Presence-absence', 'Basal area')]
+kable(SAC, row.names = F)
+```
+
+Table 3: Spatial autocorrelation measured by Moran's I for community weighted means and variances.
+*Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations.*
+
+```{r 11 lm & null models}
+reg.abiot <- '~ Slope + Curvature + Wetness + SouthWesterness'
+reg.biot <- '~ Canopy + BA'
+# LM to simplify with loop or lapply
+LM <- lapply(com, function(x) {with(x@data, list(LES.abiot.cwm = lm(formula(paste('LES.cwm', reg.abiot))), # Abiotic CWM
+                                                 logSLA.abiot.cwm = lm(formula(paste('log(SLA.cwm)', reg.abiot))), 
+                                                 WES.abiot.cwm = lm(formula(paste('WES.cwm', reg.abiot))),
+                                                 WD.abiot.cwm = lm(formula(paste('WD.cwm', reg.abiot))),
+                                                 LDMC.abiot.cwm = lm(formula(paste('LDMC.cwm', reg.abiot))),
+                                                 logLA.abiot.cwm = lm(formula(paste('log(LA.cwm)', reg.abiot))),
+                                                 Thick.abiot.cwm = lm(formula(paste('Thick.cwm', reg.abiot))),
+                                                 LES.abiot.cwv = lm(formula(paste('LES.cwv', reg.abiot))), # Abiotic CWV
+                                                 logSLA.abiot.cwv = lm(formula(paste('log(SLA.cwv)', reg.abiot))), 
+                                                 WES.abiot.cwv = lm(formula(paste('WES.cwv', reg.abiot))),
+                                                 WD.abiot.cwv = lm(formula(paste('WD.cwv', reg.abiot))),
+                                                 LDMC.abiot.cwv = lm(formula(paste('LDMC.cwv', reg.abiot))),
+                                                 logLA.abiot.cwv = lm(formula(paste('log(LA.cwv)', reg.abiot))),
+                                                 Thick.abiot.cwv = lm(formula(paste('Thick.cwv', reg.abiot))),
+                                                 LES.biot.cwm = lm(formula(paste('LES.cwm', reg.biot))), # Biotic CWM
+                                                 logSLA.biot.cwm = lm(formula(paste('log(SLA.cwm)', reg.biot))),
+                                                 WES.biot.cwm = lm(formula(paste('WES.cwm', reg.biot))),
+                                                 WD.biot.cwm = lm(formula(paste('WD.cwm', reg.biot))),
+                                                 LDMC.biot.cwm = lm(formula(paste('LDMC.cwm', reg.biot))),
+                                                 logLA.biot.cwm = lm(formula(paste('log(LA.cwm)', reg.biot))),
+                                                 Thick.biot.cwm = lm(formula(paste('Thick.cwm', reg.biot)))
+                                                 ))})
+LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
+model <- null_model(com, Trees, PFT_sp, lapply(LM, function(y){lapply(y, function(x){x$call})}), SpCode, n = 999, verbose = F, time = T)
+```
 
@@
 
-```{r topo results}
-reg <- '~ Slope + Curvature + Wetness + SouthWesterness'
-LM <- lapply(list(com_abund, com_bin, com_ba),
-             function(x) {with(x@data, list(LES = lm(formula(paste('LES', reg))),
-                                            logSLA = lm(formula(paste('log(SLA)', reg))),
-                                            WES = lm(formula(paste('WES', reg))),
-                                            WD = lm(formula(paste('WD', reg))),
-                                            LDMC = lm(formula(paste('LDMC', reg))),
-                                            logLA = lm(formula(paste('log(LA)', reg))),
-                                            Thick = lm(formula(paste('Thick', reg)))))})
-names(LM) <- c('Abundance', 'Presence/Absence', 'Basal area')
-LM <- lapply(LM, function(y){lapply(y, function(x){
-  step(x, direction = 'both', trace = 0)})})
-TreesL <- list(Trees, Trees_bin, TreesBA)
-comL <- list(com_abund, com_bin, com_ba)
-baL <- list(F, F, T)
-model <- mapply(function(lm, trees, com, ba){
-  null_model(formulas = lapply(lm, function(x){as.character(x$call)[2]}),
-             ba = ba, test = 'lm deviance', PFT_sp = PFT_sp, Trees = trees,
-             com = com, SpCode = SpCode, n = 1000, plot = F, count = F, time = F)},
-  LM, TreesL, comL, baL, SIMPLIFY = F)
-
-kable(regression_table(LM, model, reg))
+```{r CWM abiotic results}
+kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwm'))
 ```
 
-Table 3: Step mesure on abiotic factors and null models summary for traits community weighted means.
+Table 4: Step mesure on abiotic factors and null models summary for traits community weighted means.
 *Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW).
@@
 
+```{r SAC residuals}
+library(ape)
+weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
+diag(weights) <- 0
+SAC <- lapply(LM, function(x){lapply(x, function(y){Moran.I(residuals(y), weights)})})
+SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
+SAC <- data.frame(lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))}))
+names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
+row.names(SAC) <- row.names(model$p.value)
+# kable(SAC)
+```
+
+Model residuals spatial autocorrelation (SAC) is null confirming that our models explain fully the spatial structur of our 5 traits and 2 strategy axes (compare to table 3).
+At the exception of thickness that still shows a significant spatial autocorrelation, suggesting that our model missed spatially structured factors in this case.
+
+```{r CWV abiotic results}
+kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwv'))
+```
+
+Table X: Abiotic CWV
+
 ## Biotic filters
 
-```{r biotic results}
-reg <- '~ Canopy + BA'
-LM <- lapply(list(com_abund, com_bin, com_ba),
-             function(x) {with(x@data, list(LES = lm(formula(paste('LES', reg))),
-                                            logSLA = lm(formula(paste('log(SLA)', reg))),
-                                            WES = lm(formula(paste('WES', reg))),
-                                            WD = lm(formula(paste('WD', reg))),
-                                            LDMC = lm(formula(paste('LDMC', reg))),
-                                            logLA = lm(formula(paste('log(LA)', reg))),
-                                            Thick = lm(formula(paste('Thick', reg)))))})
-names(LM) <- c('Abundance', 'Presence/Absence', 'Basal area')
-LM <- lapply(LM, function(y){lapply(y, function(x){
-  step(x, direction = 'both', trace = 0)})})
-TreesL <- list(Trees, Trees_bin, TreesBA)
-comL <- list(com_abund, com_bin, com_ba)
-baL <- list(F, F, T)
-model <- mapply(function(lm, trees, com, ba){
-  null_model(formulas = lapply(lm, function(x){as.character(x$call)[2]}),
-             ba = ba, test = 'lm deviance', PFT_sp = PFT_sp, Trees = trees,
-             com = com, SpCode = SpCode, n = 1000, plot = F, count = F, time = F)},
-  LM, TreesL, comL, baL, SIMPLIFY = F)
-
-kable(regression_table(LM, model, reg))
+```{r CWM biotic results}
+kable(regression_table(LM, model, reg.biot, select = '.biot.cwm'))
 ```
 
-Table 4: Step mesure on abiotic factors and null models summary for traits community weighted means.
+Table 5: Step mesure on abiotic factors and null models summary for traits community weighted means.
 *Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with canopy height (canopy), basal area (BA), and stem density (StD).
@@
 
+An assessment of spatial autocorrelation (SAC) of model residuals confirm that most of the spatial structure of our traits is not explained by our models because the spatial organization stay the same as the one found on the community weighted mean shown in table 3.
+
+```{r CWV biotic results}
+kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwv'))
+```
+
+Table X: Abiotic CWV
+
+## Individual filters
+
+Absence of significant results with biotic factors and null results with abiotic factors on communities weighted mean brought us to study environmental filtering to the individual level.
+To do so we realized a fourthcorner (our RLQ) analysis with three matrices: *L*, the $Individuals*Species$ matrix linking the two other matrices, *Q*, the $Trait*Species$ matrix representing our functionnal traits, and *R* the $Environment*Individuals$ matrix representing both biotic and abiotic filtering.
+As a result, fourthcorner analysis brought no significant adjusted p-values between our functionnal traits and biotic and abiotic filters at the individual levels.
+Whereas communities show environmental filtering on tree functionnal traits and strategies, environmental filtering doesn't strongly affect trees on the individual level.
+
+```{r RLQ}
+library(ade4)
+Deriv <- DEMderiv(Env$DEM, c('slope', 'curvature', 'aspect', 'wetness'))
+Deriv$SW <- cos(Deriv$Aspect + 220)
+Deriv <- Deriv[[-3]]
+
+# Individuals
+TreesNA <- Trees[-which(is.na(Trees$X)),]
+TreesNA <- TreesNA[-which(is.na(TreesNA$`2013_girth`)),]
+TreesNA <- TreesNA[-which(is.na(extract(Env$DEM, TreesNA[2:3], df = T)[2])),]
+TreesNA <- TreesNA[-which(TreesNA$SpCode %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
+
+# Matrices
+Lm <- TreesNA[1]
+Lm <- data.frame(tapply(rep(1, length(Lm[,1])), list(row.names(Lm), Lm$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Qm <- PFT_sp[which(names(PFT_sp) %in% c('SLA', 'LDMC', 'WD', 'Thick', 'LA'))]
+row.names(Qm) <- PFT_sp$Sp_Code
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Qm <- Qm[names(Lm),]
+Rm <- cbind(extract(Deriv, TreesNA[2:3], df = T)[-1], TreesNA$`2013_girth`)
+names(Rm)[5] <- 'Girth'
+row.names(Rm) <- row.names(Lm)
+
+# RLQ
+L <- dudi.coa(Lm, scannf = FALSE)
+R <- dudi.pca(Rm, scannf = FALSE, row.w = L$lw)
+Q <- dudi.pca(Qm, scannf = FALSE, row.w = L$cw)
+rlq <- rlq(R, L, Q, scannf = FALSE)
+rd <- randtest(rlq)
+fq <- fourthcorner.rlq(rlq, type = "Q.axes")
+fr <- fourthcorner.rlq(rlq, type = "R.axes")
+fqr <- fourthcorner(Rm, Lm, Qm)
+fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
+
+# Results
+# summary(rlq)
+# plot(rlq)
+# s.arrow(rlq$l1)
+# s.arrow(rlq$c1)
+# s.label(rlq$lQ, boxes = FALSE)
+# plot(fqr.adj, alpha = 0.5, stat = "D2")
+# plot(fqr.adj, x.rlq = rlq, alpha = 0.5, stat = "D2", type = "biplot")
+# plot(fq, alpha = 0.1, stat = "D2")
+# plot(fr, alpha = 0.1, stat = "D2")
+# print(fq, stat = 'D')
+# print(fr, stat = 'D')
+```
+
+## Phylogenetic signal
+
+Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 measured traits and the two strategy axes and by looking at the phylogenetic correlogram wiht the `phylosignal` R package (cite).
+Thickness, leaf dry matter content, and wood economic spectrum were the only one to be significantly non null with a positive correlation for all of theim (between 0.01 and 0.03) in a short phylogenetic distance between 0 and approximatively 100.
+
+```{r phylosignal, include=FALSE}
+library(phylosignal)
+library(phytools)
+library(phylobase)
+
+# Building species list
+sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), 2:3]
+sp$genus <- unlist(lapply(strsplit(sp$LatinName, ' ', fixed = T), function(x){x[1]}))
+sp$species <- sub(' ', '_', sp$LatinName, fixed = T)
+sp$family <- sp$Family
+sp <- sp[c('species', 'genus', 'family')]
+sp[which(sp$family == 'Flacourtiaceae'),3] <- 'Salicaceae'
+sp[which(sp$family == 'Steruliaceae'),3] <- 'Malvaceae'
+sp[which(sp$family == 'Cesaliniaceae'),3] <- 'Fabaaceae'
+# sp['cato','family'] <- 'Lamiaceae'
+# sp['clvi',] <- c('Clerodendrum_infortunatum', 'Clerodendrum', 'Lamiaceae')
+# sp['glma','family'] <-'Phyllanthaceae'
+# sp['gote','family'] <- 'Stemonuraceae'
+# sp['goca','family'] <- 'Annonaceae'
+# sp['liol','family'] <- 'Lauraceae'
+# sp['mewi','family'] <- 'Melastomataceae'
+# sp['blde',] <- c('Blachia_andamanica', 'Blachia', 'Euphorbiaceae')
+sp['hepa','family'] <- 'Malvaceae'
+# sp['lemo','family'] <- 'Malvaceae'
+# sp['hubr','family'] <- 'Leguminosae' # Humboldtia is unmatched !
+sp <- sp[-which(row.names(sp) == 'hubr'),]
+sp <- sp[-which(row.names(sp) %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
+
+# Getting phylogeny
+path <- system.file('phylogeny', package = 'Uppangala')
+source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
+phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
+nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
+results <- S.PhyloMaker(sp, phylo, nodes)
+
+# Phylosignal
+tree <- results$Scenario.1
+data <- PFT_sp
+row.names(data) <- sub(' ', '_', row.names(data), fixed = T)
+data <- data[tree$tip.label,c(2:5,7,9:10)]
+p4d <- phylo4d(tree, data)
+barplot.phylo4d(p4d, tree.type = "phylo", tree.ladderize = TRUE)
+# phylosim <- phyloSim(tree = tre, method = "all", nsim = 100, reps = 99)
+# plot.phylosim(phylosim, what = "pval", stacked.methods = TRUE)
+phyCor <- lapply(as.list(names(data)), function(x){phyloCorrelogram(p4d, trait = as.character(x))})
+# par(mfrow = c(2,4))
+# for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
+lipa <- lipaMoran(p4d)
+# barplot.phylo4d(p4d, bar.col=(lipa$p.value < 0.05) + 1, center = F, scale = F)
+```
+
 # *Future directions*
 
-**As a conclusion, some question need to be discussed:**
-
-- **Inclusion or not of the Leaf Tissue Density (LTD) trait**
-- **Interpretation of our PCA axes**
-- **The study of covariogram between community weighted mean and topographical variables**
-- **And more generally, the way of building the analysis (is the present trajectory okay)**
+- Add tree estimated height as biotic factors
+- Add species growth as biotic factors
+- Eveness and other FD package metrics
 

### 2016-06-20 R/CWT.R

@@ -1,0 +1,95 @@
+#' @importFrom plyr ddply
+#' @importFrom sp merge
+#' @importFrom Hmisc wtd.var wtd.mean
+NULL
+
+#' Community weighted trait
+#'
+#' Compute a community weighted metric (mean or variance) with traits
+#'
+#' @param Trees data.frame. Trees data frame
+#' @param com SpatialGridDataFrame. Communities with their environment (abiotic
+#'   and biotic)
+#' @param traits char. selected traits
+#' @param metric char. selected metric to be computed on community
+#' @param weights char. selected weights
+#' @param env char. selected environmental variables to be kept
+#'
+#' @return a list with a community grid data frame for each weighting methods
+#'   containing the environmental values, and the community weighted traits by
+#'   defined metrics
+#'
+#' @export
+#'
+#' @examples
+#'
+CWT <- function(Trees, com,
+                traits = c('LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'),
+                metric = c('mean', 'variance'),
+                weights = c('abundance', 'presence-absence', 'basal area'),
+                env = c('Slope', 'Curvature', 'Wetness', 'SouthWesterness', 'Canopy', 'BA')){
+
+  weights <- unlist(lapply(as.list(weights), function(x){switch (x,
+                                                                 'abundance' = 'ab',
+                                                                 'presence-absence' = 'pa',
+                                                                 'basal area' = 'ba')}))
+  com <- com[c('id', env)]
+  cwt.l <- list()
+
+  for (wgt in weights) {
+
+    # Preparing result
+    i <- (length(cwt.l) + 1)
+    cwt.l[[i]] <- com
+    names(cwt.l)[i] <- switch (wgt,
+                                 'ab' = 'abundance',
+                                 'pa' = 'presence-absence',
+                                 'ba' = 'basal area'
+    )
+
+    # Trees table
+    trees <- switch (wgt,
+                     'ab' = Trees,
+                     'pa' = Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),],
+                     'ba' = TreesBA <- Trees[-which(is.na(Trees$`2013_girth`)),]
+    )
+
+    for (m in metric) {
+
+      # Computing CWT
+      if(wgt %in% c('ab', 'pa')){
+
+        fun <- switch (m,
+                       'mean' = base::mean,
[…]

### 2016-06-20 R/null_model.R

@@ -1,113 +1,87 @@
-#' @importFrom car Anova
-#' @importFrom plyr ddply
-NULL
-
 #' Null model
 #'
-#' Function to compute null model
+#' Function to compute null models
 #'
-#' @param formula char. test formula
-#' @param test char. name of the used test (Anova Fvalue or lm R2)
-#' @param ba logical. True if basal area wighted mean
-#' @param PFT_sp data.frame. PFT values by species
-#' @param Trees data.frame. trees table
-#' @param com SpatialGridDataFrame. communities
-#' @param tcol int. column to keep in trees table
-#' @param ccol int. communities column to keep
+#' @param com list. List of communities
+#' @param PFT data.frame. PFT values by species
+#' @param formulas list. Model formulas
+#' @param SpCode dta.frame. Link between species nd they're code
+#' @param traits char. Choosen traits and strategy axes
+#' @param env char. Choosen environmental variables
+#' @param type char. Communti weighted trait metric
 #' @param n int. number of repetitions
-#' @param plot logical. do all null communitier need to be plot
-#' @param count logical. do the number of the realized null model need to be counted
-#' @param time logical. do the time of the operation need to be measured
+#' @param time logical. Use a timer
+#' @param verbose logical. Use a counter
 #'
-#' @return a list with the formula, the test, the null values, and the community value
-#'
+#' @return List of communities null models
 #' @export
 #'
 #' @examples
-#' # No example yet see Uppangala_PFT vignette
 #'
-null_model = function(formulas,
-                      test,
-                      ba = F,
-                      PFT_sp,
-                      Trees,
-                      com,
-                      SpCode,
-                      tcol = NULL,
-                      cwmcol = c(1:1),
-                      ccol = c(1:15),
-                      n = 1000,
-                      plot = F,
-                      count = F,
-                      time = T){
-  if(is.null(tcol)){
-    tcol <- which(names(Trees) %in% c('SpCode', 'com', '2013_girth'))
-  }
-  formul <- formulas
[…]

### 2016-06-20 R/phylogeny.R

@@ -1,0 +1,11 @@
+#' Phylogeny data
+#'
+#' To work, the package need external data from the Appendix of Qian, H. and Y.
+#' Jin (2015). "An updated megaphylogeny of plants, a tool for generating plant
+#' phylogenies, and an analysis of phylogenetic community structure." Journal of
+#' Plant Ecology.
+#'
+#' @name phylogeny
+NULL
+
+#' => NULL

### 2016-06-20 R/plot.null_model.R

@@ -1,37 +1,0 @@
-#' Plot null model
-#'
-#' Function to plot null model and get null model result
-#'
-#' @param model null_model list. result from null_model function (could do an object)
-#' @param formula char. formula of the model to be plotted in the object
-#' @param model int. indice of the model to be plotted in the object
-#'
-#' @return nothing but a plot of the result
-#'
-#' @export
-#'
-#' @examples
-#' # No example yet see Uppangala_PFT vignette
-#'
-plot_model = function(model, formula = NULL, index = NULL){
-  if(!is.null(formula)){
-    index <- which(model$formula == formula)
-  }
-  model$formula <- model$formula[index]
-  model$nullvalues <- model$nullvalues[,index]
-  model$value <- model$value[[index]]
-  model$pvalue <- model$pvalue[index]
-  model$lim <- model$lim[index]
-  hist(model$nullvalues, freq = F, col = 'lightgrey', border = F, xlab = "", breaks = 20,
-       main = as.character(model$formula), xlim = range(c(model$nullvalues, (model$value+0.2))))
-  lines(density(model$nullvalues),
-       main = paste('Null Model with', model$test, ':', model$formula),
-       xlim = range(c(model$nullvalues, (model$value+0.2))))
-  lines(c(model$value, model$value), c(0,1), col = 'red', lwd = 2)
-  text(model$value + 0.2, 0.1, 'Com\nval', col = 'red')
-  mtext(paste('p-value =', model$pvalue, stars(model$pvalue)), 3)
-  mtext(paste('Repetitions =', model$rep), 1)
-  x <- model$lim
-  lines(c(x, x), c(0,1), lty = 2)
-  text(x + 0.2, 0.1, '0.95', cex = 0.7)
-}

### 2016-06-20 R/regression_table.R

@@ -3,6 +3,7 @@
 #' @param LM list. List of tested regressions
 #' @param reg char. Initial regression formula
 #' @param model list. List of null models issued from null_model function
+#' @param select char. subset of models to summarize
 #'
 #' @return A table summarizing all the regression
 #' @export
@@ -10,7 +11,16 @@
 #' @examples
 #' # No example yet
 #'
-regression_table <- function(LM, model, reg){
+regression_table <- function(LM, model, reg, select = '.abiot.cwm'){
+
+  # Results subsetting
+  LM <- lapply(LM, function(x){x[grep(select, names(x), fixed = T)]})
+  model$formula <- lapply(model$formula, function(x){x[grep(select, names(x), fixed = T)]})
+  model$nstat <- lapply(model$nstat, function(x){x[,grep(select, colnames(x), fixed = T)]})
+  model$stat <- model$stat[grep(select, row.names(model$stat), fixed = T),]
+  model$p.value <- model$p.value[grep(select, row.names(model$p.value), fixed = T),]
+
+  # Regression coefficients
   var <- strsplit(substr(reg,3,nchar(reg)), ' + ', fixed = T)[[1]]
   names(var) <- var
   var_table <- data.frame(matrix(nrow = length(LM[[1]])*3, ncol = (length(var)+2)))
@@ -35,15 +45,20 @@ regression_table <- function(LM, model, reg){
     var["SouthWesterness"] <- 'SW'
   }
   names(var_table) <- c('Trait', 'CWM', var)
+  var_table$Trait <- sub(select, '', var_table$Trait)
+
+  # LM R squarred
   var_table$`R-squared` <- NA
   var_table$`R-squared`[c(1,((1:length(LM[[1]]))*3+1)[-length(LM[[1]])])] <- lapply(LM[[1]], function(x){round(summary(x)$r.squared,2)})
   var_table$`R-squared`[c(1,((1:length(LM[[1]]))*3+2)[-length(LM[[1]])])] <- lapply(LM[[2]], function(x){round(summary(x)$r.squared,2)})
   var_table$`R-squared`[c(3,((1:length(LM[[1]]))*3+3)[-length(LM[[1]])])] <- lapply(LM[[3]], function(x){round(summary(x)$r.squared,2)})
[…]

### 2016-06-23 inst/doc/PreliminaryResults.Rmd
v0.1.9013 FD, all null models and result tables, and tree heights

@@
 # *News*
-Add from the june 20th version :
+Add from the `r Sys.Date()` version :
 
@@
 - Structural Equation Modelling (SEM)
+- Species growth too much unprecised for analysis
+- Eveness and other FD package metrics
 
@@
 Figure 1: Principal component analysis (PCA) performed on individual trait means.
-*Left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
+*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
 See table 1 for abbreviation.*
@@
 
-We can see that as wood density ans specific leaf area are orthogonal following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$).
+We can see that as wood density and specific leaf area are orthogonal following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$).
 Leaf area is almost opposed to wood density and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007.
@@
 com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$SouthWesterness <- sin((com@data$Aspect + 225) / 180 * pi)
-com@data$NorthEasterness <- cos((com@data$Aspect + 225) / 180 * pi)
+com@data$SW <- sin((com@data$Aspect + 225) / 180 * pi)
+com@data$NE <- cos((com@data$Aspect + 225) / 180 * pi)
 com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
@@
 com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
+# com@data$Height <- aggregate(Trees$height, list(Trees$com), mean, na.rm = T)[,2]
 rm(Deriv, XY)
@@
 
-To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (aspect in direction of south-west), and distance to the nearest potential river.
+To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
 Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
@@
 Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
-As a conclusion, we choose to reduce our variable to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness to represent monsoon winds [@bunyan_effect_2015] and slope.
+As a conclusion, we choose to reduce our variable to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness (SW) to represent monsoon winds [@bunyan_effect_2015] and slope.
 
 ```{r PCA Env}
-pca.env <- princomp(~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SouthWesterness, data = com@data, cor = T)
+pca.env <- princomp(~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
 lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
@@
 Figure 2: Principal component analysis (PCA) performed on topgraphical variables.
-*Left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).*
+*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
+SW stand for southwesterness.*
 
@@
 
-To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density (StD) *(we may add tree height and growth in the future)*.
+To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density *(we may add tree height and growth in the future)*.
 Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$.
@@
 ```{r 5 Com}
+library(FD)
 com <- CWT(Trees, com)
+x <- data.frame(PFT_sp[-1], row.names = PFT_sp$Sp_Code)
+a <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+a[is.na(a)] <- 0
+x <- x[names(a),]
+FD <- as.data.frame(dbFD(x, a, corr = 'cailliez', calc.CWM = F, print.pco = F, messages = F))[-c(1:2,4)]
+names(FD) <- paste0(names(FD), '.fd')
+FD$id <- row.names(FD)
+com$abundance <- merge(com$abundance, FD)
 ```
@@
 Weighted mean and variances were realized by abundance, presence-absence, or basal area.
+In addition functionnal diversity indices weighted by abundances were computed with the `FD` package [@laliberte_distance-based_2010] for our communities: functional richness (FRic), functional evenness (FEve), and functional divergence (FDiv) and Rao's quadratic entropy (Q) (cite Villéger et al.
+(2008) et Botta-Dukát 2005).
 
 Spatial variograms of community weighted means and variances didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$.
+Functionnal diversity indices show also a plateau around $70 m$.
 
@@
 m1 <- '
-Abiotic =~ Slope+Curvature+Wetness+SouthWesterness
-Traits =~ WD.cwm+Thick.cwm+LA.cwm+LDMC.cwm+SLA.cwm
-Biotic =~ BA+Canopy
-Traits ~ Abiotic+Biotic
+Abiotic =~ Slope + Curvature + Wetness + SW
+Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
+Biotic =~ BA + Canopy
+Traits ~ Abiotic + Biotic
 '
 m2 <- '
-Abiotic =~ Slope+Curvature+Wetness+SouthWesterness
-Traits =~ WD.cwm+Thick.cwm+LA.cwm+LDMC.cwm+SLA.cwm
-Biotic =~ BA+Canopy
+Abiotic =~ Slope + Curvature + Wetness + SW
+Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
+Biotic =~ BA + Canopy
 Biotic ~ Traits +Abiotic
@@
 
-Figure 5: Path diagram for structural equation modelling. *Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic*.
+Figure 5: Path diagram for structural equation modelling.
+*Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic.
+Arrow values are indicating standardized parameter estimates*.
 
@@
 Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 4 and 5 (respectively for community weighted means and variances).
+Same procedure was also applied to functionnal diversity indices (FD) from our ocmmunities, results are presented in table 6.
 
-An assessment of spatial autocorrelation (SAC) of the community weighted means was conducted using Moran's I statistic.
-Most of community weighed means, especially in abundance, were significantly positively spatially autocorrelated (see table 3).
-But half of presence-absence and basal area community weighted means spatial autocrorrelation didn't significantly differ from 0 and the other were all low (Moran's I from 0.01 to 0.03).
-Specific leaf area and wood density are slightly spatially correlated suggesting a spatial filter due to environment or local history pressure (as dispersion limitation).
-Leaf dry mater content and more generally the wood economic spectrum possess a stochastic spatial distribution implying that they're not shaped by factors with a spatial organization.
-Finally the leaf economic spectrum, thickness and area are spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter.
+An assessment of spatial autocorrelation (SAC) of the community weighted means and variances and functionnal diversity indices was conducted using Moran's I statistic.
+Only half of community weighted means and variances were significantly positively spatially autocorrelated (see table 3) with low values (Moran's I from 0.01 to 0.04).
+Leaf economic spectrum and dry matter content weighted means as wood density weighted variances are slightly spatially correlated suggesting a spatial filter due to environment or local history pressure (as dispersion limitation).
+On the opposite, wood density and thickness community weighted means as wood economic spectrum community weighted variances possess a stochastic spatial distribution implying that they're not shaped by factors with a spatial organization.
+Finally remaining traits community weighted means and variances are spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter.
+As a conclusion both community weighted means and variances are most of the time stochastic or driven by dispersion.
+Moreover all functionnal diversity indices (FD) are spatially correlated.
+
+Table 3: Spatial autocorrelation measured by Moran's I for community weighted means and variances.
+*Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations.*
 
@@
 traits <- c(paste0(traits, '.cwm'), paste0(traits, '.cwv'))
+traits <- list(c(traits, 'FRic.fd', 'FEve.fd', 'FDiv.fd', 'FDis.fd', 'RaoQ.fd'), traits, traits)
 weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
 diag(weights) <- 0
-SAC <- lapply(com, function(x){lapply(as.list(x@data[traits]), Moran.I, weights)})
+SAC <- mapply(function(x,y){lapply(as.list(x@data[y]), Moran.I, weights)}, x =com, y = traits)
 SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
-SAC <- data.frame(lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))}))
+SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
+SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
+SAC <- data.frame(SAC)
 names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
-SAC$Trait <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA')
+SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
 SAC$Type <- ''
-SAC$Type[c(1,8)] <- c('CWM', 'CWV')
-SAC <- SAC[c('Type', 'Trait', 'Abundance', 'Presence-absence', 'Basal area')]
+SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
+SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
 kable(SAC, row.names = F)
@@
 
-Table 3: Spatial autocorrelation measured by Moran's I for community weighted means and variances.
-*Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations.*
-
-```{r 11 lm & null models}
-reg.abiot <- '~ Slope + Curvature + Wetness + SouthWesterness'
-reg.biot <- '~ Canopy + BA'
+```{r 6 lm & null models}
+reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
 # LM to simplify with loop or lapply
-LM <- lapply(com, function(x) {with(x@data, list(LES.abiot.cwm = lm(formula(paste('LES.cwm', reg.abiot))), # Abiotic CWM
-                                                 logSLA.abiot.cwm = lm(formula(paste('log(SLA.cwm)', reg.abiot))), 
-                                                 WES.abiot.cwm = lm(formula(paste('WES.cwm', reg.abiot))),
-                                                 WD.abiot.cwm = lm(formula(paste('WD.cwm', reg.abiot))),
-                                                 LDMC.abiot.cwm = lm(formula(paste('LDMC.cwm', reg.abiot))),
-                                                 logLA.abiot.cwm = lm(formula(paste('log(LA.cwm)', reg.abiot))),
-                                                 Thick.abiot.cwm = lm(formula(paste('Thick.cwm', reg.abiot))),
-                                                 LES.abiot.cwv = lm(formula(paste('LES.cwv', reg.abiot))), # Abiotic CWV
-                                                 logSLA.abiot.cwv = lm(formula(paste('log(SLA.cwv)', reg.abiot))), 
-                                                 WES.abiot.cwv = lm(formula(paste('WES.cwv', reg.abiot))),
-                                                 WD.abiot.cwv = lm(formula(paste('WD.cwv', reg.abiot))),
-                                                 LDMC.abiot.cwv = lm(formula(paste('LDMC.cwv', reg.abiot))),
-                                                 logLA.abiot.cwv = lm(formula(paste('log(LA.cwv)', reg.abiot))),
-                                                 Thick.abiot.cwv = lm(formula(paste('Thick.cwv', reg.abiot))),
-                                                 LES.biot.cwm = lm(formula(paste('LES.cwm', reg.biot))), # Biotic CWM
-                                                 logSLA.biot.cwm = lm(formula(paste('log(SLA.cwm)', reg.biot))),
-                                                 WES.biot.cwm = lm(formula(paste('WES.cwm', reg.biot))),
-                                                 WD.biot.cwm = lm(formula(paste('WD.cwm', reg.biot))),
-                                                 LDMC.biot.cwm = lm(formula(paste('LDMC.cwm', reg.biot))),
-                                                 logLA.biot.cwm = lm(formula(paste('log(LA.cwm)', reg.biot))),
-                                                 Thick.biot.cwm = lm(formula(paste('Thick.cwm', reg.biot)))
-                                                 ))})
+LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
+                                                 logSLA.cwm = lm(formula(paste('log(SLA.cwm)', reg))), 
+                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
+                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
+                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
+                                                 logLA.cwm = lm(formula(paste('log(LA.cwm)', reg))),
+                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
+                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
+                                                 logSLA.cwv = lm(formula(paste('log(SLA.cwv)', reg))), 
+                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
+                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
+                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
+                                                 logLA.cwv = lm(formula(paste('log(LA.cwv)', reg))),
+                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
+LM$abundance <- c(LM$abundance, with(com$abundance@data, list(FRic.fd = lm(formula(paste('FRic.fd', reg))),
+                                                              FEve.fd = lm(formula(paste('FEve.fd', reg))),
+                                                              FDiv.fd = lm(formula(paste('FDiv.fd', reg))),
+                                                              FDis.fd = lm(formula(paste('FDis.fd', reg))),
+                                                              RaoQ.fd = lm(formula(paste('RaoQ.fd', reg))))))
 LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
@@
 
-## Abiotic filters
+## Community weighted means
 
-```{r CWM abiotic results}
-kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwm'))
+Table 4: Summary for linear models and null models for traits community weighted means.
+*Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+See table 1 for abbreviations.*
+
+```{r CWM}
+kable(regression_table(LM, model, reg, select = '.cwm'))
 ```
 
-Table 4: Step mesure on abiotic factors and null models summary for traits community weighted means.
-*Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW).
-Model deviance was compared to 999 null models deviance in which traits were randomized among species.*
+### Abiotic factors
 
@@
 Leaf economic spectrum (LES) and specific leaf area (SLA) is associated non-randomly with convex and dry surfaces, as the thickness for basal area community weighted mean.
-Trees will tend to increase their surface compare to their carbon investment in areas where the water will flow but not accumulate nor stay, and where the soil will tend to accumulate and be stabilized.
+Trees will tend to increase their surface compare to their carbon invest ment in areas where the water will flow but not accumulate nor stay, and where the soil will tend to accumulate and be stabilized.
 In favourable water and soil conditions, trees will tend to invest more in leaf ressource capture at the expense of leaf defense, adopting the acquisitive [@sterck_functional_2011] our fast growing strategy [@reich_world-wide_2014].
 
-Wood density (WD) in presence absence shows a non-random association with southwest steep slope and concave surfaces with a relatively important explanation level ($R^2=0.23$).
+Wood density (WD) in presence absence shows a non-random association with southwest steep slope and concave surfaces with a relatively important explanation level ($R^2=0.25$).
 Trees will increase their wood density in thin soil and water and soil departure conditions under potential monsoon winds *(We need to decide or not if regarding the plot condition we can accept or not the southwest orientation as being linked to monsoon winds)*.
 In water stress and mechanical hazard conditions, trees will compensate by increasing their water transport safety and mechanical resistance and resilience with an increased wood density [@chave_towards_2009].
+
+### Biotic factors
+
+Most of community weighted means don't show association with biotic factors, and most of the rare association are not significantly different from null models randomly generated.
+Still leaf area (LA) in basal area is associated with high trees.
+Voluminous individuals tend to increase their leaf surface in heigher canopy condition to compensate understory shadow.
+And wood density tends also to increase with tree height for rare species, suggesting that with higher canopy denser wood species will tend to grow in a community, as shade species.
+
+## Community weighted variances
+
+Table 5: Summary for linear models and null models for traits community weighted variances.
+*Linear model results for each trait community weighted variances (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+See table 1 for abbreviations.*
+
+```{r CWV}
+kable(regression_table(LM, model, reg, select = '.cwv'))
+```
+
+## Functionnal diversity
+
+Table 6: Summary for linear models and null models for functionnal diversity indices (FD).
+*Linear model results for each community functionnal indices weighted by abundances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.*
+
+```{r FD}
+kable(regression_table(LM, model, reg, select = '.fd'))
+```
+
+## Models residuals
+
+Table 6: Spatial autocorrelation measured by Moran's I for linear regression model residuals from community weighted means (CWM) and variances (CWV) and functionnal diversity indices (FD).
+*Spatial autocorrelation for residuals from each trait community weighted means and variances (by abundance, presence-absence and basal area) linear models and for functionnal diversity (FD) linear models, see table 1 for abbreviations.*
 
@@
 SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
-SAC <- data.frame(lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))}))
+SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
+SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
+SAC <- data.frame(SAC)
 names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
-row.names(SAC) <- row.names(model$p.value)
-# kable(SAC)
+SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
+SAC$Type <- ''
+SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
+SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
+kable(SAC, row.names = F)
 ```
 
-Model residuals spatial autocorrelation (SAC) is null confirming that our models explain fully the spatial structur of our 5 traits and 2 strategy axes (compare to table 3).
+Model residuals spatial autocorrelation (SAC)  for our community weighted means (CWM) is null confirming that our models explain fully the spatial structure of our 5 traits and 2 strategy axes (compare to table 3).
 At the exception of thickness that still shows a significant spatial autocorrelation, suggesting that our model missed spatially structured factors in this case.
-
-```{r CWV abiotic results}
-kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwv'))
-```
-
-Table X: Abiotic CWV
-
-## Biotic filters
-
-```{r CWM biotic results}
-kable(regression_table(LM, model, reg.biot, select = '.biot.cwm'))
-```
-
-Table 5: Step mesure on abiotic factors and null models summary for traits community weighted means.
-*Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with canopy height (canopy), basal area (BA), and stem density (StD).
-Model deviance was compared to 999 null models deviance in which traits were randomized among species.*
-
-Most of community weighted means don't show association with biotic factors, and most of the rare association are not significantly different from null models randomly generated.
-Still leaf area (LA) in basal area is associated with high trees.
-Voluminous individuals tend to increase their leaf surface in heigher canopy condition to compensate understory shadow.
-
-An assessment of spatial autocorrelation (SAC) of model residuals confirm that most of the spatial structure of our traits is not explained by our models because the spatial organization stay the same as the one found on the community weighted mean shown in table 3.
-
-```{r CWV biotic results}
-kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwv'))
-```
-
-Table X: Abiotic CWV
+For models residuals from community weighted variances, specific leaf area, wood economic spectrum and leaf dry matter content still show a spatial structure.
+This spatial structure in local trait variances linked to alpha diversity may be explained by local history and stochasticty associated with a null distribution of individuals not directly shaped by their environment.
+Finally we find back this unexplained by model spatial structure for functionnal diversity indices.
 
@@
 
-Absence of significant results with biotic factors and null results with abiotic factors on communities weighted mean brought us to study environmental filtering to the individual level.
+Absence of significant results with biotic factors and null results with abiotic factors on communities weighted mean brought us to study environmental filtering at the individual level.
 To do so we realized a fourthcorner (our RLQ) analysis with three matrices: *L*, the $Individuals*Species$ matrix linking the two other matrices, *Q*, the $Trait*Species$ matrix representing our functionnal traits, and *R* the $Environment*Individuals$ matrix representing both biotic and abiotic filtering.
@@
 # plot(fqr.adj, alpha = 0.5, stat = "D2")
-# plot(fqr.adj, x.rlq = rlq, alpha = 0.5, stat = "D2", type = "biplot")
+# plot(fqr.adj, x.rlq = rlq, alpha = 0.05, stat = "D2", type = "biplot")
 # plot(fq, alpha = 0.1, stat = "D2")
@@
 
-Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 measured traits and the two strategy axes and by looking at the phylogenetic correlogram wiht the `phylosignal` R package (cite).
+Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 measured traits and the two strategy axes and by looking at the phylogenetic correlogram wiht the `phylosignal` R package [@keck_phylosignal:_2016].
 Thickness, leaf dry matter content, and wood economic spectrum were the only one to be significantly non null with a positive correlation for all of theim (between 0.01 and 0.03) in a short phylogenetic distance between 0 and approximatively 100.
@@
 
-- Add tree estimated height as biotic factors
-- Add species growth as biotic factors
-- Eveness and other FD package metrics
+- Tree estimated height with `BIOMASS`
 

### 2016-06-23 R/CWT.R

@@ -27,7 +27,7 @@ CWT <- function(Trees, com,
                 traits = c('LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'),
                 metric = c('mean', 'variance'),
                 weights = c('abundance', 'presence-absence', 'basal area'),
-                env = c('Slope', 'Curvature', 'Wetness', 'SouthWesterness', 'Canopy', 'BA')){
+                env = c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')){
 
   weights <- unlist(lapply(as.list(weights), function(x){switch (x,
                                                                  'abundance' = 'ab',

### 2016-06-23 R/draft.R

@@ -170,3 +170,40 @@
 # plot(results$Scenario.1,cex=1.1,main="Scenarion One")
 # plot(results$Scenario.2,cex=1.1,main="Scenarion Two")
 # plot(results$Scenario.3,cex=1.1,main="Scenarion Three")
+
+# ## FD
+#
+# # Building
+# x <- data.frame(PFT_sp[-1], row.names = PFT_sp$Sp_Code)
+# a <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+# a[is.na(a)] <- 0
+# x <- x[names(a),]
+# FD <- as.data.frame(dbFD(x, a, corr = 'cailliez', calc.CWM = F, print.pco = F, messages = F))[-c(1:2,4)]
+# FD$id <- row.names(FD)
+# com_FD <- merge(com$abundance, FD)
+# plot(stack(com_FD)[[22:26]])
+#
+# # Variogram
+# library(gstat)
+# dist <- variogram(FRic~1, as(com_FD, 'SpatialPointsDataFrame'))$dist
+# traits <- c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ')
+# plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', xlab = 'distance (m)', ylab = 'relative variance (%)')
+# var <- data.frame(lapply(lapply(as.list(traits), function(x){variogram(formula(paste(x, '~ 1')), as(com_FD, 'SpatialPointsDataFrame'))}), function(x){x[3]}))
+# var <- data.frame(apply(var, 2, function(x){x/max(x)}))
+# names(var) <- traits
+# # colramp <- rainbow(length(var))
+# # for(i in 1:length(var)){
+# #   lines(var[,i] ~ dist, col = colramp[i])
+# # }
+# # legend('bottomright', names(var), fill = colramp, cex = 0.7)
+#
+# # SAC
+# library(ape)
+# traits <- c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ')
+# weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
+# diag(weights) <- 0
+# SAC <- do.call(rbind.data.frame, lapply(as.list(com_FD@data[traits]), Moran.I, weights))
+# SAC <- paste(round(SAC$observed, 2), unlist(lapply(as.list(SAC$p.value), stars)))
[…]

### 2016-06-23 R/genSpecies.R

@@ -9,12 +9,13 @@
 #' summary(genSpecies())
 #'
 genSpecies <- function(){
+
   # Opening
-  path <- system.file('extdata', 'Species.txt', package = 'Uppangala')
-  species <- read.table(path,sep=",",dec=".", stringsAsFactors =F) # Dans BDD Uppangala
+  path <- system.file('extdata', 'Species.csv', package = 'Uppangala')
+  species <- read.table(path,sep=",",dec=".", stringsAsFactors =F, h = T) # Dans BDD Uppangala
   # Formatting table
-  species <- species[-1,c(3,4,10,5,9,11:13)]
-  colnames(species) <- c("SpCode","LatinName","Family","Authority","VoucherRef","PotStrata","SpError","SpNote")
+  species <- species[c('spco', 'Latin_name', 'Family', 'PotStrata')]
+  colnames(species) <- c("SpCode","LatinName","Family","PotStrata")
   species <- species[-which(is.na(species$LatinName)),]
   species$SpCode <- tolower(species$SpCode)
   rownames(species) <- species$SpCode
@@ -36,5 +37,6 @@ genSpecies <- function(){
   species$Strata[species$ESIV==1] <- 'ESIV'
   species$Strata <- as.factor(species$Strata)
   species <- species[-which(names(species) %in% levels(species$Strata))]
+
   return(species)
 }

### 2016-06-23 R/genTrees.R

@@ -1,3 +1,6 @@
+#' @importFrom BIOMASS modelHD predictHeight
+NULL
+
 #' Trees data generation
 #'
 #' Generates Trees data set from external data extracted from the database
@@ -40,6 +43,29 @@ genTrees <- function(){
   # Opening girth 5 ha plots
   path <- system.file('extdata', 'Girth_5Ha.csv', package = 'Uppangala')
   girth5 <- read.table(path, header = T, sep = ",", dec = ".")
+  girth5$ID.date <- paste(girth5$Tree_5Ha_ID,
+                          substr(as.character(girth5$Census_date), start = 7, stop = 11),
+                          substr(as.character(girth5$Census_date), start = 4, stop = 5),
+                          sep = '_')
+  # Opening height 5 ha plots
+  path <- system.file('extdata', 'Height_5Ha.csv', package = 'Uppangala')
+  height5 <- read.table(path, header = T, sep = ",", dec = ".")
+  height5$ID.date <- paste(height5$Tree_5Ha_ID,
+                          substr(as.character(height5$Census_date), start = 7, stop = 11),
+                          substr(as.character(height5$Census_date), start = 4, stop = 5),
+                          sep = '_')
+  # Modelling height girth relation
+  IDs <- intersect(height5$ID.date, girth5$ID.date)
+  H <- data.frame(H = height5$Total_height, row.names = height5$ID.date)
+  H[H == -999] <- NA
+  H <- H[IDs,]
+  D <- data.frame(D = girth5$Girth/2*pi, ID = girth5$ID.date)
+  D <- D[which(!duplicated(D$ID)),]
+  row.names(D) <- D$ID
+  D[D == -999] <- NA
+  D <- D[IDs,1]
+  m <- modelHD(D, H, method = 'michaelis')
+  # Preparing girth
   girth5$ID <- girth5$Tree_5Ha_ID
   girth5$Date <- paste(substr(as.character(girth5$Census_date), start = 7, stop = 11),
                        substr(as.character(girth5$Census_date), start = 4, stop = 5),
@@ -71,6 +97,7 @@ genTrees <- function(){
   tree <- tree[-which(is.na(tree$SpCode)),]
   row.names(tree) <- tree$ID
   tree <- tree[-1]
+  tree$height <- predictHeight(tree$`2013_girth`, m)
   summary(tree)
 
   return(tree)

### 2016-06-23 R/null_model.R

@@ -1,3 +1,6 @@
+#' @importFrom FD dbFD
+NULL
+
 #' Null model
 #'
 #' Function to compute null models
@@ -19,21 +22,21 @@
 #' @examples
 #'
 null_model <- function(com,
-                        Trees,
-                        PFT,
-                        formulas,
-                        SpCode,
-                        traits = c('LES', 'SLA', 'WES','WD', 'LDMC', 'Thick', 'LA'),
-                        env = c('Slope', 'Curvature', 'Wetness', 'SouthWesterness', 'Canopy', 'BA'),
-                        type = c('CWM', 'CWV'),
-                        n = 999,
-                        time = T,
-                        verbose = T){
+                       Trees,
+                       PFT,
+                       formulas,
+                       SpCode,
+                       traits = c('LES', 'SLA', 'WES','WD', 'LDMC', 'Thick', 'LA'),
+                       env = c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA'),
+                       type = c('CWM', 'CWV'),
+                       n = 999,
+                       time = T,
+                       verbose = T){
 
   t0 <- Sys.time()
   nstat <- list(abundance = matrix(ncol = length(formulas[[1]]), nrow = n, dimnames = list(1:n, names(formulas[[1]]))),
-                `presence-absence` = matrix(ncol = length(formulas[[1]]), nrow = n, dimnames = list(1:n, names(formulas[[1]]))),
-                `basal area` = matrix(ncol = length(formulas[[1]]), nrow = n, dimnames = list(1:n, names(formulas[[1]])))
+                `presence-absence` = matrix(ncol = length(formulas[[2]]), nrow = n, dimnames = list(1:n, names(formulas[[2]]))),
+                `basal area` = matrix(ncol = length(formulas[[3]]), nrow = n, dimnames = list(1:n, names(formulas[[3]])))
                 )
 
   # Null Communities
@@ -54,6 +57,15 @@ null_model <- function(com,
                                                         'CWV' = 'variance'
     )})
     ncom <- CWT(nTrees, com[[1]], traits = traits, metric = metric, weights = names(com), env = env)
+    # Addind FD values
[…]

### 2016-06-23 R/regression_table.R

@@ -11,54 +11,63 @@
 #' @examples
 #' # No example yet
 #'
-regression_table <- function(LM, model, reg, select = '.abiot.cwm'){
+regression_table <- function(LM, model, reg, select = '.cwm'){
 
   # Results subsetting
   LM <- lapply(LM, function(x){x[grep(select, names(x), fixed = T)]})
   model$formula <- lapply(model$formula, function(x){x[grep(select, names(x), fixed = T)]})
   model$nstat <- lapply(model$nstat, function(x){x[,grep(select, colnames(x), fixed = T)]})
-  model$stat <- model$stat[grep(select, row.names(model$stat), fixed = T),]
-  model$p.value <- model$p.value[grep(select, row.names(model$p.value), fixed = T),]
+  model$stat <- lapply(model$stat, function(x){x[grep(select, names(x), fixed = T)]})
+  model$p.value <- lapply(model$p.value, function(x){x[grep(select, names(x), fixed = T)]})
 
   # Regression coefficients
   var <- strsplit(substr(reg,3,nchar(reg)), ' + ', fixed = T)[[1]]
   names(var) <- var
   var_table <- data.frame(matrix(nrow = length(LM[[1]])*3, ncol = (length(var)+2)))
-  names(var_table) <- c('Trait', 'CWM', var)
+  names(var_table) <- c('Trait', toupper(substr(select,2,4)), var)
   var_table$Trait[c(1,((1:length(LM[[1]]))*3+1)[-length(LM[[1]])])] <- names(LM[[1]])
-  var_table$CWM <- rep(names(LM), length(LM[[1]]))
+  var_table[,2] <- rep(c('AB', 'PA', 'BA'), length(LM[[1]]))
   for(j in 1:length(LM)){
+    if(select == '.fd'){
+      j <-  1
+    }
     for(i in 1:length(LM[[1]])){
       for(v in var){
         row <- which(var_table$Trait == names(LM[[1]])[i]) + j - 1
         if(v %in% row.names(summary(LM[[j]][[i]])$coefficients)){
           var_table[row,v] <- paste(format(summary(LM[[j]][[i]])$coefficients[v,1],
-                                         scientific = T, digits = 2),
-                                  stars(summary(LM[[j]][[i]])$coefficients[v,4]))
[…]

### 2016-06-27 inst/doc/PreliminaryResults.Rmd
v0.1.9014 Article draft beginning

@@
 Biotic =~ BA + Canopy
-Biotic ~ Traits +Abiotic
+Biotic ~ Traits + Abiotic
 '
@@
 Table 3: Spatial autocorrelation measured by Moran's I for community weighted means and variances.
-*Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations.*
+_Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations.
+FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 Table 4: Summary for linear models and null models for traits community weighted means.
-*Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+_Linear model results for each trait community weighted means (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
 Model deviance was compared to 999 null model deviances in which traits were randomized among species.
-See table 1 for abbreviations.*
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 Table 5: Summary for linear models and null models for traits community weighted variances.
-*Linear model results for each trait community weighted variances (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+_Linear model results for each trait community weighted variances (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
 Model deviance was compared to 999 null model deviances in which traits were randomized among species.
-See table 1 for abbreviations.*
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 
+Most of traits and tree strategy axes didn't show linear models for community weighted variances (CWV) significantyl differently affected by both abiotic and biotic environment compare to null models.
+Only the leaf economic spectrum (LES) and wood density (WD) showed linear models significantly different from null models for their community weighted variances (CWV) for all wieghts (abundance, presence-absence, and basal area).
+Leaf area community weighted variance by basal area showed also a significantly positive linear model with slope compare to null models.
+
+Leaf economic spectra (LES) converge with better soil and water conditions in convex situations, and for abundant species oriented under monsoon wind LES diverge suggesting different adaptive strategies from species.
+Whereas wood density from abundant and rare species also diverged with the southwest orientation to monsoon winds, wood density from voluminous species converge with slope and diverge with the canopy height.
+Effectively under high canopy trees will be under a lot of shade situation, implying differing growth strategies and further more different wood densities.
+Leaf area also converge for voluminous species with slope.
+
 ## Functionnal diversity
 
+Whereas functionnal richness (FRic) standardized by the *'global FRic'* that include all species is low and relatively stable (mean = 0.05, sd = 0.04);
+functionnal eveness is high and relatively stable (mean = `r round(mean(com$abundance@data$FEve.fd), 2)`, sd = `r round(sd(com$abundance@data$FEve.fd), 2)`) in our communities as the functionnal diversity (mean = `r round(mean(com$abundance@data$FDiv.fd), 2)`, sd = `r round(sd(com$abundance@data$FDiv.fd), 2)`).
+Our communties are relatively poor in functionnal values but with an even and diverse trait distribution.
+
 Table 6: Summary for linear models and null models for functionnal diversity indices (FD).
-*Linear model results for each community functionnal indices weighted by abundances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model deviance was compared to 999 null model deviances in which traits were randomized among species.*
+_Linear model results for each community functionnal indices weighted by abundances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
 ```{r FD}
-kable(regression_table(LM, model, reg, select = '.fd'))
+kable(regression_table(LM, model, reg, select = '.fd')[-2], row.names = F)
 ```
+
+At the exception of the functionnal richness (FRic), all linear models for functionnal diversity indices didn't differ from null models suggesting that our bioctic and abitotic factors are not responsible for their distribution in our communities.
+Still functionnal richness increases with slope, suggesting as commmunity weighted variances (CWV) that our traits diverge in different adaptive strategies in steeper slopes.
 
@@
 Table 6: Spatial autocorrelation measured by Moran's I for linear regression model residuals from community weighted means (CWM) and variances (CWV) and functionnal diversity indices (FD).
-*Spatial autocorrelation for residuals from each trait community weighted means and variances (by abundance, presence-absence and basal area) linear models and for functionnal diversity (FD) linear models, see table 1 for abbreviations.*
+_Spatial autocorrelation for residuals from each trait community weighted means and variances (by abundance, presence-absence and basal area) linear models and for functionnal diversity (FD) linear models, see table 1 for abbreviations.
+FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 This spatial structure in local trait variances linked to alpha diversity may be explained by local history and stochasticty associated with a null distribution of individuals not directly shaped by their environment.
-Finally we find back this unexplained by model spatial structure for functionnal diversity indices.
+Finally we find back this spatial structure unexplained by model for functionnal diversity indices.
 

### 2016-06-27 inst/doc/draft.Rmd
v0.1.9014 Article draft beginning

---
title: "Uppangala tree functionnal traits: Draft"
author: "Sylvain Schmitt"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: refs.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

# Abstract

Keywords: functionnal trait, tropical evergreen forest, Western Ghats, India...

# Introduction

# Material and methods

## Study site

The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village. The permanent plot was established in 1990 by the French Institute of Pondicherry and described by Pascal and Pelissier [@pascal_structure_1996]. The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergree forest type [@pascal_wet_1988]. Climate and geology of the 10-ha plot can be considered as homogeneous. General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].

*Grith monitoring design ?*

## Sampling

We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates. Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals. Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot. For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013. Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy. Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.

- Unsampled species beanch wood 
- Leaves manipulation conditions.

Among leaves, 5 were weighted fresh, scanned [cite scan], and their thickness were measured with a microgauge [cite]. Those leaves were kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying processed was stopped after dry leaf weights were stable in time. Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias. Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues. Branch wood samples were then sink in water to evaluate their volume by water displacement. Samples were after dried for 48 hours at 80 degrees celsius and weighted.

## Traits

- Traits measurement (handbook)
- 5 Basic traits description
- Associated strategies axes with PCA (+ fortunel)
- Associated covariation (following litterature)
- Diversity indices (FD)

## Environmental variables

- Abitoic factors
- PCA reduction
- Biotic factors
- Bitoic uncorrelation
- Performance variables
- Linking traits, biotic, abiotic factors and performance variables with sem

## Analyses

- Community defintions
- Community spatial organization investigation (variorgam + SAC)
- Link with environment: step measure, linear regression, null model, residuals SAC
- Individual filtering (RLQ)
- Phylogenetic signal

# Results

## Community weighted means

### Abiotic filtering

### Biotic filtering

## Community weighted variances

### Abiotic filtering

### Biotic filtering

## Functionnal diversity

## Individual filtering

## Phylogenetic signal

# Discussion

# *Acknowledgement*

# References

### 2016-06-29 inst/doc/PreliminaryResults.Rmd
v0.1.9015 cosaspect DEMderiv.R #AV

@@
 
+```{r RLQ2, eval=FALSE, include=FALSE}
+library(ade4)
+
+# Matrices
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+# Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum)) #BA
+Lm[is.na(Lm)] <- 0
+Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
+row.names(Qm) <- PFT_sp$Sp_Code
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Qm <- Qm[names(Lm),]
+Rm <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+row.names(Rm) <- com@data$id
+
+# Consistency
+all(row.names(Lm) == row.names(Rm))
+all(row.names(Lm) %in% row.names(Rm))
+Lm <- Lm[row.names(Rm),]
+all(names(Lm) == row.names(Qm))
+Qm <- Qm[names(Lm),]
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Lm <- Lm[row.names(Qm)]
+Lm[Lm > 0] <-  1 #Presence-absence
+Lm <- Lm / rowSums(Lm)
+
+# RLQ
+L <- dudi.coa(Lm, scannf = F)
+R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
+Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
+rlq <- rlq(R, L, Q, scannf = F)
+model = 6
+rd <- randtest(rlq, modeltype = model)
+fq <- fourthcorner.rlq(rlq, type = "Q.axes", modeltype = model)
+fr <- fourthcorner.rlq(rlq, type = "R.axes", modeltype = model)
+fqr <- fourthcorner(Rm, Lm, Qm, modeltype = model)
+fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
+
+# Results
+summary(rlq)
+plot(rlq)
+s.arrow(rlq$l1)
+s.arrow(rlq$c1)
+s.label(rlq$lQ, boxes = FALSE)
+rd
+plot(rd)
+alpha = 0.5
+plot(fq, alpha = alpha, stat = "D2")
+plot(fr, alpha = alpha, stat = "D2")
+plot(fqr.adj, alpha = alpha, stat = "D2")
+plot(fqr.adj, x.rlq = rlq, alpha = alpha, stat = "D2", type = "biplot")
+```
+
 ## Communities
@@
 fit2 <- sem(m2, data = com$abundance@data)
-semPlot::semPaths(fit1,what='path',whatLabels='std',intercept=FALSE,residuals=FALSE,thresholds=FALSE,nCharNodes=0,nCharEdges=0,groups='manlat',pastel=TRUE,title=TRUE,curveAdjacent=TRUE,ask=FALSE)
+semPlot::semPaths(fit1,
+                  what = 'col',
+                  whatLabels = 'std',
+                  style = 'lisrel',
+                  residuals = T,
+                  nCharNodes = 0,
+                  nCharEdges = 0,
+                  groups='manlat',
+                  pastel = T,
+                  ask = F)
 ```

### 2016-06-29 inst/doc/draft.Rmd
v0.1.9015 cosaspect DEMderiv.R #AV

@@
 ---
-title: "Uppangala tree functionnal traits: Draft"
+title: 'Uppangala tree functionnal traits: Draft'
 author: "Sylvain Schmitt"
-date: "`r Sys.Date()`"
+date: '`r Sys.Date()`'
 output: pdf_document

### 2016-06-29 R/draft.R

@@ -207,3 +207,9 @@
 # names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
 # names(SAC) <- traits
 # # kable(as.data.frame(SAC))
+
+# trt <- Qm
+# spl <- TreesNA
+# spl <- tapply(rep(1, length(spl[,1])), list(spl$SpCode, spl$com), sum)
+# spl[is.na(spl)] <- 0
+# phy <- tree

### 2016-06-29 R/topo_deriv.R

@@ -7,6 +7,7 @@ NULL
 #' @param DEM raster. digital elevation model
 #' @param choices char. choices of DEM derivative to compute
 #' @param minslope num. for minslope, see help(rsaga.fill.sinks)
+#' @param dir int. Direction for cosapspect computation in degree
 #' @param workspace char. current workspace
 #' @param path char. path is the folder of the SAGA software
 #' @param method char. for methods see help(rsaga.curvature)
@@ -22,9 +23,9 @@ NULL
 #' DEMderiv(DEM)
 #'
 DEMderiv = function(DEM, choices = c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect',
-                                     'wetness', 'flowdir', 'hillshade'),
-                    minslope = 0.5, workspace = NULL, path = "C:/Program Files/QGIS Essen/apps/saga",
-                     method = "poly2zevenbergen") {
+                                     'cosaspect', 'wetness', 'flowdir', 'hillshade'),
+                    minslope = 0.5, dir = 225, workspace = NULL, method = "poly2zevenbergen",
+                    path = "C:/Program Files/QGIS Essen/apps/saga") {
 
   if(is.null(workspace)){
     dir.create('.tmp')
@@ -75,18 +76,25 @@ DEMderiv = function(DEM, choices = c('slope', 'curvature', 'plancurvature', 'pro
   }
 
   ##########################
-  ##### Aspect
-  if('aspect' %in% choices){
+  ##### Aspect & Cosaspect
+  if('aspect' %in% choices || 'cosaspect' %in% choices){
     rsaga.aspect(in.dem="dem.sgrd", out.aspect="aspect.sgrd",env=myenv,
                  method=method, show.output.on.console = F)
     Aspect <- raster(file.path(workspace, 'aspect.sdat'))
     Aspect <- Aspect / pi * 180
     names(Aspect) <- 'Aspect'
-    Deriv <- stack(Deriv, Aspect)
+    if('aspect' %in% choices){
+      Deriv <- stack(Deriv, Aspect)
+    }
+    if('coaspect' %in% choices){
+      Cosaspect <- cos((Aspect + dir) / 180 * pi)
+      names(Aspect) <- 'Cosaspect'
+      Deriv <- stack(Deriv, Cosaspect)
+    }
   }
 
   ##########################
-  ##### wetness
+  ##### Wetness
   if('wetness' %in% choices){
[…]

### 2016-07-01 inst/doc/PreliminaryResults.Rmd
v0.1.9016 RLQ.Rmd + cosaspect fix

--- 2016-06-27 inst/doc/PreliminaryResults.Rmd
@@
 ```
-
-# *News*
-Add from the `r Sys.Date()` version :
-
-- RLQ to investigate individuals filters
-- Spatial autocorrelation for CWM and models residuals with Moran's I
-- Phylosignal
-- CWV
-- Structural Equation Modelling (SEM)
-- Species growth too much unprecised for analysis
-- Eveness and other FD package metrics
 
@@
 
-Slope was separated in three categories: plane surface (below 20 degrees), mean slope (between 20 and 30degrees) and steep slope (above 30 degrees).
-Plane surface were then discriminated on curvature: concave (positive curvature) and convex (negative curvature).
-Whereas slopes where discriminated by their south-west or north east aspect.
-Finally, 6 topographical classes were defined: concave plane surface (CvP), convex plane surface (CxP), north-east mean slope (NEMS), north-east steep slope (NESS), south-west mean slope (SWMS), and south-west steep slope (SWSS), see figure 3.
-The predominance of north-east slope (62%) in our study site is noticeable and should be took into account.
-
-```{r topo classes}
-col <- c("#008B00", "#00CD00", "#009ACD", "#00BFFF", "#CD9B1D", "#FFC125")
-plot(raster(com, 'topo'), col = col, legend = F)
-contour(Env$DEM, add = T)
-legend('bottomleft', fill = col, legend = levels(com@data$topo), bg = 'white')
-```
-
-Figure 3: Uppangala 10 hectares plot topographical classes.*6 classes were defined one slope, curvature, and south-westerness: concave plane surface (CvP) in dark green, convexe plane surface (CxP) in light green, north-east mean slope (NEMS) in dark blue, north-east steep slope (NESS) in light blue, south-west mean slope (SWMS) in dark orange, and south-west steep slope (SWSS) in ligh orange.*
-
 To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density *(we may add tree height and growth in the future)*.
@@
 
-## Individual filters
-
-Absence of significant results with biotic factors and null results with abiotic factors on communities weighted mean brought us to study environmental filtering at the individual level.
-To do so we realized a fourthcorner (our RLQ) analysis with three matrices: *L*, the $Individuals*Species$ matrix linking the two other matrices, *Q*, the $Trait*Species$ matrix representing our functionnal traits, and *R* the $Environment*Individuals$ matrix representing both biotic and abiotic filtering.
-As a result, fourthcorner analysis brought no significant adjusted p-values between our functionnal traits and biotic and abiotic filters at the individual levels.
-Whereas communities show environmental filtering on tree functionnal traits and strategies, environmental filtering doesn't strongly affect trees on the individual level.
-
-```{r RLQ}
-library(ade4)
-Deriv <- DEMderiv(Env$DEM, c('slope', 'curvature', 'aspect', 'wetness'))
-Deriv$SW <- cos(Deriv$Aspect + 220)
-Deriv <- Deriv[[-3]]
-
-# Individuals
-TreesNA <- Trees[-which(is.na(Trees$X)),]
-TreesNA <- TreesNA[-which(is.na(TreesNA$`2013_girth`)),]
-TreesNA <- TreesNA[-which(is.na(extract(Env$DEM, TreesNA[2:3], df = T)[2])),]
-TreesNA <- TreesNA[-which(TreesNA$SpCode %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
-
-# Matrices
-Lm <- TreesNA[1]
-Lm <- data.frame(tapply(rep(1, length(Lm[,1])), list(row.names(Lm), Lm$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Qm <- PFT_sp[which(names(PFT_sp) %in% c('SLA', 'LDMC', 'WD', 'Thick', 'LA'))]
-row.names(Qm) <- PFT_sp$Sp_Code
-Qm <- Qm[-which(is.na(Qm$WD)),]
-Qm <- Qm[names(Lm),]
-Rm <- cbind(extract(Deriv, TreesNA[2:3], df = T)[-1], TreesNA$`2013_girth`)
-names(Rm)[5] <- 'Girth'
-row.names(Rm) <- row.names(Lm)
-
-# RLQ
-L <- dudi.coa(Lm, scannf = FALSE)
-R <- dudi.pca(Rm, scannf = FALSE, row.w = L$lw)
-Q <- dudi.pca(Qm, scannf = FALSE, row.w = L$cw)
-rlq <- rlq(R, L, Q, scannf = FALSE)
-rd <- randtest(rlq)
-fq <- fourthcorner.rlq(rlq, type = "Q.axes")
-fr <- fourthcorner.rlq(rlq, type = "R.axes")
-fqr <- fourthcorner(Rm, Lm, Qm)
-fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
-
-# Results
-# summary(rlq)
-# plot(rlq)
-# s.arrow(rlq$l1)
-# s.arrow(rlq$c1)
-# s.label(rlq$lQ, boxes = FALSE)
-# plot(fqr.adj, alpha = 0.5, stat = "D2")
-# plot(fqr.adj, x.rlq = rlq, alpha = 0.05, stat = "D2", type = "biplot")
-# plot(fq, alpha = 0.1, stat = "D2")
-# plot(fr, alpha = 0.1, stat = "D2")
-# print(fq, stat = 'D')
-# print(fr, stat = 'D')
-```
-
 ## Phylogenetic signal
@@
 
-# *Future directions*
-
-- Tree estimated height with `BIOMASS`
-
 # References

### 2016-07-01 inst/doc/RLQ.Rmd
v0.1.9016 RLQ.Rmd + cosaspect fix

--- 2016-07-01 inst/doc/PreliminaryResults.Rmd
@@
 ---
-title: 'Uppangala tree functionnal traits: Preliminray Results'
+title: 'Uppangala tree functionnal traits: RLQ and Fourth cornr analysis'
 author: "Sylvain Schmitt"
@@
 
-## Communities
+## RLQ analysis
 
-```{r 5 Com}
-library(FD)
-com <- CWT(Trees, com)
-x <- data.frame(PFT_sp[-1], row.names = PFT_sp$Sp_Code)
-a <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-a[is.na(a)] <- 0
-x <- x[names(a),]
-FD <- as.data.frame(dbFD(x, a, corr = 'cailliez', calc.CWM = F, print.pco = F, messages = F))[-c(1:2,4)]
-names(FD) <- paste0(names(FD), '.fd')
-FD$id <- row.names(FD)
-com$abundance <- merge(com$abundance, FD)
+Following @dray_testing_2008, we conducting both RLQ and fourthcorner analysis to test the effect of environment on our traits.
+Model used to compare our results with a null model in both RLQ Monte-Carlo test and in the fourth corner analysis is the model 6 proposed by @ter_braak_improved_2012.
+
+```{r RLQ}
+library(ade4)
+
+# Matrices
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
+row.names(Qm) <- PFT_sp$Sp_Code
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Qm <- Qm[names(Lm),]
+Rm <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+row.names(Rm) <- com@data$id
+
+# Consistency
+# all(row.names(Lm) == row.names(Rm))
+# all(row.names(Lm) %in% row.names(Rm))
+Lm <- Lm[row.names(Rm),]
+# all(names(Lm) == row.names(Qm))
+Qm <- Qm[names(Lm),]
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Lm <- Lm[row.names(Qm)]
+Lm <- Lm / rowSums(Lm)
+
+# RLQ
+L <- dudi.coa(Lm, scannf = F)
+R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
+Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
+rlq <- rlq(R, L, Q, scannf = F)
+
+# Rersults
+par(mfrow = c(1,2))
+s.arrow(rlq$l1)
+s.arrow(rlq$c1)
+# plot(rlq)
 ```
 
-Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
-For all those community, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES).
-Weighted mean and variances were realized by abundance, presence-absence, or basal area.
-In addition functionnal diversity indices weighted by abundances were computed with the `FD` package [@laliberte_distance-based_2010] for our communities: functional richness (FRic), functional evenness (FEve), and functional divergence (FDiv) and Rao's quadratic entropy (Q) (cite Villéger et al.
-(2008) et Botta-Dukát 2005).
+Figure 3: Environmental variables (left) and traits (right) ordination with RLQ analysis results for L matrix in abundances.
 
-Spatial variograms of community weighted means and variances didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$.
-Functionnal diversity indices show also a plateau around $70 m$.
-
-```{r Variograms}
-library(gstat)
-dist <- variogram(WD.cwm~1, as(com$abundance, 'SpatialPointsDataFrame'))$dist
-par(mfrow = c(2,1))
-for(i in 1:2){
-  traits <- paste0(c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES'), c('.cwm','.cwv')[i])
-  plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', main = c('(a)', '(b)')[i], xlab = 'distance (m)', ylab = 'relative variance (%)')
-  for(j in 1:length(com)){
-    var <- data.frame(lapply(lapply(as.list(traits), function(x){variogram(formula(paste(x, '~ 1')), as(com[[j]], 'SpatialPointsDataFrame'))}), function(x){x[3]}))
-    var <- data.frame(apply(var, 2, function(x){x/max(x)}))
-    names(var) <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
-    colramp <- rainbow(length(var))
-    for(i in 1:length(var)){
-      lines(var[,i] ~ dist, col = colramp[i], lty = j + 1)
-    }
-  }
-legend('bottomright', names(var), fill = colramp, cex = 0.7)
-legend('bottomleft', names(com), lty = c(2:4), cex = 0.7)
-}
+```{r randtest}
+rd <- ade4::randtest(rlq, modeltype = 6)
+plot(rd)
 ```
 
-Figure 4: Trait variograms for all communities.*Variograms were initiated on* $20*20 m^2$ *communities for all traits with both community weighted means* **(a)** *and variances* **(b)** *(by abundance, presence-absence and basal area).
-Some variaograms seem to reach a plateau around* $70 m$.
+Figure 4: RLQ Monte-Carlo test results for model 6 proposed by @ter_braak_improved_2012.
 
-In order to see if environmental biotic factors were influencing functionnal traits or if environmental biotic factors were the results of functionnal traits we used structural equation modelling on community weighted means by abundance.
-Three latent variables were defined: *Abiotic environment* defined with our 4 selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
-Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
-Retained model was the first one (**m1**) with a significantly lower Aikaike's Criterion Indice on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)
+## Fourth corner analysis
 
-```{r SEM}
-library(lavaan)
-m1 <- '
-Abiotic =~ Slope + Curvature + Wetness + SW
-Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
-Biotic =~ BA + Canopy
-Traits ~ Abiotic + Biotic
-'
-m2 <- '
-Abiotic =~ Slope + Curvature + Wetness + SW
-Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
-Biotic =~ BA + Canopy
-Biotic ~ Traits + Abiotic
-'
-fit1 <- sem(m1, data = com$abundance@data)
-fit2 <- sem(m2, data = com$abundance@data)
-semPlot::semPaths(fit1,
-                  what = 'col',
-                  whatLabels = 'std',
-                  style = 'lisrel',
-                  residuals = T,
-                  nCharNodes = 0,
-                  nCharEdges = 0,
-                  groups='manlat',
-                  pastel = T,
-                  ask = F)
+```{r 4thQ AB}
+library(ade4)
+# FQR
+fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
+fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
+
+# Results
+par(mfrow = c(1,2))
+plot(fqr)
+plot(fqr.adj)
+# print(fqr.adj)
 ```
 
-Figure 5: Path diagram for structural equation modelling.
-*Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic.
-Arrow values are indicating standardized parameter estimates*.
+Figure 5: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species abundances in communities with non adjusted p-values (left) and adjusted p-values (right).
 
-Links between both community weighted means and variances with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
-Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables (slope, curvature, wetness and southwesterness for abiotic factors, and canopy height and basal area for biotic one).
-A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix).
-Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance.
-Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 4 and 5 (respectively for community weighted means and variances).
-Same procedure was also applied to functionnal diversity indices (FD) from our ocmmunities, results are presented in table 6.
+```{r 4thQ BIN}
+library(ade4)
 
-An assessment of spatial autocorrelation (SAC) of the community weighted means and variances and functionnal diversity indices was conducted using Moran's I statistic.
-Only half of community weighted means and variances were significantly positively spatially autocorrelated (see table 3) with low values (Moran's I from 0.01 to 0.04).
-Leaf economic spectrum and dry matter content weighted means as wood density weighted variances are slightly spatially correlated suggesting a spatial filter due to environment or local history pressure (as dispersion limitation).
-On the opposite, wood density and thickness community weighted means as wood economic spectrum community weighted variances possess a stochastic spatial distribution implying that they're not shaped by factors with a spatial organization.
-Finally remaining traits community weighted means and variances are spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter.
-As a conclusion both community weighted means and variances are most of the time stochastic or driven by dispersion.
-Moreover all functionnal diversity indices (FD) are spatially correlated.
+# Matrices
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
 
-Table 3: Spatial autocorrelation measured by Moran's I for community weighted means and variances.
-_Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations.
-FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+# Consistency
+Lm <- Lm[row.names(Rm),]
+Lm <- Lm[row.names(Qm)]
+Lm[Lm > 0] <-  1 #Presence-absence
+Lm <- Lm / rowSums(Lm)
 
-```{r SAC}
-library(ape)
-traits <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
-traits <- c(paste0(traits, '.cwm'), paste0(traits, '.cwv'))
-traits <- list(c(traits, 'FRic.fd', 'FEve.fd', 'FDiv.fd', 'FDis.fd', 'RaoQ.fd'), traits, traits)
-weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
-diag(weights) <- 0
-SAC <- mapply(function(x,y){lapply(as.list(x@data[y]), Moran.I, weights)}, x =com, y = traits)
-SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
-SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
-SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
-SAC <- data.frame(SAC)
-names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
-SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
-SAC$Type <- ''
-SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
-SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
-kable(SAC, row.names = F)
+# RLQ
+L <- dudi.coa(Lm, scannf = F)
+R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
+Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
+rlq <- rlq(R, L, Q, scannf = F)
+fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
+fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
+
+# Results
+par(mfrow = c(1,2))
+plot(fqr)
+plot(fqr.adj)
+# print(fqr.adj)
 ```
 
-```{r 6 lm & null models}
-reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
-# LM to simplify with loop or lapply
-LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
-                                                 logSLA.cwm = lm(formula(paste('log(SLA.cwm)', reg))), 
-                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
-                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
-                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
-                                                 logLA.cwm = lm(formula(paste('log(LA.cwm)', reg))),
-                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
-                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
-                                                 logSLA.cwv = lm(formula(paste('log(SLA.cwv)', reg))), 
-                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
-                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
-                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
-                                                 logLA.cwv = lm(formula(paste('log(LA.cwv)', reg))),
-                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
-LM$abundance <- c(LM$abundance, with(com$abundance@data, list(FRic.fd = lm(formula(paste('FRic.fd', reg))),
-                                                              FEve.fd = lm(formula(paste('FEve.fd', reg))),
-                                                              FDiv.fd = lm(formula(paste('FDiv.fd', reg))),
-                                                              FDis.fd = lm(formula(paste('FDis.fd', reg))),
-                                                              RaoQ.fd = lm(formula(paste('RaoQ.fd', reg))))))
-LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
-model <- null_model(com, Trees, PFT_sp, lapply(LM, function(y){lapply(y, function(x){x$call})}), SpCode, n = 999, verbose = F, time = T)
+Figure 6: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species presence absence in communities with non adjusted p-values (left) and adjusted p-values (right).
+
+```{r 4thQ BA}
+library(ade4)
+
+# Matrices
+Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum)) #BA
+Lm[is.na(Lm)] <- 0
+
+# Consistency
+Lm <- Lm[row.names(Rm),]
+Qm <- Qm[names(Lm),]
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Lm <- Lm[row.names(Qm)]
+Lm <- Lm / rowSums(Lm)
+
+# RLQ
+L <- dudi.coa(Lm, scannf = F)
+R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
+Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
+rlq <- rlq(R, L, Q, scannf = F)
+fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
+fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
+
+# Results
+par(mfrow = c(1,2))
+plot(fqr)
+plot(fqr.adj)
+# print(fqr.adj)
 ```
 
-## Community weighted means
-
-Table 4: Summary for linear models and null models for traits community weighted means.
-_Linear model results for each trait community weighted means (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model deviance was compared to 999 null model deviances in which traits were randomized among species.
-See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-```{r CWM}
-kable(regression_table(LM, model, reg, select = '.cwm'))
-```
-
-### Abiotic factors
-
-First we can notice that, whereas the leaf economic spectrum shows significantly non-random links with abiotic factors, most of weighted means associated to the wood economic spectrum on our principal component analysis (see figure 1) didn't differ from null models.
-A spatial structure due to local individual history and dispersion is responsible for the observed links between abiotic factors and weighted mean associated to the wood economic spectrum.
-Outside from the leaf economic spectrum (LES), topography doesn't localy afect Uppangala plot trees as an environmental filter.
-
-Leaf economic spectrum (LES) and specific leaf area (SLA) is associated non-randomly with convex and dry surfaces, as the thickness for basal area community weighted mean.
-Trees will tend to increase their surface compare to their carbon invest ment in areas where the water will flow but not accumulate nor stay, and where the soil will tend to accumulate and be stabilized.
-In favourable water and soil conditions, trees will tend to invest more in leaf ressource capture at the expense of leaf defense, adopting the acquisitive [@sterck_functional_2011] our fast growing strategy [@reich_world-wide_2014].
-
-Wood density (WD) in presence absence shows a non-random association with southwest steep slope and concave surfaces with a relatively important explanation level ($R^2=0.25$).
-Trees will increase their wood density in thin soil and water and soil departure conditions under potential monsoon winds *(We need to decide or not if regarding the plot condition we can accept or not the southwest orientation as being linked to monsoon winds)*.
-In water stress and mechanical hazard conditions, trees will compensate by increasing their water transport safety and mechanical resistance and resilience with an increased wood density [@chave_towards_2009].
-
-### Biotic factors
-
-Most of community weighted means don't show association with biotic factors, and most of the rare association are not significantly different from null models randomly generated.
-Still leaf area (LA) in basal area is associated with high trees.
-Voluminous individuals tend to increase their leaf surface in heigher canopy condition to compensate understory shadow.
-And wood density tends also to increase with tree height for rare species, suggesting that with higher canopy denser wood species will tend to grow in a community, as shade species.
-
-## Community weighted variances
-
-Table 5: Summary for linear models and null models for traits community weighted variances.
-_Linear model results for each trait community weighted variances (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model deviance was compared to 999 null model deviances in which traits were randomized among species.
-See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-```{r CWV}
-kable(regression_table(LM, model, reg, select = '.cwv'))
-```
-
-Most of traits and tree strategy axes didn't show linear models for community weighted variances (CWV) significantyl differently affected by both abiotic and biotic environment compare to null models.
-Only the leaf economic spectrum (LES) and wood density (WD) showed linear models significantly different from null models for their community weighted variances (CWV) for all wieghts (abundance, presence-absence, and basal area).
-Leaf area community weighted variance by basal area showed also a significantly positive linear model with slope compare to null models.
-
-Leaf economic spectra (LES) converge with better soil and water conditions in convex situations, and for abundant species oriented under monsoon wind LES diverge suggesting different adaptive strategies from species.
-Whereas wood density from abundant and rare species also diverged with the southwest orientation to monsoon winds, wood density from voluminous species converge with slope and diverge with the canopy height.
-Effectively under high canopy trees will be under a lot of shade situation, implying differing growth strategies and further more different wood densities.
-Leaf area also converge for voluminous species with slope.
-
-## Functionnal diversity
-
-Whereas functionnal richness (FRic) standardized by the *'global FRic'* that include all species is low and relatively stable (mean = 0.05, sd = 0.04);
-functionnal eveness is high and relatively stable (mean = `r round(mean(com$abundance@data$FEve.fd), 2)`, sd = `r round(sd(com$abundance@data$FEve.fd), 2)`) in our communities as the functionnal diversity (mean = `r round(mean(com$abundance@data$FDiv.fd), 2)`, sd = `r round(sd(com$abundance@data$FDiv.fd), 2)`).
-Our communties are relatively poor in functionnal values but with an even and diverse trait distribution.
-
-Table 6: Summary for linear models and null models for functionnal diversity indices (FD).
-_Linear model results for each community functionnal indices weighted by abundances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model deviance was compared to 999 null model deviances in which traits were randomized among species.
-FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-```{r FD}
-kable(regression_table(LM, model, reg, select = '.fd')[-2], row.names = F)
-```
-
-At the exception of the functionnal richness (FRic), all linear models for functionnal diversity indices didn't differ from null models suggesting that our bioctic and abitotic factors are not responsible for their distribution in our communities.
-Still functionnal richness increases with slope, suggesting as commmunity weighted variances (CWV) that our traits diverge in different adaptive strategies in steeper slopes.
-
-## Models residuals
-
-Table 6: Spatial autocorrelation measured by Moran's I for linear regression model residuals from community weighted means (CWM) and variances (CWV) and functionnal diversity indices (FD).
-_Spatial autocorrelation for residuals from each trait community weighted means and variances (by abundance, presence-absence and basal area) linear models and for functionnal diversity (FD) linear models, see table 1 for abbreviations.
-FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-```{r SAC residuals}
-library(ape)
-weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
-diag(weights) <- 0
-SAC <- lapply(LM, function(x){lapply(x, function(y){Moran.I(residuals(y), weights)})})
-SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
-SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
-SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
-SAC <- data.frame(SAC)
-names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
-SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
-SAC$Type <- ''
-SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
-SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
-kable(SAC, row.names = F)
-```
-
-Model residuals spatial autocorrelation (SAC)  for our community weighted means (CWM) is null confirming that our models explain fully the spatial structure of our 5 traits and 2 strategy axes (compare to table 3).
-At the exception of thickness that still shows a significant spatial autocorrelation, suggesting that our model missed spatially structured factors in this case.
-For models residuals from community weighted variances, specific leaf area, wood economic spectrum and leaf dry matter content still show a spatial structure.
-This spatial structure in local trait variances linked to alpha diversity may be explained by local history and stochasticty associated with a null distribution of individuals not directly shaped by their environment.
-Finally we find back this spatial structure unexplained by model for functionnal diversity indices.
+Figure 7: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species basal area in communities with non adjusted p-values (left) and adjusted p-values (right).
 
@@
 
-```{r phylosignal, include=FALSE}
+```{r phylosignal}
 library(phylosignal)
@@
 p4d <- phylo4d(tree, data)
-barplot.phylo4d(p4d, tree.type = "phylo", tree.ladderize = TRUE)
+# barplot.phylo4d(p4d, tree.type = "phylo", tree.ladderize = TRUE)
 # phylosim <- phyloSim(tree = tre, method = "all", nsim = 100, reps = 99)
@@
 phyCor <- lapply(as.list(names(data)), function(x){phyloCorrelogram(p4d, trait = as.character(x))})
-# par(mfrow = c(2,4))
-# for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
+par(mfrow = c(2,4))
+for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
 lipa <- lipaMoran(p4d)
@@
 
+Once the phylogenetic structure revealed we wanted to test it in our community with the Bst and PIst metric proposed by @hardy_characterizing_2007 with the null model 1s suggested by [cite] to see if we were under phylogenetic clustering or overdispersion.
+
+```{r phylo structure}
+library(spacodiR)
+
+# Phylogeny renaming
+tree$tip.label <- gsub('_', ' ', tree$tip.label, fixed = T)
+tree$tip.label <- as.character(SpCode[tree$tip.label])
+
+# Matrices
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Rm <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+row.names(Rm) <- com@data$id
+
+# Consistency
+# all(row.names(Lm) == row.names(Rm))
+# all(row.names(Lm) %in% row.names(Rm))
+Lm <- Lm[row.names(Rm),]
+# all(names(Lm) == tree$tip.label)
+Lm <- Lm[tree$tip.label]
+
+# Phylogeny structure
+spafun <- spacodi.calc(sp.plot = t(Lm), sp.traits = Qm)
+spaphy <- spacodi.calc(sp.plot = t(Lm), phy = tree)
+permut <- lapply(as.list(names(spaphy[1:4])), function(x){spacodi.by.nodes(t(Lm), tree, sp.parm = x, n.rep = 100, method = "1s")})
+names(permut) <- names(spaphy[1:4])
+par(mfrow=c(2,1))
+spacodi.permutplot(permut$Bst)
+spacodi.permutplot(permut$PIst)
+```
+
+The phylogenetic structure among our communities was then evaluated by an RLQ analysis with cophenetic matrix of our phylogenetic tree distances as the Q matrix reduced by a principal coordinate analysis.
+
+```{r phylo RLQ}
+library(ade4)
+library(adephylo)
+
+# RLQ
+L <- dudi.coa(Lm, scannf = F)
+R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
+Qd <- distTips(tree)
+Q <- dudi.pco(Qd, scannf = F, row.w = L$cw)
+rlq <- rlq(R, L, Q, scannf = F)
+
+# Random RLQs
+tree$node.label <- NULL
+ntrees <- rep(list(tree), 999)
+ntrees <- lapply(ntrees, resamp.phy)
+nQd <- lapply(ntrees, distTips)
+nrlq <- lapply(ntrees, function(x){
+  Lm <- Lm[x$tip.label]
+  L <- dudi.coa(Lm, scannf = F)
+  R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
+  Q <- dudi.pco(Qd, scannf = F, row.w = L$cw)
+  rlq(R, L, Q, scannf = F)})
+
+# Randtest
+pval <- rank(c(sum(rlq$eig), unlist(lapply(nrlq, function(x){sum(x$eig)}))))[1]/1000
+hist(unlist(lapply(nrlq, function(x){sum(x$eig)})), main = paste('pval =', round(pval, 3)))
+lines(c(sum(rlq$eig), sum(rlq$eig)), c(0,500), col = 'red', lwd = 2)
+```
+
+Figure 8: RLQ test by model nul results with phylogenetic tree distances as the Q matrix for model 1s proposed by [cite].
+
 # References

### 2016-07-01 inst/doc/draft.Rmd
v0.1.9016 RLQ.Rmd + cosaspect fix

@@
 ---
-title: 'Uppangala tree functionnal traits: Draft'
-author: "Sylvain Schmitt"
-date: '`r Sys.Date()`'
+title: 'Uppangala tree functionnal: Article draft'
+author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
 output: pdf_document
@@
 The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village.
-The permanent plot was established in 1990 by the French Institute of Pondicherry and described by Pascal and Pelissier [@pascal_structure_1996].
+The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996.
 The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergree forest type [@pascal_wet_1988].
@@
 Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.
+Leaves and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
+For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).
 
-- Unsampled species beanch wood 
-- Leaves manipulation conditions.
-
-Among leaves, 5 were weighted fresh, scanned [cite scan], and their thickness were measured with a microgauge [cite].
-Those leaves were kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying processed was stopped after dry leaf weights were stable in time.
+Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite].
+Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time.
 Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias.
@@
 Samples were after dried for 48 hours at 80 degrees celsius and weighted.
+
+- Removed petiole, petiolule and rachis
 

### 2016-07-01 R/draft.R

@@ -208,8 +208,21 @@
 # names(SAC) <- traits
 # # kable(as.data.frame(SAC))
 
+# library(spacodiR)
 # trt <- Qm
 # spl <- TreesNA
 # spl <- tapply(rep(1, length(spl[,1])), list(spl$SpCode, spl$com), sum)
 # spl[is.na(spl)] <- 0
 # phy <- tree
+# phy$tip.label <- gsub('_', ' ', phy$tip.label, fixed = T)
+# phy$tip.label <- as.character(SpCode[phy$tip.label])
+# spafun <- spacodi.calc(sp.plot = spl, sp.traits = trt)
+# spaphy <- spacodi.calc(sp.plot = spl, phy = phy)
+# permut <- lapply(as.list(names(spaphy[1:4])), function(x){spacodi.by.nodes(spl, phy, sp.parm = x, n.rep = 100, method = "1s")})
+# names(permut) <- names(spaphy[1:4])
+# par(mfrow=c(2,1))
+# spacodi.permutplot(permut$Bst)
+# spacodi.permutplot(permut$PIst)
+
+# plot(dist(coordinates(com)), betapart::beta.pair(Lm)$beta.sor, xlab = 'distance (m)', ylab = 'fraction species shared', main = 'Sørensen similarity index between pairs of 20*20 m2 communities')
+

### 2016-07-01 R/topo_deriv.R

@@ -86,9 +86,9 @@ DEMderiv = function(DEM, choices = c('slope', 'curvature', 'plancurvature', 'pro
     if('aspect' %in% choices){
       Deriv <- stack(Deriv, Aspect)
     }
-    if('coaspect' %in% choices){
+    if('cosaspect' %in% choices){
       Cosaspect <- cos((Aspect + dir) / 180 * pi)
-      names(Aspect) <- 'Cosaspect'
+      names(Cosaspect) <- 'Cosaspect'
       Deriv <- stack(Deriv, Cosaspect)
     }
   }

### 2016-07-08 inst/doc/RLQ.Rmd
v0.1.9017 RLQ.Rmd + plan

@@
 ---
-title: 'Uppangala tree functionnal traits: RLQ and Fourth cornr analysis'
-author: "Sylvain Schmitt"
-date: '`r Sys.Date()`'
+title: 'Uppangala tree functionnal: RLQ draft'
+author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
 output: pdf_document
@@
 library(knitr)
+library(pander)
 opts_chunk$set(
@@
 
-# Traits
+# Abstract
 
-## Measured traits
+Keywords: functionnal trait, tropical evergreen forest, Western Ghats, India...
 
-Table 1: Functional traits measured. *Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development.*
+# Introduction
+
+# Material and methods
+
+## Study site
+
+The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village.
+The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996.
+The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988].
+Climate and geology of the 10-ha plot can be considered as homogeneous.
+General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
+Each tree girth is monitored by a two-year census.
+
+## Sampling
+
+We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates.
+Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals.
+Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot.
+For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013.
+Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy.
+Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.
+Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
+For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).
+
+Leaf petioles, rachis and petiolules were removed for all measurements.
+Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite].
+Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time.
+Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias.
+Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues.
+Branch wood samples were sink in water to evaluate their volume by water displacement.
+Samples were then dried for 48 hours at 80 degrees celsius and weighted.
+
+## Traits
+
+With leaf and branch wood samples we calculated five functionnal traits: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the wood sample volume in $cm^3$ divided by the wood sample dry weight in $g$ (see table 1).
+Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
+For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+
+Table 1: Functional traits measured.
+*Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development from @baraloto_decoupled_2010.*
 
@@
 
-Uppangala measured functional traits vary in the ranges of tropical species values that we can find in literature.
-We can notice in table 1 that wood density has a relatively high mean as we could expect in an old untouched forest.
-Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010], and can be as well explained by environment condition.
-For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
-We could also have added a sixth trait, called leaf tissue density (LTD) with the formula: $LTD=Leaf dry weight/Thickness*Leaf area$.
-LTD is highly correlated to LDMC [@hodgson_is_2011], but can still bring extra information **To be discussed**.
+## Environmental variables
 
-## Traits distribution and variation
+To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
+Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
+Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
+Distance to the nearest potential river was caluclated under ArcGis (cite).
+
+To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density.
+Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$.
+Both basal area and stem density are calculated with 2013 girth inventory from Uppangala plot on stem with a diameter above 10 centimeters at breast height ($DBH>10$).
+Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$).
+So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
+
+## Analyses
+
+### Covariation in traits
+
+Traits variation was first investigated in three levels with the coefficient of variation ($cv=sd/mean$) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
+For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
+To test the correlation and covariation among traits we performed a principal component analysis (PCA).
+We investigated peculiar traits covariation highlighted with the PCA with a Spearman's rank correlation test.
+
+### Environment definition
+
+In order to avoid multi-colinearity in future analysis, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
+In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling.
+Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
+Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
+Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).
+
+### Community analysis
+
+Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
+For all communities, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes.
+We used the following formula to compute each community weighted parameter (CWP):
+$$CWP = \sum_{k=1}^{nj} W_{k,j} * P_{k,j} ~ (1)$$ 
+where $nj$ is the number of species sampled in community $j$, $W_{k,j}$ is the weight of species $k$ in community $j$ used to compute CWP (relative abundance, presence-absence, or relative basal area), and $P_{k,j}$ is the parameter of species $k$ in community $j$ (mean or the variance).
+In addition 5 functionnal diversity indices (FD) weighted by abundances were computed: functional richness (FRic), functional evenness (FEve), functional divergence (FDiv) and Rao's quadratic entropy (Q) (@villeger_new_2008, @botta-dukat_raos_2005).
+Spatial variograms were used to investigate our CWM and CWV spatial organization.
+CWM, CWV, and FD spatial autocorrelation (SAC) was also assessed by Moran's I statistic.
+
+Links between both CWM and CWV with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
+Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables.
+A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix).
+Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance.
+Same procedure was also applied to functionnal diversity indices (FD) from our communities.
+Finally spatial autocorrelation (SAC) of residuals was investigated to assess selected models.
+
+### Fourth-corner analysis
+
+In addition of our community analysis by regression, we conducted both RLQ and fourth-corner analyses to test the effect of environments on Uppangala tree functionnal traits following @dray_testing_2008.
+Both matrices R (`Environment*Sites`) with selected features and Q (`Traits*Species`) were reduced by a principal component analysis.
+We used a Monte-Carlo test to compare fourth-corner and RLQ results to 999 nul models built with the model *m6* proposed by @ter_braak_improved_2012.
+Fourth corner analysis was tested with 3 types of L matrices ($Sites*Species$) with species relative abundance, presence absence or relative basal area in each sites; whereas RLQ analysis was only conducted with relative abundance.
+
+### Phylogenetic signal
+
+Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
+Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 functionnal traits, the two PCA axes and in the phylogenetic tree.
+Phylogenetic signal structure was assessed by a tree tip distances correlogram.
+In addition, $\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate phylogenetic signal spatial organization with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
+Finally phylogenetic association to habitat was also studied by an RLQ analysis with tree tip distances as the Q matrix reduced by a principal coordinate analysis.
+RLQ result with phylogenetic distance matrix was compared to 999 nul models were phylogenetic tree tips were randomized.
+
+All analysis were conducted in the R 3.3.0  free software environment for statistical computing and graphics [@r_core_team_r:_2016] using the pacakge 
+`ade4` [@dray_ade4_2007], 
+`lavaan` [@rosseel_lavaan:_2012], 
+`FD` [@laliberte_distance-based_2010], 
+`gstat` [@pebesma_multivariable_2004], 
+`ape` [@paradis_ape:_2004], 
+`phylosignal` [@keck_phylosignal:_2015] 
+and `spacodiR` [@eastman_spacodir:_2013].
+
+# Results
+
+## Are functionnal composition of communities linked to the environment ?
 
@@
 
-Some outliers are still to be corrected in individual leaf trait value, but thickness outliers and extreme values are due to too close secondary veins.
-Table 2 shows a intra-individual variation lower than intra-specific and inter-specific variation, indicating that leaves are even in morphology inside an individual but can start to show some variation between trees of the same species.
-More specifically, whereas traits like thickness and wood density seems to stay stable in intra-individual and intra-specific, leaf area shows a high coefficient of variation for both intra-individual and intra-specific values.
+Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
+Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$).
 
-## Covariations in traits
+The multivariate analysis of trait correlation shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
+The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood.
+Whereas the second PCA axis (PCA2) covary mainly with SLA representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
+PCA first axis might be associated with the WD covariation to the wood economic spectrum [WES, @chave_towards_2009].
+Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
+Our results integrate less traits (5 against 14), and less species (92 against 758), but still capture the same strategy axes.
 
-```{r 3 PCA PFT}
+Figure 1 shows WD orthogonal to SLA following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$).
+LA is almost opposed to WD and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007.
+Finally, we found the positive relation suggested by @diaz_global_2015 between LDMC and WD (Spearman's $rho = 0.24$, $p-value < 0.001$).
+
+```{r 3 RLQ}
 library(vegan)
-pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_ind, cor = T)
-lambda <- pca$sdev * sqrt(nrow(pca$scores))
-pca.var = apply(pca$score, 2, var)
-PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
-names(PCAscores) <- c('WES', 'LES')
-PCAscores$ID <- row.names(PCAscores)
-plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
-     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
-     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
-     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
-# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
-par(new=T)
-Rot <- t(t(pca$loadings)*lambda)
-XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
-XLIM <- XLIM+(XLIM*0.2)
-plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
-lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
-lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
-legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
-       text.col = c("green", "firebrick"))
+library(ade4)
+library(raster)
 
-arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
-text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
-Hmisc::subplot(function(){
-  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
-          cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(pca.var)), col = 'red') ;
-  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
-         box.col = 'white')},
-  x=grconvertX(c(0.75,1), from='npc'),
-  y=grconvertY(c(0.05,0.30), from='npc')
-)
-axis (3)
-axis (4)
-PFT_ind$ID <- row.names(PFT_ind)
-PFT_ind <- merge(PFT_ind, PCAscores, all = T)
-row.names(PFT_ind) <- PFT_ind$ID
-PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
-```
-
-Figure 1: Principal component analysis (PCA) performed on individual trait means.
-*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
-See table 1 for abbreviation.*
-
-We realized a principal component analysis on individual trait means (see figure 1), where first axis (PCA1) covary mainly with leaf thickness, dry matter content and wood density.
-Whereas second axis (PCA2) covary mainly with specific leaf area representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
-PCA1 might be associated with the wood density covariation to the wood economic spectrum [WES, @chave_towards_2009].
-Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
-Our results integrate less traits (5/14), and less species (92/758), but still capture the same strategy axes.
-
-We can see that as wood density and specific leaf area are orthogonal following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$).
-Leaf area is almost opposed to wood density and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007.
-This relationship could be explained by an equilibrium in tree hydraulics (Coomes et al, 2008).
-Finally, we found the positive relation suggested by @diaz_global_2015 between leaf dry matter content and wood density (Spearman's $rho = 0.24$, $p-value < 0.001$).
-
-# Environmental filters
-
-## Environment and trees
-
-We collected trait information on 92 over 117 (79%) species focusing in priority on abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates (*Warning : one of our 92 species is not referenced yet among currently known Uppangala species, Agrosystachis indica, one of 117 species is not removed yet, Toona ciliata.*).
-Thus we generated data for 98.8% of the tree individuals present in Uppangala plot.
-
-```{r 4 env-com}
-library(raster)
-Env <- genEnv()
-Env$Rivers <- readAll(Env$Rivers)
+# Data
 Trees <- genTrees()
 Species <- genSpecies()
-PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
-row.names(PFT_sp) <- PFT_sp$Group.1
-PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
-PFT_sp <- PFT_sp[-c(1:2,4)]
+Env <- genEnv()
+com <- quadrats()
+com@data$id <- paste0('C', com@data$id)
+
+# L matrix
 Trees$SpCode <- tolower(Trees$SpCode)
 Trees$ID <- row.names(Trees)
-Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
 row.names(Trees) <- Trees$ID
 Trees <- Trees[-which(names(Trees) == 'ID')]
-com <- quadrats()
 Trees$com <- NA
@@
 Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
-Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
-com@data$id <- paste0('C', com@data$id)
-com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+
+# R matrix
+com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'cosaspect'))
 com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
+names(com@data)[5] <- 'SW'
 com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$SW <- sin((com@data$Aspect + 225) / 180 * pi)
-com@data$NE <- cos((com@data$Aspect + 225) / 180 * pi)
-com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
 com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
 com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
-com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
-# com@data$Height <- aggregate(Trees$height, list(Trees$com), mean, na.rm = T)[,2]
-rm(Deriv, XY)
-```
+Rm <- com@data[-1]
+row.names(Rm) <- com@data$id
 
-To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
-Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
-Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect in degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
-Distance to the nearest potential river was caluclated under ArcGis (cite).
-
-In order to avoid multi-colinearity in future regressions, we reduced our explanatory variables via a principal component analysis, see figure 2.
-Curvature, profile curvature, plan curvature, and distance to potential rivers covary together and with the principal component analysis first axis (PCA1).
-Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
-As a conclusion, we choose to reduce our variable to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness (SW) to represent monsoon winds [@bunyan_effect_2015] and slope.
-
-```{r PCA Env}
-pca.env <- princomp(~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
-lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
-pca.env.var = apply(pca.env$score, 2, var)
-plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
-     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
-     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
-     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
-Hmisc::subplot(function(){
-  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
-          cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(pca.env.var)), col = 'red') ;
-  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
-         box.col = 'white')},
-  x=grconvertX(c(0.05,0.35), from='npc'),
-  y=grconvertY(c(0.025,0.30), from='npc')
-)
-par(new=T)
-Rot <- t(t(pca.env$loadings)*lambda)
-XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
-XLIM <- XLIM+(XLIM*0.2)
-plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
-lines(XLIM, c(0,0), lty = 2, lwd = 2)
-lines(c(0,0), XLIM, lty = 2, lwd = 2)
-col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
-arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
-        Rot[,1],Rot[,2],lwd=2,col=col)
-legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
-       text.col = col, cex = 1.2)
-text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 1), rownames(Rot),cex=1.2,col=col)
-axis (3)
-axis (4)
-```
-
-Figure 2: Principal component analysis (PCA) performed on topgraphical variables.
-*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
-SW stand for southwesterness.*
-
-Among all topographical variables, slope, curvature and south-westerness were chosen to define 6 topographical classes.
-We rejected elevation because our plot is situated in a mountain slope and doesn't include valley.
-Consequently elevation mainly inform on the position in the mountain general slope.
-And because the elevation range between the top of the plot and the bottom is low (325-490), we assumed it won't affect directly tree ecology.
-
-To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density *(we may add tree height and growth in the future)*.
-Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$.
-Both basal area and stem density are calculated with 2013 Uppangala plots girth inventory only on stem with a diameter above 10 centimeters at breast height ($DBH>10$).
-Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$).
-So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
-
-For each species the potential strata was defined.
-Because species were defined in several strata, we kept the lowest strata as the species strata.
-We tested the strata effect on each trait with an anova combined to a tukey test, but no trait show a significant variation (all F values weren't significanlty different from 0, $p-values>0.1$).
-
-```{r strata effect}
-AOV <- list(SLA = aov(PFT_sp$SLA ~ Species[PFT_sp$Sp_Code,]$Strata), 
-            WD = aov(PFT_sp$WD ~ Species[PFT_sp$Sp_Code,]$Strata),
-            LDMC = aov(PFT_sp$LDMC ~ Species[PFT_sp$Sp_Code,]$Strata),
-            LA = aov(PFT_sp$LA ~ Species[PFT_sp$Sp_Code,]$Strata),
-            Thick = aov(PFT_sp$Thick ~ Species[PFT_sp$Sp_Code,]$Strata))
-# lapply(AOV, summary)
-# lapply(AOV, TukeyHSD)
-```
-
-## RLQ analysis
-
-Following @dray_testing_2008, we conducting both RLQ and fourthcorner analysis to test the effect of environment on our traits.
-Model used to compare our results with a null model in both RLQ Monte-Carlo test and in the fourth corner analysis is the model 6 proposed by @ter_braak_improved_2012.
-
-```{r RLQ}
-library(ade4)
-
-# Matrices
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
+# Q matrix
+PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
+row.names(PFT_sp) <- PFT_sp$Group.1
+PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
+PFT_sp <- PFT_sp[-c(1:2,4)]
+PFT_sp$SLA <- log(PFT_sp$SLA)
+PFT_sp$LA <- log(PFT_sp$LA)
 Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
 row.names(Qm) <- PFT_sp$Sp_Code
+
+# Matrices adjustments
+Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
 Qm <- Qm[-which(is.na(Qm$WD)),]
-Qm <- Qm[names(Lm),]
-Rm <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
-row.names(Rm) <- com@data$id
-
-# Consistency
-# all(row.names(Lm) == row.names(Rm))
-# all(row.names(Lm) %in% row.names(Rm))
 Lm <- Lm[row.names(Rm),]
-# all(names(Lm) == row.names(Qm))
-Qm <- Qm[names(Lm),]
-Qm <- Qm[-which(is.na(Qm$WD)),]
-Lm <- Lm[row.names(Qm)]
+Lm <- Lm[intersect(row.names(Qm), names(Lm))]
 Lm <- Lm / rowSums(Lm)
@@
 
-# Rersults
-par(mfrow = c(1,2))
-s.arrow(rlq$l1)
-s.arrow(rlq$c1)
-# plot(rlq)
+# Q plot
+plot(t(t(Q$li)/Q$eig[1:2]),pch=16, col = 'lightgrey',
+     xlab = paste('Axe 1 -', round(Q$eig[1]/sum(Q$eig)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(Q$eig[2]/sum(Q$eig)*100, 2), '%'))
+par(new=T)
+XLIM <- c(-max(abs(Q$co[,1])),max(abs(Q$co[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Q$co,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
+lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
+# legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
+       # text.col = c("green", "firebrick"))
+Hmisc::subplot(function(){
+  b <- barplot((Q$eig / sum(Q$eig)), las = 2, ylim = c(0,range(bstick(7))[2]),
+          cex.names = 0.7, axes = T, names.arg = NA) ;
+  lines(b, bstick(length(Q$eig)), col = 'red') ;
+  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
+         box.col = 'white')},
+  x=grconvertX(c(0.65,0.95), from='npc'),
+  y=grconvertY(c(0.01,0.26), from='npc')
+)
+axis (3)
+axis (4)
+# arrows (rep(0,nrow(R$co)),rep(0,nrow(R$co)),R$co[,1],R$co[,2],lwd=1, col = 'blue')
+# text(R$co, pos = c(4, 2, 4, 2, 3), rownames(R$co),cex=0.9,col='blue')
+arrows (rep(0,nrow(Q$co)),rep(0,nrow(Q$co)),Q$co[,1],Q$co[,2],lwd=2)
+text (Q$co, pos = c(2, 2, 2, 4, 4), rownames(Q$co),cex=1.2)
 ```
 
-Figure 3: Environmental variables (left) and traits (right) ordination with RLQ analysis results for L matrix in abundances.
+Figure 1: Principal component analysis (PCA) performed on individual trait means.
+*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
+See table 1 for abbreviation.*
 
-```{r randtest}
-rd <- ade4::randtest(rlq, modeltype = 6)
-plot(rd)
+Multi-colinearity was avoided by reducing our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see figure 2).
+Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
+Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
+Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary.
+Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)
+
+Whereas total co-inertia from RLQ analysis results was significanlty different for the model *m2*, it didn't differ from null model *m4* in model *m6* proposed by @ter_braak_improved_2012.
+So because both test should be rejected to reject model *m6* hypothesis, therefore the null hypothesis *H0* is not rejected.
+
+Still fourth-corner analysis with adjusted p-values or not show significant trait-environment association at a level $\alpha = 0.05$ (see table 7).
+WD is negatively associated with southwesterness for all L matrices and positively associated with canopy height in presence absence.
+LDMC and LA are respectively negatively and positively associated with canopy height in relative abundance.
+
+Table 7: Summary of fourth-corner analysis.
+*Result of Monte-Carlo test between fourth-corner result and 999 nul models following the model m6 suggested by @ter_braak_improved_2012.* _L matrices are filled with relative abundance (AB), presence-absence (PA) or relative basal area (BA)).
+The 5 functionnal traits are tested with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+See table 1 for trait abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r FQR}
+library(ade4)
+rd <- randtest(rlq, modeltype = 6)
+
+# L matrices list
+Lm.l <- list(abundance = Lm)
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Lm <- Lm[row.names(Rm),]
+Lm <- Lm[intersect(row.names(Qm), names(Lm))]
+Lm[Lm > 0] <-  1
+Lm <- Lm / rowSums(Lm)
+Lm.l$`presence-absence` = Lm
+Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Lm <- Lm[row.names(Rm),]
+Lm <- Lm[intersect(row.names(Qm), names(Lm))]
+Lm <- Lm / rowSums(Lm)
+Lm.l$`basal area` = Lm
+
+# Full # -WD
+RLQm.l <- list(full = list(Rm.l = rep(list(Rm),3), Lm.l = Lm.l, Qm.l = rep(list(Qm),3)))
+Qm <- PFT_sp[c('SLA', 'LDMC', 'Thick', 'LA')]
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+row.names(Qm) <- PFT_sp$Sp_Code
+Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
+Lm <- Lm[row.names(Rm),]
+Lm <- Lm[intersect(row.names(Qm), names(Lm))]
+Lm <- Lm / rowSums(Lm)
+Lm.l <- list(abundance = Lm)
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Lm <- Lm[row.names(Rm),]
+Lm <- Lm[intersect(row.names(Qm), names(Lm))]
+Lm[Lm > 0] <-  1
+Lm <- Lm / rowSums(Lm)
+Lm.l$`presence-absence` = Lm
+Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Lm <- Lm[row.names(Rm),]
+Lm <- Lm[intersect(row.names(Qm), names(Lm))]
+Lm <- Lm / rowSums(Lm)
+Lm.l$`basal area` = Lm
+RLQm.l$`-WD` <- list(Rm.l = rep(list(Rm),3), Lm.l = Lm.l, Qm.l = rep(list(Qm),3))
+# str(RLQm.l,3)
+
+# FQR
+fqr <- lapply(RLQm.l, function(x){mapply(function(r,l,q){fourthcorner(r, l, q, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")}, r = x$Rm.l, l = x$Lm.l, q = x$Qm.l, SIMPLIFY = F)})
+names(fqr$full) <- names(Lm.l) ; names(fqr$`-WD`) <- names(Lm.l)
+
+# Result
+# par(mfrow = c(2,3))
+# lapply(fqr, function(x){lapply(x, plot)})
+fqr.df.l <- mapply(function(y,z){
+  fqr.df <- with(y$tabG, cbind(data.frame(t(matrix(unlist(strsplit(names, ' / ')), nrow = 2)), stringsAsFactors = F), obs, pvalue))
+  fqr.wd <- with(z$tabG, cbind(data.frame(t(matrix(unlist(strsplit(names, ' / ')), nrow = 2)), stringsAsFactors = F), obs, pvalue))
+  fqr.df <- rbind(fqr.df, fqr.wd[which(fqr.wd$X2 == 'WD'),])
+  names(fqr.df)[1:2] <- c('Factor', 'Trait') 
+  fqr.df$res <- paste(round(fqr.df$obs, 3), '/', round(fqr.df$pvalue, 3), unlist(lapply(as.list(fqr.df$pvalue), stars, ns = ' ')))
+  fqr.df <- tapply(fqr.df$res, INDEX = list(fqr.df$Trait, fqr.df$Factor), paste, simplify = F)
+  return(fqr.df)
+  }, y = fqr$`-WD`, z = fqr$full, SIMPLIFY = F)
+```
+```{r FQR abundance}
+kable(t(fqr.df.l$abundance))
 ```
 
-Figure 4: RLQ Monte-Carlo test results for model 6 proposed by @ter_braak_improved_2012.
+## Are phylogenetic composition of communities linked to the environment ?
 
-## Fourth corner analysis
+Only Thick, LDMC, WES and at a fewer level LES shown a phylogenetic signal positively structured in the phylogenetic tree at a phylogenetic distance between 0 and 100 (see supplementary material S3).
+$\beta st$ and $\pi st$ indicates both for respectivelly abundance and presence absence data that our phylogeny is overdispersed inside communities: species within communities are less phylogenetically related on average than are specis from distinct communities.
+But compared to the model null *1s* suggested by @hardy_characterizing_2007 , the phylogenetic overdispersion is almost null except in recent years (from 50 Myrs).
+Finally, total inertia from RLQ analysis with tip distances of phylogenetic tree as matrix Q didn't differ from a null model were all phylogenetic tree tips were randomized ($p-value = 0.5344$ with 999 repetitions.)
 
-```{r 4thQ AB}
-library(ade4)
-# FQR
-fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
-fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
-
-# Results
-par(mfrow = c(1,2))
-plot(fqr)
-plot(fqr.adj)
-# print(fqr.adj)
-```
-
-Figure 5: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species abundances in communities with non adjusted p-values (left) and adjusted p-values (right).
-
-```{r 4thQ BIN}
-library(ade4)
-
-# Matrices
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-
-# Consistency
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[row.names(Qm)]
-Lm[Lm > 0] <-  1 #Presence-absence
-Lm <- Lm / rowSums(Lm)
-
-# RLQ
-L <- dudi.coa(Lm, scannf = F)
-R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
-Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
-rlq <- rlq(R, L, Q, scannf = F)
-fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
-fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
-
-# Results
-par(mfrow = c(1,2))
-plot(fqr)
-plot(fqr.adj)
-# print(fqr.adj)
-```
-
-Figure 6: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species presence absence in communities with non adjusted p-values (left) and adjusted p-values (right).
-
-```{r 4thQ BA}
-library(ade4)
-
-# Matrices
-Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum)) #BA
-Lm[is.na(Lm)] <- 0
-
-# Consistency
-Lm <- Lm[row.names(Rm),]
-Qm <- Qm[names(Lm),]
-Qm <- Qm[-which(is.na(Qm$WD)),]
-Lm <- Lm[row.names(Qm)]
-Lm <- Lm / rowSums(Lm)
-
-# RLQ
-L <- dudi.coa(Lm, scannf = F)
-R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
-Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
-rlq <- rlq(R, L, Q, scannf = F)
-fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
-fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
-
-# Results
-par(mfrow = c(1,2))
-plot(fqr)
-plot(fqr.adj)
-# print(fqr.adj)
-```
-
-Figure 7: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species basal area in communities with non adjusted p-values (left) and adjusted p-values (right).
-
-## Phylogenetic signal
-
-Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 measured traits and the two strategy axes and by looking at the phylogenetic correlogram wiht the `phylosignal` R package [@keck_phylosignal:_2016].
-Thickness, leaf dry matter content, and wood economic spectrum were the only one to be significantly non null with a positive correlation for all of theim (between 0.01 and 0.03) in a short phylogenetic distance between 0 and approximatively 100.
-
-```{r phylosignal}
+```{r PhyloMaker}
 library(phylosignal)
@@
 # Building species list
-sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), 2:3]
+l <- PFT_sp
+# setdiff(row.names(l), Species$LatinName)
+row.names(l)[which(row.names(l) == 'Actinodaphne tadulingami')] <- 'Actinodaphne tadulingamii'
+Species['agin',] <- c('agin', "Agrostistachys indica", 'Euphorbiaceae', NA, NA)
+Species['anme',] <- c('anme', "Antidesma menasu", 'Euphorbiaceae', NA, NA)
+Species['anmo',] <- c('anme', "Antidesma montanum", 'Euphorbiaceae', NA, NA)
+row.names(l)[which(row.names(l) == 'Blepharistemma membranifolia')] <- 'Blepharistemma serratum '
+row.names(l)[which(row.names(l) == 'Casearia ovata')] <- 'Casearia ovata '
+row.names(l)[which(row.names(l) == 'Cinnamomum malabathrum')] <- 'Cinnamomum malabatrum'
+row.names(l)[which(row.names(l) == 'Cinnamomum sp1')] <- 'Cinnamomum sp.'
+row.names(l)[which(row.names(l) == 'Dimocarpus longan')] <- 'Dimocarpus longan '
+row.names(l)[which(row.names(l) == 'Drypetes confertiflorus')] <- 'Drypetes confertiflora'
+row.names(l)[which(row.names(l) == 'Glochidion malabaricum')] <- 'Glochidion ellipticum'
+row.names(l)[which(row.names(l) == 'Homalium zeylanicum')] <- 'Homalium zeylanicum '
+row.names(l)[which(row.names(l) == 'Ixora nigricams')] <- 'Ixora nigricans  '
+row.names(l)[which(row.names(l) == 'Knema attenuata')] <- 'Knema attenuata  '
+row.names(l)[which(row.names(l) == 'Leptonychia moacurroides')] <- 'Leptonychia caudata '
+row.names(l)[which(row.names(l) == 'Myristica malabarica')] <- 'Myristica malabarica '
+row.names(l)[which(row.names(l) == 'Olea dioica')] <- 'Olea dioica '
+row.names(l)[which(row.names(l) == 'Reinwardtiodendron anaimalaiense')] <- 'Reinwardtiodendron anamalaiense'
+row.names(l)[which(row.names(l) == 'Strombosia ceylanica')] <- 'Strombosia ceylanica '
+row.names(l)[which(row.names(l) == 'Symplocos cochinensis')] <- 'Symplocos cochinchinensis'
+row.names(l)[which(row.names(l) == 'Syzygium laetum')] <- 'Syzygium laetum '
+row.names(l)[which(row.names(l) == 'Vepris bilocularis')] <- 'Vepris bilocularis '
+sp <- Species[which(Species$LatinName %in% row.names(l)), 2:3]
 sp$genus <- unlist(lapply(strsplit(sp$LatinName, ' ', fixed = T), function(x){x[1]}))
@@
 sp[which(sp$family == 'Cesaliniaceae'),3] <- 'Fabaaceae'
-# sp['cato','family'] <- 'Lamiaceae'
-# sp['clvi',] <- c('Clerodendrum_infortunatum', 'Clerodendrum', 'Lamiaceae')
-# sp['glma','family'] <-'Phyllanthaceae'
-# sp['gote','family'] <- 'Stemonuraceae'
-# sp['goca','family'] <- 'Annonaceae'
-# sp['liol','family'] <- 'Lauraceae'
-# sp['mewi','family'] <- 'Melastomataceae'
-# sp['blde',] <- c('Blachia_andamanica', 'Blachia', 'Euphorbiaceae')
 sp['hepa','family'] <- 'Malvaceae'
-# sp['lemo','family'] <- 'Malvaceae'
-# sp['hubr','family'] <- 'Leguminosae' # Humboldtia is unmatched !
-sp <- sp[-which(row.names(sp) == 'hubr'),]
-sp <- sp[-which(row.names(sp) %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
+# sp <- sp[-which(row.names(sp) == 'hubr'),]
+# sp <- sp[-which(row.names(sp) %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
 
@@
 results <- S.PhyloMaker(sp, phylo, nodes)
-
-# Phylosignal
 tree <- results$Scenario.1
-data <- PFT_sp
-row.names(data) <- sub(' ', '_', row.names(data), fixed = T)
-data <- data[tree$tip.label,c(2:5,7,9:10)]
-p4d <- phylo4d(tree, data)
-# barplot.phylo4d(p4d, tree.type = "phylo", tree.ladderize = TRUE)
-# phylosim <- phyloSim(tree = tre, method = "all", nsim = 100, reps = 99)
-# plot.phylosim(phylosim, what = "pval", stacked.methods = TRUE)
-phyCor <- lapply(as.list(names(data)), function(x){phyloCorrelogram(p4d, trait = as.character(x))})
-par(mfrow = c(2,4))
-for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
-lipa <- lipaMoran(p4d)
-# barplot.phylo4d(p4d, bar.col=(lipa$p.value < 0.05) + 1, center = F, scale = F)
 ```
 
-Once the phylogenetic structure revealed we wanted to test it in our community with the Bst and PIst metric proposed by @hardy_characterizing_2007 with the null model 1s suggested by [cite] to see if we were under phylogenetic clustering or overdispersion.
+```{r Phylogenetic Tree, eval=FALSE, include=FALSE}
+# Completing the phylogenetic tree
+# Humboldtia: Close to Kingiodendron pinnatum according to Bruneau et al. 2001 Systematic Botany 26(3):487-514
+# Humboldtia and Kingiodendron are our only two Leguminosae
+# zoom(tree, 70:75)
+node <- which(tree$tip.label == "Kingiodendron_pinnatum")
+position<- 0.1*tree$edge.length[which(tree$edge[,2]==node)]
+tree <- bind.tip(tree,"Humboldtia_brunonis",where = node, position = position)
+# zoom(tree, 60:80)
+```
 
-```{r phylo structure}
+```{r Phylosignal}
+data <- l
+row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
+tips <- row.names(data)[which(is.na(l$WD))]
+phy <- list(full = drop.tip(tree, tips), `-WD` = tree)
+data <- list(full = data[phy$full$tip.label,c(2:5,7)], `-WD` = data[phy$`-WD`$tip.label,c(3:5,7)])
+p4d <- mapply(function(p,d){phylo4d(p,d)}, p = phy, d = data)
+# lapply(p4d, barplot)
+phyloS <- lapply(p4d, phyloSignal, methods = 'I')
+phyloS <- lapply(phyloS, function(x){cbind(x$stat, x$pvalue)})
+phyloS <- rbind(phyloS$`-WD`, phyloS$full['WD',])
+names(phyloS) <- c('Moran\'s I', 'p-value')
+phyloS$` ` <- lapply(as.list(phyloS$`p-value`), stars)
+phyCor <- mapply(function(d,p){lapply(as.list(names(d)), function(x){phyloCorrelogram(p, trait = as.character(x))})}, d = data, p = p4d)
+phyCor <- c(phyCor$`-WD`, list(phyCor$full[[1]]))
+names(phyCor) <- unlist(lapply(phyCor, function(x){x$trait}))
+lipa <- lapply(p4d, lipaMoran)
+# mapply(function(x,y){barplot.phylo4d(x, bar.col=(y$p.value < 0.05) + 1)}, x = p4d, y = lipa)
+kable(phyloS)
+```
+
+```{r Phylo RLQ, fig.height=4}
+library(adephylo)
 library(spacodiR)
 
-# Phylogeny renaming
-tree$tip.label <- gsub('_', ' ', tree$tip.label, fixed = T)
-tree$tip.label <- as.character(SpCode[tree$tip.label])
+# RLQ
+for(i in 1:2) phy[[i]]$tip.label <- gsub('_', ' ', phy[[i]]$tip.label, fixed = T)
+for(i in 1:2) phy[[i]]$tip.label <- as.character(l$Sp_Code[match(phy[[i]]$tip.label, row.names(l))])
+Lm <- lapply(phy, function(x){Lm[intersect(names(Lm), x$tip.label)]})
+L <- lapply(Lm, dudi.coa, scannf = F)
+R <- mapply(function(x,y){dudi.pca(x, scannf = F, row.w = y$lw)}, x = rep(list(Rm), 2), y = L, SIMPLIFY = F)
+phyRLQ <- mapply(function(x,y){drop.tip(x,setdiff(x$tip.label, names(y)))}, x = phy, y = Lm, SIMPLIFY = F)
+Qd <- lapply(phyRLQ, distTips)
+Q <- mapply(function(x,y){dudi.pco(x, scannf = F, row.w = y$cw)}, x = Qd, y = L, SIMPLIFY = F)
+rlq <- mapply(function(x,y,z){rlq(x, y, z, scannf = F)}, x = R, y = L, z = Q, SIMPLIFY = F)
 
-# Matrices
+# Random RLQs
+phyRD <- lapply(phy, function(x){x$node.label <- NULL ; return(x)})
+phyRD <- lapply(phyRD, function(x){rep(list(x), 999)})
+phyRD <- lapply(phyRD, function(x){lapply(x, resamp.phy)})
+LmRD <- lapply(Lm, function(x){rep(list(x), 999)})
+LmRD <- mapply(function(x,y){mapply(function(r,s){r[intersect(names(r), s$tip.label)]}, r = x, s = y, SIMPLIFY = F)}, x = LmRD, y = phyRD, SIMPLIFY = F)
+phyRD <- mapply(function(x,y){mapply(function(r,s){drop.tip(r, setdiff(r$tip.label, names(s)))}, r = x, s= y, SIMPLIFY = F)}, x = phyRD, y = LmRD, SIMPLIFY = F)
+QdRD <- lapply(phyRD, function(x){lapply(x, distTips)})
+LRD <- lapply(LmRD, function(x){lapply(x, dudi.coa, scannf = F)})
+RmRD <- lapply(rep(list(full = Rm, `-WD` = Rm)), function(x){rep(list(x), 999)})
+RRD <- mapply(function(x,y){mapply(function(r,s){dudi.pca(r, scannf = F, row.w = s$lw)}, r = x, s = y, SIMPLIFY = F)}, x = RmRD, y = LRD, SIMPLIFY = F)
+QRD <- mapply(function(x,y){mapply(function(r,s){dudi.pco(r, scannf = F, row.w = s$cw)}, r = x, s = y, SIMPLIFY = F)}, x = QdRD, y = LRD, SIMPLIFY = F)
+rlqRD <- mapply(function(x,y,z){mapply(function(r,s,t){rlq(r, s, t, scannf = F)}, r = x, s = y, t = z, SIMPLIFY = F)}, x = RRD, y = LRD, z = QRD, SIMPLIFY = F)
+
+# Randtest
+pval <- mapply(function(x,y){rank(c(sum(x$eig), unlist(lapply(y, function(z){sum(z$eig)}))))[1]/1000}, x = rlq, y = rlqRD)
+par(mfrow = c(1,2))  
+invisible(mapply(function(x,y,z,l){
+  hist(unlist(lapply(y, function(r){sum(r$eig)})), main = paste('Model', l, ':', 'pval =', round(z, 3)), xlab = 'Coinertia', col = 'lightgrey', border = 'white')
+  lines(c(sum(x$eig), sum(x$eig)), c(0,500), col = 'red', lwd = 2)
+}, x = rlq, y = rlqRD, z = as.list(pval), l = as.list(names(rlqRD))))
+```
+
+## Are communities spatially structured ?
+
+```{r Spatial structure}
+library(spacodiR)
+
+# Traits structure
 Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
 Lm[is.na(Lm)] <- 0
-Rm <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
-row.names(Rm) <- com@data$id
+Lm.l <- list(full = Lm, `-WD` = Lm)
+Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
+row.names(Qm) <- PFT_sp$Sp_Code
+Qm.l <- list(full = Qm, `-WD` = Qm[-3])
+Qm.l <- mapply(function(x,y){x[intersect(row.names(x), names(y)),]}, x = Qm.l, y = Lm.l, SIMPLIFY = F)
+Qm.l$full <- Qm.l$full[-which(is.na(Qm.l$full$WD)),]
+Lm.l <- mapply(function(x,y){y[intersect(row.names(x), names(y))]}, x = Qm.l, y = Lm.l, SIMPLIFY = F)
+spafun <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), sp.traits = x)}, x = Qm.l, y = Lm.l, SIMPLIFY = F)
+spafun <- do.call('rbind.data.frame', lapply(spafun, function(x){x[1:4]}))
 
-# Consistency
-# all(row.names(Lm) == row.names(Rm))
-# all(row.names(Lm) %in% row.names(Rm))
-Lm <- Lm[row.names(Rm),]
-# all(names(Lm) == tree$tip.label)
-Lm <- Lm[tree$tip.label]
+Qm.lRD <- lapply(Qm.l, function(x){rep(list(x),999)})
+Qm.lRD <- lapply(Qm.lRD, function(x){lapply(x, function(y){y[sample(row.names(y)),]})})
+Lm.lRD <- lapply(Lm.l, function(x){rep(list(x),999)})
+Lm.lRD <- mapply(function(x,y){mapply(function(r,s){r[row.names(s)]}, r = x, s = y, SIMPLIFY = F)}, x = Lm.lRD, y = Qm.lRD, SIMPLIFY = F)
+spafun.l <- mapply(function(x,y){mapply(function(r,s){spacodi.calc(sp.plot = t(s), sp.traits = r)}, r = x, s = y, SIMPLIFY = F)}, x = Qm.lRD, y = Lm.lRD, SIMPLIFY = F)
+spafun.l <- lapply(spafun.l, function(x){lapply(x, function(y){y[1:4]})})
+spafunRD <- list(Ist = lapply(spafun.l, function(x){unlist(lapply(x, function(y){y$Ist}))}),
+                 Tst = lapply(spafun.l, function(x){unlist(lapply(x, function(y){y$Tst}))}),
+                 Ust = lapply(spafun.l, function(x){unlist(lapply(x, function(y){y$Ust}))}),
+                 TAUst = lapply(spafun.l, function(x){unlist(lapply(x, function(y){y$TAUst}))}))
+pvalues <- mapply(function(obs,rd){
+  mapply(function(o,r){rank(c(o, r))[1]/1000}, o = obs, r = rd)
+  }, obs = lapply(as.list(spafun), as.list), rd = spafunRD)
+row.names(pvalues) <- row.names(spafun)
 
 # Phylogeny structure
-spafun <- spacodi.calc(sp.plot = t(Lm), sp.traits = Qm)
-spaphy <- spacodi.calc(sp.plot = t(Lm), phy = tree)
-permut <- lapply(as.list(names(spaphy[1:4])), function(x){spacodi.by.nodes(t(Lm), tree, sp.parm = x, n.rep = 100, method = "1s")})
-names(permut) <- names(spaphy[1:4])
-par(mfrow=c(2,1))
-spacodi.permutplot(permut$Bst)
-spacodi.permutplot(permut$PIst)
+spaphy <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), phy = x)}, x = phy, y = Lm.l, SIMPLIFY = F)
+spaphy <- do.call('rbind.data.frame', lapply(spaphy, function(x){x[1:4]}))
+# permut <- list(Bst = mapply(function(x,y){spacodi.by.nodes(sp.plot = t(y), x, sp.parm = "Bst", method = "1s")}, x = phy, y = Lm.l, SIMPLIFY = F), PIst = mapply(function(x,y){spacodi.by.nodes(sp.plot = t(y), x, sp.parm = "PIst", method = "1s")}, x = phy, y = Lm.l, SIMPLIFY = F))
+# par(mfrow=c(2,2)) ; invisible(lapply(permut, function(x){lapply(x, spacodi.permutplot)}))
 ```
 
-The phylogenetic structure among our communities was then evaluated by an RLQ analysis with cophenetic matrix of our phylogenetic tree distances as the Q matrix reduced by a principal coordinate analysis.
+# Discussion
 
-```{r phylo RLQ}
-library(ade4)
-library(adephylo)
+# *Acknowledgement*
 
-# RLQ
-L <- dudi.coa(Lm, scannf = F)
-R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
-Qd <- distTips(tree)
-Q <- dudi.pco(Qd, scannf = F, row.w = L$cw)
-rlq <- rlq(R, L, Q, scannf = F)
+# Supplementary materials
 
-# Random RLQs
-tree$node.label <- NULL
-ntrees <- rep(list(tree), 999)
-ntrees <- lapply(ntrees, resamp.phy)
-nQd <- lapply(ntrees, distTips)
-nrlq <- lapply(ntrees, function(x){
-  Lm <- Lm[x$tip.label]
-  L <- dudi.coa(Lm, scannf = F)
-  R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
-  Q <- dudi.pco(Qd, scannf = F, row.w = L$cw)
-  rlq(R, L, Q, scannf = F)})
+*I'm aware that Supplementary materials goes after References (as figure and tables), but it's not possible for the moment with Rmarkdown and knitr.*
 
-# Randtest
-pval <- rank(c(sum(rlq$eig), unlist(lapply(nrlq, function(x){sum(x$eig)}))))[1]/1000
-hist(unlist(lapply(nrlq, function(x){sum(x$eig)})), main = paste('pval =', round(pval, 3)))
-lines(c(sum(rlq$eig), sum(rlq$eig)), c(0,500), col = 'red', lwd = 2)
+## S1: Fourth corner analysis results in presence-absence and relative basal area
+
+```{r FQR others}
+kable(t(fqr.df.l$`presence-absence`))
+kable(t(fqr.df.l$`basal area`))
 ```
 
-Figure 8: RLQ test by model nul results with phylogenetic tree distances as the Q matrix for model 1s proposed by [cite].
+## S2: Phylogenetic signal among functionnal traits by correlogram
+
+Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits
+
+```{r phylosignal plots}
+par(mfrow = c(2,3))
+invisible(lapply(phyCor, function(x){plot(x, main = x$trait)}))
+```
 

### 2016-07-08 inst/doc/draft.Rmd
v0.1.9017 RLQ.Rmd + plan

@@
 The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996.
-The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergree forest type [@pascal_wet_1988].
+The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988].
 Climate and geology of the 10-ha plot can be considered as homogeneous.
 General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
-
-*Grith monitoring design ?*
+Each tree girth is monitored by a two-year census.
 
@@
 Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.
-Leaves and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
+Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
 For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).
 
+Leaf petioles, rachis and petiolules were removed for all measurements.
 Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite].
@@
 Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues.
-Branch wood samples were then sink in water to evaluate their volume by water displacement.
-Samples were after dried for 48 hours at 80 degrees celsius and weighted.
-
-- Removed petiole, petiolule and rachis
+Branch wood samples were sink in water to evaluate their volume by water displacement.
+Samples were then dried for 48 hours at 80 degrees celsius and weighted.
 
@@
 
-- Traits measurement (handbook)
-- 5 Basic traits description
-- Associated strategies axes with PCA (+ fortunel)
-- Associated covariation (following litterature)
-- Diversity indices (FD)
+With leaf and branch wood samples we calculated five functionnal traits: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the wood sample volume in $cm^3$ divided by the wood sample dry weight in $g$ (see table 1).
+Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
+For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+
+Table 1: Functional traits measured.
+*Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development from @baraloto_decoupled_2010.*
+
+```{r 1 Traits}
+rm(list = ls()); invisible(gc())
+library(Uppangala)
+PFT <- genPFT()
+captions <- genCaptions()
+table <- data.frame(
+  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
+  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
+  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
+            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
+            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
+            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
+            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
+  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
+                         round(sd(PFT$Thick, na.rm = T)), ')'),
+                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
+                         round(sd(PFT$LA, na.rm = T)), ')'),
+                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
+                         round(sd(PFT$LDMC, na.rm = T)), ')'),
+                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
+                         round(sd(PFT$SLA, na.rm = T),2), ')'),
+                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
+                         round(sd(PFT$WD, na.rm = T),2), ')')),
+  unit = c('um', 'mm2', 'mg.g-1', 'm2.kg-1', 'g.cm-3'),
+  strategy = c('Leaf defense vs. investment',
+               'Leaf ressource capture vs.investment',
+               'Leaf defense vs. investment',
+               'Leaf ressource acquisiton vs. defense',
+               'Stem transport, structure and defense')
+  )
+kable(table,  label="traits", row.names = F)
+```
 
@@
 
-- Abitoic factors
-- PCA reduction
-- Biotic factors
-- Bitoic uncorrelation
-- Performance variables
-- Linking traits, biotic, abiotic factors and performance variables with sem
+To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
+Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
+Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
+Distance to the nearest potential river was caluclated under ArcGis (cite).
+
+To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density.
+Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$.
+Both basal area and stem density are calculated with 2013 girth inventory from Uppangala plot on stem with a diameter above 10 centimeters at breast height ($DBH>10$).
+Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$).
+So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
 
@@
 
-- Community defintions
-- Community spatial organization investigation (variorgam + SAC)
-- Link with environment: step measure, linear regression, null model, residuals SAC
-- Individual filtering (RLQ)
-- Phylogenetic signal
+### Covariation in traits
+
+Traits variation was first investigated in three levels with the coefficient of variation ($cv=sd/mean$) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
+For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
+To test the correlation and covariation among traits we performed a principal component analysis (PCA).
+We investigated peculiar traits covariation highlighted with the PCA with a Spearman's rank correlation test.
+
+### Environment definition
+
+In order to avoid multi-colinearity in future analysis, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
+In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling.
+Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
+Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
+Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).
+
+### Community analysis
+
+Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
+For all communities, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes.
+We used the following formula to compute each community weighted parameter (CWP):
+$$CWP = \sum_{k=1}^{nj} W_{k,j} * P_{k,j} ~ (1)$$ 
+where $nj$ is the number of species sampled in community $j$, $W_{k,j}$ is the weight of species $k$ in community $j$ used to compute CWP (relative abundance, presence-absence, or relative basal area), and $P_{k,j}$ is the parameter of species $k$ in community $j$ (mean or the variance).
+In addition 5 functionnal diversity indices (FD) weighted by abundances were computed: functional richness (FRic), functional evenness (FEve), functional divergence (FDiv) and Rao's quadratic entropy (Q) (@villeger_new_2008, @botta-dukat_raos_2005).
+Spatial variograms were used to investigate our CWM and CWV spatial organization.
+CWM, CWV, and FD spatial autocorrelation (SAC) was also assessed by Moran's I statistic.
+
+Links between both CWM and CWV with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
+Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables.
+A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix).
+Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance.
+Same procedure was also applied to functionnal diversity indices (FD) from our communities.
+Finally spatial autocorrelation (SAC) of residuals was investigated to assess selected models.
+
+### Fourth-corner analysis
+
+In addition of our community analysis by regression, we conducted both RLQ and fourth-corner analyses to test the effect of environments on Uppangala tree functionnal traits following @dray_testing_2008.
+Both matrices R (`Environment*Sites`) with selected features and Q (`Traits*Species`) were reduced by a principal component analysis.
+We used a Monte-Carlo test to compare fourth-corner and RLQ results to 999 nul models built with the model *m6* proposed by @ter_braak_improved_2012.
+Fourth corner analysis was tested with 3 types of L matrices ($Sites*Species$) with species relative abundance, presence absence or relative basal area in each sites; whereas RLQ analysis was only conducted with relative abundance.
+
+### Phylogenetic signal
+
+Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
+Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 functionnal traits, the two PCA axes and in the phylogenetic tree.
+Phylogenetic signal structure was assessed by a tree tip distances correlogram.
+In addition, $\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate phylogenetic signal spatial organization with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
+Finally phylogenetic association to habitat was also studied by an RLQ analysis with tree tip distances as the Q matrix reduced by a principal coordinate analysis.
+RLQ result with phylogenetic distance matrix was compared to 999 nul models were phylogenetic tree tips were randomized.
+
+All analysis were conducted in the R 3.3.0  free software environment for statistical computing and graphics [@r_core_team_r:_2016] using the pacakge 
+`ade4` [@dray_ade4_2007], 
+`lavaan` [@rosseel_lavaan:_2012], 
+`FD` [@laliberte_distance-based_2010], 
+`gstat` [@pebesma_multivariable_2004], 
+`ape` [@paradis_ape:_2004], 
+`phylosignal` [@keck_phylosignal:_2015] 
+and `spacodiR` [@eastman_spacodir:_2013].
 
@@
 
-## Community weighted means
+## Traits and environment
 
-### Abiotic filtering
+Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific.
+*Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
+See table 1 for abbreviation.*
 
-### Biotic filtering
+```{r 2 CV}
+PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
+row.names(PFT_ind) <- PFT_ind$Group.1
+SpT <- unique(PFT[1:2])
+Sp <- SpT$SP
+names(Sp) <- SpT$Tree
+SpCodeT <- unique(PFT[2:3])
+SpCode <- SpCodeT$Sp_Code
+names(SpCode) <- SpCodeT$SP
+CET <- unique(PFT[c(1,4)])
+CE <- CET$CE
+names(CE) <- CET$Tree
+PFT_ind$SP <- Sp[row.names(PFT_ind)]
+PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
+PFT_ind$CE <- CE[row.names(PFT_ind)]
+PFT_ind <- PFT_ind[-c(1:2)]
+cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
+cv$Individual <- unlist(lapply(
+  as.list(row.names(cv)),
+  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
+cv$Species <- unlist(lapply(
+  as.list(row.names(cv)),
+  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
+cv$Community <- unlist(
+  lapply(as.list(row.names(cv)),
+         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
+             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
+cv <- round(cv, 2)
+cv['WD','Individual'] <- ''
+kable(cv)
+```
 
-## Community weighted variances
+Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
+Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$).
 
-### Abiotic filtering
+The multivariate analysis of trait correlation shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
+The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood.
+Whereas the second PCA axis (PCA2) covary mainly with SLA representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
+PCA first axis might be associated with the WD covariation to the wood economic spectrum [WES, @chave_towards_2009].
+Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
+Our results integrate less traits (5 against 14), and less species (92 against 758), but still capture the same strategy axes.
 
-### Biotic filtering
+Figure 1 shows WD orthogonal to SLA following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$).
+LA is almost opposed to WD and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007.
+Finally, we found the positive relation suggested by @diaz_global_2015 between LDMC and WD (Spearman's $rho = 0.24$, $p-value < 0.001$).
 
-## Functionnal diversity
+```{r 3 PCA PFT}
+library(vegan)
+pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_ind, cor = T)
+lambda <- pca$sdev * sqrt(nrow(pca$scores))
+pca.var = apply(pca$score, 2, var)
+PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
+names(PCAscores) <- c('WES', 'LES')
+PCAscores$ID <- row.names(PCAscores)
+plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
+     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
+     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
+     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
+# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
+par(new=T)
+Rot <- t(t(pca$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
+lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
+legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
+       text.col = c("green", "firebrick"))
 
-## Individual filtering
+arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
+text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
+Hmisc::subplot(function(){
+  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
+          cex.names = 0.7, axes = T, names.arg = NA) ;
+  lines(b, bstick(length(pca.var)), col = 'red') ;
+  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
+         box.col = 'white')},
+  x=grconvertX(c(0.75,1), from='npc'),
+  y=grconvertY(c(0.05,0.30), from='npc')
+)
+axis (3)
+axis (4)
+PFT_ind$ID <- row.names(PFT_ind)
+PFT_ind <- merge(PFT_ind, PCAscores, all = T)
+row.names(PFT_ind) <- PFT_ind$ID
+PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
+```
+
+Figure 1: Principal component analysis (PCA) performed on individual trait means.
+*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
+See table 1 for abbreviation.*
+
+Multi-colinearity was avoided by reducing our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see figure 2).
+Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
+Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
+Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary.
+Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)
+
+```{r 4 env-com}
+library(raster)
+Env <- genEnv()
+Env$Rivers <- readAll(Env$Rivers)
+Trees <- genTrees()
+Species <- genSpecies()
+PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
+row.names(PFT_sp) <- PFT_sp$Group.1
+PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
+PFT_sp <- PFT_sp[-c(1:2,4)]
+Trees$SpCode <- tolower(Trees$SpCode)
+Trees$ID <- row.names(Trees)
+Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
+row.names(Trees) <- Trees$ID
+Trees <- Trees[-which(names(Trees) == 'ID')]
+com <- quadrats()
+Trees$com <- NA
+XY <- data.frame(X = Trees$X, Y = Trees$Y)
+XY <- XY[-which(is.na(XY$X)),] # Removing NA
+coordinates(XY) <- ~ X + Y
+proj4string(XY) <- crs(com)
+Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
+Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
+com@data$id <- paste0('C', com@data$id)
+com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
+com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
+com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$SW <- sin((com@data$Aspect + 225) / 180 * pi)
+com@data$NE <- cos((com@data$Aspect + 225) / 180 * pi)
+com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
+com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
+com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
+# com@data$Height <- aggregate(Trees$height, list(Trees$com), mean, na.rm = T)[,2]
+rm(Deriv, XY)
+pca.env <- princomp(~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
+lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
+pca.env.var = apply(pca.env$score, 2, var)
+plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
+     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
+     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
+     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
+Hmisc::subplot(function(){
+  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
+          cex.names = 0.7, axes = T, names.arg = NA) ;
+  lines(b, bstick(length(pca.env.var)), col = 'red') ;
+  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
+         box.col = 'white')},
+  x=grconvertX(c(0.07,0.37), from='npc'),
+  y=grconvertY(c(0.025,0.30), from='npc')
+)
+par(new=T)
+Rot <- t(t(pca.env$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, lwd = 2)
+lines(c(0,0), XLIM, lty = 2, lwd = 2)
+col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
+arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
+        Rot[,1],Rot[,2],lwd=2,col=col)
+legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
+       text.col = col, cex = 1.2)
+text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 1), rownames(Rot),cex=1.2,col=col)
+axis (3)
+axis (4)
+```
+
+Figure 2: Principal component analysis (PCA) performed on topgraphical variables.
+*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
+SW stand for southwesterness.*
+
+## Community analysis
+
+Spatial variograms of CWM and CWV didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$. Functionnal diversity indices show also a plateau around $70 m$.
+
+```{r 5 Com, fig.height=4}
+library(FD)
+com <- CWT(Trees, com)
+x <- data.frame(PFT_sp[-1], row.names = PFT_sp$Sp_Code)
+a <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+a[is.na(a)] <- 0
+x <- x[names(a),]
+FD <- as.data.frame(dbFD(x, a, corr = 'cailliez', calc.CWM = F, print.pco = F, messages = F))[-c(1:2,4)]
+names(FD) <- paste0(names(FD), '.fd')
+FD$id <- row.names(FD)
+com$abundance <- merge(com$abundance, FD)
+library(gstat)
+dist <- variogram(WD.cwm~1, as(com$abundance, 'SpatialPointsDataFrame'))$dist
+par(mfrow = c(1,2))
+for(i in 1:2){
+  traits <- paste0(c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES'), c('.cwm','.cwv')[i])
+  plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', main = c('(a)', '(b)')[i], xlab = 'distance (m)', ylab = 'relative variance (%)')
+  for(j in 1:length(com)){
+    var <- data.frame(lapply(lapply(as.list(traits), function(x){variogram(formula(paste(x, '~ 1')), as(com[[j]], 'SpatialPointsDataFrame'))}), function(x){x[3]}))
+    var <- data.frame(apply(var, 2, function(x){x/max(x)}))
+    names(var) <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
+    colramp <- rainbow(length(var))
+    for(i in 1:length(var)){
+      lines(var[,i] ~ dist, col = colramp[i], lty = j + 1)
+    }
+  }
+legend('bottomright', names(var), fill = colramp, cex = 0.7)
+legend('bottomleft', names(com), lty = c(2:4), cex = 0.7)
+}
+```
+
+Figure 4: Trait variograms for all communities.*Variograms were initiated on* $20*20 m^2$ *communities for all traits with both community weighted means* **(a)** *and variances* **(b)** *(by abundance, presence-absence and basal area).
+Some variaograms seem to reach a plateau around* $70 m$.
+
+Only half of CWMs and CWVs were significantly positively spatially autocorrelated (see table 3) with low values (Moran's I from 0.01 to 0.04).
+CWMs of LES and LDMC as CWV of WD are slightly spatially correlated suggesting a spatial filter.
+Remaining traits community weighted means and variances are either stochasticaclly spatially distributed either spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter.
+Moreover all functionnal diversity indices (FD) are spatially correlated.
+
+Table 3: Spatial autocorrelation measured by Moran's I for community weighted means and variances.
+_Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations.
+FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r SAC}
+library(ape)
+traits <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
+traits <- c(paste0(traits, '.cwm'), paste0(traits, '.cwv'))
+traits <- list(c(traits, 'FRic.fd', 'FEve.fd', 'FDiv.fd', 'FDis.fd', 'RaoQ.fd'), traits, traits)
+weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
+diag(weights) <- 0
+SAC <- mapply(function(x,y){lapply(as.list(x@data[y]), Moran.I, weights)}, x =com, y = traits)
+SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
+SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
+SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
+SAC <- data.frame(SAC)
+names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
+SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
+SAC$Type <- ''
+SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
+SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
+kable(SAC, row.names = F)
+```
+
+```{r 6 lm & null models}
+reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
+# LM to simplify with loop or lapply
+LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
+                                                 logSLA.cwm = lm(formula(paste('log(SLA.cwm)', reg))), 
+                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
+                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
+                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
+                                                 logLA.cwm = lm(formula(paste('log(LA.cwm)', reg))),
+                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
+                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
+                                                 logSLA.cwv = lm(formula(paste('log(SLA.cwv)', reg))), 
+                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
+                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
+                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
+                                                 logLA.cwv = lm(formula(paste('log(LA.cwv)', reg))),
+                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
+LM$abundance <- c(LM$abundance, with(com$abundance@data, list(FRic.fd = lm(formula(paste('FRic.fd', reg))),
+                                                              FEve.fd = lm(formula(paste('FEve.fd', reg))),
+                                                              FDiv.fd = lm(formula(paste('FDiv.fd', reg))),
+                                                              FDis.fd = lm(formula(paste('FDis.fd', reg))),
+                                                              RaoQ.fd = lm(formula(paste('RaoQ.fd', reg))))))
+LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
+model <- null_model(com, Trees, PFT_sp, lapply(LM, function(y){lapply(y, function(x){x$call})}), SpCode, n = 999, verbose = F, time = T)
+```
+
+Whereas the CWM of the LES, and therefore SLA, show significantly non-random associations with abiotic factors ($p-values < 0.03$, see table 4), most of CWMs associated to the WES on the PCA didn't differ from a null distribution ($p-values > 0.1$).
+CWMs of LES and SLA are associated with convex and dry surfaces, as CWM of Thick weighted by relative basal area.
+CWM of WD weighted by presence absence is also significantly ($p-value = 0.014$) non-randomly associated to southwest steep slopes with concave surfaces, with a relatively important explanation level ($R^2_{adjusted} = 0.25$).
+Regarding biotic factors only CWMs of LA and WD, respectively weighted by relative basal area and presence absence, show significant ($p-values < 0.05$) association with canopy height.
+
+Table 4: Summary for linear models and null models for traits community weighted means.
+_Linear model results for each trait community weighted means (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r CWM}
+kable(regression_table(LM, model, reg, select = '.cwm'))
+```
+
+Only CWVs of LES and WD are significantly affected by both abiotic and biotic environment compare to null models ($p-values < 0.05$, see table 5).
+LES converge with better soil and water conditions in convex situations, and for abundant species oriented under monsoon wind LES diverge suggesting different adaptive strategies from species.
+Whereas CWVs of WD, weighted by relative abundance and presence absence, also diverged with the southwest orientation to monsoon winds, CWV of WD weighted by relative basal area converge with slope and diverge with the canopy height.
+Finally, CWV of LA weighted by relative basal area show a significantly positive linear model with slope compare to null models ($p-value = 0.04$).
+
+Table 5: Summary for linear models and null models for traits community weighted variances.
+_Linear model results for each trait community weighted variances (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r CWV}
+kable(regression_table(LM, model, reg, select = '.cwv'))
+```
+
+FD indices show communities relatively functionnaly poor but with an even and diverse trait distribution ($mean( \frac{FRic}{global~Fric} ) = 0.054$, $mean(FEve) = 0.794$, and $mean() = 0.72$) relatively stable ($0.04 \ge \sigma _FD \ge 0.07$).
+Except FRic, all FD indices didn't differ from null models suggesting that our bioctic and abitotic factors are not responsible for their distribution in our communities ($p-values > 0.1$, see table 6).
+Still FRic increases with slope, suggesting as CWV that our traits diverge in different adaptive strategies in steeper slopes.
+
+Table 6: Summary for linear models and null models for functionnal diversity indices (FD).
+_Linear model results for each community functionnal indices weighted by abundances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r FD}
+kable(regression_table(LM, model, reg, select = '.fd')[-2], row.names = F)
+```
+
+Spatial autocorrelation (SAC) of model residuals for CWMs is mostly null confirming that our models explain fully the spatial structure of our 5 traits (see supplementary materials S2).
+Beside CWMs of Thick still show a spatial organization ($I_Moran = 0.01$ $p-value < 0.01$) suggesting that our model missed spatially structured factor(s) in this case.
+SAC of models residuals for CWV show also a spatial organization for SLA, WES, and LDMC.
+Finally all FD show a significantly non null SAC for their model residuals.
+
+## Fourth-corner analysis
+
+Whereas total co-inertia from RLQ analysis results was significanlty different for the model *m2*, it didn't differ from null model *m4* in model *m6* proposed by @ter_braak_improved_2012.
+So because both test should be rejected to reject model *m6* hypothesis, therefore the null hypothesis *H0* is not rejected.
+
+Still fourth-corner analysis with adjusted p-values or not show significant trait-environment association at a level $\alpha = 0.05$ (see table 7).
+WD is negatively associated with southwesterness for all L matrices and positively associated with canopy height in presence absence.
+LDMC and LA are respectively negatively and positively associated with canopy height in relative abundance.
+
+Table 7: Summary of fourth-corner analysis.
+*Result of Monte-Carlo test between fourth-corner result and 999 nul models following the model m6 suggested by @ter_braak_improved_2012.* _L matrices are filled with relative abundance (AB), presence-absence (PA) or relative basal area (BA)).
+The 5 functionnal traits are tested with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+See table 1 for trait abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r FQR, fig.height=4}
+library(ade4)
+
+# Abundance
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
+row.names(Qm) <- PFT_sp$Sp_Code
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Qm <- Qm[names(Lm),]
+Rm <- com$abundance@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+row.names(Rm) <- com$abundance@data$id
+Lm <- Lm[row.names(Rm),]
+Qm <- Qm[names(Lm),]
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Lm <- Lm[row.names(Qm)]
+Lm <- Lm / rowSums(Lm)
+Lm.l = list(abundance = Lm)
+
+# RLQ
+L <- dudi.coa(Lm, scannf = F)
+R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
+Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
+rlq <- rlq(R, L, Q, scannf = F)
+rd <- randtest(rlq, modeltype = 6)
+
+# Presence absence
+Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Lm <- Lm[row.names(Rm),]
+Lm <- Lm[row.names(Qm)]
+Lm[Lm > 0] <-  1
+Lm <- Lm / rowSums(Lm)
+Lm.l$`presence-absence` = Lm
+
+# Basal area
+Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
+Lm[is.na(Lm)] <- 0
+Lm <- Lm[row.names(Rm),]
+Qm <- Qm[names(Lm),]
+Qm <- Qm[-which(is.na(Qm$WD)),]
+Lm <- Lm[row.names(Qm)]
+Lm <- Lm / rowSums(Lm)
+Lm.l$`basal area` = Lm
+
+# FQR
+fqr <- mapply(function(r,l,q){fourthcorner(r, l, q, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")}, r = rep(list(Rm), 3), l = Lm.l, q = rep(list(Qm), 3), SIMPLIFY = F)
+names(fqr) <- names(Lm.l)
+fqr.adj <- lapply(fqr, p.adjust.4thcorner, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")
+
+# Result
+par(mfrow = c(1,3))
+lapply(fqr, plot)
+```
+
+**WAITING FOR RESULTS AS TABLE**
 
 ## Phylogenetic signal
+
+Only Thick, LDMC, WES and at a fewer level LES shown a phylogenetic signal positively structured in the phylogenetic tree at a phylogenetic distance between 0 and 100 (see supplementary material S3).
+$\beta st$ and $\pi st$ indicates both for respectivelly abundance and presence absence data that our phylogeny is overdispersed inside communities: species within communities are less phylogenetically related on average than are specis from distinct communities.
+But compared to the model null *1s* suggested by @hardy_characterizing_2007 , the phylogenetic overdispersion is almost null except in recent years (from 50 Myrs).
+Finally, total inertia from RLQ analysis with tip distances of phylogenetic tree as matrix Q didn't differ from a null model were all phylogenetic tree tips were randomized ($p-value = 0.5344$ with 999 repetitions.)
+
+-*Taxomonic analysis of traits clustering among the phylogenetic tree (LIPA)*
+
+```{r phylosignal}
+library(phylosignal)
+library(phytools)
+library(phylobase)
+
+# Building species list
+sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), 2:3]
+sp$genus <- unlist(lapply(strsplit(sp$LatinName, ' ', fixed = T), function(x){x[1]}))
+sp$species <- sub(' ', '_', sp$LatinName, fixed = T)
+sp$family <- sp$Family
+sp <- sp[c('species', 'genus', 'family')]
+sp[which(sp$family == 'Flacourtiaceae'),3] <- 'Salicaceae'
+sp[which(sp$family == 'Steruliaceae'),3] <- 'Malvaceae'
+sp[which(sp$family == 'Cesaliniaceae'),3] <- 'Fabaaceae'
+sp['hepa','family'] <- 'Malvaceae'
+sp <- sp[-which(row.names(sp) == 'hubr'),]
+sp <- sp[-which(row.names(sp) %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
+
+# Getting phylogeny
+path <- system.file('phylogeny', package = 'Uppangala')
+source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
+phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
+nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
+results <- S.PhyloMaker(sp, phylo, nodes)
+
+# Phylosignal
+tree <- results$Scenario.1
+data <- PFT_sp
+row.names(data) <- sub(' ', '_', row.names(data), fixed = T)
+data <- data[tree$tip.label,c(2:5,7,9:10)]
+p4d <- phylo4d(tree, data)
+phyCor <- lapply(as.list(names(data)), function(x){phyloCorrelogram(p4d, trait = as.character(x))})
+```
 
@@
 
+# Supplementary materials
+
+*I'm aware that Supplementary materials goes after References (as figure and tables), but it's not possible for the moment with Rmarkdown and knitr.*
+
+## S1: Structural equation model
+
+```{r SEM}
+library(lavaan)
+m1 <- '
+Abiotic =~ Slope + Curvature + Wetness + SW
+Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
+Biotic =~ BA + Canopy
+Traits ~ Abiotic + Biotic
+'
+m2 <- '
+Abiotic =~ Slope + Curvature + Wetness + SW
+Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
+Biotic =~ BA + Canopy
+Biotic ~ Traits + Abiotic
+'
+fit1 <- sem(m1, data = com$abundance@data)
+fit2 <- sem(m2, data = com$abundance@data)
+semPlot::semPaths(fit1,
+                  what = 'col',
+                  whatLabels = 'std',
+                  style = 'lisrel',
+                  residuals = T,
+                  nCharNodes = 0,
+                  nCharEdges = 0,
+                  groups='manlat',
+                  pastel = T,
+                  ask = F)
+```
+
+Supplementary material S1: Path diagram for structural equation modelling.
+*Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic.
+Arrow values are indicating standardized parameter estimates*.
+
+## S2: Spatial autocorrelation of models residuals
+
+Supplementary materials S2: Spatial autocorrelation measured by Moran's I for linear regression model residuals from community weighted means (CWM) and variances (CWV) and functionnal diversity indices (FD).
+_Spatial autocorrelation for residuals from each trait community weighted means and variances (by abundance, presence-absence and basal area) linear models and for functionnal diversity (FD) linear models, see table 1 for abbreviations.
+FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r SAC residuals}
+library(ape)
+weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
+diag(weights) <- 0
+SAC <- lapply(LM, function(x){lapply(x, function(y){Moran.I(residuals(y), weights)})})
+SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
+SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
+SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
+SAC <- data.frame(SAC)
+names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
+SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
+SAC$Type <- ''
+SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
+SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
+kable(SAC, row.names = F)
+```
+
+## S3: Phylogenetic signal among functionnal traits by correlogram
+
+Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits.
+```{r phylosignal plots}
+par(mfrow = c(2,4))
+for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
+```
+
 # References

### 2016-07-08 R/genEnv.R

@@ -31,5 +31,7 @@ genEnv <- function(area = data.frame(x = c(571675.8, 572011.3),
   Rivers <- crop(Rivers, extent)
   Canopy <- raster(system.file('extdata', 'Canopy.tif', package = 'Uppangala'))
   Canopy <- crop(Canopy, extent)
-  return(list(DEM = DEM, Rivers = Rivers, Canopy = Canopy))
+  return(list(DEM = DEM,
+              Rivers = Rivers,
+              Canopy = Canopy))
 }

### 2016-07-08 R/stars.R

@@ -1,6 +1,7 @@
 #' Stars
 #'
 #' @param val num. p-value to convert in stars
+#' @param ns char. non significant character
 #'
 #' @return char. Number of stars corresponding to the p-value
 #' @export
@@ -8,7 +9,7 @@
 #' @examples
 #' stars(0.049)
 #'
-stars = function(val){
+stars = function(val, ns = 'n.s.'){
   if(val < 0.001){
     s = '*** '
   } else if (val < 0.01) {
@@ -18,7 +19,7 @@ stars = function(val){
   } else if (val < 0.1) {
     s = '.   '
   } else {
-    s = 'n.s.'
+    s = ns
   }
   return(s)
 }

### 2016-07-18 inst/doc/RLQ.Rmd
v0.1.9018 Pres & Francois cleaning

@@
 ---
-title: 'Uppangala tree functionnal: RLQ draft'
+title: 'Uppangala tree functional structure: RLQ draft'
 author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
@@
 library(knitr)
-library(pander)
 opts_chunk$set(
@@
 
+Things to include :
+
+- Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
+For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+- Leaf investment definition
+- Analyses scheme presentation:
+    + Are functionnal composition of communities linked to the environment ?
+        - Do covariations in functionnal traits illustrate ecological strategy axes ?
+    + Are phylogenetic composition of communities linked to the environment
+        - Are functionnal traits phylogenetically conserved ?
+    + Are communities spatially structured ?
+        - Are functionnal composition of communities spatially structured ?
+        - Are phylogenetic composition of communities spatially structured ?
+
 # Material and methods
@@
 Climate and geology of the 10-ha plot can be considered as homogeneous.
-General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
+General topography of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as a major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
 Each tree girth is monitored by a two-year census.
+**!
+Developper girth census !**
 
@@
 Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
-For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).
+For some individuals we were unable to collect an appropriate branch sample (58 over 267 sampled trees representing 12 species over 91 and 1.3% of individuals).
 
@@
 
-## Traits
+## Trait measurements
 
-With leaf and branch wood samples we calculated five functionnal traits: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the wood sample volume in $cm^3$ divided by the wood sample dry weight in $g$ (see table 1).
-Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
-For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+We calculate five functional traits of leaf and wood samples: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the wood sample dry weight in $g$ divided by the wood sample volume in $cm^3$ (see table 1).
 
@@
                'Leaf defense vs. investment',
-               'Leaf ressource acquisiton vs. defense',
+               'Leaf ressource acquisition vs. defense',
                'Stem transport, structure and defense')
@@
 
-To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
+To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness index (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
 Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
@@
 To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density.
-Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$.
+Canopy height is calculated from a digital canopy model built with a LIDAR campaign on Uppangala **date ?** plots with a resolution $1*1 m^2$.
 Both basal area and stem density are calculated with 2013 girth inventory from Uppangala plot on stem with a diameter above 10 centimeters at breast height ($DBH>10$).
 Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$).
-So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
+So we removed stem density to avoid multicolinearity, and we kept canopy height that was not correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
 
@@
 
-### Covariation in traits
+*Title are just here to help remembering the structure but might be removed*
 
-Traits variation was first investigated in three levels with the coefficient of variation ($cv=sd/mean$) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
+### Functionnal composition of communities and environment
+
+### Phylogenetic composition of communities and environment
+
+### Communitie compositions spatial structure
+
+<!-- Traits variation was first investigated in three levels with the coefficient of variation ($cv=sd/mean$) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
 For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
@@
 We investigated peculiar traits covariation highlighted with the PCA with a Spearman's rank correlation test.
+-->
 
-### Environment definition
+<!-- ### Environment definition -->
 
-In order to avoid multi-colinearity in future analysis, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
+<!-- In order to avoid multi-colinearity in future analysis, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
 In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling.
@@
 Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).
+-->
 
-### Community analysis
+<!-- ### Community analysis -->
 
-Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
+<!-- Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
 For all communities, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes.
-We used the following formula to compute each community weighted parameter (CWP):
-$$CWP = \sum_{k=1}^{nj} W_{k,j} * P_{k,j} ~ (1)$$ 
-where $nj$ is the number of species sampled in community $j$, $W_{k,j}$ is the weight of species $k$ in community $j$ used to compute CWP (relative abundance, presence-absence, or relative basal area), and $P_{k,j}$ is the parameter of species $k$ in community $j$ (mean or the variance).
+We used the following formula to compute each community weighted parameter (CWP): -->
+<!-- $$CWP = \sum_{k=1}^{nj} W_{k,j} * P_{k,j} ~ (1)$$  -->
+<!-- where $nj$ is the number of species sampled in community $j$, $W_{k,j}$ is the weight of species $k$ in community $j$ used to compute CWP (relative abundance, presence-absence, or relative basal area), and $P_{k,j}$ is the parameter of species $k$ in community $j$ (mean or the variance).
 In addition 5 functionnal diversity indices (FD) weighted by abundances were computed: functional richness (FRic), functional evenness (FEve), functional divergence (FDiv) and Rao's quadratic entropy (Q) (@villeger_new_2008, @botta-dukat_raos_2005).
@@
 CWM, CWV, and FD spatial autocorrelation (SAC) was also assessed by Moran's I statistic.
+-->
 
-Links between both CWM and CWV with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
+<!-- Links between both CWM and CWV with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
 Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables.
@@
 Finally spatial autocorrelation (SAC) of residuals was investigated to assess selected models.
+-->
 
-### Fourth-corner analysis
+<!-- ### Fourth-corner analysis -->
 
-In addition of our community analysis by regression, we conducted both RLQ and fourth-corner analyses to test the effect of environments on Uppangala tree functionnal traits following @dray_testing_2008.
+<!-- In addition of our community analysis by regression, we conducted both RLQ and fourth-corner analyses to test the effect of environments on Uppangala tree functionnal traits following @dray_testing_2008.
 Both matrices R (`Environment*Sites`) with selected features and Q (`Traits*Species`) were reduced by a principal component analysis.
@@
 Fourth corner analysis was tested with 3 types of L matrices ($Sites*Species$) with species relative abundance, presence absence or relative basal area in each sites; whereas RLQ analysis was only conducted with relative abundance.
+-->
 
-### Phylogenetic signal
+<!-- ### Phylogenetic signal -->
 
-Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
+<!-- Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
 Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 functionnal traits, the two PCA axes and in the phylogenetic tree.
@@
 RLQ result with phylogenetic distance matrix was compared to 999 nul models were phylogenetic tree tips were randomized.
+-->
 
-All analysis were conducted in the R 3.3.0  free software environment for statistical computing and graphics [@r_core_team_r:_2016] using the pacakge 
+All analysis were conducted in the R 3.3.0  free software environment for statistical analyses and graphics [@r_core_team_r:_2016] using the package 
 `ade4` [@dray_ade4_2007], 
@@
 Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
-Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$).
+Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
 
@@
 Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
-Qm <- Qm[-which(is.na(Qm$WD)),]
 Lm <- Lm[row.names(Rm),]
@@
 
-# RLQ
-L <- dudi.coa(Lm, scannf = F)
-R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
+# Q plot
 Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
-rlq <- rlq(R, L, Q, scannf = F)
-
-# Q plot
 plot(t(t(Q$li)/Q$eig[1:2]),pch=16, col = 'lightgrey',
@@
 lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
-lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
-# legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
-       # text.col = c("green", "firebrick"))
+legend('topright', c('Plant economic spectrum'), text.col = c("green", "firebrick"))
 Hmisc::subplot(function(){
@@
 axis (4)
-# arrows (rep(0,nrow(R$co)),rep(0,nrow(R$co)),R$co[,1],R$co[,2],lwd=1, col = 'blue')
-# text(R$co, pos = c(4, 2, 4, 2, 3), rownames(R$co),cex=0.9,col='blue')
 arrows (rep(0,nrow(Q$co)),rep(0,nrow(Q$co)),Q$co[,1],Q$co[,2],lwd=2)
@@
 
-Table 7: Summary of fourth-corner analysis.
-*Result of Monte-Carlo test between fourth-corner result and 999 nul models following the model m6 suggested by @ter_braak_improved_2012.* _L matrices are filled with relative abundance (AB), presence-absence (PA) or relative basal area (BA)).
-The 5 functionnal traits are tested with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-See table 1 for trait abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+Table 7: Summary of CWM et CWV analyses.
 
 ```{r FQR}
-library(ade4)
-rd <- randtest(rlq, modeltype = 6)
-
 # L matrices list
@@
 Lm.l$`presence-absence` = Lm
-Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm <- Lm / rowSums(Lm)
-Lm.l$`basal area` = Lm
 
@@
 Lm.l$`presence-absence` = Lm
-Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm <- Lm / rowSums(Lm)
-Lm.l$`basal area` = Lm
-RLQm.l$`-WD` <- list(Rm.l = rep(list(Rm),3), Lm.l = Lm.l, Qm.l = rep(list(Qm),3))
-# str(RLQm.l,3)
-
-# FQR
-fqr <- lapply(RLQm.l, function(x){mapply(function(r,l,q){fourthcorner(r, l, q, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")}, r = x$Rm.l, l = x$Lm.l, q = x$Qm.l, SIMPLIFY = F)})
-names(fqr$full) <- names(Lm.l) ; names(fqr$`-WD`) <- names(Lm.l)
-
-# Result
-# par(mfrow = c(2,3))
-# lapply(fqr, function(x){lapply(x, plot)})
-fqr.df.l <- mapply(function(y,z){
-  fqr.df <- with(y$tabG, cbind(data.frame(t(matrix(unlist(strsplit(names, ' / ')), nrow = 2)), stringsAsFactors = F), obs, pvalue))
-  fqr.wd <- with(z$tabG, cbind(data.frame(t(matrix(unlist(strsplit(names, ' / ')), nrow = 2)), stringsAsFactors = F), obs, pvalue))
-  fqr.df <- rbind(fqr.df, fqr.wd[which(fqr.wd$X2 == 'WD'),])
-  names(fqr.df)[1:2] <- c('Factor', 'Trait') 
-  fqr.df$res <- paste(round(fqr.df$obs, 3), '/', round(fqr.df$pvalue, 3), unlist(lapply(as.list(fqr.df$pvalue), stars, ns = ' ')))
-  fqr.df <- tapply(fqr.df$res, INDEX = list(fqr.df$Trait, fqr.df$Factor), paste, simplify = F)
-  return(fqr.df)
-  }, y = fqr$`-WD`, z = fqr$full, SIMPLIFY = F)
-```
-```{r FQR abundance}
-kable(t(fqr.df.l$abundance))
 ```
@@
 
-```{r Spatial structure}
+```{r Functionnal spatial structure}
 library(spacodiR)
@@
 
-# Phylogeny structure
+pvalues <- as.data.frame(pvalues)[3:4]
+pvalues$Ust <-  1- pvalues$Ust
+spafun.res <- format(spafun[3:4], scientific = T, digits = 2)
+pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
+spafun.res$Ust <- paste(spafun.res$Ust, '/', pvalues$Ust)
+spafun.res$TAUst <- paste(spafun.res$TAUst, '/', pvalues$TAUst)
+kable(spafun.res)
+```
+
+```{r Phylogenetic spatial structure}
 spaphy <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), phy = x)}, x = phy, y = Lm.l, SIMPLIFY = F)
 spaphy <- do.call('rbind.data.frame', lapply(spaphy, function(x){x[1:4]}))
+
+spaphy.l <- mapply(function(x,y){mapply(function(r,s){spacodi.calc(t(r), s)}, r = x, s = y, SIMPLIFY = F)},x = LmRD, y = phyRD, SIMPLIFY = F)
+spaphy.l <- lapply(spaphy.l, function(x){lapply(x, function(y){y[1:4]})})
+spaphyRD <- list(Ist = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Ist}))}),
+                 Tst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Tst}))}),
+                 Bst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Bst}))}),
+                 PIst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$PIst}))}))
+pvalues <- mapply(function(obs,rd){
+  mapply(function(o,r){rank(c(o, r))[1]/1000}, o = obs, r = rd)
+  }, obs = lapply(as.list(spaphy), as.list), rd = spaphyRD)
+row.names(pvalues) <- row.names(spaphy)
+
+pvalues <- as.data.frame(pvalues)[3:4]
+pvalues$Bst <-  rep(NaN,2)
+spaphy.res <- format(spaphy[3:4], scientific = T, digits = 2)
+pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
+spaphy.res$Bst <- paste(spaphy.res$Bst, '/', pvalues$Bst)
+spaphy.res$PIst <- paste(spaphy.res$PIst, '/', pvalues$PIst)
+kable(spaphy.res)
+
 # permut <- list(Bst = mapply(function(x,y){spacodi.by.nodes(sp.plot = t(y), x, sp.parm = "Bst", method = "1s")}, x = phy, y = Lm.l, SIMPLIFY = F), PIst = mapply(function(x,y){spacodi.by.nodes(sp.plot = t(y), x, sp.parm = "PIst", method = "1s")}, x = phy, y = Lm.l, SIMPLIFY = F))
@@
 # Supplementary materials
-
-*I'm aware that Supplementary materials goes after References (as figure and tables), but it's not possible for the moment with Rmarkdown and knitr.*
-
-## S1: Fourth corner analysis results in presence-absence and relative basal area
-
-```{r FQR others}
-kable(t(fqr.df.l$`presence-absence`))
-kable(t(fqr.df.l$`basal area`))
-```
 

### 2016-07-18 inst/doc/RLQpres.Rmd
v0.1.9018 Pres & Francois cleaning

--- 2016-07-18 inst/doc/RLQ.Rmd
@@
 ---
-title: 'Uppangala tree functional structure: RLQ draft'
+title: Study of spatial structure of taxonomic and functional diversity in a tropical forest tree community
 author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
-output: pdf_document
+output: 
+  ioslides_presentation: default
+  beamer_presentation:
+    fig_caption: yes
+    fonttheme: professionalfonts
+  slidy_presentation: default
 bibliography: refs.bib
@@
 opts_chunk$set(
-  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
+  echo = F, message = F, warning = F, fig.height = 5, fig.width = 8,
   cache = T, cache.lazy = F)
@@
 
-# Abstract
+## Functional ecology
 
-Keywords: functionnal trait, tropical evergreen forest, Western Ghats, India...
+```{r PFT scheme}
+library(png)
+library(grid)
+img <- readPNG("RLQpres_files/figure-html/PFT.png")
+ grid.raster(img)
+```
 
-# Introduction
+## Community ecology
 
-Things to include :
+```{r Community scheme}
+library(png)
+library(grid)
+img <- readPNG("RLQpres_files/figure-html/community.png")
+ grid.raster(img)
+```
 
-- Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
-For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
-- Leaf investment definition
-- Analyses scheme presentation:
-    + Are functionnal composition of communities linked to the environment ?
-        - Do covariations in functionnal traits illustrate ecological strategy axes ?
-    + Are phylogenetic composition of communities linked to the environment
-        - Are functionnal traits phylogenetically conserved ?
-    + Are communities spatially structured ?
-        - Are functionnal composition of communities spatially structured ?
-        - Are phylogenetic composition of communities spatially structured ?
+## Analysis scheme
 
-# Material and methods
-
-## Study site
-
-The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village.
-The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996.
-The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988].
-Climate and geology of the 10-ha plot can be considered as homogeneous.
-General topography of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as a major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
-Each tree girth is monitored by a two-year census.
-**!
-Developper girth census !**
+```{r Analysis scheme}
+library(png)
+library(grid)
+img <- readPNG("RLQpres_files/figure-html/Uppangala6.png")
+ grid.raster(img)
+```
 
@@
 
-We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates.
-Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals.
-Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot.
-For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013.
-Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy.
-Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.
-Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
-For some individuals we were unable to collect an appropriate branch sample (58 over 267 sampled trees representing 12 species over 91 and 1.3% of individuals).
+- 92 species over 117 (79%)
+- Representing 98.8% of individuals
+- 272 trees
+- 1362 dried leaves
+- 272 fresh leaves
+- 250 fresh branch wood sample
 
-Leaf petioles, rachis and petiolules were removed for all measurements.
-Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite].
-Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time.
-Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias.
-Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues.
-Branch wood samples were sink in water to evaluate their volume by water displacement.
-Samples were then dried for 48 hours at 80 degrees celsius and weighted.
+## Direct measurements
 
-## Trait measurements
+- Tree crown exposure
+- Leaf fresh weight
+- Leaf thickness
+- Leaf area
+- Leaf dry weight
+- Wood volume
+- Wood dry weight
 
-We calculate five functional traits of leaf and wood samples: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the wood sample dry weight in $g$ divided by the wood sample volume in $cm^3$ (see table 1).
-
-Table 1: Functional traits measured.
-*Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development from @baraloto_decoupled_2010.*
+## Functional traits
 
@@
                'Leaf defense vs. investment',
-               'Leaf ressource acquisition vs. defense',
+               'Leaf ressource acquisiton vs. defense',
                'Stem transport, structure and defense')
@@
 
-## Environmental variables
+Table 1: Functional traits measured.
 
-To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness index (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
-Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
-Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
-Distance to the nearest potential river was caluclated under ArcGis (cite).
+## Traits coefficients of variation
 
-To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density.
-Canopy height is calculated from a digital canopy model built with a LIDAR campaign on Uppangala **date ?** plots with a resolution $1*1 m^2$.
-Both basal area and stem density are calculated with 2013 girth inventory from Uppangala plot on stem with a diameter above 10 centimeters at breast height ($DBH>10$).
-Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$).
-So we removed stem density to avoid multicolinearity, and we kept canopy height that was not correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
-
-## Analyses
-
-*Title are just here to help remembering the structure but might be removed*
-
-### Functionnal composition of communities and environment
-
-### Phylogenetic composition of communities and environment
-
-### Communitie compositions spatial structure
-
-<!-- Traits variation was first investigated in three levels with the coefficient of variation ($cv=sd/mean$) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
-For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
-To test the correlation and covariation among traits we performed a principal component analysis (PCA).
-We investigated peculiar traits covariation highlighted with the PCA with a Spearman's rank correlation test.
--->
-
-<!-- ### Environment definition -->
-
-<!-- In order to avoid multi-colinearity in future analysis, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
-In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling.
-Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
-Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
-Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).
--->
-
-<!-- ### Community analysis -->
-
-<!-- Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
-For all communities, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes.
-We used the following formula to compute each community weighted parameter (CWP): -->
-<!-- $$CWP = \sum_{k=1}^{nj} W_{k,j} * P_{k,j} ~ (1)$$  -->
-<!-- where $nj$ is the number of species sampled in community $j$, $W_{k,j}$ is the weight of species $k$ in community $j$ used to compute CWP (relative abundance, presence-absence, or relative basal area), and $P_{k,j}$ is the parameter of species $k$ in community $j$ (mean or the variance).
-In addition 5 functionnal diversity indices (FD) weighted by abundances were computed: functional richness (FRic), functional evenness (FEve), functional divergence (FDiv) and Rao's quadratic entropy (Q) (@villeger_new_2008, @botta-dukat_raos_2005).
-Spatial variograms were used to investigate our CWM and CWV spatial organization.
-CWM, CWV, and FD spatial autocorrelation (SAC) was also assessed by Moran's I statistic.
--->
-
-<!-- Links between both CWM and CWV with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
-Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables.
-A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix).
-Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance.
-Same procedure was also applied to functionnal diversity indices (FD) from our communities.
-Finally spatial autocorrelation (SAC) of residuals was investigated to assess selected models.
--->
-
-<!-- ### Fourth-corner analysis -->
-
-<!-- In addition of our community analysis by regression, we conducted both RLQ and fourth-corner analyses to test the effect of environments on Uppangala tree functionnal traits following @dray_testing_2008.
-Both matrices R (`Environment*Sites`) with selected features and Q (`Traits*Species`) were reduced by a principal component analysis.
-We used a Monte-Carlo test to compare fourth-corner and RLQ results to 999 nul models built with the model *m6* proposed by @ter_braak_improved_2012.
-Fourth corner analysis was tested with 3 types of L matrices ($Sites*Species$) with species relative abundance, presence absence or relative basal area in each sites; whereas RLQ analysis was only conducted with relative abundance.
--->
-
-<!-- ### Phylogenetic signal -->
-
-<!-- Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
-Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 functionnal traits, the two PCA axes and in the phylogenetic tree.
-Phylogenetic signal structure was assessed by a tree tip distances correlogram.
-In addition, $\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate phylogenetic signal spatial organization with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
-Finally phylogenetic association to habitat was also studied by an RLQ analysis with tree tip distances as the Q matrix reduced by a principal coordinate analysis.
-RLQ result with phylogenetic distance matrix was compared to 999 nul models were phylogenetic tree tips were randomized.
--->
-
-All analysis were conducted in the R 3.3.0  free software environment for statistical analyses and graphics [@r_core_team_r:_2016] using the package 
-`ade4` [@dray_ade4_2007], 
-`lavaan` [@rosseel_lavaan:_2012], 
-`FD` [@laliberte_distance-based_2010], 
-`gstat` [@pebesma_multivariable_2004], 
-`ape` [@paradis_ape:_2004], 
-`phylosignal` [@keck_phylosignal:_2015] 
-and `spacodiR` [@eastman_spacodir:_2013].
-
-# Results
-
-## Are functionnal composition of communities linked to the environment ?
-
-Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific.
-*Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
-See table 1 for abbreviation.*
+*How vary functional traits ?*
 
@@
 
-Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
-Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
+Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific.
 
-The multivariate analysis of trait correlation shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
-The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood.
-Whereas the second PCA axis (PCA2) covary mainly with SLA representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
-PCA first axis might be associated with the WD covariation to the wood economic spectrum [WES, @chave_towards_2009].
-Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
-Our results integrate less traits (5 against 14), and less species (92 against 758), but still capture the same strategy axes.
+## Functional traits PCA
 
-Figure 1 shows WD orthogonal to SLA following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$).
-LA is almost opposed to WD and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007.
-Finally, we found the positive relation suggested by @diaz_global_2015 between LDMC and WD (Spearman's $rho = 0.24$, $p-value < 0.001$).
+*Do covariation in functional traits illustrate ecological strategy axes ?*
 
-```{r 3 RLQ}
+```{r 3 PCA PFT}
 library(vegan)
-library(ade4)
+
+PFT_ind$LA <- log(PFT_ind$LA)
+PFT_ind$SLA <- log(PFT_ind$SLA)
+PFT_ind$ID <- row.names(PFT_ind)
+row.names(PFT_ind) <- PFT_ind$ID
+PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
+Species <- genSpecies()
+PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
+row.names(PFT_sp) <- PFT_sp$Group.1
+PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
+PFT_sp <- PFT_sp[-c(1:2,4)]
+pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_sp, cor = T)
+
+lambda <- pca$sdev * sqrt(nrow(pca$scores))
+pca.var = apply(pca$score, 2, var)
+PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
+names(PCAscores) <- c('WES', 'LES')
+PCAscores$ID <- row.names(PCAscores)
+plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
+     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
+     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
+     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
+# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
+par(new=T)
+Rot <- t(t(pca$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
+lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
+legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
+       text.col = c("green", "firebrick"))
+
+arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
+text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
+Hmisc::subplot(function(){
+  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
+          cex.names = 0.7, axes = T, names.arg = NA) ;
+  lines(b, bstick(length(pca.var)), col = 'red') ;
+  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
+         box.col = 'white')},
+  x=grconvertX(c(0.75,1), from='npc'),
+  y=grconvertY(c(0.05,0.30), from='npc')
+)
+axis (3)
+axis (4)
+
+PCAscores <- as.data.frame(pca$scores)[1:2]
+names(PCAscores) <- c('WES', 'LES')
+PCAscores$ID <- row.names(PCAscores)
+PFT_sp$ID <- row.names(PFT_sp)
+PFT_sp <- merge(PFT_sp, PCAscores, all = T)
+row.names(PFT_sp) <- PFT_sp$ID
+PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
+```
+
+Figure 1: Principal component analysis on traits
+
+## Environment features PCA
+
+*Do environment variables covariate ?*
+
+```{r 4 env-com}
 library(raster)
-
-# Data
+Env <- genEnv()
+Env$Rivers <- readAll(Env$Rivers)
 Trees <- genTrees()
-Species <- genSpecies()
-Env <- genEnv()
-com <- quadrats()
-com@data$id <- paste0('C', com@data$id)
-
-# L matrix
 Trees$SpCode <- tolower(Trees$SpCode)
 Trees$ID <- row.names(Trees)
+Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
 row.names(Trees) <- Trees$ID
 Trees <- Trees[-which(names(Trees) == 'ID')]
+com <- quadrats()
 Trees$com <- NA
@@
 Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-
-# R matrix
+Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
+com@data$id <- paste0('C', com@data$id)
 com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'cosaspect'))
+Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
+names(Deriv)[5] <- 'SW'
 com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
-names(com@data)[5] <- 'SW'
 com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
 com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
 com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
-Rm <- com@data[-1]
-row.names(Rm) <- com@data$id
-
-# Q matrix
-PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
-row.names(PFT_sp) <- PFT_sp$Group.1
-PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
-PFT_sp <- PFT_sp[-c(1:2,4)]
-PFT_sp$SLA <- log(PFT_sp$SLA)
-PFT_sp$LA <- log(PFT_sp$LA)
-Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
-row.names(Qm) <- PFT_sp$Sp_Code
-
-# Matrices adjustments
-Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm <- Lm / rowSums(Lm)
-
-# Q plot
-Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
-plot(t(t(Q$li)/Q$eig[1:2]),pch=16, col = 'lightgrey',
-     xlab = paste('Axe 1 -', round(Q$eig[1]/sum(Q$eig)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(Q$eig[2]/sum(Q$eig)*100, 2), '%'))
-par(new=T)
-XLIM <- c(-max(abs(Q$co[,1])),max(abs(Q$co[,1])))
-XLIM <- XLIM+(XLIM*0.2)
-plot(Q$co,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
-lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
-legend('topright', c('Plant economic spectrum'), text.col = c("green", "firebrick"))
+com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
+rm(Deriv, XY)
+pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
+lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
+pca.env.var = apply(pca.env$score, 2, var)
+plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
+     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
 Hmisc::subplot(function(){
-  b <- barplot((Q$eig / sum(Q$eig)), las = 2, ylim = c(0,range(bstick(7))[2]),
+  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
           cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(Q$eig)), col = 'red') ;
+  lines(b, bstick(length(pca.env.var)), col = 'red') ;
   legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
          box.col = 'white')},
-  x=grconvertX(c(0.65,0.95), from='npc'),
-  y=grconvertY(c(0.01,0.26), from='npc')
+  x=grconvertX(c(0.07,0.37), from='npc'),
+  y=grconvertY(c(0.025,0.30), from='npc')
 )
+par(new=T)
+Rot <- t(t(pca.env$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, lwd = 2)
+lines(c(0,0), XLIM, lty = 2, lwd = 2)
+col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
+arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
+        Rot[,1],Rot[,2],lwd=2,col=col)
+legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
+       text.col = col, cex = 1.2)
+text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
 axis (3)
 axis (4)
-arrows (rep(0,nrow(Q$co)),rep(0,nrow(Q$co)),Q$co[,1],Q$co[,2],lwd=2)
-text (Q$co, pos = c(2, 2, 2, 4, 4), rownames(Q$co),cex=1.2)
 ```
 
-Figure 1: Principal component analysis (PCA) performed on individual trait means.
-*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
-See table 1 for abbreviation.*
+Figure 2: Principal component analysis performed on environment
 
-Multi-colinearity was avoided by reducing our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see figure 2).
-Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
-Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
-Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary.
-Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)
+## Communities definition
 
-Whereas total co-inertia from RLQ analysis results was significanlty different for the model *m2*, it didn't differ from null model *m4* in model *m6* proposed by @ter_braak_improved_2012.
-So because both test should be rejected to reject model *m6* hypothesis, therefore the null hypothesis *H0* is not rejected.
-
-Still fourth-corner analysis with adjusted p-values or not show significant trait-environment association at a level $\alpha = 0.05$ (see table 7).
-WD is negatively associated with southwesterness for all L matrices and positively associated with canopy height in presence absence.
-LDMC and LA are respectively negatively and positively associated with canopy height in relative abundance.
-
-Table 7: Summary of CWM et CWV analyses.
-
-```{r FQR}
-# L matrices list
-Lm.l <- list(abundance = Lm)
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm[Lm > 0] <-  1
-Lm <- Lm / rowSums(Lm)
-Lm.l$`presence-absence` = Lm
-
-# Full # -WD
-RLQm.l <- list(full = list(Rm.l = rep(list(Rm),3), Lm.l = Lm.l, Qm.l = rep(list(Qm),3)))
-Qm <- PFT_sp[c('SLA', 'LDMC', 'Thick', 'LA')]
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-row.names(Qm) <- PFT_sp$Sp_Code
-Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm <- Lm / rowSums(Lm)
-Lm.l <- list(abundance = Lm)
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm[Lm > 0] <-  1
-Lm <- Lm / rowSums(Lm)
-Lm.l$`presence-absence` = Lm
+```{r 5 Com, fig.height=4}
+com <- CWT(Trees, com)
 ```
 
-## Are phylogenetic composition of communities linked to the environment ?
-
-Only Thick, LDMC, WES and at a fewer level LES shown a phylogenetic signal positively structured in the phylogenetic tree at a phylogenetic distance between 0 and 100 (see supplementary material S3).
-$\beta st$ and $\pi st$ indicates both for respectivelly abundance and presence absence data that our phylogeny is overdispersed inside communities: species within communities are less phylogenetically related on average than are specis from distinct communities.
-But compared to the model null *1s* suggested by @hardy_characterizing_2007 , the phylogenetic overdispersion is almost null except in recent years (from 50 Myrs).
-Finally, total inertia from RLQ analysis with tip distances of phylogenetic tree as matrix Q didn't differ from a null model were all phylogenetic tree tips were randomized ($p-value = 0.5344$ with 999 repetitions.)
-
-```{r PhyloMaker}
-library(phylosignal)
-library(phytools)
-library(phylobase)
-
-# Building species list
-l <- PFT_sp
-# setdiff(row.names(l), Species$LatinName)
-row.names(l)[which(row.names(l) == 'Actinodaphne tadulingami')] <- 'Actinodaphne tadulingamii'
-Species['agin',] <- c('agin', "Agrostistachys indica", 'Euphorbiaceae', NA, NA)
-Species['anme',] <- c('anme', "Antidesma menasu", 'Euphorbiaceae', NA, NA)
-Species['anmo',] <- c('anme', "Antidesma montanum", 'Euphorbiaceae', NA, NA)
-row.names(l)[which(row.names(l) == 'Blepharistemma membranifolia')] <- 'Blepharistemma serratum '
-row.names(l)[which(row.names(l) == 'Casearia ovata')] <- 'Casearia ovata '
-row.names(l)[which(row.names(l) == 'Cinnamomum malabathrum')] <- 'Cinnamomum malabatrum'
-row.names(l)[which(row.names(l) == 'Cinnamomum sp1')] <- 'Cinnamomum sp.'
-row.names(l)[which(row.names(l) == 'Dimocarpus longan')] <- 'Dimocarpus longan '
-row.names(l)[which(row.names(l) == 'Drypetes confertiflorus')] <- 'Drypetes confertiflora'
-row.names(l)[which(row.names(l) == 'Glochidion malabaricum')] <- 'Glochidion ellipticum'
-row.names(l)[which(row.names(l) == 'Homalium zeylanicum')] <- 'Homalium zeylanicum '
-row.names(l)[which(row.names(l) == 'Ixora nigricams')] <- 'Ixora nigricans  '
-row.names(l)[which(row.names(l) == 'Knema attenuata')] <- 'Knema attenuata  '
-row.names(l)[which(row.names(l) == 'Leptonychia moacurroides')] <- 'Leptonychia caudata '
-row.names(l)[which(row.names(l) == 'Myristica malabarica')] <- 'Myristica malabarica '
-row.names(l)[which(row.names(l) == 'Olea dioica')] <- 'Olea dioica '
-row.names(l)[which(row.names(l) == 'Reinwardtiodendron anaimalaiense')] <- 'Reinwardtiodendron anamalaiense'
-row.names(l)[which(row.names(l) == 'Strombosia ceylanica')] <- 'Strombosia ceylanica '
-row.names(l)[which(row.names(l) == 'Symplocos cochinensis')] <- 'Symplocos cochinchinensis'
-row.names(l)[which(row.names(l) == 'Syzygium laetum')] <- 'Syzygium laetum '
-row.names(l)[which(row.names(l) == 'Vepris bilocularis')] <- 'Vepris bilocularis '
-sp <- Species[which(Species$LatinName %in% row.names(l)), 2:3]
-sp$genus <- unlist(lapply(strsplit(sp$LatinName, ' ', fixed = T), function(x){x[1]}))
-sp$species <- sub(' ', '_', sp$LatinName, fixed = T)
-sp$family <- sp$Family
-sp <- sp[c('species', 'genus', 'family')]
-sp[which(sp$family == 'Flacourtiaceae'),3] <- 'Salicaceae'
-sp[which(sp$family == 'Steruliaceae'),3] <- 'Malvaceae'
-sp[which(sp$family == 'Cesaliniaceae'),3] <- 'Fabaaceae'
-sp['hepa','family'] <- 'Malvaceae'
-# sp <- sp[-which(row.names(sp) == 'hubr'),]
-# sp <- sp[-which(row.names(sp) %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
-
-# Getting phylogeny
-path <- system.file('phylogeny', package = 'Uppangala')
-source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
-phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
-nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
-results <- S.PhyloMaker(sp, phylo, nodes)
-tree <- results$Scenario.1
+```{r com ex}
+plot(raster(com$abundance, 'WD.cwm'))
 ```
 
-```{r Phylogenetic Tree, eval=FALSE, include=FALSE}
-# Completing the phylogenetic tree
-# Humboldtia: Close to Kingiodendron pinnatum according to Bruneau et al. 2001 Systematic Botany 26(3):487-514
-# Humboldtia and Kingiodendron are our only two Leguminosae
-# zoom(tree, 70:75)
-node <- which(tree$tip.label == "Kingiodendron_pinnatum")
-position<- 0.1*tree$edge.length[which(tree$edge[,2]==node)]
-tree <- bind.tip(tree,"Humboldtia_brunonis",where = node, position = position)
-# zoom(tree, 60:80)
+Figure 3: CWM for wood density in $20*20m^2$ communities
+
+## Structural equation modelling
+
+*Abiotic factors result of or influence functional traits ?*
+
+```{r SEM}
+library(lavaan)
+m1 <- '
+Abiotic =~ Slope + Curvature + Wetness + SW
+Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
+Biotic =~ BA + Canopy
+Traits ~ Abiotic + Biotic
+'
+m2 <- '
+Abiotic =~ Slope + Curvature + Wetness + SW
+Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
+Biotic =~ BA + Canopy
+Biotic ~ Traits + Abiotic
+'
+fit1 <- sem(m1, data = com$abundance@data)
+fit2 <- sem(m2, data = com$abundance@data)
+semPlot::semPaths(fit1,
+                  what = 'col',
+                  whatLabels = 'std',
+                  style = 'lisrel',
+                  residuals = T,
+                  nCharNodes = 0,
+                  nCharEdges = 0,
+                  groups='manlat',
+                  pastel = T,
+                  ask = F)
 ```
+
+Figure 4: Path diagram for structural equation modelling.
+
+```{r 6 lm & null models}
+reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
+LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
+                                                 SLA.cwm = lm(formula(paste('SLA.cwm', reg))), 
+                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
+                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
+                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
+                                                 LA.cwm = lm(formula(paste('LA.cwm', reg))),
+                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
+                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
+                                                 SLA.cwv = lm(formula(paste('SLA.cwv', reg))), 
+                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
+                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
+                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
+                                                 LA.cwv = lm(formula(paste('LA.cwv', reg))),
+                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
+LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
+Fval <- lapply(LM, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
+Fval <- lapply(Fval, function(x){do.call('rbind.data.frame', x)})
+Fval <- data.frame(Fval)
+names(Fval) <- names(LM)
+
+n = 999
+nPFT <- rep(list(PFT_sp), n)
+nPFT <- lapply(nPFT, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
+nPFT <- lapply(nPFT, function(x){x$Sp_Code <- SpCode[row.names(x)] ; return(x)})
+nTrees <- rep(list(Trees[which(names(Trees) %in% c('SpCode', 'com', '2013_girth'))]), n)
+nTrees <- lapply(nTrees, function(x){x$ID <- row.names(x) ; return(x)})
+nTrees <- mapply(function(x,y){merge(x, y, by.x = 'SpCode', by.y = 'Sp_Code')}, x = nTrees, y = nPFT, SIMPLIFY = F)
+nTrees <- lapply(nTrees, function(x){row.names(x) <- x$ID ; return(x)})
+nTrees <- lapply(nTrees, function(x){x <- x[-which(names(x) == 'ID')] ; return(x)})
+ncom <- lapply(nTrees, function(x){CWT(x, com[[1]])})
+ncom <- lapply(ncom, `[`, names(ncom[[1]]))
+ncom <- apply(do.call(rbind, ncom), 2, as.list) 
+nLM <- rep(list(LM), n)
+nLM <- lapply(nLM, `[`, names(nLM[[1]]))
+nLM <- apply(do.call(rbind, nLM), 2, as.list)
+ncom <- lapply(ncom, function(c){lapply(c, function(x){x <- rep(list(x), 14) ; return(x)})})
+nLM <- mapply(function(x,y){
+  mapply(function(r,s){
+    mapply(function(p,q){
+      lm(p$call,q)
+    }, p = r, q = s, SIMPLIFY = F)
+  }, r = x, s = y, SIMPLIFY = F)
+}, x = nLM, y = ncom, SIMPLIFY = F)
+nFval <- lapply(nLM, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
+nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
+nFval <- lapply(nFval, function(f){lapply(f, unlist)})
+pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
+row.names(pval) <- row.names(Fval)
+```
+
+## Null model: traits randomization
+
+```{r null com ex}
+library(raster)
+comp <- stack(raster(com$abundance, 'WD.cwm'),raster(ncom$abundance[[100]][[1]], 'WD.cwm'))
+names(comp) <- c('Uppangala communities', 'Null communities')
+plot(comp)
+```
+
+Figure 5: CWM for wood density in $20*20m^2$ null communities
+
+## Null model: model comparison
+
+```{r null ex}
+hist(nFval$abundance$WD.cwm, col = 'lightgrey', border = 'white', freq = F, main = 'Null models distribution for WD CWM in abundance', xlab = ' ')
+lines(density(as.numeric(nFval$abundance$WD.cwm)))
+lines(rep(Fval['WD.cwm','abundance'],2), c(0, 0.13), col = 'red', lwd = 2)
+mtext(paste('p-value =', round(pval['WD.cwm','abundance'],3), '**'), col = 'red')
+```
+
+Figure 6: Null models example
+
+## CWM in abundance
+
+*Are functional composition of communities linked to the environment ?*
+
+```{r CWM ab}
+kable(regression_table(LM, reg, pval, weight = 'abundance', metric = '.cwm'))
+```
+
+## CWM in presence-absence
+
+*Are functional composition of communities linked to the environment ?*
+
+```{r CWM pa}
+kable(regression_table(LM, reg, pval, weight = 'presence-absence', metric = '.cwm'))
+```
+
+## Phylogenetic tree
+
+```{r tree}
+library(png)
+library(grid)
+img <- readPNG("RLQpres_files/figure-html/tree.png")
+ grid.raster(img)
+```
+
+## Phylosignal
+
+*Are functionnal traits phylogenetically conserved ?*
 
@@
 kable(phyloS)
+```
+
+Table 7: Phylogenetic tree autoccorelation with Moran's I.
+
+## Phylosignal variograms
+
+*Are functionnal traits phylogenetically conserved ?*
+
+```{r phylosignal plots}
+par(mfrow = c(2,3))
+invisible(lapply(phyCor, function(x){plot(x, main = x$trait)}))
+```
+
+## Phylosignal variograms
+
+```{r lipa plot}
+phylosignal::barplot.phylo4d(p4d$full, bar.col=(lipa$full$p.value < 0.05) + 1)
 ```
@@
 
-## Are communities spatially structured ?
+## Functional spatial structure
+
+*Are functional composition of communities spatially structured ?*
 
@@
 
+Table 9: Functional spatial structure with Hardy's I Ust and TAUst.
+
+## Phylogenetic spatial structure
+
+*Are phylogenetic structure composition spatially structured ?*
+
 ```{r Phylogenetic spatial structure}
@@
 
-# Discussion
+Table 10: Phylogenetic spatial structure with Hardy's I Bst and PIst.
 
-# *Acknowledgement*
+## Discussion: Question 1
 
-# Supplementary materials
+- *Are functional composition of communities linked to the environment ?*
+    - **WD** increases with **SW** oriented and **concave** areas for **abundant**
+    - **SLA** increases with **SW** oriented, **dry** and **convex** areas for **abundant**
+    - **LA** decreases with **SW** oriented and **concave** areas for **rare**
+    - **WES** increases with **SW** oriented areas for **rare**
+    - Globally **environment filtering** is **weak** at this scale
+    - *Do covariation in functional traits illustrate ecological strategy axes ?*
+        - **Wood and leaf economic spectra**
+    - *Are functional traits phylogenetically conserved ?*
+        - **Short significant phylogenetic signal except for LDMC**
 
-## S2: Phylogenetic signal among functionnal traits by correlogram
+## Discussion: Question 1
 
-Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits
+- *Are functional composition of communities linked to the environment ?*
+    - **Abundant** species on **SW** slopes have a **denser wood** increasing with **concave** surfaces and a **higher SLA** increasing with **convex** surfaces
+    - **Rare** species on **SW** slopes have **smaller leaves** reducing with **concave** surfaces and **acquisitive wood strategy**
+    - Globally **environment filtering** is **weak** at this scale
+    - *Do covariation in functional traits illustrate ecological strategy axes ?*
+        - **Wood and leaf economic spectra**
+    - *Are functional traits phylogenetically conserved ?*
+        - **Short significant** phylogenetic **signal** except for LDMC
 
-```{r phylosignal plots}
-par(mfrow = c(2,3))
-invisible(lapply(phyCor, function(x){plot(x, main = x$trait)}))
-```
+## Discussion: Question 2
 
-# References
+- *Are communities spatially structured ?*
+    - Globally **not**
+    - *Are functional composition of communities spatially structured ?*
+        - A small **divergence** for **abundant** species 
+    - *Are phylogenetic structure composition spatially structured ?*
+        - **Not** for abundant nor for rare species

### 2016-07-18 inst/doc/draft2.Rmd
v0.1.9018 Pres & Francois cleaning

--- 2016-07-08 inst/doc/RLQ.Rmd
@@
 ---
-title: 'Uppangala tree functionnal: RLQ draft'
+title: 'Uppangala tree functionnal: Article draft'
 author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
@@
 
+Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific.
+*Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
+See table 1 for abbreviation.*
+
+```{r 2 CV}
+PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
+row.names(PFT_ind) <- PFT_ind$Group.1
+SpT <- unique(PFT[1:2])
+Sp <- SpT$SP
+names(Sp) <- SpT$Tree
+SpCodeT <- unique(PFT[2:3])
+SpCode <- SpCodeT$Sp_Code
+names(SpCode) <- SpCodeT$SP
+CET <- unique(PFT[c(1,4)])
+CE <- CET$CE
+names(CE) <- CET$Tree
+PFT_ind$SP <- Sp[row.names(PFT_ind)]
+PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
+PFT_ind$CE <- CE[row.names(PFT_ind)]
+PFT_ind <- PFT_ind[-c(1:2)]
+cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
+cv$Individual <- unlist(lapply(
+  as.list(row.names(cv)),
+  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
+cv$Species <- unlist(lapply(
+  as.list(row.names(cv)),
+  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
+cv$Community <- unlist(
+  lapply(as.list(row.names(cv)),
+         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
+             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
+cv <- round(cv, 2)
+cv['WD','Individual'] <- ''
+kable(cv)
+```
+
+Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
+Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$).
+
 ### Environment definition
@@
 
-## Are functionnal composition of communities linked to the environment ?
-
-Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific.
-*Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
-See table 1 for abbreviation.*
-
-```{r 2 CV}
-PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
-row.names(PFT_ind) <- PFT_ind$Group.1
-SpT <- unique(PFT[1:2])
-Sp <- SpT$SP
-names(Sp) <- SpT$Tree
-SpCodeT <- unique(PFT[2:3])
-SpCode <- SpCodeT$Sp_Code
-names(SpCode) <- SpCodeT$SP
-CET <- unique(PFT[c(1,4)])
-CE <- CET$CE
-names(CE) <- CET$Tree
-PFT_ind$SP <- Sp[row.names(PFT_ind)]
-PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
-PFT_ind$CE <- CE[row.names(PFT_ind)]
-PFT_ind <- PFT_ind[-c(1:2)]
-cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
-cv$Individual <- unlist(lapply(
-  as.list(row.names(cv)),
-  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
-cv$Species <- unlist(lapply(
-  as.list(row.names(cv)),
-  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
-cv$Community <- unlist(
-  lapply(as.list(row.names(cv)),
-         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
-             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
-cv <- round(cv, 2)
-cv['WD','Individual'] <- ''
-kable(cv)
-```
-
-Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
-Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$).
+## Traits and environment
 
@@
 
-```{r 3 RLQ}
+```{r 3 PCA PFT}
 library(vegan)
-library(ade4)
-library(raster)
 
-# Data
-Trees <- genTrees()
+PFT_ind$LA <- log(PFT_ind$LA)
+PFT_ind$SLA <- log(PFT_ind$SLA)
+PFT_ind$ID <- row.names(PFT_ind)
+row.names(PFT_ind) <- PFT_ind$ID
+PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
 Species <- genSpecies()
-Env <- genEnv()
-com <- quadrats()
-com@data$id <- paste0('C', com@data$id)
-
-# L matrix
-Trees$SpCode <- tolower(Trees$SpCode)
-Trees$ID <- row.names(Trees)
-row.names(Trees) <- Trees$ID
-Trees <- Trees[-which(names(Trees) == 'ID')]
-Trees$com <- NA
-XY <- data.frame(X = Trees$X, Y = Trees$Y)
-XY <- XY[-which(is.na(XY$X)),] # Removing NA
-coordinates(XY) <- ~ X + Y
-proj4string(XY) <- crs(com)
-Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-
-# R matrix
-com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'cosaspect'))
-com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
-names(com@data)[5] <- 'SW'
-com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
-Rm <- com@data[-1]
-row.names(Rm) <- com@data$id
-
-# Q matrix
 PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
@@
 PFT_sp <- PFT_sp[-c(1:2,4)]
-PFT_sp$SLA <- log(PFT_sp$SLA)
-PFT_sp$LA <- log(PFT_sp$LA)
-Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
-row.names(Qm) <- PFT_sp$Sp_Code
+pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_sp, cor = T)
 
-# Matrices adjustments
-Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
-Qm <- Qm[-which(is.na(Qm$WD)),]
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm <- Lm / rowSums(Lm)
-
-# RLQ
-L <- dudi.coa(Lm, scannf = F)
-R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
-Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
-rlq <- rlq(R, L, Q, scannf = F)
-
-# Q plot
-plot(t(t(Q$li)/Q$eig[1:2]),pch=16, col = 'lightgrey',
-     xlab = paste('Axe 1 -', round(Q$eig[1]/sum(Q$eig)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(Q$eig[2]/sum(Q$eig)*100, 2), '%'))
+lambda <- pca$sdev * sqrt(nrow(pca$scores))
+pca.var = apply(pca$score, 2, var)
+PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
+names(PCAscores) <- c('WES', 'LES')
+PCAscores$ID <- row.names(PCAscores)
+plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
+     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
+     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
+     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
+# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
 par(new=T)
-XLIM <- c(-max(abs(Q$co[,1])),max(abs(Q$co[,1])))
+Rot <- t(t(pca$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
 XLIM <- XLIM+(XLIM*0.2)
-plot(Q$co,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
 lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
 lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
-# legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
-       # text.col = c("green", "firebrick"))
+legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
+       text.col = c("green", "firebrick"))
+
+arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
+text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
 Hmisc::subplot(function(){
-  b <- barplot((Q$eig / sum(Q$eig)), las = 2, ylim = c(0,range(bstick(7))[2]),
+  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
           cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(Q$eig)), col = 'red') ;
+  lines(b, bstick(length(pca.var)), col = 'red') ;
   legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
          box.col = 'white')},
-  x=grconvertX(c(0.65,0.95), from='npc'),
-  y=grconvertY(c(0.01,0.26), from='npc')
+  x=grconvertX(c(0.75,1), from='npc'),
+  y=grconvertY(c(0.05,0.30), from='npc')
 )
@@
 axis (4)
-# arrows (rep(0,nrow(R$co)),rep(0,nrow(R$co)),R$co[,1],R$co[,2],lwd=1, col = 'blue')
-# text(R$co, pos = c(4, 2, 4, 2, 3), rownames(R$co),cex=0.9,col='blue')
-arrows (rep(0,nrow(Q$co)),rep(0,nrow(Q$co)),Q$co[,1],Q$co[,2],lwd=2)
-text (Q$co, pos = c(2, 2, 2, 4, 4), rownames(Q$co),cex=1.2)
+
+PCAscores <- as.data.frame(pca$scores)[1:2]
+names(PCAscores) <- c('WES', 'LES')
+PCAscores$ID <- row.names(PCAscores)
+PFT_sp$ID <- row.names(PFT_sp)
+PFT_sp <- merge(PFT_sp, PCAscores, all = T)
+row.names(PFT_sp) <- PFT_sp$ID
+PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
 ```
 
-Figure 1: Principal component analysis (PCA) performed on individual trait means.
+Figure 1: Principal component analysis (PCA) performed on species trait means.
 *Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
@@
 
-Whereas total co-inertia from RLQ analysis results was significanlty different for the model *m2*, it didn't differ from null model *m4* in model *m6* proposed by @ter_braak_improved_2012.
-So because both test should be rejected to reject model *m6* hypothesis, therefore the null hypothesis *H0* is not rejected.
+```{r 4 env-com}
+library(raster)
+Env <- genEnv()
+Env$Rivers <- readAll(Env$Rivers)
+Trees <- genTrees()
+Trees$SpCode <- tolower(Trees$SpCode)
+Trees$ID <- row.names(Trees)
+Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
+row.names(Trees) <- Trees$ID
+Trees <- Trees[-which(names(Trees) == 'ID')]
+com <- quadrats()
+Trees$com <- NA
+XY <- data.frame(X = Trees$X, Y = Trees$Y)
+XY <- XY[-which(is.na(XY$X)),] # Removing NA
+coordinates(XY) <- ~ X + Y
+proj4string(XY) <- crs(com)
+Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
+Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
+com@data$id <- paste0('C', com@data$id)
+com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
+names(Deriv)[5] <- 'SW'
+com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
+com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
+com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
+rm(Deriv, XY)
+pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
+lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
+pca.env.var = apply(pca.env$score, 2, var)
+plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
+     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
+Hmisc::subplot(function(){
+  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
+          cex.names = 0.7, axes = T, names.arg = NA) ;
+  lines(b, bstick(length(pca.env.var)), col = 'red') ;
+  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
+         box.col = 'white')},
+  x=grconvertX(c(0.07,0.37), from='npc'),
+  y=grconvertY(c(0.025,0.30), from='npc')
+)
+par(new=T)
+Rot <- t(t(pca.env$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, lwd = 2)
+lines(c(0,0), XLIM, lty = 2, lwd = 2)
+col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
+arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
+        Rot[,1],Rot[,2],lwd=2,col=col)
+legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
+       text.col = col, cex = 1.2)
+text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
+axis (3)
+axis (4)
+```
 
-Still fourth-corner analysis with adjusted p-values or not show significant trait-environment association at a level $\alpha = 0.05$ (see table 7).
-WD is negatively associated with southwesterness for all L matrices and positively associated with canopy height in presence absence.
-LDMC and LA are respectively negatively and positively associated with canopy height in relative abundance.
+Figure 2: Principal component analysis (PCA) performed on topgraphical variables.
+*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
+SW stand for southwesterness.*
 
-Table 7: Summary of fourth-corner analysis.
-*Result of Monte-Carlo test between fourth-corner result and 999 nul models following the model m6 suggested by @ter_braak_improved_2012.* _L matrices are filled with relative abundance (AB), presence-absence (PA) or relative basal area (BA)).
-The 5 functionnal traits are tested with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-See table 1 for trait abbreviations.
+## Community analysis
+
+Spatial variograms of CWM and CWV didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$. Functionnal diversity indices show also a plateau around $70 m$.
+
+```{r 5 Com, fig.height=4}
+com <- CWT(Trees, com)
+```
+
+```{r 6 lm & null models}
+reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
+LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
+                                                 SLA.cwm = lm(formula(paste('SLA.cwm', reg))), 
+                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
+                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
+                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
+                                                 LA.cwm = lm(formula(paste('LA.cwm', reg))),
+                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
+                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
+                                                 SLA.cwv = lm(formula(paste('SLA.cwv', reg))), 
+                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
+                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
+                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
+                                                 LA.cwv = lm(formula(paste('LA.cwv', reg))),
+                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
+LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
+Fval <- lapply(LM, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
+Fval <- lapply(Fval, function(x){do.call('rbind.data.frame', x)})
+Fval <- data.frame(Fval)
+names(Fval) <- names(LM)
+
+n = 999
+nPFT <- rep(list(PFT_sp), n)
+nPFT <- lapply(nPFT, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
+nPFT <- lapply(nPFT, function(x){x$Sp_Code <- SpCode[row.names(x)] ; return(x)})
+nTrees <- rep(list(Trees[which(names(Trees) %in% c('SpCode', 'com', '2013_girth'))]), n)
+nTrees <- lapply(nTrees, function(x){x$ID <- row.names(x) ; return(x)})
+nTrees <- mapply(function(x,y){merge(x, y, by.x = 'SpCode', by.y = 'Sp_Code')}, x = nTrees, y = nPFT, SIMPLIFY = F)
+nTrees <- lapply(nTrees, function(x){row.names(x) <- x$ID ; return(x)})
+nTrees <- lapply(nTrees, function(x){x <- x[-which(names(x) == 'ID')] ; return(x)})
+ncom <- lapply(nTrees, function(x){CWT(x, com[[1]])})
+ncom <- lapply(ncom, `[`, names(ncom[[1]]))
+ncom <- apply(do.call(rbind, ncom), 2, as.list) 
+nLM <- rep(list(LM), n)
+nLM <- lapply(nLM, `[`, names(nLM[[1]]))
+nLM <- apply(do.call(rbind, nLM), 2, as.list)
+ncom <- lapply(ncom, function(c){lapply(c, function(x){x <- rep(list(x), 14) ; return(x)})})
+nLM <- mapply(function(x,y){
+  mapply(function(r,s){
+    mapply(function(p,q){
+      lm(p$call,q)
+    }, p = r, q = s, SIMPLIFY = F)
+  }, r = x, s = y, SIMPLIFY = F)
+}, x = nLM, y = ncom, SIMPLIFY = F)
+nFval <- lapply(nLM, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
+nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
+nFval <- lapply(nFval, function(f){lapply(f, unlist)})
+pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
+row.names(pval) <- row.names(Fval)
+```
+
+Table 4: Summary for linear models and null models for traits community weighted means.
+_Linear model results for each trait community weighted means (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+See table 1 for abbreviations.
 The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
-```{r FQR}
-library(ade4)
-rd <- randtest(rlq, modeltype = 6)
-
-# L matrices list
-Lm.l <- list(abundance = Lm)
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm[Lm > 0] <-  1
-Lm <- Lm / rowSums(Lm)
-Lm.l$`presence-absence` = Lm
-Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm <- Lm / rowSums(Lm)
-Lm.l$`basal area` = Lm
-
-# Full # -WD
-RLQm.l <- list(full = list(Rm.l = rep(list(Rm),3), Lm.l = Lm.l, Qm.l = rep(list(Qm),3)))
-Qm <- PFT_sp[c('SLA', 'LDMC', 'Thick', 'LA')]
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-row.names(Qm) <- PFT_sp$Sp_Code
-Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm <- Lm / rowSums(Lm)
-Lm.l <- list(abundance = Lm)
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm[Lm > 0] <-  1
-Lm <- Lm / rowSums(Lm)
-Lm.l$`presence-absence` = Lm
-Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm <- Lm[row.names(Rm),]
-Lm <- Lm[intersect(row.names(Qm), names(Lm))]
-Lm <- Lm / rowSums(Lm)
-Lm.l$`basal area` = Lm
-RLQm.l$`-WD` <- list(Rm.l = rep(list(Rm),3), Lm.l = Lm.l, Qm.l = rep(list(Qm),3))
-# str(RLQm.l,3)
-
-# FQR
-fqr <- lapply(RLQm.l, function(x){mapply(function(r,l,q){fourthcorner(r, l, q, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")}, r = x$Rm.l, l = x$Lm.l, q = x$Qm.l, SIMPLIFY = F)})
-names(fqr$full) <- names(Lm.l) ; names(fqr$`-WD`) <- names(Lm.l)
-
-# Result
-# par(mfrow = c(2,3))
-# lapply(fqr, function(x){lapply(x, plot)})
-fqr.df.l <- mapply(function(y,z){
-  fqr.df <- with(y$tabG, cbind(data.frame(t(matrix(unlist(strsplit(names, ' / ')), nrow = 2)), stringsAsFactors = F), obs, pvalue))
-  fqr.wd <- with(z$tabG, cbind(data.frame(t(matrix(unlist(strsplit(names, ' / ')), nrow = 2)), stringsAsFactors = F), obs, pvalue))
-  fqr.df <- rbind(fqr.df, fqr.wd[which(fqr.wd$X2 == 'WD'),])
-  names(fqr.df)[1:2] <- c('Factor', 'Trait') 
-  fqr.df$res <- paste(round(fqr.df$obs, 3), '/', round(fqr.df$pvalue, 3), unlist(lapply(as.list(fqr.df$pvalue), stars, ns = ' ')))
-  fqr.df <- tapply(fqr.df$res, INDEX = list(fqr.df$Trait, fqr.df$Factor), paste, simplify = F)
-  return(fqr.df)
-  }, y = fqr$`-WD`, z = fqr$full, SIMPLIFY = F)
-```
-```{r FQR abundance}
-kable(t(fqr.df.l$abundance))
+```{r CWM}
+kable(regression_table(LM, reg, pval, weight = 'abundance', metric = '.cwm'))
+kable(regression_table(LM, reg, pval, weight = 'presence-absence', metric = '.cwm'))
 ```
 
-## Are phylogenetic composition of communities linked to the environment ?
+Table 5: Summary for linear models and null models for traits community weighted variances.
+_Linear model results for each trait community weighted variances (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
-Only Thick, LDMC, WES and at a fewer level LES shown a phylogenetic signal positively structured in the phylogenetic tree at a phylogenetic distance between 0 and 100 (see supplementary material S3).
-$\beta st$ and $\pi st$ indicates both for respectivelly abundance and presence absence data that our phylogeny is overdispersed inside communities: species within communities are less phylogenetically related on average than are specis from distinct communities.
-But compared to the model null *1s* suggested by @hardy_characterizing_2007 , the phylogenetic overdispersion is almost null except in recent years (from 50 Myrs).
-Finally, total inertia from RLQ analysis with tip distances of phylogenetic tree as matrix Q didn't differ from a null model were all phylogenetic tree tips were randomized ($p-value = 0.5344$ with 999 repetitions.)
+```{r CWV}
+kable(regression_table(LM, reg, pval, weight = 'abundance', metric = '.cwv'))
+kable(regression_table(LM, reg, pval, weight = 'presence-absence', metric = '.cwv'))
+```
+
+## Phylogenetic signal
 
@@
 
-```{r Phylogenetic Tree, eval=FALSE, include=FALSE}
-# Completing the phylogenetic tree
-# Humboldtia: Close to Kingiodendron pinnatum according to Bruneau et al. 2001 Systematic Botany 26(3):487-514
-# Humboldtia and Kingiodendron are our only two Leguminosae
-# zoom(tree, 70:75)
-node <- which(tree$tip.label == "Kingiodendron_pinnatum")
-position<- 0.1*tree$edge.length[which(tree$edge[,2]==node)]
-tree <- bind.tip(tree,"Humboldtia_brunonis",where = node, position = position)
-# zoom(tree, 60:80)
-```
-
 ```{r Phylosignal}
@@
 
-```{r Phylo RLQ, fig.height=4}
-library(adephylo)
-library(spacodiR)
-
-# RLQ
-for(i in 1:2) phy[[i]]$tip.label <- gsub('_', ' ', phy[[i]]$tip.label, fixed = T)
-for(i in 1:2) phy[[i]]$tip.label <- as.character(l$Sp_Code[match(phy[[i]]$tip.label, row.names(l))])
-Lm <- lapply(phy, function(x){Lm[intersect(names(Lm), x$tip.label)]})
-L <- lapply(Lm, dudi.coa, scannf = F)
-R <- mapply(function(x,y){dudi.pca(x, scannf = F, row.w = y$lw)}, x = rep(list(Rm), 2), y = L, SIMPLIFY = F)
-phyRLQ <- mapply(function(x,y){drop.tip(x,setdiff(x$tip.label, names(y)))}, x = phy, y = Lm, SIMPLIFY = F)
-Qd <- lapply(phyRLQ, distTips)
-Q <- mapply(function(x,y){dudi.pco(x, scannf = F, row.w = y$cw)}, x = Qd, y = L, SIMPLIFY = F)
-rlq <- mapply(function(x,y,z){rlq(x, y, z, scannf = F)}, x = R, y = L, z = Q, SIMPLIFY = F)
-
-# Random RLQs
-phyRD <- lapply(phy, function(x){x$node.label <- NULL ; return(x)})
-phyRD <- lapply(phyRD, function(x){rep(list(x), 999)})
-phyRD <- lapply(phyRD, function(x){lapply(x, resamp.phy)})
-LmRD <- lapply(Lm, function(x){rep(list(x), 999)})
-LmRD <- mapply(function(x,y){mapply(function(r,s){r[intersect(names(r), s$tip.label)]}, r = x, s = y, SIMPLIFY = F)}, x = LmRD, y = phyRD, SIMPLIFY = F)
-phyRD <- mapply(function(x,y){mapply(function(r,s){drop.tip(r, setdiff(r$tip.label, names(s)))}, r = x, s= y, SIMPLIFY = F)}, x = phyRD, y = LmRD, SIMPLIFY = F)
-QdRD <- lapply(phyRD, function(x){lapply(x, distTips)})
-LRD <- lapply(LmRD, function(x){lapply(x, dudi.coa, scannf = F)})
-RmRD <- lapply(rep(list(full = Rm, `-WD` = Rm)), function(x){rep(list(x), 999)})
-RRD <- mapply(function(x,y){mapply(function(r,s){dudi.pca(r, scannf = F, row.w = s$lw)}, r = x, s = y, SIMPLIFY = F)}, x = RmRD, y = LRD, SIMPLIFY = F)
-QRD <- mapply(function(x,y){mapply(function(r,s){dudi.pco(r, scannf = F, row.w = s$cw)}, r = x, s = y, SIMPLIFY = F)}, x = QdRD, y = LRD, SIMPLIFY = F)
-rlqRD <- mapply(function(x,y,z){mapply(function(r,s,t){rlq(r, s, t, scannf = F)}, r = x, s = y, t = z, SIMPLIFY = F)}, x = RRD, y = LRD, z = QRD, SIMPLIFY = F)
-
-# Randtest
-pval <- mapply(function(x,y){rank(c(sum(x$eig), unlist(lapply(y, function(z){sum(z$eig)}))))[1]/1000}, x = rlq, y = rlqRD)
-par(mfrow = c(1,2))  
-invisible(mapply(function(x,y,z,l){
-  hist(unlist(lapply(y, function(r){sum(r$eig)})), main = paste('Model', l, ':', 'pval =', round(z, 3)), xlab = 'Coinertia', col = 'lightgrey', border = 'white')
-  lines(c(sum(x$eig), sum(x$eig)), c(0,500), col = 'red', lwd = 2)
-}, x = rlq, y = rlqRD, z = as.list(pval), l = as.list(names(rlqRD))))
-```
-
 ## Are communities spatially structured ?
 
-```{r Spatial structure}
+```{r Functionnal spatial structure}
 library(spacodiR)
@@
 
-# Phylogeny structure
+pvalues <- as.data.frame(pvalues)[3:4]
+pvalues$Ust <-  1- pvalues$Ust
+spafun.res <- format(spafun[3:4], scientific = T, digits = 2)
+pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
+spafun.res$Ust <- paste(spafun.res$Ust, '/', pvalues$Ust)
+spafun.res$TAUst <- paste(spafun.res$TAUst, '/', pvalues$TAUst)
+kable(spafun.res)
+```
+
+```{r Phylogenetic spatial structure}
+for(i in 1:2) phy[[i]]$tip.label <- gsub('_', ' ', phy[[i]]$tip.label, fixed = T)
+for(i in 1:2) phy[[i]]$tip.label <- as.character(l$Sp_Code[match(phy[[i]]$tip.label, row.names(l))])
+Lm <- lapply(phy, function(x){Lm[intersect(names(Lm), x$tip.label)]})
+phyRD <- lapply(phy, function(x){x$node.label <- NULL ; return(x)})
+phyRD <- lapply(phyRD, function(x){rep(list(x), 999)})
+phyRD <- lapply(phyRD, function(x){lapply(x, resamp.phy)})
+LmRD <- lapply(Lm, function(x){rep(list(x), 999)})
+LmRD <- mapply(function(x,y){mapply(function(r,s){r[intersect(names(r), s$tip.label)]}, r = x, s = y, SIMPLIFY = F)}, x = LmRD, y = phyRD, SIMPLIFY = F)
+phyRD <- mapply(function(x,y){mapply(function(r,s){drop.tip(r, setdiff(r$tip.label, names(s)))}, r = x, s= y, SIMPLIFY = F)}, x = phyRD, y = LmRD, SIMPLIFY = F)
+
 spaphy <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), phy = x)}, x = phy, y = Lm.l, SIMPLIFY = F)
 spaphy <- do.call('rbind.data.frame', lapply(spaphy, function(x){x[1:4]}))
+
+spaphy.l <- mapply(function(x,y){mapply(function(r,s){spacodi.calc(t(r), s)}, r = x, s = y, SIMPLIFY = F)},x = LmRD, y = phyRD, SIMPLIFY = F)
+spaphy.l <- lapply(spaphy.l, function(x){lapply(x, function(y){y[1:4]})})
+spaphyRD <- list(Ist = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Ist}))}),
+                 Tst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Tst}))}),
+                 Bst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Bst}))}),
+                 PIst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$PIst}))}))
+pvalues <- mapply(function(obs,rd){
+  mapply(function(o,r){rank(c(o, r))[1]/1000}, o = obs, r = rd)
+  }, obs = lapply(as.list(spaphy), as.list), rd = spaphyRD)
+row.names(pvalues) <- row.names(spaphy)
+
+pvalues <- as.data.frame(pvalues)[3:4]
+pvalues$Bst <-  rep(NaN,2)
+spaphy.res <- format(spaphy[3:4], scientific = T, digits = 2)
+pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
+spaphy.res$Bst <- paste(spaphy.res$Bst, '/', pvalues$Bst)
+spaphy.res$PIst <- paste(spaphy.res$PIst, '/', pvalues$PIst)
+kable(spaphy.res)
+
 # permut <- list(Bst = mapply(function(x,y){spacodi.by.nodes(sp.plot = t(y), x, sp.parm = "Bst", method = "1s")}, x = phy, y = Lm.l, SIMPLIFY = F), PIst = mapply(function(x,y){spacodi.by.nodes(sp.plot = t(y), x, sp.parm = "PIst", method = "1s")}, x = phy, y = Lm.l, SIMPLIFY = F))
@@
 
-## S1: Fourth corner analysis results in presence-absence and relative basal area
+## S1: Structural equation model
 
-```{r FQR others}
-kable(t(fqr.df.l$`presence-absence`))
-kable(t(fqr.df.l$`basal area`))
+```{r SEM}
+library(lavaan)
+m1 <- '
+Abiotic =~ Slope + Curvature + Wetness + SW
+Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
+Biotic =~ BA + Canopy
+Traits ~ Abiotic + Biotic
+'
+m2 <- '
+Abiotic =~ Slope + Curvature + Wetness + SW
+Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
+Biotic =~ BA + Canopy
+Biotic ~ Traits + Abiotic
+'
+fit1 <- sem(m1, data = com$abundance@data)
+fit2 <- sem(m2, data = com$abundance@data)
+semPlot::semPaths(fit1,
+                  what = 'col',
+                  whatLabels = 'std',
+                  style = 'lisrel',
+                  residuals = T,
+                  nCharNodes = 0,
+                  nCharEdges = 0,
+                  groups='manlat',
+                  pastel = T,
+                  ask = F)
 ```
 
-## S2: Phylogenetic signal among functionnal traits by correlogram
+Supplementary material S1: Path diagram for structural equation modelling.
+*Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic.
+Arrow values are indicating standardized parameter estimates*.
 
-Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits
+## S3: Phylogenetic signal among functionnal traits by correlogram
 
-```{r phylosignal plots}
-par(mfrow = c(2,3))
-invisible(lapply(phyCor, function(x){plot(x, main = x$trait)}))
+Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits.
+```{r phylosignal plots, eval=FALSE, include=FALSE}
+par(mfrow = c(2,4))
+for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
 ```

### 2016-07-18 R/CWT.R

@@ -26,13 +26,12 @@ NULL
 CWT <- function(Trees, com,
                 traits = c('LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'),
                 metric = c('mean', 'variance'),
-                weights = c('abundance', 'presence-absence', 'basal area'),
+                weights = c('abundance', 'presence-absence'),
                 env = c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')){
 
   weights <- unlist(lapply(as.list(weights), function(x){switch (x,
                                                                  'abundance' = 'ab',
-                                                                 'presence-absence' = 'pa',
-                                                                 'basal area' = 'ba')}))
+                                                                 'presence-absence' = 'pa')}))
   com <- com[c('id', env)]
   cwt.l <- list()
 
@@ -43,44 +42,21 @@ CWT <- function(Trees, com,
     cwt.l[[i]] <- com
     names(cwt.l)[i] <- switch (wgt,
                                  'ab' = 'abundance',
-                                 'pa' = 'presence-absence',
-                                 'ba' = 'basal area'
-    )
+                                 'pa' = 'presence-absence')
 
     # Trees table
     trees <- switch (wgt,
                      'ab' = Trees,
-                     'pa' = Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),],
-                     'ba' = TreesBA <- Trees[-which(is.na(Trees$`2013_girth`)),]
-    )
+                     'pa' = Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),])
 
     for (m in metric) {
 
-      # Computing CWT
-      if(wgt %in% c('ab', 'pa')){
-
-        fun <- switch (m,
-                       'mean' = base::mean,
-                       'variance' = stats::var
-        )
-        cwt <- aggregate(trees, by = list(trees$com), fun, na.rm = T)
-        names(cwt)[1] <- 'id'
-        cwt <- cwt[which(names(cwt) %in% c('id', traits))]
-
-      } else if (wgt == 'ba'){
-
-        cwt <- switch (m,
-                       'mean' = data.frame(lapply(as.list(traits),
-                                                  function(t){ddply(trees, ~ com,
-                                                                    function(x){Hmisc::wtd.mean(x[,which(names(x) == t)], x$`2013_girth`/2*pi, na.rm = T)})[,2]})),
[…]

### 2016-07-18 R/extdata.R

@@ -10,6 +10,8 @@
 #' \item{DEM.tif}{Digital elevation model}
 #' \item{Girth_10Ha.csv}{DB extraction for 10 ha plots girth}
 #' \item{Girth_5Ha.csv}{DB extraction for 5 ha plots girth}
+#' \item{Height_5Ha.csv}{DB extraction for 5 ha plots height}
+#' \item{Rivers.tif}{Distance to the nearest potential river model}
 #' \item{Sp_trees_allPlot.csv}{DB extraction for tree species}
 #' \item{Species.txt}{DB extraction for species data}
 #' }

### 2016-07-18 R/null_model.R

@@ -1,99 +1,0 @@
-#' @importFrom FD dbFD
-NULL
-
-#' Null model
-#'
-#' Function to compute null models
-#'
-#' @param com list. List of communities
-#' @param PFT data.frame. PFT values by species
-#' @param formulas list. Model formulas
-#' @param SpCode dta.frame. Link between species nd they're code
-#' @param traits char. Choosen traits and strategy axes
-#' @param env char. Choosen environmental variables
-#' @param type char. Communti weighted trait metric
-#' @param n int. number of repetitions
-#' @param time logical. Use a timer
-#' @param verbose logical. Use a counter
-#'
-#' @return List of communities null models
-#' @export
-#'
-#' @examples
-#'
-null_model <- function(com,
-                       Trees,
-                       PFT,
-                       formulas,
-                       SpCode,
-                       traits = c('LES', 'SLA', 'WES','WD', 'LDMC', 'Thick', 'LA'),
-                       env = c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA'),
-                       type = c('CWM', 'CWV'),
-                       n = 999,
-                       time = T,
-                       verbose = T){
-
-  t0 <- Sys.time()
-  nstat <- list(abundance = matrix(ncol = length(formulas[[1]]), nrow = n, dimnames = list(1:n, names(formulas[[1]]))),
-                `presence-absence` = matrix(ncol = length(formulas[[2]]), nrow = n, dimnames = list(1:n, names(formulas[[2]]))),
-                `basal area` = matrix(ncol = length(formulas[[3]]), nrow = n, dimnames = list(1:n, names(formulas[[3]])))
-                )
-
-  # Null Communities
-  for(i in 1:n){
-    # Creating random PFT data
-    nPFT <- PFT
-    row.names(nPFT) <- sample(row.names(nPFT))
-    nPFT$Sp_Code <- SpCode[row.names(nPFT)]
-    # Merging data to trees
-    nTrees <- Trees[which(names(Trees) %in% c('SpCode', 'com', '2013_girth'))]
-    nTrees$ID <- row.names(Trees)
-    nTrees <- merge(nTrees, nPFT, by.x = 'SpCode', by.y = 'Sp_Code')
-    row.names(nTrees) <- nTrees$ID
-    nTrees <- nTrees[-which(names(nTrees) == 'ID')]
[…]

### 2016-07-18 R/regression_table.R

@@ -11,63 +11,39 @@
 #' @examples
 #' # No example yet
 #'
-regression_table <- function(LM, model, reg, select = '.cwm'){
+regression_table <- function(LM, reg, pval, weight = 'abundance', metric = '.cwm'){
 
   # Results subsetting
-  LM <- lapply(LM, function(x){x[grep(select, names(x), fixed = T)]})
-  model$formula <- lapply(model$formula, function(x){x[grep(select, names(x), fixed = T)]})
-  model$nstat <- lapply(model$nstat, function(x){x[,grep(select, colnames(x), fixed = T)]})
-  model$stat <- lapply(model$stat, function(x){x[grep(select, names(x), fixed = T)]})
-  model$p.value <- lapply(model$p.value, function(x){x[grep(select, names(x), fixed = T)]})
+  LM <- LM[[weight]]
+  LM <- LM[grep(metric, names(LM), fixed = T)]
+  pval <- as.data.frame(pval)[weight]
+  pval <- pval[grep(metric, row.names(pval), fixed = T),]
 
   # Regression coefficients
   var <- strsplit(substr(reg,3,nchar(reg)), ' + ', fixed = T)[[1]]
   names(var) <- var
-  var_table <- data.frame(matrix(nrow = length(LM[[1]])*3, ncol = (length(var)+2)))
-  names(var_table) <- c('Trait', toupper(substr(select,2,4)), var)
-  var_table$Trait[c(1,((1:length(LM[[1]]))*3+1)[-length(LM[[1]])])] <- names(LM[[1]])
-  var_table[,2] <- rep(c('AB', 'PA', 'BA'), length(LM[[1]]))
-  for(j in 1:length(LM)){
-    if(select == '.fd'){
-      j <-  1
-    }
-    for(i in 1:length(LM[[1]])){
-      for(v in var){
-        row <- which(var_table$Trait == names(LM[[1]])[i]) + j - 1
-        if(v %in% row.names(summary(LM[[j]][[i]])$coefficients)){
-          var_table[row,v] <- paste(format(summary(LM[[j]][[i]])$coefficients[v,1],
-                                           scientific = T, digits = 2),
-                                    stars(summary(LM[[j]][[i]])$coefficients[v,4]))
-        } else {
-          var_table[row,v] <- ''
-        }
[…]

### 2016-07-18 R/stars.R

@@ -10,6 +10,9 @@
 #' stars(0.049)
 #'
 stars = function(val, ns = 'n.s.'){
+  if(!is.numeric(val) || is.nan(val)){
+    val = 1
+  }
   if(val < 0.001){
     s = '*** '
   } else if (val < 0.01) {

### 2016-07-18 inst/doc/draft2.Rmd
v0.1.9019 Rm BIOMASS NAMESPACE & genTrees.R

@@
 LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
+# Global F
 Fval <- lapply(LM, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
@@
 names(Fval) <- names(LM)
+# Partial F
+# pFval <- lapply(LM, function(y){lapply(y, function(x){car::Anova(x)['F value']})})
 
@@
 nTrees <- mapply(function(x,y){merge(x, y, by.x = 'SpCode', by.y = 'Sp_Code')}, x = nTrees, y = nPFT, SIMPLIFY = F)
+rm(nPFT)
 nTrees <- lapply(nTrees, function(x){row.names(x) <- x$ID ; return(x)})
@@
 ncom <- lapply(nTrees, function(x){CWT(x, com[[1]])})
+rm(nTrees)
 ncom <- lapply(ncom, `[`, names(ncom[[1]]))
@@
 }, x = nLM, y = ncom, SIMPLIFY = F)
+rm(ncom) ; gc()
+
+# Global model
 nFval <- lapply(nLM, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
@@
 nFval <- lapply(nFval, function(f){lapply(f, unlist)})
+
+# nDev <- lapply(nLM, function(lm){lapply(lm, function(y){lapply(y, function(x){deviance(x)})})})
+# nDev <- lapply(nDev, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
+# nDev <- lapply(nDev, function(f){lapply(f, unlist)})
+# invisible(mapply(function(x,y,z){plot(x, y, xlab = 'Fval', ylab = 'Dev', main = z)}, x = nFval$abundance, y = nDev$abundance, z = as.list(names(nFval$abundance))))
+
 pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
 row.names(pval) <- row.names(Fval)
+
+# Partial contribution
+
 ```

### 2016-07-18 R/genTrees.R

@@ -1,6 +1,3 @@
-#' @importFrom BIOMASS modelHD predictHeight
-NULL
-
 #' Trees data generation
 #'
 #' Generates Trees data set from external data extracted from the database
@@ -54,17 +51,17 @@ genTrees <- function(){
                           substr(as.character(height5$Census_date), start = 7, stop = 11),
                           substr(as.character(height5$Census_date), start = 4, stop = 5),
                           sep = '_')
-  # Modelling height girth relation
-  IDs <- intersect(height5$ID.date, girth5$ID.date)
-  H <- data.frame(H = height5$Total_height, row.names = height5$ID.date)
-  H[H == -999] <- NA
-  H <- H[IDs,]
-  D <- data.frame(D = girth5$Girth/2*pi, ID = girth5$ID.date)
-  D <- D[which(!duplicated(D$ID)),]
-  row.names(D) <- D$ID
-  D[D == -999] <- NA
-  D <- D[IDs,1]
-  m <- modelHD(D, H, method = 'michaelis')
+  # # Modelling height girth relation
+  # IDs <- intersect(height5$ID.date, girth5$ID.date)
+  # H <- data.frame(H = height5$Total_height, row.names = height5$ID.date)
+  # H[H == -999] <- NA
+  # H <- H[IDs,]
+  # D <- data.frame(D = girth5$Girth/2*pi, ID = girth5$ID.date)
+  # D <- D[which(!duplicated(D$ID)),]
+  # row.names(D) <- D$ID
+  # D[D == -999] <- NA
+  # D <- D[IDs,1]
+  # m <- modelHD(D, H, method = 'michaelis')
   # Preparing girth
   girth5$ID <- girth5$Tree_5Ha_ID
   girth5$Date <- paste(substr(as.character(girth5$Census_date), start = 7, stop = 11),
@@ -97,7 +94,7 @@ genTrees <- function(){
   tree <- tree[-which(is.na(tree$SpCode)),]
   row.names(tree) <- tree$ID
   tree <- tree[-1]
-  tree$height <- predictHeight(tree$`2013_girth`, m)
+  # tree$height <- predictHeight(tree$`2013_girth`, m)
   summary(tree)
 
   return(tree)

### 2016-07-20 inst/doc/draft2.Rmd
v0.1.9020 Draft with CWM 7 CVNND

--- 2016-07-18 inst/doc/draft2.Rmd
@@
 ---
-title: 'Uppangala tree functionnal: Article draft'
+title: 'Uppangala tree functional traits: Article draft'
 author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
@@
 library(knitr)
+library(parallel)
 opts_chunk$set(
@@
 
+**To add:**
+
+* Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
+For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+* Leaf investment definition
+* Leaf ressource acuisition definition
+* Analyses scheme presentation:
+    + Are functionnal composition of communities linked to the environment ?
+        - Do covariations in functionnal traits illustrate ecological strategy axes ?
+        - Are functionnal traits phylogenetically conserved ?
+    + Are communities spatially structured ?
+        - Are functionnal composition of communities spatially structured ?
+        - Are phylogenetic composition of communities spatially structured ?
+
 # Material and methods
@@
 Climate and geology of the 10-ha plot can be considered as homogeneous.
-General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
+General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
 Each tree girth is monitored by a two-year census.
+**Girth monitoring to develop**
 
-## Sampling
+## Wood and leaf sampling
 
@@
 Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
-For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).
+For some individuals we were unable to collect an appropriate branch sample (58 over 267 individuals representing 12 species over 91).
 
@@
 Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias.
-Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues.
+The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues.
 Branch wood samples were sink in water to evaluate their volume by water displacement.
@@
 
-## Traits
+## Trait measurements
 
-With leaf and branch wood samples we calculated five functionnal traits: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the wood sample volume in $cm^3$ divided by the wood sample dry weight in $g$ (see table 1).
-Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
-For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+We calculated five functionnal traits of leaf and wood samples: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the the wood sample dry weight in $g$ divided by wood sample volume in $cm^3$ (see table 1).
 
@@
 kable(table,  label="traits", row.names = F)
+rm(table)
 ```
@@
 
-To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
+To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
 Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
 Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
-Distance to the nearest potential river was caluclated under ArcGis (cite).
+Distance to the nearest potential river was using ArcGis [@desktop_arcgis_2011].
+
+In order to avoid multi-colinearity in future analyses, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
+We reduced our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see supplementary materials S1).
 
 To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density.
-Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$.
-Both basal area and stem density are calculated with 2013 girth inventory from Uppangala plot on stem with a diameter above 10 centimeters at breast height ($DBH>10$).
+Canopy height is calculated from a digital canopy model built with a LIDAR campaign on Uppangala **date ?** plots with a resolution $1*1 m^2$.
+Both basal area and stem density are calculated from 2013 girth inventory of Uppangala plot on stemw with a diameter above 10 centimeters at breast height ($DBH>10$).
 Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$).
-So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
+So we removed stem density to avoid multicolinearity, and we kept canopy height that was not correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
 
@@
 
-### Covariation in traits
+*Title are just here to help remembering the structure but might be removed*
 
-Traits variation was first investigated in three levels with the coefficient of variation ($cv=sd/mean$) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
-For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
-To test the correlation and covariation among traits we performed a principal component analysis (PCA).
-We investigated peculiar traits covariation highlighted with the PCA with a Spearman's rank correlation test.
+### Functionnal composition of communities regarding environment
 
-Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific.
+#### Covariation in traits
+
+Traits variation was first investigated in three levels with the coefficient of variation ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
+Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
+Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
+
+Table 2: Coefficient of intra-individual, intra-specific and inter-specific trait variation.
 *Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
@@
 kable(cv)
+rm(PFT, CET, cv, CE, Sp, SpT)
 ```
 
-Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
-Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$).
+For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
+To test the correlation and covariation among traits we performed a principal component analysis (PCA).
 
-### Environment definition
-
-In order to avoid multi-colinearity in future analysis, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
-In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling.
-Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
-Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
-Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).
-
-### Community analysis
+#### Community analysis
 
 Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
-For all communities, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes.
-We used the following formula to compute each community weighted parameter (CWP):
-$$CWP = \sum_{k=1}^{nj} W_{k,j} * P_{k,j} ~ (1)$$ 
-where $nj$ is the number of species sampled in community $j$, $W_{k,j}$ is the weight of species $k$ in community $j$ used to compute CWP (relative abundance, presence-absence, or relative basal area), and $P_{k,j}$ is the parameter of species $k$ in community $j$ (mean or the variance).
-In addition 5 functionnal diversity indices (FD) weighted by abundances were computed: functional richness (FRic), functional evenness (FEve), functional divergence (FDiv) and Rao's quadratic entropy (Q) (@villeger_new_2008, @botta-dukat_raos_2005).
-Spatial variograms were used to investigate our CWM and CWV spatial organization.
-CWM, CWV, and FD spatial autocorrelation (SAC) was also assessed by Moran's I statistic.
+To investigate traits distribution inside communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010].
+Analyses were conducted on the five functionnal traits and the two economic spectra defined with the species scores obtained on the 2 first PCA axes.
 
-Links between both CWM and CWV with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
-Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables.
-A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix).
-Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance.
-Same procedure was also applied to functionnal diversity indices (FD) from our communities.
-Finally spatial autocorrelation (SAC) of residuals was investigated to assess selected models.
+CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
+abundance, $p_{i}$:
+$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
+CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution.
+We tested links between CWMs and environment with a first null model randomizing traits among species (permuting raw in `Trait*Species` matrix).
+**Precise what this model imply**.
+Linear regressions between CWMs and biotic and abiotic factors after step measure were compared to corresponding 999 null linear regressions from our null models with F-statistic.
 
-### Fourth-corner analysis
+CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
+abundance, $p_{i}$:
+$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
+We tested links between CWVs and environment with a second null model randomizing abundances among species inside each community but keeping the observed list of species and trait values inside communities.
+**Precise what this model imply**.
+The CWV standardized effect size (SES) was then computed comparing each community to their corresponding null communities with the formula :
+$$SES = \frac{(CWV_{obs} - CWV_{null})}{\sigma _{null}}$$
+where $CWV_{obs}$ is the observed CWV and $CWV_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distribution CWVs.
+Finally proportion of of CWVs differing from null expectaions was tested with a double tails test; and links between SES of CWVs and environment was investigated by linear regression.
 
-In addition of our community analysis by regression, we conducted both RLQ and fourth-corner analyses to test the effect of environments on Uppangala tree functionnal traits following @dray_testing_2008.
-Both matrices R (`Environment*Sites`) with selected features and Q (`Traits*Species`) were reduced by a principal component analysis.
-We used a Monte-Carlo test to compare fourth-corner and RLQ results to 999 nul models built with the model *m6* proposed by @ter_braak_improved_2012.
-Fourth corner analysis was tested with 3 types of L matrices ($Sites*Species$) with species relative abundance, presence absence or relative basal area in each sites; whereas RLQ analysis was only conducted with relative abundance.
+In parrallel to CWVs, eveness of communities was investigated with CVNND with the same null model randomizing abundances among species inside each community but keeping the observed list of species and trait values inside communities.
+SES were also calculated for CVNNDs.
+Finally proportion of of CVNNDs differing from null expectaions was tested with a double tails test; and links between SES of CVNNDs and environment was investigated by linear regression.
 
-### Phylogenetic signal
+#### Phylogenetic signal
 
 Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
-Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 functionnal traits, the two PCA axes and in the phylogenetic tree.
+Phylogenetic signal was globaly assessed for each functionnal trait among the phylogenetic tree by Moran's I statistic.
 Phylogenetic signal structure was assessed by a tree tip distances correlogram.
-In addition, $\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate phylogenetic signal spatial organization with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
-Finally phylogenetic association to habitat was also studied by an RLQ analysis with tree tip distances as the Q matrix reduced by a principal coordinate analysis.
-RLQ result with phylogenetic distance matrix was compared to 999 nul models were phylogenetic tree tips were randomized.
 
-All analysis were conducted in the R 3.3.0  free software environment for statistical computing and graphics [@r_core_team_r:_2016] using the pacakge 
+### Spatial structure of community compositions
+
+$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions.
+Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
+In parrallel spatial structure of community functionnal composition was assessed by $U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007.
+Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix).
+
+All analysis were conducted in the R 3.3.0  free software environment for statistical analyses and graphics [@r_core_team_r:_2016] using the package 
 `ade4` [@dray_ade4_2007], 
-`lavaan` [@rosseel_lavaan:_2012], 
-`FD` [@laliberte_distance-based_2010], 
-`gstat` [@pebesma_multivariable_2004], 
 `ape` [@paradis_ape:_2004], 
@@
 
-## Traits and environment
+## Functionnal composition of communities regarding environment
+
+### Traits and environment
 
@@
 Our results integrate less traits (5 against 14), and less species (92 against 758), but still capture the same strategy axes.
-
-Figure 1 shows WD orthogonal to SLA following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$).
-LA is almost opposed to WD and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007.
-Finally, we found the positive relation suggested by @diaz_global_2015 between LDMC and WD (Spearman's $rho = 0.24$, $p-value < 0.001$).
 
@@
 PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
+Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')]
+row.names(Q) <- PFT_sp$Sp_Code
+rm(PCAscores, PFT_ind, SpCodeT, lambda, pca, pca.var, Rot, XLIM)
 ```
@@
 See table 1 for abbreviation.*
-
-Multi-colinearity was avoided by reducing our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see figure 2).
-Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
-Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
-Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary.
-Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)
 
@@
 com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
-rm(Deriv, XY)
 pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
-lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
-pca.env.var = apply(pca.env$score, 2, var)
-plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
-     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
-Hmisc::subplot(function(){
-  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
-          cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(pca.env.var)), col = 'red') ;
-  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
-         box.col = 'white')},
-  x=grconvertX(c(0.07,0.37), from='npc'),
-  y=grconvertY(c(0.025,0.30), from='npc')
-)
-par(new=T)
-Rot <- t(t(pca.env$loadings)*lambda)
-XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
-XLIM <- XLIM+(XLIM*0.2)
-plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
-lines(XLIM, c(0,0), lty = 2, lwd = 2)
-lines(c(0,0), XLIM, lty = 2, lwd = 2)
-col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
-arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
-        Rot[,1],Rot[,2],lwd=2,col=col)
-legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
-       text.col = col, cex = 1.2)
-text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
-axis (3)
-axis (4)
+R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+row.names(R) <- com@data$id
+L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+L[is.na(L)] <- 0
+sel <- intersect(row.names(Q), names(L))
+Q <- Q[sel,]
+L <- L[sel]
+R <- R[row.names(L),]
+L <- list(abundance = L, `presence-absence` = L)
+L$`presence-absence`[L$`presence-absence` > 1]  <- 1
+rm(Deriv, XY, com, Env, Trees, sel, PFT_sp)
 ```
 
-Figure 2: Principal component analysis (PCA) performed on topgraphical variables.
-*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
-SW stand for southwesterness.*
+### Community analysis
 
-## Community analysis
+Table 4: Summary of linear models and null models for traits community weighted means.
+_Linear model results for each trait community weighted means by species relative abundance after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model global and partial F-statistics were compared to 999 null model F-statistics in which traits were randomized among species.
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
-Spatial variograms of CWM and CWV didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$. Functionnal diversity indices show also a plateau around $70 m$.
+**Second table are the result in presence-absence. They are juste here for discussion. But they probably won't be in the final report under this shape.**
 
-```{r 5 Com, fig.height=4}
-com <- CWT(Trees, com)
-```
-
-```{r 6 lm & null models}
-reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
-LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
-                                                 SLA.cwm = lm(formula(paste('SLA.cwm', reg))), 
-                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
-                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
-                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
-                                                 LA.cwm = lm(formula(paste('LA.cwm', reg))),
-                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
-                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
-                                                 SLA.cwv = lm(formula(paste('SLA.cwv', reg))), 
-                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
-                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
-                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
-                                                 LA.cwv = lm(formula(paste('LA.cwv', reg))),
-                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
-LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
-Fval <- lapply(LM, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
+```{r CWM}
+# Observed CWMs
+CWM <- lapply(L, function(l){lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){apply(l, 1, function(x){weighted.mean(Q[,v], x, na.rm=T)})})})
+CWM <- lapply(CWM, function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})
+## Regressions
+reg <- paste('~', paste(names(R), collapse = ' + '))
+CWM.lm <- lapply(CWM, function(x){lapply(x, function(y){lm(y ~., R)})})
+CWM.lm <- lapply(CWM.lm, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
+## Global F
+Fval <- lapply(CWM.lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
 Fval <- lapply(Fval, function(x){do.call('rbind.data.frame', x)})
 Fval <- data.frame(Fval)
-names(Fval) <- names(LM)
+names(Fval) <- names(CWM.lm)
+## Partial F
+pFval <- lapply(CWM.lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})
 
+# Simulated CWMs
 n = 999
-nPFT <- rep(list(PFT_sp), n)
-nPFT <- lapply(nPFT, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
-nPFT <- lapply(nPFT, function(x){x$Sp_Code <- SpCode[row.names(x)] ; return(x)})
-nTrees <- rep(list(Trees[which(names(Trees) %in% c('SpCode', 'com', '2013_girth'))]), n)
-nTrees <- lapply(nTrees, function(x){x$ID <- row.names(x) ; return(x)})
-nTrees <- mapply(function(x,y){merge(x, y, by.x = 'SpCode', by.y = 'Sp_Code')}, x = nTrees, y = nPFT, SIMPLIFY = F)
-nTrees <- lapply(nTrees, function(x){row.names(x) <- x$ID ; return(x)})
-nTrees <- lapply(nTrees, function(x){x <- x[-which(names(x) == 'ID')] ; return(x)})
-ncom <- lapply(nTrees, function(x){CWT(x, com[[1]])})
-ncom <- lapply(ncom, `[`, names(ncom[[1]]))
-ncom <- apply(do.call(rbind, ncom), 2, as.list) 
-nLM <- rep(list(LM), n)
-nLM <- lapply(nLM, `[`, names(nLM[[1]]))
-nLM <- apply(do.call(rbind, nLM), 2, as.list)
-ncom <- lapply(ncom, function(c){lapply(c, function(x){x <- rep(list(x), 14) ; return(x)})})
-nLM <- mapply(function(x,y){
+nQ <- rep(list(Q), n)
+nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
+nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
+cl <- makeCluster(3)
+clusterExport(cl, list('L', 'nQ'))
+nCWM <- clusterApply(cl, L, function(l){
+  lapply(nQ, function(q){
+    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){
+      apply(l, 1, function(x){
+        weighted.mean(q[,v], x, na.rm=T)
+      })
+    })
+  })
+})
+stopCluster(cl)
+rm(nQ, cl)
+nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})})
+## Regressions
+nCWM.lm <- mapply(function(x,y){
+  lapply(y, function(z){
     mapply(function(r,s){
-    mapply(function(p,q){
-      lm(p$call,q)
-    }, p = r, q = s, SIMPLIFY = F)
-  }, r = x, s = y, SIMPLIFY = F)
-}, x = nLM, y = ncom, SIMPLIFY = F)
-nFval <- lapply(nLM, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
+      lm(formula(paste0('s ~ ', as.character(r$terms)[3])), cbind(s,R))
+    }, r = x, s = z, SIMPLIFY = F)
+  })
+}, x = CWM.lm, y = nCWM, SIMPLIFY = F)
+rm(nCWM) ; invisible(gc())
+## Global F
+nFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
 nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
@@
 row.names(pval) <- row.names(Fval)
+rm(Fval, nFval)
+# Partial F
+npFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})})
+npFval <- lapply(npFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
+npFval <- lapply(npFval, function(f){lapply(f, function(x){do.call('cbind', x)})})
+ppval <- mapply(function(x,y){mapply(function(r,s){mapply(function(p,q){rank(c(p,q))[1]/(n+1)}, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))}, r = x, s = y)}, x = pFval, y = npFval, SIMPLIFY = F)
+ppval <- lapply(ppval, function(y){lapply(y, function(x){data.frame(t(x))})})
+ppval <- lapply(ppval, function(y){data.frame(data.table::rbindlist(y, fill = T))})
+ppval <- lapply(ppval, function(y){row.names(y) <- names(CWM.lm$abundance) ; return(y)})
+ppval$abundance <- ppval$abundance[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy')]
+ppval$abundance$BA <- NA
+ppval$`presence-absence` <- ppval$`presence-absence`[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+ppval <- lapply(ppval, function(x){1 - x})
+rm(pFval, npFval)
+
+# Result tables
+kable(regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]]))
+kable(regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]]))
+rm(CWM.lm, nCWM.lm, pval, ppval, reg) ; invisible(gc())
 ```
 
-Table 4: Summary for linear models and null models for traits community weighted means.
-_Linear model results for each trait community weighted means (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+Table 5: Summary of linear models for traits standardized effect size of community weighted variances.
+_Linear model results for each trait standardized effect size of community weighted variances species relative abundance after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
 See table 1 for abbreviations.
@@
 
-```{r CWM}
-kable(regression_table(LM, reg, pval, weight = 'abundance', metric = '.cwm'))
-kable(regression_table(LM, reg, pval, weight = 'presence-absence', metric = '.cwm'))
+**First output is the result of two tails test on communities CWVs for each traits. They are juste here for discussion. But they probably won't be in the final report under this shape.**
+
+```{r CWV, fig.height=4}
+# Observed CWVs
+CWV <- lapply(names(CWM$abundance), function(v){apply(cbind(data.frame(CWM$abundance[[which(names(CWM$abundance) == v)]]),L$abundance), 1, function(x){sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))})})
+names(CWV) <- names(CWM$abundance)
+
+# Simulated CWVs 
+nL <- rep(list(L$abundance), n)
+nL <- lapply(nL,function(l){apply(l, 1, function(x){x[x != 0] <- sample(x[x != 0]) ; return(x)})}) 
+nL <- lapply(nL, function(x){t(x)})
+nL <- lapply(nL, data.frame)
+
+cl <- makeCluster(3)
+clusterExport(cl, list('nL', 'CWM', 'Q'))
+nCWM.cwv <- clusterApply(cl, nL, function(l){
+  lapply(as.list(names(CWM$abundance)), function(v){
+    apply(l, 1, function(x){
+      weighted.mean(Q[,v], x, na.rm=T)
+    })
+  })
+})
+stopCluster(cl)
+nCWM.cwv <- lapply(nCWM.cwv, function(x){names(x) <- names(CWM$abundance) ; return(x)})
+nCWV <- mapply(function(l,cwm){
+  lapply(names(CWM$abundance), function(v){
+    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
+      sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))
+    })
+  })
+}, l = nL, cwm = nCWM.cwv, SIMPLIFY = F)
+nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
+rm(nCWM.cwv, cl)
+# ## Regressions
+# nCWV.lm <- lapply(nCWV, function(cwv){
+#   mapply(function(x,y){
+#     lm(formula(paste0('y ~ ', as.character(x$terms)[3])), cbind(y,R))
+#   }, x = CWV.lm, y = cwv, SIMPLIFY = F)
+# })
+## Non null CWVs
+nCWV <- lapply(nCWV, `[`, names(nCWV[[1]]))
+nCWV <- apply(do.call(rbind, nCWV), 2, as.list)
+nCWV <- lapply(nCWV, function(x){do.call('cbind', x)})
+nCWV <- lapply(nCWV, data.frame)
+CWV.pval <- data.frame(mapply(function(x,y){
+  mapply(function(p,q){
+    rank(c(p,q))[1]/(n+1)
+  }, p = as.list(x), q = as.list(data.frame(t(y))))
+}, x = CWV, y = nCWV))
+CWV.pval.res <- data.frame(inf = apply(CWV.pval, 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(CWV.pval, 2, function(x){length(which(x > 0.975))/length(x)}))
+## CWV SES
+CWV.ses <- data.frame(mapply(function(x,y){
+  mapply(function(p,q){
+    (p - mean(q))/sd(q)
+  }, p = as.list(x), q = as.list(data.frame(t(y))))
+}, x = CWV, y = nCWV))
+CWV.ses.lm <- apply(CWV.ses, 2, function(y){lm(y ~., R)})
+CWV.ses.lm <- lapply(CWV.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
+CWV.ses.res <- lapply(CWV.ses.lm, function(x){summary(x)$coefficients})
+CWV.ses.res <- lapply(CWV.ses.res, data.frame)
+CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
+CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
+CWV.ses.res <- CWV.ses.res[-1]
+CWV.ses.res$Curvature <- NA
+CWV.ses.res <- CWV.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
+row.names(CWV.ses.res) <- names(CWV.ses)
+CWV.ses.res[is.na(CWV.ses.res)] <- ''
+CWV.ses.res <- data.frame(CWV.ses.res)
+CWV.ses.res$R2 <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)
+
+# Results
+barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
+kable(CWV.ses.res)
+rm(CWV.pval, CWV.pval.res, CWV.ses.res, CWV.ses.lm, nCWV)
 ```
 
-Table 5: Summary for linear models and null models for traits community weighted variances.
-_Linear model results for each trait community weighted variances (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model deviance was compared to 999 null model deviances in which traits were randomized among species.
+Table 6: Summary of linear models for traits standardized effect size of coefficient of variation of nearest neighbour distances.
+_Linear model results for each trait standardized effect size of coefficient of variation of nearest neighbour distances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
 See table 1 for abbreviations.
@@
 
-```{r CWV}
-kable(regression_table(LM, reg, pval, weight = 'abundance', metric = '.cwv'))
-kable(regression_table(LM, reg, pval, weight = 'presence-absence', metric = '.cwv'))
+**First output is the result of two tails test on communities CWVs for each traits. They are juste here for discussion. But they probably won't be in the final report under this shape.**
+
+```{r CVNND, fig.height=4}
+# Observed CVNND
+CVNND <- apply(L$abundance, 1, function(l){apply(Q, 2, function(q){cati::CVNND(rep(q, l), na.rm = T)})})
+
+# Simulated CVNNDs 
+cl <- makeCluster(3)
+clusterExport(cl, list('nL', 'Q'))
+nCVNND <- clusterApply(cl, nL, function(x){
+  apply(x, 1, function(l){
+    apply(Q, 2, function(q){
+      cati::CVNND(rep(q,l), na.rm = T)
+    })
+  })
+})
+stopCluster(cl) ; rm(cl)
+rm(nL) ; invisible(gc)
+## Non null CVNNDs
+nCVNND <- lapply(nCVNND, function(x){as.list(data.frame(t(x)))})
+nCVNND <- lapply(nCVNND, function(x){lapply(x, function(y){names(y) <- colnames(CVNND) ; return(y)})})
+nCVNND <- lapply(nCVNND, `[`, names(nCVNND[[1]]))
+nCVNND <- apply(do.call(rbind, nCVNND), 2, as.list)
+nCVNND <- lapply(nCVNND, function(x){do.call('cbind', x)})
+nCVNND <- lapply(nCVNND, data.frame)
+CVNND.pval <- data.frame(mapply(function(x,y){
+  mapply(function(p,q){
+    rank(c(p,q))[1]/(n+1)
+  }, p = as.list(x), q = as.list(data.frame(t(y))))
+}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
+row.names(CVNND.pval) <- colnames(CVNND)
+CVNND.pval.res <- data.frame(inf = apply(CVNND.pval, 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(CVNND.pval, 2, function(x){length(which(x > 0.975))/length(x)}))
+## CVNND SES
+CVNND.ses <- data.frame(mapply(function(x,y){
+  mapply(function(p,q){
+    (p - mean(q))/sd(q)
+  }, p = as.list(x), q = as.list(data.frame(t(y))))
+}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
+row.names(CVNND.ses) <- colnames(CVNND)
+CVNND.ses.lm <- apply(CVNND.ses, 2, function(y){lm(y ~., R)})
+CVNND.ses.lm <- lapply(CVNND.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
+CVNND.ses.res <- lapply(CVNND.ses.lm, function(x){summary(x)$coefficients})
+CVNND.ses.res <- lapply(CVNND.ses.res, data.frame)
+CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
+CVNND.ses.res <- data.frame(data.table::rbindlist(lapply(CVNND.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
+CVNND.ses.res <- CVNND.ses.res[-1]
+CVNND.ses.res$Curvature <- NA
+CVNND.ses.res <- CVNND.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+CVNND.ses.res <- apply(CVNND.ses.res, 2, as.character)
+row.names(CVNND.ses.res) <- names(CVNND.ses)
+CVNND.ses.res[is.na(CVNND.ses.res)] <- ''
+CVNND.ses.res <- data.frame(CVNND.ses.res)
+CVNND.ses.res$R2 <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)
+
+# Results
+barplot(t(CVNND.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
+kable(CVNND.ses.res)
+rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
 ```
 
-## Phylogenetic signal
+### Phylogenetic signal
+
+**Artefacts of phylogenetic tree building to be removed**
 
@@
 
+Table 7: Summary of Moran’s I statistic and p-values for phylogenetic signal of functionnal traits.
+_See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
 ```{r Phylosignal}
@@
 
-## Are communities spatially structured ?
+## Spatial structure of community compositions
+
+Spatial structure of both community phylogenetic and functionnal communities don't show significant organization, no convergence nor divergence ($U_{st} = -1.1*10^-3$ with $p-value = 0.114$, $\tau _{st} = 1.8*10^-3$ with $p-value = 0.079$, $\beta _{st} = -8.3*10^-4$ with $p-value = NaN$, and $\pi _{st} = -1.8*10^-3$ with $p-value = 0.247$).
+
+**Following tables will probably be removed in final report by simply including results in text. They are juste here for discussion.**
+
+Table 8: Summary of spatial structure of community functionnal compositions.
+_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions.
+Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 ```
+
+Table 9: Summary of spatial structure of community phylogenetic compositions.
+_$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions.
+Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix).
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 
-*I'm aware that Supplementary materials goes after References (as figure and tables), but it's not possible for the moment with Rmarkdown and knitr.*
+**I'm aware that Supplementary materials goes after References (as figure and tables), but it's not possible for the moment with Rmarkdown and knitr.**
 
-## S1: Structural equation model
+## S1: Principal component analysis performed on topgraphical variables
+
+Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
+Elevation and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
+
+```{r PCA Env}
+lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
+pca.env.var = apply(pca.env$score, 2, var)
+plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
+     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
+Hmisc::subplot(function(){
+  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
+          cex.names = 0.7, axes = T, names.arg = NA) ;
+  lines(b, bstick(length(pca.env.var)), col = 'red') ;
+  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
+         box.col = 'white')},
+  x=grconvertX(c(0.07,0.37), from='npc'),
+  y=grconvertY(c(0.025,0.30), from='npc')
+)
+par(new=T)
+Rot <- t(t(pca.env$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, lwd = 2)
+lines(c(0,0), XLIM, lty = 2, lwd = 2)
+col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
+arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
+        Rot[,1],Rot[,2],lwd=2,col=col)
+legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
+       text.col = col, cex = 1.2)
+text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
+axis (3)
+axis (4)
+```
+
+Supplementary material S1: Principal component analysis (PCA) performed on topgraphical variables.
+*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
+SW stand for southwesterness.*
+
+## S2: Structural equation model
+
+In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling.
+Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
+Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
+Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).
+
+Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary.
+Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$).
 
@@
 
-Supplementary material S1: Path diagram for structural equation modelling.
+Supplementary material S2: Path diagram for structural equation modelling.
 *Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic.
@@
 
-Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits.
-```{r phylosignal plots, eval=FALSE, include=FALSE}
+```{r phylosignal plots}
 par(mfrow = c(2,4))
@@
 
+Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits.
+
 # References

### 2016-07-20 R/regression_table.R

@@ -11,13 +11,7 @@
 #' @examples
 #' # No example yet
 #'
-regression_table <- function(LM, reg, pval, weight = 'abundance', metric = '.cwm'){
-
-  # Results subsetting
-  LM <- LM[[weight]]
-  LM <- LM[grep(metric, names(LM), fixed = T)]
-  pval <- as.data.frame(pval)[weight]
-  pval <- pval[grep(metric, row.names(pval), fixed = T),]
+regression_table <- function(LM, reg, pval, ppval){
 
   # Regression coefficients
   var <- strsplit(substr(reg,3,nchar(reg)), ' + ', fixed = T)[[1]]
@@ -29,21 +23,19 @@ regression_table <- function(LM, reg, pval, weight = 'abundance', metric = '.cwm
     for(v in var){
       if(v %in% row.names(summary(LM[[i]])$coefficients)){
         var_table[i,v] <- paste(format(summary(LM[[i]])$coefficients[v,1], scientific = T, digits = 2),
-                                  stars(summary(LM[[i]])$coefficients[v,4]))
+                                  stars(ppval[i,v], ns = '    '))
       } else {
         var_table[i,v] <- ''
       }
     }
   }
-  var_table$Trait <- gsub(metric, '', var_table$Trait, fixed = T)[1:7]
-
   # LM R squarred
   var_table$R2 <- unlist(lapply(LM, function(x){round(summary(x)$r.squared,2)}))
 
   # Null models
   var_table$`Null model` <- round(pval,3)
   var_table$` ` <- unlist(lapply(pval, stars, ns = ' '))
-  # var_table[which(var_table$` ` == ' '),2:7] <- ' '
+  var_table[which(var_table$` ` == ' '),2:7] <- ' '
 
   return(var_table)
 }

### 2016-07-22 inst/doc/draft2.Rmd
v0.1.9021 Pre tree data changes

@@
 
-Keywords: functionnal trait, tropical evergreen forest, Western Ghats, India...
+Keywords: functional trait, tropical evergreen forest, Western Ghats, India...
 
@@
 
+**Study scheme:**
+
+1. Do variations and covariations in functionnal traits reveal species ecological strategies ?
+    i. Do functionnal traits vary among species ?
+    ii. Do covariations in functional traits represent axes of ecological strategy ?
+2. Do functional composition reflect the influence of abiotic and biotic drivers of community dynamics ?
+    i. locally
+    ii. between communities
+
 **To add:**
 
-* Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
-For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+* Functional traits can be linked to trade-offs in tree strategy for functionning, survival and reproduction [@baraloto_decoupled_2010, and see table 1], which vary depending on environment condition.
+For instance, SLA varies along a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
 * Leaf investment definition
-* Leaf ressource acuisition definition
-* Analyses scheme presentation:
-    + Are functionnal composition of communities linked to the environment ?
-        - Do covariations in functionnal traits illustrate ecological strategy axes ?
-        - Are functionnal traits phylogenetically conserved ?
-    + Are communities spatially structured ?
-        - Are functionnal composition of communities spatially structured ?
-        - Are phylogenetic composition of communities spatially structured ?
+* Leaf ressource acquisition definition
 
@@
 Each tree girth is monitored by a two-year census.
-**Girth monitoring to develop**
+**Girth monitoring to develop shortly**
 
@@
 
-We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates.
+We collected trait data on 91 over 117 (78%) species.
+We sampled five individuals per species for most abundant species and we completed with less abundant and rare species with one to five individuals.
 Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals.
@@
 Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
-For some individuals we were unable to collect an appropriate branch sample (58 over 267 individuals representing 12 species over 91).
+For some individuals we were unable to collect an appropriate branch sample (58 over 267 individuals representing 12 missing species over 91).
+
+## Trait measurements
 
@@
 Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time.
-Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias.
+Once dried, the leaves were weighted directly after been extracted from the oven to avoid humidity bias.
 The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues.
@@
 
-## Trait measurements
-
-We calculated five functionnal traits of leaf and wood samples: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the the wood sample dry weight in $g$ divided by wood sample volume in $cm^3$ (see table 1).
+We calculated five functional traits of leaf and wood samples: leaf thickness (Thick in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1).
 
 Table 1: Functional traits measured.
-*Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development from @baraloto_decoupled_2010.*
+*Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*
 
@@
 
-To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
-Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
+We selected 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
+Digital elevation model is issued from a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
 Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
-Distance to the nearest potential river was using ArcGis [@desktop_arcgis_2011].
+Distance to the nearest potential river was computed using ArcGis [@desktop_arcgis_2011].
 
-In order to avoid multi-colinearity in future analyses, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
-We reduced our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see supplementary materials S1).
+In order to avoid multi-colinearity in future analyses, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning.
+We reduced our explanatory variables to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness to represent monsoon winds [@bunyan_effect_2015] and slope.
 
-To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density.
-Canopy height is calculated from a digital canopy model built with a LIDAR campaign on Uppangala **date ?** plots with a resolution $1*1 m^2$.
-Both basal area and stem density are calculated from 2013 girth inventory of Uppangala plot on stemw with a diameter above 10 centimeters at breast height ($DBH>10$).
-Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$).
-So we removed stem density to avoid multicolinearity, and we kept canopy height that was not correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
+We selected two uncorrelated biotic factors: canopy height and basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
+Canopy height was calculated from a digital canopy model built with the 2013 LIDAR campaign on Uppangala plots with resolution $1*1 m^2$.
+Basal area was calculated from 2013 girth inventory of Uppangala plot on stems with a diameter above 10 centimeters at breast height ($DBH>10~cm$).
+<!-- Use of canopy height and basal area as biotic factors of our 5 functional traits was justified by structural equation modelling (see supplementary material S2). -->
 
@@
 
-*Title are just here to help remembering the structure but might be removed*
+### Traits variation and covariation
 
-### Functionnal composition of communities regarding environment
+Traits variation was first investigated in three levels with the coefficient of variation ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species (inter-specific).
 
-#### Covariation in traits
+For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
+To test the correlation and covariation among traits we performed a principal component analysis (PCA).
 
-Traits variation was first investigated in three levels with the coefficient of variation ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
-Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
+### Functional composition of communities
+
+Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
+To test for pattern in functional composition of communities we used null model approach.
+A first null model (NM1) randomizes traits among species (permuting raw in `Trait*Species` matrix).
+The second model (NM2) randomizes abundances among species inside each community but keeps the observed list of species and trait values inside communities.
+NM1 allow to assess environmental filtering and biotic process on Uppangala plot species.
+NM2 bring more insight on environmental and biotic effects on species distribution inside communities [@bernard-verdier_community_2012].
+
+To investigate traits distribution of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010].
+CWV lower than expected under a null model will be evidence for convergence of traits inside the community whereas higher CWV will show divergence in trait distribution.
+CVNND is used to detect niche differenciation patterns in case of a lower CVNND as observed under a null model.
+
+CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
+abundance, $p_{i}$:
+$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
+CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution.
+CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
+abundance, $p_{i}$:
+$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
+CVNND was computed in abundance inside each community. For both CWV and CVNND, we calculated the standardized effect size (SES) as: 
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}}$$
+where $I_{obs}$ is the observed metric and $I_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.
+
+Linear regressions between CWMs and biotic and abiotic factors after step measure were compared to corresponding 999 null linear regressions from NM1 with F-statistic.
+For both CWV and CVNND, proportion of of communities differing from null expectaions was tested with a two-tailed test; and links between SES and environment was investigated by linear regression.
+Analyses were conducted on the five functional traits and the two economic spectra defined with the species scores obtained on the 2 first PCA axes.
+
+### Composition variation between communities
+
+Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
+Phylogenetic signal was globaly assessed for each functional trait among the phylogenetic tree by Moran's I statistic.
+Positive Moran's I value will indicate convergence of trait value inside the phylogeny whereas negative Moran's I will show divergence.
+
+$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities.
+Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
+In parrallel variation of functional composition between communities was assessed by $U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007.
+Observed values were compared to NM1.
+Negative Hardy's metrics significate that species are functionaly or phylogenetically
+
+All analysis were conducted in the R 3.3.0  free software environment for statistical analyses and graphics [@r_core_team_r:_2016] using the package 
+`ade4` [@dray_ade4_2007], 
+`ape` [@paradis_ape:_2004], 
+`phylosignal` [@keck_phylosignal:_2015] 
+and `spacodiR` [@eastman_spacodir:_2013].
+
+# Results
+
+## Traits variation and covariation
+
+Trait coefficients of variation analysis reveals a general increase of functional trait variation among scales (from intra-individual to inter-specific, see table 2).
 Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
@@
 ```
-
-For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
-To test the correlation and covariation among traits we performed a principal component analysis (PCA).
-
-#### Community analysis
-
-Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
-To investigate traits distribution inside communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010].
-Analyses were conducted on the five functionnal traits and the two economic spectra defined with the species scores obtained on the 2 first PCA axes.
-
-CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
-abundance, $p_{i}$:
-$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
-CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution.
-We tested links between CWMs and environment with a first null model randomizing traits among species (permuting raw in `Trait*Species` matrix).
-**Precise what this model imply**.
-Linear regressions between CWMs and biotic and abiotic factors after step measure were compared to corresponding 999 null linear regressions from our null models with F-statistic.
-
-CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
-abundance, $p_{i}$:
-$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
-We tested links between CWVs and environment with a second null model randomizing abundances among species inside each community but keeping the observed list of species and trait values inside communities.
-**Precise what this model imply**.
-The CWV standardized effect size (SES) was then computed comparing each community to their corresponding null communities with the formula :
-$$SES = \frac{(CWV_{obs} - CWV_{null})}{\sigma _{null}}$$
-where $CWV_{obs}$ is the observed CWV and $CWV_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distribution CWVs.
-Finally proportion of of CWVs differing from null expectaions was tested with a double tails test; and links between SES of CWVs and environment was investigated by linear regression.
-
-In parrallel to CWVs, eveness of communities was investigated with CVNND with the same null model randomizing abundances among species inside each community but keeping the observed list of species and trait values inside communities.
-SES were also calculated for CVNNDs.
-Finally proportion of of CVNNDs differing from null expectaions was tested with a double tails test; and links between SES of CVNNDs and environment was investigated by linear regression.
-
-#### Phylogenetic signal
-
-Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
-Phylogenetic signal was globaly assessed for each functionnal trait among the phylogenetic tree by Moran's I statistic.
-Phylogenetic signal structure was assessed by a tree tip distances correlogram.
-
-### Spatial structure of community compositions
-
-$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions.
-Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
-In parrallel spatial structure of community functionnal composition was assessed by $U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007.
-Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix).
-
-All analysis were conducted in the R 3.3.0  free software environment for statistical analyses and graphics [@r_core_team_r:_2016] using the package 
-`ade4` [@dray_ade4_2007], 
-`ape` [@paradis_ape:_2004], 
-`phylosignal` [@keck_phylosignal:_2015] 
-and `spacodiR` [@eastman_spacodir:_2013].
-
-# Results
-
-## Functionnal composition of communities regarding environment
-
-### Traits and environment
 
@@
 com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
+com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})[,2]
 com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
@@
 L$`presence-absence`[L$`presence-absence` > 1]  <- 1
-rm(Deriv, XY, com, Env, Trees, sel, PFT_sp)
+# rm(Deriv, XY, com, Env, Trees, sel, PFT_sp)
 ```
 
-### Community analysis
+## Functional composition of communities
 
@@
 stopCluster(cl)
-rm(nQ, cl)
+rm(cl)
 nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})})
@@
 nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
-rm(nCWM.cwv, cl)
-# ## Regressions
-# nCWV.lm <- lapply(nCWV, function(cwv){
-#   mapply(function(x,y){
-#     lm(formula(paste0('y ~ ', as.character(x$terms)[3])), cbind(y,R))
-#   }, x = CWV.lm, y = cwv, SIMPLIFY = F)
-# })
+rm(nCWM.cwv, cl, nL) ; invisible(gc())
 ## Non null CWVs
@@
 cl <- makeCluster(3)
-clusterExport(cl, list('nL', 'Q'))
-nCVNND <- clusterApply(cl, nL, function(x){
-  apply(x, 1, function(l){
-    apply(Q, 2, function(q){
+clusterExport(cl, list('nQ', 'L'))
+nCVNND <- clusterApply(cl, nQ, function(x){
+  apply(L$abundance, 1, function(l){
+    apply(x, 2, function(q){
       cati::CVNND(rep(q,l), na.rm = T)
@@
 stopCluster(cl) ; rm(cl)
-rm(nL) ; invisible(gc)
+rm(nQ) ; invisible(gc)
 ## Non null CVNNDs
@@
 CVNND.ses.res <- CVNND.ses.res[-1]
-CVNND.ses.res$Curvature <- NA
+CVNND.ses.res$Slope <- NA
+CVNND.ses.res$BA <- NA
 CVNND.ses.res <- CVNND.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
@@
 
-### Phylogenetic signal
+## Composition variation between communities
 
@@
 
-Table 7: Summary of Moran’s I statistic and p-values for phylogenetic signal of functionnal traits.
+Table 7: Summary of Moran’s I statistic and p-values for phylogenetic signal of functional traits.
 _See table 1 for abbreviations.
@@
 
-## Spatial structure of community compositions
+Variations of both functional and phylogenetic compositions between communities don't significantly differ from null models ($U_{st} = -1.1*10^-3$ with $p-value = 0.114$, $\tau _{st} = 1.8*10^-3$ with $p-value = 0.079$, $\beta _{st} = -8.3*10^-4$ with $p-value = NaN$, and $\pi _{st} = -1.8*10^-3$ with $p-value = 0.247$).
 
-Spatial structure of both community phylogenetic and functionnal communities don't show significant organization, no convergence nor divergence ($U_{st} = -1.1*10^-3$ with $p-value = 0.114$, $\tau _{st} = 1.8*10^-3$ with $p-value = 0.079$, $\beta _{st} = -8.3*10^-4$ with $p-value = NaN$, and $\pi _{st} = -1.8*10^-3$ with $p-value = 0.247$).
+**Following tables will probably be removed in final report by simply including results in text. They are just here for discussion.**
 
-**Following tables will probably be removed in final report by simply including results in text. They are juste here for discussion.**
-
-Table 8: Summary of spatial structure of community functionnal compositions.
-_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions.
+Table 8: Summary of variation of functional composition between communities.
+_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities.
 Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008.
@@
 
-Table 9: Summary of spatial structure of community phylogenetic compositions.
-_$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate spatial structure of community phylogenetic compositions.
+Table 9: Summary of variation of phylogenetic composition between communities.
+_$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities.
 Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix).
@@
 # Discussion
+
+## Strategy axes
+## Community assembly
+## Process scales
+## Limits and perspectives
 
@@
 
-In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling.
-Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
-Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
+In order to see if or if environmental biotic factors were the results of functional traits we used structural equation modelling.
+Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *functional traits* defined with our 5 functional traits.
+Two models were compared, a first one with the functional traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functional traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
 Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).
 
-Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary.
+Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functional traits and not the contrary.
 Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$).
@@
 
-## S3: Phylogenetic signal among functionnal traits by correlogram
+## S3: Phylogenetic signal among functional traits by correlogram
 
@@
 
-Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits.
+Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functional traits.
 

### 2016-07-22 inst/doc/draft.Rmd
v0.1.9022 New data, BA corrected & Mantel beginning

--- 2016-07-01 inst/doc/draft.Rmd
@@
 library(knitr)
+library(parallel)
+library(Uppangala)
 opts_chunk$set(
@@
 
-Keywords: functionnal trait, tropical evergreen forest, Western Ghats, India...
+Keywords: functional trait, tropical evergreen forest, Western Ghats, India...
 
 # Introduction
+
+**Study scheme:**
+
+1. Do variations and covariations in functionnal traits can reveal species ecological strategies ?
+    i. Do functionnal traits vary among species ?
+    ii. Do covariations in functional traits represent axes of ecological strategy ?
+2. Do functional composition reflect the influence of abiotic and biotic drivers of community dynamics ?
+    i. Do local functional composition reflect the influence of abiotic and biotic drivers of community dynamics ?
+    ii. Do functional composition variation between communities reflect the influence of abiotic and biotic drivers of community dynamics ?
+
+**To add:**
+
+* Functional traits can be linked to trade-offs in tree strategy for functionning, survival and reproduction [@baraloto_decoupled_2010, and see table 1], which vary depending on environment condition.
+For instance, SLA varies along a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+* Leaf investment definition
+* Leaf ressource acquisition definition
 
@@
 The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996.
-The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergree forest type [@pascal_wet_1988].
+The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988].
 Climate and geology of the 10-ha plot can be considered as homogeneous.
-General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
+General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
+Each tree girth is monitored by a two-year census.
+**Girth monitoring to develop shortly**
 
-*Grith monitoring design ?*
+## Wood and leaf sampling
 
-## Sampling
-
-We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates.
+We collected trait data on 91 over 117 (78%) species.
+We sampled five individuals per species for most abundant species and we completed with less abundant and rare species with one to five individuals.
 Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals.
@@
 Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.
-Leaves and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
-For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).
+Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
+For some individuals we were unable to collect an appropriate branch sample (58 over 267 individuals representing 12 missing species over 91).
 
+## Trait measurements
+
+Leaf petioles, rachis and petiolules were removed for all measurements.
 Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite].
 Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time.
-Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias.
-Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues.
-Branch wood samples were then sink in water to evaluate their volume by water displacement.
-Samples were after dried for 48 hours at 80 degrees celsius and weighted.
+Once dried, the leaves were weighted directly after been extracted from the oven to avoid humidity bias.
+The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues.
+Branch wood samples were sink in water to evaluate their volume by water displacement.
+Samples were then dried for 48 hours at 80 degrees celsius and weighted.
 
-- Removed petiole, petiolule and rachis
+We calculated five functional traits of leaf and wood samples: leaf thickness (Thick in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1).
 
-## Traits
+Table 1: Functional traits measured.
+*Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*
 
-- Traits measurement (handbook)
-- 5 Basic traits description
-- Associated strategies axes with PCA (+ fortunel)
-- Associated covariation (following litterature)
-- Diversity indices (FD)
+```{r 1 Traits}
+rm(list = ls()); invisible(gc())
+PFT <- genPFT()
+table <- data.frame(
+  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
+  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
+  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
+            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
+            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
+            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
+            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
+  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
+                         round(sd(PFT$Thick, na.rm = T)), ')'),
+                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
+                         round(sd(PFT$LA, na.rm = T)), ')'),
+                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
+                         round(sd(PFT$LDMC, na.rm = T)), ')'),
+                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
+                         round(sd(PFT$SLA, na.rm = T),2), ')'),
+                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
+                         round(sd(PFT$WD, na.rm = T),2), ')')),
+  unit = c('um', 'mm2', 'mg.g-1', 'm2.kg-1', 'g.cm-3'),
+  strategy = c('Leaf defense vs. investment',
+               'Leaf ressource capture vs.investment',
+               'Leaf defense vs. investment',
+               'Leaf ressource acquisiton vs. defense',
+               'Stem transport, structure and defense')
+  )
+kable(table,  label="traits", row.names = F)
+rm(table)
+```
 
@@
 
-- Abitoic factors
-- PCA reduction
-- Biotic factors
-- Bitoic uncorrelation
-- Performance variables
-- Linking traits, biotic, abiotic factors and performance variables with sem
+We selected 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
+Digital elevation model is issued from a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
+Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
+Distance to the nearest potential river was computed using ArcGis [@desktop_arcgis_2011].
+
+In order to avoid multi-colinearity in future analyses, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning.
+We reduced our explanatory variables to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness to represent monsoon winds [@bunyan_effect_2015] and slope.
+
+We selected two uncorrelated biotic factors: canopy height and basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
+Canopy height was calculated from a digital canopy model built with the 2013 LIDAR campaign on Uppangala plots with resolution $1*1 m^2$.
+Basal area was calculated from 2013 girth inventory of Uppangala plot on stems with a diameter above 10 centimeters at breast height ($DBH>10~cm$).
 
@@
 
-- Community defintions
-- Community spatial organization investigation (variorgam + SAC)
-- Link with environment: step measure, linear regression, null model, residuals SAC
-- Individual filtering (RLQ)
-- Phylogenetic signal
+### Traits variation and covariation
+
+Traits variation was first investigated in three levels with the coefficient of variation ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species (inter-specific).
+
+For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
+To test the correlation and covariation among traits we performed a principal component analysis (PCA).
+
+### Functional composition of communities
+
+Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
+To test for pattern in functional composition of communities we used null model approach.
+A first null model (NM1) randomizes traits among species (permuting raw in `Trait*Species` matrix).
+The second model (NM2) randomizes abundances among species inside each community but keeps the observed list of species and trait values inside communities.
+NM1 allow to assess environmental filtering and biotic process on Uppangala plot species.
+NM2 bring more insight on environmental and biotic effects on species distribution inside communities [@bernard-verdier_community_2012].
+
+To investigate traits distribution of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010].
+CWV lower than expected under a null model will be evidence for convergence of traits inside the community whereas higher CWV will show divergence in trait distribution.
+CVNND is used to detect niche differenciation patterns in case of a lower CVNND as observed under a null model.
+
+CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
+abundance, $p_{i}$:
+$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
+CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution.
+CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
+abundance, $p_{i}$:
+$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
+CVNND was computed in abundance inside each community. For both CWV and CVNND, we calculated the standardized effect size (SES) as: 
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}}$$
+where $I_{obs}$ is the observed metric and $I_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.
+
+Linear regressions between CWMs and biotic and abiotic factors after step measure were compared to corresponding 999 null linear regressions from NM1 with F-statistic.
+For both CWV and CVNND, proportion of of communities differing from null expectaions was tested with a two-tailed test; and links between SES and environment was investigated by linear regression.
+Analyses were conducted on the five functional traits and the two economic spectra defined with the species scores obtained on the 2 first PCA axes.
+
+### Composition variation between communities
+
+Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
+Phylogenetic signal was globaly assessed for each functional trait among the phylogenetic tree by Moran's I statistic.
+Positive Moran's I value will indicate convergence of trait value inside the phylogeny whereas negative Moran's I will show divergence.
+
+$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities.
+Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
+In parrallel variation of functional composition between communities was assessed by $U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007.
+Observed values were compared to NM1.
+Negative Hardy's metrics significate that species are functionaly or phylogenetically
+
+All analysis were conducted in the R 3.3.0  free software environment for statistical analyses and graphics [@r_core_team_r:_2016] using the package 
+`ade4` [@dray_ade4_2007], 
+`ape` [@paradis_ape:_2004], 
+`phylosignal` [@keck_phylosignal:_2015] 
+and `spacodiR` [@eastman_spacodir:_2013].
 
@@
 
-## Community weighted means
+## Traits variation and covariation
 
-### Abiotic filtering
+Trait coefficients of variation analysis reveals a general increase of functional trait variation among scales (from intra-individual to inter-specific, see table 2).
+Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
 
-### Biotic filtering
+Table 2: Coefficient of intra-individual, intra-specific and inter-specific trait variation.
+*Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
+See table 1 for abbreviation.*
 
-## Community weighted variances
+```{r 2 CV}
+PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
+row.names(PFT_ind) <- PFT_ind$Group.1
+SpT <- unique(PFT[1:2])
+Sp <- SpT$Sp
+names(Sp) <- SpT$Tree
+SpCodeT <- unique(PFT[2:3])
+SpCode <- SpCodeT$SpCode
+names(SpCode) <- SpCodeT$Sp
+CET <- unique(PFT[c(1,4)])
+CE <- CET$CE
+names(CE) <- CET$Tree
+PFT_ind$Sp <- Sp[row.names(PFT_ind)]
+PFT_ind$SpCode <- SpCode[as.character(PFT_ind$Sp)]
+PFT_ind$CE <- CE[row.names(PFT_ind)]
+PFT_ind <- PFT_ind[-c(1:2)]
+cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
+cv$Individual <- unlist(lapply(
+  as.list(row.names(cv)),
+  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
+cv$Species <- unlist(lapply(
+  as.list(row.names(cv)),
+  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$Sp), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
+cv$Community <- unlist(
+  lapply(as.list(row.names(cv)),
+         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
+             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
+cv <- round(cv, 2)
+cv['WD','Individual'] <- ''
+kable(cv)
+rm(PFT, CET, cv, CE, Sp, SpT)
+```
 
-### Abiotic filtering
+The multivariate analysis of trait correlation shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
+The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood.
+Whereas the second PCA axis (PCA2) covary mainly with SLA representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
+PCA first axis might be associated with the WD covariation to the wood economic spectrum [WES, @chave_towards_2009].
+Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
+Our results integrate less traits (5 against 14), and less species (92 against 758), but still capture the same strategy axes.
 
-### Biotic filtering
+```{r 3 PCA PFT}
+library(vegan)
 
-## Functionnal diversity
+PFT_ind$LA <- log(PFT_ind$LA)
+PFT_ind$SLA <- log(PFT_ind$SLA)
+PFT_ind$ID <- row.names(PFT_ind)
+row.names(PFT_ind) <- PFT_ind$ID
+PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
+Species <- genSpecies()
+PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$Sp), FUN = mean, na.rm = T)
+row.names(PFT_sp) <- PFT_sp$Group.1
+PFT_sp$SpCode <- SpCode[row.names(PFT_sp)]
+PFT_sp <- PFT_sp[-c(1:2,4)]
+pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_sp, cor = T)
 
-## Individual filtering
+lambda <- pca$sdev * sqrt(nrow(pca$scores))
+pca.var = apply(pca$score, 2, var)
+PCAscores <- as.data.frame(pca$scores)[1:2]
+names(PCAscores) <- c('WES', 'LES')
+PCAscores$ID <- row.names(PCAscores)
+plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
+      xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
+par(new=T)
+Rot <- t(t(pca$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
+lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
+legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
+       text.col = c("green", "firebrick"))
 
-## Phylogenetic signal
+arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
+text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
+Hmisc::subplot(function(){
+  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
+          cex.names = 0.7, axes = T, names.arg = NA) ;
+  lines(b, bstick(length(pca.var)), col = 'red') ;
+  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
+         box.col = 'white')},
+  x=grconvertX(c(0.75,1), from='npc'),
+  y=grconvertY(c(0.05,0.30), from='npc')
+)
+axis (3)
+axis (4)
+
+PCAscores <- as.data.frame(pca$scores)[1:2]
+names(PCAscores) <- c('WES', 'LES')
+PCAscores$ID <- row.names(PCAscores)
+PFT_sp$ID <- row.names(PFT_sp)
+PFT_sp <- merge(PFT_sp, PCAscores, all = T)
+row.names(PFT_sp) <- PFT_sp$ID
+PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
+Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')]
+row.names(Q) <- PFT_sp$SpCode
+rm(PCAscores, PFT_ind, SpCodeT, lambda, pca, pca.var, Rot, XLIM)
+```
+
+Figure 1: Principal component analysis (PCA) performed on species trait means.
+*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
+See table 1 for abbreviation.*
+
+```{r 4 env-com}
+library(raster)
+Env <- genEnv()
+Trees <- genTrees()
+Trees$ID <- row.names(Trees)
+Trees <- merge(Trees, PFT_sp)
+row.names(Trees) <- Trees$ID
+Trees <- Trees[-which(names(Trees) == 'ID')]
+com <- quadrats()
+XY <- Trees[c('x', 'y')]
+coordinates(XY) <- ~ x + y
+proj4string(XY) <- crs(com)
+Trees$com <- (XY %over% com)[,1]
+com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
+names(Deriv)[5] <- 'SW'
+com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
+com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
+com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
+pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
+R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+row.names(R) <- com@data$id
+L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+L[is.na(L)] <- 0
+sel <- intersect(row.names(Q), names(L))
+Q <- Q[sel,]
+L <- L[sel]
+R <- R[row.names(L),]
+L <- list(abundance = L, `presence-absence` = L)
+L$`presence-absence`[L$`presence-absence` > 1]  <- 1
+rm(Deriv, XY, com, Env, Trees, sel, BA)
+```
+
+## Functional composition of communities
+
+Table 4: Summary of linear models and null models for traits community weighted means.
+_Linear model results for each trait community weighted means by species relative abundance after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Model global and partial F-statistics were compared to 999 null model F-statistics in which traits were randomized among species.
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+**Second table are the result in presence-absence. They are juste here for discussion. But they probably won't be in the final report under this shape.**
+
+```{r CWM}
+# Observed CWMs
+CWM <- lapply(L, function(l){lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){apply(l, 1, function(x){weighted.mean(Q[,v], x, na.rm=T)})})})
+CWM <- lapply(CWM, function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})
+## Regressions
+reg <- paste('~', paste(names(R), collapse = ' + '))
+CWM.lm <- lapply(CWM, function(x){lapply(x, function(y){lm(y ~., R)})})
+CWM.lm <- lapply(CWM.lm, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
+## Global F
+Fval <- lapply(CWM.lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
+Fval <- lapply(Fval, function(x){do.call('rbind.data.frame', x)})
+Fval <- data.frame(Fval)
+names(Fval) <- names(CWM.lm)
+## Partial F
+pFval <- lapply(CWM.lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})
+
+# Simulated CWMs
+n = 999
+nQ <- rep(list(Q), n)
+nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
+nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
+cl <- makeCluster(3)
+clusterExport(cl, list('L', 'nQ'))
+nCWM <- clusterApply(cl, L, function(l){
+  lapply(nQ, function(q){
+    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){
+      apply(l, 1, function(x){
+        weighted.mean(q[,v], x, na.rm=T)
+      })
+    })
+  })
+})
+stopCluster(cl)
+rm(cl)
+nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})})
+## Regressions
+nCWM.lm <- mapply(function(x,y){
+  lapply(y, function(z){
+    mapply(function(r,s){
+      lm(formula(paste0('s ~ ', as.character(r$terms)[3])), cbind(s,R))
+    }, r = x, s = z, SIMPLIFY = F)
+  })
+}, x = CWM.lm, y = nCWM, SIMPLIFY = F)
+rm(nCWM) ; invisible(gc())
+## Global F
+nFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
+nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
+nFval <- lapply(nFval, function(f){lapply(f, unlist)})
+pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
+row.names(pval) <- row.names(Fval)
+rm(Fval, nFval)
+# Partial F
+npFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})})
+npFval <- lapply(npFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
+npFval <- lapply(npFval, function(f){lapply(f, function(x){do.call('cbind', x)})})
+ppval <- mapply(function(x,y){mapply(function(r,s){mapply(function(p,q){rank(c(p,q))[1]/(n+1)}, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))}, r = x, s = y)}, x = pFval, y = npFval, SIMPLIFY = F)
+ppval <- lapply(ppval, function(y){lapply(y, function(x){data.frame(t(x))})})
+ppval <- lapply(ppval, function(y){data.frame(data.table::rbindlist(y, fill = T))})
+ppval <- lapply(ppval, function(y){row.names(y) <- names(CWM.lm$abundance) ; return(y)})
+ppval$abundance <- ppval$abundance[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy')]
+ppval$abundance$BA <- NA
+ppval$`presence-absence` <- ppval$`presence-absence`[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+ppval <- lapply(ppval, function(x){1 - x})
+rm(pFval, npFval)
+
+# Result tables
+kable(regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]]))
+kable(regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]]))
+rm(CWM.lm, nCWM.lm, pval, ppval, reg) ; invisible(gc())
+```
+
+Table 5: Summary of linear models for traits standardized effect size of community weighted variances.
+_Linear model results for each trait standardized effect size of community weighted variances species relative abundance after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+**First output is the result of two tails test on communities CWVs for each traits. They are juste here for discussion. But they probably won't be in the final report under this shape.**
+
+```{r CWV, fig.height=4}
+# Observed CWVs
+CWV <- lapply(names(CWM$abundance), function(v){apply(cbind(data.frame(CWM$abundance[[which(names(CWM$abundance) == v)]]),L$abundance), 1, function(x){sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))})})
+names(CWV) <- names(CWM$abundance)
+
+# Simulated CWVs 
+nL <- rep(list(L$abundance), n)
+nL <- lapply(nL,function(l){apply(l, 1, function(x){x[x != 0] <- sample(x[x != 0]) ; return(x)})}) 
+nL <- lapply(nL, function(x){t(x)})
+nL <- lapply(nL, data.frame)
+
+cl <- makeCluster(3)
+clusterExport(cl, list('nL', 'CWM', 'Q'))
+nCWM.cwv <- clusterApply(cl, nL, function(l){
+  lapply(as.list(names(CWM$abundance)), function(v){
+    apply(l, 1, function(x){
+      weighted.mean(Q[,v], x, na.rm=T)
+    })
+  })
+})
+stopCluster(cl)
+nCWM.cwv <- lapply(nCWM.cwv, function(x){names(x) <- names(CWM$abundance) ; return(x)})
+nCWV <- mapply(function(l,cwm){
+  lapply(names(CWM$abundance), function(v){
+    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
+      sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))
+    })
+  })
+}, l = nL, cwm = nCWM.cwv, SIMPLIFY = F)
+nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
+rm(nCWM.cwv, cl, nL) ; invisible(gc())
+## Non null CWVs
+nCWV <- lapply(nCWV, `[`, names(nCWV[[1]]))
+nCWV <- apply(do.call(rbind, nCWV), 2, as.list)
+nCWV <- lapply(nCWV, function(x){do.call('cbind', x)})
+nCWV <- lapply(nCWV, data.frame)
+CWV.pval <- data.frame(mapply(function(x,y){
+  mapply(function(p,q){
+    rank(c(p,q))[1]/(n+1)
+  }, p = as.list(x), q = as.list(data.frame(t(y))))
+}, x = CWV, y = nCWV))
+CWV.pval.res <- data.frame(inf = apply(CWV.pval, 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(CWV.pval, 2, function(x){length(which(x > 0.975))/length(x)}))
+## CWV SES
+CWV.ses <- data.frame(mapply(function(x,y){
+  mapply(function(p,q){
+    (p - mean(q))/sd(q)
+  }, p = as.list(x), q = as.list(data.frame(t(y))))
+}, x = CWV, y = nCWV))
+CWV.ses.lm <- apply(CWV.ses, 2, function(y){lm(y ~., R)})
+CWV.ses.lm <- lapply(CWV.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
+CWV.ses.res <- lapply(CWV.ses.lm, function(x){summary(x)$coefficients})
+CWV.ses.res <- lapply(CWV.ses.res, data.frame)
+CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
+CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
+CWV.ses.res <- CWV.ses.res[-1]
+CWV.ses.res$Slope <- NA
+CWV.ses.res <- CWV.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
+row.names(CWV.ses.res) <- names(CWV.ses)
+CWV.ses.res[is.na(CWV.ses.res)] <- ''
+CWV.ses.res <- data.frame(CWV.ses.res)
+CWV.ses.res$R2 <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)
+
+# Results
+barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
+kable(CWV.ses.res)
+rm(CWV.pval, CWV.pval.res, CWV.ses.res, CWV.ses.lm, nCWV)
+```
+
+Table 6: Summary of linear models for traits standardized effect size of coefficient of variation of nearest neighbour distances.
+_Linear model results for each trait standardized effect size of coefficient of variation of nearest neighbour distances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+**First output is the result of two tails test on communities CWVs for each traits. They are juste here for discussion. But they probably won't be in the final report under this shape.**
+
+```{r CVNND, fig.height=4}
+# Observed CVNND
+CVNND <- apply(L$abundance, 1, function(l){apply(Q, 2, function(q){cati::CVNND(rep(q, l), na.rm = T)})})
+
+# Simulated CVNNDs 
+cl <- makeCluster(3)
+clusterExport(cl, list('nQ', 'L'))
+nCVNND <- clusterApply(cl, nQ, function(x){
+  apply(L$abundance, 1, function(l){
+    apply(x, 2, function(q){
+      cati::CVNND(rep(q,l), na.rm = T)
+    })
+  })
+})
+stopCluster(cl) ; rm(cl)
+rm(nQ) ; invisible(gc)
+## Non null CVNNDs
+nCVNND <- lapply(nCVNND, function(x){as.list(data.frame(t(x)))})
+nCVNND <- lapply(nCVNND, function(x){lapply(x, function(y){names(y) <- colnames(CVNND) ; return(y)})})
+nCVNND <- lapply(nCVNND, `[`, names(nCVNND[[1]]))
+nCVNND <- apply(do.call(rbind, nCVNND), 2, as.list)
+nCVNND <- lapply(nCVNND, function(x){do.call('cbind', x)})
+nCVNND <- lapply(nCVNND, data.frame)
+CVNND.pval <- data.frame(mapply(function(x,y){
+  mapply(function(p,q){
+    rank(c(p,q))[1]/(n+1)
+  }, p = as.list(x), q = as.list(data.frame(t(y))))
+}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
+row.names(CVNND.pval) <- colnames(CVNND)
+CVNND.pval.res <- data.frame(inf = apply(CVNND.pval, 2, function(x){length(which(x < 0.05))/length(x)}))
+## CVNND SES
+CVNND.ses <- data.frame(mapply(function(x,y){
+  mapply(function(p,q){
+    (p - mean(q))/sd(q)
+  }, p = as.list(x), q = as.list(data.frame(t(y))))
+}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
+row.names(CVNND.ses) <- colnames(CVNND)
+CVNND.ses.lm <- apply(CVNND.ses, 2, function(y){lm(y ~., R)})
+CVNND.ses.lm <- lapply(CVNND.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
+CVNND.ses.res <- lapply(CVNND.ses.lm, function(x){summary(x)$coefficients})
+CVNND.ses.res <- lapply(CVNND.ses.res, data.frame)
+CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
+CVNND.ses.res <- data.frame(data.table::rbindlist(lapply(CVNND.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
+CVNND.ses.res <- CVNND.ses.res[-1]
+CVNND.ses.res$Slope <- NA
+CVNND.ses.res <- CVNND.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+CVNND.ses.res <- apply(CVNND.ses.res, 2, as.character)
+row.names(CVNND.ses.res) <- names(CVNND.ses)
+CVNND.ses.res[is.na(CVNND.ses.res)] <- ''
+CVNND.ses.res <- data.frame(CVNND.ses.res)
+CVNND.ses.res$R2 <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)
+
+# Results
+barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
+kable(CVNND.ses.res)
+rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
+```
+
+## Composition variation between communities
+
+**Artefacts of phylogenetic tree building to be removed**
+
+```{r PhyloMaker}
+library(phylosignal)
+library(phytools)
+library(phylobase)
+path <- system.file('phylogeny', package = 'Uppangala')
+source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
+phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
+nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
+sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), c('LatinName', 'Genus', 'Family')]
+names(sp) <- c('species', 'genus', 'family')
+results <- invisible(S.PhyloMaker(sp, phylo, nodes)) # Humboldtia_brunonis unmatched
+tree <- results$Scenario.1
+rm(phylor, results, modes, sp, S.PhyloMaker, path)
+```
+
+Table 7: Summary of Moran’s I statistic and p-values for phylogenetic signal of functional traits.
+_See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r Phylosignal}
+data <- PFT_sp
+row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
+tips <- row.names(data)[which(is.na(data$WD))]
+phy <- drop.tip(tree, tips)
+data <- data[phy$tip.label,c(2:5,7)]
+p4d <- phylo4d(phy,data)
+phyloS <- phyloSignal(p4d, methods = 'I') 
+phyloS <- cbind(phyloS$stat, phyloS$pvalue)
+names(phyloS) <- c('Moran\'s I', 'p-value')
+phyloS$` ` <- lapply(as.list(phyloS$`p-value`), stars, ns = ' ')
+kable(phyloS)
+rm(data, tips, p4d, phyloS)
+```
+
+Variations of both functional and phylogenetic compositions between communities don't significantly differ from null models ($U_{st} = -1.1*10^-3$ with $p-value = 0.114$, $\tau _{st} = 1.8*10^-3$ with $p-value = 0.079$, $\beta _{st} = -8.3*10^-4$ with $p-value = NaN$, and $\pi _{st} = -1.8*10^-3$ with $p-value = 0.247$).
+
+**Following tables will probably be removed in final report by simply including results in text. They are just here for discussion.**
+
+Table 8: Summary of variation of functional composition between communities.
+_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities.
+Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r Functionnal spatial structure, eval=FALSE, include=FALSE}
+library(spacodiR)
+mtest <- dist(as(quadrats(), 'SpatialPoints')@coords)
+
+# Observed Hardys
+spafun <- lapply(as.list(1:length(Q)), function(i){
+  if (length(which(is.na(Q[i]))) > 0)
+    q.names <- row.names(Q[-which(is.na(Q[i])),])
+  else
+    q.names <- row.names(Q)
+  sel <- intersect(names(L$abundance), q.names)
+  return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(Q[sel,i], row.names = sel), pairwise = T))
+})
+names(spafun) <- names(Q)
+spafun <- lapply(spafun, function(x){x[c('pairwise.Ist','pairwise.Ust', 'pairwise.TAUst')]})
+## Mantel values
+spafun.mantel <- lapply(spafun, function(t){lapply(t, function(m){vegan::mantel(m, mtest, permutations = 0)$statistic})})
+
+# Simulated Hardys
+cl <- makeCluster(3)
+clusterExport(cl, list('nQ', 'L', 'spacodi.calc'))
+nspafun <- clusterApply(cl, nQ, function(q){
+  lapply(as.list(1:length(q)), function(i){
+    if (length(which(is.na(q[i]))) > 0)
+      q.names <- row.names(q[-which(is.na(q[i])),])
+    else
+      q.names <- row.names(q)
+    sel <- intersect(names(L$abundance), q.names)
+    return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(q[sel,i], row.names = sel), pairwise = T))
+  })
+})
+stopCluster()
+rm(cl, nQ) ; invisible(gc())
+## Simulated mantels
+
+# Result
+kable(spafun.res)
+```
+
+Table 9: Summary of variation of phylogenetic composition between communities.
+_$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities.
+Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix).
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r Phylogenetic spatial structure, eval=FALSE, include=FALSE}
+for(i in 1:2) phy[[i]]$tip.label <- gsub('_', ' ', phy[[i]]$tip.label, fixed = T)
+for(i in 1:2) phy[[i]]$tip.label <- as.character(l$Sp_Code[match(phy[[i]]$tip.label, row.names(l))])
+Lm <- lapply(phy, function(x){Lm[intersect(names(Lm), x$tip.label)]})
+phyRD <- lapply(phy, function(x){x$node.label <- NULL ; return(x)})
+phyRD <- lapply(phyRD, function(x){rep(list(x), 999)})
+phyRD <- lapply(phyRD, function(x){lapply(x, resamp.phy)})
+LmRD <- lapply(Lm, function(x){rep(list(x), 999)})
+LmRD <- mapply(function(x,y){mapply(function(r,s){r[intersect(names(r), s$tip.label)]}, r = x, s = y, SIMPLIFY = F)}, x = LmRD, y = phyRD, SIMPLIFY = F)
+phyRD <- mapply(function(x,y){mapply(function(r,s){drop.tip(r, setdiff(r$tip.label, names(s)))}, r = x, s= y, SIMPLIFY = F)}, x = phyRD, y = LmRD, SIMPLIFY = F)
+
+spaphy <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), phy = x)}, x = phy, y = Lm.l, SIMPLIFY = F)
+spaphy <- do.call('rbind.data.frame', lapply(spaphy, function(x){x[1:4]}))
+
+spaphy.l <- mapply(function(x,y){mapply(function(r,s){spacodi.calc(t(r), s)}, r = x, s = y, SIMPLIFY = F)},x = LmRD, y = phyRD, SIMPLIFY = F)
+spaphy.l <- lapply(spaphy.l, function(x){lapply(x, function(y){y[1:4]})})
+spaphyRD <- list(Ist = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Ist}))}),
+                 Tst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Tst}))}),
+                 Bst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Bst}))}),
+                 PIst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$PIst}))}))
+pvalues <- mapply(function(obs,rd){
+  mapply(function(o,r){rank(c(o, r))[1]/1000}, o = obs, r = rd)
+  }, obs = lapply(as.list(spaphy), as.list), rd = spaphyRD)
+row.names(pvalues) <- row.names(spaphy)
+
+pvalues <- as.data.frame(pvalues)[3:4]
+pvalues$Bst <-  rep(NaN,2)
+spaphy.res <- format(spaphy[3:4], scientific = T, digits = 2)
+pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
+spaphy.res$Bst <- paste(spaphy.res$Bst, '/', pvalues$Bst)
+spaphy.res$PIst <- paste(spaphy.res$PIst, '/', pvalues$PIst)
+kable(spaphy.res)
+```
 
@@
 
+## Strategy axes
+## Community assembly
+## Process scales
+## Limits and perspectives
+
 # *Acknowledgement*
 
+# Supplementary materials
+
+**I'm aware that Supplementary materials goes after References (as figure and tables), but it's not possible for the moment with Rmarkdown and knitr.**
+
+## S1: Principal component analysis performed on topgraphical variables
+
+Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
+Elevation and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
+
+```{r PCA Env}
+lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
+pca.env.var = apply(pca.env$score, 2, var)
+plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
+     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
+     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
+Hmisc::subplot(function(){
+  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
+          cex.names = 0.7, axes = T, names.arg = NA) ;
+  lines(b, bstick(length(pca.env.var)), col = 'red') ;
+  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
+         box.col = 'white')},
+  x=grconvertX(c(0.07,0.37), from='npc'),
+  y=grconvertY(c(0.025,0.30), from='npc')
+)
+par(new=T)
+Rot <- t(t(pca.env$loadings)*lambda)
+XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
+XLIM <- XLIM+(XLIM*0.2)
+plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
+lines(XLIM, c(0,0), lty = 2, lwd = 2)
+lines(c(0,0), XLIM, lty = 2, lwd = 2)
+col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
+arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
+        Rot[,1],Rot[,2],lwd=2,col=col)
+legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
+       text.col = col, cex = 1.2)
+text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
+axis (3)
+axis (4)
+```
+
+Supplementary material S1: Principal component analysis (PCA) performed on topgraphical variables.
+*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
+SW stand for southwesterness.*
+
 # References

### 2016-07-22 R/CWT.R

@@ -1,71 +1,0 @@
-#' @importFrom plyr ddply
-#' @importFrom sp merge
-#' @importFrom Hmisc wtd.var wtd.mean
-NULL
-
-#' Community weighted trait
-#'
-#' Compute a community weighted metric (mean or variance) with traits
-#'
-#' @param Trees data.frame. Trees data frame
-#' @param com SpatialGridDataFrame. Communities with their environment (abiotic
-#'   and biotic)
-#' @param traits char. selected traits
-#' @param metric char. selected metric to be computed on community
-#' @param weights char. selected weights
-#' @param env char. selected environmental variables to be kept
-#'
-#' @return a list with a community grid data frame for each weighting methods
-#'   containing the environmental values, and the community weighted traits by
-#'   defined metrics
-#'
-#' @export
-#'
-#' @examples
-#'
-CWT <- function(Trees, com,
-                traits = c('LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'),
-                metric = c('mean', 'variance'),
-                weights = c('abundance', 'presence-absence'),
-                env = c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')){
-
-  weights <- unlist(lapply(as.list(weights), function(x){switch (x,
-                                                                 'abundance' = 'ab',
-                                                                 'presence-absence' = 'pa')}))
-  com <- com[c('id', env)]
-  cwt.l <- list()
-
-  for (wgt in weights) {
-
-    # Preparing result
-    i <- (length(cwt.l) + 1)
-    cwt.l[[i]] <- com
-    names(cwt.l)[i] <- switch (wgt,
-                                 'ab' = 'abundance',
-                                 'pa' = 'presence-absence')
-
-    # Trees table
-    trees <- switch (wgt,
-                     'ab' = Trees,
-                     'pa' = Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),])
-
-    for (m in metric) {
-
-      fun <- switch (m,
-                     'mean' = base::mean,
-                     'variance' = stats::var)
-      cwt <- aggregate(trees, by = list(trees$com), fun, na.rm = T)
-      names(cwt)[1] <- 'id'
-      cwt <- cwt[which(names(cwt) %in% c('id', traits))]
-
-      # Adding result
-      end <- switch (m,
[…]

### 2016-07-22 R/PFTrel.R

@@ -1,27 +1,0 @@
-#' @importFrom smatr sma
-NULL
-
-#' PFT relation
-#'
-#' @param PFT1 char. Name of the first PFT
-#' @param PFT2 char. Name of the second PFT
-#' @param data data.frame. PFT data set
-#' @param captions vector. Captions for PFT
-#' @param ... other argument pass to plot
-#'
-#' @return Graphs of relation between the two PFT
-#' @export
-#'
-#' @examples
-#' data(PFT)
-#' data(captions)
-#' PFTrel('SLA', 'WD', PFT, captions)
-#'
-PFTrel = function(PFT1, PFT2, data, captions, ...){
-  plot(formula(paste(PFT1, '~', PFT2)), data = data,
-       main = paste(PFT1, "~", PFT2), xlab = captions[PFT2], ylab = captions[PFT1], ...)
-  abline(sma(formula(paste(PFT1, '~', PFT2)), data), col = 'red')
-  mtext(paste('Standardized Major Axis : R\u00b2 =',round(sma(formula(paste(PFT1, '~', PFT2)), data)$r2[[1]], 2)), 1, col = 'red')
-  c = cor.test(formula(paste('~', PFT1, '+', PFT2)), data, method = 'spearman')
-  mtext(paste('Spearman\'s correlation : cor =', round(c$estimate, 2), stars(c$p.value)) , 3, col = 'blue')
-}

### 2016-07-22 R/genCaptions.R

@@ -1,45 +1,0 @@
-#' PFT data captions generation
-#'
-#' Generates PFT data captions
-#'
-#' @return PFT data captions
-#' @export
-#'
-#' @examples
-#' genCaptions()
-#'
-genCaptions <- function(){
-  captions <- c('Leaf ID',
-                'Tree ID',
-                'Species',
-                'Species code',
-                'Crown exposure (dawkins 13 index)',
-                'Simplified crown exposure (dawkins 5 index)',
-                'Thickness (\u00b5m)',
-                'Leaf area (mm\u00b2)',
-                'Logarithm of leaf area (mm\u00b2)',
-                'Leaf dry matter content (mg.g\u207b\u00b9)',
-                'Leaf tissue density (g.mm\u207b\u00b3)',
-                'Specific leaf area (m\u00b2.kg\u207b\u00b9)',
-                'Logarithm of specific leaf area (m\u00b2.kg\u207b\u00b9)',
-                'Specific holes area (mm\u00b2.mm\u207b\u00b2)',
-                'Logarithm of specific holes area (mm\u00b2.mm\u207b\u00b2)',
-                'Wood density (g.cm\u207b\u00b3)')
-  names(captions) = c('ID',
-                      'Tree',
-                      'SP',
-                      'Species code',
-                      'CE',
-                      'CEs',
-                      'Thick',
-                      'LA',
-                      'log(LA)',
-                      'LDMC',
-                      'LTD',
-                      'SLA',
-                      'log(SLA)',
-                      'SHA',
-                      'log(SHA)',
-                      'WD')
-  return(captions)
-}

### 2016-07-22 R/genEnv.R

@@ -21,17 +21,7 @@ genEnv <- function(area = data.frame(x = c(571675.8, 572011.3),
   extent <- extent(t(as.matrix(area)))
   DEM <- raster(system.file('extdata', 'DEM.tif', package = 'Uppangala'))
   DEM <- crop(DEM, extent)
-  # Distance to rivers generation (kept code but no just loading raster)
-  # rivers <- shapefile('~/data/Env/Rivers/stream_order1.shp') # Generated under ArcGis
-  # rivers.r <- rasterize(rivers, Env$DEM, field=1)
-  # rivdist.r <- distance(rivers.r)
-  # rivdist.r <- log(reclassify(rivdist.r, c(100,Inf,100)))
-  # writeRaster(rivdist.r, 'inst/extdata/Rivers.tif', 'GTiff')
-  Rivers <- raster(system.file('extdata', 'Rivers.tif', package = 'Uppangala'))
-  Rivers <- crop(Rivers, extent)
   Canopy <- raster(system.file('extdata', 'Canopy.tif', package = 'Uppangala'))
   Canopy <- crop(Canopy, extent)
-  return(list(DEM = DEM,
-              Rivers = Rivers,
-              Canopy = Canopy))
+  return(list(DEM = DEM, Canopy = Canopy))
 }

### 2016-07-22 R/genPFT.R

@@ -1,3 +1,6 @@
+#' @include  genSpecies.R
+NULL
+
 #' PFT data generation
 #'
 #' Generates PFT data set from external data collected on the field
@@ -10,14 +13,13 @@
 #'
 genPFT <- function(){
   # Opening
-  path <- system.file('extdata', 'data.csv', package = 'Uppangala')
+  path <- system.file('extdata', 'PFT.csv', package = 'Uppangala')
   PFT <- read.csv(path, stringsAsFactors = F, na.strings = c('#N/A', 'NA'))
   PFT <- PFT[-c(13:14)] # Deleting SD & VD in waiting of measurements
+  names(PFT)[3:4] <- c('Sp', 'SpCode')
   # Factors
   PFT$Thick <- as.numeric(PFT$Thick)
   PFT$CE <- as.factor(tolower(PFT$CE))
-  PFT$SP <- as.factor(PFT$SP)
-  PFT$Sp_Code <- as.factor(PFT$Sp_Code)
   # Variables modification
   PFT$CEs <- as.numeric(as.character(substr(PFT$CE, 1, 1)))
   PFT$LTD <- as.numeric(PFT$LTD)
@@ -27,6 +29,11 @@ genPFT <- function(){
   names(PFT)[which(names(PFT) == 'HA')] = 'SHA' # Specific holes area
   # Errors modification
   PFT$WD[which(PFT$WD > 1)] <- NA # Wood sample too small for correct WD measurement
+  species <- genSpecies()
+  PFT$SpCode[!(PFT$SpCode %in% species$SpCode)] <- species$SpCode[match(PFT$SpCode[!(PFT$SpCode %in% species$SpCode)], species$OldCode)]
+  PFT$Sp <- species$LatinName[match(PFT$SpCode, species$SpCode)]
+  PFT$Sp <- as.factor(PFT$Sp)
+  PFT$SpCode <- as.factor(PFT$SpCode)
   # Final
   PFT <- PFT[-which(duplicated(PFT)),]
   row.names(PFT) <- PFT$ID

### 2016-07-22 R/genSpecies.R

@@ -14,10 +14,10 @@ genSpecies <- function(){
   path <- system.file('extdata', 'Species.csv', package = 'Uppangala')
   species <- read.table(path,sep=",",dec=".", stringsAsFactors =F, h = T) # Dans BDD Uppangala
   # Formatting table
-  species <- species[c('spco', 'Latin_name', 'Family', 'PotStrata')]
-  colnames(species) <- c("SpCode","LatinName","Family","PotStrata")
-  species <- species[-which(is.na(species$LatinName)),]
+  species <- species[c('Sp_code', 'Sp_code_large', 'Latin_name', 'Family', 'Genus', 'Species', 'Pot_strata', 'old_Sp_code')]
+  colnames(species) <- c("SpCode", "SpCodeL", "LatinName", "Family", 'Genus', 'Species', 'PotStrata', 'OldCode')
   species$SpCode <- tolower(species$SpCode)
+  species[duplicated(species$SpCode),1] <- species[duplicated(species$SpCode),8]
   rownames(species) <- species$SpCode
   # Expliciting strata
   species$ESI <- 0; species[grep(x=species$PotStrata,pattern="\\bESI\\b"),]$ESI <- 1
@@ -37,6 +37,11 @@ genSpecies <- function(){
   species$Strata[species$ESIV==1] <- 'ESIV'
   species$Strata <- as.factor(species$Strata)
   species <- species[-which(names(species) %in% levels(species$Strata))]
+  # Errors corrections
+  species['agin',] <- c('agin', 'Agrind', "Agrostistachys indica", 'Euphorbiaceae', 'Agrostistachys', 'indica', NA, 'agin', NA)
+  species[which(species$Family == 'Flacourtiaceae'),4] <- 'Salicaceae'
+  species[which(species$Family == 'Steruliaceae'),4] <- 'Malvaceae'
+  species[which(species$Family == 'Cesaliniaceae'),4] <- 'Fabaaceae'
 
   return(species)
 }

### 2016-07-22 R/genTrees.R

@@ -1,3 +1,6 @@
+#' @include genSpecies.R
+NULL
+
 #' Trees data generation
 #'
 #' Generates Trees data set from external data extracted from the database
@@ -9,9 +12,10 @@
 #' summary(genTrees())
 #'
 genTrees <- function(){
+
   # Opening trees
-  path <- system.file('extdata', 'Sp_trees_allPlot.csv', package = 'Uppangala')
-  tree <- read.csv(path, header = T, stringsAsFactors = F)
+  path <- system.file('extdata', 'UppangalaCoordDataset.csv', package = 'Uppangala')
+  tree <- read.table(path, header = T, sep = ",", dec = ".")
   # ID creations
   tree$ID <- NA
   tree$ID[which(is.na(tree$Plot_10Ha_ID))] <- as.character(tree$Tree_10Ha_Field_ID[which(is.na(tree$Plot_10Ha_ID))])
@@ -20,82 +24,29 @@ genTrees <- function(){
   tree[which(duplicated(tree$ID)),] # rh47 duplicated
   tree <- tree[-which(duplicated(tree$ID)),] # Remove
   # Formatting table
-  tree <- data.frame(SpCode = tree$Sp_code, row.names = tree$ID)
-
-  # Opening locations
-  path <- system.file('extdata', 'Coord_trees_allPlot.csv', package = 'Uppangala')
-  location <- read.table(path, header = T, sep = ",", dec = ".")
-  # ID creations
-  location$ID <- NA
-  location$ID[which(is.na(location$Plot_10Ha_ID))] <- as.character(location$Tree_10Ha_Field_ID[which(is.na(location$Plot_10Ha_ID))])
-  location$ID[which(!is.na(location$Plot_10Ha_ID))] <- as.character(location$Tree_10Ha_ID[which(!is.na(location$Plot_10Ha_ID))])
-  # ID issues
-  location[which(duplicated(location$ID)),] # rh47 duplicated
-  location <- location[-which(duplicated(location$ID)),] # Remove
-  # Formatting table
-  row.names(location) <- location$ID
-  location <- location[7:8]
-  names(location) <- c('X', 'Y')
-
-  # Opening girth 5 ha plots
-  path <- system.file('extdata', 'Girth_5Ha.csv', package = 'Uppangala')
-  girth5 <- read.table(path, header = T, sep = ",", dec = ".")
-  girth5$ID.date <- paste(girth5$Tree_5Ha_ID,
[…]

### 2016-07-22 R/quadrats.R

@@ -29,6 +29,7 @@ quadrats <- function(cs = c(20,20),
   cd <- floor(diff(t(bb))/cs)  # number of cells per direction
   grd <- GridTopology(cellcentre.offset = cc, cellsize = cs, cells.dim = cd)
   grd <- SpatialGridDataFrame(grd,  dat = data.frame(id=1:prod(cd)), proj4string = proj)
+  grd@data$id <- paste0('C', grd@data$id)
 
   return(grd)
 }

### 2016-07-22 R/stars.R

@@ -10,7 +10,7 @@
 #' stars(0.049)
 #'
 stars = function(val, ns = 'n.s.'){
-  if(!is.numeric(val) || is.nan(val)){
+  if(!is.numeric(val) || is.nan(val) || is.na(val)){
     val = 1
   }
   if(val < 0.001){

### 2016-07-22 R/topo_classes.R

@@ -1,46 +1,0 @@
-#' Topographical classes
-#'
-#' @param DEM num. Digital elevation model values
-#' @param Slope num. Slope model values
-#' @param Curvature num. Curvature model values
-#' @param elev num. Discriminant elevation, autoset with NULL
-#' @param degree num. Discriminant slope, autoset with NULL
-#' @param dir num. Aspect direction to classify aspect in two categories (default SW for monsoon wind)
-#'
-#' @return char. Vector of topographical classes
-#' @export
-#'
-#' @examples
-#' data("Env")
-#' data("Trees")
-#' topo_classes(Trees$X, Trees$Y, Env$DEM$Ground, Env$DEM$Slope, Env$DEM$Curvature)
-#'
-topo_classes = function(Slope, Curvature, Aspect, degree = c(20,30), dir = 225){
-
-  # Factorizing aspect
-  AspectDir <- sin((Aspect + dir) / 180 * pi)
-  AspectDir <- c(0, 1)[findInterval(AspectDir, c(-1, 0, 1))] # To change with wind true direction
-
-  # Attributing classes
-  topo <- data.frame(d = Slope, c = Curvature, a = AspectDir)
-  topo$class <- NA
-  topo$class[topo$d < degree[1] &  topo$c >= 0] <- 'CvP' # Concave plane surface
-  topo$class[topo$d < degree[1] &  topo$c < 0] <- 'CxP' # Convex plane surface
-  topo$class[topo$d >= degree[2] &  topo$a == 1] <- 'SWSS' # South - West steep slope
-  topo$class[topo$d >= degree[2] &  topo$a == 0] <- 'NESS' # North - East steep slope
-  topo$class[topo$d >= degree[1] & topo$d < degree[2] &  topo$a == 1] <- 'SWMS' # South - West mean slope
-  topo$class[topo$d >= degree[1] & topo$d <  degree[2] & topo$a == 0] <- 'NEMS' # North - East mean slope
-
-#   # Classes Gunatileke et al, 2006
-#   topo$class = NA
-#   topo$class[topo$z < elev & topo$deg < degree & topo$curv < 0] = 1  # USS
-#   topo$class[topo$z < elev & topo$deg < degree & topo$curv >= 0] = 2 # USG
-#   topo$class[topo$z < elev & topo$deg >= degree & topo$curv < 0] = 3 # ULS
[…]

### 2016-07-26 inst/doc/draft.Rmd
v0.1.9023 Mantel results and Ist NM3

@@
 title: 'Uppangala tree functional traits: Article draft'
-author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
-output: pdf_document
+author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
+output: word_document
 bibliography: refs.bib
@@
 
-We selected two uncorrelated biotic factors: canopy height and basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
+We selected two biotic factors: canopy height and basal area.
 Canopy height was calculated from a digital canopy model built with the 2013 LIDAR campaign on Uppangala plots with resolution $1*1 m^2$.
@@
 
-### Functional composition of communities
+### Local functionnal composition of communities
 
@@
 $$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
-CVNND was computed in abundance inside each community. For both CWV and CVNND, we calculated the standardized effect size (SES) as: 
-$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}}$$
+CVNND for each trait $t$ was calculated as nearest neigbor distances (NND) variance, $\sigma _{NND}$, of the $i$ individuals in each community, divided by its mean, $\mu _{NND}$; and NND for one individual $i$ is calculated as the minimal distance $d$ between the $j$ other individuals of the community :
+$$CVNND = \frac{\sigma _{NDD}}{\mu _{NDD}} ~ with ~ NND_{i} = min(d_{i,j})_{i \neq j} ~ (3)$$
+For both CWV and CVNND, we calculated the standardized effect size (SES) as: 
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
 where $I_{obs}$ is the observed metric and $I_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.
@@
 
-### Composition variation between communities
+### Functionnal and phylogenetic composition between communities
 
@@
 
-$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities.
-Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
-In parrallel variation of functional composition between communities was assessed by $U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007.
-Observed values were compared to NM1.
-Negative Hardy's metrics significate that species are functionaly or phylogenetically
+$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between community pairs.
+In parrallel variation of functional composition between community pairs was assessed by $U st$ and $\tau st$ metric.
+Environmental distances between community pairs was also calculated as distances between communities on the plan composed by the two first axes of the principal component analysis of environmental variables (see suppplementary materials S1).
+Finally spatial distances between community pairs was also assessed.
+We used Mantel test to sudy the link between functionnal and phylogenetic pairwise compositions with hardy's metrics and environmental distances between community pairs.
+A second mantel test was realized to assess those pairwise links between environment and functionnal and phylogenetic structure knowing spatial structure with the spatial distances between community pairs matrix.
+
+Observed mantel correlations for phylogenetic composition were compared to 999 obtained with the null model *"1s"*" suggested by @hardy_testing_2008 randomizing phylogenetic tree tips.
+Observed mantel correlations for functionnal composition were compared to NM1, randomizing traits among species.
 
@@
 # Result tables
-kable(regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]]))
-kable(regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]]))
+kable(bold(regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]])))
+kable(bold(regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]])))
 rm(CWM.lm, nCWM.lm, pval, ppval, reg) ; invisible(gc())
@@
 CWV.ses.res <- lapply(CWV.ses.res, data.frame)
-CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
+CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
 CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
@@
 CWV.ses.res <- data.frame(CWV.ses.res)
-CWV.ses.res$R2 <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)
+CWV.ses.res$`$R^2$` <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)
 
@@
 barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
-kable(CWV.ses.res)
+kable(bold(CWV.ses.res))
 rm(CWV.pval, CWV.pval.res, CWV.ses.res, CWV.ses.lm, nCWV)
@@
 CVNND.ses.res <- lapply(CVNND.ses.res, data.frame)
-CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
+CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
 CVNND.ses.res <- data.frame(data.table::rbindlist(lapply(CVNND.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
@@
 CVNND.ses.res <- data.frame(CVNND.ses.res)
-CVNND.ses.res$R2 <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)
+CVNND.ses.res$`$R^2$` <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)
 
@@
 barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
-kable(CVNND.ses.res)
+kable(bold(CVNND.ses.res))
 rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
@@
 ## Composition variation between communities
-
-**Artefacts of phylogenetic tree building to be removed**
 
@@
 tree <- results$Scenario.1
-rm(phylor, results, modes, sp, S.PhyloMaker, path)
+rm(phylo, results, modes, sp, S.PhyloMaker, path)
 ```
@@
 phyloS <- phyloSignal(p4d, methods = 'I') 
-phyloS <- cbind(phyloS$stat, phyloS$pvalue)
-names(phyloS) <- c('Moran\'s I', 'p-value')
-phyloS$` ` <- lapply(as.list(phyloS$`p-value`), stars, ns = ' ')
-kable(phyloS)
-rm(data, tips, p4d, phyloS)
+phyloS <- data.frame(paste(round(phyloS$stat[,1], 3), 
+                           unlist(lapply(as.list(phyloS$pvalue[,1]), stars, ns = ' '))),
+                     row.names = row.names(phyloS$stat))
+names(phyloS) <- 'Moran\'s I'
+kable(bold(phyloS))
+rm(data, tips, p4d, phyloS, phy)
 ```
 
-Variations of both functional and phylogenetic compositions between communities don't significantly differ from null models ($U_{st} = -1.1*10^-3$ with $p-value = 0.114$, $\tau _{st} = 1.8*10^-3$ with $p-value = 0.079$, $\beta _{st} = -8.3*10^-4$ with $p-value = NaN$, and $\pi _{st} = -1.8*10^-3$ with $p-value = 0.247$).
-
-**Following tables will probably be removed in final report by simply including results in text. They are just here for discussion.**
-
-Table 8: Summary of variation of functional composition between communities.
-_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities.
-Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008.
+Table 8: Summary of Mantel test between floristic composition, environment and spatial distances between communities.
+_$Ist$  was used to investigate variation of floristic composition between communities.
+Observed values were compared to a null model randomizing individuals among subplots while preserving the regional abundance of species, and the number of individuals per community.
+E is the mantel test of the metric with the environmental distance matrix issued from PCA two first axes.
+S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix.
 The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
-```{r Functionnal spatial structure, eval=FALSE, include=FALSE}
+```{r Ist}
 library(spacodiR)
-mtest <- dist(as(quadrats(), 'SpatialPoints')@coords)
 
+# Environmental matrices
+XY <- data.frame(as(quadrats(), 'SpatialPoints')@coords)
+row.names(XY) <- paste0('C',row.names(XY))
+XY <- XY[row.names(L$abundance),]
+Ms <- dist(XY)
+XY <- pca.env$scores[,1:2]
+row.names(XY) <- paste0('C',row.names(XY))
+XY <- XY[row.names(L$abundance),]
+Me <- dist(XY)
+rm(XY)
+
+# Observed Ist
+spaflo <- spacodi.calc(sp.plot = t(L$abundance), pairwise = T)$'pairwise.Ist'
+## Mantel values
+spaflo.mantel <- list(`E` = vegan::mantel(spaflo, Me, permutations = 0)$statistic, 
+                      `S | E`= vegan::mantel.partial(spaflo, Ms, Me, permutations = 0)$statistic)
+
+# Simulated mantel
+nL <- vegan::permatfull(L$abundance, fixedmar = 'both', mtype = 'count', times = n, shuffle = 'both')
+cl <- makeCluster(3)
+clusterExport(cl, list('nL', 'spacodi.calc', 'Me', 'Ms'))
+nspaflo <- lapply(nL$perm, function(l){spacodi.calc(sp.plot = t(l), pairwise = T)$'pairwise.Ist'})
+nspaflo.mantel <- lapply(nspaflo, function(ns){
+  list(`E` = vegan::mantel(ns, Me, permutations = 0)$statistic, 
+       `S | E`= vegan::mantel.partial(ns, Ms, Me, permutations = 0)$statistic)})
+stopCluster(cl)
+rm(cl, nL) ; invisible(gc())
+## Non null mantels
+nspaflo.mantel <- lapply(nspaflo.mantel, `[`, names(nspaflo.mantel[[1]]))
+nspaflo.mantel <- apply(do.call(rbind, nspaflo.mantel), 2, as.list)
+nspaflo.mantel <- lapply(nspaflo.mantel, unlist)
+spaflo.res <- mapply(function(s, ns){
+    paste(format(s, scientific = T, digits = 3),
+          Uppangala::stars(min(1-rank(c(s,ns))[1]/(n+1),rank(c(s,ns))[1]/(n+1)), ns = '   '))
+}, s = spaflo.mantel, ns = nspaflo.mantel)
+spaflo.res <- data.frame(spaflo.res)
+names(spaflo.res) <- c('$I_{st}$')
+
+# Result
+kable(bold(spaflo.res))
+rm(mtest, spaflo.mantel, nspaflo, nspaflo.mantel, spaflo.res)
+```
+
+Table 9: Summary of Mantel test between functionnal composition, environment and spatial distances between communities.
+_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities.
+Observed values were compared to the null model NM1 randomizing traits among species (permuting raw in `Trait*Species` matrix).
+E is the mantel test of the metric with the environmental distance matrix issued from PCA two first axes.
+S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r Ust & TAUst}
 # Observed Hardys
 spafun <- lapply(as.list(1:length(Q)), function(i){
-  if (length(which(is.na(Q[i]))) > 0)
-    q.names <- row.names(Q[-which(is.na(Q[i])),])
-  else
-    q.names <- row.names(Q)
-  sel <- intersect(names(L$abundance), q.names)
-  return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(Q[sel,i], row.names = sel), pairwise = T))
+  if (length(which(is.na(Q[i]))) > 0) {
+    sel <- intersect(names(L$abundance), row.names(Q[-which(is.na(Q[i])),]))
+    return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(Q[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
+  } else {
+    return(spacodi.calc(sp.plot = t(L$abundance), sp.traits = Q, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
+  }
 })
 names(spafun) <- names(Q)
-spafun <- lapply(spafun, function(x){x[c('pairwise.Ist','pairwise.Ust', 'pairwise.TAUst')]})
 ## Mantel values
-spafun.mantel <- lapply(spafun, function(t){lapply(t, function(m){vegan::mantel(m, mtest, permutations = 0)$statistic})})
+spafun.mantel <- lapply(spafun, function(t){list(`E` = lapply(t, function(M){vegan::mantel(M, Me, permutations = 0)$statistic}), `S | E` = lapply(t, function(M){vegan::mantel.partial(M, Ms, Me, permutations = 0)$statistic}))})
 
-# Simulated Hardys
+# Simulated mantel
 cl <- makeCluster(3)
-clusterExport(cl, list('nQ', 'L', 'spacodi.calc'))
-nspafun <- clusterApply(cl, nQ, function(q){
-  lapply(as.list(1:length(q)), function(i){
-    if (length(which(is.na(q[i]))) > 0)
-      q.names <- row.names(q[-which(is.na(q[i])),])
-    else
-      q.names <- row.names(q)
-    sel <- intersect(names(L$abundance), q.names)
-    return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(q[sel,i], row.names = sel), pairwise = T))
+clusterExport(cl, list('nQ', 'L', 'Me', 'Ms', 'spacodi.calc'))
+nspafun.mantel <- clusterApply(cl, nQ, function(q){
+  s <- lapply(as.list(1:length(q)), function(i){
+    if (length(which(is.na(q[i]))) > 0) {
+      sel <- intersect(names(L$abundance), row.names(q[-which(is.na(q[i])),]))
+      return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(q[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
+    } else {
+      return(spacodi.calc(sp.plot = t(L$abundance), sp.traits = q, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
+    }
+  })
+  names(s) <- names(q)
+  res <- list(
+    `E` = lapply(s, function(t){
+      lapply(t, function(M){
+        try(vegan::mantel(M, Me, permutations = 0)$statistic)
+        })
+      }),
+    `S | E` = lapply(s, function(t){
+      lapply(t, function(M){
+        try(vegan::mantel.partial(M, Ms, Me, permutations = 0)$statistic)
         })
       })
-stopCluster()
+    )
+  res <- lapply(res, function(r){names(r) <- names(q) ; return(r)})
+  res <- lapply(res, `[`, names(res[[1]]))
+  res <- apply(do.call(rbind, res), 2, as.list)
+  return(res)
+})
+stopCluster(cl)
 rm(cl, nQ) ; invisible(gc())
-## Simulated mantels
+
+## Non null mantels
+nspafun.mantel <- lapply(nspafun.mantel, `[`, names(nspafun.mantel[[1]]))
+nspafun.mantel <- apply(do.call(rbind, nspafun.mantel), 2, as.list)
+nspafun.mantel <- lapply(nspafun.mantel, function(s){lapply(s, `[`, names(s[[1]]))})
+nspafun.mantel <- lapply(nspafun.mantel, function(s){apply(do.call(rbind, s), 2, as.list)})
+nspafun.mantel <- lapply(nspafun.mantel, function(s){lapply(s, function(m){lapply(m, `[`, names(m[[1]]))})})
+nspafun.mantel <- lapply(nspafun.mantel, function(s){lapply(s, function(m){apply(do.call(rbind, m), 2, as.list)})})
+nspafun.mantel <- lapply(nspafun.mantel, function(s){lapply(s, function(x){lapply(x, function(x){as.numeric(unlist(x))})})})
+spafun.res <- data.frame(mapply(function(x,y){
+  mapply(function(r,s){
+    mapply(function(p,q){
+      paste(format(p, scientific = T, digits = 3), 
+            Uppangala::stars(min(rank(c(p,q))[1]/(n+1), 1-rank(c(p,q))[1]/(n+1)), ns = '   '))
+    }, p = r, q = s, SIMPLIFY = F)
+  }, r = x, s= y)
+}, x = spafun.mantel, y = nspafun.mantel))
+spafun.res <- cbind(data.frame('Mantel test' = c('E', ' ', 'S|E', ' '), 'Metric' = rep(c('$U_{st}$', '$\\tau _{st}$'), 2)), spafun.res)
 
 # Result
-kable(spafun.res)
+kable(bold(spafun.res))
+rm(spafun.mantel, nspafun.mantel, spafun.res)
 ```
 
-Table 9: Summary of variation of phylogenetic composition between communities.
+Table 10: Summary of Mantel test between phylogenetic composition, environment and spatial distances between communities.
 _$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities.
-Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix).
+Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 randomizong phylogenetic tree tips.
+E is the mantel test of the metric with the environmental distance matrix issued from PCA two first axes.
+S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix.
 The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
-```{r Phylogenetic spatial structure, eval=FALSE, include=FALSE}
-for(i in 1:2) phy[[i]]$tip.label <- gsub('_', ' ', phy[[i]]$tip.label, fixed = T)
-for(i in 1:2) phy[[i]]$tip.label <- as.character(l$Sp_Code[match(phy[[i]]$tip.label, row.names(l))])
-Lm <- lapply(phy, function(x){Lm[intersect(names(Lm), x$tip.label)]})
-phyRD <- lapply(phy, function(x){x$node.label <- NULL ; return(x)})
-phyRD <- lapply(phyRD, function(x){rep(list(x), 999)})
-phyRD <- lapply(phyRD, function(x){lapply(x, resamp.phy)})
-LmRD <- lapply(Lm, function(x){rep(list(x), 999)})
-LmRD <- mapply(function(x,y){mapply(function(r,s){r[intersect(names(r), s$tip.label)]}, r = x, s = y, SIMPLIFY = F)}, x = LmRD, y = phyRD, SIMPLIFY = F)
-phyRD <- mapply(function(x,y){mapply(function(r,s){drop.tip(r, setdiff(r$tip.label, names(s)))}, r = x, s= y, SIMPLIFY = F)}, x = phyRD, y = LmRD, SIMPLIFY = F)
+```{r BETAst $ PIst}
+# Observed Hardys
+phy <- tree
+phy$tip.label <- Species$SpCode[match(gsub('_', ' ',phy$tip.label), Species$LatinName)]
+tips <- c(setdiff(phy$tip.label, names(L$abundance)), setdiff(names(L$abundance), phy$tip.label))
+phy <- drop.tip(phy, tips)
+phy$node.label <- NULL
+spaphy <- spacodi.calc(sp.plot = t(L$abundance), phy = phy, pairwise = T)[c('pairwise.Bst', 'pairwise.PIst')]
+## Mantel values
+spaphy.mantel <- list(`E` = lapply(spaphy, function(m){vegan::mantel(m, Me, permutations = 0)$statistic}), 
+                      `S | E`= lapply(spaphy, function(m){vegan::mantel.partial(m, Ms, Me, permutations = 0)$statistic}))
 
-spaphy <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), phy = x)}, x = phy, y = Lm.l, SIMPLIFY = F)
-spaphy <- do.call('rbind.data.frame', lapply(spaphy, function(x){x[1:4]}))
+# Simulated mantel
+nphy <- rep(list(phy), 999)
+cl <- makeCluster(3)
+clusterExport(cl, list('nphy', 'L', 'spacodi.calc', 'Ntip', 'resamp.phy', 'branching.times'))
+nphy <- clusterApply(cl, nphy, resamp.phy)
+stopCluster(cl)
+nspaphy <- lapply(nphy, function(p){spacodi.calc(sp.plot = t(L$abundance), phy = p, pairwise = T)[c('pairwise.Bst', 'pairwise.PIst')]})
+rm(cl, nphy) ; invisible(gc())
+nspaphy.mantel <- list(`E` =lapply(nspaphy, function(x){lapply(x, function(M){vegan::mantel(M, Me, permutations = 0)$statistic})}), `S | E` = lapply(nspaphy, function(x){lapply(x, function(M){vegan::mantel.partial(M, Ms, Me, permutations = 0)$statistic})}))
+## Non null mantels
+nspaphy.mantel <- lapply(nspaphy.mantel, function(s){lapply(s, `[`, names(s[[1]]))})
+nspaphy.mantel <- lapply(nspaphy.mantel, function(s){apply(do.call(rbind, s), 2, as.list)})
+nspaphy.mantel <- lapply(nspaphy.mantel, function(s){lapply(s, unlist)})
+spaphy.res <- mapply(function(s, ns){
+  mapply(function(x,y){
+    paste(format(x, scientific = T, digits = 3), 
+          Uppangala::stars(min(1-rank(c(x,y))[1]/(n+1),rank(c(x,y))[1]/(n+1)), ns = '   '))
+  }, x = s, y = ns)
+}, s = spaphy.mantel, ns = nspaphy.mantel)
+row.names(spaphy.res) <- c('$\\beta_{st}$', '$\\pi_{st}$')
 
-spaphy.l <- mapply(function(x,y){mapply(function(r,s){spacodi.calc(t(r), s)}, r = x, s = y, SIMPLIFY = F)},x = LmRD, y = phyRD, SIMPLIFY = F)
-spaphy.l <- lapply(spaphy.l, function(x){lapply(x, function(y){y[1:4]})})
-spaphyRD <- list(Ist = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Ist}))}),
-                 Tst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Tst}))}),
-                 Bst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Bst}))}),
-                 PIst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$PIst}))}))
-pvalues <- mapply(function(obs,rd){
-  mapply(function(o,r){rank(c(o, r))[1]/1000}, o = obs, r = rd)
-  }, obs = lapply(as.list(spaphy), as.list), rd = spaphyRD)
-row.names(pvalues) <- row.names(spaphy)
-
-pvalues <- as.data.frame(pvalues)[3:4]
-pvalues$Bst <-  rep(NaN,2)
-spaphy.res <- format(spaphy[3:4], scientific = T, digits = 2)
-pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
-spaphy.res$Bst <- paste(spaphy.res$Bst, '/', pvalues$Bst)
-spaphy.res$PIst <- paste(spaphy.res$PIst, '/', pvalues$PIst)
-kable(spaphy.res)
+# Result
+kable(bold(spaphy.res))
+rm(mtest, spaphy.mantel, nspaphy, nspaphy.mantel, spaphy.res)
 ```
@@
 
-**I'm aware that Supplementary materials goes after References (as figure and tables), but it's not possible for the moment with Rmarkdown and knitr.**
-
 ## S1: Principal component analysis performed on topgraphical variables
-
-Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
-Elevation and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
 
@@
 
+Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
+Elevation and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
+
 # References

### 2016-07-26 R/bold.R

@@ -1,0 +1,26 @@
+#' Bold
+#'
+#' Render bold cell in Rmarkdown depending on a pattern (to put in blod
+#' significant results for instance)
+#'
+#' @param table data.frame. Table on which to highlight cell in bold.
+#' @param pattern char. Pattern to be detected to highlight cells.
+#'
+#' @return data.frame
+#'
+#' @export
+#'
+#' @examples
+#' # See draft.Rmd
+#'
+bold <- function(table, pattern = '*'){
+  table <- apply(table, 2, function(col){
+    unlist(lapply(as.list(col), function(cell){
+      if(grepl(pattern, cell, fixed = T)){
+        cell <- paste0('__', cell, '__')
+      }
+      return(cell)
+    }))
+  })
+  return(table)
+}

### 2016-07-26 R/regression_table.R

@@ -30,7 +30,7 @@ regression_table <- function(LM, reg, pval, ppval){
     }
   }
   # LM R squarred
-  var_table$R2 <- unlist(lapply(LM, function(x){round(summary(x)$r.squared,2)}))
+  var_table$`$R^2$` <- unlist(lapply(LM, function(x){round(summary(x)$r.squared,2)}))
 
   # Null models
   var_table$`Null model` <- round(pval,3)

### 2016-07-29 inst/doc/draft.Rmd
v0.1.9024 M&M + Results v1

@@
 author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
-output: word_document
+output: pdf_document
 bibliography: refs.bib
@@
 
-1. Do variations and covariations in functionnal traits can reveal species ecological strategies ?
-    i. Do functionnal traits vary among species ?
-    ii. Do covariations in functional traits represent axes of ecological strategy ?
-2. Do functional composition reflect the influence of abiotic and biotic drivers of community dynamics ?
-    i. Do local functional composition reflect the influence of abiotic and biotic drivers of community dynamics ?
-    ii. Do functional composition variation between communities reflect the influence of abiotic and biotic drivers of community dynamics ?
+1. Do variations and covariations in functional traits reflect distinct ecological strategies within and among species ?
+    i. Do and how functional traits vary among species ?
+    ii. Do covariations in functional traits represent axes of ecological strategy within species ?
+2. Does functional composition reflect the influence of local abiotic and biotic drivers of community dynamics ?
+    i. Do local functional composition ($\alpha$ scale) reflect the influence of abiotic and biotic drivers of community dynamics ?
+    ii. Do functional composition variation between communities ($\beta$ scale) reflect the changing of abiotic and biotic conditions ?
 
@@
 * Functional traits can be linked to trade-offs in tree strategy for functionning, survival and reproduction [@baraloto_decoupled_2010, and see table 1], which vary depending on environment condition.
-For instance, SLA varies along a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
+* For instance, SLA varies along a soil fertility - shade index whereas LDMC is basically linked to soil fertility [@hodgson_is_2011]. **! In european temperate context !**
 * Leaf investment definition
@@
 
-The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village.
+The study was carried out in the the 10-ha "Uppangala plot" of undisturbed, old-growth wet evergreen Diptrocarp forest.
+The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India) in India's Western Ghats biodiversity hotspot.
+Climate is characterized by a peak of heavy rainfall during monsoon and a marked dry season.
+Annual precipitation averages 5108 mm with 90% between June and October and an average dry season (monthly rainfall $\leq 45$ mm) of 3.8 months [@pelissier_twenty_2011].
+Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996].
+Climate and geology of the 10-ha plot can be considered as homogeneous.
+General topography of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
 The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996.
-The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988].
-Climate and geology of the 10-ha plot can be considered as homogeneous.
-General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
-Each tree girth is monitored by a two-year census.
-**Girth monitoring to develop shortly**
+Since 1989, all trees $\geq 30$ cm girth at breast height (gbh) were monitored in an area progressively extended to 10ha in 2014.
+All trees were identified, tagged, mapped, and provided with permanent dendrometer bands censused every 2-5 years.
 
@@
 
-We collected trait data on 91 over 117 (78%) species.
-We sampled five individuals per species for most abundant species and we completed with less abundant and rare species with one to five individuals.
-Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals.
-Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot.
-For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013.
-Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy.
-Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.
+We collected trait data on 89 over 111 (80.1%) species. The 32 most abundant species represent more than 80% of Uppangala tree individuals.
+We sampled five individuals per species for most abundant species and we completed with less abundant and rare species with one to five individuals (in total of 267 individuals).
+Thus we got trait data for 99.0% of the tree individuals present in the Uppangala plot.
+For each individual we collected 6 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation from @perez-harguindeguy_new_2013.
+Individuals and leaves were selected in consistent light condition, development stage and sanitary state.
+Dawkins crown exposure index [@dawkins_management_1958] was assessed for each tree sampled.
 Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
-For some individuals we were unable to collect an appropriate branch sample (58 over 267 individuals representing 12 missing species over 91).
+For some individuals, we were unable to collect an appropriate branch sample (58 over 267 individuals representing 11 missing species over 89 accounitg for 1% of individuals present in the 10ha plot).
 
@@
 Leaf petioles, rachis and petiolules were removed for all measurements.
-Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite].
-Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time.
+Leaf fresh weights, area and thickness were measured, using a scanner **[cite]** and a microgauge **[cite]**.
+Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$, the drying process was stopped after dry leaf weights were stable in time.
 Once dried, the leaves were weighted directly after been extracted from the oven to avoid humidity bias.
@@
                          round(sd(PFT$WD, na.rm = T),2), ')')),
-  unit = c('um', 'mm2', 'mg.g-1', 'm2.kg-1', 'g.cm-3'),
+  unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
   strategy = c('Leaf defense vs. investment',
@@
 We selected 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
-Digital elevation model is issued from a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
-Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
+Digital elevation model was provided by a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
+Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) were all derived from the digital elevation model using the RSAGA package [@brenning_statistical_2008].
 Distance to the nearest potential river was computed using ArcGis [@desktop_arcgis_2011].
 
-In order to avoid multi-colinearity in future analyses, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning.
-We reduced our explanatory variables to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness to represent monsoon winds [@bunyan_effect_2015] and slope.
+In order to identify main source of environmental variation and to avoid multi-colinearity in future analyses, we reduced our abiotic explanatory variables via a principal component analysis (PCA).
+We selected less correlated variables with the most ecological meaning.
+Thus we reduced our explanatory variables to 4: curvature to represent the water flow paths and soil accumulation areas, wetness to highlight water accumulation areas, southwesterness to represent aspect [@bunyan_effect_2015] and slope.
 
-We selected two biotic factors: canopy height and basal area.
+We selected two biotic factors related to local forest structure: canopy height and basal area.
 Canopy height was calculated from a digital canopy model built with the 2013 LIDAR campaign on Uppangala plots with resolution $1*1 m^2$.
@@
 
-Traits variation was first investigated in three levels with the coefficient of variation ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species (inter-specific).
+The coefficient of variation of traits was investigated at three levels ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean): between leaves of the same individual (intra-individual), between trees of the same species (intra-specific for species with more than one individual), and between species (inter-specific).
+We performed a principal component analysis (PCA) of the mean value of traits per species.
+Subsequent community-level analysis were based on trait values averaged per species.
 
-For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
-To test the correlation and covariation among traits we performed a principal component analysis (PCA).
+### Local functional composition of communities
 
-### Local functionnal composition of communities
+Across the whole 10 hectare plot from Uppangala, 240 local communities where defined in $20*20 m^2$ quadrats.
+To investigate functional composition of communities we used a null model approach.
+A first null model (NM1) randomizes trait values among species (permuting raw in `Trait*Species` matrix).
+The second model (NM2) randomizes abundances among species inside each community but keeps the observed list of species and trait values inside communities [@bernard-verdier_community_2012].
 
-Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
-To test for pattern in functional composition of communities we used null model approach.
-A first null model (NM1) randomizes traits among species (permuting raw in `Trait*Species` matrix).
-The second model (NM2) randomizes abundances among species inside each community but keeps the observed list of species and trait values inside communities.
-NM1 allow to assess environmental filtering and biotic process on Uppangala plot species.
-NM2 bring more insight on environmental and biotic effects on species distribution inside communities [@bernard-verdier_community_2012].
+To investigate local functional composition of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010].
+CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative abundance, $p_{i}$:
+$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
+CWM was also computed in presence-absence, to further emphasize role of rare species in community trait distribution.
+Linear regressions between CWMs and biotic and abiotic factors ($CWM_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$), after stepwise selection, were compared to corresponding 999 null linear regressions from NM1 with F-statistic.
 
-To investigate traits distribution of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010].
-CWV lower than expected under a null model will be evidence for convergence of traits inside the community whereas higher CWV will show divergence in trait distribution.
-CVNND is used to detect niche differenciation patterns in case of a lower CVNND as observed under a null model.
-
-CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
-abundance, $p_{i}$:
-$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
-CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution.
-CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
+CWV for each trait $t$ was calculated as the mean distance of trait values, $t_{i}$, to $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
 abundance, $p_{i}$:
 $$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
-CVNND for each trait $t$ was calculated as nearest neigbor distances (NND) variance, $\sigma _{NND}$, of the $i$ individuals in each community, divided by its mean, $\mu _{NND}$; and NND for one individual $i$ is calculated as the minimal distance $d$ between the $j$ other individuals of the community :
-$$CVNND = \frac{\sigma _{NDD}}{\mu _{NDD}} ~ with ~ NND_{i} = min(d_{i,j})_{i \neq j} ~ (3)$$
-For both CWV and CVNND, we calculated the standardized effect size (SES) as: 
-$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
+CVNND for each trait $t$, scaled with its mean and variance, was calculated as nearest neigbor distances (NND) standard deviation, $\sigma _{NND}$, of the $i$ individuals in each community, divided by its mean, $\mu _{NND}$; and NND for one individual $i$ is calculated as the minimal distance $d$ between the $j$ other individuals of the community :
+$$CVNND = \frac{\sigma _{NND}}{\mu _{NND}} ~ , ~ ~ ~ NND = [NND_{i} = min(d_{i,j})_{i \neq j}] ~ (3)$$
+For both CWV and CVNND, proportion of of communities differing from null expectaions was tested with a two-tailed test with adjusted p-values by Benjamini-Hochberg correction.
+CWV lower than expected under a null model will be evidence for convergence of traits inside the community whereas higher CWV will show divergence in trait distribution.
+CVNND is used to detect niche differenciation patterns in case of a lower CVNND as observed under a null model [@jung_intraspecific_2010].
+We then calculated the standardized effect size (SES) as: $$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
 where $I_{obs}$ is the observed metric and $I_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.
+Effect of environment on SES was investigated by linear regression with selected biotic and abiotic factors.
+Analyses were conducted on the five functional traits and the species scores obtained on the 2 first PCA axes.
 
-Linear regressions between CWMs and biotic and abiotic factors after step measure were compared to corresponding 999 null linear regressions from NM1 with F-statistic.
-For both CWV and CVNND, proportion of of communities differing from null expectaions was tested with a two-tailed test; and links between SES and environment was investigated by linear regression.
-Analyses were conducted on the five functional traits and the two economic spectra defined with the species scores obtained on the 2 first PCA axes.
+### Floristic, functional and phylogenetic composition between communities
 
-### Functionnal and phylogenetic composition between communities
+Phylogenetic tree of our 89 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
+Phylogenetic signal was globaly assessed for each functional trait among the phylogenetic tree by Moran's I statistic.
+Positive Moran's I value indicates convergence of trait values within the phylogeny whereas negative Moran's I will show divergence.
 
-Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
-Phylogenetic signal was globaly assessed for each functional trait among the phylogenetic tree by Moran's I statistic.
-Positive Moran's I value will indicate convergence of trait value inside the phylogeny whereas negative Moran's I will show divergence.
+To investigate floristic, functional, and phylogenetic composition between communities we used $I_{st}$, for floristic composition, $U_{st}$ and $\tau _{st}$, for functional composition in abundance and presence-absence respectively, and $\beta _{st}$ and $\pi _{st}$, for phylogenetic composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
 
-$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between community pairs.
-In parrallel variation of functional composition between community pairs was assessed by $U st$ and $\tau st$ metric.
-Environmental distances between community pairs was also calculated as distances between communities on the plan composed by the two first axes of the principal component analysis of environmental variables (see suppplementary materials S1).
-Finally spatial distances between community pairs was also assessed.
-We used Mantel test to sudy the link between functionnal and phylogenetic pairwise compositions with hardy's metrics and environmental distances between community pairs.
-A second mantel test was realized to assess those pairwise links between environment and functionnal and phylogenetic structure knowing spatial structure with the spatial distances between community pairs matrix.
+Variation of composition between community pairs with those metrics was translated as a floristic, functional and phylogenetic distance, and compared to environmental and spatial distances among community pairs.
+Environmental distances were calculated as geometric distances between communities on the plan composed by the two first axes of environmental variables PCA (see supplementary material S1).
+Correlations with environment and space were assessed with two Mantel tests successively with floristic, functional, and phylogenetic community compositions.
+A firt full Mantel statistic evaluates the correlation between each metric distance matrix and the environmental distance matrix (E).
+The second partial Mantel statistic assess the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S), representing spatial residuals that can influence community composition once the environment effect taken into account.
 
-Observed mantel correlations for phylogenetic composition were compared to 999 obtained with the null model *"1s"*" suggested by @hardy_testing_2008 randomizing phylogenetic tree tips.
-Observed mantel correlations for functionnal composition were compared to NM1, randomizing traits among species.
+We used null models to test if observed correlation with Mantel statistic differed significantly from expectations under null hypothesis.
+$I_{st}$ correlation to environment and space was tested with 999 repetions of the model nul described by @myers_beta-diversity_2013 which randomizes individuals among subplots while preserving the regional abundance of species, and the number of individuals per community.
+$U st$ and $\tau st$ were tested by the nul model NM1 which randomizes traits among species.
+Finally $\beta st$ and $\pi st$ were tested against the nul model *"1s"* suggested by @hardy_testing_2008 which randomizes phylogenetic tree tips.
 
-All analysis were conducted in the R 3.3.0  free software environment for statistical analyses and graphics [@r_core_team_r:_2016] using the package 
+All analysis were conducted using the R 3.3.0  software environment [@r_core_team_r:_2016] with packages 
 `ade4` [@dray_ade4_2007], 
 `ape` [@paradis_ape:_2004], 
-`phylosignal` [@keck_phylosignal:_2015] 
-and `spacodiR` [@eastman_spacodir:_2013].
+`phylosignal` [@keck_phylosignal:_2015], 
+`spacodiR` [@eastman_spacodir:_2013],
+`vegan` [@oksanen_vegan:_2016].
 
@@
 
-Trait coefficients of variation analysis reveals a general increase of functional trait variation among scales (from intra-individual to inter-specific, see table 2).
-Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
+We found a general increase of functional trait variation from intra-individual to inter-specific level (see table 2).
+Whereas LA shows already high variabililty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
 
 Table 2: Coefficient of intra-individual, intra-specific and inter-specific trait variation.
-*Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
+*Mean of coefficient of variation for leaves in all individuals (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
 See table 1 for abbreviation.*
@@
 
-The multivariate analysis of trait correlation shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
+The principal component nalysis (PCA) shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
 The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood.
-Whereas the second PCA axis (PCA2) covary mainly with SLA representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
-PCA first axis might be associated with the WD covariation to the wood economic spectrum [WES, @chave_towards_2009].
-Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
-Our results integrate less traits (5 against 14), and less species (92 against 758), but still capture the same strategy axes.
+PCA1 axis associates WD to leaf structural traits variation linked to hydraulic conductivity efficiency associated with the wood economic spectrum [WES, @chave_towards_2009].
+Whereas the PCA2 covary mainly positively with SLA and negatively with LDMC representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
 
@@
 Figure 1: Principal component analysis (PCA) performed on species trait means.
-*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
+*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 5 components (red line).
 See table 1 for abbreviation.*
@@
 
-## Functional composition of communities
+## Local functional composition of communities
 
-Table 4: Summary of linear models and null models for traits community weighted means.
-_Linear model results for each trait community weighted means by species relative abundance after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model global and partial F-statistics were compared to 999 null model F-statistics in which traits were randomized among species.
+Community weighted means (CWM) only show few significant trends along environmental gradients (see table 3).
+CWMs reveal that local mean WD will significantly increase under South-West; whereas SLA and LA, in presence-absence only, tend to decrease with curvature.
+In addition, LDMC in presence-absence is decreasing with canopy height.
+
+Table 3: Summary of linear regressions and null models for CWMs of traits.
+_Linear regression results for each trait CWMs by species relative abundance and presence after a variable stepwise selection with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Linear regressions global and partial F-statistics were compared to 999 null linear regressions built with null model NM1 which randomizes traits among species.
 See table 1 for abbreviations.
 The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-**Second table are the result in presence-absence. They are juste here for discussion. But they probably won't be in the final report under this shape.**
 
@@
 
-Table 5: Summary of linear models for traits standardized effect size of community weighted variances.
-_Linear model results for each trait standardized effect size of community weighted variances species relative abundance after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Traits divergence is measured with community weighted variances (CWV).
+Overall, observed CWVs doesn't differ from null expectations for all traits at the exception of one community for WD (over 240 communities).
+Still we can observe some trends between standardized effect size (SES) of CWVs and environment (see table 4).
+Effectively South-West orientation will make local functional composition of communities converge towards a mean LES and LDMC whereas LA will diverge inside community under South-West orientation.
+LA and LDMC values will also diverge with increasing curvature.
+Observed CVNNDs doesn't reveal niche differenciation among communities for all traits.
+Still we can observe some trends between standardized effect size (SES) of CVNNDs and environment (see table 4).
+Species tend to differentiate their LA with lower curvature and wetness and their WD opppositely to SouthWest orientation.
+
+Table 4: Summary of linear regressions for traits standardized effect size (SES) of CWVs and CVNNDs.
+_Linear regression results for each trait SES of CWVs and CVNNDs after a variablestepwise selection with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
 See table 1 for abbreviations.
@@
 
-**First output is the result of two tails test on communities CWVs for each traits. They are juste here for discussion. But they probably won't be in the final report under this shape.**
-
-```{r CWV, fig.height=4}
+```{r CWV}
 # Observed CWVs
@@
 }, x = CWV, y = nCWV))
-CWV.pval.res <- data.frame(inf = apply(CWV.pval, 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(CWV.pval, 2, function(x){length(which(x > 0.975))/length(x)}))
+CWV.pval.res <- data.frame(inf = apply(apply(CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(apply(1 - CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}))
 ## CWV SES
@@
 CWV.ses.res <- CWV.ses.res[-1]
-CWV.ses.res$Slope <- NA
-CWV.ses.res <- CWV.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
+CWV.ses.res[,setdiff(var, names(CWV.ses.res))] <- NA
+CWV.ses.res <- CWV.ses.res[var]
 CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
@@
 # Results
-barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
+# barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
 kable(bold(CWV.ses.res))
@@
 
-Table 6: Summary of linear models for traits standardized effect size of coefficient of variation of nearest neighbour distances.
-_Linear model results for each trait standardized effect size of coefficient of variation of nearest neighbour distances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-**First output is the result of two tails test on communities CWVs for each traits. They are juste here for discussion. But they probably won't be in the final report under this shape.**
-
-```{r CVNND, fig.height=4}
+```{r CVNND}
 # Observed CVNND
@@
 row.names(CVNND.pval) <- colnames(CVNND)
+CVNND.pval <- apply(CVNND.pval, 2, p.adjust, method = 'fdr')
 CVNND.pval.res <- data.frame(inf = apply(CVNND.pval, 2, function(x){length(which(x < 0.05))/length(x)}))
@@
 CVNND.ses.res <- CVNND.ses.res[-1]
-CVNND.ses.res$Slope <- NA
-CVNND.ses.res <- CVNND.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+CVNND.ses.res[,setdiff(var, names(CVNND.ses.res))] <- NA
+CVNND.ses.res <- CVNND.ses.res[var]
 CVNND.ses.res <- apply(CVNND.ses.res, 2, as.character)
@@
 # Results
-barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
+# barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
 kable(bold(CVNND.ses.res))
@@
 
-## Composition variation between communities
+## Floristic, functional and phylogenetic composition between communities
 
-```{r PhyloMaker}
-library(phylosignal)
-library(phytools)
-library(phylobase)
-path <- system.file('phylogeny', package = 'Uppangala')
-source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
-phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
-nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
-sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), c('LatinName', 'Genus', 'Family')]
-names(sp) <- c('species', 'genus', 'family')
-results <- invisible(S.PhyloMaker(sp, phylo, nodes)) # Humboldtia_brunonis unmatched
-tree <- results$Scenario.1
-rm(phylo, results, modes, sp, S.PhyloMaker, path)
-```
+Floristic composition between communities is correlated significantly to environement spatial organization with Mantel test ($r_{I_{st},E} = 0.23$ with $p-value < 0.001$) and without any spatial residuals represented with partial Mantel test knowing environment ($r_{I_{st},S|E} = 0.01$ with $p-value > 0.1$).
 
-Table 7: Summary of Moran’s I statistic and p-values for phylogenetic signal of functional traits.
-_See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+Overall, functional composition between communities is not correlated to environment nor to space knowing the environment effect (see table 5).
+Still WD between communities, in abundance as well as in presence absence, is positively significantly correlated to environment and is even negatively correlated in presence absence to spatial structure knowing environment spatial effect.
 
-```{r Phylosignal}
-data <- PFT_sp
-row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
-tips <- row.names(data)[which(is.na(data$WD))]
-phy <- drop.tip(tree, tips)
-data <- data[phy$tip.label,c(2:5,7)]
-p4d <- phylo4d(phy,data)
-phyloS <- phyloSignal(p4d, methods = 'I') 
-phyloS <- data.frame(paste(round(phyloS$stat[,1], 3), 
-                           unlist(lapply(as.list(phyloS$pvalue[,1]), stars, ns = ' '))),
-                     row.names = row.names(phyloS$stat))
-names(phyloS) <- 'Moran\'s I'
-kable(bold(phyloS))
-rm(data, tips, p4d, phyloS, phy)
-```
+Whereas LDMC is not phylogenetically convergent nor divergent (Moran's I of 0 with a $p-value > 0.1$), LA and Thick are significantly convergent among the phylogenetic tree (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA are weakly convergent (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
+Phylogenetic composition between communities is not correlated to environement spatial organization with Mantel test ($r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor with spatial residuals represented with partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$ respectively).
 
-Table 8: Summary of Mantel test between floristic composition, environment and spatial distances between communities.
+<!-- **Table to replace by text**: Summary of Mantel test between floristic composition, environment and spatial distances between communities.
 _$Ist$  was used to investigate variation of floristic composition between communities.
@@
 S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).._ -->
 
@@
 # Result
-kable(bold(spaflo.res))
+# kable(bold(spaflo.res))
 rm(mtest, spaflo.mantel, nspaflo, nspaflo.mantel, spaflo.res)
@@
 
-Table 9: Summary of Mantel test between functionnal composition, environment and spatial distances between communities.
+Table 5: Summary of Mantel test between functional composition, environment and spatial distances between communities.
 _$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities.
@@
     mapply(function(p,q){
-      paste(format(p, scientific = T, digits = 3), 
-            Uppangala::stars(min(rank(c(p,q))[1]/(n+1), 1-rank(c(p,q))[1]/(n+1)), ns = '   '))
+      paste(round(p, 3), Uppangala::stars(min(rank(c(p,q))[1]/(n+1), 1-rank(c(p,q))[1]/(n+1)), ns = '   '))
     }, p = r, q = s, SIMPLIFY = F)
@@
 }, x = spafun.mantel, y = nspafun.mantel))
-spafun.res <- cbind(data.frame('Mantel test' = c('E', ' ', 'S|E', ' '), 'Metric' = rep(c('$U_{st}$', '$\\tau _{st}$'), 2)), spafun.res)
+spafun.res <- cbind(data.frame('Mantel statistic' = c('E', ' ', 'S|E', ' '), 'Metric' = rep(c('$U_{st}$', '$\\tau _{st}$'), 2)), spafun.res)
 
@@
 
-Table 10: Summary of Mantel test between phylogenetic composition, environment and spatial distances between communities.
+```{r PhyloMaker}
+library(phylosignal)
+library(phytools)
+library(phylobase)
+path <- system.file('phylogeny', package = 'Uppangala')
+source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
+phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
+nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
+sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), c('LatinName', 'Genus', 'Family')]
+names(sp) <- c('species', 'genus', 'family')
+results <- invisible(S.PhyloMaker(sp, phylo, nodes)) # Humboldtia_brunonis unmatched
+tree <- results$Scenario.1
+rm(phylo, results, modes, sp, S.PhyloMaker, path)
+```
+
+<!-- **Table to replace by text**: Summary of Moran’s I statistic and p-values for phylogenetic signal of functional traits.
+_See table 1 for abbreviations.
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._ -->
+
+```{r Phylosignal}
+data <- PFT_sp
+row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
+tips <- row.names(data)[which(is.na(data$WD))]
+phy <- drop.tip(tree, tips)
+data <- data[phy$tip.label,c(2:5,7)]
+p4d <- phylo4d(phy,data)
+phyloS <- phyloSignal(p4d, methods = 'I') 
+phyloS <- data.frame(paste(round(phyloS$stat[,1], 3), 
+                           unlist(lapply(as.list(phyloS$pvalue[,1]), stars, ns = ' '))),
+                     row.names = row.names(phyloS$stat))
+names(phyloS) <- 'Moran\'s I'
+# kable(bold(phyloS))
+rm(data, tips, p4d, phyloS, phy)
+```
+
+<!-- **Table to replace by text**: Summary of Mantel test between phylogenetic composition, environment and spatial distances between communities.
 _$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities.
@@
 S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._ -->
 
@@
 # Result
-kable(bold(spaphy.res))
+# kable(bold(spaphy.res))
 rm(mtest, spaphy.mantel, nspaphy, nspaphy.mantel, spaphy.res)
@@
 ## Strategy axes
+
+**To add:**
+
+* Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
+Our results integrate less traits (5 against 14), and less species (89 against 758), at a smaller scale, but still capture the same strategy axes.
+
 ## Community assembly
@@
 Supplementary material S1: Principal component analysis (PCA) performed on topgraphical variables.
-*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
+*Botom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 7 components (red line).
 SW stand for southwesterness.*

### 2016-07-29 R/draft.R

@@ -1,228 +1,136 @@
-# # Wood desnity & Dryades DB
-# wood <- read.csv('~/data/Cleaning/test/wood.csv')
-# str(wood)
-# wood <- wood[c(2,4,3,5)]
-# names(wood) <- c('ID', 'Sp', 'WD', 'Ref')
-# wood$Ref <- as.numeric(as.character(wood$Ref))
-# wood <- wood[-which(is.na(wood$Ref)),]
-# wood
-# # test <- wilcox.test(wood$WD, wood$Ref, paired = T)
-# test <- cor.test(wood$WD, wood$Ref, method = 'spearman')
-# plot(Ref ~ WD, wood, xlim = c(0,1), ylim = c(0,1), main = test$method)
-# abline(0, 1)
-# mtext(paste('p-value =', test$p.value, ', alternative hypothesis: true rho is not equal to 0'))
-
-# Env <- genEnv()
-# Topo <- stack(Env$DEM, DEMderiv(Env$DEM, choices = c('slope', 'curvature', 'plancurvature', 'aspect')))
-# Topo$Aspect <- reclassify(Topo$Aspect, c(0, 180, 1, 180, 360, 2))
-# writeRaster(Topo$Aspect, '/Users/sylvain/Desktop/Asp.tif', 'GTiff')
-# plot(reclassify(Topo$Slope, c(0, 20, 0, 20, 30, 1, 30, Inf, 2)), main = 'Slope')
-# contour(Topo$DEM, add = T)
-
-# col <- terrain.colors(7)[c(1,2,5,6,3,4)]
-# plot(raster(com, 'topo'), col = col, main = 'Topographical classes')
-# contour(Env$DEM, add = T)
-# legend('bottomleft', fill = col, legend = c('Concave surface', 'Convex plane surface', 'West steep slope', 'East steep slope', 'West mean slope', 'East mean slope'), cex = 0.7)
-# col = heat.colors(100)[as.numeric(cut(Trees$WD, breaks = 100))]
-# points(Y ~ X, Trees, col = "black", cex = `2013_girth`/100, pch = 20)
-# points(Y ~ X, Trees, cex = `2013_girth`/100, pch = 20, col = col)
-
-# plot(raster(com, 'WD'), main = 'CWMbin(WD)')
-# contour(raster(com, 'topo'), levels = c(0:6), add = T)
-
-# library(vegan)
[…]

### 2016-07-29 R/regression_table.R

@@ -35,7 +35,7 @@ regression_table <- function(LM, reg, pval, ppval){
   # Null models
   var_table$`Null model` <- round(pval,3)
   var_table$` ` <- unlist(lapply(pval, stars, ns = ' '))
-  var_table[which(var_table$` ` == ' '),2:7] <- ' '
+  var_table <- var_table[-which(var_table$` ` == ' '),]
 
   return(var_table)
 }

### 2016-08-01 inst/doc/draft.Rmd
v0.1.9025 Discussion ~ v1

@@
 
-The principal component nalysis (PCA) shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
+The principal component analysis (PCA) shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
 The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood.
@@
 
-```{r PhyloMaker}
+```{r PhyloMaker, include=FALSE}
 library(phylosignal)
@@
 
-## Strategy axes
+## Ecological strategies
 
-**To add:**
+Plants distinguish theimselves in functional trait composition betwen individuals and even more between species.
+Variation of species functional composition results from functional trade-offs that compose major variation axes of functional traits called economic spectra.
+We found back the leaf economic spectrum of @wright_worldwide_2004 opposing in our case SLA to LDMC in a trade-off between productivity, with less carbon investment for higher photosynthesis rates, opposed to persistance, with more leaf defense and investment.
+Independantly to that spectrum, we detected a second major axis of trait variation opposing WD to LA and Thick.
+This axis opposes more resistant to stress and disturbance species, thanks to denser wood, to competitive species, with thicker and larger leaves for a more efficient water conductivity.
+This spectrum correspond to the wood economic spectrum described by @chave_neutral_2004, and more precisely this opposition between small thin leave with dense wood to large thick leaves with light wood correspond to unexplained results of @baraloto_decoupled_2010.
 
-* Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
+Results are similar to the one found by @fortunel_leaf_2012 where PCA first and second axes were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
 Our results integrate less traits (5 against 14), and less species (89 against 758), at a smaller scale, but still capture the same strategy axes.
 
-## Community assembly
-## Process scales
+## Local functional composition
+
+The leaf economic specrtum is majorly affected by aspect with a convergence of value toward optimal functional traits under South-West orientation, especially for rare species with LDMC.
+Aspect, and more specifically south-west, effect in tropical forest is scarcely documented but highlighted in discussion by @gunatilleke_specieshabitat_2006.
+@bunyan_effect_2015 observed a similar effect on shola distribution of Western Ghats and suggested it could result from monsoon winds.
+Effectively monsoon winds arise fom south-west direction in this region, and bring more stressfull and disturbed conditions to species.
+Curvature is also affecting the leaf economy.
+Convex area will accumulate soil and attract water flux increasing soil fertiliy conditions for species that will invest more in leaf ressource acquisition toward a productive strategy with higher SLA and a convergent optimal LDMC.
+SLA and LDMC with curvature effect seems to reflect well soil fertility [@hodgson_is_2011].
+**LDMC rare species <- Canopy height**
+
+The wood economic spectrum defined by WD, LA, and Thick is only partially affected by environment at our 10 ha plot scale.
+Neutral hypothesis of funcional equivalence [@hubbell_neutral_2005] are most of the times sufficient to build back similar links between topography and functional compoisiton of communities.
+Still aspect wiht south-west orientation and curvature are the two main ecological factors filtering functional traits from the wood economic spectrum.
+South-west orientation, with probably more disturbance and stress, is globaly increasing community WD mean values while reducing eveness by convergence toward different optimal WD.
+In parrallel, LA is broaden with south-west direction, pehraps to answer to hydraulic diversity induced by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008].
+Fertility decrease with curvature increase results in smaller leaves inside communities with a distribution broaden by multimodality reducing eveness.
+Wetness is increasing with higher curvature and lower LA eveness, indicating that multimodality is happening more in wettest low elevation area.
+Wetness could compensate water departure trigerred by positive curvature in lower areas.
+
+## Composition between communities
+
+Limited dispersal, assumed by neutral theory, does not affect florisitic composition of communities at the 10 ha plot scale due to the absence of links  with spatial residuals.
+Florisitic composition of communities is only shaped by environment.
+On opposite, functional composition is overall not structured with environment nor with spatial residuals.
+Whereas at the community level environment is partially shapping florisitic composition, at the 10 ha level functional variation between communities is stochastic.
+Measured functional traits should vary to a larger scale [@messier_trait_2016].
+Still WD is positively correlated to $\beta$ structure.
+WD of rare species is even diverging with spatial residuals supposing an unaccounted spatially organized variable opposing close community in WD composition.
+Topography removed from spatial residuals, and climate and pedology being constant in the 10 ha plot, this unaccouted variation is assumed to be due to stochastic event in plot history as death.
+Unshowed analysis reveals that this spatial divergence of WD is not explained by tree death events from the two last decades but certainly to older events.
+Finally, variation in distance between taxon wasn't translated in phylogenetic variation whereas most of measured functional traits appeared phylogenetically conserved and clustered.
+
 ## Limits and perspectives
 
+Ecological strategies underline two econmic spectra by comparisons to previous study results.
+But the multivariate analysis of our functional traits lack trait diversity for correct interpretations of functional trade-offs.
+Measured traits were all morphological and from the leaf except for one.
+Anatomical and chemical traits as leaf vein patterns and leaf chemical content (N, P, K) could brought broader perspective on ecological strategies of Uppangala species.
+Still the study reveal that few commonly used traits can highlight well known strategy axes.
+
+Besides patterns have been detected between functional composition and environment at both local ($\alpha$) and between-community ($\beta$) scales.
+Lot of relation observed didn't differ from those obtain with null hypothesis reflecting neutral theory [as functional equivalence, @hubbell_neutral_2005].
+Ecological gradient and spatial scale sizes can be at the origin of this lack of environmental filtering.
+Effectively the abiotic variation is mainly topographic with a constant climate and pedology.
+Study highlighting environmental filtering often present a higher environmental variation [@bernard-verdier_community_2012].
+Similar to this, the 10 ha plot scale is really fine and partially responsible for the small environmental variation.
+@kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25 ha tropical forest plots including large flat valleys.
+This mid-low elevation study of functional traits from western ghats forest could be reproduced at higher elevation or in dryer areas as eastern Ghats for better insight of environmental filtering at broad scale.
+On the other hand WD as shown to be plastic to environment at fine scale both in $\alpha$ and $\beta$ levels.
+Species WD seems highly adptated to environment in western Gahts forests in link to monsoon events.
+
 # *Acknowledgement*
+
+French Institute of Pondicherry (FIP)
 

### 2016-08-01 R/draft.R

@@ -134,3 +134,29 @@
 # plot(as(com, 'SpatialPolygons'), add = T, border = 'grey')
 # text(coordinates(com), com$id, col = rainbow(n)[com$group], cex = 0.7)
 
+# ## Light to soil and canopy height ##
+# Canopy <- raster('./inst/extdata/Canopy.tif')
+# terrain_rays <- read.csv("C:/Users/sylvain/Desktop/terrain_rays.csv", header=FALSE)
+# light <- terrain_rays[c(1,2,4)]
+# names(light) <- c('x','y','light')
+# light <- rasterFromXYZ(light)
+# Canopy <- crop(Canopy, extent(light))
+# Canopy <- resample(Canopy, light)
+# Comp <- stack(Canopy, light)
+# plot(Comp)
+# summary(lm(values(light) ~ values(Canopy)))
+# plot(values(light) ~ values(Canopy))
+
+# ## LDMC - Canopy ##
+# com <- row.names(which(R == max(R$Canopy), arr.ind = T))
+# hist(CWM$`presence-absence`$LDMC)
+# CWM$`presence-absence`$LDMC[com]
+# comp <- colnames(L$`presence-absence`[com,L$`presence-absence`[com,] > 0])
+# Species$LatinName[match(comp, Species$SpCode)]
+# barplot(t(Q[match(comp, row.names(Q)),][5]), las = 2)
+# plot(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)]), xlab = 'Strata', ylab = 'LDMC')
+# abline(lm(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)])), col = 'red')
+# mtext(paste('LDMC ~ Strata***, R2 =', round(summary(lm(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)])), col = 'red')$r.squared, 3)))
+# plot(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)]), xlab = 'Strata', ylab = 'LDMC')
+# abline(lm(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)])), col = 'red')
+# mtext(paste('LDMC ~ Strata***, R2 =', round(summary(lm(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)])), col = 'red')$r.squared, 3)))

### 2016-08-03 inst/doc/draft.Rmd
v0.1.9026 Intro beginning

@@
 ---
-title: 'Uppangala tree functional traits: Article draft'
+title: 'Seasonal filtering at community scale of a tropical evergreen forest from Western Ghats, India'
 author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
@@
 
-Keywords: functional trait, tropical evergreen forest, Western Ghats, India...
+## Keywords
+
+functional trait; tropical evergreen forest; Western Ghats; India...
 
@@
 
-**Study scheme:**
+Ecology offers different theories describing community composition of tropical forest.
+Niche theory invoke ecological filtering responsible for species assembly inside communities [@chave_comparing_2002].
+On the other hand neutral theory [@hubbell_neutral_2005] assume functional equivalence of species, implying species distributions shaped by stochastic events as dispersion and death.
+@lortie_rethinking_2004 offered a conceptual framewok that reconciliates the two approach by defining integrated community concept.
+Integrated communities are build on three levels: $\gamma$ diversity as the initial regional species pool filtered by stochastic neutral events, $\beta$ diversity between communities resulting from abiotic filtering and stochastic events among communities, and $\alpha$ diversity defining the community assembly after local abiotic and biotic filtering.
+
+Communities are composed from species spanning a range of functional traits responsible for their growth, survival and reproduction.
+Functional traits can thus be linked to trade-offs representative of species ecological strategies.
+For instance, specific leaf area varies along a soil fertility - shade index representative of a trade-off between ressource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011].
+Due to common ecological strategies, functional traits will covary in so called economic spectra.
+@wright_worldwide_2004 showed the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009.
+The leaf economic spectrum opposes productive ecological strategy to conservative persistent species.
+The wood economic spectrum is separating species between three pole close to @grime_evidence_1977 primary strategy axes: (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength.
+A debated question is links between species economic spectra and strategy axes.
+Whereas @baraloto_decoupled_2010 highlighted decoupled economic spectra allowing tree species to adjust differently their ecological strategies; @reich_world-wide_2014 advocate more for a whole plant economic spectrum identified by @freschet_evidence_2010 opposing fast growing species to more conservative ones.
+
+Stochastic process and environmental filtering will be highly scale dependent.
+Effectively some functional traits will be more variable at the individual level with plasticity to micro-environmental conditions whereas others will be more affected on global environmental gradient with phylogeny.
+For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they question the existence of the leaf economic spectrum at community scale.
+
+<!-- Tropical evergreen forest are composed by tree keeping their foliage all year around.
+Yet species will observe leaves fall events ruled by their leaf lifespan.
+Leaf lifespan will tend to decrease in tropical evergreen forest seasonally dry [@poorter_leaf_2009; @prior_leaf_2003].
+-->
+
++ *Move to India Western Ghats context*
++ *No study of community ecology with functional traits*
++ *Peculiar case of highly seasonal moist evergreen forest*
++ *Move to previous questionning*
++ *Quid from ecological strategies at small scale in Uppangala forest*
++ *Quid from partitionning between stochastic event and environmental filtering on community structure*
 
@@
     ii. Do functional composition variation between communities ($\beta$ scale) reflect the changing of abiotic and biotic conditions ?
-
-**To add:**
-
-* Functional traits can be linked to trade-offs in tree strategy for functionning, survival and reproduction [@baraloto_decoupled_2010, and see table 1], which vary depending on environment condition.
-* For instance, SLA varies along a soil fertility - shade index whereas LDMC is basically linked to soil fertility [@hodgson_is_2011]. **! In european temperate context !**
-* Leaf investment definition
-* Leaf ressource acquisition definition
 
@@
 Leaf petioles, rachis and petiolules were removed for all measurements.
-Leaf fresh weights, area and thickness were measured, using a scanner **[cite]** and a microgauge **[cite]**.
+Leaf fresh weights, area and thickness were measured, using a bend scanner and a micrometer.
 Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$, the drying process was stopped after dry leaf weights were stable in time.
@@
 This axis opposes more resistant to stress and disturbance species, thanks to denser wood, to competitive species, with thicker and larger leaves for a more efficient water conductivity.
-This spectrum correspond to the wood economic spectrum described by @chave_neutral_2004, and more precisely this opposition between small thin leave with dense wood to large thick leaves with light wood correspond to unexplained results of @baraloto_decoupled_2010.
+This spectrum correspond to the wood economic spectrum described by @chave_towards_2009, and more precisely this opposition between small thin leave with dense wood to large thick leaves with light wood correspond to unexplained results of @baraloto_decoupled_2010.
 

### 2016-08-03 R/draft.R

@@ -160,3 +160,31 @@
 # plot(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)]), xlab = 'Strata', ylab = 'LDMC')
 # abline(lm(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)])), col = 'red')
 # mtext(paste('LDMC ~ Strata***, R2 =', round(summary(lm(Q$LDMC ~ as.numeric(Species$Strata[match(row.names(Q), Species$SpCode)])), col = 'red')$r.squared, 3)))
+
+# ## RLQ bis ##
+# library(ade4)
+# Q <- Q[-which(is.na(Q$WD)),-c(1,3)]
+# L <- L$abundance[row.names(Q)]
+# L.coa <- dudi.coa(L, scannf = F)
+# R.pca <- dudi.pca(R, scannf = F, row.w = L.coa$lw)
+# Q.pca <- dudi.pca(Q, scannf = F, row.w = L.coa$cw)
+# rlq <- rlq(R.pca, L.coa, Q.pca, scannf = T)
+# rlq$l1
+# s.label(rlq$mQ, xax = 4, yax = 5, boxes = F)
+# s.arrow(rlq$l1, xax = 4, yax = 5, add.plot = T)
+# FamilyFact <- as.factor(Species$Family[match(row.names(rlq$mQ), Species$SpCode)])
+# col <- rainbow(length(levels(FamilyFact)))[as.numeric(FamilyFact)]
+# s.class(rlq$mQ, FamilyFact, clabel = 0, xax = 4, yax = 5, cellipse = 0, label = rep(' ',72), col = col, add.plot = T)
+# legend('topleft', levels(FamilyFact), fill = col)
+# selsp <- c('anmo', 'baco', 'blse', 'desi', 'dila', 'mrin')
+# col = rep(1, length(Q$LDMC))
+# col[match(selsp, row.names(Q))] <- 2
+# barplot(Q$LDMC, col = c('grey','firebrick')[col], las = 2, names.arg = row.names(Q))
+# s.label(rlq$mR, xax = 4, yax = 5, boxes = F)
+# s.arrow(rlq$l1, xax = 4, yax = 5, add.plot = T)
+#
+# L['C224',which(L['C224',] > 0)]
[…]

### 2016-08-04 inst/doc/draft.Rmd
v0.1.9027 LDMC ~ Canopy

@@
 ---
-title: 'Seasonal filtering at community scale of a tropical evergreen forest from Western Ghats, India'
+title: Functional assembly into and among community of a tropical evergreen forest from Western Ghats, India
 author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
@@
 library(parallel)
+cores <- detectCores() - 1
 library(Uppangala)
@@
 Due to common ecological strategies, functional traits will covary in so called economic spectra.
-@wright_worldwide_2004 showed the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009.
+@wright_worldwide_2004 discovered the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009.
 The leaf economic spectrum opposes productive ecological strategy to conservative persistent species.
@@
 
-<!-- Tropical evergreen forest are composed by tree keeping their foliage all year around.
-Yet species will observe leaves fall events ruled by their leaf lifespan.
-Leaf lifespan will tend to decrease in tropical evergreen forest seasonally dry [@poorter_leaf_2009; @prior_leaf_2003].
--->
-
 + *Move to India Western Ghats context*
-+ *No study of community ecology with functional traits*
-+ *Peculiar case of highly seasonal moist evergreen forest*
+    - *No previous study of community ecology with functional traits*
+    - *Peculiar case of highly seasonal moist evergreen forest*
 + *Move to previous questionning*
@@
 PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
-Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')]
-row.names(Q) <- PFT_sp$SpCode
 rm(PCAscores, PFT_ind, SpCodeT, lambda, pca, pca.var, Rot, XLIM)
@@
 L[is.na(L)] <- 0
+Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')]
+row.names(Q) <- PFT_sp$SpCode
 sel <- intersect(row.names(Q), names(L))
@@
 nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
-cl <- makeCluster(3)
+cl <- makeCluster(cores)
 clusterExport(cl, list('L', 'nQ'))
@@
 
-cl <- makeCluster(3)
+cl <- makeCluster(cores)
 clusterExport(cl, list('nL', 'CWM', 'Q'))
@@
 # Simulated CVNNDs 
-cl <- makeCluster(3)
+cl <- makeCluster(cores)
 clusterExport(cl, list('nQ', 'L'))
@@
 nL <- vegan::permatfull(L$abundance, fixedmar = 'both', mtype = 'count', times = n, shuffle = 'both')
-cl <- makeCluster(3)
+cl <- makeCluster(cores)
 clusterExport(cl, list('nL', 'spacodi.calc', 'Me', 'Ms'))
@@
 # Simulated mantel
-cl <- makeCluster(3)
+cl <- makeCluster(cores)
 clusterExport(cl, list('nQ', 'L', 'Me', 'Ms', 'spacodi.calc'))
@@
 nphy <- rep(list(phy), 999)
-cl <- makeCluster(3)
+cl <- makeCluster(cores)
 clusterExport(cl, list('nphy', 'L', 'spacodi.calc', 'Ntip', 'resamp.phy', 'branching.times'))
@@
 
-The leaf economic specrtum is majorly affected by aspect with a convergence of value toward optimal functional traits under South-West orientation, especially for rare species with LDMC.
+The leaf economic spectrum is majorly affected by aspect with a convergence of value toward optimal functional traits under South-West orientation, especially for rare species with LDMC.
 Aspect, and more specifically south-west, effect in tropical forest is scarcely documented but highlighted in discussion by @gunatilleke_specieshabitat_2006.
 @bunyan_effect_2015 observed a similar effect on shola distribution of Western Ghats and suggested it could result from monsoon winds.
-Effectively monsoon winds arise fom south-west direction in this region, and bring more stressfull and disturbed conditions to species.
+Effectively monsoon winds arise fom south-west direction in this region, and bring more stressfull and disturbed conditions to species [@soman_aspects_1990].
 Curvature is also affecting the leaf economy.
@@
 SLA and LDMC with curvature effect seems to reflect well soil fertility [@hodgson_is_2011].
-**LDMC rare species <- Canopy height**
+
+Biotic factors also affect the LES through the negative effect of canopy height on LDMC.
+The canopy height effect in fact is null on understorey species and only apply on species that take part in the canopy composition or emerge (results unshown).
+We could have assumed that understorey species would be investing in higher LDMC for an higher leaf lifespan under shade stress [@kitajima_tissue-level_2010].
+But both the high variation of canopy height resulting in a relatively light open canopy and the dry season of 3.8 months suggest that leaf lifespan investment is not so obvious.
+Instead of a growth-survival trade-off shade tolerant species may adopt a more low-high light growth trade-off strategy [@poorter_leaf_2009] strongly affecting SLA but not driving LDMC.
+Emergent and high canopy species will have a lower LDMC, resulting from a more productive strategy toward leaf resource acquisition at the expense of leaf carbon investment in defense and resistance.
+Thus canopy species with the lowest LDMC will grow the fastest and then reach the highest canopy.
+Measurement of LDMC on species classified as canopy and emergent are realized most of the time on shadow leaves from understorey individuals.
+Still we observed previous relation indicating that species phylogenetic determinism is stronger than micro-environmental variation plasticity for LDMC, as suggested by the intra-specific variation of LDMC 1.5 times bigger than intra-individuals variation.
 
@@
 In parrallel, LA is broaden with south-west direction, pehraps to answer to hydraulic diversity induced by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008].
-Fertility decrease with curvature increase results in smaller leaves inside communities with a distribution broaden by multimodality reducing eveness.
+Fertility decrease, with curvature increase, results in smaller leaves inside communities with a distribution broaden by multimodality reducing eveness.
 Wetness is increasing with higher curvature and lower LA eveness, indicating that multimodality is happening more in wettest low elevation area.
-Wetness could compensate water departure trigerred by positive curvature in lower areas.
+Wetness could compensate water departure trigerred by positive curvature in lower elevation areas.
 
@@
 
-Limited dispersal, assumed by neutral theory, does not affect florisitic composition of communities at the 10 ha plot scale due to the absence of links  with spatial residuals.
+Limited dispersal, assumed by neutral theory [@condit_beta-diversity_2002], does not affect florisitic composition of communities at the 10 ha plot scale due to the absence of links  with spatial residuals.
 Florisitic composition of communities is only shaped by environment.
-On opposite, functional composition is overall not structured with environment nor with spatial residuals.
-Whereas at the community level environment is partially shapping florisitic composition, at the 10 ha level functional variation between communities is stochastic.
+On opposite, functional composition is overall not structured with environment nor with spatial residuals.At the 10 ha level functional variation between communities is stochastic; whereas at the community level environment is partially shapping florisitic composition.
 Measured functional traits should vary to a larger scale [@messier_trait_2016].
 Still WD is positively correlated to $\beta$ structure.
+**To developp**.
 WD of rare species is even diverging with spatial residuals supposing an unaccounted spatially organized variable opposing close community in WD composition.
@@
 
-Ecological strategies underline two econmic spectra by comparisons to previous study results.
-But the multivariate analysis of our functional traits lack trait diversity for correct interpretations of functional trade-offs.
+Ecological strategies underline two econmic spectra as shown by @baraloto_decoupled_2010.
+Whereas the multivariate analysis of our functional traits lack trait diversity for correct interpretations of functional trade-offs.
 Measured traits were all morphological and from the leaf except for one.

### 2016-08-04 R/draft.R

@@ -188,3 +188,13 @@
 # bca$c1
 # biplot(bca, xax = 5, yax = 3)
 # randtest(bca)
+
+PFT_sp0 <- PFT_sp
+PFT_sp <- PFT_sp0
+sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] %in% c('ESII'))
+PFT_sp <- PFT_sp[sel,]
+par(mfrow = c(1,2))
+hist(PFT_sp$LDMC[sel], xlim = range(PFT_sp$LDMC), col = 'firebrick', xlab = 'LDMC', main = "ESI species LDMC")
+hist(PFT_sp$LDMC[sel], xlim = range(PFT_sp$LDMC), col = 'blue', xlab = 'LDMC', main = "ESII species LDMC")
+# CWMII <- CWM
+# summary(CWM.lm$abundance$LDMC)

### 2016-08-05 inst/doc/draft.Rmd
v0.1.9028 Intro + Schema discussion

@@
 ---
-title: Functional assembly into and among community of a tropical evergreen forest from Western Ghats, India
+title: Functional assembly into and among community of a tropical evergreen forest
+  from Western Ghats, India
 author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
-output: pdf_document
+date: '`r Sys.Date()`'
+output: word_document
 bibliography: refs.bib
@@
 ```
+
+Built with `Uppangala` package version `r packageVersion('Uppangala')`.
 
@@
 Whereas @baraloto_decoupled_2010 highlighted decoupled economic spectra allowing tree species to adjust differently their ecological strategies; @reich_world-wide_2014 advocate more for a whole plant economic spectrum identified by @freschet_evidence_2010 opposing fast growing species to more conservative ones.
-
 Stochastic process and environmental filtering will be highly scale dependent.
@@
 
-+ *Move to India Western Ghats context*
-    - *No previous study of community ecology with functional traits*
-    - *Peculiar case of highly seasonal moist evergreen forest*
-+ *Move to previous questionning*
-+ *Quid from ecological strategies at small scale in Uppangala forest*
-+ *Quid from partitionning between stochastic event and environmental filtering on community structure*
+Functional traits in community ecology are linking ecological strategies, community assembly theory and functional diversity [@westoby_land-plant_2006].
+In this paper we present the use of traits-based approach combined to community ecology to encompass functional assembly into and among community of a tropical evergreen forest from Western Ghats, India.
+Few studies investigated functional ecology whithin the Western Ghats indian biodiversity hotspot [@myers_biodiversity_2000].
+Western Ghats present atypic moist tropical evergreen forest though seasonnaly dry and afffected by strong monsoon events.
+Such environmental conditions can be highly interesting to discuss and evaluate community assembly theory resulting in functional assembly of those forests.
+To do so we organised our study framework in two main questions :
 
@@
 ```{r 1 Traits}
-rm(list = ls()); invisible(gc())
+rm(list = ls()[-which(ls() == "cores")]); invisible(gc())
 PFT <- genPFT()
 table <- data.frame(
-  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
-  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
-  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
+  Trait = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
+  Abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
+  Range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
             paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
@@
             paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
-  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
+  m = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
                          round(sd(PFT$Thick, na.rm = T)), ')'),
@@
                          round(sd(PFT$WD, na.rm = T),2), ')')),
-  unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
-  strategy = c('Leaf defense vs. investment',
+  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
+  s = c('Leaf defense vs. investment',
                'Leaf ressource capture vs.investment',
@@
   )
-kable(table,  label="traits", row.names = F)
+names(table)[c(4,6)] <- c('Mean (sd)', 'Trade-off')
+kable(table, row.names = F)
 rm(table)
@@
 
-To investigate local functional composition of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010].
+To investigate local functional composition of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010].
 CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative abundance, $p_{i}$:
@@
 `spacodiR` [@eastman_spacodir:_2013],
-`vegan` [@oksanen_vegan:_2016].
+and `vegan` [@oksanen_vegan:_2016].
 
@@
 Community weighted means (CWM) only show few significant trends along environmental gradients (see table 3).
-CWMs reveal that local mean WD will significantly increase under South-West; whereas SLA and LA, in presence-absence only, tend to decrease with curvature.
+CWMs reveal that local mean WD will significantly increase under south-west; whereas SLA and LA, in presence-absence only, tend to decrease with curvature.
 In addition, LDMC in presence-absence is decreasing with canopy height.
@@
 Table 3: Summary of linear regressions and null models for CWMs of traits.
-_Linear regression results for each trait CWMs by species relative abundance and presence after a variable stepwise selection with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+_Linear regression results for each trait CWMs weighted by species relative abundance (AB) and presence-absence (PA) after a variable stepwise selection with slope, curvature, wetness, southwesterness (SW), canopy height (Canopy) and basal area (BA).
 Linear regressions global and partial F-statistics were compared to 999 null linear regressions built with null model NM1 which randomizes traits among species.
@@
 # Result tables
-kable(bold(regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]])))
-kable(bold(regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]])))
-rm(CWM.lm, nCWM.lm, pval, ppval, reg) ; invisible(gc())
+t1 <- regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]])
+t1 <- cbind(c('AB', rep(' ', length(t1[,1])-1)), t1)
+names(t1)[1] <- ' '
+t2 <- regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]])
+t2 <- cbind(c('PA', rep(' ', length(t2[,2])-1)), t2)
+names(t2)[1] <- ' '
+kable(bold(rbind(t1,t2)), row.names = F)
+rm(CWM.lm, nCWM.lm, pval, ppval, reg, t1, t2) ; invisible(gc())
 ```
@@
 Still we can observe some trends between standardized effect size (SES) of CWVs and environment (see table 4).
-Effectively South-West orientation will make local functional composition of communities converge towards a mean LES and LDMC whereas LA will diverge inside community under South-West orientation.
+Effectively south-west orientation will make local functional composition of communities converge towards a mean LES and LDMC whereas LA will diverge inside community under south-west orientation.
 LA and LDMC values will also diverge with increasing curvature.
@@
 Table 4: Summary of linear regressions for traits standardized effect size (SES) of CWVs and CVNNDs.
-_Linear regression results for each trait SES of CWVs and CVNNDs after a variablestepwise selection with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
+_Linear regression results for each trait SES of CWVs and CVNNDs after a variable stepwise selection with slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
 See table 1 for abbreviations.
@@
 # barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
-kable(bold(CWV.ses.res))
-rm(CWV.pval, CWV.pval.res, CWV.ses.res, CWV.ses.lm, nCWV)
+CWV.ses.res <- cbind(c('CWV', rep(' ', length(CWV.ses.res[,1])-1)), row.names(CWV.ses.res), CWV.ses.res)
+names(CWV.ses.res)[1:2] <- c(' ', 'Trait')
+rm(CWV.pval, CWV.pval.res, CWV.ses.lm, nCWV)
 ```
@@
 # barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
-kable(bold(CVNND.ses.res))
+CVNND.ses.res <- cbind(c('CVNND', rep(' ', length(CVNND.ses.res[,1])-1)), row.names(CVNND.ses.res), CVNND.ses.res)
+names(CVNND.ses.res)[1:2] <- c(' ', 'Trait')
+kable(bold(rbind(CWV.ses.res,CVNND.ses.res)), row.names = F)
 rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
@@
 Whereas LDMC is not phylogenetically convergent nor divergent (Moran's I of 0 with a $p-value > 0.1$), LA and Thick are significantly convergent among the phylogenetic tree (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA are weakly convergent (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
-Phylogenetic composition between communities is not correlated to environement spatial organization with Mantel test ($r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor with spatial residuals represented with partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$ respectively).
+Phylogenetic composition between communities is not correlated to environement spatial organization with Mantel test ($r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor with spatial residuals represented with partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).
 
@@
 This axis opposes more resistant to stress and disturbance species, thanks to denser wood, to competitive species, with thicker and larger leaves for a more efficient water conductivity.
-This spectrum correspond to the wood economic spectrum described by @chave_towards_2009, and more precisely this opposition between small thin leave with dense wood to large thick leaves with light wood correspond to unexplained results of @baraloto_decoupled_2010.
+This spectrum correspond to the wood economic spectrum described by @chave_towards_2009, and more precisely this opposition between small thin leave with dense wood to large thick leaves with light wood correspond to the additional gradient uncovered by  @baraloto_decoupled_2010.
 
@@
 
-The leaf economic spectrum is majorly affected by aspect with a convergence of value toward optimal functional traits under South-West orientation, especially for rare species with LDMC.
+The leaf economic spectrum is majorly affected by aspect with a convergence of value toward optimal functional traits under south-west orientation, especially for LDMC of rare species.
 Aspect, and more specifically south-west, effect in tropical forest is scarcely documented but highlighted in discussion by @gunatilleke_specieshabitat_2006.
@@
 Biotic factors also affect the LES through the negative effect of canopy height on LDMC.
-The canopy height effect in fact is null on understorey species and only apply on species that take part in the canopy composition or emerge (results unshown).
-We could have assumed that understorey species would be investing in higher LDMC for an higher leaf lifespan under shade stress [@kitajima_tissue-level_2010].
+The canopy height effect is null on understorey species and only apply on species that take part in the canopy composition or emerge (results unshown).
+We would assumed that understorey species would be investing in higher LDMC for an higher leaf lifespan under shade stress [@kitajima_tissue-level_2010].
 But both the high variation of canopy height resulting in a relatively light open canopy and the dry season of 3.8 months suggest that leaf lifespan investment is not so obvious.
@@
 Still aspect wiht south-west orientation and curvature are the two main ecological factors filtering functional traits from the wood economic spectrum.
-South-west orientation, with probably more disturbance and stress, is globaly increasing community WD mean values while reducing eveness by convergence toward different optimal WD.
-In parrallel, LA is broaden with south-west direction, pehraps to answer to hydraulic diversity induced by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008].
-Fertility decrease, with curvature increase, results in smaller leaves inside communities with a distribution broaden by multimodality reducing eveness.
+South-west orientation, with probably more disturbance and stress from monsoon, result in denser wood with reduced eveness by convergence toward different optimal density.
+In parrallel, south-west direction broaden leaf size strategies, pehraps to answer to hydraulic diversity induced by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008].
+Fertility decrease, with concavity, results in smaller leaves inside communities with a distribution broaden by multimodality reducing eveness.
 Wetness is increasing with higher curvature and lower LA eveness, indicating that multimodality is happening more in wettest low elevation area.
@@
 
-Limited dispersal, assumed by neutral theory [@condit_beta-diversity_2002], does not affect florisitic composition of communities at the 10 ha plot scale due to the absence of links  with spatial residuals.
+Limited dispersal, assumed by neutral theory [@condit_beta-diversity_2002], does not affect florisitic composition of communities at the 10 ha plot scale revealed by the absence of links  with spatial residuals.
 Florisitic composition of communities is only shaped by environment.
-On opposite, functional composition is overall not structured with environment nor with spatial residuals.At the 10 ha level functional variation between communities is stochastic; whereas at the community level environment is partially shapping florisitic composition.
+On opposite, functional composition is overall not structured with environment nor with spatial residuals.
+At the 10 ha level functional variation between communities is stochastic; whereas at the community level environment is partially shapping florisitic composition.
 Measured functional traits should vary to a larger scale [@messier_trait_2016].
-Still WD is positively correlated to $\beta$ structure.
-**To developp**.
+Still environment is shaping community resistance to disturbance and stress by affecting wood density $\beta$ structure at fine scale.
 WD of rare species is even diverging with spatial residuals supposing an unaccounted spatially organized variable opposing close community in WD composition.
@@
 Measured traits were all morphological and from the leaf except for one.
-Anatomical and chemical traits as leaf vein patterns and leaf chemical content (N, P, K) could brought broader perspective on ecological strategies of Uppangala species.
+Anatomical and chemical traits as leaf vein patterns and leaf chemical content (N, P, K) could brought broader perspective on ecological strategies of Western Ghats species.
 Still the study reveal that few commonly used traits can highlight well known strategy axes.
@@
 Study highlighting environmental filtering often present a higher environmental variation [@bernard-verdier_community_2012].
-Similar to this, the 10 ha plot scale is really fine and partially responsible for the small environmental variation.
+In parallel, the 10 ha plot scale is really low and partially responsible for the small environmental variation.
 @kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25 ha tropical forest plots including large flat valleys.
-This mid-low elevation study of functional traits from western ghats forest could be reproduced at higher elevation or in dryer areas as eastern Ghats for better insight of environmental filtering at broad scale.
-On the other hand WD as shown to be plastic to environment at fine scale both in $\alpha$ and $\beta$ levels.
-Species WD seems highly adptated to environment in western Gahts forests in link to monsoon events.
+This mid-low elevation study of functional traits from Western Ghats forest could be reproduced at higher elevation or in dryer areas as Eastern Ghats for better insight of environmental filtering at broad scale.
+On the other hand, WD as shown to be filtered by environment at fine scale both in $\alpha$ and $\beta$ levels.
+Species wood seems highly adpted to environment in Western Gahts forests in link to monsoon events.
 

### 2016-08-08 inst/doc/draft.Rmd
v0.1.9029 Corrected intro

@@
 ---
-title: Functional assembly into and among community of a tropical evergreen forest
-  from Western Ghats, India
+title: 'Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India : vertical stratification reflects guild-depedent dynamics'
 author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
 date: '`r Sys.Date()`'
-output: word_document
+output: pdf_document
 bibliography: refs.bib
@@
 
-functional trait; tropical evergreen forest; Western Ghats; India...
+functional trait; rainforest; dry season; forest strata; Western Ghats; ...
 
@@
 
-Ecology offers different theories describing community composition of tropical forest.
-Niche theory invoke ecological filtering responsible for species assembly inside communities [@chave_comparing_2002].
-On the other hand neutral theory [@hubbell_neutral_2005] assume functional equivalence of species, implying species distributions shaped by stochastic events as dispersion and death.
-@lortie_rethinking_2004 offered a conceptual framewok that reconciliates the two approach by defining integrated community concept.
-Integrated communities are build on three levels: $\gamma$ diversity as the initial regional species pool filtered by stochastic neutral events, $\beta$ diversity between communities resulting from abiotic filtering and stochastic events among communities, and $\alpha$ diversity defining the community assembly after local abiotic and biotic filtering.
+Understanding the determinants of species occurrence and co-occurrence is a long standing issue in ecology [@diamond_ecology_1975].
+Tropical rainforests have especially fascinated standing issue in ecology [@connell_diversity_1978].
+Two concurrent theories have been proposed to explain how species coexistence can be maintained in ecological communities.
+Niche theory states that the niche differences of species determine wether they can establish and coexist in communities [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that species coexistence is possible even when species are functionaly equivalent, under the sole influence of stochasticity in life, death, reproduction and dispersal [@hubbell_unified_2001].
+Both theory can provide realistic expectations of taxonomic diversity [@chave_comparing_2002], and some recent approaches have advocated investigating patterns of functional [@westoby_land-plant_2006; @mcgill_rebuilding_2006] and phylogenetic diversity [@mouquet_ecophylogenetics:_2012] to trace the signature of ecological processes related to specific biological attributes and adaptations.
+These approaches have been applied to resolve the conundrum og high tree diversity in tropical forests [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012], but studies on these ecosystems still remain scarce.
+We here investigate taxonomic, functional and phylogenetic diversity in a low elevation rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], where no study has addressed the relative influence of neutral and niche-based dynamics on local forest composition.
 
-Communities are composed from species spanning a range of functional traits responsible for their growth, survival and reproduction.
-Functional traits can thus be linked to trade-offs representative of species ecological strategies.
-For instance, specific leaf area varies along a soil fertility - shade index representative of a trade-off between ressource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011].
-Due to common ecological strategies, functional traits will covary in so called economic spectra.
-@wright_worldwide_2004 discovered the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009.
-The leaf economic spectrum opposes productive ecological strategy to conservative persistent species.
-The wood economic spectrum is separating species between three pole close to @grime_evidence_1977 primary strategy axes: (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength.
-A debated question is links between species economic spectra and strategy axes.
-Whereas @baraloto_decoupled_2010 highlighted decoupled economic spectra allowing tree species to adjust differently their ecological strategies; @reich_world-wide_2014 advocate more for a whole plant economic spectrum identified by @freschet_evidence_2010 opposing fast growing species to more conservative ones.
-Stochastic process and environmental filtering will be highly scale dependent.
-Effectively some functional traits will be more variable at the individual level with plasticity to micro-environmental conditions whereas others will be more affected on global environmental gradient with phylogeny.
-For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they question the existence of the leaf economic spectrum at community scale.
+The variation of functional traits among species reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007].
+Thes variation thus reflect fundamental trade-offs in ecological strategies along environmental gradients [@mark_westoby_plant_2002].
+For instance, specific leaf area varies along a soil fertility-shade index and represent a trade-off between resource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011].
+Due to common ecological strategies, functional traits will covary along basic leaf [@wright_worldwide_2004] and wood [@chave_towards_2009] economic spectra.
+The leaf economic spectrum opposes productive ecological strategy to conservative persistent strategy.
+The wood economic spectrum distinguish species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977: (i) competitive species, wiht high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength.
+@baraloto_decoupled_2010 highlighted decoupled wood and leaf strategies allowing independant variation of these strtegies, but @reich_world-wide_2014 conversely advocated that the two specrta have to integrated in a comprehensive whole plant economic spectrum [@freschet_evidence_2010] opposing fast growing species to more conservative ones.
+Here we address how varying wood and leaf ecological strategies determine species community assembly in an Indian rainforest.
+Therefore a multiscale framework is needed to represent the hierarchy of ecological processes from regional to local scale [@lortie_rethinking_2004].
+Determining the extent of neutral and niche-based dynamics then implies (i) decomposing diversity patterns within and among communities [@ackerly_trait-based_2007], and (ii) designing appropriate statistical tests to disentangle the influence of local and larger-scale dynamics [@bernard-verdier_community_2012].
+We will decompose diversity patterns in a large rainforest plot into sources of variation within and between communities [@kraft_disentangling_2011], and investigate the fine-scale and plot-scale dynamics using appropriate null models.
 
-Functional traits in community ecology are linking ecological strategies, community assembly theory and functional diversity [@westoby_land-plant_2006].
-In this paper we present the use of traits-based approach combined to community ecology to encompass functional assembly into and among community of a tropical evergreen forest from Western Ghats, India.
-Few studies investigated functional ecology whithin the Western Ghats indian biodiversity hotspot [@myers_biodiversity_2000].
-Western Ghats present atypic moist tropical evergreen forest though seasonnaly dry and afffected by strong monsoon events.
-Such environmental conditions can be highly interesting to discuss and evaluate community assembly theory resulting in functional assembly of those forests.
-To do so we organised our study framework in two main questions :
+Apart from scale-dependent dynamics in space whithin and among communities, ecological guilds can coexist and occupy distinct niches within communities, while the species within guilds are ecologically similar [@wilson_guilds_1999; @vergnon_niches_2009].
+The neutral theory has been basically applied in rainforests to adress diversity patterns in the guild of canopy trees [@hubbell_unified_2001], thus ignoring dynamics of understorey trees.
+However basic ecological constraints drive resource use strategies and biotic interactions in this ecosystems and imply considering functional strategies and relationship among guilds as well.
+Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988].
+It thus provide an ideal context to analyze guilod-dependent dynamics and ecological strategies in tropical rainforests.
+We will thus address diversity patterns and community assembly at guild-level.
+
+In this paper, we used a trait-based approach to adress guild and community dynamics within and among local tree assemblages in a tropical rain forest from Western Ghats, India.
+Although of wet evergreen type, the forest undergoes a long dry season and an intense rainy season under the influence of monsoon.
+In this context, we expect water stress and competition for resource to be key drivers of forest dynamics.
+We thus adressed two main questions:
 
@@
 
-The study was carried out in the the 10-ha "Uppangala plot" of undisturbed, old-growth wet evergreen Diptrocarp forest.
+The study was carried out in the 10-ha "Uppangala plot" of undisturbed, old-growth wet evergreen Diptrocarp forest.
 The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India) in India's Western Ghats biodiversity hotspot.

### 2016-08-08 inst/doc/draft.Rmd
v0.1.9030 Maxime version

@@
 ---
-title: 'Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India : vertical stratification reflects guild-depedent dynamics'
+title: 'Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India: vertical stratification reflects guild-dependent dynamics'
 author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
 date: '`r Sys.Date()`'
-output: pdf_document
+output: word_document
+csl: mee.csl
 bibliography: refs.bib
@@
 Both theory can provide realistic expectations of taxonomic diversity [@chave_comparing_2002], and some recent approaches have advocated investigating patterns of functional [@westoby_land-plant_2006; @mcgill_rebuilding_2006] and phylogenetic diversity [@mouquet_ecophylogenetics:_2012] to trace the signature of ecological processes related to specific biological attributes and adaptations.
-These approaches have been applied to resolve the conundrum og high tree diversity in tropical forests [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012], but studies on these ecosystems still remain scarce.
+These approaches have been applied to resolve the conundrum of high tree diversity in tropical forests [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012], but studies on these ecosystems still remain scarce.
 We here investigate taxonomic, functional and phylogenetic diversity in a low elevation rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], where no study has addressed the relative influence of neutral and niche-based dynamics on local forest composition.
@@
 The variation of functional traits among species reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007].
-Thes variation thus reflect fundamental trade-offs in ecological strategies along environmental gradients [@mark_westoby_plant_2002].
+These variation thus reflect fundamental trade-offs in ecological strategies along environmental gradients [@mark_westoby_plant_2002].
 For instance, specific leaf area varies along a soil fertility-shade index and represent a trade-off between resource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011].
@@
 The leaf economic spectrum opposes productive ecological strategy to conservative persistent strategy.
-The wood economic spectrum distinguish species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977: (i) competitive species, wiht high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength.
-@baraloto_decoupled_2010 highlighted decoupled wood and leaf strategies allowing independant variation of these strtegies, but @reich_world-wide_2014 conversely advocated that the two specrta have to integrated in a comprehensive whole plant economic spectrum [@freschet_evidence_2010] opposing fast growing species to more conservative ones.
+The wood economic spectrum distinguish species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, wiht high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength.
+@baraloto_decoupled_2010 highlighted decoupled wood and leaf strategies allowing independant variation of these strategies, but @reich_world-wide_2014 conversely advocated that the two spectra have to integrated in a comprehensive whole plant economic spectrum [@freschet_evidence_2010] opposing fast growing species to more conservative ones.
 Here we address how varying wood and leaf ecological strategies determine species community assembly in an Indian rainforest.
+
+A major challenge in addressing community assembly is to relate local coexistence to a broader spatial context of species dynamics [@ricklefs_disintegration_2008; @munoz_neutral_2016].
+Niche-based dynamics can entail community variation along broad environmental gradients, while local community dynamics can be neutral [@leibold_coexistence_2006].
+Local community dynamics can further be influenced by niche differences under the influence of competition [@macarthur_limiting_1967].
+In this perspective, @messier_trait_2016 have underlined that the significance of the leaf economic spectrum is basically scale-dependent, which can explain conflicting results across studies performed at different scales.
+In the same logic, a spatial hierarchy of ecological dynamics can entail a hierarchy of variation in functional traits from broad variation among species along large-scale environmental gradients, to intraspecific response to micro-environmental conditions at very fine scale [@albert_when_2011].
 Therefore a multiscale framework is needed to represent the hierarchy of ecological processes from regional to local scale [@lortie_rethinking_2004].
@@
 We will decompose diversity patterns in a large rainforest plot into sources of variation within and between communities [@kraft_disentangling_2011], and investigate the fine-scale and plot-scale dynamics using appropriate null models.
+For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they question the existence of the leaf economic spectrum at community scale.
 
@@
 Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988].
-It thus provide an ideal context to analyze guilod-dependent dynamics and ecological strategies in tropical rainforests.
+It thus provide an ideal context to analyze guild-dependent dynamics and ecological strategies in tropical rainforests.
 We will thus address diversity patterns and community assembly at guild-level.
@@
 
-The study was carried out in the 10-ha "Uppangala plot" of undisturbed, old-growth wet evergreen Diptrocarp forest.
-The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India) in India's Western Ghats biodiversity hotspot.
-Climate is characterized by a peak of heavy rainfall during monsoon and a marked dry season.
-Annual precipitation averages 5108 mm with 90% between June and October and an average dry season (monthly rainfall $\leq 45$ mm) of 3.8 months [@pelissier_twenty_2011].
+The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry of an undisturbed, old-growth wet evergreen Diptrocarp forest [@pascal_structure_1996].
+The plot is located near Uppangala village ("Uppangala plot") in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India) in India's Western Ghats biodiversity hotspot.
+Climate is characterized by both heavy rainfall and very dry seasons.
+Annual precipitation averages 5108 mm per year with 90% between June and October during the monsoon, while the dry season (monthly rainfall $\leq 45$ mm) lasts on average 3.8 months [@pelissier_twenty_2011].
 Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996].
 Climate and geology of the 10-ha plot can be considered as homogeneous.
-General topography of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
-The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996.
-Since 1989, all trees $\geq 30$ cm girth at breast height (gbh) were monitored in an area progressively extended to 10ha in 2014.
-All trees were identified, tagged, mapped, and provided with permanent dendrometer bands censused every 2-5 years.
+General topography of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as a major source of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998].
+We hypothesized that topographic variation should influence community variation depending on tree ecological strategies [@condit_spatial_2000].
 
@@
 
-We collected trait data on 89 over 111 (80.1%) species. The 32 most abundant species represent more than 80% of Uppangala tree individuals.
-We sampled five individuals per species for most abundant species and we completed with less abundant and rare species with one to five individuals (in total of 267 individuals).
-Thus we got trait data for 99.0% of the tree individuals present in the Uppangala plot.
-For each individual we collected 6 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation from @perez-harguindeguy_new_2013.
-Individuals and leaves were selected in consistent light condition, development stage and sanitary state.
+Since 1989, all trees $\geq 30$ cm girth at breast height (gbh) have been monitored in an area progressively extended to 10 ha in 2014.
+All trees were identified, tagged, mapped, and provided with permanent dendrometer bands censused every 2-5 years [@pelissier_twenty_2011].
+111 different species were recorded.
+We collected trait data on 89 over 111 (80.1%) species, which represented 99.0% of the tree individuals present in the Uppangala plot.
+With the 32 most abundant species, we reached  more than 80% of Uppangala tree individuals.
+We sampled five individuals per species for these species, and sampled one to five individuals of the less abundant and rare species (in total 267 individuals were sampled).
+For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation from @perez-harguindeguy_new_2013.
+Individuals and leaves were selected in consistent light conditions, development stage and sanitary state.
 Dawkins crown exposure index [@dawkins_management_1958] was assessed for each tree sampled.
 Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
-For some individuals, we were unable to collect an appropriate branch sample (58 over 267 individuals representing 11 missing species over 89 accounitg for 1% of individuals present in the 10ha plot).
+For some individuals, we were unable to collect an appropriate branch sample (11 missing species over 89 representing 1% of individuals present in the 10 ha plot).
 
@@
 Leaf petioles, rachis and petiolules were removed for all measurements.
-Leaf fresh weights, area and thickness were measured, using a bend scanner and a micrometer.
-Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$, the drying process was stopped after dry leaf weights were stable in time.
+Leaf fresh weights, area and thickness were measured, using a weighing scale, a bend scanner and a micrometer respectively.
+Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$.
 Once dried, the leaves were weighted directly after been extracted from the oven to avoid humidity bias.
@@
 Branch wood samples were sink in water to evaluate their volume by water displacement.
-Samples were then dried for 48 hours at 80 degrees celsius and weighted.
+Samples were then dried for 48 hours at 80 $^\circ C$ and weighted.
 
-We calculated five functional traits of leaf and wood samples: leaf thickness (Thick in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1).
+We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1).
+**Justify trait choice**.
 
@@
 table <- data.frame(
-  Trait = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
-  Abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
-  Range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
+  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
+  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
+  Range = c(paste(range(PFT$LT, na.rm = T), collapse = '-'),
             paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
@@
             paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
-  m = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
-                         round(sd(PFT$Thick, na.rm = T)), ')'),
+  m = c(paste0(round(mean(PFT$LT, na.rm = T)), ' (',
+                         round(sd(PFT$LT, na.rm = T)), ')'),
                   paste0(round(mean(PFT$LA, na.rm = T)), ' (',
@@
   s = c('Leaf defense vs. investment',
-               'Leaf ressource capture vs.investment',
+               'Leaf resource capture vs.investment',
                'Leaf defense vs. investment',
-               'Leaf ressource acquisiton vs. defense',
+               'Leaf resource acquisiton vs. defense',
                'Stem transport, structure and defense')
@@
 
-We selected 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
+We selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from the digital elevation model using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed using ArcGis [@desktop_arcgis_2011].
 Digital elevation model was provided by a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
-Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) were all derived from the digital elevation model using the RSAGA package [@brenning_statistical_2008].
-Distance to the nearest potential river was computed using ArcGis [@desktop_arcgis_2011].
+**Justify variable choice**.
 
-In order to identify main source of environmental variation and to avoid multi-colinearity in future analyses, we reduced our abiotic explanatory variables via a principal component analysis (PCA).
-We selected less correlated variables with the most ecological meaning.
-Thus we reduced our explanatory variables to 4: curvature to represent the water flow paths and soil accumulation areas, wetness to highlight water accumulation areas, southwesterness to represent aspect [@bunyan_effect_2015] and slope.
+In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the environmental variables.
+We selected the less correlated variables with the most ecological meaning.
+Thus we reduced our explanatory variables to 4 variables: curvature representing the water flow paths and soil accumulation areas, wetness highlighting water accumulation areas, southwesterness representing aspect [@bunyan_effect_2015] and slope.
 
@@
 The coefficient of variation of traits was investigated at three levels ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean): between leaves of the same individual (intra-individual), between trees of the same species (intra-specific for species with more than one individual), and between species (inter-specific).
-We performed a principal component analysis (PCA) of the mean value of traits per species.
+In addition, we performed a principal component analysis (PCA) of the mean value of traits per species, to characterize basic dimensions of ecological strategies.
 Subsequent community-level analysis were based on trait values averaged per species.
 
-### Local functional composition of communities
+### Community trait variability
 
-Across the whole 10 hectare plot from Uppangala, 240 local communities where defined in $20*20 m^2$ quadrats.
-To investigate functional composition of communities we used a null model approach.
-A first null model (NM1) randomizes trait values among species (permuting raw in `Trait*Species` matrix).
-The second model (NM2) randomizes abundances among species inside each community but keeps the observed list of species and trait values inside communities [@bernard-verdier_community_2012].
-
+Across the whole 10 hectare plot, $20*20 m^2$ quadrats were defined and represented 240 local communities.
 To investigate local functional composition of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010].
@@
 $$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
-CWM was also computed in presence-absence, to further emphasize role of rare species in community trait distribution.
-Linear regressions between CWMs and biotic and abiotic factors ($CWM_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$), after stepwise selection, were compared to corresponding 999 null linear regressions from NM1 with F-statistic.
-
+We also calculated the average of trait values based on species presence-absence, to further emphasize the contribution of rare species in community trait distribution [@cingolani_filtering_2007].
 CWV for each trait $t$ was calculated as the mean distance of trait values, $t_{i}$, to $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
@@
 $$CVNND = \frac{\sigma _{NND}}{\mu _{NND}} ~ , ~ ~ ~ NND = [NND_{i} = min(d_{i,j})_{i \neq j}] ~ (3)$$
-For both CWV and CVNND, proportion of of communities differing from null expectaions was tested with a two-tailed test with adjusted p-values by Benjamini-Hochberg correction.
-CWV lower than expected under a null model will be evidence for convergence of traits inside the community whereas higher CWV will show divergence in trait distribution.
-CVNND is used to detect niche differenciation patterns in case of a lower CVNND as observed under a null model [@jung_intraspecific_2010].
+
+To investigate functional composition of communities we used a null model approach.
+A first null model (NM1) randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix).
+The second model (NM2) randomized abundances among species within each community but kept the observed list of species and trait values inside communities [@bernard-verdier_community_2012].
+Linear regressions between CWMs and biotic and abiotic factors ($CWM_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$), after stepwise selection, were compared to corresponding 999 null linear regressions from NM1 with F-statistic.
+For both CWV and CVNND, we performed a two-tailed test of the proportion of of communities differing from null expectation with adjusted p-values by Benjamini-Hochberg correction.
+CWV lower than expected under a null model represented convergence of trait values within the community, whereas CWV higher than null evidenced divergence in trait distribution [@bernard-verdier_community_2012].
+CVNND was used to detect niche differentiation among coexisting species, which occured when the observed CVNND was lower than under a null model [@jung_intraspecific_2010].
 We then calculated the standardized effect size (SES) as: $$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
-where $I_{obs}$ is the observed metric and $I_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.
-Effect of environment on SES was investigated by linear regression with selected biotic and abiotic factors.
+where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.
+The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$).
 Analyses were conducted on the five functional traits and the species scores obtained on the 2 first PCA axes.
 
-### Floristic, functional and phylogenetic composition between communities
+### Taxonomic, functional and phylogenetic variation among communities
 
-Phylogenetic tree of our 89 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
-Phylogenetic signal was globaly assessed for each functional trait among the phylogenetic tree by Moran's I statistic.
-Positive Moran's I value indicates convergence of trait values within the phylogeny whereas negative Moran's I will show divergence.
+A phylogenetic tree of our 89 species was built with `S.PhyloMaker` function proposed by @qian_updated_2016-1.
+We assessed phylogenetic signal of each functional using the Moran's I statistic.
+Positive Moran's I value indicated convergence of trait values within the phylogeny whereas negative Moran's I will showed divergence.
 
-To investigate floristic, functional, and phylogenetic composition between communities we used $I_{st}$, for floristic composition, $U_{st}$ and $\tau _{st}$, for functional composition in abundance and presence-absence respectively, and $\beta _{st}$ and $\pi _{st}$, for phylogenetic composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
+To investigate taxonomic, functional, and phylogenetic variation among communities we calculated $I_{st}$, for taxonomic composition, $U_{st}$ and $\tau _{st}$, for functional composition in abundance and presence-absence respectively, and $\beta _{st}$ and $\pi _{st}$, for phylogenetic composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
+**Develop indices**.
 
-Variation of composition between community pairs with those metrics was translated as a floristic, functional and phylogenetic distance, and compared to environmental and spatial distances among community pairs.
+Variation of composition between community pairs with those metrics was translated as a taxonomic, functional and phylogenetic distance, and compared to environmental and spatial distances among community pairs.
 Environmental distances were calculated as geometric distances between communities on the plan composed by the two first axes of environmental variables PCA (see supplementary material S1).
-Correlations with environment and space were assessed with two Mantel tests successively with floristic, functional, and phylogenetic community compositions.
-A firt full Mantel statistic evaluates the correlation between each metric distance matrix and the environmental distance matrix (E).
-The second partial Mantel statistic assess the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S), representing spatial residuals that can influence community composition once the environment effect taken into account.
+Correlations with environment and space were assessed based on two Mantel tests successively with taxonomic, functional, and phylogenetic community compositions.
+A firt Mantel statistic assessed the correlation between each metric distance matrix and the environmental distance matrix (E).
+The second partial Mantel statistic assessed the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S), to represent remaining spatial dependency in community composition once the influence of environmental variation is taken into account.
 
@@
 Finally $\beta st$ and $\pi st$ were tested against the nul model *"1s"* suggested by @hardy_testing_2008 which randomizes phylogenetic tree tips.
+**One global null model section ?**
 
@@
 Whereas LA shows already high variabililty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
+**Conclude on low $cv_{intra}$ justifying species mean trait**.
+**Give Species example of extreme range and compare to litterature**.
 
@@
 *Mean of coefficient of variation for leaves in all individuals (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
-See table 1 for abbreviation.*
+See table 1 for abbreviation.* **Move range, mean and sd columns here*
 
@@
 PFT_ind <- PFT_ind[-c(1:2)]
-cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
+cv <- data.frame(row.names = c('LT', 'LA', 'LDMC', 'SLA', 'WD'))
 cv$Individual <- unlist(lapply(
@@
 
-The principal component analysis (PCA) shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
-The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood.
-PCA1 axis associates WD to leaf structural traits variation linked to hydraulic conductivity efficiency associated with the wood economic spectrum [WES, @chave_towards_2009].
-Whereas the PCA2 covary mainly positively with SLA and negatively with LDMC representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
+The principal component analysis (PCA) shows WD associated with structural leaf traits orthogonal to the SLA (figure 1).
+The first PCA axis (PCA1) covary mainly with LT, LDMC and WD; opposing thick big leaves and lightwood to dense leaves and hardwood.
+PCA1 axis associated WD to leaf structural traits linked to hydraulic conductive efficiency along the wood economic spectrum [WES, @chave_towards_2009].
+The PCA2 covaried mainly positively with SLA and negatively with LDMC along a leaf economic spectrum [LES, @wright_worldwide_2004].
+**Develop productive against persistent strategy**.
+**High WD LDMC relation indicating structural investment**.
 
@@
 PFT_sp <- PFT_sp[-c(1:2,4)]
-pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_sp, cor = T)
+pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT_sp, cor = T)
 
@@
 
-Figure 1: Principal component analysis (PCA) performed on species trait means.
-*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 5 components (red line).
+Figure 1: Principal Component Analysis (PCA) performed on species mean traot values.
+*The bottom right subplot represents the eigen values (grey bars) compare to a broken stick distribution for 5 components (red line).
 See table 1 for abbreviation.*
@@
 L[is.na(L)] <- 0
-Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')]
+Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')]
 row.names(Q) <- PFT_sp$SpCode
@@
 
-## Local functional composition of communities
+## Functional composition of local communities
 
-Community weighted means (CWM) only show few significant trends along environmental gradients (see table 3).
-CWMs reveal that local mean WD will significantly increase under south-west; whereas SLA and LA, in presence-absence only, tend to decrease with curvature.
-In addition, LDMC in presence-absence is decreasing with canopy height.
+Community weighted means (CWM), for both abundant and rare species, showed some significant variations along environmental gradients (see table 3).
+Community mean values of WD significantly increased under south-west direction, both in abundance and presence-absence CWM,  whereas SLA and LA, in presence-absence only, decreased with curvature.
+In addition, LDMC in presence-absence was decreasing with canopy height.
+**Develop meaning of AB et PA, specifically in term of species**.
 
-Table 3: Summary of linear regressions and null models for CWMs of traits.
-_Linear regression results for each trait CWMs weighted by species relative abundance (AB) and presence-absence (PA) after a variable stepwise selection with slope, curvature, wetness, southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Linear regressions global and partial F-statistics were compared to 999 null linear regressions built with null model NM1 which randomizes traits among species.
+Table 3: Summary of linear models of community-level trait values.
+_The results are shown for either CWM values weighted either by species relative abundance (AB) or in presence-absence (PA) after stepwise selection of the predictors including slope, curvature, wetness, southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Global and partial F-statistics of final models were compared to statistics of 999 null linear regressions built with null model NM1, which randomizes traits among species.
 See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+The number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 # Observed CWMs
-CWM <- lapply(L, function(l){lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){apply(l, 1, function(x){weighted.mean(Q[,v], x, na.rm=T)})})})
-CWM <- lapply(CWM, function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})
+CWM <- lapply(L, function(l){lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){apply(l, 1, function(x){weighted.mean(Q[,v], x, na.rm=T)})})})
+CWM <- lapply(CWM, function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT') ; return(x)})
 ## Regressions
@@
   lapply(nQ, function(q){
-    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick')), function(v){
+    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
       apply(l, 1, function(x){
@@
 rm(cl)
-nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'Thick') ; return(x)})})
+nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT') ; return(x)})})
 ## Regressions
@@
 
-Traits divergence is measured with community weighted variances (CWV).
-Overall, observed CWVs doesn't differ from null expectations for all traits at the exception of one community for WD (over 240 communities).
-Still we can observe some trends between standardized effect size (SES) of CWVs and environment (see table 4).
-Effectively south-west orientation will make local functional composition of communities converge towards a mean LES and LDMC whereas LA will diverge inside community under south-west orientation.
-LA and LDMC values will also diverge with increasing curvature.
-Observed CVNNDs doesn't reveal niche differenciation among communities for all traits.
-Still we can observe some trends between standardized effect size (SES) of CVNNDs and environment (see table 4).
-Species tend to differentiate their LA with lower curvature and wetness and their WD opppositely to SouthWest orientation.
+Traits convergence/divergence was measured with community weighted variances (CWV).
+Overall, observed CWVs DID NOT differ from null expectations for all traits.
+We could found significant relationships between standardized effect size (SES) of CWVs and environmental variables (see table 4).
+Convergence of LES and LDMC and divergence of LA were found under south-west orientation.
+LA and LDMC values also diverged in communities with increasing curvature.
+Observed CVNNDs did not reveal niche differentiation among communities for all traits.
+But we also found significant relationships of their standardized effect size (SES) of CVNNDs with environmental variables (see table 4).
+niche differentiation of LA increased in lower curvature and wetness contexts, while differentiation of WD decreased in southWest orientation.
 
-Table 4: Summary of linear regressions for traits standardized effect size (SES) of CWVs and CVNNDs.
-_Linear regression results for each trait SES of CWVs and CVNNDs after a variable stepwise selection with slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Table 4: Summary of linear models of traits standardized effect size (SES) of CWVs and CVNNDs.
+_Linear regression results for each trait SES of CWVs and CVNNDs after stepwise selection of the predictors oncluding slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
 See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+The number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 
-## Floristic, functional and phylogenetic composition between communities
+## Taxonomic, functional and phylogenetic varitions among communities
 
-Floristic composition between communities is correlated significantly to environement spatial organization with Mantel test ($r_{I_{st},E} = 0.23$ with $p-value < 0.001$) and without any spatial residuals represented with partial Mantel test knowing environment ($r_{I_{st},S|E} = 0.01$ with $p-value > 0.1$).
+Taxonomic variation among communities was correlated significantly to environment variation (Mantel test, $r_{I_{st},E} = 0.23$ with $p-value < 0.001$), but not with remaining spatial dependency in community composition (Mantel test of space knowing environment, $r_{I_{st},S|E} = 0.01$ with $p-value > 0.1$).
 
-Overall, functional composition between communities is not correlated to environment nor to space knowing the environment effect (see table 5).
-Still WD between communities, in abundance as well as in presence absence, is positively significantly correlated to environment and is even negatively correlated in presence absence to spatial structure knowing environment spatial effect.
+Overall, the global functional variation among communities was neither correlated to environment nor to spactial distance knowing the environment effect (see table 5).
+However, the variation of WD, in abundance as well as in presence absence, was positively correlated to environmental variation and is negatively correlated in presence absence to the spatial structure knowing environment spatial effect.
 
-Whereas LDMC is not phylogenetically convergent nor divergent (Moran's I of 0 with a $p-value > 0.1$), LA and Thick are significantly convergent among the phylogenetic tree (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA are weakly convergent (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
-Phylogenetic composition between communities is not correlated to environement spatial organization with Mantel test ($r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor with spatial residuals represented with partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).
-
-<!-- **Table to replace by text**: Summary of Mantel test between floristic composition, environment and spatial distances between communities.
-_$Ist$  was used to investigate variation of floristic composition between communities.
-Observed values were compared to a null model randomizing individuals among subplots while preserving the regional abundance of species, and the number of individuals per community.
-E is the mantel test of the metric with the environmental distance matrix issued from PCA two first axes.
-S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).._ -->
+While LDMC did not phylogenetic structure (Moran's I of 0 with a $p-value > 0.1$), LA and LT were phylogenetically clustered (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA were weakly clustered (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
+Phylogenetic variation among communities was neither correlated to environmental variation (Mantel test, $r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor to remaining spatial variation in partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).
 
@@
 
-Table 5: Summary of Mantel test between functional composition, environment and spatial distances between communities.
-_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities.
-Observed values were compared to the null model NM1 randomizing traits among species (permuting raw in `Trait*Species` matrix).
-E is the mantel test of the metric with the environmental distance matrix issued from PCA two first axes.
-S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+Table 5: Summary of Mantel tests between functional composition, environment and spatial distances between communities.
+_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to measure functional composition variation among communities.
+Observed values were compared to null values obtained with model NM1 randomizing traits among species (permuting raw in `Trait*Species` matrix).
+E is the Mantel test of this metric according to the environmental distance derived from communtiy scores in the environmental PCA (two first axes).
+S|E is the partial Mantel test of this metric according to spatial distance conditionally to environment variation.
+The number of asterisk indicates significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 
-<!-- **Table to replace by text**: Summary of Moran’s I statistic and p-values for phylogenetic signal of functional traits.
-_See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._ -->
-
 ```{r Phylosignal}
@@
 ```
-
-<!-- **Table to replace by text**: Summary of Mantel test between phylogenetic composition, environment and spatial distances between communities.
-_$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities.
-Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 randomizong phylogenetic tree tips.
-E is the mantel test of the metric with the environmental distance matrix issued from PCA two first axes.
-S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._ -->
 
@@
 
-Plants distinguish theimselves in functional trait composition betwen individuals and even more between species.
-Variation of species functional composition results from functional trade-offs that compose major variation axes of functional traits called economic spectra.
-We found back the leaf economic spectrum of @wright_worldwide_2004 opposing in our case SLA to LDMC in a trade-off between productivity, with less carbon investment for higher photosynthesis rates, opposed to persistance, with more leaf defense and investment.
-Independantly to that spectrum, we detected a second major axis of trait variation opposing WD to LA and Thick.
-This axis opposes more resistant to stress and disturbance species, thanks to denser wood, to competitive species, with thicker and larger leaves for a more efficient water conductivity.
-This spectrum correspond to the wood economic spectrum described by @chave_towards_2009, and more precisely this opposition between small thin leave with dense wood to large thick leaves with light wood correspond to the additional gradient uncovered by  @baraloto_decoupled_2010.
+We found increasing trait variation within individuals, between individuals within species, and between speces (table 2).
+It entailed that the primary source of trait variation within and among communities should be among species.
+The PCA of functional trait variation among species displayed two main dimensions (figure 1), which basically represented a wood and a leaf economic spectrum, respectively.
+We found opposing SLA to LDMC values along the leaf economic spectrum, which expressed a trade-off between productivity, with less carbon investment for higher photosynthesis rates, and persistence, with more leaf defense and investment [PCA2, @wright_worldwide_2004].
+A second major axis of trait variation opposed WD to LA and LT.
+It thus opposed stress tolerant species, with denser wood, to competitive species, with thicker and larger leaves for more efficient water conductivity.
+It represented the wood economic spectrum as defined by @chave_towards_2009, and the opposition between small thin leave and hardwood and large thick leaves with light wood corresponded to the additional gradient uncovered by @baraloto_decoupled_2010.
 
-Results are similar to the one found by @fortunel_leaf_2012 where PCA first and second axes were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
-Our results integrate less traits (5 against 14), and less species (89 against 758), at a smaller scale, but still capture the same strategy axes.
+Furthermore, our results are similar to the between-species variations found @fortunel_leaf_2012, where basic PCA axes also represented leaf and wood economic spectra (inverted compare to us).
+Noticeably though, our analysis integrated less traits (5 against 14), less species (89 against 758), and considered a regional set of species found in a local forest-plot scale while @fortunel_leaf_2012 considered a regional set of species.
+Therefore, we could capture basic economic spectra found at regional scale in Neotropics at a far more local scale in an Indian rainforest, even though such downscaling does not hold in any case [@messier_trait_2016].
+It suggests that these spectra indeed represent major drivers of ecological strategies in local forest dynamics in our context.
 
-## Local functional composition
+## Communtiy-level trait values
 
-The leaf economic spectrum is majorly affected by aspect with a convergence of value toward optimal functional traits under south-west orientation, especially for LDMC of rare species.
-Aspect, and more specifically south-west, effect in tropical forest is scarcely documented but highlighted in discussion by @gunatilleke_specieshabitat_2006.
-@bunyan_effect_2015 observed a similar effect on shola distribution of Western Ghats and suggested it could result from monsoon winds.
-Effectively monsoon winds arise fom south-west direction in this region, and bring more stressfull and disturbed conditions to species [@soman_aspects_1990].
-Curvature is also affecting the leaf economy.
-Convex area will accumulate soil and attract water flux increasing soil fertiliy conditions for species that will invest more in leaf ressource acquisition toward a productive strategy with higher SLA and a convergent optimal LDMC.
-SLA and LDMC with curvature effect seems to reflect well soil fertility [@hodgson_is_2011].
+We analyzed functional composition of $20*20m^2$ communities (CWM), as well as the variation of this composition with environment across the 10 ha forest plot.
+We found that the traits of the leaf economic spectrum (LES) were mainly affected by aspect with more convergent community composition (lower CWV) under south-west orientation, especially for LDMC of less common species.
+The influence of aspect is little documented in tropical forests, but still highlighted by @gunatilleke_specieshabitat_2006 in monsoon rainforests.
+@bunyan_effect_2015 also observed similar influence in mountain rainforests of Western Ghats, and suggested that it could result from monsoon winds.
+In our context, monsoon winds arise from south-west direction, and bring more stressful and disturbed conditions to species [@soman_aspects_1990] **To develop, see maxime disturbance hypothesis on opposite side**.
+Curvature is also affecting the variation of leaf economics in community.
+A locally convex context would accumulate more soil and attract more water, increasing soil fertility and allowing persistence of species investing more in leaf resource acquisition, with higher SLA and a convergence in LDMC values.
+The variation of SLA and LDMC with curvature, is consistent with the influence of changing soil fertility [@hodgson_is_2011].
 
-Biotic factors also affect the LES through the negative effect of canopy height on LDMC.
-The canopy height effect is null on understorey species and only apply on species that take part in the canopy composition or emerge (results unshown).
-We would assumed that understorey species would be investing in higher LDMC for an higher leaf lifespan under shade stress [@kitajima_tissue-level_2010].
-But both the high variation of canopy height resulting in a relatively light open canopy and the dry season of 3.8 months suggest that leaf lifespan investment is not so obvious.
-Instead of a growth-survival trade-off shade tolerant species may adopt a more low-high light growth trade-off strategy [@poorter_leaf_2009] strongly affecting SLA but not driving LDMC.
+Biotic factors also affected the LES through a negative effect of canopy height on LDMC.
+There was no effect of canopy height on understorey species, while it influenced leaf traits of canopy and emergent trees (results unshown **To include in the analysis**).
+A basic expectation is that understorey species would invest in higher LDMC for higher leaf lifespan and durability under limited light and water resources [@kitajima_tissue-level_2010].
+But the Uppangala forest is characterized by heterogeneity of canopy height allowing light to penetrate more in lower strata and a dry season of 3.8 months.
+Comsequently leaf lifespan investment may not be so important resulting in a low-high light growth trade-off strategy instead of a growth-survival trade-off [@poorter_leaf_2009], affecting SLA but not LDMC.
 Emergent and high canopy species will have a lower LDMC, resulting from a more productive strategy toward leaf resource acquisition at the expense of leaf carbon investment in defense and resistance.
-Thus canopy species with the lowest LDMC will grow the fastest and then reach the highest canopy.
-Measurement of LDMC on species classified as canopy and emergent are realized most of the time on shadow leaves from understorey individuals.
-Still we observed previous relation indicating that species phylogenetic determinism is stronger than micro-environmental variation plasticity for LDMC, as suggested by the intra-specific variation of LDMC 1.5 times bigger than intra-individuals variation.
+**Second hyptohesis: canopy and emergent tree are more affected by water stress than understorey species preserved by microclimatic conditions buffered by canopy**.
+Thus canopy species with the lowest LDMC will grow more quickly and then reach the highest canopy.
+Measurement of LDMC on species classified as canopy and emergent were realized most of the time on shadow leaves from understorey individuals.
+Still we observed previous relationship indicating that species phylogenetic determinism is stronger than micro-environmental variation plasticity for LDMC, as suggested by the intra-specific variation of LDMC 1.5 times larger than intra-individuals variation.
 
-The wood economic spectrum defined by WD, LA, and Thick is only partially affected by environment at our 10 ha plot scale.
-Neutral hypothesis of funcional equivalence [@hubbell_neutral_2005] are most of the times sufficient to build back similar links between topography and functional compoisiton of communities.
-Still aspect wiht south-west orientation and curvature are the two main ecological factors filtering functional traits from the wood economic spectrum.
-South-west orientation, with probably more disturbance and stress from monsoon, result in denser wood with reduced eveness by convergence toward different optimal density.
-In parrallel, south-west direction broaden leaf size strategies, pehraps to answer to hydraulic diversity induced by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008].
-Fertility decrease, with concavity, results in smaller leaves inside communities with a distribution broaden by multimodality reducing eveness.
-Wetness is increasing with higher curvature and lower LA eveness, indicating that multimodality is happening more in wettest low elevation area.
+The wood economic spectrum defined by WD, LA, and LT was mainly affected at community level by south-west orientation and curvature.
+South-west orientation, with probably more disturbance and stress due to exposure to monsoon winds **Maxime: counter effect by topo ?**, resulted in denser wood.
+However lower evenness than random could be related to multimodality.
+In addition, south-west orientation also related to a broader range of leaf sizes, which could be consistent with more diverse hydraulic strategies reflected by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008] **To develop, especially for a multimodality under southwest**.
+Decreasing fertility , with concavity was related to smaller leaves inside communities with a distribution broaden by multimodality reducing evenness **to develop, eventually illustrated with one example**.
+Wetness increased with higher curvature and lower LA evenness, indicating that multimodality is happening more in wettest low elevation area.
 Wetness could compensate water departure trigerred by positive curvature in lower elevation areas.
+**Repetition with first paragraph, to merge ?**.
 
-## Composition between communities
+## Variations among communities
 
-Limited dispersal, assumed by neutral theory [@condit_beta-diversity_2002], does not affect florisitic composition of communities at the 10 ha plot scale revealed by the absence of links  with spatial residuals.
-Florisitic composition of communities is only shaped by environment.
-On opposite, functional composition is overall not structured with environment nor with spatial residuals.
+Taxonomic variation in community composition was only dependent on environmental distance.
+Because we did not find any residual pattern of increasing taxonomic dissimilarity with spatial distance, it could indicate that environmental heterogeneity was a primary driver of species distributions, and that dispersal limitation did not influence these distributions at the scale of our forest plot.
+Conversely, functional variation did not depend on environmental distance or on spatial distance **Paradoxal ?
+Is it due to environmental aggregation in one variable or to a power loss of the test ?**.
 At the 10 ha level functional variation between communities is stochastic; whereas at the community level environment is partially shapping florisitic composition.
@@
 Still environment is shaping community resistance to disturbance and stress by affecting wood density $\beta$ structure at fine scale.
-WD of rare species is even diverging with spatial residuals supposing an unaccounted spatially organized variable opposing close community in WD composition.
-Topography removed from spatial residuals, and climate and pedology being constant in the 10 ha plot, this unaccouted variation is assumed to be due to stochastic event in plot history as death.
-Unshowed analysis reveals that this spatial divergence of WD is not explained by tree death events from the two last decades but certainly to older events.
-Finally, variation in distance between taxon wasn't translated in phylogenetic variation whereas most of measured functional traits appeared phylogenetically conserved and clustered.
+WD of less abundant species is even diverging with spatial residuals supposing an unaccounted spatially organized variable opposing close community in WD composition.
+Topography removed from spatial residuals, and climate and pedology being constant in the 10 ha plot, this unaccounted variation is assumed to be due to stochastic event in plot history as death.
+Unshown analysis reveals that this spatial divergence of WD is not explained by tree death events from the two last decades but certainly to older events.
+**To redraft**.
+Finally, variation in distance between taxon was not translated in phylogenetic variation whereas most of measured functional traits appeared phylogenetically conserved and clustered.
+**To discuss**.
 
@@
 Still the study reveal that few commonly used traits can highlight well known strategy axes.
+**Develop what would have brought this other traits**.
 
@@
 On the other hand, WD as shown to be filtered by environment at fine scale both in $\alpha$ and $\beta$ levels.
-Species wood seems highly adpted to environment in Western Gahts forests in link to monsoon events.
+Species wood seems highly adpted to environment in Western Gahts forests in link to monsoon events **To brought sooner, break conclusion**.
 
@@
 
-## S1: Principal component analysis performed on topgraphical variables
+## S1: Principal component analysis of topographical variables
 
@@
 
-Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
-Elevation and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
+Curvature covaries with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
+Elevation and wetness are opposed in the principal component analysis and slope and south-westerness does not seem to covary much with other variables **Not precise**.
 

### 2016-08-08 R/draft.R

@@ -189,12 +189,12 @@
 # biplot(bca, xax = 5, yax = 3)
 # randtest(bca)
 
-PFT_sp0 <- PFT_sp
-PFT_sp <- PFT_sp0
-sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] %in% c('ESII'))
-PFT_sp <- PFT_sp[sel,]
-par(mfrow = c(1,2))
-hist(PFT_sp$LDMC[sel], xlim = range(PFT_sp$LDMC), col = 'firebrick', xlab = 'LDMC', main = "ESI species LDMC")
-hist(PFT_sp$LDMC[sel], xlim = range(PFT_sp$LDMC), col = 'blue', xlab = 'LDMC', main = "ESII species LDMC")
+# PFT_sp0 <- PFT_sp
+# PFT_sp <- PFT_sp0
+# sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] %in% c('ESII'))
+# PFT_sp <- PFT_sp[sel,]
+# par(mfrow = c(1,2))
+# hist(PFT_sp$LDMC[sel], xlim = range(PFT_sp$LDMC), col = 'firebrick', xlab = 'LDMC', main = "ESI species LDMC")
+# hist(PFT_sp$LDMC[sel], xlim = range(PFT_sp$LDMC), col = 'blue', xlab = 'LDMC', main = "ESII species LDMC")
 # CWMII <- CWM
 # summary(CWM.lm$abundance$LDMC)

### 2016-08-08 R/genPFT.R

@@ -27,6 +27,7 @@ genPFT <- function(){
   PFT$SLA = PFT$SLA / 1000 # from mm2.g-1 to m2.kg-1
   PFT$HA = PFT$HA / PFT$LA
   names(PFT)[which(names(PFT) == 'HA')] = 'SHA' # Specific holes area
+  names(PFT)[which(names(PFT) == 'Thick')] = 'LT'
   # Errors modification
   PFT$WD[which(PFT$WD > 1)] <- NA # Wood sample too small for correct WD measurement
   species <- genSpecies()

### 2016-08-09 inst/doc/draft.Rmd
v0.1.9031 Strata analysis

@@
 date: '`r Sys.Date()`'
-output: word_document
+output: pdf_document
 csl: mee.csl
@@
 row.names(Q) <- PFT_sp$SpCode
+sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] %in% c('ESI', 'ESII'))
+Q <- list(C = Q[sel,], U = Q[-sel,])
+sel <- lapply(Q, function(q){intersect(row.names(q), names(L))})
+Q <- mapply(function(q,s){q[s,]}, q = Q, s = sel, SIMPLIFY = F)
+L <- lapply(sel, function(s){L[s]})
+R <- lapply(L, function(l){R[row.names(l),]})
+L <- lapply(L, function(l){list(abundance = l, `presence-absence` = l)})
+L <- lapply(L, function(l){l$`presence-absence`[l$`presence-absence` > 1]  <- 1 ; return(l)})
+rm(Deriv, XY, Env, sel, BA)
+```
+
+## Functional composition of local communities
+
+Community weighted means (CWM), for both abundant and rare species, showed some significant variations along environmental gradients (see table 3).
+Community mean values of WD significantly increased under south-west direction, both in abundance and presence-absence CWM,  whereas SLA and LA, in presence-absence only, decreased with curvature.
+In addition, LDMC in presence-absence was decreasing with canopy height.
+**Develop meaning of AB et PA, specifically in term of species**.
+
+Table 3: Summary of linear models of community-level trait values.
+_The results are shown for either CWM values weighted either by species relative abundance (AB) or in presence-absence (PA) after stepwise selection of the predictors including slope, curvature, wetness, southwesterness (SW), canopy height (Canopy) and basal area (BA).
+Global and partial F-statistics of final models were compared to statistics of 999 null linear regressions built with null model NM1, which randomizes traits among species.
+See table 1 for abbreviations.
+The number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r CWM}
+# Observed CWMs
+CWM <- mapply(function(Ls,Qs){
+  lapply(Ls, function(l){
+    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
+      apply(l, 1, function(x){
+        weighted.mean(Qs[,v], x, na.rm=T)
+      })
+    })
+  })
+}, Ls = L, Qs = Q, SIMPLIFY = F)
+CWM <- lapply(CWM, function(s){
+  lapply(s, function(x){
+    names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')
+    return(x)
+  })
+})
+## Regressions
+CWM.lm <- mapply(function(CWMs, Rs){
+  lapply(CWMs, function(p){
+    lapply(p, function(t){
+      lm(t ~., Rs)
+    })
+  })
+}, CWMs = CWM, Rs = R, SIMPLIFY = F)
+CWM.lm <- mapply(function(s, Rs){
+  lapply(s, function(p){
+    lapply(p, function(t){
+      step(t, direction = 'both', trace = 0)
+    })
+  })
+}, s = CWM.lm, Rs = R, SIMPLIFY = F)
+## Global F
+Fval <- lapply(CWM.lm, function(s){
+  lapply(s, function(y){
+    lapply(y, function(x){
+      summary(x)$fstatistic['value']
+    })
+  })
+})
+Fval <- lapply(Fval, function(s){
+  lapply(s, function(x){
+    do.call('rbind.data.frame', x)
+  })
+})
+Fval <- lapply(Fval, function(s){
+  s <- data.frame(s)
+  names(s) <- names(CWM.lm$C)
+  return(s)
+})
+Fval$U <- Fval$U[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),]
+## Partial F
+pFval <- lapply(CWM.lm, function(s){
+  lapply(s, function(y){
+    lapply(y, function(x){
+      car::Anova(x)['F value']
+    })
+  })
+})
+
+# Simulated CWMs
+n = 999
+nQ <- lapply(Q, function(s){
+  rep(list(s), n)
+})
+nQ <- lapply(nQ, function(s){ 
+  lapply(s, function(nq){
+      row.names(nq) <- sample(row.names(nq))
+      return(nq)
+  })
+})
+nQ <- mapply(function(nQs, Ls){
+  lapply(nQs, function(nq){
+    nq[names(Ls$abundance),]
+  })
+}, nQs = nQ, Ls = L, SIMPLIFY = F)
+cl <- makeCluster(cores)
+clusterExport(cl, list('L', 'nQ'))
+nCWM <- clusterMap(cl, function(nQs, Ls){
+  lapply(nQs, function(nq){
+    lapply(Ls, function(l){
+      lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
+        apply(l, 1, function(x){
+          weighted.mean(nq[,v], x, na.rm=T)
+        })
+      })
+    })
+  })
+}, Ls = L, nQs = nQ)
+stopCluster(cl)
+rm(cl)
+nCWM <- lapply(nCWM, function(s){
+  lapply(s, function(ns){
+  lapply(ns,function(ab){
+    names(ab) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')
+    return(ab)
+    })
+  })
+})
+## Regressions
+nCWM.lm <- mapply(function(CWM.lms, nCWMs, Rs){
+  lapply(nCWMs, function(ncwm){
+    mapply(function(CWM.lmab, ncwmab){
+      mapply(function(CWM.lmt, ncwmt){
+        lm(formula(paste0('ncwmt ~ ', as.character(CWM.lmt$terms)[3])), cbind(ncwmt,Rs))
+      }, CWM.lmt = CWM.lmab, ncwmt = ncwmab, SIMPLIFY = F)
+    }, CWM.lmab = CWM.lms, ncwmab = ncwm, SIMPLIFY = F)
+  })
+}, CWM.lms = CWM.lm, nCWMs = nCWM, Rs = R, SIMPLIFY = F)
+rm(nCWM)
+invisible(gc())
+## Global F
+nFval <- lapply(nCWM.lm, function(nCWM.lms){
+  lapply(nCWM.lms, function(lm){
+    lapply(lm, function(y){
+      lapply(y, function(x){
+        summary(x)$fstatistic['value']
+      })
+    })
+  })
+})
+nFval <- lapply(nFval, function(x){
+  lapply(x, `[`, names(x[[1]]))
+  apply(do.call(rbind, x), 2, as.list)
+})
+nFval <- lapply(nFval, function(s){
+  lapply(s, function(x){
+    lapply(x, `[`, names(x[[1]]))
+    apply(do.call(rbind, x), 2, as.list)
+  })
+})
+nFval <- lapply(nFval, function(s){
+  lapply(s, function(ab){
+    lapply(ab, unlist)
+  })
+})
+pval <- mapply(function(Fvals, nFvals){
+  1 -  mapply(function(Fvalab, nFvalab){
+    mapply(function(Fvalt,nFvalt){
+      if(!is.na(Fvalt))
+        rank(c(Fvalt, nFvalt))[1]/(n+1)
+      else
+        NA
+    }, Fvalt = Fvalab, nFvalt = nFvalab)
+  }, Fvalab = lapply(as.list(Fvals), as.list), nFvalab = nFvals)
+}, Fvals = Fval, nFvals = nFval, SIMPLIFY = F)
+pval <- lapply(pval, function(s){
+  row.names(s) <- row.names(Fval$C)
+  return(s)
+})
+rm(Fval, nFval)
+# Partial F
+npFval <- lapply(nCWM.lm, function(s){
+  lapply(s, function(lm){
+    lapply(lm, function(y){
+      lapply(y, function(x){
+        car::Anova(x)['F value']
+      })
+    })
+  })
+})
+npFval <- lapply(npFval, function(x){
+  lapply(x, `[`, names(x[[1]]))
+  apply(do.call(rbind, x), 2, as.list)
+})
+npFval <- lapply(npFval, function(s){
+  lapply(s, function(x){
+    lapply(x, `[`, names(x[[1]]))
+    apply(do.call(rbind, x), 2, as.list)
+  })
+})
+npFval <- lapply(npFval, function(s){
+  lapply(s, function(ab){
+    lapply(ab, function(t){
+      do.call('cbind', t)
+    })
+  })
+})
+ppval <- mapply(function(pFvals, npFvals){
+  mapply(function(x,y){
+    mapply(function(r,s){
+      mapply(function(p,q){
+        rank(c(p,q))[1]/(n+1)
+      }, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))
+    }, r = x, s = y)
+  }, x = pFvals, y = npFvals, SIMPLIFY = F)
+}, pFvals = pFval, npFvals = npFval, SIMPLIFY = F)
+ppval <- lapply(ppval, function(s){
+  lapply(s, function(y){
+    lapply(y, function(x){
+      data.frame(t(x))
+    })
+  })
+})
+ppval <- lapply(ppval, function(s){
+  lapply(s, function(y){
+    data.frame(data.table::rbindlist(y, fill = T))
+  })
+})
+ppval <- lapply(ppval, function(s){
+  lapply(s, function(y){
+    row.names(y) <- names(CWM.lm$C$abundance)
+    return(y)
+  })
+})
+var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
+ppval <- lapply(ppval, function(s){
+  lapply(s, function(ab){
+    ab[,setdiff(var, names(ab))] <- NA
+    ab <- ab[var]
+    return(ab)
+  })
+})
+ppval <- lapply(ppval, function(s){
+  lapply(s, function(x){
+    1 - x
+  })
+})
+rm(pFval, npFval)
+
+# Result tables
+reg <- paste('~', paste(names(R$C), collapse = ' + '))
+res <- mapply(function(CWM.lms, pvals, ppvals){
+  mapply(function(CWM.lmab, pvalab, ppvalab){
+    regression_table(CWM.lmab, reg, pvalab, ppvalab)
+  }, CWM.lmab = CWM.lms, pvalab = as.list(data.frame(pvals)), ppvalab = ppvals, SIMPLIFY = F)
+}, CWM.lms = CWM.lm, pvals = pval, ppvals = ppval, SIMPLIFY = F)
+res <- lapply(res, function(s){
+  mapply(function(ab, name){
+    ab <- cbind(c(name, rep(' ', length(ab[,1])-1)), ab)
+    names(ab)[1] <- 'Type'
+    return(ab)
+  }, ab = s, name = as.list(c('AB', 'PA')), SIMPLIFY = F)
+})
+res <- lapply(res, function(s){
+  do.call('rbind', s)
+})
+res <- mapply(function(s, name){
+  s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
+  names(s)[1] <- 'Strata'
+  return(s)
+}, s = res, name = as.list(c('C', 'U')), SIMPLIFY = F)
+res <- do.call('rbind', res)
+kable(bold(res), row.names = F)
+rm(CWM.lm, nCWM.lm, pval, ppval, reg, res) ; invisible(gc())
+```
+
+Traits convergence/divergence was measured with community weighted variances (CWV).
+Overall, observed CWVs did not differ from null expectations for all traits.
+We could found significant relationships between standardized effect size (SES) of CWVs and environmental variables (see table 4).
+Convergence of LES and LDMC and divergence of LA were found under south-west orientation.
+LA and LDMC values also diverged in communities with increasing curvature.
+Observed CVNNDs did not reveal niche differentiation among communities for all traits.
+But we also found significant relationships of their standardized effect size (SES) of CVNNDs with environmental variables (see table 4).
+niche differentiation of LA increased in lower curvature and wetness contexts, while differentiation of WD decreased in southwest orientation.
+
+Table 4: Summary of linear models of traits standardized effect size (SES) of CWVs and CVNNDs.
+_Linear regression results for each trait SES of CWVs and CVNNDs after stepwise selection of the predictors oncluding slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
+See table 1 for abbreviations.
+The number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+
+```{r CWV}
+# Observed CWVs
+CWV <- mapply(function(Qs, CWMs, Ls){
+  lapply(names(CWMs$abundance), function(v){
+    apply(cbind(data.frame(CWMs$abundance[[which(names(CWMs$abundance) == v)]]),Ls$abundance), 1, function(x){
+      sqrt(weighted.mean((Qs[,v]-x[1])^2, x[-1], na.rm=T))
+    })
+  })
+}, Qs = Q, CWMs = CWM, Ls = L, SIMPLIFY = F)
+CWV <- lapply(CWV, function(s){
+  names(s) <- names(CWM$C$abundance)
+  return(s)
+})
+
+# Simulated CWVs 
+nL <- list(C = rep(list(L$C$abundance), n),
+           U = rep(list(L$U$abundance), n))
+nL <- lapply(nL, function(s){
+  lapply(s,function(l){
+    apply(l, 1, function(x){
+      x[x != 0] <- sample(x[x != 0])
+      return(x)
+    })
+  }) 
+})
+nL <- lapply(nL, function(s){
+  lapply(s, function(x){
+    t(x)
+  })
+})
+nL <- lapply(nL, function(s){
+  lapply(s, data.frame)
+})
+cl <- makeCluster(cores)
+clusterExport(cl, list('nL', 'CWM', 'Q'))
+nCWM.cwv <- clusterMap(cl, function(nLs, CWMs, Qs){
+  lapply(nLs, function(nl){
+    lapply(as.list(names(CWMs$abundance)), function(v){
+      apply(nl, 1, function(x){
+        weighted.mean(Qs[,v], x, na.rm=T)
+      })
+    })
+  })
+}, nLs = nL, CWMs = CWM, Qs = Q, SIMPLIFY = F)
+stopCluster(cl)
+nCWM.cwv <- lapply(nCWM.cwv, function(s){
+  lapply(s, function(x){
+    names(x) <- names(CWM$C$abundance)
+    return(x)
+  })
+})
+nCWV <- mapply(function(nLs, nCWM.cwvs, Qs){
+  mapply(function(l,cwm){
+  lapply(names(CWM$C$abundance), function(v){
+    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
+      sqrt(weighted.mean((Qs[,v]-x[1])^2, x[-1], na.rm=T))
+    })
+  })
+}, l = nLs, cwm = nCWM.cwvs, SIMPLIFY = F)
+}, nLs = nL, nCWM.cwvs = nCWM.cwv, Qs = Q, SIMPLIFY = F)
+nCWV <- lapply(nCWV, function(s){
+  lapply(s, function(x){
+    names(x) <- names(CWM$C$abundance)
+    return(x)
+  })
+})
+rm(nCWM.cwv, cl, nL)
+invisible(gc())
+## Non null CWVs
+nCWV <- lapply(nCWV, function(s){
+  lapply(s, `[`, names(s[[1]]))
+  apply(do.call(rbind, s), 2, as.list)
+})
+nCWV <- lapply(nCWV, function(s){
+  lapply(s, function(x){
+    do.call('cbind', x)
+  })
+})
+nCWV <- lapply(nCWV, function(s){
+  lapply(s, data.frame)
+})
+CWV.pval <- mapply(function(CWVs, nCWVs){
+  data.frame(
+  mapply(function(x,y){
+    mapply(function(p,q){
+      rank(c(p,q))[1]/(n+1)
+    }, p = as.list(x), q = as.list(data.frame(t(y))))
+  }, x = CWVs, y = nCWVs)
+)
+}, CWVs = CWV, nCWVs = nCWV, SIMPLIFY = F)
+CWV.pval.res <- lapply(CWV.pval, function(s){
+  data.frame(
+    inf = apply(apply(s, 2, p.adjust, method = 'fdr'), 2, function(x){
+      length(which(x < 0.025))/length(x)
+    }), 
+    sup = apply(apply(1 - s, 2, p.adjust, method = 'fdr'), 2, function(x){
+      length(which(x < 0.025))/length(x)
+    }))
+})
+## CWV SES
+CWV.ses <- mapply(function(CWVs, nCWVs){data.frame(
+  mapply(function(x,y){
+    mapply(function(p,q){
+      (p - mean(q))/sd(q)
+    }, p = as.list(x), q = as.list(data.frame(t(y))))
+  }, x = CWVs, y = nCWVs)
+)}, CWVs = CWV, nCWVs = nCWV, SIMPLIFY = F)
+CWV.ses.lm <- mapply(function(s, Rs){
+  apply(s, 2, function(y){
+    lm(y ~., Rs)
+  })
+}, s = CWV.ses, Rs = R, SIMPLIFY = F)
+CWV.ses.lm <- mapply(function(s, Rs){
+  lapply(s, function(x){
+    step(x, direction = 'both', trace = 0)
+  })
+}, s= CWV.ses.lm, Rs = R, SIMPLIFY = F)
+CWV.ses.res <- lapply(CWV.ses.lm, function(s){
+  lapply(s, function(x){
+    summary(x)$coefficients
+  })
+})
+CWV.ses.res <- lapply(CWV.ses.res, function(s){
+  lapply(s, data.frame)
+})
+CWV.ses.res <- lapply(CWV.ses.res, function(s){
+  lapply(s, function(x){
+    x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   '))
+    return(x)
+  })
+})
+CWV.ses.res <- lapply(CWV.ses.res, function(s){
+  data.frame(data.table::rbindlist(lapply(s, function(x){
+    data.frame(t(x[5]))
+  }), fill = T))
+})
+var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
+CWV.ses.res <- lapply(CWV.ses.res, function(s){
+  s[,setdiff(var, names(s))] <- NA
+  s <- s[var]
+  s <- apply(s, 2, as.character)
+  row.names(s) <- names(CWV.ses$C)
+  s[is.na(s)] <- ''
+  s <- data.frame(s)
+  return(s)
+})
+CWV.ses.res <- mapply(function(s, CWV.ses.lms){
+  s$`$R^2$` <- round(unlist(lapply(CWV.ses.lms, function(x){
+    summary(x)$r.squared
+  })),3)
+  return(s)
+}, s = CWV.ses.res, CWV.ses.lms = CWV.ses.lm, SIMPLIFY = F)
+CWV.ses.res <- lapply(CWV.ses.res, function(s){
+  s <- cbind(row.names(s), s)
+  names(s)[1] <- 'Trait'
+  return(s)
+})
+CWV.ses.res <- lapply(CWV.ses.res, function(s){
+  sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
+  any(grepl('*', t, fixed = T))
+})))
+  return(s[sel,])
+})
+CWV.ses.res <- mapply(function(s, name){
+  s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
+  names(s)[1] <- 'Strata'
+  return(s)
+}, s = CWV.ses.res, name = as.list(c('C', 'U')), SIMPLIFY = F)
+CWV.ses.res <- do.call('rbind', CWV.ses.res)
+rm(CWV.pval, CWV.pval.res, CWV.ses.lm, nCWV)
+```
+
+```{r CVNND}
+# Observed CVNND
+CVNND <- mapply(function(Ls, Qs){
+  apply(Ls$abundance, 1, function(l){
+  apply(Qs, 2, function(q){
+    cati::CVNND(rep(q, l), na.rm = T)
+    })
+  })
+}, Ls = L, Qs = Q, SIMPLIFY = F)
+
+# Simulated CVNNDs 
+cl <- makeCluster(cores)
+clusterExport(cl, list('nQ', 'L'))
+nCVNND <- clusterMap(cl, function(nQs, Ls){ 
+  lapply(nQs, function(x){
+    apply(Ls$abundance, 1, function(l){
+      apply(x, 2, function(q){
+        cati::CVNND(rep(q,l), na.rm = T)
+      })
+    })
+  })
+}, nQs = nQ, Ls = L)
+stopCluster(cl)
+rm(cl, nQ)
+invisible(gc)
+
+## Non null CVNNDs
+nCVNND <- lapply(nCVNND, function(s){
+  lapply(s, function(x){
+    as.list(data.frame(t(x)))
+  })
+})
+nCVNND <- lapply(nCVNND, function(s){
+  lapply(s, function(x){
+    lapply(x, function(y){
+      names(y) <- colnames(CVNND$C)
+      return(y)
+    })
+  })
+})
+nCVNND <- lapply(nCVNND, function(s){
+  lapply(s, `[`, names(s[[1]]))
+  apply(do.call(rbind, s), 2, as.list)
+})
+nCVNND <- lapply(nCVNND, function(s){
+  lapply(s, function(x){
+    do.call('cbind', x)
+  })
+})
+nCVNND <- lapply(nCVNND, function(s){
+  lapply(s, data.frame)
+})
+CVNND.pval <- mapply(function(CVNNDs, nCVNNDs){
+  data.frame(
+    mapply(function(x,y){
+      mapply(function(p,q){
+        rank(c(p,q))[1]/(n+1)
+      }, p = as.list(x), q = as.list(data.frame(t(y))))
+    }, x = as.list(data.frame(t(CVNNDs))), y = nCVNNDs)
+  )
+}, CVNNDs = CVNND, nCVNNDs = nCVNND, SIMPLIFY = F)
+CVNND.pval <- lapply(CVNND.pval, function(s){
+  row.names(s) <- colnames(CVNND)
+  apply(s, 2, p.adjust, method = 'fdr')
+  return(s)
+})
+CVNND.pval.res <- lapply(CVNND.pval, function(s){data.frame(
+  inf = apply(s, 2, function(x){
+    length(which(x < 0.05))/length(x)
+  })
+)})
+CVNND.pval.res <- do.call('cbind', CVNND.pval.res)
+names(CVNND.pval.res) <- names(CVNND.pval)
+print(round(CVNND.pval.res, 3))
+
+## CVNND SES
+CVNND.ses <- mapply(function(CVNNDs, nCVNNDs){
+  data.frame(
+    mapply(function(x,y){
+      mapply(function(p,q){
+        (p - mean(q))/sd(q)
+      }, p = as.list(x), q = as.list(data.frame(t(y))))
+    }, x = as.list(data.frame(t(CVNNDs))), y = nCVNNDs)
+  )
+}, CVNNDs = CVNND, nCVNNDs = nCVNND, SIMPLIFY = F)
+CVNND.ses <- lapply(CVNND.ses, function(s){
+  row.names(s) <- colnames(CVNND)
+  return(s)
+})
+CVNND.ses.lm <- mapply(function(s, Rs){
+  apply(s, 2, function(y){
+    lm(y ~., Rs)
+  })
+}, s = CVNND.ses, Rs = R, SIMPLIFY = F)
+CVNND.ses.lm <- mapply(function(s, Rs){
+  lapply(s, function(x){
+    step(x, direction = 'both', trace = 0)
+  })
+}, s= CVNND.ses.lm, Rs = R, SIMPLIFY = F)
+CVNND.ses.res <- lapply(CVNND.ses.lm, function(s){
+  lapply(s, function(x){
+    summary(x)$coefficients
+  })
+})
+CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
+  lapply(s, data.frame)
+})
+CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
+  lapply(s, function(x){
+    x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   '))
+    return(x)
+  })
+})
+CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
+  data.frame(data.table::rbindlist(lapply(s, function(x){
+    data.frame(t(x[5]))
+  }), fill = T))
+})
+var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
+CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
+  s[,setdiff(var, names(s))] <- NA
+  s <- s[var]
+  s <- apply(s, 2, as.character)
+  row.names(s) <- names(CVNND.ses$C)
+  s[is.na(s)] <- ''
+  s <- data.frame(s)
+  return(s)
+})
+CVNND.ses.res <- mapply(function(s, CVNND.ses.lms){
+  s$`$R^2$` <- round(unlist(lapply(CVNND.ses.lms, function(x){
+    summary(x)$r.squared
+  })),3)
+  return(s)
+}, s = CVNND.ses.res, CVNND.ses.lms = CVNND.ses.lm, SIMPLIFY = F)
+CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
+  s <- cbind(row.names(s), s)
+  names(s)[1] <- 'Trait'
+  return(s)
+})
+CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
+  sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
+  any(grepl('*', t, fixed = T))
+})))
+  return(s[sel,])
+})
+CVNND.ses.res <- mapply(function(s, name){
+  s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
+  names(s)[1] <- 'Strata'
+  return(s)
+}, s = CVNND.ses.res, name = as.list(c('C', 'U')), SIMPLIFY = F)
+CVNND.ses.res <- do.call('rbind', CVNND.ses.res)
+rm(CVNND.pval, CVNND.pval.res, CVNND.ses.lm, nCVNND)
+invisible(gc())
+```
+
+```{r CWV & CVNND table}
+res <- list(CWV = CWV.ses.res, CVNND = CVNND.ses.res)
+res <- mapply(function(s, name){
+  s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
+  names(s)[1] <- 'Type'
+  return(s)
+}, s = res, name = as.list(c('CWV', 'CVNND')), SIMPLIFY = F)
+res <- do.call('rbind', res)
+kable(bold(res), row.names = F)
+rm(CWV.ses.res, CVNND.ses.res, res)
+```
+
+## Taxonomic, functional and phylogenetic varitions among communities
+
+Taxonomic variation among communities was correlated significantly to environment variation (Mantel test, $r_{I_{st},E} = 0.23$ with $p-value < 0.001$), but not with remaining spatial dependency in community composition (Mantel test of space knowing environment, $r_{I_{st},S|E} = 0.01$ with $p-value > 0.1$).
+
+Overall, the global functional variation among communities was neither correlated to environment nor to spactial distance knowing the environment effect (see table 5).
+However, the variation of WD, in abundance as well as in presence absence, was positively correlated to environmental variation and is negatively correlated in presence absence to the spatial structure knowing environment spatial effect.
+
+While LDMC did not phylogenetic structure (Moran's I of 0 with a $p-value > 0.1$), LA and LT were phylogenetically clustered (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA were weakly clustered (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
+Phylogenetic variation among communities was neither correlated to environmental variation (Mantel test, $r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor to remaining spatial variation in partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).
+
+```{r Ist}
+library(spacodiR)
+
+# R,L,Q wihtout strata
+R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+row.names(R) <- com@data$id
+L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+L[is.na(L)] <- 0
+Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')]
+row.names(Q) <- PFT_sp$SpCode
 sel <- intersect(row.names(Q), names(L))
@@
 L$`presence-absence`[L$`presence-absence` > 1]  <- 1
-rm(Deriv, XY, com, Env, Trees, sel, BA)
-```
-
-## Functional composition of local communities
-
-Community weighted means (CWM), for both abundant and rare species, showed some significant variations along environmental gradients (see table 3).
-Community mean values of WD significantly increased under south-west direction, both in abundance and presence-absence CWM,  whereas SLA and LA, in presence-absence only, decreased with curvature.
-In addition, LDMC in presence-absence was decreasing with canopy height.
-**Develop meaning of AB et PA, specifically in term of species**.
-
-Table 3: Summary of linear models of community-level trait values.
-_The results are shown for either CWM values weighted either by species relative abundance (AB) or in presence-absence (PA) after stepwise selection of the predictors including slope, curvature, wetness, southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Global and partial F-statistics of final models were compared to statistics of 999 null linear regressions built with null model NM1, which randomizes traits among species.
-See table 1 for abbreviations.
-The number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-```{r CWM}
-# Observed CWMs
-CWM <- lapply(L, function(l){lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){apply(l, 1, function(x){weighted.mean(Q[,v], x, na.rm=T)})})})
-CWM <- lapply(CWM, function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT') ; return(x)})
-## Regressions
-reg <- paste('~', paste(names(R), collapse = ' + '))
-CWM.lm <- lapply(CWM, function(x){lapply(x, function(y){lm(y ~., R)})})
-CWM.lm <- lapply(CWM.lm, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
-## Global F
-Fval <- lapply(CWM.lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
-Fval <- lapply(Fval, function(x){do.call('rbind.data.frame', x)})
-Fval <- data.frame(Fval)
-names(Fval) <- names(CWM.lm)
-## Partial F
-pFval <- lapply(CWM.lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})
-
-# Simulated CWMs
-n = 999
-nQ <- rep(list(Q), n)
-nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
-nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
-cl <- makeCluster(cores)
-clusterExport(cl, list('L', 'nQ'))
-nCWM <- clusterApply(cl, L, function(l){
-  lapply(nQ, function(q){
-    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
-      apply(l, 1, function(x){
-        weighted.mean(q[,v], x, na.rm=T)
-      })
-    })
-  })
-})
-stopCluster(cl)
-rm(cl)
-nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT') ; return(x)})})
-## Regressions
-nCWM.lm <- mapply(function(x,y){
-  lapply(y, function(z){
-    mapply(function(r,s){
-      lm(formula(paste0('s ~ ', as.character(r$terms)[3])), cbind(s,R))
-    }, r = x, s = z, SIMPLIFY = F)
-  })
-}, x = CWM.lm, y = nCWM, SIMPLIFY = F)
-rm(nCWM) ; invisible(gc())
-## Global F
-nFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
-nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
-nFval <- lapply(nFval, function(f){lapply(f, unlist)})
-pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
-row.names(pval) <- row.names(Fval)
-rm(Fval, nFval)
-# Partial F
-npFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})})
-npFval <- lapply(npFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
-npFval <- lapply(npFval, function(f){lapply(f, function(x){do.call('cbind', x)})})
-ppval <- mapply(function(x,y){mapply(function(r,s){mapply(function(p,q){rank(c(p,q))[1]/(n+1)}, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))}, r = x, s = y)}, x = pFval, y = npFval, SIMPLIFY = F)
-ppval <- lapply(ppval, function(y){lapply(y, function(x){data.frame(t(x))})})
-ppval <- lapply(ppval, function(y){data.frame(data.table::rbindlist(y, fill = T))})
-ppval <- lapply(ppval, function(y){row.names(y) <- names(CWM.lm$abundance) ; return(y)})
-ppval$abundance <- ppval$abundance[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy')]
-ppval$abundance$BA <- NA
-ppval$`presence-absence` <- ppval$`presence-absence`[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
-ppval <- lapply(ppval, function(x){1 - x})
-rm(pFval, npFval)
-
-# Result tables
-t1 <- regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]])
-t1 <- cbind(c('AB', rep(' ', length(t1[,1])-1)), t1)
-names(t1)[1] <- ' '
-t2 <- regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]])
-t2 <- cbind(c('PA', rep(' ', length(t2[,2])-1)), t2)
-names(t2)[1] <- ' '
-kable(bold(rbind(t1,t2)), row.names = F)
-rm(CWM.lm, nCWM.lm, pval, ppval, reg, t1, t2) ; invisible(gc())
-```
-
-Traits convergence/divergence was measured with community weighted variances (CWV).
-Overall, observed CWVs DID NOT differ from null expectations for all traits.
-We could found significant relationships between standardized effect size (SES) of CWVs and environmental variables (see table 4).
-Convergence of LES and LDMC and divergence of LA were found under south-west orientation.
-LA and LDMC values also diverged in communities with increasing curvature.
-Observed CVNNDs did not reveal niche differentiation among communities for all traits.
-But we also found significant relationships of their standardized effect size (SES) of CVNNDs with environmental variables (see table 4).
-niche differentiation of LA increased in lower curvature and wetness contexts, while differentiation of WD decreased in southWest orientation.
-
-Table 4: Summary of linear models of traits standardized effect size (SES) of CWVs and CVNNDs.
-_Linear regression results for each trait SES of CWVs and CVNNDs after stepwise selection of the predictors oncluding slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
-See table 1 for abbreviations.
-The number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-```{r CWV}
-# Observed CWVs
-CWV <- lapply(names(CWM$abundance), function(v){apply(cbind(data.frame(CWM$abundance[[which(names(CWM$abundance) == v)]]),L$abundance), 1, function(x){sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))})})
-names(CWV) <- names(CWM$abundance)
-
-# Simulated CWVs 
-nL <- rep(list(L$abundance), n)
-nL <- lapply(nL,function(l){apply(l, 1, function(x){x[x != 0] <- sample(x[x != 0]) ; return(x)})}) 
-nL <- lapply(nL, function(x){t(x)})
-nL <- lapply(nL, data.frame)
-
-cl <- makeCluster(cores)
-clusterExport(cl, list('nL', 'CWM', 'Q'))
-nCWM.cwv <- clusterApply(cl, nL, function(l){
-  lapply(as.list(names(CWM$abundance)), function(v){
-    apply(l, 1, function(x){
-      weighted.mean(Q[,v], x, na.rm=T)
-    })
-  })
-})
-stopCluster(cl)
-nCWM.cwv <- lapply(nCWM.cwv, function(x){names(x) <- names(CWM$abundance) ; return(x)})
-nCWV <- mapply(function(l,cwm){
-  lapply(names(CWM$abundance), function(v){
-    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
-      sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))
-    })
-  })
-}, l = nL, cwm = nCWM.cwv, SIMPLIFY = F)
-nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
-rm(nCWM.cwv, cl, nL) ; invisible(gc())
-## Non null CWVs
-nCWV <- lapply(nCWV, `[`, names(nCWV[[1]]))
-nCWV <- apply(do.call(rbind, nCWV), 2, as.list)
-nCWV <- lapply(nCWV, function(x){do.call('cbind', x)})
-nCWV <- lapply(nCWV, data.frame)
-CWV.pval <- data.frame(mapply(function(x,y){
-  mapply(function(p,q){
-    rank(c(p,q))[1]/(n+1)
-  }, p = as.list(x), q = as.list(data.frame(t(y))))
-}, x = CWV, y = nCWV))
-CWV.pval.res <- data.frame(inf = apply(apply(CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(apply(1 - CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}))
-## CWV SES
-CWV.ses <- data.frame(mapply(function(x,y){
-  mapply(function(p,q){
-    (p - mean(q))/sd(q)
-  }, p = as.list(x), q = as.list(data.frame(t(y))))
-}, x = CWV, y = nCWV))
-CWV.ses.lm <- apply(CWV.ses, 2, function(y){lm(y ~., R)})
-CWV.ses.lm <- lapply(CWV.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
-CWV.ses.res <- lapply(CWV.ses.lm, function(x){summary(x)$coefficients})
-CWV.ses.res <- lapply(CWV.ses.res, data.frame)
-CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
-CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
-CWV.ses.res <- CWV.ses.res[-1]
-var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
-CWV.ses.res[,setdiff(var, names(CWV.ses.res))] <- NA
-CWV.ses.res <- CWV.ses.res[var]
-CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
-row.names(CWV.ses.res) <- names(CWV.ses)
-CWV.ses.res[is.na(CWV.ses.res)] <- ''
-CWV.ses.res <- data.frame(CWV.ses.res)
-CWV.ses.res$`$R^2$` <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)
-
-# Results
-# barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
-CWV.ses.res <- cbind(c('CWV', rep(' ', length(CWV.ses.res[,1])-1)), row.names(CWV.ses.res), CWV.ses.res)
-names(CWV.ses.res)[1:2] <- c(' ', 'Trait')
-rm(CWV.pval, CWV.pval.res, CWV.ses.lm, nCWV)
-```
-
-```{r CVNND}
-# Observed CVNND
-CVNND <- apply(L$abundance, 1, function(l){apply(Q, 2, function(q){cati::CVNND(rep(q, l), na.rm = T)})})
-
-# Simulated CVNNDs 
-cl <- makeCluster(cores)
-clusterExport(cl, list('nQ', 'L'))
-nCVNND <- clusterApply(cl, nQ, function(x){
-  apply(L$abundance, 1, function(l){
-    apply(x, 2, function(q){
-      cati::CVNND(rep(q,l), na.rm = T)
-    })
-  })
-})
-stopCluster(cl) ; rm(cl)
-rm(nQ) ; invisible(gc)
-## Non null CVNNDs
-nCVNND <- lapply(nCVNND, function(x){as.list(data.frame(t(x)))})
-nCVNND <- lapply(nCVNND, function(x){lapply(x, function(y){names(y) <- colnames(CVNND) ; return(y)})})
-nCVNND <- lapply(nCVNND, `[`, names(nCVNND[[1]]))
-nCVNND <- apply(do.call(rbind, nCVNND), 2, as.list)
-nCVNND <- lapply(nCVNND, function(x){do.call('cbind', x)})
-nCVNND <- lapply(nCVNND, data.frame)
-CVNND.pval <- data.frame(mapply(function(x,y){
-  mapply(function(p,q){
-    rank(c(p,q))[1]/(n+1)
-  }, p = as.list(x), q = as.list(data.frame(t(y))))
-}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
-row.names(CVNND.pval) <- colnames(CVNND)
-CVNND.pval <- apply(CVNND.pval, 2, p.adjust, method = 'fdr')
-CVNND.pval.res <- data.frame(inf = apply(CVNND.pval, 2, function(x){length(which(x < 0.05))/length(x)}))
-## CVNND SES
-CVNND.ses <- data.frame(mapply(function(x,y){
-  mapply(function(p,q){
-    (p - mean(q))/sd(q)
-  }, p = as.list(x), q = as.list(data.frame(t(y))))
-}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
-row.names(CVNND.ses) <- colnames(CVNND)
-CVNND.ses.lm <- apply(CVNND.ses, 2, function(y){lm(y ~., R)})
-CVNND.ses.lm <- lapply(CVNND.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
-CVNND.ses.res <- lapply(CVNND.ses.lm, function(x){summary(x)$coefficients})
-CVNND.ses.res <- lapply(CVNND.ses.res, data.frame)
-CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
-CVNND.ses.res <- data.frame(data.table::rbindlist(lapply(CVNND.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
-CVNND.ses.res <- CVNND.ses.res[-1]
-CVNND.ses.res[,setdiff(var, names(CVNND.ses.res))] <- NA
-CVNND.ses.res <- CVNND.ses.res[var]
-CVNND.ses.res <- apply(CVNND.ses.res, 2, as.character)
-row.names(CVNND.ses.res) <- names(CVNND.ses)
-CVNND.ses.res[is.na(CVNND.ses.res)] <- ''
-CVNND.ses.res <- data.frame(CVNND.ses.res)
-CVNND.ses.res$`$R^2$` <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)
-
-# Results
-# barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
-CVNND.ses.res <- cbind(c('CVNND', rep(' ', length(CVNND.ses.res[,1])-1)), row.names(CVNND.ses.res), CVNND.ses.res)
-names(CVNND.ses.res)[1:2] <- c(' ', 'Trait')
-kable(bold(rbind(CWV.ses.res,CVNND.ses.res)), row.names = F)
-rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
-```
-
-## Taxonomic, functional and phylogenetic varitions among communities
-
-Taxonomic variation among communities was correlated significantly to environment variation (Mantel test, $r_{I_{st},E} = 0.23$ with $p-value < 0.001$), but not with remaining spatial dependency in community composition (Mantel test of space knowing environment, $r_{I_{st},S|E} = 0.01$ with $p-value > 0.1$).
-
-Overall, the global functional variation among communities was neither correlated to environment nor to spactial distance knowing the environment effect (see table 5).
-However, the variation of WD, in abundance as well as in presence absence, was positively correlated to environmental variation and is negatively correlated in presence absence to the spatial structure knowing environment spatial effect.
-
-While LDMC did not phylogenetic structure (Moran's I of 0 with a $p-value > 0.1$), LA and LT were phylogenetically clustered (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA were weakly clustered (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
-Phylogenetic variation among communities was neither correlated to environmental variation (Mantel test, $r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor to remaining spatial variation in partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).
-
-```{r Ist}
-library(spacodiR)
+rm(com, Trees)
+invisible(gc())
 
@@
 ```{r Ust & TAUst}
+# nQ without strata
+n = 999
+nQ <- rep(list(Q), n)
+nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
+nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
+
 # Observed Hardys

### 2016-08-09 R/regression_table.R

@@ -36,6 +36,8 @@ regression_table <- function(LM, reg, pval, ppval){
   var_table$`Null model` <- round(pval,3)
   var_table$` ` <- unlist(lapply(pval, stars, ns = ' '))
   var_table <- var_table[-which(var_table$` ` == ' '),]
+  var_table$`Null model` <- paste(var_table$`Null model`, var_table$` `)
+  var_table <- var_table[-length(names(var_table))]
 
   return(var_table)
 }

### 2016-08-11 inst/doc/draft.Rmd
v0.1.9032 Corrected Strata and 30*30m2 analysis

@@
 We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1).
-**Justify trait choice**.
+We choose those five traits because they have been widely used, besides there easier measurement, and they have been shown to be good indicator of species abilities to grow, survive and reproduce [@wilson_specific_1999].
 
 Table 1: Functional traits measured.
-*Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*
+*Traits with their abbreviation, standard unit, and associated trade-off in tree strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*
 
@@
   Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
-  Range = c(paste(range(PFT$LT, na.rm = T), collapse = '-'),
-            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
-            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
-            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
-            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
-  m = c(paste0(round(mean(PFT$LT, na.rm = T)), ' (',
-                         round(sd(PFT$LT, na.rm = T)), ')'),
-                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
-                         round(sd(PFT$LA, na.rm = T)), ')'),
-                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
-                         round(sd(PFT$LDMC, na.rm = T)), ')'),
-                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
-                         round(sd(PFT$SLA, na.rm = T),2), ')'),
-                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
-                         round(sd(PFT$WD, na.rm = T),2), ')')),
   Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
@@
 )
-names(table)[c(4,6)] <- c('Mean (sd)', 'Trade-off')
+names(table)[4] <- 'Trade-off'
 kable(table, row.names = F)
@@
 
-We selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from the digital elevation model using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed using ArcGis [@desktop_arcgis_2011].
-Digital elevation model was provided by a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
-**Justify variable choice**.
+Because topography was assumed as a major source of environmental variation [@gimaret-carpentier_sampling_1998], we selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from the digital elevation model using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed using ArcGis [@desktop_arcgis_2011].
+<!-- Digital elevation model was provided by a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. -->
 
@@
 
-Across the whole 10 hectare plot, $20*20 m^2$ quadrats were defined and represented 240 local communities.
-To investigate local functional composition of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010].
+Across the whole 10 hectare plot, $30*30 m^2$ quadrats were defined and represented 110 local communities.
+Communities where split in two guilds representing canopy and emergent species and understorey species following their potential maximum strata as defined by **cite**.
+To investigate local functional composition of canopy and understorey communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010].
 CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative abundance, $p_{i}$:
 $$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
-We also calculated the average of trait values based on species presence-absence, to further emphasize the contribution of rare species in community trait distribution [@cingolani_filtering_2007].
+We also calculated the average of trait values based on species presence-absence, to further emphasize the contribution of less abundant and rare species in community trait distribution [@cingolani_filtering_2007].
 CWV for each trait $t$ was calculated as the mean distance of trait values, $t_{i}$, to $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
@@
 To investigate taxonomic, functional, and phylogenetic variation among communities we calculated $I_{st}$, for taxonomic composition, $U_{st}$ and $\tau _{st}$, for functional composition in abundance and presence-absence respectively, and $\beta _{st}$ and $\pi _{st}$, for phylogenetic composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
-**Develop indices**.
+$I_{st}$ measures species turnover accounting for local abundance.
+$U_{st}$ and $\beta _{st}$ integrate only phylogenetic and functional turnover, respectively, weighting species by their local abundances; whereas $\tau _{st}$ and $\pi _{st}$ express phylogenetic and functional turnover, respectively, without accounting for species local abundances.
 
@@
 Finally $\beta st$ and $\pi st$ were tested against the nul model *"1s"* suggested by @hardy_testing_2008 which randomizes phylogenetic tree tips.
-**One global null model section ?**
 
@@
 Whereas LA shows already high variabililty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
-**Conclude on low $cv_{intra}$ justifying species mean trait**.
-**Give Species example of extreme range and compare to litterature**.
+Finally the relatively low intra-individual variation ($cv_{intra}$ range from 0.06 to 0.09) allow us to consider the mean trait values per species for subsequent connumity-level analysis.
 
 Table 2: Coefficient of intra-individual, intra-specific and inter-specific trait variation.
-*Mean of coefficient of variation for leaves in all individuals (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
-See table 1 for abbreviation.* **Move range, mean and sd columns here*
+*Traits with their range of values, mean and standard deviation, mean of coefficient of variation for leaves in all individuals (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
+See table 1 for abbreviation.*
 
@@
 PFT_ind <- PFT_ind[-c(1:2)]
-cv <- data.frame(row.names = c('LT', 'LA', 'LDMC', 'SLA', 'WD'))
-cv$Individual <- unlist(lapply(
-  as.list(row.names(cv)),
-  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
-cv$Species <- unlist(lapply(
-  as.list(row.names(cv)),
-  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$Sp), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
-cv$Community <- unlist(
-  lapply(as.list(row.names(cv)),
-         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
-             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
-cv <- round(cv, 2)
+trait <- c('LT', 'LA', 'LDMC', 'SLA', 'WD')
+cv <- data.frame(
+  row.names = trait,
+  Range = c(paste(range(PFT$LT, na.rm = T), collapse = '-'),
+            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
+            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
+            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
+            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
+  m = c(paste0(round(mean(PFT$LT, na.rm = T)), ' (',
+               round(sd(PFT$LT, na.rm = T)), ')'),
+        paste0(round(mean(PFT$LA, na.rm = T)), ' (',
+               round(sd(PFT$LA, na.rm = T)), ')'),
+        paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
+               round(sd(PFT$LDMC, na.rm = T)), ')'),
+        paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
+               round(sd(PFT$SLA, na.rm = T),2), ')'),
+        paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
+               round(sd(PFT$WD, na.rm = T),2), ')')),
+  Individual = round(unlist(lapply(as.list(trait), function(y){
+    mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){
+      sd(x, na.rm = T) / mean(x, na.rm = T)
+    })[,2], na.rm = T)
+  })) ,2),
+  Species = round(unlist(lapply(as.list(trait), function(y){
+    mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$Sp), function(x){
+      sd(x, na.rm = T) / mean(x, na.rm = T)
+    })[,2], na.rm = T)
+  })), 2),
+  Community = round(unlist(
+    lapply(as.list(trait), function(y){
+      sd(PFT[,which(names(PFT) ==  y)], na.rm = T) / mean(PFT[,which(names(PFT) ==  y)], na.rm = T)
+    })), 2)
+)
+names(cv)[2] <- 'Mean (sd)' 
 cv['WD','Individual'] <- ''
 kable(cv)
-rm(PFT, CET, cv, CE, Sp, SpT)
+rm(PFT, CET, cv, triat, CE, Sp, SpT)
 ```
@@
 The principal component analysis (PCA) shows WD associated with structural leaf traits orthogonal to the SLA (figure 1).
-The first PCA axis (PCA1) covary mainly with LT, LDMC and WD; opposing thick big leaves and lightwood to dense leaves and hardwood.
+The first PCA axis (PCA1) covary mainly with LT, LA, LDMC and WD; opposing thick big leaves and lightwood to dense leaves and hardwood.
 PCA1 axis associated WD to leaf structural traits linked to hydraulic conductive efficiency along the wood economic spectrum [WES, @chave_towards_2009].
 The PCA2 covaried mainly positively with SLA and negatively with LDMC along a leaf economic spectrum [LES, @wright_worldwide_2004].
-**Develop productive against persistent strategy**.
-**High WD LDMC relation indicating structural investment**.
+The leaf econmic spectrum opposes high SLA leaves for a strong productivity capacity, with high phyotosynthesis compare to carbon investment, and high LDMC leaves to ensure persistence in time and against attacks.
+PCA also highlight the strong correlation between LDMC and WD (Pearson's correlation test, $r = 0.25$ with $p-value < 0.001$), indicating a coordinated structural investment of species among stem and leaf.
 
@@
 
-Figure 1: Principal Component Analysis (PCA) performed on species mean traot values.
+Figure 1: Principal Component Analysis (PCA) performed on species mean trait values.
 *The bottom right subplot represents the eigen values (grey bars) compare to a broken stick distribution for 5 components (red line).
@@
 Trees <- Trees[-which(names(Trees) == 'ID')]
-com <- quadrats()
+com <- quadrats(cs = c(30,30))
 XY <- Trees[c('x', 'y')]
@@
 
-Community weighted means (CWM), for both abundant and rare species, showed some significant variations along environmental gradients (see table 3).
+Community weighted means (CWM), both in abundance or presence-absence, showed some significant variations along environmental gradients (see table 3).
 Community mean values of WD significantly increased under south-west direction, both in abundance and presence-absence CWM,  whereas SLA and LA, in presence-absence only, decreased with curvature.
 In addition, LDMC in presence-absence was decreasing with canopy height.
-**Develop meaning of AB et PA, specifically in term of species**.
+Trait values that allow species existence is not always the same that trait values determining their ability to dominate a given environment.
+For instance here both abundant and less abundant species increased their WD under southwest orientation, whereas SLA and LA decrease with curvature is only affecting less abundant species ability to survive in concave environments.
 
@@
 })
-Fval$U <- Fval$U[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),]
 ## Partial F
@@
 names(CVNND.pval.res) <- names(CVNND.pval)
-print(round(CVNND.pval.res, 3))
 
@@
 CVNND.ses.res <- do.call('rbind', CVNND.ses.res)
-rm(CVNND.pval, CVNND.pval.res, CVNND.ses.lm, nCVNND)
+rm(CVNND.pval, CVNND.ses.lm, nCVNND)
 invisible(gc())
@@
 ```{r CWV & CVNND table}
+print(round(CVNND.pval.res, 3))
 res <- list(CWV = CWV.ses.res, CVNND = CVNND.ses.res)
@@
   s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
-  names(s)[1] <- 'Type'
+  names(s)[1:2] <- c('T', 'S')
   return(s)
@@
 kable(bold(res), row.names = F)
-rm(CWV.ses.res, CVNND.ses.res, res)
+rm(CWV.ses.res, CVNND.ses.res, CVNND.pval.res, res)
 ```
@@
 @bunyan_effect_2015 also observed similar influence in mountain rainforests of Western Ghats, and suggested that it could result from monsoon winds.
-In our context, monsoon winds arise from south-west direction, and bring more stressful and disturbed conditions to species [@soman_aspects_1990] **To develop, see maxime disturbance hypothesis on opposite side**.
+In our context, monsoon winds arise from south-west direction, and could bring more stressful and disturbed conditions to species [@soman_aspects_1990].
+Uppangala plot is situated inside a low elevation slope separated from west coast by a higher crest orientated toward southwest.
+In this context it is diffcult to predict wind flow and orientation inside the 10 ha plot topgraphy.
+We hyptohesis to explanation, hardwood could be selected for mechanical resistance in more disturbed souwhtwest oriented area or topography provoke a turbulences on northeast oriented areas responsible of more disturbance and death events in the forest history resulting in lightwood filtering [@larjavaara_rethinking_2010].
 Curvature is also affecting the variation of leaf economics in community.
@@
 Biotic factors also affected the LES through a negative effect of canopy height on LDMC.
-There was no effect of canopy height on understorey species, while it influenced leaf traits of canopy and emergent trees (results unshown **To include in the analysis**).
+There was no effect of canopy height on understorey species, while it influenced leaf traits of canopy and emergent trees.
 A basic expectation is that understorey species would invest in higher LDMC for higher leaf lifespan and durability under limited light and water resources [@kitajima_tissue-level_2010].
@@
 However lower evenness than random could be related to multimodality.
-In addition, south-west orientation also related to a broader range of leaf sizes, which could be consistent with more diverse hydraulic strategies reflected by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008] **To develop, especially for a multimodality under southwest**.
+In addition, south-west orientation also related to a broader range of leaf sizes, which could be consistent with more diverse hydraulic strategies reflected by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008].
 Decreasing fertility , with concavity was related to smaller leaves inside communities with a distribution broaden by multimodality reducing evenness **to develop, eventually illustrated with one example**.

### 2016-08-11 R/genSpecies.R

@@ -25,10 +25,10 @@ genSpecies <- function(){
   species$ESIII <- 0; species[grep(x=species$PotStrata,pattern=", III"),]$ESIII <- 1; species[grep(x=species$PotStrata,pattern="ESIII"),]$ESIII <- 1
   species$ESIV <- 0; species[grep(x=species$PotStrata,pattern=", IV"),]$ESIV <- 1; species[grep(x=species$PotStrata,pattern="ESIV"),]$ESIV <- 1
   # Lot of species in different strata
-  # Ayyappan suggest to take the lowest value
-  species[species$ESI==1 & species$ESII==1,]$ESI <- 0
-  species[species$ESII==1 & species$ESIII==1,]$ESII <- 0
-  species[species$ESIII==1 & species$ESIV==1,]$ESIII <- 0
+  # Ayyappan suggest to keep the highest potential strata
+  species[species$ESI==1 & species$ESII==1,]$ESII <- 0
+  species[species$ESII==1 & species$ESIII==1,]$ESIII <- 0
+  species[species$ESIII==1 & species$ESIV==1,]$ESIV <- 0
   # Strata as factor
   species$Strata <- NA
   species$Strata[species$ESI==1] <- 'ESI'

### 2016-08-11 inst/doc/draft.Rmd
v0.1.9033 New strata pelissier, 30*30m2 com, & SES CWM

@@
 Across the whole 10 hectare plot, $30*30 m^2$ quadrats were defined and represented 110 local communities.
-Communities where split in two guilds representing canopy and emergent species and understorey species following their potential maximum strata as defined by **cite**.
+@pascal_wet_1988 defined four categories of strata groups: emergent below 39m, canopy tree between 25 and 39m, sub-canopy tree between 15 and 25m, and understorey tree below 15m.
+Due to topography, sub-canopy tree from Uppangala are often under light exposure and participates to canopy, especially in slopes.
+We defined two guilds to represent vertical stratification called canopy (C), composed of emergent, canopy and sub-canopy tree, and understorey (U) trees.
+The 110 local communities where split by those two guilds and communities accounting for only one species were removed from the subsequent analysis.
+
 To investigate local functional composition of canopy and understorey communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010].
@@
 ```{r 2 CV}
+# Individual aggregation
 PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
@@
 PFT_ind <- PFT_ind[-c(1:2)]
+PFT_ind$ID <- row.names(PFT_ind)
+row.names(PFT_ind) <- PFT_ind$ID
+PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
+# Species aggregation
+Species <- genSpecies()
+PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$Sp), FUN = mean, na.rm = T)
+row.names(PFT_sp) <- PFT_sp$Group.1
+PFT_sp$SpCode <- SpCode[row.names(PFT_sp)]
+PFT_sp <- PFT_sp[-c(1:2,4)]
+# Coefficient of variation
 trait <- c('LT', 'LA', 'LDMC', 'SLA', 'WD')
@@
                round(sd(PFT$WD, na.rm = T),2), ')')),
-  Individual = round(unlist(lapply(as.list(trait), function(y){
+  m1 = round(unlist(lapply(as.list(trait), function(y){
     mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){
@@
   })) ,2),
-  Species = round(unlist(lapply(as.list(trait), function(y){
+  m2 = round(unlist(lapply(as.list(trait), function(y){
     mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$Sp), function(x){
@@
   })), 2),
-  Community = round(unlist(
+  m3 = round(unlist(
     lapply(as.list(trait), function(y){
-      sd(PFT[,which(names(PFT) ==  y)], na.rm = T) / mean(PFT[,which(names(PFT) ==  y)], na.rm = T)
+      sd(PFT_sp[,which(names(PFT_sp) ==  y)], na.rm = T) / mean(PFT_sp[,which(names(PFT_sp) ==  y)], na.rm = T)
     })), 2)
 )
-names(cv)[2] <- 'Mean (sd)' 
-cv['WD','Individual'] <- ''
+names(cv)[2:5] <- c('Mean (sd)' , 'Intra-individual', 'Intra-species', 'Inter-species')
+cv['WD','Intra-individual'] <- ''
 kable(cv)
-rm(PFT, CET, cv, triat, CE, Sp, SpT)
+rm(PFT, CET, cv, CE, Sp, SpT, SpCodeT)
 ```
@@
 
+# Indovidual to log
 PFT_ind$LA <- log(PFT_ind$LA)
@@
 PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
+# Species aggregation
 Species <- genSpecies()
@@
 PFT_sp <- PFT_sp[-c(1:2,4)]
+# PCA
 pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT_sp, cor = T)
-
 lambda <- pca$sdev * sqrt(nrow(pca$scores))
@@
 
+# LES and WES
 PCAscores <- as.data.frame(pca$scores)[1:2]
@@
 library(raster)
+
+# Building community and environmental datasets
 Env <- genEnv()
@@
 pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
+
+# Building RLQ matrices
 R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
@@
 row.names(Q) <- PFT_sp$SpCode
-sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] %in% c('ESI', 'ESII'))
-Q <- list(C = Q[sel,], U = Q[-sel,])
+sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] == 3)
+Q <- list(C = Q[-sel,], U = Q[sel,])
 sel <- lapply(Q, function(q){intersect(row.names(q), names(L))})
@@
 L <- lapply(L, function(l){l$`presence-absence`[l$`presence-absence` > 1]  <- 1 ; return(l)})
-rm(Deriv, XY, Env, sel, BA)
+
+rm(Deriv, XY, Env, sel, BA, s)
 ```
@@
     return(x)
-  })
-})
-## Regressions
-CWM.lm <- mapply(function(CWMs, Rs){
-  lapply(CWMs, function(p){
-    lapply(p, function(t){
-      lm(t ~., Rs)
-    })
-  })
-}, CWMs = CWM, Rs = R, SIMPLIFY = F)
-CWM.lm <- mapply(function(s, Rs){
-  lapply(s, function(p){
-    lapply(p, function(t){
-      step(t, direction = 'both', trace = 0)
-    })
-  })
-}, s = CWM.lm, Rs = R, SIMPLIFY = F)
-## Global F
-Fval <- lapply(CWM.lm, function(s){
-  lapply(s, function(y){
-    lapply(y, function(x){
-      summary(x)$fstatistic['value']
-    })
-  })
-})
-Fval <- lapply(Fval, function(s){
-  lapply(s, function(x){
-    do.call('rbind.data.frame', x)
-  })
-})
-Fval <- lapply(Fval, function(s){
-  s <- data.frame(s)
-  names(s) <- names(CWM.lm$C)
-  return(s)
-})
-## Partial F
-pFval <- lapply(CWM.lm, function(s){
-  lapply(s, function(y){
-    lapply(y, function(x){
-      car::Anova(x)['F value']
-    })
   })
@@
 })
-## Regressions
-nCWM.lm <- mapply(function(CWM.lms, nCWMs, Rs){
-  lapply(nCWMs, function(ncwm){
-    mapply(function(CWM.lmab, ncwmab){
-      mapply(function(CWM.lmt, ncwmt){
-        lm(formula(paste0('ncwmt ~ ', as.character(CWM.lmt$terms)[3])), cbind(ncwmt,Rs))
-      }, CWM.lmt = CWM.lmab, ncwmt = ncwmab, SIMPLIFY = F)
-    }, CWM.lmab = CWM.lms, ncwmab = ncwm, SIMPLIFY = F)
-  })
-}, CWM.lms = CWM.lm, nCWMs = nCWM, Rs = R, SIMPLIFY = F)
-rm(nCWM)
-invisible(gc())
-## Global F
-nFval <- lapply(nCWM.lm, function(nCWM.lms){
-  lapply(nCWM.lms, function(lm){
-    lapply(lm, function(y){
-      lapply(y, function(x){
-        summary(x)$fstatistic['value']
-      })
-    })
-  })
-})
-nFval <- lapply(nFval, function(x){
-  lapply(x, `[`, names(x[[1]]))
-  apply(do.call(rbind, x), 2, as.list)
-})
-nFval <- lapply(nFval, function(s){
-  lapply(s, function(x){
-    lapply(x, `[`, names(x[[1]]))
-    apply(do.call(rbind, x), 2, as.list)
-  })
-})
-nFval <- lapply(nFval, function(s){
+nCWM <- lapply(nCWM, function(s){
+  lapply(s, `[`, names(s[[1]]))
+  apply(do.call(rbind, s), 2, as.list)
+})
+nCWM <- lapply(nCWM, function(s){
   lapply(s, function(ab){
-    lapply(ab, unlist)
-  })
-})
-pval <- mapply(function(Fvals, nFvals){
-  1 -  mapply(function(Fvalab, nFvalab){
-    mapply(function(Fvalt,nFvalt){
-      if(!is.na(Fvalt))
-        rank(c(Fvalt, nFvalt))[1]/(n+1)
-      else
-        NA
-    }, Fvalt = Fvalab, nFvalt = nFvalab)
-  }, Fvalab = lapply(as.list(Fvals), as.list), nFvalab = nFvals)
-}, Fvals = Fval, nFvals = nFval, SIMPLIFY = F)
-pval <- lapply(pval, function(s){
-  row.names(s) <- row.names(Fval$C)
-  return(s)
-})
-rm(Fval, nFval)
-# Partial F
-npFval <- lapply(nCWM.lm, function(s){
-  lapply(s, function(lm){
-    lapply(lm, function(y){
-      lapply(y, function(x){
-        car::Anova(x)['F value']
-      })
-    })
-  })
-})
-npFval <- lapply(npFval, function(x){
-  lapply(x, `[`, names(x[[1]]))
-  apply(do.call(rbind, x), 2, as.list)
-})
-npFval <- lapply(npFval, function(s){
-  lapply(s, function(x){
-    lapply(x, `[`, names(x[[1]]))
-    apply(do.call(rbind, x), 2, as.list)
-  })
-})
-npFval <- lapply(npFval, function(s){
+    lapply(ab, `[`, names(ab[[1]]))
+    apply(do.call(rbind, ab), 2, as.list)
+  })
+})
+nCWM <- lapply(nCWM, function(s){
   lapply(s, function(ab){
-    lapply(ab, function(t){
-      do.call('cbind', t)
-    })
-  })
-})
-ppval <- mapply(function(pFvals, npFvals){
+    lapply(ab, function(x){
+      do.call('cbind', x)
+    })
+  })
+})
+nCWM <- lapply(nCWM, function(s){
+  lapply(s, function(ab){
+    lapply(ab, data.frame)
+  })
+})
+
+## CWM SES
+CWM.ses <- mapply(function(CWMs, nCWMs){
+  mapply(function(CWMab, nCWMab){
+    data.frame(
       mapply(function(x,y){
-    mapply(function(r,s){
         mapply(function(p,q){
-        rank(c(p,q))[1]/(n+1)
-      }, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))
-    }, r = x, s = y)
-  }, x = pFvals, y = npFvals, SIMPLIFY = F)
-}, pFvals = pFval, npFvals = npFval, SIMPLIFY = F)
-ppval <- lapply(ppval, function(s){
-  lapply(s, function(y){
-    lapply(y, function(x){
-      data.frame(t(x))
-    })
-  })
-})
-ppval <- lapply(ppval, function(s){
-  lapply(s, function(y){
-    data.frame(data.table::rbindlist(y, fill = T))
-  })
-})
-ppval <- lapply(ppval, function(s){
-  lapply(s, function(y){
-    row.names(y) <- names(CWM.lm$C$abundance)
-    return(y)
+          (p - mean(q))/sd(q)
+        }, p = as.list(x), q = as.list(data.frame(t(y))))
+      }, x = CWMab, y = nCWMab)
+    )
+  }, CWMab = CWMs, nCWMab = nCWMs, SIMPLIFY = F)
+}, CWMs = CWM, nCWMs = nCWM, SIMPLIFY = F)
+CWM.ses.lm <- mapply(function(s, Rs){
+  lapply(s, function(ab){
+    apply(ab, 2, function(y){
+      step(lm(y ~., Rs), direction = 'both', trace = 0)
+    })
+  })
+}, s = CWM.ses, Rs = R, SIMPLIFY = F)
+CWM.ses.res <- lapply(CWM.ses.lm, function(s){
+  lapply(s, function(ab){
+    lapply(ab, function(x){
+      summary(x)$coefficients
+    })
+  })
+})
+CWM.ses.res <- lapply(CWM.ses.res, function(s){
+  lapply(s, function(ab){
+    lapply(ab, data.frame)
+  })
+})
+CWM.ses.res <- lapply(CWM.ses.res, function(s){
+  lapply(s, function(ab){
+    lapply(ab, function(x){
+      x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   '))
+      return(x)
+    })
+  })
+})
+CWM.ses.res <- lapply(CWM.ses.res, function(s){
+  lapply(s, function(ab){
+    data.frame(data.table::rbindlist(lapply(ab, function(x){
+      data.frame(t(x[5]))
+    }), fill = T))
   })
@@
 var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
-ppval <- lapply(ppval, function(s){
+CWM.ses.res <- lapply(CWM.ses.res, function(s){
   lapply(s, function(ab){
@@
   ab <- ab[var]
+  ab <- apply(ab, 2, as.character)
+  row.names(ab) <- names(CWM.ses$C$abundance)
+  ab[is.na(ab)] <- ''
+  ab <- data.frame(ab)
   return(ab)
@@
 })
-ppval <- lapply(ppval, function(s){
-  lapply(s, function(x){
-    1 - x
-  })
-})
-rm(pFval, npFval)
-
-# Result tables
-reg <- paste('~', paste(names(R$C), collapse = ' + '))
-res <- mapply(function(CWM.lms, pvals, ppvals){
-  mapply(function(CWM.lmab, pvalab, ppvalab){
-    regression_table(CWM.lmab, reg, pvalab, ppvalab)
-  }, CWM.lmab = CWM.lms, pvalab = as.list(data.frame(pvals)), ppvalab = ppvals, SIMPLIFY = F)
-}, CWM.lms = CWM.lm, pvals = pval, ppvals = ppval, SIMPLIFY = F)
-res <- lapply(res, function(s){
-  mapply(function(ab, name){
+CWM.ses.res <- mapply(function(s, CWM.ses.lms){
+  mapply(function(ab, CWM.ses.lmab){
+    ab$`$R^2$` <- round(unlist(lapply(CWM.ses.lmab, function(x){
+      summary(x)$r.squared
+    })),3)
+    return(ab)
+  }, ab = s, CWM.ses.lmab = CWM.ses.lms, SIMPLIFY = F)
+}, s = CWM.ses.res, CWM.ses.lms = CWM.ses.lm, SIMPLIFY = F)
+CWM.ses.res <- lapply(CWM.ses.res, function(s){
+  lapply(s, function(ab){
+    ab <- cbind(row.names(ab), ab)
+    names(ab)[1] <- 'Trait'
+    return(ab)
+  })
+})
+# CWM.ses.res <- lapply(CWM.ses.res, function(s){
+#   lapply(s, function(ab){
+#     sel <- which(unlist(lapply(as.list(data.frame(t(ab))), function(t){
+#   any(grepl('*', t, fixed = T))
+# })))
+#   return(ab[sel,])
+#   })
+# })
+CWM.ses.res <- mapply(function(s, name){
+  lapply(s, function(ab){
       ab <- cbind(c(name, rep(' ', length(ab[,1])-1)), ab)
-    names(ab)[1] <- 'Type'
+      names(ab)[1] <- 'T'
       return(ab)
-  }, ab = s, name = as.list(c('AB', 'PA')), SIMPLIFY = F)
-})
-res <- lapply(res, function(s){
+  })
+}, s = CWM.ses.res, name = as.list(c('AB', 'PA')), SIMPLIFY = F)
+CWM.ses.res <- lapply(CWM.ses.res, function(s){
   do.call('rbind', s)
 })
-res <- mapply(function(s, name){
+CWM.ses.res <- mapply(function(s, name){
   s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
-  names(s)[1] <- 'Strata'
+  names(s)[1] <- 'S'
   return(s)
-}, s = res, name = as.list(c('C', 'U')), SIMPLIFY = F)
-res <- do.call('rbind', res)
-kable(bold(res), row.names = F)
-rm(CWM.lm, nCWM.lm, pval, ppval, reg, res) ; invisible(gc())
+}, s = CWM.ses.res, name = as.list(c('C', 'U')), SIMPLIFY = F)
+CWM.ses.res <- do.call('rbind', CWM.ses.res)
+kable(bold(CWM.ses.res), row.names = F)
+rm(nCWM, CWM.ses, CWM.ses.lm, CWM.ses.res)
+invisible(gc())
 ```
@@
 ```{r CWV}
+# Removing communities with only one species
+for(s in names(L)){
+  sel <- which(rowSums(L[[s]]$`presence-absence`) < 2)
+  if(length(sel) > 1){
+    L[[s]] <- lapply(L[[s]], function(ab){
+      return(ab <- ab[-sel,])
+    })
+    R[[s]] <- R[[s]][-sel,]
+  }
+}
+
 # Observed CWVs
@@
 invisible(gc())
-## Non null CWVs
 nCWV <- lapply(nCWV, function(s){
@@
     lapply(s, data.frame)
-})
-CWV.pval <- mapply(function(CWVs, nCWVs){
-  data.frame(
-  mapply(function(x,y){
-    mapply(function(p,q){
-      rank(c(p,q))[1]/(n+1)
-    }, p = as.list(x), q = as.list(data.frame(t(y))))
-  }, x = CWVs, y = nCWVs)
-)
-}, CWVs = CWV, nCWVs = nCWV, SIMPLIFY = F)
-CWV.pval.res <- lapply(CWV.pval, function(s){
-  data.frame(
-    inf = apply(apply(s, 2, p.adjust, method = 'fdr'), 2, function(x){
-      length(which(x < 0.025))/length(x)
-    }), 
-    sup = apply(apply(1 - s, 2, p.adjust, method = 'fdr'), 2, function(x){
-      length(which(x < 0.025))/length(x)
-    }))
 })
@@
 })
-CWV.ses.res <- lapply(CWV.ses.res, function(s){
-  sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
-  any(grepl('*', t, fixed = T))
-})))
-  return(s[sel,])
-})
+# CWV.ses.res <- lapply(CWV.ses.res, function(s){
+#   sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
+#   any(grepl('*', t, fixed = T))
+# })))
+#   return(s[sel,])
+# })
 CWV.ses.res <- mapply(function(s, name){
+  if(dim(s)[1] > 0){
     s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
@@
     return(s) 
+  }
 }, s = CWV.ses.res, name = as.list(c('C', 'U')), SIMPLIFY = F)
 CWV.ses.res <- do.call('rbind', CWV.ses.res)
-rm(CWV.pval, CWV.pval.res, CWV.ses.lm, nCWV)
+rm(CWV.pval.res, CWV.ses.lm, nCWV)
 ```
@@
 ```{r CVNND}
+# Removing communities with only two species
+for(s in names(L)){
+  sel <- which(rowSums(L[[s]]$`presence-absence`) < 3)
+  if(length(sel) > 1){
+    L[[s]] <- lapply(L[[s]], function(ab){
+      return(ab <- ab[-sel,])
+    })
+    R[[s]] <- R[[s]][-sel,]
+  }
+}
+
 # Observed CVNND
@@
 }, Ls = L, Qs = Q, SIMPLIFY = F)
+## NaN CVNND$U
+sel <- which(is.na(CVNND$U[1,]))
+CVNND$U <- CVNND$U[,-sel]
+R$U <- R$U[-sel,]
 
@@
 invisible(gc)
-
-## Non null CVNNDs
+## NaN CVNND$U
+nCVNND$U <- lapply(nCVNND$U, function(x){
+  x <- x[,-sel]
+})
+
 nCVNND <- lapply(nCVNND, function(s){
   lapply(s, function(x){
-    as.list(data.frame(t(x)))
-  })
-})
-nCVNND <- lapply(nCVNND, function(s){
-  lapply(s, function(x){
-    lapply(x, function(y){
-      names(y) <- colnames(CVNND$C)
-      return(y)
-    })
+    x <- as.list(data.frame(t(x)))
   })
@@
 nCVNND <- lapply(nCVNND, function(s){
-  lapply(s, function(x){
-    do.call('cbind', x)
-  })
-})
-nCVNND <- lapply(nCVNND, function(s){
-  lapply(s, data.frame)
-})
-CVNND.pval <- mapply(function(CVNNDs, nCVNNDs){
-  data.frame(
-    mapply(function(x,y){
-      mapply(function(p,q){
-        rank(c(p,q))[1]/(n+1)
-      }, p = as.list(x), q = as.list(data.frame(t(y))))
-    }, x = as.list(data.frame(t(CVNNDs))), y = nCVNNDs)
-  )
-}, CVNNDs = CVNND, nCVNNDs = nCVNND, SIMPLIFY = F)
-CVNND.pval <- lapply(CVNND.pval, function(s){
-  row.names(s) <- colnames(CVNND)
-  apply(s, 2, p.adjust, method = 'fdr')
-  return(s)
-})
-CVNND.pval.res <- lapply(CVNND.pval, function(s){data.frame(
-  inf = apply(s, 2, function(x){
-    length(which(x < 0.05))/length(x)
-  })
-)})
-CVNND.pval.res <- do.call('cbind', CVNND.pval.res)
-names(CVNND.pval.res) <- names(CVNND.pval)
+  lapply(s, function(t){
+      t <- data.frame(lapply(t, cbind))
+  })
+})
 
@@
   apply(s, 2, function(y){
-    lm(y ~., Rs)
+    step(lm(y ~., Rs), direction = 'both', trace = 0)
   })
 }, s = CVNND.ses, Rs = R, SIMPLIFY = F)
-CVNND.ses.lm <- mapply(function(s, Rs){
-  lapply(s, function(x){
-    step(x, direction = 'both', trace = 0)
-  })
-}, s= CVNND.ses.lm, Rs = R, SIMPLIFY = F)
 CVNND.ses.res <- lapply(CVNND.ses.lm, function(s){
@@
 })
-CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
-  sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
-  any(grepl('*', t, fixed = T))
-})))
-  return(s[sel,])
-})
+# CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
+#   sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
+#   any(grepl('*', t, fixed = T))
+# })))
+#   return(s[sel,])
+# })
 CVNND.ses.res <- mapply(function(s, name){
+  if(dim(s)[1] > 0){
     s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
@@
     return(s)
+  }
 }, s = CVNND.ses.res, name = as.list(c('C', 'U')), SIMPLIFY = F)
@@
 ```{r CWV & CVNND table}
-print(round(CVNND.pval.res, 3))
 res <- list(CWV = CWV.ses.res, CVNND = CVNND.ses.res)

### 2016-08-11 R/genSpecies.R

@@ -19,24 +19,26 @@ genSpecies <- function(){
   species$SpCode <- tolower(species$SpCode)
   species[duplicated(species$SpCode),1] <- species[duplicated(species$SpCode),8]
   rownames(species) <- species$SpCode
-  # Expliciting strata
-  species$ESI <- 0; species[grep(x=species$PotStrata,pattern="\\bESI\\b"),]$ESI <- 1
-  species$ESII <- 0; species[grep(x=species$PotStrata,pattern=", II"),]$ESII <- 1; species[grep(x=species$PotStrata,pattern="\\bESII\\b"),]$ESII <- 1
-  species$ESIII <- 0; species[grep(x=species$PotStrata,pattern=", III"),]$ESIII <- 1; species[grep(x=species$PotStrata,pattern="ESIII"),]$ESIII <- 1
-  species$ESIV <- 0; species[grep(x=species$PotStrata,pattern=", IV"),]$ESIV <- 1; species[grep(x=species$PotStrata,pattern="ESIV"),]$ESIV <- 1
-  # Lot of species in different strata
-  # Ayyappan suggest to keep the highest potential strata
-  species[species$ESI==1 & species$ESII==1,]$ESII <- 0
-  species[species$ESII==1 & species$ESIII==1,]$ESIII <- 0
-  species[species$ESIII==1 & species$ESIV==1,]$ESIV <- 0
-  # Strata as factor
-  species$Strata <- NA
-  species$Strata[species$ESI==1] <- 'ESI'
-  species$Strata[species$ESII==1] <- 'ESII'
-  species$Strata[species$ESIII==1] <- 'ESIII'
-  species$Strata[species$ESIV==1] <- 'ESIV'
-  species$Strata <- as.factor(species$Strata)
-  species <- species[-which(names(species) %in% levels(species$Strata))]
+  # Opening strata
+  path <- system.file('extdata', 'UPSP_Species_list.csv', package = 'Uppangala')
+  strata <- read.table(path,sep=",",dec=".", stringsAsFactors =F, h = T) # Pelissier et al, 2011
+  strata <- strata[c(1:100),]
+  row.names(strata) <- strata$SpCode
+  # Linking strata
+  setdiff(row.names(species), strata$SpCode)
+  setdiff(strata$SpCode, row.names(species))
+  strata['agja',]$SpCode <- 'agla'
[…]

### 2016-08-12 inst/doc/draft.Rmd
v0.1.9034 Corrected strata beta results

@@
 Understanding the determinants of species occurrence and co-occurrence is a long standing issue in ecology [@diamond_ecology_1975].
-Tropical rainforests have especially fascinated standing issue in ecology [@connell_diversity_1978].
-Two concurrent theories have been proposed to explain how species coexistence can be maintained in ecological communities.
+Tropical rainforests have especially fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
+Many theories have been proposed to explain how species coexistence can be maintained in ecological communities .
 Niche theory states that the niche differences of species determine wether they can establish and coexist in communities [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that species coexistence is possible even when species are functionaly equivalent, under the sole influence of stochasticity in life, death, reproduction and dispersal [@hubbell_unified_2001].
@@
 
-### Taxonomic, functional and phylogenetic variation among communities
-
-A phylogenetic tree of our 89 species was built with `S.PhyloMaker` function proposed by @qian_updated_2016-1.
-We assessed phylogenetic signal of each functional using the Moran's I statistic.
-Positive Moran's I value indicated convergence of trait values within the phylogeny whereas negative Moran's I will showed divergence.
-
-To investigate taxonomic, functional, and phylogenetic variation among communities we calculated $I_{st}$, for taxonomic composition, $U_{st}$ and $\tau _{st}$, for functional composition in abundance and presence-absence respectively, and $\beta _{st}$ and $\pi _{st}$, for phylogenetic composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
+### Taxonomic and functional variation among communities
+
+To investigate taxonomic and functional variation among communities we calculated $I_{st}$, for taxonomic composition and $U_{st}$ and $\tau _{st}$, for functional composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
 $I_{st}$ measures species turnover accounting for local abundance.
-$U_{st}$ and $\beta _{st}$ integrate only phylogenetic and functional turnover, respectively, weighting species by their local abundances; whereas $\tau _{st}$ and $\pi _{st}$ express phylogenetic and functional turnover, respectively, without accounting for species local abundances.
+$U_{st}$ integrate only functional turnover, weighting species by their local abundances; whereas $\tau _{st}$ express functional turnover, without accounting for species local abundances.
 
@@
 $U st$ and $\tau st$ were tested by the nul model NM1 which randomizes traits among species.
-Finally $\beta st$ and $\pi st$ were tested against the nul model *"1s"* suggested by @hardy_testing_2008 which randomizes phylogenetic tree tips.
 
@@
 
-Table 3: Summary of linear models of community-level trait values.
-_The results are shown for either CWM values weighted either by species relative abundance (AB) or in presence-absence (PA) after stepwise selection of the predictors including slope, curvature, wetness, southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Global and partial F-statistics of final models were compared to statistics of 999 null linear regressions built with null model NM1, which randomizes traits among species.
+Table 3: Summary of linear models of traits standardized effect size (SES) of CWMs.
+_Linear regression results for each trait SES of CWMs, for canopy (C) and understorey (U) guilds, in abundance (AB) and presence-absence (PA), after stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
 See table 1 for abbreviations.
@@
 Traits convergence/divergence was measured with community weighted variances (CWV).
-Overall, observed CWVs did not differ from null expectations for all traits.
 We could found significant relationships between standardized effect size (SES) of CWVs and environmental variables (see table 4).
-Convergence of LES and LDMC and divergence of LA were found under south-west orientation.
-LA and LDMC values also diverged in communities with increasing curvature.
-Observed CVNNDs did not reveal niche differentiation among communities for all traits.
-But we also found significant relationships of their standardized effect size (SES) of CVNNDs with environmental variables (see table 4).
-niche differentiation of LA increased in lower curvature and wetness contexts, while differentiation of WD decreased in southwest orientation.
+Convergence of WES and LDMC in the canopy guild  and convergence of LA in the undersotrey guild were found under concave areas.
+LDMC of the canopy guild converged also with canopy height.
+LES of understorey guild converged in slopes and SLA of understorey guild diverged with increasing BA.
+We also found significant relationships of their standardized effect size (SES) of CVNNDs with environmental variables (see table 4).
+Niche differentiation of WES and LDMC in canopy guild increased in lower basal area context.
+Differentiation of canopy guild LA increased in convex and high canopy contexts, whereas LT of the canopy guild differentiated in concave context.
+Niche differentiation of LES in understorey guild increased in concave areas, while differentiation of LT of the understorey guild increased in wet context with a high basal area.
 
 Table 4: Summary of linear models of traits standardized effect size (SES) of CWVs and CVNNDs.
-_Linear regression results for each trait SES of CWVs and CVNNDs after stepwise selection of the predictors oncluding slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
+_Linear regression results for each trait SES of CWVs and CVNNDs, for canopy (C) and understorey (U) guilds, after stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
 See table 1 for abbreviations.
@@
 kable(bold(res), row.names = F)
-rm(CWV.ses.res, CVNND.ses.res, CVNND.pval.res, res)
+rm(CWV.ses.res, CVNND.ses.res, CVNND.pval.res, sel, res)
 ```
 
-## Taxonomic, functional and phylogenetic varitions among communities
-
-Taxonomic variation among communities was correlated significantly to environment variation (Mantel test, $r_{I_{st},E} = 0.23$ with $p-value < 0.001$), but not with remaining spatial dependency in community composition (Mantel test of space knowing environment, $r_{I_{st},S|E} = 0.01$ with $p-value > 0.1$).
-
-Overall, the global functional variation among communities was neither correlated to environment nor to spactial distance knowing the environment effect (see table 5).
-However, the variation of WD, in abundance as well as in presence absence, was positively correlated to environmental variation and is negatively correlated in presence absence to the spatial structure knowing environment spatial effect.
-
-While LDMC did not phylogenetic structure (Moran's I of 0 with a $p-value > 0.1$), LA and LT were phylogenetically clustered (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA were weakly clustered (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
-Phylogenetic variation among communities was neither correlated to environmental variation (Mantel test, $r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor to remaining spatial variation in partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).
-
-```{r Ist}
+## Taxonomic and functional varitions among communities
+
+Taxonomic variation among communities was correlated significantly to environment variation for both canopy and understorey guilds (Mantel test for canopy and understorey guild, $r_{I_{st},E,C} = 0.22$ and $r_{I_{st},E,U} = 0.23$, respectively, with $p-value < 0.01$), but not with remaining spatial dependency in community composition for canopy guild (Mantel test of space knowing environment, $r_{I_{st},S|E,C} = 0.01$ with $p-value > 0.1$), whereas understorey guild taaxonomic variation was correlated to remaining spatial dependency in community composition (Mantel test of space knowing environment, $r_{I_{st},S|E,U} = 0.11$ with $p-value < 0.01$).
+
+```{r Ist, eval=FALSE, include=FALSE}
 library(spacodiR)
 
-# R,L,Q wihtout strata
-R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
-row.names(R) <- com@data$id
-L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-L[is.na(L)] <- 0
-Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')]
-row.names(Q) <- PFT_sp$SpCode
-sel <- intersect(row.names(Q), names(L))
-Q <- Q[sel,]
-L <- L[sel]
-R <- R[row.names(L),]
-L <- list(abundance = L, `presence-absence` = L)
-L$`presence-absence`[L$`presence-absence` > 1]  <- 1
-rm(com, Trees)
-invisible(gc())
-
 # Environmental matrices
-XY <- data.frame(as(quadrats(), 'SpatialPoints')@coords)
+XY <- data.frame(as(quadrats(c(30,30)), 'SpatialPoints')@coords)
 row.names(XY) <- paste0('C',row.names(XY))
-XY <- XY[row.names(L$abundance),]
-Ms <- dist(XY)
+XY <- lapply(L, function(s){
+  XY[row.names(s$abundance),]
+})
+Ms <- lapply(XY, function(s){
+  dist(s)
+})
 XY <- pca.env$scores[,1:2]
 row.names(XY) <- paste0('C',row.names(XY))
-XY <- XY[row.names(L$abundance),]
-Me <- dist(XY)
+XY <- lapply(L, function(s){
+  XY[row.names(s$abundance),]
+})
+Me <- lapply(XY, function(s){
+  dist(s)
+})
 rm(XY)
@@
 # Observed Ist
-spaflo <- spacodi.calc(sp.plot = t(L$abundance), pairwise = T)$'pairwise.Ist'
+spaflo <- lapply(L, function(s){
+  spacodi.calc(sp.plot = t(s$abundance), pairwise = T)$'pairwise.Ist'
+})
 ## Mantel values
-spaflo.mantel <- list(`E` = vegan::mantel(spaflo, Me, permutations = 0)$statistic, 
-                      `S | E`= vegan::mantel.partial(spaflo, Ms, Me, permutations = 0)$statistic)
+spaflo.mantel <- mapply(function(s, Mes, Mss){
+  list(
+    `E` = vegan::mantel(s, Mes, permutations = 0)$statistic, 
+    `S | E`= vegan::mantel.partial(s, Mss, Mes, permutations = 0)$statistic
+  )
+}, s = spaflo, Mes = Me, Mss = Ms, SIMPLIFY = F)
 
 # Simulated mantel
-nL <- vegan::permatfull(L$abundance, fixedmar = 'both', mtype = 'count', times = n, shuffle = 'both')
-cl <- makeCluster(cores)
-clusterExport(cl, list('nL', 'spacodi.calc', 'Me', 'Ms'))
-nspaflo <- lapply(nL$perm, function(l){spacodi.calc(sp.plot = t(l), pairwise = T)$'pairwise.Ist'})
-nspaflo.mantel <- lapply(nspaflo, function(ns){
-  list(`E` = vegan::mantel(ns, Me, permutations = 0)$statistic, 
-       `S | E`= vegan::mantel.partial(ns, Ms, Me, permutations = 0)$statistic)})
-stopCluster(cl)
-rm(cl, nL) ; invisible(gc())
+nL <- lapply(L, function(s){
+  vegan::permatfull(s$abundance, fixedmar = 'both', mtype = 'count', times = n, shuffle = 'both')
+})
+nspaflo <- lapply(nL, function(s){
+  lapply(s$perm, function(l){
+    spacodi.calc(sp.plot = t(l), pairwise = T)$'pairwise.Ist'
+  })
+})
+nspaflo.mantel <- mapply(function(s, Mes, Mss){
+  lapply(s, function(x){
+    try(list(
+      `E` = vegan::mantel(x, Mes, permutations = 0)$statistic, 
+      `S | E`= vegan::mantel.partial(x, Mss, Mes, permutations = 0)$statistic
+    ))
+  })
+}, s = nspaflo, Mes = Me, Mss = Ms, SIMPLIFY = F)
+sel <- which(unlist(lapply(nspaflo.mantel$U, inherits, 'try-error')))
+nspaflo.mantel$U <- nspaflo.mantel$U[-sel]
 ## Non null mantels
-nspaflo.mantel <- lapply(nspaflo.mantel, `[`, names(nspaflo.mantel[[1]]))
-nspaflo.mantel <- apply(do.call(rbind, nspaflo.mantel), 2, as.list)
-nspaflo.mantel <- lapply(nspaflo.mantel, unlist)
+nspaflo.mantel <- lapply(nspaflo.mantel, function(s){
+  lapply(s, `[`, names(s[[1]]))
+  apply(do.call(rbind, s), 2, as.list)
+})
+nspaflo.mantel <- lapply(nspaflo.mantel, function(s){
+  lapply(s, unlist)
+})
 spaflo.res <- mapply(function(s, ns){
-    paste(format(s, scientific = T, digits = 3),
-          Uppangala::stars(min(1-rank(c(s,ns))[1]/(n+1),rank(c(s,ns))[1]/(n+1)), ns = '   '))
+  mapply(function(x, nx){
+    paste(format(x, scientific = T, digits = 3),
+          Uppangala::stars(min(1-rank(c(x,nx))[1]/(n+1),rank(c(x,nx))[1]/(n+1)), ns = '   '))
+  }, x =s, nx = ns)
 }, s = spaflo.mantel, ns = nspaflo.mantel)
-spaflo.res <- data.frame(spaflo.res)
-names(spaflo.res) <- c('$I_{st}$')
 
 # Result
-# kable(bold(spaflo.res))
-rm(mtest, spaflo.mantel, nspaflo, nspaflo.mantel, spaflo.res)
+kable(bold(spaflo.res))
+rm(spaflo.mantel, nspaflo, nspaflo.mantel, spaflo.res)
 ```
 
+Most of trait functional variation among communities was neither correlated to environment nor to spactial distance knowing the environment effect (see table 5).
+However, the variation of WD, in abundance as well as in presence absence, was positively correlated to environmental variation and is negatively correlated in presence absence to the spatial structure knowing environment spatial effect.
+
 Table 5: Summary of Mantel tests between functional composition, environment and spatial distances between communities.
-_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to measure functional composition variation among communities.
+_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to measure functional composition variation among communities, for canopy (C) and understorey (U) guilds.
 Observed values were compared to null values obtained with model NM1 randomizing traits among species (permuting raw in `Trait*Species` matrix).
@@
 ```{r Ust & TAUst}
-# nQ without strata
-n = 999
-nQ <- rep(list(Q), n)
-nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
-nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
-
 # Observed Hardys
-spafun <- lapply(as.list(1:length(Q)), function(i){
-  if (length(which(is.na(Q[i]))) > 0) {
-    sel <- intersect(names(L$abundance), row.names(Q[-which(is.na(Q[i])),]))
-    return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(Q[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
+spafun <- mapply(function(s, Ls){
+  lapply(as.list(1:length(s)), function(i){
+    if (length(which(is.na(s[i]))) > 0) {
+      sel <- intersect(names(Ls$abundance), row.names(s[-which(is.na(s[i])),]))
+      return(spacodi.calc(sp.plot = t(Ls$abundance[sel]), sp.traits = data.frame(s[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
     } else {
-    return(spacodi.calc(sp.plot = t(L$abundance), sp.traits = Q, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
+      return(spacodi.calc(sp.plot = t(Ls$abundance), sp.traits = s, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
     }
   })
-names(spafun) <- names(Q)
+}, s = Q, Ls = L, SIMPLIFY = F)
+spafun <- lapply(spafun, function(s){
+  names(s) <- names(Q$C)
+  return(s)
+})
 ## Mantel values
-spafun.mantel <- lapply(spafun, function(t){list(`E` = lapply(t, function(M){vegan::mantel(M, Me, permutations = 0)$statistic}), `S | E` = lapply(t, function(M){vegan::mantel.partial(M, Ms, Me, permutations = 0)$statistic}))})
+spafun.mantel <- mapply(function(s, Mes, Mss){
+  lapply(s, function(t){
+    list(
+      `E` = lapply(t, function(M){
+        vegan::mantel(M, Mes, permutations = 0)$statistic
+      }), 
+      `S | E` = lapply(t, function(M){
+        vegan::mantel.partial(M, Mss, Mes, permutations = 0)$statistic
+      })
+    )
+  })
+}, s = spafun, Mes = Me, Mss = Ms, SIMPLIFY = F)
 
@@
 clusterExport(cl, list('nQ', 'L', 'Me', 'Ms', 'spacodi.calc'))
-nspafun.mantel <- clusterApply(cl, nQ, function(q){
+nspafun.mantel <- clusterMap(cl, function(nQs, Ls, Mes, Mss){ 
+  lapply(nQs, function(q){
     s <- lapply(as.list(1:length(q)), function(i){
       if (length(which(is.na(q[i]))) > 0) {
-      sel <- intersect(names(L$abundance), row.names(q[-which(is.na(q[i])),]))
-      return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(q[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
+        sel <- intersect(names(Ls$abundance), row.names(q[-which(is.na(q[i])),]))
+        return(spacodi.calc(sp.plot = t(Ls$abundance[sel]), sp.traits = data.frame(q[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
       } else {
-      return(spacodi.calc(sp.plot = t(L$abundance), sp.traits = q, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
+        return(spacodi.calc(sp.plot = t(Ls$abundance), sp.traits = q, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
       }
@@
         lapply(t, function(M){
-        try(vegan::mantel(M, Me, permutations = 0)$statistic)
+          try(vegan::mantel(M, Mes, permutations = 0)$statistic)
         })
@@
         lapply(t, function(M){
-        try(vegan::mantel.partial(M, Ms, Me, permutations = 0)$statistic)
+          try(vegan::mantel.partial(M, Mss, Mes, permutations = 0)$statistic)
         })
@@
   })
+}, nQs = nQ, Ls = L, Mes = Me, Mss = Ms, SIMPLIFY = F)
 stopCluster(cl)
@@
 ## Non null mantels
-nspafun.mantel <- lapply(nspafun.mantel, `[`, names(nspafun.mantel[[1]]))
-nspafun.mantel <- apply(do.call(rbind, nspafun.mantel), 2, as.list)
-nspafun.mantel <- lapply(nspafun.mantel, function(s){lapply(s, `[`, names(s[[1]]))})
-nspafun.mantel <- lapply(nspafun.mantel, function(s){apply(do.call(rbind, s), 2, as.list)})
-nspafun.mantel <- lapply(nspafun.mantel, function(s){lapply(s, function(m){lapply(m, `[`, names(m[[1]]))})})
-nspafun.mantel <- lapply(nspafun.mantel, function(s){lapply(s, function(m){apply(do.call(rbind, m), 2, as.list)})})
-nspafun.mantel <- lapply(nspafun.mantel, function(s){lapply(s, function(x){lapply(x, function(x){as.numeric(unlist(x))})})})
-spafun.res <- data.frame(mapply(function(x,y){
-  mapply(function(r,s){
-    mapply(function(p,q){
-      paste(round(p, 3), Uppangala::stars(min(rank(c(p,q))[1]/(n+1), 1-rank(c(p,q))[1]/(n+1)), ns = '   '))
-    }, p = r, q = s, SIMPLIFY = F)
-  }, r = x, s= y)
-}, x = spafun.mantel, y = nspafun.mantel))
-spafun.res <- cbind(data.frame('Mantel statistic' = c('E', ' ', 'S|E', ' '), 'Metric' = rep(c('$U_{st}$', '$\\tau _{st}$'), 2)), spafun.res)
+nspafun.mantel <- lapply(nspafun.mantel, function(s){
+  lapply(s, `[`, names(s[[1]]))
+  apply(do.call(rbind, s), 2, as.list)
+})
+nspafun.mantel <- lapply(nspafun.mantel, function(s){
+  lapply(s, function(m){
+    lapply(m, `[`, names(m[[1]]))
+    apply(do.call(rbind, m), 2, as.list)
+  })
+})
+nspafun.mantel <- lapply(nspafun.mantel, function(s){
+  lapply(s, function(m){
+    lapply(m, function(u){
+      lapply(u, `[`, names(u[[1]]))
+      apply(do.call(rbind, u), 2, as.list)
+    })
+  })
+})
+nspafun.mantel <- lapply(nspafun.mantel, function(s){
+  lapply(s, function(m){
+    lapply(m, function(u){
+      lapply(u, unlist)
+    })
+  })
+})
+
+spafun.res <- mapply(function(s, ns){
+  mapply(function(t, nt){
+    mapply(function(m, nm){
+      mapply(function(u, nu){
+        paste(round(u, 3), Uppangala::stars(min(rank(c(u,nu))[1]/(n+1), 1-rank(c(u,nu))[1]/(n+1)), ns = '   '))
+      }, u = m, nu = nm)
+    }, m = t, nm = nt)
+  }, t = s, nt = ns, SIMPLIFY = F)
+}, s = spafun.mantel, ns = nspafun.mantel, SIMPLIFY = F)
+trait <- names(spafun.res$C)
+spafun.res <- lapply(spafun.res, function(s){
+  mapply(function(t, tr){
+    t <- data.frame(t)
+    t <- cbind(c(tr, ' '), c('$U_{st}$', '$\\tau _{st}$'),  t)
+    names(t) <- c('T', 'M', 'E', 'S|E')
+    return(t)
+  }, t = s, tr = trait, SIMPLIFY = F)
+})
+spafun.res <-lapply(spafun.res, function(s){
+  s <- do.call('rbind', s)
+})
+spafun.res <- mapply(function(s, str){
+  s <- cbind(c(str, rep(' ', dim(s)[1]-1)), s)
+  names(s)[1] <- 'S'
+  return(s)
+}, s = spafun.res, str = c('C', 'U'), SIMPLIFY = F)
+spafun.res <- do.call('rbind', spafun.res)
 
 # Result
-kable(bold(spafun.res))
+kable(bold(spafun.res), row.names = F)
 rm(spafun.mantel, nspafun.mantel, spafun.res)
-```
-
-```{r PhyloMaker, include=FALSE}
-library(phylosignal)
-library(phytools)
-library(phylobase)
-path <- system.file('phylogeny', package = 'Uppangala')
-source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
-phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
-nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
-sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), c('LatinName', 'Genus', 'Family')]
-names(sp) <- c('species', 'genus', 'family')
-results <- invisible(S.PhyloMaker(sp, phylo, nodes)) # Humboldtia_brunonis unmatched
-tree <- results$Scenario.1
-rm(phylo, results, modes, sp, S.PhyloMaker, path)
-```
-
-```{r Phylosignal}
-data <- PFT_sp
-row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
-tips <- row.names(data)[which(is.na(data$WD))]
-phy <- drop.tip(tree, tips)
-data <- data[phy$tip.label,c(2:5,7)]
-p4d <- phylo4d(phy,data)
-phyloS <- phyloSignal(p4d, methods = 'I') 
-phyloS <- data.frame(paste(round(phyloS$stat[,1], 3), 
-                           unlist(lapply(as.list(phyloS$pvalue[,1]), stars, ns = ' '))),
-                     row.names = row.names(phyloS$stat))
-names(phyloS) <- 'Moran\'s I'
-# kable(bold(phyloS))
-rm(data, tips, p4d, phyloS, phy)
-```
-
-```{r BETAst $ PIst}
-# Observed Hardys
-phy <- tree
-phy$tip.label <- Species$SpCode[match(gsub('_', ' ',phy$tip.label), Species$LatinName)]
-tips <- c(setdiff(phy$tip.label, names(L$abundance)), setdiff(names(L$abundance), phy$tip.label))
-phy <- drop.tip(phy, tips)
-phy$node.label <- NULL
-spaphy <- spacodi.calc(sp.plot = t(L$abundance), phy = phy, pairwise = T)[c('pairwise.Bst', 'pairwise.PIst')]
-## Mantel values
-spaphy.mantel <- list(`E` = lapply(spaphy, function(m){vegan::mantel(m, Me, permutations = 0)$statistic}), 
-                      `S | E`= lapply(spaphy, function(m){vegan::mantel.partial(m, Ms, Me, permutations = 0)$statistic}))
-
-# Simulated mantel
-nphy <- rep(list(phy), 999)
-cl <- makeCluster(cores)
-clusterExport(cl, list('nphy', 'L', 'spacodi.calc', 'Ntip', 'resamp.phy', 'branching.times'))
-nphy <- clusterApply(cl, nphy, resamp.phy)
-stopCluster(cl)
-nspaphy <- lapply(nphy, function(p){spacodi.calc(sp.plot = t(L$abundance), phy = p, pairwise = T)[c('pairwise.Bst', 'pairwise.PIst')]})
-rm(cl, nphy) ; invisible(gc())
-nspaphy.mantel <- list(`E` =lapply(nspaphy, function(x){lapply(x, function(M){vegan::mantel(M, Me, permutations = 0)$statistic})}), `S | E` = lapply(nspaphy, function(x){lapply(x, function(M){vegan::mantel.partial(M, Ms, Me, permutations = 0)$statistic})}))
-## Non null mantels
-nspaphy.mantel <- lapply(nspaphy.mantel, function(s){lapply(s, `[`, names(s[[1]]))})
-nspaphy.mantel <- lapply(nspaphy.mantel, function(s){apply(do.call(rbind, s), 2, as.list)})
-nspaphy.mantel <- lapply(nspaphy.mantel, function(s){lapply(s, unlist)})
-spaphy.res <- mapply(function(s, ns){
-  mapply(function(x,y){
-    paste(format(x, scientific = T, digits = 3), 
-          Uppangala::stars(min(1-rank(c(x,y))[1]/(n+1),rank(c(x,y))[1]/(n+1)), ns = '   '))
-  }, x = s, y = ns)
-}, s = spaphy.mantel, ns = nspaphy.mantel)
-row.names(spaphy.res) <- c('$\\beta_{st}$', '$\\pi_{st}$')
-
-# Result
-# kable(bold(spaphy.res))
-rm(mtest, spaphy.mantel, nspaphy, nspaphy.mantel, spaphy.res)
 ```
@@
 
+## S2: Phylogenetic varitions among communities
+
+A phylogenetic tree of our 89 species was built with `S.PhyloMaker` function proposed by @qian_updated_2016-1.
+We assessed phylogenetic signal of each functional using the Moran's I statistic.
+Positive Moran's I value indicated convergence of trait values within the phylogeny whereas negative Moran's I will showed divergence.
+
+To investigate phylogenetic variation among communities we calculated $\beta _{st}$ and $\pi _{st}$, for phylogenetic composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
+$\beta _{st}$ integrate only phylogenetic turnover, weighting species by their local abundances; whereas $\pi _{st}$ express phylogenetic turnover, without accounting for species local abundances.
+
+Variation of composition between community pairs with those metrics was translated as a taxonomic, functional and phylogenetic distance, and compared to environmental and spatial distances among community pairs.
+Environmental distances were calculated as geometric distances between communities on the plan composed by the two first axes of environmental variables PCA (see supplementary material S1).
+Correlations with environment and space were assessed based on two Mantel tests successively with taxonomic, functional, and phylogenetic community compositions.
+A firt Mantel statistic assessed the correlation between each metric distance matrix and the environmental distance matrix (E).
+The second partial Mantel statistic assessed the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S), to represent remaining spatial dependency in community composition once the influence of environmental variation is taken into account.
+
+We used null models to test if observed correlation with Mantel statistic differed significantly from expectations under null hypothesis.
+$\beta st$ and $\pi st$ were tested against the nul model *"1s"* suggested by @hardy_testing_2008 which randomizes phylogenetic tree tips.
+
+```{r PhyloMaker, include=FALSE}
+library(phylosignal)
+library(phytools)
+library(phylobase)
+path <- system.file('phylogeny', package = 'Uppangala')
+source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
+phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
+nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
+sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), c('LatinName', 'Genus', 'Family')]
+names(sp) <- c('species', 'genus', 'family')
+results <- invisible(S.PhyloMaker(sp, phylo, nodes)) # Humboldtia_brunonis unmatched
+tree <- results$Scenario.1
+rm(phylo, results, modes, sp, S.PhyloMaker, path)
+```
+
+```{r Phylosignal}
+data <- PFT_sp
+row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
+tips <- row.names(data)[which(is.na(data$WD))]
+phy <- drop.tip(tree, tips)
+data <- data[phy$tip.label,c(2:5,7)]
+p4d <- phylo4d(phy,data)
+phyloS <- phyloSignal(p4d, methods = 'I') 
+phyloS <- data.frame(paste(round(phyloS$stat[,1], 3), 
+                           unlist(lapply(as.list(phyloS$pvalue[,1]), stars, ns = ' '))),
+                     row.names = row.names(phyloS$stat))
+names(phyloS) <- 'Moran\'s I'
+# kable(bold(phyloS))
+rm(data, tips, p4d, phyloS, phy)
+```
+
+```{r BETAst $ PIst}
+# Observed Hardys
+phy <- tree
+phy$tip.label <- Species$SpCode[match(gsub('_', ' ',phy$tip.label), Species$LatinName)]
+tips <- lapply(L, function(s){
+  c(setdiff(phy$tip.label, names(s$abundance)), setdiff(names(s$abundance), phy$tip.label))
+})
+phy <- lapply(tips, function(s){
+  drop.tip(phy, s)
+})
+phy <- lapply(phy, function(s){
+  s$node.label <- NULL
+  return(s)
+})
+spaphy <- mapply(function(Ls, phys){
+  spacodi.calc(sp.plot = t(Ls$abundance), phy = phys, pairwise = T)[c('pairwise.Bst', 'pairwise.PIst')]
+}, Ls = L, phys = phy, SIMPLIFY = F)
+## Mantel values
+spaphy.mantel <- mapply(function(s, Mes, Mss){
+  list(
+    `E` = lapply(s, function(m){
+      vegan::mantel(m, Mes, permutations = 0)$statistic
+      }), 
+    `S | E`= lapply(s, function(m){
+      vegan::mantel.partial(m, Mss, Mes, permutations = 0)$statistic
+      })
+    )
+}, s = spaphy, Mes = Me, Mss = Ms, SIMPLIFY = F)
+
+# Simulated mantel
+nphy <- lapply(phy, function(s){
+  rep(list(s), 999)
+})
+cl <- makeCluster(cores)
+clusterExport(cl, list('nphy', 'L', 'spacodi.calc', 'Ntip', 'resamp.phy', 'branching.times'))
+nphy <- clusterApply(cl, nphy, function(s){
+  lapply(s, resamp.phy)
+})
+stopCluster(cl)
+nspaphy <- mapply(function(s, Ls){
+  lapply(s, function(p){
+  spacodi.calc(sp.plot = t(Ls$abundance), phy = p, pairwise = T)[c('pairwise.Bst', 'pairwise.PIst')]
+  })
+}, s = nphy, Ls = L, SIMPLIFY = F)
+rm(cl, nphy)
+invisible(gc())
+nspaphy.mantel <- mapply(function(s, Mes, Mss){
+  list(
+    `E` =lapply(s, function(x){
+      lapply(x, function(M){
+        vegan::mantel(M, Mes, permutations = 0)$statistic})}), 
+    `S | E` = lapply(s, function(x){
+      lapply(x, function(M){
+        vegan::mantel.partial(M, Mss, Mes, permutations = 0)$statistic
+      })
+    })
+  )
+}, s = nspaphy, Mes = Me, Mss = Ms, SIMPLIFY = F)
+## Non null mantels
+names(nspaphy.mantel) <- c('C', 'U')
+nspaphy.mantel <- lapply(nspaphy.mantel, function(s){
+  lapply(s, function(m){
+    lapply(m, `[`, names(m[[1]]))
+    apply(do.call(rbind, m), 2, as.list)
+  })
+})
+nspaphy.mantel <- lapply(nspaphy.mantel, function(s){
+  lapply(s, function(m){
+    lapply(m, unlist)
+  })
+})
+spaphy.res <- mapply(function(s, ns){
+  mapply(function(m, nm){
+    mapply(function(b, nb){
+      paste(format(b, scientific = T, digits = 3), 
+            Uppangala::stars(min(1-rank(c(b,nb))[1]/(n+1),rank(c(b,nb))[1]/(n+1)), ns = '   '))
+    }, b = m, nb = nm, SIMPLIFY = T)
+  }, m = s, nm = ns, SIMPLIFY = F)
+}, s = spaphy.mantel, ns = nspaphy.mantel, SIMPLIFY = F)
+spaphy.res <- mapply(function(s, str){
+  s <- data.frame(s)
+  s <- cbind(c(str, ' '), c('$U_{st}$', '$\\tau _{st}$'), s)
+  names(s) <- c('S', 'M', 'E', 'S|E')
+  return(s)
+}, s = spaphy.res, str = c('C', 'U'), SIMPLIFY = F)
+spaphy.res <- do.call('rbind', spaphy.res)
+
+# Result
+kable(bold(spaphy.res), row.names = F)
+rm(mtest, spaphy.mantel, nspaphy, nspaphy.mantel, spaphy.res)
+```
+
+While LDMC did not phylogenetic structure (Moran's I of 0 with a $p-value > 0.1$), LA and LT were phylogenetically clustered (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA were weakly clustered (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
+Phylogenetic variation among communities was neither correlated to environmental variation (Mantel test, $r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor to remaining spatial variation in partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).
+
 # References

### 2016-08-12 inst/doc/draft.Rmd
v0.1.9035 Internship end

@@
 Many theories have been proposed to explain how species coexistence can be maintained in ecological communities .
-Niche theory states that the niche differences of species determine wether they can establish and coexist in communities [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that species coexistence is possible even when species are functionaly equivalent, under the sole influence of stochasticity in life, death, reproduction and dispersal [@hubbell_unified_2001].
+Among them, two have been hotly debated and tested those last years.
+First niche theory states that the niche differences of species determine wether they can establish and coexist in communities [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that species coexistence is possible even when species are functionaly equivalent, under the sole influence of stochasticity in life, death, reproduction and dispersal [@hubbell_unified_2001].
 Both theory can provide realistic expectations of taxonomic diversity [@chave_comparing_2002], and some recent approaches have advocated investigating patterns of functional [@westoby_land-plant_2006; @mcgill_rebuilding_2006] and phylogenetic diversity [@mouquet_ecophylogenetics:_2012] to trace the signature of ecological processes related to specific biological attributes and adaptations.
 These approaches have been applied to resolve the conundrum of high tree diversity in tropical forests [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012], but studies on these ecosystems still remain scarce.
-We here investigate taxonomic, functional and phylogenetic diversity in a low elevation rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], where no study has addressed the relative influence of neutral and niche-based dynamics on local forest composition.
-
-The variation of functional traits among species reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007].
+We here investigate taxonomic, functional and phylogenetic diversity in a low elevation rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], where few studies have assessed the relative influence of neutral and niche-based processes on local forest composition.
+
+The variation in functional traits among species reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007].
 These variation thus reflect fundamental trade-offs in ecological strategies along environmental gradients [@mark_westoby_plant_2002].
 For instance, specific leaf area varies along a soil fertility-shade index and represent a trade-off between resource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011].
-Due to common ecological strategies, functional traits will covary along basic leaf [@wright_worldwide_2004] and wood [@chave_towards_2009] economic spectra.
-The leaf economic spectrum opposes productive ecological strategy to conservative persistent strategy.
-The wood economic spectrum distinguish species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, wiht high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength.
-@baraloto_decoupled_2010 highlighted decoupled wood and leaf strategies allowing independant variation of these strategies, but @reich_world-wide_2014 conversely advocated that the two spectra have to integrated in a comprehensive whole plant economic spectrum [@freschet_evidence_2010] opposing fast growing species to more conservative ones.
+Due to shared ecological strategies, functional traits will covary along basic leaf [@wright_worldwide_2004 but see @osnas_global_2013 and @lloyd_photosynthetically_2013] and wood [@chave_towards_2009] economics spectra.
+The leaf economics spectrum opposes productive ecological strategy to conservative persistent strategy.
+The wood economics spectrum distinguish species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, wiht high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength [@chave_towards_2009].
+@baraloto_decoupled_2010 highlighted decoupled wood and leaf strategies allowing independant variation of these strategies, but @reich_world-wide_2014 conversely advocated that the two spectra have to integrated in a comprehensive whole plant economics spectrum [@freschet_evidence_2010] opposing fast growing species to more conservative ones.
 Here we address how varying wood and leaf ecological strategies determine species community assembly in an Indian rainforest.
@@
 Local community dynamics can further be influenced by niche differences under the influence of competition [@macarthur_limiting_1967].
-In this perspective, @messier_trait_2016 have underlined that the significance of the leaf economic spectrum is basically scale-dependent, which can explain conflicting results across studies performed at different scales.
+In this perspective, @messier_trait_2016 have underlined that the significance of the leaf economics spectrum is basically scale-dependent, which can explain conflicting results across studies performed at different scales.
 In the same logic, a spatial hierarchy of ecological dynamics can entail a hierarchy of variation in functional traits from broad variation among species along large-scale environmental gradients, to intraspecific response to micro-environmental conditions at very fine scale [@albert_when_2011].
 Therefore a multiscale framework is needed to represent the hierarchy of ecological processes from regional to local scale [@lortie_rethinking_2004].
-Determining the extent of neutral and niche-based dynamics then implies (i) decomposing diversity patterns within and among communities [@ackerly_trait-based_2007], and (ii) designing appropriate statistical tests to disentangle the influence of local and larger-scale dynamics [@bernard-verdier_community_2012].
-We will decompose diversity patterns in a large rainforest plot into sources of variation within and between communities [@kraft_disentangling_2011], and investigate the fine-scale and plot-scale dynamics using appropriate null models.
-For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they question the existence of the leaf economic spectrum at community scale.
-
-Apart from scale-dependent dynamics in space whithin and among communities, ecological guilds can coexist and occupy distinct niches within communities, while the species within guilds are ecologically similar [@wilson_guilds_1999; @vergnon_niches_2009].
+Determining the extent of neutral and niche-based dynamics thus implies decomposing diversity patterns within and among communities [@ackerly_trait-based_2007], and designing appropriate statistical tests to disentangle the influence of local and larger-scale dynamics [@bernard-verdier_community_2012].
+In this paper, we will decompose diversity patterns in a large rainforest plot into sources of variation within and between communities [@kraft_disentangling_2011], and investigate the fine-scale and plot-scale dynamics using appropriate null models.
+For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they challenged the existence of the leaf economics spectrum at community scale.
+
+Apart from scale-dependent dynamics in space within and among communities, ecological guilds can coexist and occupy distinct niches within communities, while the species within guilds are ecologically similar [@wilson_guilds_1999; @vergnon_niches_2009].
 The neutral theory has been basically applied in rainforests to adress diversity patterns in the guild of canopy trees [@hubbell_unified_2001], thus ignoring dynamics of understorey trees.
-However basic ecological constraints drive resource use strategies and biotic interactions in this ecosystems and imply considering functional strategies and relationship among guilds as well.
+However basic ecological constraints drive resource use strategies and biotic interactions in these ecosystems and imply considering functional strategies and relationship among guilds as well.
 Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988].
@@
 
-In this paper, we used a trait-based approach to adress guild and community dynamics within and among local tree assemblages in a tropical rain forest from Western Ghats, India.
-Although of wet evergreen type, the forest undergoes a long dry season and an intense rainy season under the influence of monsoon.
-In this context, we expect water stress and competition for resource to be key drivers of forest dynamics.
+In this paper, we used a trait-based approach to adress within-guild community dynamics within and among local tree assemblages in a tropical rain forest from Western Ghats, India.
+Although considered as a wet evergreen forest, the forest undergoes a long dry season and an intense rainy season under the influence of monsoon.
+In this context, we expect water stress and competition for resource to be the key drivers of forest dynamics.
 We thus adressed two main questions:
@@
 1. Do variations and covariations in functional traits reflect distinct ecological strategies within and among species ?
-    i. Do and how functional traits vary among species ?
-    ii. Do covariations in functional traits represent axes of ecological strategy within species ?
-2. Does functional composition reflect the influence of local abiotic and biotic drivers of community dynamics ?
+    i. How functional traits vary among species ?
+    ii. Do covariations in functional traits reflect axes of ecological strategy ?
+2. Does functional composition depends on potential local abiotic and biotic drivers of community dynamics ?
     i. Do local functional composition ($\alpha$ scale) reflect the influence of abiotic and biotic drivers of community dynamics ?
@@
 
-The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry of an undisturbed, old-growth wet evergreen Diptrocarp forest [@pascal_structure_1996].
-The plot is located near Uppangala village ("Uppangala plot") in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India) in India's Western Ghats biodiversity hotspot.
-Climate is characterized by both heavy rainfall and very dry seasons.
-Annual precipitation averages 5108 mm per year with 90% between June and October during the monsoon, while the dry season (monthly rainfall $\leq 45$ mm) lasts on average 3.8 months [@pelissier_twenty_2011].
+The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry in an undisturbed, old-growth wet evergreen Dipterocarp forest [@pascal_structure_1996].
+The plot is located near Uppangala village, and thus refers to as "Uppangala plot", in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India).
+Annual precipitation averages 5108 mm per year but is highly variable in time with heavy rainfall from June to October (90% of the rainfall during the monsoon) and a severe dry season (monthly rainfall $\leq 45$ mm) lasting 3.8 months [@pelissier_twenty_2011].
 Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996].
-Climate and geology of the 10-ha plot can be considered as homogeneous.
-General topography of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as a major source of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998].
+Climate and pedology of the 10-ha plot can be considered as homogeneous.
+The topography of the site is highly heterogeneous with a regular east-west alternation between thalwegs and ridges separated by steep slopes, so that topography can be assumed as a major source of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998].
 We hypothesized that topographic variation should influence community variation depending on tree ecological strategies [@condit_spatial_2000].
 
-## Wood and leaf sampling
-
-Since 1989, all trees $\geq 30$ cm girth at breast height (gbh) have been monitored in an area progressively extended to 10 ha in 2014.
-All trees were identified, tagged, mapped, and provided with permanent dendrometer bands censused every 2-5 years [@pelissier_twenty_2011].
+## Floristic and trait data
+
+Since 1989, all trees $\geq 30$ cm in girth at breast height (gbh) have been monitored in an area progressively extended to a contiguous 10-ha plot up to 2014.
+<!-- Need to add Cedric Vega in Acknowledgments for that.
+--> All trees were identified, tagged and mapped over the plot [@pelissier_twenty_2011].
 111 different species were recorded.
 We collected trait data on 89 over 111 (80.1%) species, which represented 99.0% of the tree individuals present in the Uppangala plot.
-With the 32 most abundant species, we reached  more than 80% of Uppangala tree individuals.
-We sampled five individuals per species for these species, and sampled one to five individuals of the less abundant and rare species (in total 267 individuals were sampled).
+Note that the 32 most abundant species, represent  more than 80% of Uppangala tree individuals.
+We sampled five individuals per species for these 32 species, and sampled one to five individuals from the remaining species (in total 267 individuals were sampled).
 For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation from @perez-harguindeguy_new_2013.
 Individuals and leaves were selected in consistent light conditions, development stage and sanitary state.
-Dawkins crown exposure index [@dawkins_management_1958] was assessed for each tree sampled.
-Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
+Dawkins crown exposure index [@dawkins_management_1958] was recorder for each tree sampled.
+Leaf and branch samples were kept in humidified ziplock filled with enriched $CO_{2}$ air inside a black bag untill their measurement in the following 24 hours.
 For some individuals, we were unable to collect an appropriate branch sample (11 missing species over 89 representing 1% of individuals present in the 10 ha plot).
@@
 We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1).
-We choose those five traits because they have been widely used, besides there easier measurement, and they have been shown to be good indicator of species abilities to grow, survive and reproduce [@wilson_specific_1999].
-
-Table 1: Functional traits measured.
-*Traits with their abbreviation, standard unit, and associated trade-off in tree strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*
+We choose those five traits because they have been widely used, in addition of there easier measurement, and they have been shown to be good indicator of species abilities to grow, survive and reproduce [@wilson_specific_1999].
+
+Table 1: Functional traits measured. *Traits with their abbreviation, standard unit, and associated species strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*
 
@@
   Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
-  s = c('Leaf defense vs. investment',
+  Strategy = c('Leaf defense vs. investment',
         'Leaf resource capture vs.investment',
@@
 )
-names(table)[4] <- 'Trade-off'
 kable(table, row.names = F)
@@
 
-Because topography was assumed as a major source of environmental variation [@gimaret-carpentier_sampling_1998], we selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from the digital elevation model using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed using ArcGis [@desktop_arcgis_2011].
-<!-- Digital elevation model was provided by a 2013 LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. -->
-
-In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the environmental variables.
-We selected the less correlated variables with the most ecological meaning.
-Thus we reduced our explanatory variables to 4 variables: curvature representing the water flow paths and soil accumulation areas, wetness highlighting water accumulation areas, southwesterness representing aspect [@bunyan_effect_2015] and slope.
+Because topography was assumed as a major source of environmental variation [@gimaret-carpentier_sampling_1998], we selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed using ArcGis [@desktop_arcgis_2011].
+
+In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the environmental variables (see supplementary materials S1).
+We selected the less correlated variables having the most ecological meaning.
+Thus we reduced our explanatory variables to 4 variables: curvature representing the water flow paths and soil accumulation areas, wetness highlighting water accumulation areas, southwesterness representing aspect [@bunyan_effect_2015] and slope, which may have both a strong effect on forest dynamics and on mutrient/water availability **ref**.
 
 We selected two biotic factors related to local forest structure: canopy height and basal area.
-Canopy height was calculated from a digital canopy model built with the 2013 LIDAR campaign on Uppangala plots with resolution $1*1 m^2$.
-Basal area was calculated from 2013 girth inventory of Uppangala plot on stems with a diameter above 10 centimeters at breast height ($DBH>10~cm$).
+Canopy height, i.e.
+the maximum local height of the canopy, was calculated from a digital canopy model built from a LiDAR (Light Detection And Ranging) campaign done in 2013 over the Uppangala plot with a resolution of $1*1 m^2$.
+Basal area was calculated from 2013 and 2014 girth inventories of Uppangala plot on stems with a diameter above 30 centimeters at breast height ($DBH>30~cm$).
 
@@
 
-The coefficient of variation of traits was investigated at three levels ($cv = \frac{\sigma}{\mu}$ with $\sigma$ the standard deviation and $\mu$ the mean): between leaves of the same individual (intra-individual), between trees of the same species (intra-specific for species with more than one individual), and between species (inter-specific).
-In addition, we performed a principal component analysis (PCA) of the mean value of traits per species, to characterize basic dimensions of ecological strategies.
+To quantify trait variation whithin and between individuals and between species we calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values between leaves of the same individual (intra-individual), between trees of the same species (intra-specific for species with more than one individual), and between species (inter-specific).
+In addition, to assess the covariation of trait within species, and thus potential axes of strategy, we performed a principal component analysis (PCA) of the mean value of traits per species.
 Subsequent community-level analysis were based on trait values averaged per species.
@@
 We defined two guilds to represent vertical stratification called canopy (C), composed of emergent, canopy and sub-canopy tree, and understorey (U) trees.
-The 110 local communities where split by those two guilds and communities accounting for only one species were removed from the subsequent analysis.
-
-To investigate local functional composition of canopy and understorey communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010].
-CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative abundance, $p_{i}$:
-$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
-We also calculated the average of trait values based on species presence-absence, to further emphasize the contribution of less abundant and rare species in community trait distribution [@cingolani_filtering_2007].
-CWV for each trait $t$ was calculated as the mean distance of trait values, $t_{i}$, to $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
+The 110 local communities where split by those two guilds.
+
+To characterize the local functional composition of canopy and understorey communities we used the community weighted mean (CWM), the community weighted variance (CWV), and the coefficient of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010].
+CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $i$ species in the quadrat $j$, with each species, $i$, weighted by its relative abundance, $p_{i}$:
+$$CWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
+We also calculated the average of trait values based on species presence-absence, to further emphasize the contribution of less abundant and rare species in community trait composition [@cingolani_filtering_2007].
+CWV for each trait $t$ was calculated as the mean distance of trait values, $t_{i}$, to $CWM$, of the $S$ species in the quadrat $j$, with each species, $i$, weighted by its relative
 abundance, $p_{i}$:
-$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
-CVNND for each trait $t$, scaled with its mean and variance, was calculated as nearest neigbor distances (NND) standard deviation, $\sigma _{NND}$, of the $i$ individuals in each community, divided by its mean, $\mu _{NND}$; and NND for one individual $i$ is calculated as the minimal distance $d$ between the $j$ other individuals of the community :
-$$CVNND = \frac{\sigma _{NND}}{\mu _{NND}} ~ , ~ ~ ~ NND = [NND_{i} = min(d_{i,j})_{i \neq j}] ~ (3)$$
+$$CWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - CWM_{j})^2 ~ (2)$$
+CVNND for each trait $t$, was calculated as the standard deviation of the nearest neighbor distances (NND), $\sigma _{NND}$, of the $i$ individuals in each community, divided by the mean NND, $\mu _{NND}$; and NND for one individual $i$ is calculated as the minimal distance $d$ between the $j$ other individuals of the community :
+$$CVNND = \frac{\sigma _{NND}}{\mu _{NND}} ~ , ~ ~ ~ NND = {NND_{i} = min(d_{i,j})_{i \neq j}} ~ (3)$$
 
@@
 
-Variation of composition between community pairs with those metrics was translated as a taxonomic, functional and phylogenetic distance, and compared to environmental and spatial distances among community pairs.
-Environmental distances were calculated as geometric distances between communities on the plan composed by the two first axes of environmental variables PCA (see supplementary material S1).
-Correlations with environment and space were assessed based on two Mantel tests successively with taxonomic, functional, and phylogenetic community compositions.
-A firt Mantel statistic assessed the correlation between each metric distance matrix and the environmental distance matrix (E).
-The second partial Mantel statistic assessed the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S), to represent remaining spatial dependency in community composition once the influence of environmental variation is taken into account.
-
-We used null models to test if observed correlation with Mantel statistic differed significantly from expectations under null hypothesis.
-$I_{st}$ correlation to environment and space was tested with 999 repetions of the model nul described by @myers_beta-diversity_2013 which randomizes individuals among subplots while preserving the regional abundance of species, and the number of individuals per community.
+Variation in composition between community pairs with those metrics was translated as a taxonomic, functional and phylogenetic distance, and compared to the corresponding environmental and spatial distances.
+Environmental distances were calculated as euclidean distances between communities on the plan composed by the two first axes of environmental variables PCA (see supplementary material S1).
+Correlations with environment and space were assessed with two Mantel tests successively with taxonomic and functional, community composition.
+The firt Mantel test consisted in assessing the correlation between each metric distance matrix and the environmental distance matrix (E) and the second test, a partial Mantel test, assessed the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S).
+This second test thus test the remaining spatial dependency in community composition once the influence of environmental variation is removed.
+
+We used null models to test if the observed correlations differed significantly from expectations under null hypothesis.
+$I_{st}$ correlations with environment and space were tested with 999 repetions of the nul model described by @myers_beta-diversity_2013 which randomizes individuals among quadrat while preserving the regional abundance of species, and the number of individuals per community.
 $U st$ and $\tau st$ were tested by the nul model NM1 which randomizes traits among species.
 
-All analysis were conducted using the R 3.3.0  software environment [@r_core_team_r:_2016] with packages 
+All analyses were conducted using the R 3.3.0  software environment [@r_core_team_r:_2016] with packages 
 `ade4` [@dray_ade4_2007], 
@@
 
-Table 2: Coefficient of intra-individual, intra-specific and inter-specific trait variation.
+Table 2: Coefficient of variation of intra-individual, intra-specific and inter-specific trait values.
 *Traits with their range of values, mean and standard deviation, mean of coefficient of variation for leaves in all individuals (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
@@
 
-The principal component analysis (PCA) shows WD associated with structural leaf traits orthogonal to the SLA (figure 1).
-The first PCA axis (PCA1) covary mainly with LT, LA, LDMC and WD; opposing thick big leaves and lightwood to dense leaves and hardwood.
-PCA1 axis associated WD to leaf structural traits linked to hydraulic conductive efficiency along the wood economic spectrum [WES, @chave_towards_2009].
-The PCA2 covaried mainly positively with SLA and negatively with LDMC along a leaf economic spectrum [LES, @wright_worldwide_2004].
-The leaf econmic spectrum opposes high SLA leaves for a strong productivity capacity, with high phyotosynthesis compare to carbon investment, and high LDMC leaves to ensure persistence in time and against attacks.
-PCA also highlight the strong correlation between LDMC and WD (Pearson's correlation test, $r = 0.25$ with $p-value < 0.001$), indicating a coordinated structural investment of species among stem and leaf.
+The principal component analysis (PCA) revealed a first axis along which both wood density and structural leaf traits (LT, LDMC) correlate, thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood.
+Interestingly, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between these two variables.
+The second axis of the PCA2 covaried mainly positively with SLA and negatively with LDMC.
+PCA also highlight the strong correlation between LDMC and WD (Pearson's correlation test, $r = 0.25$ with $p-value < 0.001$).
 
@@
 lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
-legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
-       text.col = c("green", "firebrick"))
+# legend('topright', c('Leaf economics spectrum', 'Wood economics spectrum'),
+       # text.col = c("green", "firebrick"))
 
@@
 Figure 1: Principal Component Analysis (PCA) performed on species mean trait values.
-*The bottom right subplot represents the eigen values (grey bars) compare to a broken stick distribution for 5 components (red line).
+*The bottom right subplot represents the eigen values (grey bars) compare to a broken stick distribution for five components (red line).
 See table 1 for abbreviation.*
@@
 See table 1 for abbreviations.
-The number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 See table 1 for abbreviations.
-The number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+Bold and  number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 
@@
 
-Taxonomic variation among communities was correlated significantly to environment variation for both canopy and understorey guilds (Mantel test for canopy and understorey guild, $r_{I_{st},E,C} = 0.22$ and $r_{I_{st},E,U} = 0.23$, respectively, with $p-value < 0.01$), but not with remaining spatial dependency in community composition for canopy guild (Mantel test of space knowing environment, $r_{I_{st},S|E,C} = 0.01$ with $p-value > 0.1$), whereas understorey guild taaxonomic variation was correlated to remaining spatial dependency in community composition (Mantel test of space knowing environment, $r_{I_{st},S|E,U} = 0.11$ with $p-value < 0.01$).
-
-```{r Ist, eval=FALSE, include=FALSE}
+Taxonomic variation among communities was correlated significantly with environmental variation for both canopy and understorey guilds (Mantel test for canopy and understorey guild, $r_{I_{st},E,C} = 0.22$ and $r_{I_{st},E,U} = 0.23$, respectively, with $p-value < 0.01$), but not with remaining spatial dependency in community composition for canopy guild (Mantel test of space knowing environment, $r_{I_{st},S|E,C} = 0.01$ with $p-value > 0.1$), whereas understorey guild taaxonomic variation was correlated to remaining spatial dependency in community composition (Mantel test of space knowing environment, $r_{I_{st},S|E,U} = 0.11$ with $p-value < 0.01$).
+
+```{r Ist, include=FALSE}
 library(spacodiR)
@@
 It entailed that the primary source of trait variation within and among communities should be among species.
-The PCA of functional trait variation among species displayed two main dimensions (figure 1), which basically represented a wood and a leaf economic spectrum, respectively.
-We found opposing SLA to LDMC values along the leaf economic spectrum, which expressed a trade-off between productivity, with less carbon investment for higher photosynthesis rates, and persistence, with more leaf defense and investment [PCA2, @wright_worldwide_2004].
-A second major axis of trait variation opposed WD to LA and LT.
-It thus opposed stress tolerant species, with denser wood, to competitive species, with thicker and larger leaves for more efficient water conductivity.
-It represented the wood economic spectrum as defined by @chave_towards_2009, and the opposition between small thin leave and hardwood and large thick leaves with light wood corresponded to the additional gradient uncovered by @baraloto_decoupled_2010.
-
-Furthermore, our results are similar to the between-species variations found @fortunel_leaf_2012, where basic PCA axes also represented leaf and wood economic spectra (inverted compare to us).
+The PCA of functional trait variation among species displayed two main dimensions (figure 1), which may be interpreted as independent wood and leaf economics spectra.
+We found opposing SLA to LDMC values along the leaf economics spectrum, which expressed a trade-off between productivity, with less carbon investment for higher photosynthesis rates, and persistence, with more leaf defense and investment [@wright_worldwide_2004].
+A second major axis of trait variation opposed wood density to leaf area and thickness.
+It may thus be interpreted as an axis opposing stress tolerant species, with denser wood, to competitive species, with thicker and larger leaves, which confer a higher efficiency in water conductivity.
+It may thus be interpreted as the wood economics spectrum firstly introduced by @chave_towards_2009, and then discussed by @baraloto_decoupled_2010.
+Interestingly, @baraloto_decoupled_2010 found an independence between the leaf and wood economics spectra and their wood economics spectrum also showed an opposition between hardwood species having small thin leaves and light wood species having large thick leaves.
+
+Furthermore, our results are similar to the ones found by @fortunel_leaf_2012, where basic PCA axes also represented leaf and wood economics spectra (inverted compare to us).
 Noticeably though, our analysis integrated less traits (5 against 14), less species (89 against 758), and considered a regional set of species found in a local forest-plot scale while @fortunel_leaf_2012 considered a regional set of species.
-Therefore, we could capture basic economic spectra found at regional scale in Neotropics at a far more local scale in an Indian rainforest, even though such downscaling does not hold in any case [@messier_trait_2016].
+Therefore, we could capture basic economics spectra found at regional scale in Neotropics at a far more local scale in an Indian rainforest, even though such downscaling does not hold in any case [@messier_trait_2016].
 It suggests that these spectra indeed represent major drivers of ecological strategies in local forest dynamics in our context.
 
-## Communtiy-level trait values
+## Community-level trait values
 
 We analyzed functional composition of $20*20m^2$ communities (CWM), as well as the variation of this composition with environment across the 10 ha forest plot.
-We found that the traits of the leaf economic spectrum (LES) were mainly affected by aspect with more convergent community composition (lower CWV) under south-west orientation, especially for LDMC of less common species.
+We found that the traits of the leaf economics spectrum (LES) were mainly affected by aspect with more convergent community composition (lower CWV) under south-west orientation, especially for LDMC of less common species.
 The influence of aspect is little documented in tropical forests, but still highlighted by @gunatilleke_specieshabitat_2006 in monsoon rainforests.
@@
 
-The wood economic spectrum defined by WD, LA, and LT was mainly affected at community level by south-west orientation and curvature.
+The wood economics spectrum defined by WD, LA, and LT was mainly affected at community level by south-west orientation and curvature.
 South-west orientation, with probably more disturbance and stress due to exposure to monsoon winds **Maxime: counter effect by topo ?**, resulted in denser wood.
@@
 
-A phylogenetic tree of our 89 species was built with `S.PhyloMaker` function proposed by @qian_updated_2016-1.
-We assessed phylogenetic signal of each functional using the Moran's I statistic.
-Positive Moran's I value indicated convergence of trait values within the phylogeny whereas negative Moran's I will showed divergence.
+A phylogenetic tree of the 89 species was built with the R `S.PhyloMaker` function proposed by @qian_updated_2016-1.
+We assessed the phylogenetic signal of each functional trait using the Moran's I statistic.
+Positive Moran's I value indicated convergence of trait values within the phylogeny (i.e.
+trait conversatism) whereas negative Moran's I will showed divergence (i.e.
+trait differentiation).
 

### 2017-03-01 inst/doc/draft_mrm.Rmd
Uppangala 0.1.9037 draft mrm version

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation
  rainforest of India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output: pdf_document
csl: mee.csl
bibliography: refs.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(parallel)
cores <- detectCores() - 1
library(Uppangala)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

Built with `Uppangala` package version `r packageVersion('Uppangala')`.

# Abstract

## Keywordsand supplementary materials S2

functional trait; rainforest; vertical forest strata; Western Ghats; environmental filtering; niche differentiation; ecological guild

# Introduction

Understanding the determinants of species occurrence and co-occurrence is a long standing issue in ecology [@diamond_ecology_1975]. Tropical rainforests have especially fascinated ecologists due to their outstanding diversity [@connell_diversity_1978]. Niche theory explains species coexistence through niche differences that limit competitive exclusion [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that maintenance of high diversity is possible even for functionally equivalent species, under the influence of stochastic life, death, reproduction and dispersal dynamics [@hubbell_unified_2001]. Both theories can provide realistic expectations of taxonomic diversity [@chave_comparing_2002] so that inferring ecological processes from taxonomical composition pattern is challenging in community ecology.

Investigating patterns of functional or phylogenetic diversity may reveal signatures of ecological processes that are specific to the niche theory [@mcgill_rebuilding_2006; @westoby_land-plant_2006; @mouquet_ecophylogenetics:_2012]. Indeed, if taxonomic composition is expected to be spatially structured in both theories, due to the spatial structure of environment or to dispersal limitation, functional or phylogenetic assemblage are expected to be spatially structured only under the niche theory [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012; @parmentier_prevalence_2014].

The variation in functional traits among species is expected to reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007] and expresses fundamental trade-offs along environmental gradients [@mark_westoby_plant_2002]. For instance, specific leaf area varies along the soil fertility-shade gradient and represents a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@hodgson_is_2011]. Fundamental trade-offs entail functional traits covariation along basic leaf [@wright_worldwide_2004 but see @osnas_global_2013 and @lloyd_photosynthetically_2013] and wood [@chave_towards_2009] economics spectra. The leaf economics spectrum opposes acquisitive to conservative ecological strategies, basically expressed at leaf level. The wood economics spectrum distinguishes species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength [@chave_towards_2009]. @baraloto_decoupled_2010 found decoupled wood and leaf economics of trees in neotropical forest, but @reich_world-wide_2014 conversely posited that both spectra have to be integrated in a comprehensive whole plant economics spectrum [@freschet_evidence_2010] opposing fast growing acquisitive species to slow growing and more conservative ones.

The trait values of coexisting species represent a non-random sample of trait values found in a broader area, du to specific niche-based dynamics in the local community [@lortie_rethinking_2004; @bernard-verdier_community_2012]. Then the signature of local niche-based dynamics can be gasped by assessing the deviation of local community structure from a larger "species pool", classicaly using null models [@kraft_functional_2008]. In addition, the nature of functional traits and ecological strategies driving species coexistence can change across scales. For instance, @messier_trait_2016 have underlined that the leaf economics spectrum basically expresses changes in ecological strategies along broad-scale environmental gradients, while it is not necessarily a primary driver of local coexistence.

The concept of guilds refer to groups of functionally similar species [@wilson_guilds_1999]. Ecological guilds can coexist and occupy distinct niches within communities [@vergnon_niches_2009]. In tropical forests, the assembly of tree communities has been mostly studied irrespectivelly of guilds of canopy and understorey tree [@wright2002plant]. However, distinctive ecological constraints are expected to drive different resource use strategies and biotic interactions in these guilds, which should translate into different functional composition and response to environmental factors. For instance, in Colombian Amazonie, @duque2002different found some contrasted response of the composition of canopy and understorey guilds along environmental gradients.

In this paper, we used a trait-based approach to study separately the assembly of canopy and understorey tree species along topographical and biotic gradients in a large forest plot (10ha) loacted in Western Ghats biodiversity hotspot [@myers_biodiversity_2000], India. Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988]. It thus provides an appropriate context to adress guild-dependent assembly rules and ecological strategies in tropical rainforests. Although considered as wet evergreen, the forest undergoes contrasted long dry and intense rainy seasons under the influence of monsoon. In this context, we expected water stress and competition for resource to be important drivers of tree functional composition. We investigated how both canopy and understorey tree species composition varied with local environmental conditions, using appropriate null models. We specifically addressed the following questions:

- Does functional traits variation and covariation reflect basic dimensions of ecological strategy in the rainforest ?
- Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?

# Material and methods

## Study site

The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry in an undisturbed, old-growth wet evergreen Dipterocarp forest [@pascal_structure_1996]. The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), near the Uppangala village, and thus refers to as "Uppangala plot". Annual precipitation averages 5108 mm per year but is highly variable in time with heavy rainfall from June to October (90% of the rainfall during the monsoon) and a severe dry season (monthly rainfall $\leq 45$ mm) lasting 3.8 months [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996]. Climate and geology of the 10-ha plot can be considered as homogeneous. The topography of the site is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes, so that topography is expected to be a major component of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998]. Therefore, we hypothesized that topographic variation should influence community assembly and yield changing ecological strategies and functional composition [@condit_spatial_2000].

## Taxonomic composition and trait sampling

From 1989 to 2014, the trees $\geq 10$ cm in diameter at breast height (dbh) were identified to species level, tagged, mapped and monitored in an area progressively extended to a contiguous $300 * 330 m^2$ (ca. 10-ha) plot [@pelissier_twenty_2011]. The plot included 111 different tree species. We collected trait data from 89 over 111 (80.1%) species, which represented 99.0% of tree individuals present in the plot. Note that the 32 most abundant species represented more than 80% of tree individuals. We sampled five individuals per species for these 32 species, and sampled one to five individuals for the remaining species (in total 267 individuals). For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation of @perez-harguindeguy_new_2013. Individuals and leaves were selected in consistent light conditions, development stage and sanitary state. Dawkins crown exposure index [@dawkins_management_1958] was recorded for each tree sampled to control for any sun-exposure effect. Leaf and branch samples were kept in humidified ziplock filled with enriched $CO_{2}$ air inside a black bag until measurement within the next 24 hours. For some individuals, we were unable to collect suitable branch sample, i.e. with a diameter $\leq 1$ cm (11 missing species over 89 representing 1% of individuals present).

## Trait measurements

Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weights, thickness and area were measured, using a weighing scale (**add detail on brand and precision**), a micrometer (**add detail on brand and precision**) and a bend scanner with ImageJ software [**ref**] respectively. Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$. Once dried, the leaves were weighted directly after been extracted from the oven to avoid any humidity bias. The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues. Branch wood samples were immersed in water to evaluate their volume by water displacement. Wood samples were then dried for 48 hours at 80 $^\circ C$ and weighted.

We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1). We choose those five traits because they proved to be good proxies of the major dimensions of ecological strategies in plants [@mark_westoby_plant_2002; @diaz_global_2015] and were relatively easy to measure on the field.

Table 1: Functional traits measured. *Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].*

```{r 1 Traits}
rm(list = ls()[-which(ls() == "cores")]); invisible(gc())
PFT <- genPFT()
table <- data.frame(
  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  Strategy = c('Leaf defense and investment',
        'Leaf resource capture vs.investment',
        'Leaf defense and investment',
        'Leaf resource acquisiton vs. defense',
        'Stem transport, structure and defense')
)
kable(table, row.names = F)
rm(table)
```

## Environmental variables

Topography was expected to be a major environmental factor influencing the assemby of tree communities [@gimaret-carpentier_sampling_1998]. We selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed in ArcGis [**ref**].

In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the abiotic environmental variables (see Supplementary Materials S1). We selected four little-correlated variables: curvature representing local water flow paths and soil accumulation areas, wetness highlighting water accumulation areas over the whole study area, southwesterness representing aspect [@bunyan_effect_2015] and slope, which may have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher].

In addition, we selected two biotic factors related to local forest structure: canopy height and basal area. Canopy height, i.e. the maximum local height of the canopy, was calculated from a 1-m resolution digital canopy model built from a LiDAR (Light Detection And Ranging) dataset acquired in 2013 over the Uppangala plot. Basal area was calculated from the girth measurements taken within the 10-ha plot (censuses done in 2013 and 2014) for stems $\geq 10$ cm dbh.

## Analyses

### Guilds

@pascal_wet_1988 defined four vertical strata attained by adult trees: emergent above 39m, canopy trees between 25 and 39m, sub-canopy trees between 15 and 25m, and understorey trees below 15m. Despite the vertical hierarchy, steep topography could allow sub-canopy trees receiving more light, especially in slopes, and thus participates to canopy. We thus defined two basic guilds called canopy (C), composed of emergent, canopy and sub-canopy species (n = 55), and understorey (U) species (n = 34).

### Traits variation and covariation

We first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among leaves of the same individual (intra-individual), among trees of the same species (intra-specific for species with more than one individual), and among species (inter-specific). We then performed a principal component analysis (PCA) of the mean value of traits per species to identify main dimensions of ecological strategies along which traits can covary among guilds. We tested more specifically association of tree species from each guild to identified ecological strategies. To do so, we compared variance of the mean value of traits per species between main ecological strategies with Levene's test **ref**. Subsequent community-level analyses were based on trait values averaged per species.

### Drivers of community dynamics within guilds

We subdivided the 10-ha plot into $30*30 m^2$ quadrats leading to 110 local communities. We analyze the functionnal composition of the guilds in the 110 communities. Guilds in the communities have a mean number of 28 individuals representing 10 species, with less in understorey than in canopy. We calculated weighted mean [GWM, where G stands for "guild"; @violle_let_2007] for each guild in each community. In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of species $i$ weighted by species relative abundance $p_{i}$:
$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
where $S$ is the number of species in the quadrat $j$. We also calculated the average trait values based on presence-absence values to assess the influence of less abundant and rare species on community trait composition [@cingolani_filtering_2007].

To test for any significant changes in functional composition among guilds, we devised a null model approach. A null model (NM1) randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix) [@bernard-verdier_community_2012]. Note that, spatial correlations in tree abundance are conserved limiting associated type I error. We performed 999 randomizations of null model NM1 and we derived a two-tailed test of the proportion of guilds deviating from the null expectation, with adjusted p-values by Benjamini-Hochberg correction **ref**. Shifts in GWM values with changing environmental conditions could reflect the shift of species strategies across assemblages [@ackerly_trait-based_2007].

To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as
$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (2)$$
where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of the corresponding null distribution. The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$). Analyses were conducted on the five functional traits and the species scores obtained on the two first PCA axes done on the `Trait*Species` matrix .

All analyses were conducted using the R 3.3.0  software [@r_core_team_r:_2016] with packages 
`ade4` [@dray_ade4_2007],
and `vegan` [@oksanen_vegan:_2016].

# Results

## Traits variation and covariation

We found an overall increase of functional trait variation from intra-individual to inter-specific levels (see table 2), even though LA showed almost as large variability at intra-individual than at intra-specific levels ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$). given the relatively low intra-individual and intra species variation, compared to inter-species variation, using species trait average values for community analyses is warranted.

Table 2: Intra-individual, intra-specific and inter-specific component of trait variation. *The overall range, mean and standard deviation are provided for each trait. The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species. See table 1 for abbreviation.*

```{r 2 CV}
# Individual aggregation
PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
row.names(PFT_ind) <- PFT_ind$Group.1
SpT <- unique(PFT[1:2])
Sp <- SpT$Sp
names(Sp) <- SpT$Tree
SpCodeT <- unique(PFT[2:3])
SpCode <- SpCodeT$SpCode
names(SpCode) <- SpCodeT$Sp
CET <- unique(PFT[c(1,4)])
CE <- CET$CE
names(CE) <- CET$Tree
PFT_ind$Sp <- Sp[row.names(PFT_ind)]
PFT_ind$SpCode <- SpCode[as.character(PFT_ind$Sp)]
PFT_ind$CE <- CE[row.names(PFT_ind)]
PFT_ind <- PFT_ind[-c(1:2)]
PFT_ind$ID <- row.names(PFT_ind)
row.names(PFT_ind) <- PFT_ind$ID
PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
# Species aggregation
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$Sp), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$SpCode <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
# Coefficient of variation
trait <- c('LT', 'LA', 'LDMC', 'SLA', 'WD')
cv <- data.frame(
  row.names = trait,
  Range = c(paste(range(PFT$LT, na.rm = T), collapse = '-'),
            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
  m = c(paste0(round(mean(PFT$LT, na.rm = T)), ' (',
               round(sd(PFT$LT, na.rm = T)), ')'),
        paste0(round(mean(PFT$LA, na.rm = T)), ' (',
               round(sd(PFT$LA, na.rm = T)), ')'),
        paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
               round(sd(PFT$LDMC, na.rm = T)), ')'),
        paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
               round(sd(PFT$SLA, na.rm = T),2), ')'),
        paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
               round(sd(PFT$WD, na.rm = T),2), ')')),
  m1 = round(unlist(lapply(as.list(trait), function(y){
    mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){
      sd(x, na.rm = T) / mean(x, na.rm = T)
    })[,2], na.rm = T)
  })) ,2),
  m2 = round(unlist(lapply(as.list(trait), function(y){
    mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$Sp), function(x){
      sd(x, na.rm = T) / mean(x, na.rm = T)
    })[,2], na.rm = T)
  })), 2),
  m3 = round(unlist(
    lapply(as.list(trait), function(y){
      sd(PFT_sp[,which(names(PFT_sp) ==  y)], na.rm = T) / mean(PFT_sp[,which(names(PFT_sp) ==  y)], na.rm = T)
    })), 2)
)
names(cv)[2:5] <- c('Mean (sd)' , 'Intra-individual', 'Intra-species', 'Inter-species')
cv['WD','Intra-individual'] <- ''
kable(cv)
rm(PFT, CET, cv, CE, Sp, SpT, SpCodeT)
```

The first axis of the principal component  analysis (PCA) of mean species trait values was basically related to wood density and structural leaf trait variations (LT, LA, LDMC), thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood. The second axis mostly opposed SLA to LDMC, and to a less extent to wood density. However, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between those two traits.

```{r 3 PCA PFT}
library(vegan)

# Individual to log
PFT_ind$LA <- log(PFT_ind$LA)
PFT_ind$SLA <- log(PFT_ind$SLA)
PFT_ind$ID <- row.names(PFT_ind)
row.names(PFT_ind) <- PFT_ind$ID
PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
# Species aggregation
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$Sp), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$SpCode <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] == 3)
PFT_sp$strata <- NA
PFT_sp[-sel,'strata'] <- 'C'
PFT_sp[sel,'strata'] <- 'U'

# PCA
pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT_sp, cor = T)
lambda <- pca$sdev * sqrt(nrow(pca$scores))
pca.var = apply(pca$score, 2, var)
PCAscores <- as.data.frame(pca$scores)[1:2]
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
plot(t(t(pca$scores)/lambda),pch=16, col = c('green','red')[as.factor(PFT_sp[row.names(PCAscores),'strata'])],
     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
par(new=T)
Rot <- t(t(pca$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
legend('topright', c('Understorey', 'Canopy'),
      fill = c("green", "red"))

arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
Hmisc::subplot(function(){
  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
               cex.names = 0.7, axes = T, names.arg = NA) ;
  # lines(b, bstick(length(pca.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.75,1), from='npc'),
  y=grconvertY(c(0.05,0.30), from='npc')
)
axis (3)
axis (4)

# LES and WES
PCAscores <- as.data.frame(pca$scores)[1:2]
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
PFT_sp$ID <- row.names(PFT_sp)
PFT_sp <- merge(PFT_sp, PCAscores, all = T)
row.names(PFT_sp) <- PFT_sp$ID
PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
rm(PCAscores, PFT_ind, SpCodeT, lambda, pca, pca.var, Rot, XLIM)
```

Figure 1: Principal Component Analysis (PCA) of species mean trait values. *The bottom right subplot represents the eigen values (grey bars) of the five components. First axis, in red, is named wood economic spectrum (WES) ; whereas second axis, in green, is named leaf economic spectrum (LES). See table 1 for abbreviation.*

Leven's tests revealed a significantly higher variance for canopy tree species on PCA first axis than second axis ($table~3$). Variance of canopy tree species is also significantly higher than variance of understorey species on PCA first axis. Other differences of variance were not significant.

Table 3: Leven's test of PCA's guilds variance. *Guilds variance inside the principal component analysis (PCA) were compared with Leven's test on PCA first and second axes.*

```{r variance PCA}
PCAscores1 <- ade4::dudi.pca(PFT_sp[-which(is.na(PFT_sp$WD)), c('LA', 'WD', 'LT', 'LDMC', 'SLA')], scannf = F)$l1
names(PCAscores1) <- c('WES', 'LES')
sel <- which(PFT_sp[row.names(PCAscores1),'strata'] == 'C')
groups <- list(C.WES = PCAscores1$WES[sel],
               U.WES = PCAscores1$WES[-sel],
               C.LES = PCAscores1$LES[sel],
               U.LES = PCAscores1$LES[-sel])
tests <- list(WES = car::leveneTest(c(groups$C.WES, groups$U.WES), 
                               c(rep(1, length(groups$C.WES)), rep(2, length(groups$U.WES)))),
              LES = car::leveneTest(c(groups$C.LES, groups$U.LES), 
                               c(rep(1, length(groups$C.LES)), rep(2, length(groups$U.LES)))),
              C = car::leveneTest(c(groups$C.WES, groups$C.LES), 
                             c(rep(1, length(groups$C.WES)), rep(2, length(groups$C.LES)))),
              U = car::leveneTest(c(groups$U.WES, groups$U.LES), 
                             c(rep(1, length(groups$U.WES)), rep(2, length(groups$U.LES)))))
tests <- lapply(tests, function(x) paste(round(x$`Pr(>F)`[1],3), stars(x$`Pr(>F)`[1])))

table <- aggregate.data.frame(PCAscores1[1:2], by = list(PFT_sp[row.names(PCAscores1),'strata']), FUN = var)
table[2:3] <- round(table[2:3], 3)
names(table) <- c('Variance', 'PCA 1', 'PCA 2')
table$Variance <- c('Canopy', 'Understorey')
table[3,] <- c('Levene\'s test', tests$WES, tests$LES)
table[4] <- c(tests$C, tests$U, '')
names(table)[4] <- 'Levene\'s test' 
kable(bold(table), row.names = F, digits = 3)
```

```{r 4 env-com}
library(raster)

# Building community and environmental datasets
Env <- genEnv()
Trees <- genTrees()
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp)
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats(cs = c(30,30))
XY <- Trees[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
Trees$com <- (XY %over% com)[,1]
com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/share/applications')
names(Deriv)[5] <- 'SW'
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)

# Building RLQ matrices
R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(R) <- com@data$id
L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
L[is.na(L)] <- 0
Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')]
row.names(Q) <- PFT_sp$SpCode
sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] == 3)
Q <- list(C = Q[-sel,], U = Q[sel,])
sel <- lapply(Q, function(q){intersect(row.names(q), names(L))})
Q <- mapply(function(q,s){q[s,]}, q = Q, s = sel, SIMPLIFY = F)
L <- lapply(sel, function(s){L[s]})
R <- lapply(L, function(l){R[row.names(l),]})
L <- lapply(L, function(l){list(abundance = l, `presence-absence` = l)})
L <- lapply(L, function(l){l$`presence-absence`[l$`presence-absence` > 1]  <- 1 ; return(l)})

rm(Deriv, XY, Env, sel, BA, s)
```

## Functional composition of guilds

The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds (see table 4). Results in presence-absence are in adequation with observed results in abundance (see supplementary materials S2). Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits, and two major differences were observed between the two guilds. The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height affected only understory guild composition.

Focusing on canopy tree species, we found an increase in wood density under south-west orientation, both in abundance and presence-absence (WES provided consistent results, Table 4). We also observed a decrease in leaf area and thickness under the south-west orientation, but only in abundance and presence-absence respectively. Leaf area decreased with curvature in both abundance and presence-absence, while wood density and LDMC increased and decreased respectively with curvature in presence-absence only.

For understorey tree species, SLA significantly decreased and wood density and leaf thickness significantly increased with curvature, both in abundance and presence-absence. We also detected a significant increase and decrease in wood economic spectrum and leaf area respectively with curvature in abundance. SLA in both abundance and presence-absence significantly increased with canopy height, while LDMC and wood density decreased with canopy height in abundance only.

Table 4: Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. _We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown. See table 1 for abbreviations. Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r CWM}
# Observed CWMs
CWM <- mapply(function(Ls,Qs){
  lapply(Ls, function(l){
    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
      apply(l, 1, function(x){
        weighted.mean(Qs[,v], x, na.rm=T)
      })
    })
  })
}, Ls = L, Qs = Q, SIMPLIFY = F)
CWM <- lapply(CWM, function(s){
  lapply(s, function(x){
    names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')
    return(x)
  })
})

# Simulated CWMs
n = 999
nQ <- lapply(Q, function(s){
  rep(list(s), n)
})
nQ <- lapply(nQ, function(s){ 
  lapply(s, function(nq){
      row.names(nq) <- sample(row.names(nq))
      return(nq)
  })
})
nQ <- mapply(function(nQs, Ls){
  lapply(nQs, function(nq){
    nq[names(Ls$abundance),]
  })
}, nQs = nQ, Ls = L, SIMPLIFY = F)
cl <- makeCluster(cores)
clusterExport(cl, list('L', 'nQ'))
nCWM <- clusterMap(cl, function(nQs, Ls){
  lapply(nQs, function(nq){
    lapply(Ls, function(l){
      lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
        apply(l, 1, function(x){
          weighted.mean(nq[,v], x, na.rm=T)
        })
      })
    })
  })
}, Ls = L, nQs = nQ)
stopCluster(cl)
rm(cl)
nCWM <- lapply(nCWM, function(s){
  lapply(s, function(ns){
  lapply(ns,function(ab){
    names(ab) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')
    return(ab)
    })
  })
})
nCWM <- lapply(nCWM, function(s){
  lapply(s, `[`, names(s[[1]]))
  apply(do.call(rbind, s), 2, as.list)
})
nCWM <- lapply(nCWM, function(s){
  lapply(s, function(ab){
    lapply(ab, `[`, names(ab[[1]]))
    apply(do.call(rbind, ab), 2, as.list)
  })
})
nCWM <- lapply(nCWM, function(s){
  lapply(s, function(ab){
    lapply(ab, function(x){
      do.call('cbind', x)
    })
  })
})
nCWM <- lapply(nCWM, function(s){
  lapply(s, function(ab){
    lapply(ab, data.frame)
  })
})

## CWM SES
CWM.ses <- mapply(function(CWMs, nCWMs){
  mapply(function(CWMab, nCWMab){
    data.frame(
      mapply(function(x,y){
        mapply(function(p,q){
          (p - mean(q))/sd(q)
        }, p = as.list(x), q = as.list(data.frame(t(y))))
      }, x = CWMab, y = nCWMab)
    )
  }, CWMab = CWMs, nCWMab = nCWMs, SIMPLIFY = F)
}, CWMs = CWM, nCWMs = nCWM, SIMPLIFY = F)
CWM.ses.lm <- mapply(function(s, Rs){
  lapply(s, function(ab){
    apply(ab, 2, function(y){
      step(lm(y ~., Rs), direction = 'both', trace = 0)
    })
  })
}, s = CWM.ses, Rs = R, SIMPLIFY = F)
CWM.ses.res <- lapply(CWM.ses.lm, function(s){
  lapply(s, function(ab){
    lapply(ab, function(x){
      summary(x)$coefficients
    })
  })
})
CWM.ses.res <- lapply(CWM.ses.res, function(s){
  lapply(s, function(ab){
    lapply(ab, data.frame)
  })
})
CWM.ses.res <- lapply(CWM.ses.res, function(s){
  lapply(s, function(ab){
    lapply(ab, function(x){
      x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   '))
      return(x)
    })
  })
})
CWM.ses.res <- lapply(CWM.ses.res, function(s){
  lapply(s, function(ab){
    data.frame(data.table::rbindlist(lapply(ab, function(x){
      data.frame(t(x[5]))
    }), fill = T))
  })
})
var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
CWM.ses.res <- lapply(CWM.ses.res, function(s){
  lapply(s, function(ab){
  ab[,setdiff(var, names(ab))] <- NA
  ab <- ab[var]
  ab <- apply(ab, 2, as.character)
  row.names(ab) <- names(CWM.ses$C$abundance)
  ab[is.na(ab)] <- ''
  ab <- data.frame(ab)
  return(ab)
  })
})
CWM.ses.res <- mapply(function(s, CWM.ses.lms){
  mapply(function(ab, CWM.ses.lmab){
    ab$`$R^2$` <- round(unlist(lapply(CWM.ses.lmab, function(x){
      summary(x)$r.squared
    })),3)
    return(ab)
  }, ab = s, CWM.ses.lmab = CWM.ses.lms, SIMPLIFY = F)
}, s = CWM.ses.res, CWM.ses.lms = CWM.ses.lm, SIMPLIFY = F)
CWM.ses.res <- lapply(CWM.ses.res, function(s){
  lapply(s, function(ab){
    ab <- cbind(row.names(ab), ab)
    names(ab)[1] <- 'Trait'
    return(ab)
  })
})
CWM.ses.res <- lapply(CWM.ses.res, function(s){
  mapply(function(ab, name){
      ab <- cbind(c(name, rep(' ', length(ab[,1])-1)), ab)
      names(ab)[1] <- 'T'
      return(ab)
  },ab = s, name = as.list(c('AB', 'PA')), SIMPLIFY = F)
})
CWM.ses.res <- lapply(CWM.ses.res, function(s){
  do.call('rbind', s)
})
CWM.ses.res <- mapply(function(s, name){
  s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
  names(s)[1] <- 'S'
  return(s)
}, s = CWM.ses.res, name = as.list(c('C', 'U')), SIMPLIFY = F)
CWM.ses.res <- do.call('rbind', CWM.ses.res)
rm(nCWM, CWM.ses.lm)
invisible(gc())
```

```{r CWM AB}
kable(bold(CWM.ses.res[c(1:7,15:21),-2]), row.names = F)
```

# Discussion

## **"Take home messages" to add :**

- **Pantropical and multiscale WES and LES ?**
- **Decoupled wood and leaf strategies among canopy and understorey tree species**
- **WES among canopy tree species**
- **LES among understorey tree species**

We found back economics spectra evidenced by @wright_worldwide_2004 and @chave_towards_2009 at finer scale than @messier_trait_2016 suggested (figure 1). This result advocate for a pantropical and multiscale oganisation of tropical plant strategies in two separated economics spectra. Still our results suggest that this two spectra are decoupled between canopy and understorey tree species, whereas existing in both guilds. Canopy tree species being the first layer of forest ecosystem interacting with external abiotic environment shows major variation in wood economics spectrum (table 3). Indeed wood economic spectrum is related to mechanical resistance, disturbance tolerance and water conductivity efficiency. Thus understorey tree will benefit from a mircoclimate **[ref?]** generated by the presence of the canopy tree. In this microclimate, leaf economics spectrum is then a major axis of strategy to answer to a changing light environment and soil fertility.

## Wood and leaf economics hold at the local scale

We found an increasing trait variation within individuals, within species, and between species (table 2). Our analyses thus reveal that the primary source of trait variation within and among communities is the inter-species variation, justifying the use of species mean traits for the study of tropical tree guild assembly. The PCA of functional trait variation among species displayed two main dimensions (figure 1), very similar to the ones found by @baraloto_decoupled_2010 and @fortunel_leaf_2012. Those authors interpreted these axes as independent wood and leaf economics spectra. We found that SLA and LDMC tend to covary negatively along the leaf economics spectrum, which be seen as a trade-off between productivity, with less carbon investment and higher photosynthesis rates, and persistence, with more leaf defense and investment [@wright_worldwide_2004]. A second major axis of trait variation opposed wood density to leaf area and thickness. It may be interpreted as an axis opposing stress tolerant species, with denser wood, to competitive species, with thicker and larger leaves, conferring a higher efficiency in water conductivity **[ref?]**. It may thus be interpreted as the wood economics spectrum firstly introduced by @chave_towards_2009, and then discussed by @baraloto_decoupled_2010. Interestingly, @baraloto_decoupled_2010 found an independence between the leaf and wood economics spectra and their wood economics spectrum also showed an opposition between hardwood species having small thin leaves and light wood species having large thick leaves.

Furthermore, our results are similar to the ones found by @fortunel_leaf_2012, where basic PCA axes also represented leaf and wood economics spectra (inverted compare to us). Noticeably, our analysis integrated less traits (5 against 14), less species (89 against 758), and considered a set of species found in a local forest-plot while @fortunel_leaf_2012 considered a regional set of species from French Guiana and Peru (i.e. the opposite ends of Amazonia). Therefore, our results shows that the economics spectra found at a regional scale in the Neotropics are also valid at a far more local scale in an Indian rainforest. This result does support the arguments by @messier_trait_2016 and suggests that the two observed spectra underline major ecological strategies in tropical forests, even locally.

## Canopy and understorey species assembly rely on different economics spectra

We found a significanlty higher variance of canopy tree species around the wood economic spectrum (WES) rather than the leaf economic spectrum (LES). Similarly understorey tree species seemed to have higher variance around leaf economic spectra (LES) rather than wood economics spectra (WES), besides it was not significant at a level $\alpha = 0.05$ (table 3). Those results advocates for decoupled strategies between species of understorey and canopy guilds [@duque2002different]. Indeed canopy tree species could deploy strategies to endorse the role of first layer against external abiotic environment. Those strategies will come through wood economic spectrum (WES) to ensure mechanical resistance, disturbance tolerance and water conductivity efficiency **[biblio ?]**. On the other hand, understorey tree species will encounter constraints directly linked to canopy guild limiting light availability. Thus they will need to deploy strategies linked to ressource acquisition coming through leaf economics spectrum (LES) **[biblio ?]**.

Furthermore, we analyzed functional composition of $30*30m^2$ guilds from canopy and understorey, as well as the variation of this composition with environment across a 10-ha forest plot. We foud that the variations in trait composition along the environmental gradients differ between canopy and understorey species, with a marked variation in traits associated to the leaf economic spectrum (LES) for understorey species and to the wood economics scpectrum (WES) for canopy species. In the former case, the curvature and the canopy height both had a significant effect on the LES-based composition of understorey tree species, whereas southwest orientation and curvature both had a significant effect on the WES-based composition of canopy tree species (table 4).

## Environment influence guild-level functionnal composition

We should note that relations found between guilds weighted means (GWM) and environment are globally weak (table 4, most of $R^2 \leq 0.1$, and see supplementary materials S4). This low explanatory power is advocating for underlying null process related to neutral theory [@hubbell_unified_2001]. Null process can effectivelly be responsible for the majority of observed variance in functionnal composition at a such fine scale **[ref?]**. Studies highlighting environmental filtering effect often present a higher environmental variation than in our study site [@bernard-verdier_community_2012]. In parallel, the 10*ha plot scale is really low and partially responsible for the small environmental variation. @kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25 ha tropical forest plots including large flat valleys.

Curvature is affecting the variation of leaf economics in understorey species and the variation of wood economics in canopy tree species. A locally convex context would accumulate more soil and attract more water, increasing soil fertility and allowing persistence of species investing more in leaf resource acquisition in understorey with higher SLA and lower LDMC, or less in wood ressources in canopy with lighter wood and bigger leaves converging more toward an homogeneous guild composition (lower CWV, unshown result). Curvature for some functional traits, as wood density, LDMC and SLA, is also doubled by a similar effect of the wetness variable expressing also the water accumulation in convex context resulting in increased soil fertility. Economics spectra are not perfectly divided between canopy and understorey, with wood and leaf respectivelly, for instance LDMC of canopy species also decreased with curvature whereas WD increased in understorey.

The influence of aspect on forest composition is little documented in tropical forests. In a similar context that our study site, @gunatilleke_specieshabitat_2006 gave some intuitive insights on the effect of aspect on forest composition: he suggested that aspect through monsoon had a potential effect on rainforests. @bunyan_effect_2015 tested the effect of aspect on the occurence of forest fragments in a forest-grassland mosaic and foud that southwest orientation strongly determined the presence of forest in mountain rainforests of Western Ghats. They suggested that it could result from monsoon winds. In our context, monsoon winds are south-west oriented, and are expected to be a major source of disturbance in forests [@soman_aspects_1990].

The Uppangala plot is situated in a topographically complex area with steep slopes mostly organized along southwest-oriented crests. In this context it is diffcult to predict wind flow and orientation inside the 10-ha plot topography. We propose two explanation, hardwood could be selected for mechanical resistance in more disturbed souwhtwest oriented area or topography provoke a turbulences on northeast oriented areas responsible of more disturbance and death events in the forest history resulting in lightwood filtering [@larjavaara_rethinking_2010]. In both case it seems reasonable that canopy tree species are the most influenced by disturbance events linked to monsoon, and may even buffer the effect on understorey species.

Biotic factors, illustrated here by the canopy height and the basal area, also had a significant effect on the LES-based composition of understorey tree species. We indeed found higher SLA values and lower LDMC values in understorey assemblages that are located under stands characterized by a tall canopy and a high basal area. This results contrasts with the expectation that understorey species invest in higher LDMC for higher leaf lifespan, at the expense of SLA, unde light-limited conditions [@kitajima_tissue-level_2010]. The Uppangala forest is characterized by a large heterogeneity of canopy height allowing light to penetrate more in lower strata, maybe especially more in high canopy context covering less closely understorey, resulting in those observation on the leaf economics variation of understorey species. **Not convincing, better explanation ?**

# *Acknowledgement*

* French Institute of Pondicherry (FIP)
* Cedric Vega for the extension of the plot

# Supplementary materials

## S1: Principal component analysis of topographical variables

```{r PCA Env}
lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
pca.env.var = apply(pca.env$score, 2, var)
plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
Hmisc::subplot(function(){
  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(pca.env.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.07,0.37), from='npc'),
  y=grconvertY(c(0.025,0.30), from='npc')
)
par(new=T)
Rot <- t(t(pca.env$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, lwd = 2)
lines(c(0,0), XLIM, lty = 2, lwd = 2)
col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
        Rot[,1],Rot[,2],lwd=2,col=col)
legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
       text.col = col, cex = 1.2)
text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
axis (3)
axis (4)
```

Supplementary material S1: Principal component analysis (PCA) performed on topgraphical variables. *Botom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 7 components (red line). SW stand for southwesterness.*

Curvature covaries with profile curvature, plan curvature, distance to potential rivers and the PCA first axis. Elevation and wetness are opposed in the principal component analysis and slope and south-westerness does not seem to covary much with other variables **Not precise**.

## S2: Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in presence-absence

```{r CWM PA}
kable(bold(CWM.ses.res[c(8:14,22:28),-2]), row.names = F)
```

Supplementary material S2: Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in presence-absence _We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown. See table 1 for abbreviations. Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

## S3: Standard effect size (SES) of wood density  (WD) for canopy (C) tree species

```{r WD}
data <- cbind(CWM.ses$C$abundance$WD, R$C)
names(data)[1] <- 'WD'
par(mfrow = c(3,2)) # Affichage des relations individuelles
plot(WD ~ Slope, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
plot(WD ~ Curvature, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
plot(WD ~ Wetness, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
plot(WD ~ SW, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
plot(WD ~ Canopy, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
plot(WD ~ BA, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
rm(data, CWM.ses)
```

Supplementary materials S3 : Standard effect size (SES) of wood density  (WD) for canopy (C) tree species. *Each subplot show  standard effect size (SES) of wood density (WD) for canopy (C) tree species along one environmental variable gradient. Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant. See table 1 for abbreviation.*

# References

### 2017-07-19 inst/doc/draft/draft2.Rmd
Reprise Kourou (draft2 v0)

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation
  rainforest of India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
csl: mee.csl
bibliography: refs.bib
knit: "bookdown::render_book"
output:
  bookdown::gitbook:
    split_by: none
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(parallel)
library(Uppangala)
library(broom)
library(vegan)
library(dplyr)
library(reshape2)
library(ggplot2)
cores <- detectCores() - 1
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
path <- '~/Documents/IFP/Uppangala/inst/doc/draft_save/'
n = 9 # Random repetitions
```

# Introduction {-}

# Data

```{r 1 Traits}
PFT <- genPFT()
kable(data.frame(
  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  Strategy = c('Leaf defense and investment',
        'Leaf resource capture vs.investment',
        'Leaf defense and investment',
        'Leaf resource acquisiton vs. defense',
        'Stem transport, structure and defense')
), row.names = F, caption = 'Functional traits measured. Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].')
```

# Trait variation and covariation

## Coefficients of variation

```{r 2 CV}
PFT_ind <- PFT %>% 
  group_by(Tree, Sp, SpCode, CE) %>% 
  summarise_each(funs(mean))
PFT_sp <- PFT_ind %>% 
  ungroup() %>% 
  dplyr::select(-Tree, -CE) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_each(funs(mean))

# Coefficient of variation
trait <- c('LT', 'LA', 'LDMC', 'SLA', 'WD')
cv <- data.frame(
  row.names = trait,
  Range = c(paste(range(PFT$LT, na.rm = T), collapse = '-'),
            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
  m = c(paste0(round(mean(PFT$LT, na.rm = T)), ' (',
               round(sd(PFT$LT, na.rm = T)), ')'),
        paste0(round(mean(PFT$LA, na.rm = T)), ' (',
               round(sd(PFT$LA, na.rm = T)), ')'),
        paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
               round(sd(PFT$LDMC, na.rm = T)), ')'),
        paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
               round(sd(PFT$SLA, na.rm = T),2), ')'),
        paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
               round(sd(PFT$WD, na.rm = T),2), ')')),
  m1 = round(unlist(lapply(as.list(trait), function(y){
    mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){
      sd(x, na.rm = T) / mean(x, na.rm = T)
    })[,2], na.rm = T)
  })) ,2),
  m2 = round(unlist(lapply(as.list(trait), function(y){
    mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$Sp), function(x){
      sd(x, na.rm = T) / mean(x, na.rm = T)
    })[,2], na.rm = T)
  })), 2),
  m3 = round(unlist(
    lapply(as.list(trait), function(y){
      sd(unlist(PFT_sp[which(names(PFT_sp) ==  y)]), na.rm = T) / mean(unlist(PFT_sp[which(names(PFT_sp) ==  y)]), na.rm = T)
    })), 2)
)
names(cv)[2:5] <- c('Mean (sd)' , 'Intra-individual', 'Intra-species', 'Inter-species')
cv['WD','Intra-individual'] <- ''
kable(cv, caption = 'Intra-individual, intra-specific and inter-specific component of trait variation. The overall range, mean and standard deviation are provided for each trait. The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species. See table 1 for abbreviation.')
```

## Principal component analysis

```{r 3 PCA PFT,  fig.cap='Principal Component Analysis (PCA) of species mean trait values. The bottom right subplot represents the eigen values (grey bars) of the five components. First axis, in red, is named wood economic spectrum (WES) ; whereas second axis, in green, is named leaf economic spectrum (LES). See table 1 for abbreviation.'}
rm(cv)
PFT_sp <- PFT_ind %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  group_by(Sp, SpCode) %>% 
  dplyr::select(-Tree, -CE) %>% 
  summarise_each(funs(mean))
# PCA
pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT_sp, cor = T)
lambda <- pca$sdev * sqrt(nrow(pca$scores))
pca.var = apply(pca$score, 2, var)
PCAscores <- as.data.frame(pca$scores)[1:2]
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
      xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
par(new=T)
Rot <- t(t(pca$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
legend('topright', c('Leaf economics spectrum', 'Wood economics spectrum'),
       text.col = c("green", "firebrick"))

arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
Hmisc::subplot(function(){
  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
          cex.names = 0.7, axes = T, names.arg = NA) ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.75,1), from='npc'),
  y=grconvertY(c(0.05,0.30), from='npc')
)
axis (3)
axis (4)
```

# Functional composition of guilds

```{r 4 env-com}
# LES and WES
PCAscores <- as.data.frame(pca$scores)[1:2]
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
PFT_sp$ID <- row.names(PFT_sp)
PFT_sp <- merge(PFT_sp, PCAscores, all = T)
row.names(PFT_sp) <- PFT_sp$ID
PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
rm(PCAscores, PFT_ind, lambda, pca, pca.var, Rot, XLIM)
# library(raster)
# Env <- genEnv()
# Trees <- genTrees()
# Species <- genSpecies()
# Trees$ID <- row.names(Trees)
# Trees <- merge(Trees, PFT_sp)
# row.names(Trees) <- Trees$ID
# Trees <- Trees[-which(names(Trees) == 'ID')]
# com <- quadrats(cs = c(30,30))
# XY <- Trees[c('x', 'y')]
# coordinates(XY) <- ~ x + y
# proj4string(XY) <- crs(com)
# Trees$com <- (XY %over% com)[,1]
# com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/share/applications')
# names(Deriv)[5] <- 'SW'
# com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
# com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
# com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
# pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
# R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
# row.names(R) <- com@data$id
# L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
# L[is.na(L)] <- 0
# Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')]
# row.names(Q) <- PFT_sp$SpCode
# sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] == 3)
# Q$strata <- 'C' ; Q$strata[sel] <- 'U'
# sel <- intersect(row.names(Q), names(L))
# Q <- Q[sel,]
# Q$SpCode <- row.names(Q)
# L <- L[sel]
# R <- R[row.names(L),]
# R$community <- row.names(R)
# Lbin <- L ; Lbin[Lbin > 1]  <- 1
# rm(Deriv, XY, Env,  sel, BA, com, Trees, Species)
# save(R, L, Lbin, Q, pca.env, file = file.path(path, 'envcom.Rdata'))
rm(PFT, PFT_sp, trait)
load(file.path(path, 'envcom.Rdata'))
```

## Guild weighted means : nul model 1

```{r GWM1}
GWM <- L %>% 
  t() %>% 
  data.frame() %>% 
  mutate(SpCode = row.names(.)) %>% 
  left_join(Q) %>% 
  select(-SpCode) %>% 
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>% 
  group_by(strata, community) %>% 
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# nQ <- array(rep(as.matrix(Q), n),
#             c(dim(Q), n),
#             list(row.names(Q), names(Q), 1:n))
# nQ <- array(apply(nQ, 3, function(x){
#   x[,'SpCode'] <- sample(x[,'SpCode'])
#   return(x)}), c(dim(Q), n),
#   list(row.names(Q), names(Q), 1:n))
# cl <- makeCluster(cores)
# clusterExport(cl, list('L', 'nQ'))
# nGWM <- parApply(cl, nQ, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   L %>%
#     t() %>%
#     data.frame() %>%
#     mutate(SpCode = row.names(.)) %>%
#     left_join(data.frame(x)) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     group_by(strata, community) %>%
#     mutate_each(funs(as.character), -abundance) %>%
#     mutate_each(funs(as.numeric), -abundance) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# })
# stopCluster(cl)
# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
#               c(dim(GWM[3:9]), n),
#               list(row.names(GWM), names(GWM[3:9]), 1:n))
# GWM_ses <- GWM
# GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
# rm(cl, nQ, nGWM)
# save(GWM_ses, file = file.path(path, 'GWM1_ses.Rdata'))
load(file.path(path, 'GWM1_ses.Rdata'))
GWM_ses %>% 
  melt(id.vars = c('strata', 'community')) %>% 
  rename(trait = variable) %>% 
  rename(SES = value) %>% 
  left_join(R) %>%
  select(-community) %>% 
  group_by(strata, trait) %>% 
  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
          direction = 'both', trace = 0))) %>% 
  group_by(strata, trait, term) %>%
  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
                          stars(p.value, ns = '   '))) %>% 
  dcast(strata + trait ~ term) %>% 
  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
  bold() %>% 
  kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown. See table 1 for abbreviations. Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
```

```{r WD, fig.cap='Exemple of standard effect size (SES) of wood density  (WD) for canopy (C) tree species. Each subplot show  standard effect size (SES) of wood density (WD) for canopy (C) tree species along one environmental variable gradient. Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant. See table 1 for abbreviation.'}
GWM_ses %>% 
  select(strata, WD, community) %>% 
  left_join(R) %>% 
  filter(strata == 'C') %>% 
  ungroup() %>% 
  select(-strata, -community) %>% 
  melt(id.vars = 'WD') %>% 
  ggplot(aes(value, WD)) +
  geom_point() +
  facet_wrap(~ variable, scales = "free") +
  geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') +
  geom_hline(yintercept = 0) +
  xlab('Environmental gradient') +
  ylab('Standard effect size of wood density')
```

## Guild weighted means : nul model 3

## Guild weighted variances : nul model 2

```{r GWV3}
GWV <- L %>% 
  t() %>% 
  data.frame() %>% 
  filter(sum(.) > 2) %>% 
  mutate(SpCode = row.names(.)) %>% 
  left_join(Q) %>% 
  select(-SpCode) %>% 
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>% 
  melt(id.vars = c('strata', 'community', 'abundance'),
       variable.name = 'trait', value.name = 'trait_value') %>% 
  left_join(melt(GWM, id.vars = c('strata', 'community'),
                 variable.name = 'trait', value.name = 'GWM')) %>% 
  group_by(strata, community, trait) %>% 
  mutate(GWV = (trait_value - GWM)^2) %>% 
  select(-trait_value, -GWM) %>% 
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  dcast(strata + community ~ trait)
# nL <- array(replicate(n, t(apply(L, 1, function(x){
#   x[x != 0] <- sample(x[which(x != 0)]) ; return(x)}))),
#   c(dim(L), n),
#   list(row.names(L), names(L), 1:n))
# cl <- makeCluster(cores)
# clusterExport(cl, list('nL', 'Q', 'GWM'))
# nGWV <- parApply(cl, nL, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   x %>%
#     t() %>%
#     data.frame() %>%
#     filter(sum(.) > 2) %>%
#     mutate(SpCode = row.names(.)) %>%
#     left_join(Q) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     melt(id.vars = c('strata', 'community', 'abundance'),
#          variable.name = 'trait', value.name = 'trait_value') %>%
#     left_join(melt(GWM, id.vars = c('strata', 'community'),
#                    variable.name = 'trait', value.name = 'GWM')) %>%
#     group_by(strata, community, trait) %>%
#     mutate(GWV = (trait_value - GWM)^2) %>%
#     select(-trait_value, -GWM) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%
#     dcast(strata + community ~ trait)
# })
# stopCluster(cl)
# nGWV <- array(unlist(lapply(nGWV, function(x) as.matrix(x[3:9]))),
#               c(dim(GWV[3:9]), n),
#               list(row.names(GWV), names(GWV[3:9]), 1:n))
# GWV_ses <- GWV
# GWV_ses[3:9] <- (GWV[3:9] - apply(nGWV, 1:2, mean, na.rm = T))/apply(nGWV, 1:2, sd, na.rm = T)
# rm(cl, nL, nGWV)
# save(GWV_ses, file = file.path(path, 'GWV_ses.Rdata'))
load(file.path(path, 'GWV_ses.Rdata'))
GWV_ses %>% 
  melt(id.vars = c('strata', 'community')) %>% 
  rename(trait = variable) %>% 
  rename(SES = value) %>% 
  left_join(R) %>%
  select(-community) %>% 
  group_by(strata, trait) %>% 
  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
          direction = 'both', trace = 0))) %>% 
  group_by(strata, trait, term) %>%
  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
                          stars(p.value, ns = '   '))) %>% 
  dcast(strata + trait ~ term) %>% 
  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
  bold() %>% 
  kable(caption = 'Linear models of the standardized effect sizes (SES) of GWV and CVNND, for canopy (C) and understorey (U) guilds. We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown. See table 1 for abbreviations. Bold and  number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
```

# Supplementary materials

## Principal component analysis of environment

```{r PCA Env, fig.cap='Principal component analysis (PCA) performed on topgraphical variables. Botom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 7 components (red line). SW stand for southwesterness.'}
lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
pca.env.var = apply(pca.env$score, 2, var)
plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
Hmisc::subplot(function(){
  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(pca.env.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.07,0.37), from='npc'),
  y=grconvertY(c(0.025,0.30), from='npc')
)
par(new=T)
Rot <- t(t(pca.env$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, lwd = 2)
lines(c(0,0), XLIM, lty = 2, lwd = 2)
col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
        Rot[,1],Rot[,2],lwd=2,col=col)
legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
       text.col = col, cex = 1.2)
text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
axis (3)
axis (4)
```

# References {-}

### 2017-07-28 inst/doc/draft/draft2.Rmd
Tabebuia

@@
 ---
-title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation
-  rainforest of India
+title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
 author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
@@
 library(ggplot2)
+library(ggfortify)
 cores <- detectCores() - 1
@@
 
-```{r 1 Traits}
+```{r Traits}
 PFT <- genPFT()
@@
         'Stem transport, structure and defense')
-), row.names = F, caption = 'Functional traits measured.
-Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].')
+), caption = 'Functional traits measured. Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].')
 ```
@@
 
-```{r 2 CV}
-PFT_ind <- PFT %>% 
+```{r CV}
+PFT %>% 
+  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
+  melt(variable.name = 'trait') %>% 
+  group_by(trait) %>% 
+  summarise_each(funs(min(.,na.rm = T), max(., na.rm = T), 
+                      mean(.,na.rm = T), sd(.,na.rm = T))) %>% 
+  left_join(PFT %>% 
+              select(-Sp, -SpCode, -CE, -SHA, -LTD) %>% 
+              melt(id.vars = 'Tree', 
+                   variable.name = 'trait', 
+                   value.name = 'Intra-individuals') %>% 
+              group_by(Tree, trait) %>% 
+              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
+              ungroup() %>% 
+              select(-Tree) %>% 
+              group_by(trait) %>% 
+              summarise_each(funs(mean(., na.rm = T)))) %>% 
+  left_join(PFT %>% 
               group_by(Tree, Sp, SpCode, CE) %>% 
-  summarise_each(funs(mean))
-PFT_sp <- PFT_ind %>% 
+              summarise_each(funs(mean)) %>% 
               ungroup() %>% 
-  dplyr::select(-Tree, -CE) %>% 
+              select(-Tree, -SpCode, -CE, -SHA, -LTD) %>% 
+              melt(id.vars = 'Sp', 
+                   variable.name = 'trait', value.name = 'Intra-species') %>% 
+              group_by(Sp, trait) %>% 
+              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
+              ungroup() %>% 
+              select(-Sp) %>% 
+              group_by(trait) %>% 
+              summarise_each(funs(mean(., na.rm = T)))) %>% 
+  left_join(PFT %>% 
+              group_by(Tree, Sp, SpCode, CE) %>% 
+              summarise_each(funs(mean)) %>% 
+              ungroup() %>% 
+              select(-Tree, -CE) %>% 
               group_by(Sp, SpCode) %>% 
-  summarise_each(funs(mean))
-
-# Coefficient of variation
-trait <- c('LT', 'LA', 'LDMC', 'SLA', 'WD')
-cv <- data.frame(
-  row.names = trait,
-  Range = c(paste(range(PFT$LT, na.rm = T), collapse = '-'),
-            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
-            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
-            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
-            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
-  m = c(paste0(round(mean(PFT$LT, na.rm = T)), ' (',
-               round(sd(PFT$LT, na.rm = T)), ')'),
-        paste0(round(mean(PFT$LA, na.rm = T)), ' (',
-               round(sd(PFT$LA, na.rm = T)), ')'),
-        paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
-               round(sd(PFT$LDMC, na.rm = T)), ')'),
-        paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
-               round(sd(PFT$SLA, na.rm = T),2), ')'),
-        paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
-               round(sd(PFT$WD, na.rm = T),2), ')')),
-  m1 = round(unlist(lapply(as.list(trait), function(y){
-    mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){
-      sd(x, na.rm = T) / mean(x, na.rm = T)
-    })[,2], na.rm = T)
-  })) ,2),
-  m2 = round(unlist(lapply(as.list(trait), function(y){
-    mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$Sp), function(x){
-      sd(x, na.rm = T) / mean(x, na.rm = T)
-    })[,2], na.rm = T)
-  })), 2),
-  m3 = round(unlist(
-    lapply(as.list(trait), function(y){
-      sd(unlist(PFT_sp[which(names(PFT_sp) ==  y)]), na.rm = T) / mean(unlist(PFT_sp[which(names(PFT_sp) ==  y)]), na.rm = T)
-    })), 2)
-)
-names(cv)[2:5] <- c('Mean (sd)' , 'Intra-individual', 'Intra-species', 'Inter-species')
-cv['WD','Intra-individual'] <- ''
-kable(cv, caption = 'Intra-individual, intra-specific and inter-specific component of trait variation.
+              summarise_each(funs(mean)) %>% 
+              ungroup() %>% 
+              select(-Sp, -SpCode, -SHA, -LTD) %>% 
+              melt(variable.name = 'trait', value.name = 'Inter-species') %>% 
+              group_by(trait) %>% 
+              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T)))) %>% 
+  kable(digits = 2, caption = 'Intra-individual, intra-specific and inter-specific component of trait variation.
 The overall range, mean and standard deviation are provided for each trait.
 The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species.
-See table 1 for abbreviation.')
+See table \\@ref(tab:Traits) for abbreviation.')
 ```
@@
 
-```{r 3 PCA PFT,  fig.cap='Principal Component Analysis (PCA) of species mean trait values.
-The bottom right subplot represents the eigen values (grey bars) of the five components.
+```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values.
 First axis, in red, is named wood economic spectrum (WES) ; whereas second axis, in green, is named leaf economic spectrum (LES).
-See table 1 for abbreviation.'}
-rm(cv)
-PFT_sp <- PFT_ind %>% 
+See table \\@ref(tab:Traits) for abbreviation.'}
+PFT <- PFT %>% 
   mutate(LA = log(LA)) %>% 
   mutate(SLA = log(SLA)) %>% 
+  group_by(Tree, Sp, SpCode, CE) %>% 
+  summarise_each(funs(mean(., na.rm = T))) %>% 
+  ungroup() %>% 
   group_by(Sp, SpCode) %>% 
-  dplyr::select(-Tree, -CE) %>% 
-  summarise_each(funs(mean))
-# PCA
-pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT_sp, cor = T)
-lambda <- pca$sdev * sqrt(nrow(pca$scores))
-pca.var = apply(pca$score, 2, var)
-PCAscores <- as.data.frame(pca$scores)[1:2]
-names(PCAscores) <- c('WES', 'LES')
-PCAscores$ID <- row.names(PCAscores)
-plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
-      xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
-par(new=T)
-Rot <- t(t(pca$loadings)*lambda)
-XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
-XLIM <- XLIM+(XLIM*0.2)
-plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
-lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
-lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
-legend('topright', c('Leaf economics spectrum', 'Wood economics spectrum'),
-       text.col = c("green", "firebrick"))
+  summarise_each(funs(mean(., na.rm = T))) %>% 
+  ungroup() %>% 
+  select(-Tree, -Sp, -CE, -LTD, -SHA)
+row.names(PFT) <- PFT$SpCode
 
-arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
-text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
-Hmisc::subplot(function(){
-  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
-          cex.names = 0.7, axes = T, names.arg = NA) ;
-  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
-         box.col = 'white')},
-  x=grconvertX(c(0.75,1), from='npc'),
-  y=grconvertY(c(0.05,0.30), from='npc')
-)
-axis (3)
-axis (4)
+PFT %>% 
+  princomp(~ LA + WD + LT + LDMC + SLA, data = ., cor = T) %>% 
+  autoplot(colour = 'grey', loadings.label.size = 6,
+           loadings.label.colour = 'black',
+           loadings = T, loadings.label = T, loadings.colour = 'black',
+           loadings.label.vjust = 1.2) +
+  coord_equal() +
+  geom_hline(aes(yintercept = 0, linetype = 'Wood'), 
+             col = 'red') +
+  geom_vline(aes(xintercept = 0, linetype = 'Leaf'), 
+             col = 'green') +
+  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
+  scale_linetype_manual(name = "Economics\nSpectrum", values = c(2, 2), 
+                        guide = guide_legend(override.aes = 
+                                            list(color = c("red", "green")))) +
+  theme(legend.position = c(1, 1), 
+       legend.justification = c(1, 1))
 ```
 
-# Functional composition of guilds
+## Hypervolumes
 
@@
 # LES and WES
-PCAscores <- as.data.frame(pca$scores)[1:2]
-names(PCAscores) <- c('WES', 'LES')
-PCAscores$ID <- row.names(PCAscores)
-PFT_sp$ID <- row.names(PFT_sp)
-PFT_sp <- merge(PFT_sp, PCAscores, all = T)
-row.names(PFT_sp) <- PFT_sp$ID
-PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
-rm(PCAscores, PFT_ind, lambda, pca, pca.var, Rot, XLIM)
+PFT <- PFT %>% 
+  princomp(~ LA + WD + LT + LDMC + SLA, data = ., cor = T) %>% 
+  .$scores %>% 
+  data.frame() %>% 
+  select(Comp.1, Comp.2) %>% 
+  rename(WES = Comp.1) %>% 
+  rename(LES = Comp.2) %>% 
+  mutate(SpCode = row.names(.)) %>% 
+  left_join(PFT)
 # library(raster)
@@
 ```
+
+```{r hypervolume, fig.cap='Functional hypervolume overlap of canopy and understorey species.'}
+bd <- Q %>% 
+  filter(!is.na(WES)) %>% 
+  select(-LES, -WES, -SpCode, -strata) %>% 
+  summarise_all(funs(sd)) %>% 
+  unlist()
+CalcOverlap <- function(traits_strata){
+  hv <- lapply(traits_strata, function(strata) 
+    hypervolume::hypervolume(strata[names(bd)], bandwidth = bd, verbose = F))
+  set <- hypervolume::hypervolume_set(hv$C, hv$U, check_memory = F, verbose = F)
+  hypervolume::hypervolume_sorensen_overlap(set)
+}
+overlap <- Q %>% 
+  filter(!is.na(WES)) %>% 
+  select(-LES, -WES, -SpCode) %>% 
+  split(., .$strata) %>% 
+  CalcOverlap()
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('Q', 'CalcOverlap', 'bd'))
+# n_overlap <- parSapply(cl, 1:n, function(i) {
+#   library(dplyr)
+#   Q %>% 
+#     filter(!is.na(WES)) %>% 
+#     select(-LES, -WES, -SpCode) %>% 
+#     mutate(strata = sample(strata)) %>% 
+#     split(., .$strata) %>% 
+#     CalcOverlap()
+# })
+# stopCluster(cl)
+# save(n_overlap, file = file.path(path, 'overlap.Rdata'))
+load(file.path(path, 'overlap.Rdata'))
+data.frame(x = n_overlap) %>% 
+  ggplot(aes(x)) +
+  geom_density() +
+  geom_vline(xintercept = overlap,
+             col = 'red') +
+  ggtitle(paste('pvalue =', 
+                round((rank(c(overlap, n_overlap))/(n+1))[1], 2))) +
+  xlab('Overlap null distribution')
+```
+
+```{r hypervolume2}
+data <- Q %>% 
+  filter(!is.na(WES)) %>% 
+  select(-LES, -WES, -SpCode) %>% 
+  split(., .$strata)
+hv <- lapply(data, function(strata) 
+    hypervolume::hypervolume(strata[names(bd)], bandwidth = bd, verbose = F))
+set <- hypervolume::hypervolume_set(hv$C, hv$U, check_memory = F, verbose = F)
+hv <- list(C = set@HVList$Unique_1,
+           U = set@HVList$Unique_2,
+           Intersect = set@HVList$Intersection)
+plot(hypervolume::hypervolume_join(hv$C, hv$U))
+hv <- list(C = data$C[
+  hypervolume::hypervolume_inclusion_test(hv$C, data$C[names(bd)]),],
+           U = data$U[
+  hypervolume::hypervolume_inclusion_test(hv$U, data$U[names(bd)]),],
+           Intersect = do.call('rbind', data)[
+  hypervolume::hypervolume_inclusion_test(hv$Intersect, 
+                                          do.call('rbind', data)[names(bd)]),])
+
+kable(do.call('rbind', lapply(hv, function(x)apply(x[names(bd)], 2, mean))),
+      caption = 'means of hypervolumes functional traits')
+rm(hv, set, data, CalcOverlap, bd)
+```
+
+# Functional composition of guilds
 
@@
 We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
-See table 1 for abbreviations.
+See table \\@ref(tab:Traits) for abbreviations.
 Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
@@
 Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
-See table 1 for abbreviation.'}
+See table \\@ref(tab:Traits) for abbreviation.'}
 GWM_ses %>% 
@@
 ## Guild weighted means : nul model 3
+
+```{r GWM3}
+GWM <- L %>% 
+  t() %>% 
+  data.frame() %>% 
+  mutate(SpCode = row.names(.)) %>% 
+  left_join(Q) %>% 
+  select(-SpCode) %>% 
+  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+       variable.name = 'community', value.name = 'abundance') %>% 
+  group_by(strata, community) %>% 
+  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# nQ <- array(rep(as.matrix(Q), n),
+#             c(dim(Q), n),
+#             list(row.names(Q), names(Q), 1:n))
+# nQ <- array(apply(nQ, 3, function(x){
+#   x[,'strata'] <- sample(x[,'strata'])
+#   return(x)}), c(dim(Q), n),
+#   list(row.names(Q), names(Q), 1:n))
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('L', 'nQ'))
+# nGWM <- parApply(cl, nQ, 3, function(x){
+#   library(dplyr)
+#   library(reshape2)
+#   L %>%
+#     t() %>%
+#     data.frame() %>%
+#     mutate(SpCode = row.names(.)) %>%
+#     left_join(data.frame(x)) %>%
+#     select(-SpCode) %>%
+#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+#          variable.name = 'community', value.name = 'abundance') %>%
+#     group_by(strata, community) %>%
+#     mutate_each(funs(as.character), -abundance) %>%
+#     mutate_each(funs(as.numeric), -abundance) %>%
+#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# })
+# stopCluster(cl)
+# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
+#               c(dim(GWM[3:9]), n),
+#               list(row.names(GWM), names(GWM[3:9]), 1:n))
+# GWM_ses <- GWM
+# GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
+# rm(cl, nQ, nGWM)
+# save(GWM_ses, file = file.path(path, 'GWM3_ses.Rdata'))
+load(file.path(path, 'GWM3_ses.Rdata'))
+GWM_ses %>% 
+  melt(id.vars = c('strata', 'community')) %>% 
+  rename(trait = variable) %>% 
+  rename(SES = value) %>% 
+  left_join(R) %>%
+  select(-community) %>% 
+  group_by(strata, trait) %>% 
+  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
+          direction = 'both', trace = 0))) %>% 
+  group_by(strata, trait, term) %>%
+  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
+                          stars(p.value, ns = '   '))) %>% 
+  dcast(strata + trait ~ term) %>% 
+  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
+  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
+  bold() %>% 
+  kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
+```
 
@@
 We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
-See table 1 for abbreviations.
+See table \\@ref(tab:Traits) for abbreviations.
 Bold and  number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
@@
 
-```{r PCA Env, fig.cap='Principal component analysis (PCA) performed on topgraphical variables.
-Botom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 7 components (red line).
-SW stand for southwesterness.'}
-lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
-pca.env.var = apply(pca.env$score, 2, var)
-plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
-     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
-Hmisc::subplot(function(){
-  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
-          cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(pca.env.var)), col = 'red') ;
-  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
-         box.col = 'white')},
-  x=grconvertX(c(0.07,0.37), from='npc'),
-  y=grconvertY(c(0.025,0.30), from='npc')
-)
-par(new=T)
-Rot <- t(t(pca.env$loadings)*lambda)
-XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
-XLIM <- XLIM+(XLIM*0.2)
-plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
-lines(XLIM, c(0,0), lty = 2, lwd = 2)
-lines(c(0,0), XLIM, lty = 2, lwd = 2)
-col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
-arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
-        Rot[,1],Rot[,2],lwd=2,col=col)
-legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
-       text.col = col, cex = 1.2)
-text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
-axis (3)
-axis (4)
+```{r PCAenv, fig.cap='Principal component analysis (PCA) performed on topgraphical variables. SW stand for southwesterness.'}
+pca.env %>% 
+  autoplot(colour = 'grey', loadings.label.size = 6,
+           loadings.label.colour = 'black',
+           loadings = T, loadings.label = T, 
+           loadings.colour = 'black',
+           loadings.label.vjust = 2) +
+  coord_equal() +
+  geom_hline(yintercept = 0, linetype = 'dashed') +
+  geom_vline(xintercept = 0, linetype = 'dashed') +
+  xlab('Axe 1 − 42.52 %') + ylab('Axe 2 − 23.44 %')
 ```

### 2017-07-28 inst/doc/draft/draft2.Rmd
New results

@@
 output:
-  bookdown::gitbook:
-    split_by: none
+  bookdown::gitbook
 ---
@@
 path <- '~/Documents/IFP/Uppangala/inst/doc/draft_save/'
-n = 9 # Random repetitions
+n = 999 # Random repetitions
 ```
@@
 
+In this study, we used a trait-based approach to study separately the assembly of canopy and understorey tree species along topographical and biotic gradients in a large forest plot (10ha) loacted in Western Ghats biodiversity hotspot (Myers et al.
+2000), India.
+Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata (Pascal 1988).
+It thus provides an appropriate context to adress guild-dependent assembly rules and ecological strategies in tropical rainforests.
+Although considered as wet evergreen, the forest undergoes contrasted long dry and intense rainy seasons under the influence of monsoon.
+In this context, we expected water stress and competition for resource to be important drivers of tree functional composition.
+We investigated how both canopy and understorey tree species composition varied with local environmental conditions, using appropriate null models.
+We specifically addressed the following questions:
+
+* Does functional traits variation and covariation reflect basic dimensions of ecological strategy in the
+rainforest ?
+* Does the local functional composition reflect the influence of abiotic and biotic drivers of community
+dynamics within guilds ?
+
+In this document I will only briefly present all analysis, one by one, with their methods and results but without any discussion.
+
 # Data
+
+We used functional traits describe in table \@ref(tab:Traits) gathered in Uppangala forest among 89 species. In subsequent analysis we used log transformation of LA and SLA.
 
@@
 ## Coefficients of variation
+
+We calculated coefficient of variations of trait values among leaves of the same individual (intra-individual), among trees of the same species (intra-specific for species with more than one individual), and among species (inter-specific) with following formula :
+$$Coefficient~of~Variation = \frac{standard~deviation}{mean} $$
+
+We found an overall increase of functional trait variation from intra-individual to inter-specific levels (see table \@ref(tab:CV)), even though LA showed almost as large variability at intra-individual than at intra-specific levels (CV individual = 0.20 and CV species = 0.24, wilcoxon rank test p − value = 0.137).
 
@@
 ## Principal component analysis
+
+We then performed a principal component analysis (PCA) of the mean value of traits per species to identify main dimensions of ecological strategies along which traits can covary among guilds (see figure \@ref(fig:PCA)).
+
+The first axis of the principal component analysis (PCA) of mean species trait values was basically related to wood density and structural leaf trait variations (LT, LA, LDMC), thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood.
+The second axis mostly opposed SLA to LDMC, and to a less extent to wood density.
+However, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between those two traits.
 
@@
 
+Finally we calculated hypervolume of functional traits for each guilds and their overlap.
+We also calculated 999 nul hypervolumes randomizing species strata.
+Contrarily to a previous document, observed hypervolume overlap between the two guilds was not significantly lower than null expectations ($p = 0.165$, see figure \@ref(fig:hypervolume)).
+This could be explained by the change of formula used to calcul hypervolume overlap in the last version of the hypervolume R package (sorensen overlap changed in version 2.1).
+We still have a methodological question, the hypervolume is built on unnormalized traits values.
+Traits standrad deviation is used to sample around the species points, so there is no issu around the bandwidth.
+But we still can question the fact that variable with higher value variations can diminish the overlap signal.
+
 ```{r hypervolume, fig.cap='Functional hypervolume overlap of canopy and understorey species.'}
@@
   ggtitle(paste('pvalue =', 
-                round((rank(c(overlap, n_overlap))/(n+1))[1], 2))) +
+                round((rank(c(overlap, n_overlap))/(n+1))[1], 3))) +
   xlab('Overlap null distribution')
 ```
+
+ Effectivelly some traits seems to show strong differentation between guilds (see figure and table \@ref(tab:hypervolume2)).
 
@@
 
+We subdivided the 10-ha plot into $30*30 m^2$ quadrats leading to 110 local communities.
+We analyze the functionnal composition of the guilds in the 110 communities.
+We calculated weighted mean [GWM, where G stands for "guild"; @violle_let_2007] for each guild in each community.
+In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of species $i$ weighted by species relative abundance $p_{i}$:
+$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
+where $S$ is the number of species in the quadrat $j$.
+We also calculated the average trait values based on presence-absence values to assess the influence of less abundant and rare species on community trait composition [@cingolani_filtering_2007].
+
+To test for any significant changes in functional composition among guilds, we devised a null model approach.
+We described further each null moels when introducing GWM or GWV results.
+We performed 999 randomizations of null models and we derived a two-tailed test of the proportion of guilds deviating from the null expectation, with adjusted p-values by Benjamini-Hochberg correction.
+Shifts in GWM or GWV values with changing environmental conditions could reflect the shift of species strategies across assemblages [@ackerly_trait-based_2007].
+
+To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (2)$$
+where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of the corresponding null distribution.
+The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$).
+Analyses were conducted on the five functional traits and the species scores obtained on the two first PCA axes done on the `Trait*Species` matrix .
+
 ## Guild weighted means : nul model 1
+
+We used a null model (NM1) which randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix) [@bernard-verdier_community_2012].
+
+We foud back the results of previous documents (see table \@ref(tab:GWM1)).
+The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds.
+Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits, and two major differences were observed between the two guilds.
+The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height affected mostly understory guild composition.
+
+I did not detailed back specific trait results. And the question to keep or not economics spectrum in the table is opened ?
 
@@
 
+We used a null model (NM3) which randomized species among guilds in the whole plot (see table \@ref(tab:GWM3)).
+
+We found that **curvature** was responsible for the guilds differenciation for most of the traits.
+Increased curvature resulted in denser lwood and leaves with high SLA in canopy, against lighter wood, low SLA, but bigger and thicker leaves in understorey.
+Some curvature results were congruent with **Wetness** variable.
+**South-West** orientation diminished the leaf area while increasing leaf density for canopy tree species.
+Finally, **Canopy** height and **basal area** affected in opposite ways SLA.
+Increased canopy height and decreased basal area resulted in lower SLA in canopy and increased SLA in understorey.
+
 ```{r GWM3}
@@
 ## Guild weighted variances : nul model 2
+
+The second model (NM2) randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012].
+
+We found significant relationships between the standardized effect size (SES) of GWVs and the environmental predictors (see table \@ref(tab:GWV3)).
+**Wetness** decreased variance of wood density in canopy and all traits in understorey.
+**Curvature** increased LDMC variance in canopy.
+**Sout-West** orientation results in decreased variance of density of wood and leaf in understorey.
+Finally, **Canopy** hieght diminished variance of leaf are and density in canopy whereas **basal aea** increased variance of SLA in both strata and increased leaf area variance in canopy.
 
@@
 
+No comment yet.
+
 ```{r PCAenv, fig.cap='Principal component analysis (PCA) performed on topgraphical variables. SW stand for southwesterness.'}

### 2017-10-04 inst/doc/Hypervolume.Rmd
hypervolume update

---
title: "Organisation des guildes autour d'axes de stratégies"
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
csl: mee.csl
bibliography: refs.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()) ; invisible(gc())
library(knitr)
library(parallel)
cores <- detectCores() - 1
library(Uppangala)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

Built with `Uppangala` package version `r packageVersion('Uppangala')`.

```{r data prep}
PFT <- genPFT()
# Individual aggregation
PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
row.names(PFT_ind) <- PFT_ind$Group.1
SpT <- unique(PFT[1:2])
Sp <- SpT$Sp
names(Sp) <- SpT$Tree
SpCodeT <- unique(PFT[2:3])
SpCode <- SpCodeT$SpCode
names(SpCode) <- SpCodeT$Sp
CET <- unique(PFT[c(1,4)])
CE <- CET$CE
names(CE) <- CET$Tree
PFT_ind$Sp <- Sp[row.names(PFT_ind)]
PFT_ind$SpCode <- SpCode[as.character(PFT_ind$Sp)]
PFT_ind$CE <- CE[row.names(PFT_ind)]
PFT_ind <- PFT_ind[-c(1:2)]
PFT_ind$ID <- row.names(PFT_ind)
row.names(PFT_ind) <- PFT_ind$ID
PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
# Individual to log
PFT_ind$LA <- log(PFT_ind$LA)
PFT_ind$SLA <- log(PFT_ind$SLA)
PFT_ind$ID <- row.names(PFT_ind)
row.names(PFT_ind) <- PFT_ind$ID
PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
# Species aggregation
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$Sp), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$SpCode <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] == 3)
PFT_sp$strata <- NA
PFT_sp[-sel,'strata'] <- 'C'
PFT_sp[sel,'strata'] <- 'U'

rm(CET, PFT, SpCodeT, SpT, CE, sel)
```

```{r hypervolume overlap, echo=TRUE}
library(hypervolume)
set.seed(42)
n <- 99 # Random repetitions

# Data prep
data <- PFT_sp[-which(is.na(PFT_sp$WD)), c('LA', 'WD', 'LT', 'LDMC', 'SLA', 'strata')]
strata <- data$strata
data <- data[-6]
bd <- apply(PFT_ind[-which(is.na(PFT_ind$WD)), c('LA', 'WD', 'LT', 'LDMC', 'SLA')], 2, sd)

# Measurement
calcOverlap <- function(strata, data, bd){
  C <- hypervolume::hypervolume(data[strata == 'C',], bandwidth = bd, verbose = F)
  U <- hypervolume::hypervolume(data[strata == 'U',], bandwidth = bd, verbose = F)
  set <- hypervolume::hypervolume_set(C, U, check_memory = F, verbose = F)
  hypervolume::hypervolume_sorensen_overlap(set)
}
overlap <- calcOverlap(strata, data, bd)
cl <- makeCluster(cores)
clusterExport(cl, list('n', 'calcOverlap', 'strata', 'data', 'bd'))
nulloverlap <- parSapply(cl, 1:n, function(i) calcOverlap(sample(strata), data, bd))
stopCluster(cl) ; rm(cl)
```

```{r Result, echo=TRUE}
# Plot
pvalue <- (rank(c(overlap, nulloverlap))/(n+1))[1]
plot(density(nulloverlap), 
     main = paste('pvalue =', round(pvalue, 2)))
abline(v = overlap, lwd = 2, col = 'red')
```

```{r hypervolume specificity, echo=TRUE}
C <- hypervolume::hypervolume(data[strata == 'C',], bandwidth = bd, verbose = F, name = 'C')
U <- hypervolume::hypervolume(data[strata == 'U',], bandwidth = bd, verbose = F, name = 'U')
set <- hypervolume::hypervolume_set(C, U, check_memory = F, verbose = F)
Cu <- set@HVList$Unique_1
Uu <- set@HVList$Unique_2
Intersect <- set@HVList$Intersection
plot(hypervolume_join(Cu, Uu))
Spu <- list(C = data[hypervolume_inclusion_test(Cu, data),],
            Intersect = data[hypervolume_inclusion_test(Intersect, data),],
            U = data[hypervolume_inclusion_test(Uu, data),])
Spu$C <- Spu$C[-which(row.names(Spu$C) %in% row.names(Spu$Intersect)),]
Spu$U <- Spu$U[-which(row.names(Spu$U) %in% row.names(Spu$Intersect)),]
diff <- data.frame(N = unlist(lapply(Spu, function(x) dim(x)[1])),
                   Ntot = c(48, 78, 30),
                   Sp = unlist(lapply(Spu, function(x) paste0(row.names(x), collapse = ', '))))
kable(diff)
kable(do.call('rbind', lapply(Spu, function(x)apply(x, 2, mean))))
kable(do.call('rbind', lapply(Spu, function(x)apply(x, 2, var))))
```

### 2017-10-04 inst/doc/draft.Rmd
hypervolume update

@@
 ---
-title: 'Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India: vertical stratification reflects guild-dependent dynamics'
+title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation
+  rainforest of India
 author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
 date: '`r Sys.Date()`'
-output: pdf_document
+output: word_document
 csl: mee.csl
@@
 
-functional trait; rainforest; dry season; forest strata; Western Ghats; ...
+functional trait; rainforest; vertical forest strata; Western Ghats; environmental filternig; niche differentiation; ecological guild
 
@@
 Tropical rainforests have especially fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
-Many theories have been proposed to explain how species coexistence can be maintained in ecological communities .
-Among them, two have been hotly debated and tested those last years.
-First niche theory states that the niche differences of species determine wether they can establish and coexist in communities [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that species coexistence is possible even when species are functionaly equivalent, under the sole influence of stochasticity in life, death, reproduction and dispersal [@hubbell_unified_2001].
-Both theory can provide realistic expectations of taxonomic diversity [@chave_comparing_2002], and some recent approaches have advocated investigating patterns of functional [@westoby_land-plant_2006; @mcgill_rebuilding_2006] and phylogenetic diversity [@mouquet_ecophylogenetics:_2012] to trace the signature of ecological processes related to specific biological attributes and adaptations.
-These approaches have been applied to resolve the conundrum of high tree diversity in tropical forests [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012], but studies on these ecosystems still remain scarce.
-We here investigate taxonomic, functional and phylogenetic diversity in a low elevation rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], where few studies have assessed the relative influence of neutral and niche-based processes on local forest composition.
-
-The variation in functional traits among species reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007].
-These variation thus reflect fundamental trade-offs in ecological strategies along environmental gradients [@mark_westoby_plant_2002].
-For instance, specific leaf area varies along a soil fertility-shade index and represent a trade-off between resource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011].
-Due to shared ecological strategies, functional traits will covary along basic leaf [@wright_worldwide_2004 but see @osnas_global_2013 and @lloyd_photosynthetically_2013] and wood [@chave_towards_2009] economics spectra.
-The leaf economics spectrum opposes productive ecological strategy to conservative persistent strategy.
-The wood economics spectrum distinguish species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, wiht high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength [@chave_towards_2009].
-@baraloto_decoupled_2010 highlighted decoupled wood and leaf strategies allowing independant variation of these strategies, but @reich_world-wide_2014 conversely advocated that the two spectra have to integrated in a comprehensive whole plant economics spectrum [@freschet_evidence_2010] opposing fast growing species to more conservative ones.
+Niche theory explains species coexistence depending on niche differences to limit competitive exclusion [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that maintenance of high diversity is possible even for functionally equivalent species, under the influence of stochastic life, death, reproduction and dispersal dynamics [@hubbell_unified_2001].
+Both theories can provide realistic expectations of taxonomic diversity [@chave_comparing_2002] so that inferring ecological processes from taxonomical composition patternis is challenging in community ecology.
+
+Investigating patterns of functional or phylogenetic diversity may reveal signatures of ecological processes that are specific to the niche theory [@mcgill_rebuilding_2006; @westoby_land-plant_2006; @mouquet_ecophylogenetics:_2012].
+Indeed, if taxonomic composition is expected to be spatially structured in both theories, due to the spatial structure of environment or to dispersal limitation, functional or phylogenetic assemblage are expected to be spatially structured only under the niche theory [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012; @parmentier_prevalence_2014].
+We here investigate taxonomic and functional diversity in a topographically complex rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], India, where few studies have assessed the ecological drivers of local composition [@munoz2007estimating; @hardy_phylogenetic_2012,; @krishnadas2016environmental].
+
+The variation in functional traits among species is expected to reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007] and expresses fundamental trade-offs along environmental gradients [@mark_westoby_plant_2002].
+For instance, the specific leaf area varies along a soil fertility-shade gradient and represents a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@hodgson_is_2011].
+Fundamental trade-offs entail functional traits covariation along basic leaf [@wright_worldwide_2004 but see @osnas_global_2013 and @lloyd_photosynthetically_2013] and wood [@chave_towards_2009] economics spectra.
+The leaf economics spectrum opposes acquisitive to conservative ecological strategies, basically expressed at leaf level.
+**WES trade-off to clarify without CSR** The wood economics spectrum distinguishes species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength [@chave_towards_2009].
+@baraloto_decoupled_2010 found decoupled wood and leaf economics of trees in neotropical forest, but @reich_world-wide_2014 conversely posited that both spectra have to be integrated in a comprehensive whole plant economics spectrum [@freschet_evidence_2010] opposing fast growing acquisitive species to slow growing and more conservative ones.
 Here we address how varying wood and leaf ecological strategies determine species community assembly in an Indian rainforest.
 
-A major challenge in addressing community assembly is to relate local coexistence to a broader spatial context of species dynamics [@ricklefs_disintegration_2008; @munoz_neutral_2016].
-Niche-based dynamics can entail community variation along broad environmental gradients, while local community dynamics can be neutral [@leibold_coexistence_2006].
-Local community dynamics can further be influenced by niche differences under the influence of competition [@macarthur_limiting_1967].
-In this perspective, @messier_trait_2016 have underlined that the significance of the leaf economics spectrum is basically scale-dependent, which can explain conflicting results across studies performed at different scales.
-In the same logic, a spatial hierarchy of ecological dynamics can entail a hierarchy of variation in functional traits from broad variation among species along large-scale environmental gradients, to intraspecific response to micro-environmental conditions at very fine scale [@albert_when_2011].
-Therefore a multiscale framework is needed to represent the hierarchy of ecological processes from regional to local scale [@lortie_rethinking_2004].
-Determining the extent of neutral and niche-based dynamics thus implies decomposing diversity patterns within and among communities [@ackerly_trait-based_2007], and designing appropriate statistical tests to disentangle the influence of local and larger-scale dynamics [@bernard-verdier_community_2012].
-In this paper, we will decompose diversity patterns in a large rainforest plot into sources of variation within and between communities [@kraft_disentangling_2011], and investigate the fine-scale and plot-scale dynamics using appropriate null models.
-For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they challenged the existence of the leaf economics spectrum at community scale.
-
-Apart from scale-dependent dynamics in space within and among communities, ecological guilds can coexist and occupy distinct niches within communities, while the species within guilds are ecologically similar [@wilson_guilds_1999; @vergnon_niches_2009].
-The neutral theory has been basically applied in rainforests to adress diversity patterns in the guild of canopy trees [@hubbell_unified_2001], thus ignoring dynamics of understorey trees.
-However basic ecological constraints drive resource use strategies and biotic interactions in these ecosystems and imply considering functional strategies and relationship among guilds as well.
+A major challenge is to relate local coexistence in an ecological community to a broader context of species dynamics [@ricklefs_disintegration_2008; @munoz_neutral_2016].
+The trait values of coexisting species represent a non-random sample of trait values found in a broader area, du to specific niche-based dynamics in the local community [@lortie_rethinking_2004; @bernard-verdier_community_2012].
+Then the signature of local niche-based dynamics can be gasped by assessing the deviation of local community structure from a larger "species pool", classicaly using null models [@kraft_functional_2008].
+Specifically, lower variance of functional traits than null is expected when a narrow range of strategies is locally successful [functional convergence, @ackerly_trait-based_2007], while larger variance and greater eveness is expected if biotic interactions entail niche differentiation [@macarthur_limiting_1967].
+In addition, the nature of functional traits and ecological strategies driving species coexistence can change across scales.
+For instance, @messier_trait_2016 have underlined that the leaf economics spectrum basically expresses changes in ecological strategies along broad-scale environmental gradients, while it is not necessarily a primary driver of local coexistence.
+In this paper we will decompose the variation of tree functional traits in a large rainforest plot into sources of variation within communities [@kraft_disentangling_2011], using appropriate null models.
+
+The concept of ecological community basically represents the spatial limits which species share resources and interact [**cite Loojien and van Andel 1999**], while the concept of guilds refer to groups of functionally similar species [@wilson_guilds_1999].
+The two concepts are not exclusive, as ecological guilds can coexist and occupy distinct niches whithin communities [@vergnon_niches_2009].
+In tropical forests, most emphasis has been put on the dynamics of the guilds of canopy trees [@hubbell_unified_2001] **Maxime disagree, Hubbel used 1 cm gbh**, but the way disctinct guilds of canopy and understorey trees coexist in communities is little addressed.
+Distinctive ecological constraints are expected to drive different resource use strategies and biotic interactions in these guilds, which should translate into different functional composition and response to environmental factors.
 Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988].
-It thus provide an ideal context to analyze guild-dependent dynamics and ecological strategies in tropical rainforests.
-We will thus address diversity patterns and community assembly at guild-level.
-
-In this paper, we used a trait-based approach to adress within-guild community dynamics within and among local tree assemblages in a tropical rain forest from Western Ghats, India.
-Although considered as a wet evergreen forest, the forest undergoes a long dry season and an intense rainy season under the influence of monsoon.
-In this context, we expect water stress and competition for resource to be the key drivers of forest dynamics.
-We thus adressed two main questions:
-
-1. Do variations and covariations in functional traits reflect distinct ecological strategies within and among species ?
-    i. How functional traits vary among species ?
-    ii. Do covariations in functional traits reflect axes of ecological strategy ?
-2. Does functional composition depends on potential local abiotic and biotic drivers of community dynamics ?
-    i. Do local functional composition ($\alpha$ scale) reflect the influence of abiotic and biotic drivers of community dynamics ?
-    ii. Do functional composition variation between communities ($\beta$ scale) reflect the changing of abiotic and biotic conditions ?
+It thus provides an appropriate context to adress guild-dependent dynamics and ecological strategies in tropical rainforests.
+We thus investigate here the diversity patterns and community assembly of rainforest trees at guild-level.
+
+In this paper, we used a trait-based approach to address guild dynamics whithn local tree assembly along topographical gradients in a tropical rainforest from the Western Ghats, India.
+Although considered as wet evergreen, the forest undergoes contrasted long dry and intense rainy seasons under the influence of monsoon.
+In this context, we expected water stress and competition for resource to be primary drivers of forest dynamics and functional composition.
+A large forest plot (10ha) was subdivided in local elementary communities and we investigated how guild composition whithin communities varied with local environmental conditions.
+We addressed the following questions:
+
+1. About the variation and covariation in functional traits within and among species,
+    i. Does functional traits variation allow identifying distinctive ecological strategies among species ?
+    ii. Does their covariations reflect basic dimensions of ecological strategy in the rainforest ?
+2. About the role of functional trait variation in guild and community assembly,
+    iii. Does the local functional composition $\alpha$ reflect the influence of guild and community dynamics ?
 
@@
 The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry in an undisturbed, old-growth wet evergreen Dipterocarp forest [@pascal_structure_1996].
-The plot is located near Uppangala village, and thus refers to as "Uppangala plot", in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India).
+The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), near the Uppangala village, and thus refers to as "Uppangala plot".
 Annual precipitation averages 5108 mm per year but is highly variable in time with heavy rainfall from June to October (90% of the rainfall during the monsoon) and a severe dry season (monthly rainfall $\leq 45$ mm) lasting 3.8 months [@pelissier_twenty_2011].
 Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996].
-Climate and pedology of the 10-ha plot can be considered as homogeneous.
-The topography of the site is highly heterogeneous with a regular east-west alternation between thalwegs and ridges separated by steep slopes, so that topography can be assumed as a major source of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998].
-We hypothesized that topographic variation should influence community variation depending on tree ecological strategies [@condit_spatial_2000].
-
-## Floristic and trait data
-
-Since 1989, all trees $\geq 30$ cm in girth at breast height (gbh) have been monitored in an area progressively extended to a contiguous 10-ha plot up to 2014.
-<!-- Need to add Cedric Vega in Acknowledgments for that.
---> All trees were identified, tagged and mapped over the plot [@pelissier_twenty_2011].
-111 different species were recorded.
-We collected trait data on 89 over 111 (80.1%) species, which represented 99.0% of the tree individuals present in the Uppangala plot.
-Note that the 32 most abundant species, represent  more than 80% of Uppangala tree individuals.
-We sampled five individuals per species for these 32 species, and sampled one to five individuals from the remaining species (in total 267 individuals were sampled).
-For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation from @perez-harguindeguy_new_2013.
+Climate and geology of the 10-ha plot can be considered as homogeneous.
+The topography of the site is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes, so that it is expected to be a major component of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998].
+Therefore, we hypothesized that topographic variation should influence community assembly and yield changing ecological strategies and functional composition [@condit_spatial_2000].
+
+## Taxonomic composition and trait sampling
+
+From 1989 to 2014, the trees $\geq 30$ cm in girth at breast height (gbh) were identified, tagged, mapped and monitored in an area progressively extended to a contiguous $300 * 330 m^2$ (ca.
+10-ha) plot [@pelissier_twenty_2011].
+The plot included 111 different tree species.
+We collected trait data from 89 over 111 (80.1%) species, which represented 99.0% of tree individuals present in the plot.
+Note that the 32 most abundant species alone, represented more than 80% of tree individuals.
+We sampled five individuals per species for these 32 species, and sampled one to five individuals for the remaining species (in total 267 individuals).
+For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation of @perez-harguindeguy_new_2013.
 Individuals and leaves were selected in consistent light conditions, development stage and sanitary state.
-Dawkins crown exposure index [@dawkins_management_1958] was recorder for each tree sampled.
-Leaf and branch samples were kept in humidified ziplock filled with enriched $CO_{2}$ air inside a black bag untill their measurement in the following 24 hours.
-For some individuals, we were unable to collect an appropriate branch sample (11 missing species over 89 representing 1% of individuals present in the 10 ha plot).
+Dawkins crown exposure index [@dawkins_management_1958] was recorded for each tree sampled.
+Leaf and branch samples were kept in humidified ziplock filled with enriched $CO_{2}$ air inside a black bag until measurement within the next 24 hours.
+For some individuals, we were unable to collect suitable branch sample (11 missing species over 89 representing 1% of individuals present).
 
@@
 Leaf petioles, rachis and petiolules were removed for all measurements.
-Leaf fresh weights, area and thickness were measured, using a weighing scale, a bend scanner and a micrometer respectively.
+Leaf fresh weights, thickness and area were measured, using a weighing scale (**add detail on brand and precision**), a micrometer (**add detail on brand and precision**) and a bend scanner with ImageJ software [**ref**] respectivelly.
 Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$.
@@
 The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues.
-Branch wood samples were sink in water to evaluate their volume by water displacement.
-Samples were then dried for 48 hours at 80 $^\circ C$ and weighted.
+Branch wood samples were immersed in water to evaluate their volume by water displacement.
+Wood samples were then dried for 48 hours at 80 $^\circ C$ and weighted.
 
 We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1).
-We choose those five traits because they have been widely used, in addition of there easier measurement, and they have been shown to be good indicator of species abilities to grow, survive and reproduce [@wilson_specific_1999].
-
-Table 1: Functional traits measured. *Traits with their abbreviation, standard unit, and associated species strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*
+We choose those five traits because they proved to be good proxies of the major dimensions of ecological strategies in plants [@mark_westoby_plant_2002; @diaz_global_2015] and were reatively easy to measure on the field.
+
+Table 1: Functional traits measured. *Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].*
 
@@
   Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
-  Strategy = c('Leaf defense vs. investment',
+  Strategy = c('Leaf defense and investment',
         'Leaf resource capture vs.investment',
-        'Leaf defense vs. investment',
+        'Leaf defense and investment',
         'Leaf resource acquisiton vs. defense',
@@
 
-Because topography was assumed as a major source of environmental variation [@gimaret-carpentier_sampling_1998], we selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed using ArcGis [@desktop_arcgis_2011].
-
-In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the environmental variables (see supplementary materials S1).
-We selected the less correlated variables having the most ecological meaning.
-Thus we reduced our explanatory variables to 4 variables: curvature representing the water flow paths and soil accumulation areas, wetness highlighting water accumulation areas, southwesterness representing aspect [@bunyan_effect_2015] and slope, which may have both a strong effect on forest dynamics and on mutrient/water availability **ref**.
-
-We selected two biotic factors related to local forest structure: canopy height and basal area.
+Topography was expected to be a major environmental factor influencing the assemby of tree communities [@gimaret-carpentier_sampling_1998].
+We selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed in ArcGis [**ref**].
+
+In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the abiotic environmental variables (see Supplementary Materials S1).
+We selected four less correlated variables: curvature representing local water flow paths and soil accumulation areas, wetness highlighting water accumulation areas over the whole study area, southwesterness representing aspect [@bunyan_effect_2015] and slope, which may have both a strong effect on forest dynamics and on nutrient/water availability **ref**.
+
+In addition, we selected two biotic factors related to local forest structure: canopy height and basal area.
 Canopy height, i.e.
-the maximum local height of the canopy, was calculated from a digital canopy model built from a LiDAR (Light Detection And Ranging) campaign done in 2013 over the Uppangala plot with a resolution of $1*1 m^2$.
-Basal area was calculated from 2013 and 2014 girth inventories of Uppangala plot on stems with a diameter above 30 centimeters at breast height ($DBH>30~cm$).
+the maximum local height of the canopy, was calculated from a 1-m resolution digital canopy model built from a LiDAR (Light Detection And Ranging) dataset acquired in 2013 over the Uppangala plot.
+Basal area was calculated the girth measurements taken within the 10-ha plot (census done in 2013 and 2014) for stems $>30~cm$ gbh.
 
@@
 
-To quantify trait variation whithin and between individuals and between species we calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values between leaves of the same individual (intra-individual), between trees of the same species (intra-specific for species with more than one individual), and between species (inter-specific).
-In addition, to assess the covariation of trait within species, and thus potential axes of strategy, we performed a principal component analysis (PCA) of the mean value of traits per species.
-Subsequent community-level analysis were based on trait values averaged per species.
-
-### Community trait variability
-
-Across the whole 10 hectare plot, $30*30 m^2$ quadrats were defined and represented 110 local communities.
-@pascal_wet_1988 defined four categories of strata groups: emergent below 39m, canopy tree between 25 and 39m, sub-canopy tree between 15 and 25m, and understorey tree below 15m.
-Due to topography, sub-canopy tree from Uppangala are often under light exposure and participates to canopy, especially in slopes.
-We defined two guilds to represent vertical stratification called canopy (C), composed of emergent, canopy and sub-canopy tree, and understorey (U) trees.
-The 110 local communities where split by those two guilds.
-
-To characterize the local functional composition of canopy and understorey communities we used the community weighted mean (CWM), the community weighted variance (CWV), and the coefficient of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010].
-CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $i$ species in the quadrat $j$, with each species, $i$, weighted by its relative abundance, $p_{i}$:
-$$CWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
-We also calculated the average of trait values based on species presence-absence, to further emphasize the contribution of less abundant and rare species in community trait composition [@cingolani_filtering_2007].
-CWV for each trait $t$ was calculated as the mean distance of trait values, $t_{i}$, to $CWM$, of the $S$ species in the quadrat $j$, with each species, $i$, weighted by its relative
-abundance, $p_{i}$:
-$$CWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - CWM_{j})^2 ~ (2)$$
-CVNND for each trait $t$, was calculated as the standard deviation of the nearest neighbor distances (NND), $\sigma _{NND}$, of the $i$ individuals in each community, divided by the mean NND, $\mu _{NND}$; and NND for one individual $i$ is calculated as the minimal distance $d$ between the $j$ other individuals of the community :
-$$CVNND = \frac{\sigma _{NND}}{\mu _{NND}} ~ , ~ ~ ~ NND = {NND_{i} = min(d_{i,j})_{i \neq j}} ~ (3)$$
-
-To investigate functional composition of communities we used a null model approach.
+We calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among leaves of the same individual (intra-individual), among trees of the same species (intra-specific for species with more than one individual), and among species (inter-specific).
+We performed a principal component analysis (PCA) of the mean value of traits per species to identify main dimensions of ecological strategies along which traits can covary.
+Subsequent community-level analyses were based on trait values averaged per species.
+
+### Guild-level analyses
+
+$30*30 m^2$ quadrats were delimited in the 10 heactare plot and represented 110 local communities.
+@pascal_wet_1988 defined four vertical strata attained by adult trees: emergent above 39m, canopy trees between 25 and 39m, sub-canopy trees between 15 and 25m, and understorey trees below 15m.
+Despite the vertical hierarchy, steep topography could allow sub-canopy trees receiving much light, expecially in slopes, and thus participates to canopy.
+We defined two basic guilds called canopy (C), composed of emergent, canopy and sub-canopy species, and understorey (U) species.
+We analyze trait composition of these guilds in the 110 communities.
+
+<!-- Metrics -->
+We calculated weighted mean (GWM, where G stands for "guild"), weighted variance (GWV), and coefficient of variation of nearest neighbour distance (CVNND) [GWM and GWV, @violle_let_2007; CVNND, @jung_intraspecific_2010] for each guild in each community.
+In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of $i$ species weighted by species relative abundance $p_{i}$:
+$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
+where $S$ is the number of species in the quadrat.
+
+We also calculated the average trait values based on species presence-absence, to addressthe influence less abundant and rare species on community trait composition [@cingolani_filtering_2007].
+GWV of trait $t$ was calculated as :
+$$GWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - GWM_{j})^2 ~ (2)$$
+
+CVNND of trait $t$ was calculated as the standard deviation of nearest neighbor distances (NND) $\sigma _{NND}$ divided by the mean NND $\mu _{NND}$; and NND was the minimal distance $d$ between a species and the most similar other species of the guild :
+$$CVNND = \frac{\sigma _{NND}}{\mu _{NND}} ~ , ~ ~ ~ NND = [NND_{i} = min(d_{i,j})_{i \neq j}] ~ (3)$$
+
+<!-- Null models -->
+To assess significant changes in functional composition among guilds, we devised a null model apporoach.
 A first null model (NM1) randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix).
-The second model (NM2) randomized abundances among species within each community but kept the observed list of species and trait values inside communities [@bernard-verdier_community_2012].
-Linear regressions between CWMs and biotic and abiotic factors ($CWM_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$), after stepwise selection, were compared to corresponding 999 null linear regressions from NM1 with F-statistic.
-For both CWV and CVNND, we performed a two-tailed test of the proportion of of communities differing from null expectation with adjusted p-values by Benjamini-Hochberg correction.
-CWV lower than expected under a null model represented convergence of trait values within the community, whereas CWV higher than null evidenced divergence in trait distribution [@bernard-verdier_community_2012].
-CVNND was used to detect niche differentiation among coexisting species, which occured when the observed CVNND was lower than under a null model [@jung_intraspecific_2010].
-We then calculated the standardized effect size (SES) as: $$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
-where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.
+The second model (NM2) randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012].
+**Spatial correlations in tree abundance are conserved limiting associated type I error.** GWM and CVNND metrics were compared to null model NM1 whereas GWV was studied with null model NM2 **To develop**.
+
+<!-- Analysis -->
+We performed 999 randomizations for each model and we derived a two-tailed test of the proportion of guilds deviating from the null expectation, with adjusted p-values by Benjamini-Hochberg correction **Citation**.
+Shifts in GWM values with changing environmental conditions could reflect the shorting of species strategies across assemblages [@ackerly_trait-based_2007].
+GWV lower/higher than null represented convergence/divergence of trait values within the guild [@bernard-verdier_community_2012].
+CVNND was used to detect niche differentiation among coexisting species, which occurred when observed CVNND was lower than null [@jung_intraspecific_2010].
+
+To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
+where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of the corresponding null distribution.
+
 The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$).
@@
 
-### Taxonomic and functional variation among communities
-
-To investigate taxonomic and functional variation among communities we calculated $I_{st}$, for taxonomic composition and $U_{st}$ and $\tau _{st}$, for functional composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
-$I_{st}$ measures species turnover accounting for local abundance.
-$U_{st}$ integrate only functional turnover, weighting species by their local abundances; whereas $\tau _{st}$ express functional turnover, without accounting for species local abundances.
-
-Variation in composition between community pairs with those metrics was translated as a taxonomic, functional and phylogenetic distance, and compared to the corresponding environmental and spatial distances.
-Environmental distances were calculated as euclidean distances between communities on the plan composed by the two first axes of environmental variables PCA (see supplementary material S1).
-Correlations with environment and space were assessed with two Mantel tests successively with taxonomic and functional, community composition.
-The firt Mantel test consisted in assessing the correlation between each metric distance matrix and the environmental distance matrix (E) and the second test, a partial Mantel test, assessed the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S).
-This second test thus test the remaining spatial dependency in community composition once the influence of environmental variation is removed.
-
-We used null models to test if the observed correlations differed significantly from expectations under null hypothesis.
-$I_{st}$ correlations with environment and space were tested with 999 repetions of the nul model described by @myers_beta-diversity_2013 which randomizes individuals among quadrat while preserving the regional abundance of species, and the number of individuals per community.
-$U st$ and $\tau st$ were tested by the nul model NM1 which randomizes traits among species.
-
-All analyses were conducted using the R 3.3.0  software environment [@r_core_team_r:_2016] with packages 
+All analyses were conducted using the R 3.3.0  software [@r_core_team_r:_2016] with packages 
 `ade4` [@dray_ade4_2007],
-`ape` [@paradis_ape:_2004], 
-`phylosignal` [@keck_phylosignal:_2015], 
-`spacodiR` [@eastman_spacodir:_2013],
 and `vegan` [@oksanen_vegan:_2016].
@@
 
+**General statistics on guilds: size, richness...**
+
 ## Traits variation and covariation
 
-We found a general increase of functional trait variation from intra-individual to inter-specific level (see table 2).
-Whereas LA shows already high variabililty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
-Finally the relatively low intra-individual variation ($cv_{intra}$ range from 0.06 to 0.09) allow us to consider the mean trait values per species for subsequent connumity-level analysis.
-
-Table 2: Coefficient of variation of intra-individual, intra-specific and inter-specific trait values.
-*Traits with their range of values, mean and standard deviation, mean of coefficient of variation for leaves in all individuals (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
+We found overall increase of functional trait variation from intra-individual to inter-specific level (see table 2), even though LA showed almost as large variability at intra-individual than at intra-specific level ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
+Relatively low intra-individual variation ($cv_{intra}$ range from 0.06 to 0.09) allowed us considering mean trait values per species in subsequent analyses of the trait composition of guilds.
+
+Table 2: Intra-individual, intra-specific and inter-specific component of trait variation.
+*The overall range, mean and standard deviation are provided for each trait.
+The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species.
 See table 1 for abbreviation.*
@@
 
-The principal component analysis (PCA) revealed a first axis along which both wood density and structural leaf traits (LT, LDMC) correlate, thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood.
-Interestingly, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between these two variables.
-The second axis of the PCA2 covaried mainly positively with SLA and negatively with LDMC.
-PCA also highlight the strong correlation between LDMC and WD (Pearson's correlation test, $r = 0.25$ with $p-value < 0.001$).
+The first axis of the principal component  analysis (PCA) of mean species trait values was basically related to wood density and structural ela trait variations (LT, LA, LDMC), thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood.
+The second axis was mainly related positively to SLA and negatively to LDMC.
+Wood density was othogonal to SLA in the plan 1-2 of the PCA, while positive correlation was found between LDMC and WD (Pearson's correlation test, $r = 0.25$ with $p-value < 0.001$).
 
@@
 
-# Indovidual to log
+# Individual to log
 PFT_ind$LA <- log(PFT_ind$LA)
@@
 lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
-# legend('topright', c('Leaf economics spectrum', 'Wood economics spectrum'),
-       # text.col = c("green", "firebrick"))
+legend('topright', c('Leaf economics spectrum', 'Wood economics spectrum'),
+       text.col = c("green", "firebrick"))
 
@@
           cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(pca.var)), col = 'red') ;
+  # lines(b, bstick(length(pca.var)), col = 'red') ;
   legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
@@
 
-Figure 1: Principal Component Analysis (PCA) performed on species mean trait values.
-*The bottom right subplot represents the eigen values (grey bars) compare to a broken stick distribution for five components (red line).
+Figure 1: Principal Component Analysis (PCA) of species mean trait values.
+*The bottom right subplot represents the eigen values (grey bars) of the five components.
+First axis, in red, is named wood economic spectrum (WES) ; whereas second axis, in green, is named leaf economic spectrum (LES).
 See table 1 for abbreviation.*
@@
 com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
+Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/share/applications')
 names(Deriv)[5] <- 'SW'
@@
 
-## Functional composition of local communities
-
-Community weighted means (CWM), both in abundance or presence-absence, showed some significant variations along environmental gradients (see table 3).
-Community mean values of WD significantly increased under south-west direction, both in abundance and presence-absence CWM,  whereas SLA and LA, in presence-absence only, decreased with curvature.
-In addition, LDMC in presence-absence was decreasing with canopy height.
-Trait values that allow species existence is not always the same that trait values determining their ability to dominate a given environment.
-For instance here both abundant and less abundant species increased their WD under southwest orientation, whereas SLA and LA decrease with curvature is only affecting less abundant species ability to survive in concave environments.
-
-Table 3: Summary of linear models of traits standardized effect size (SES) of CWMs.
-_Linear regression results for each trait SES of CWMs, for canopy (C) and understorey (U) guilds, in abundance (AB) and presence-absence (PA), after stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
+## Functional composition of guilds
+
+We found significant variation along the environmental gradients of the guild weighted means (GWM), both in abundance or presence-absence, and for both canopy (C) and understorey (U) guilds (see table 3 and supplementary materials S2).
+Functional traits of both canopy and understorey tree species were basically influenced by curvature, but only functional traits from canopy tree species were influenced by south-west orientation and mainly the functional traits of understorey tree species were influenced by canopy height.
+
+Mean value of WD, repsectively WES, in canopy tree species significantly increased, resp.
+decreased, under south-west orientation, both in abundance and presence-absence.
+GWM of LA and LT decreased under SW, in abundance and presence absence respectively.
+In addition, LA decreased with curvature in both abundance and presence-absence, while WD, resp.
+WES and LDMC, increased, resp.
+decreased, with curvature in presence-absence only.
+
+Mean value of LES and SLA significantly decreased in understorey tree species when those of WD and LT significantly increased with curvature, both in abundance and presence-absence.
+WES of understorey tree species also increased and LA decreased with curvature in abundance.
+In addition, LES and SLA in both abundance and presence-absence were increased with canopy height, while LDMC and WD decreased with canopy height in abundance.
+
+Trait values allowing species persistence are not always the same than those determining species dominance in a given environment.
+For instance here both abundant and less abundant canopy tree species had greater WD under southwest orientation, while higher mean WD was found in concave environment only in presence-absence, meaning that other factors could influence the success of more abundant species.
+
+Table 3: Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+_We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
 See table 1 for abbreviations.
@@
 })
-# CWM.ses.res <- lapply(CWM.ses.res, function(s){
-#   lapply(s, function(ab){
-#     sel <- which(unlist(lapply(as.list(data.frame(t(ab))), function(t){
-#   any(grepl('*', t, fixed = T))
-# })))
-#   return(ab[sel,])
-#   })
-# })
-CWM.ses.res <- mapply(function(s, name){
-  lapply(s, function(ab){
+CWM.ses.res <- lapply(CWM.ses.res, function(s){
+  mapply(function(ab, name){
       ab <- cbind(c(name, rep(' ', length(ab[,1])-1)), ab)
@@
       return(ab)
-  })
-}, s = CWM.ses.res, name = as.list(c('AB', 'PA')), SIMPLIFY = F)
+  },ab = s, name = as.list(c('AB', 'PA')), SIMPLIFY = F)
+})
 CWM.ses.res <- lapply(CWM.ses.res, function(s){
@@
 CWM.ses.res <- do.call('rbind', CWM.ses.res)
-kable(bold(CWM.ses.res), row.names = F)
-rm(nCWM, CWM.ses, CWM.ses.lm, CWM.ses.res)
+rm(nCWM, CWM.ses.lm)
 invisible(gc())
@@
 
-Traits convergence/divergence was measured with community weighted variances (CWV).
-We could found significant relationships between standardized effect size (SES) of CWVs and environmental variables (see table 4).
-Convergence of WES and LDMC in the canopy guild  and convergence of LA in the undersotrey guild were found under concave areas.
-LDMC of the canopy guild converged also with canopy height.
-LES of understorey guild converged in slopes and SLA of understorey guild diverged with increasing BA.
-We also found significant relationships of their standardized effect size (SES) of CVNNDs with environmental variables (see table 4).
-Niche differentiation of WES and LDMC in canopy guild increased in lower basal area context.
-Differentiation of canopy guild LA increased in convex and high canopy contexts, whereas LT of the canopy guild differentiated in concave context.
-Niche differentiation of LES in understorey guild increased in concave areas, while differentiation of LT of the understorey guild increased in wet context with a high basal area.
-
-Table 4: Summary of linear models of traits standardized effect size (SES) of CWVs and CVNNDs.
-_Linear regression results for each trait SES of CWVs and CVNNDs, for canopy (C) and understorey (U) guilds, after stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA).
+```{r CWM AB}
+kable(bold(CWM.ses.res[c(1:7,15:21),-2]), row.names = F)
+```
+
+```{r WD}
+data <- cbind(CWM.ses$C$abundance$WD, R$C)
+names(data)[1] <- 'WD'
+par(mfrow = c(3,2)) # Affichage des relations individuelles
+plot(WD ~ Slope, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
+plot(WD ~ Curvature, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
+plot(WD ~ Wetness, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
+plot(WD ~ SW, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
+plot(WD ~ Canopy, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
+plot(WD ~ BA, data) ; abline(h = c(0.95, 0, -0.95), lty = c(2, 1, 2))
+rm(data, CWM.ses)
+```
+
+Figure 2: Exemple of standard effect size (SES) of wood density  (WD) for canopy (C) tree species.
+*Each subplot show  standard effect size (SES) of wood density (WD) for canopy (C) tree species along one environmental variable gradient.
+Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
+See table 1 for abbreviation.*
+
+We assessed trait convergence/divergence based on the deviation of guild weighted variance (GWV) from null model NM2.
+We found significant relationships between the standardized effect size (SES) of GWVs and the environmental predictors (see table 4).
+Convergence of WES and LDMC in the canopy guild  and convergence of LA in the understorey guild occured in convex areas.
+LDMC values in the canopy guild converge more in taller canopy.
+We also found significant relationships of standardized effect sizes (SES) of CVNNDs with environmental predictors (see table 4).
+Niche differentiation (lower CVNND) of WES and LDMC in canopy tree species increased with lower basal area.
+Differentiation of LA in canopy tree species increased in convex context and taller canopy, whereas LT of canopy tree species showed higher differentiation in concave context.
+Niche differentiation of LES in understorey tree species increased in concave context, while LT of understorey tree species showed higher differentiation in wet context with a high basal area.
+
+Table 4: Linear models of the standardized effect sizes (SES) of GWV and CVNND, for canopy (C) and understorey (U) guilds.
+_We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
 See table 1 for abbreviations.
@@
 })
-# CWV.ses.res <- lapply(CWV.ses.res, function(s){
-#   sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
-#   any(grepl('*', t, fixed = T))
-# })))
-#   return(s[sel,])
-# })
 CWV.ses.res <- mapply(function(s, name){
@@
 })
-# CVNND.ses.res <- lapply(CVNND.ses.res, function(s){
-#   sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
-#   any(grepl('*', t, fixed = T))
-# })))
-#   return(s[sel,])
-# })
 CVNND.ses.res <- mapply(function(s, name){
@@
 
-## Taxonomic and functional varitions among communities
-
-Taxonomic variation among communities was correlated significantly with environmental variation for both canopy and understorey guilds (Mantel test for canopy and understorey guild, $r_{I_{st},E,C} = 0.22$ and $r_{I_{st},E,U} = 0.23$, respectively, with $p-value < 0.01$), but not with remaining spatial dependency in community composition for canopy guild (Mantel test of space knowing environment, $r_{I_{st},S|E,C} = 0.01$ with $p-value > 0.1$), whereas understorey guild taaxonomic variation was correlated to remaining spatial dependency in community composition (Mantel test of space knowing environment, $r_{I_{st},S|E,U} = 0.11$ with $p-value < 0.01$).
-
-```{r Ist, include=FALSE}
-library(spacodiR)
-
-# Environmental matrices
-XY <- data.frame(as(quadrats(c(30,30)), 'SpatialPoints')@coords)
-row.names(XY) <- paste0('C',row.names(XY))
-XY <- lapply(L, function(s){
-  XY[row.names(s$abundance),]
-})
-Ms <- lapply(XY, function(s){
-  dist(s)
-})
-XY <- pca.env$scores[,1:2]
-row.names(XY) <- paste0('C',row.names(XY))
-XY <- lapply(L, function(s){
-  XY[row.names(s$abundance),]
-})
-Me <- lapply(XY, function(s){
-  dist(s)
-})
-rm(XY)
-
-# Observed Ist
-spaflo <- lapply(L, function(s){
-  spacodi.calc(sp.plot = t(s$abundance), pairwise = T)$'pairwise.Ist'
-})
-## Mantel values
-spaflo.mantel <- mapply(function(s, Mes, Mss){
-  list(
-    `E` = vegan::mantel(s, Mes, permutations = 0)$statistic, 
-    `S | E`= vegan::mantel.partial(s, Mss, Mes, permutations = 0)$statistic
-  )
-}, s = spaflo, Mes = Me, Mss = Ms, SIMPLIFY = F)
-
-# Simulated mantel
-nL <- lapply(L, function(s){
-  vegan::permatfull(s$abundance, fixedmar = 'both', mtype = 'count', times = n, shuffle = 'both')
-})
-nspaflo <- lapply(nL, function(s){
-  lapply(s$perm, function(l){
-    spacodi.calc(sp.plot = t(l), pairwise = T)$'pairwise.Ist'
-  })
-})
-nspaflo.mantel <- mapply(function(s, Mes, Mss){
-  lapply(s, function(x){
-    try(list(
-      `E` = vegan::mantel(x, Mes, permutations = 0)$statistic, 
-      `S | E`= vegan::mantel.partial(x, Mss, Mes, permutations = 0)$statistic
-    ))
-  })
-}, s = nspaflo, Mes = Me, Mss = Ms, SIMPLIFY = F)
-sel <- which(unlist(lapply(nspaflo.mantel$U, inherits, 'try-error')))
-nspaflo.mantel$U <- nspaflo.mantel$U[-sel]
-## Non null mantels
-nspaflo.mantel <- lapply(nspaflo.mantel, function(s){
-  lapply(s, `[`, names(s[[1]]))
-  apply(do.call(rbind, s), 2, as.list)
-})
-nspaflo.mantel <- lapply(nspaflo.mantel, function(s){
-  lapply(s, unlist)
-})
-spaflo.res <- mapply(function(s, ns){
-  mapply(function(x, nx){
-    paste(format(x, scientific = T, digits = 3),
-          Uppangala::stars(min(1-rank(c(x,nx))[1]/(n+1),rank(c(x,nx))[1]/(n+1)), ns = '   '))
-  }, x =s, nx = ns)
-}, s = spaflo.mantel, ns = nspaflo.mantel)
-
-# Result
-kable(bold(spaflo.res))
-rm(spaflo.mantel, nspaflo, nspaflo.mantel, spaflo.res)
-```
-
-Most of trait functional variation among communities was neither correlated to environment nor to spactial distance knowing the environment effect (see table 5).
-However, the variation of WD, in abundance as well as in presence absence, was positively correlated to environmental variation and is negatively correlated in presence absence to the spatial structure knowing environment spatial effect.
-
-Table 5: Summary of Mantel tests between functional composition, environment and spatial distances between communities.
-_$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to measure functional composition variation among communities, for canopy (C) and understorey (U) guilds.
-Observed values were compared to null values obtained with model NM1 randomizing traits among species (permuting raw in `Trait*Species` matrix).
-E is the Mantel test of this metric according to the environmental distance derived from communtiy scores in the environmental PCA (two first axes).
-S|E is the partial Mantel test of this metric according to spatial distance conditionally to environment variation.
-The number of asterisk indicates significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-```{r Ust & TAUst}
-# Observed Hardys
-spafun <- mapply(function(s, Ls){
-  lapply(as.list(1:length(s)), function(i){
-    if (length(which(is.na(s[i]))) > 0) {
-      sel <- intersect(names(Ls$abundance), row.names(s[-which(is.na(s[i])),]))
-      return(spacodi.calc(sp.plot = t(Ls$abundance[sel]), sp.traits = data.frame(s[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
-    } else {
-      return(spacodi.calc(sp.plot = t(Ls$abundance), sp.traits = s, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
-    }
-  })
-}, s = Q, Ls = L, SIMPLIFY = F)
-spafun <- lapply(spafun, function(s){
-  names(s) <- names(Q$C)
-  return(s)
-})
-## Mantel values
-spafun.mantel <- mapply(function(s, Mes, Mss){
-  lapply(s, function(t){
-    list(
-      `E` = lapply(t, function(M){
-        vegan::mantel(M, Mes, permutations = 0)$statistic
-      }), 
-      `S | E` = lapply(t, function(M){
-        vegan::mantel.partial(M, Mss, Mes, permutations = 0)$statistic
-      })
-    )
-  })
-}, s = spafun, Mes = Me, Mss = Ms, SIMPLIFY = F)
-
-# Simulated mantel
-cl <- makeCluster(cores)
-clusterExport(cl, list('nQ', 'L', 'Me', 'Ms', 'spacodi.calc'))
-nspafun.mantel <- clusterMap(cl, function(nQs, Ls, Mes, Mss){ 
-  lapply(nQs, function(q){
-    s <- lapply(as.list(1:length(q)), function(i){
-      if (length(which(is.na(q[i]))) > 0) {
-        sel <- intersect(names(Ls$abundance), row.names(q[-which(is.na(q[i])),]))
-        return(spacodi.calc(sp.plot = t(Ls$abundance[sel]), sp.traits = data.frame(q[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
-      } else {
-        return(spacodi.calc(sp.plot = t(Ls$abundance), sp.traits = q, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
-      }
-    })
-    names(s) <- names(q)
-    res <- list(
-      `E` = lapply(s, function(t){
-        lapply(t, function(M){
-          try(vegan::mantel(M, Mes, permutations = 0)$statistic)
-        })
-      }),
-      `S | E` = lapply(s, function(t){
-        lapply(t, function(M){
-          try(vegan::mantel.partial(M, Mss, Mes, permutations = 0)$statistic)
-        })
-      })
-    )
-    res <- lapply(res, function(r){names(r) <- names(q) ; return(r)})
-    res <- lapply(res, `[`, names(res[[1]]))
-    res <- apply(do.call(rbind, res), 2, as.list)
-    return(res)
-  })
-}, nQs = nQ, Ls = L, Mes = Me, Mss = Ms, SIMPLIFY = F)
-stopCluster(cl)
-rm(cl, nQ) ; invisible(gc())
-
-## Non null mantels
-nspafun.mantel <- lapply(nspafun.mantel, function(s){
-  lapply(s, `[`, names(s[[1]]))
-  apply(do.call(rbind, s), 2, as.list)
-})
-nspafun.mantel <- lapply(nspafun.mantel, function(s){
-  lapply(s, function(m){
-    lapply(m, `[`, names(m[[1]]))
-    apply(do.call(rbind, m), 2, as.list)
-  })
-})
-nspafun.mantel <- lapply(nspafun.mantel, function(s){
-  lapply(s, function(m){
-    lapply(m, function(u){
-      lapply(u, `[`, names(u[[1]]))
-      apply(do.call(rbind, u), 2, as.list)
-    })
-  })
-})
-nspafun.mantel <- lapply(nspafun.mantel, function(s){
-  lapply(s, function(m){
-    lapply(m, function(u){
-      lapply(u, unlist)
-    })
-  })
-})
-
-spafun.res <- mapply(function(s, ns){
-  mapply(function(t, nt){
-    mapply(function(m, nm){
-      mapply(function(u, nu){
-        paste(round(u, 3), Uppangala::stars(min(rank(c(u,nu))[1]/(n+1), 1-rank(c(u,nu))[1]/(n+1)), ns = '   '))
-      }, u = m, nu = nm)
-    }, m = t, nm = nt)
-  }, t = s, nt = ns, SIMPLIFY = F)
-}, s = spafun.mantel, ns = nspafun.mantel, SIMPLIFY = F)
-trait <- names(spafun.res$C)
-spafun.res <- lapply(spafun.res, function(s){
-  mapply(function(t, tr){
-    t <- data.frame(t)
-    t <- cbind(c(tr, ' '), c('$U_{st}$', '$\\tau _{st}$'),  t)
-    names(t) <- c('T', 'M', 'E', 'S|E')
-    return(t)
-  }, t = s, tr = trait, SIMPLIFY = F)
-})
-spafun.res <-lapply(spafun.res, function(s){
-  s <- do.call('rbind', s)
-})
-spafun.res <- mapply(function(s, str){
-  s <- cbind(c(str, rep(' ', dim(s)[1]-1)), s)
-  names(s)[1] <- 'S'
-  return(s)
-}, s = spafun.res, str = c('C', 'U'), SIMPLIFY = F)
-spafun.res <- do.call('rbind', spafun.res)
-
-# Result
-kable(bold(spafun.res), row.names = F)
-rm(spafun.mantel, nspafun.mantel, spafun.res)
-```
-
 # Discussion
 
-## Ecological strategies
-
-We found increasing trait variation within individuals, between individuals within species, and between speces (table 2).
-It entailed that the primary source of trait variation within and among communities should be among species.
-The PCA of functional trait variation among species displayed two main dimensions (figure 1), which may be interpreted as independent wood and leaf economics spectra.
-We found opposing SLA to LDMC values along the leaf economics spectrum, which expressed a trade-off between productivity, with less carbon investment for higher photosynthesis rates, and persistence, with more leaf defense and investment [@wright_worldwide_2004].
+**"Take home messages" to add**
+
+## Wood and leaf economics hold at the local scale
+
+We found an increasing trait variation within individuals, within species, and between species (table 2).
+Our analyses thus reveal that the primary source of trait variation within and among communities is the inter-species variation, justifying the use of species mean traits for the study of tropical tree guild assembly.
+The PCA of functional trait variation among species displayed two main dimensions (figure 1), very similar to the ones found by @baraloto_decoupled_2010 and @fortunel_leaf_2012.
+those authors interpreted these axes as independent wood and leaf economics spectra.
+We found that SLA and LDMC tend to covary negatively along the leaf economics spectrum, which be seen as a trade-off between productivity, with less carbon investment and higher photosynthesis rates, and persistence, with more leaf defense and investment [@wright_worldwide_2004].
 A second major axis of trait variation opposed wood density to leaf area and thickness.
-It may thus be interpreted as an axis opposing stress tolerant species, with denser wood, to competitive species, with thicker and larger leaves, which confer a higher efficiency in water conductivity.
+It may be interpreted as an axis opposing stress tolerant species, with denser wood, to competitive species, with thicker and larger leaves, conferring a higher efficiency in water conductivity **[ref?]**.
 It may thus be interpreted as the wood economics spectrum firstly introduced by @chave_towards_2009, and then discussed by @baraloto_decoupled_2010.
@@
 Furthermore, our results are similar to the ones found by @fortunel_leaf_2012, where basic PCA axes also represented leaf and wood economics spectra (inverted compare to us).
-Noticeably though, our analysis integrated less traits (5 against 14), less species (89 against 758), and considered a regional set of species found in a local forest-plot scale while @fortunel_leaf_2012 considered a regional set of species.
-Therefore, we could capture basic economics spectra found at regional scale in Neotropics at a far more local scale in an Indian rainforest, even though such downscaling does not hold in any case [@messier_trait_2016].
-It suggests that these spectra indeed represent major drivers of ecological strategies in local forest dynamics in our context.
-
-## Community-level trait values
-
-We analyzed functional composition of $20*20m^2$ communities (CWM), as well as the variation of this composition with environment across the 10 ha forest plot.
-We found that the traits of the leaf economics spectrum (LES) were mainly affected by aspect with more convergent community composition (lower CWV) under south-west orientation, especially for LDMC of less common species.
-The influence of aspect is little documented in tropical forests, but still highlighted by @gunatilleke_specieshabitat_2006 in monsoon rainforests.
-@bunyan_effect_2015 also observed similar influence in mountain rainforests of Western Ghats, and suggested that it could result from monsoon winds.
-In our context, monsoon winds arise from south-west direction, and could bring more stressful and disturbed conditions to species [@soman_aspects_1990].
-Uppangala plot is situated inside a low elevation slope separated from west coast by a higher crest orientated toward southwest.
-In this context it is diffcult to predict wind flow and orientation inside the 10 ha plot topgraphy.
-We hyptohesis to explanation, hardwood could be selected for mechanical resistance in more disturbed souwhtwest oriented area or topography provoke a turbulences on northeast oriented areas responsible of more disturbance and death events in the forest history resulting in lightwood filtering [@larjavaara_rethinking_2010].
-Curvature is also affecting the variation of leaf economics in community.
-A locally convex context would accumulate more soil and attract more water, increasing soil fertility and allowing persistence of species investing more in leaf resource acquisition, with higher SLA and a convergence in LDMC values.
-The variation of SLA and LDMC with curvature, is consistent with the influence of changing soil fertility [@hodgson_is_2011].
-
-Biotic factors also affected the LES through a negative effect of canopy height on LDMC.
-There was no effect of canopy height on understorey species, while it influenced leaf traits of canopy and emergent trees.
-A basic expectation is that understorey species would invest in higher LDMC for higher leaf lifespan and durability under limited light and water resources [@kitajima_tissue-level_2010].
-But the Uppangala forest is characterized by heterogeneity of canopy height allowing light to penetrate more in lower strata and a dry season of 3.8 months.
-Comsequently leaf lifespan investment may not be so important resulting in a low-high light growth trade-off strategy instead of a growth-survival trade-off [@poorter_leaf_2009], affecting SLA but not LDMC.
-Emergent and high canopy species will have a lower LDMC, resulting from a more productive strategy toward leaf resource acquisition at the expense of leaf carbon investment in defense and resistance.
-**Second hyptohesis: canopy and emergent tree are more affected by water stress than understorey species preserved by microclimatic conditions buffered by canopy**.
-Thus canopy species with the lowest LDMC will grow more quickly and then reach the highest canopy.
-Measurement of LDMC on species classified as canopy and emergent were realized most of the time on shadow leaves from understorey individuals.
-Still we observed previous relationship indicating that species phylogenetic determinism is stronger than micro-environmental variation plasticity for LDMC, as suggested by the intra-specific variation of LDMC 1.5 times larger than intra-individuals variation.
-
-The wood economics spectrum defined by WD, LA, and LT was mainly affected at community level by south-west orientation and curvature.
-South-west orientation, with probably more disturbance and stress due to exposure to monsoon winds **Maxime: counter effect by topo ?**, resulted in denser wood.
-However lower evenness than random could be related to multimodality.
-In addition, south-west orientation also related to a broader range of leaf sizes, which could be consistent with more diverse hydraulic strategies reflected by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008].
-Decreasing fertility , with concavity was related to smaller leaves inside communities with a distribution broaden by multimodality reducing evenness **to develop, eventually illustrated with one example**.
-Wetness increased with higher curvature and lower LA evenness, indicating that multimodality is happening more in wettest low elevation area.
-Wetness could compensate water departure trigerred by positive curvature in lower elevation areas.
-**Repetition with first paragraph, to merge ?**.
-
-## Variations among communities
-
-Taxonomic variation in community composition was only dependent on environmental distance.
-Because we did not find any residual pattern of increasing taxonomic dissimilarity with spatial distance, it could indicate that environmental heterogeneity was a primary driver of species distributions, and that dispersal limitation did not influence these distributions at the scale of our forest plot.
-Conversely, functional variation did not depend on environmental distance or on spatial distance **Paradoxal ?
-Is it due to environmental aggregation in one variable or to a power loss of the test ?**.
-At the 10 ha level functional variation between communities is stochastic; whereas at the community level environment is partially shapping florisitic composition.
-Measured functional traits should vary to a larger scale [@messier_trait_2016].
-Still environment is shaping community resistance to disturbance and stress by affecting wood density $\beta$ structure at fine scale.
-WD of less abundant species is even diverging with spatial residuals supposing an unaccounted spatially organized variable opposing close community in WD composition.
-Topography removed from spatial residuals, and climate and pedology being constant in the 10 ha plot, this unaccounted variation is assumed to be due to stochastic event in plot history as death.
-Unshown analysis reveals that this spatial divergence of WD is not explained by tree death events from the two last decades but certainly to older events.
-**To redraft**.
-Finally, variation in distance between taxon was not translated in phylogenetic variation whereas most of measured functional traits appeared phylogenetically conserved and clustered.
-**To discuss**.
+Noticeably, our analysis integrated less traits (5 against 14), less species (89 against 758), and considered a set of species found in a local forest-plot while @fortunel_leaf_2012 considered a regional set of species from French Guiana and Peru (i.e.
+the opposite ends of Amazonia).
+Therefore, our results shows that the economics spectra found at a regional scale in the Neotropics are also valid at a far more local scale in an Indian rainforest.
+This result does support the arguments by @messier_trait_2016 and suggests that the two observed spectra underline major ecological strategies in tropical forests, even locally.
+
+## Canopy and understorey species assembly rely on different economics spectra
+
+We analyzed functional composition of $30*30m^2$ guilds from canopy and understorey, as well as the variation of this composition with environment across a 10-ha forest plot.
+We foud that the variations in trait composition along the environmental gradients differ between canopy and understorey species, with a marked variation in traits associated to the leaf economic spectrum (LES) for understorey species and to the wood economics scpectrum (WES) for canopy species.
+In the former case, the curvature and the canopy height both had a significant on the LES-based composition of understorey species, with ...
+a higher SLA on
+
+<!-- We found different influence of the environment on economics spectra betweeen canopy and understorey species.
+Traits of the wood economics spectrum (WES) were mainly affected in canopy tree species by curvature and aspect with more convergent community composition (lower CWV) under south-west orientation ; whereas understorey tree species revealed variation in the leaf economics spectrum (LES) with curvature and canopy height.
+-->
+
+Curvature is affecting the variation of leaf economics in understorey species and the variation of wood economics in canopy tree species.
+A locally convex context would accumulate more soil and attract more water, increasing soil fertility and allowing persistence of species investing more in leaf resource acquisition in understorey with higher SLA and lower LDMC, or less in wood ressources in canopy with lighter wood and bigger leaves converging more toward an homogeneous guild composition (lower CWV).
+Curvature for some functional traits, as wood density, LDMC and SLA, is also doubled by a similar effect of the wetness variable expressing also the water accumulation in convex context resulting in increased soil fertility.
+Economics spectra are not perfectly divided between canopy and understorey, with wood and leaf respectivelly, for instance LDMC of canopy species also decreased with curvature whereas WD increased in understorey.
+
+The influence of aspect on forest composition is little documented in tropical forests.
+In a similar context that our study site, @gunatilleke_specieshabitat_2006 gave some intuitive insights on the effect of aspect on forest composition: he suggested that aspect through monsoon had a potential effect on rainforests.
+@bunyan_effect_2015 tested the effect of aspect on the occurence of forest fragments in a forest-grassland mosaic and foud that southwest orientation strongly determined the presence of forest in mountain rainforests of Western Ghats.
+They suggested that it could result from monsoon winds.
+In our context, monsoon winds are south-west oriented, and are expected to be a major source of disturbance in forests [@soman_aspects_1990].
+
+The Uppangala plot is situated in a topographically complex area with steep slopes mostly organized along southwest-oriented crests.
+In this context it is diffcult to predict wind flow and orientation inside the 10-ha plot topography.
+We propose two explanation, hardwood could be selected for mechanical resistance in more disturbed souwhtwest oriented area or topography provoke a turbulences on northeast oriented areas responsible of more disturbance and death events in the forest history resulting in lightwood filtering [@larjavaara_rethinking_2010].
+In both case it seems reasonable that canopy tree species are the most influenced by disturbance events linked to monsoon, and may even buffer the effect on understorey species.
+
+Biotic factors, illustrated here by the canopy height and the basal area, also had a significant effect on the LES-based composition of understorey tree species.
+We indeed found higher SLA values and lower LDMC values in understorey assemblages that are located under stands characterized by a tall canopy and a high basal area.
+This results contrasts with the expectation that understorey species invest in higher LDMC for higher leaf lifespan, at the expense of SLA, unde light-limited conditions [@kitajima_tissue-level_2010].
+The Uppangala forest is characterized by a large heterogeneity of canopy height allowing light to penetrate more in lower strata, maybe especially more in high canopy context covering less closely understorey, resulting in those observation on the leaf economics variation of understorey species.
+**Not convincing, better explanation ?**
 
@@
 
-Ecological strategies underline two econmic spectra as shown by @baraloto_decoupled_2010.
-Whereas the multivariate analysis of our functional traits lack trait diversity for correct interpretations of functional trade-offs.
+Ecological strategies underline two economics spectra as shown by @baraloto_decoupled_2010.
+Whereas the multivariate analysis of our functional traits lack trait diversity, we were able to identify functional trade-offs.
 Measured traits were all morphological and from the leaf except for one.
-Anatomical and chemical traits as leaf vein patterns and leaf chemical content (N, P, K) could brought broader perspective on ecological strategies of Western Ghats species.
+Anatomical and chemical traits as leaf vein patterns and leaf chemical content (N, P, K) could brought broader perspective on ecological strategies of Western Ghats species, more especially about investment in leaf ressource acquisition.
 Still the study reveal that few commonly used traits can highlight well known strategy axes.
-**Develop what would have brought this other traits**.
-
-Besides patterns have been detected between functional composition and environment at both local ($\alpha$) and between-community ($\beta$) scales.
-Lot of relation observed didn't differ from those obtain with null hypothesis reflecting neutral theory [as functional equivalence, @hubbell_neutral_2005].
-Ecological gradient and spatial scale sizes can be at the origin of this lack of environmental filtering.
-Effectively the abiotic variation is mainly topographic with a constant climate and pedology.
-Study highlighting environmental filtering often present a higher environmental variation [@bernard-verdier_community_2012].
-In parallel, the 10 ha plot scale is really low and partially responsible for the small environmental variation.
+
+Although significant patterns of trait-environment association have been detected in our study, the strength of these associations was weak in most cases (small $R^2$) and many associations were rejected by the null model.
+Studies highlighting environmental filtering effect often present a higher environmental variation than in our study site [@bernard-verdier_community_2012].
+**To move in the previous part ?**.
+In parallel, the 10*ha plot scale is really low and partially responsible for the small environmental variation.
 @kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25 ha tropical forest plots including large flat valleys.
 This mid-low elevation study of functional traits from Western Ghats forest could be reproduced at higher elevation or in dryer areas as Eastern Ghats for better insight of environmental filtering at broad scale.
-On the other hand, WD as shown to be filtered by environment at fine scale both in $\alpha$ and $\beta$ levels.
-Species wood seems highly adpted to environment in Western Gahts forests in link to monsoon events **To brought sooner, break conclusion**.
 
@@
 
-French Institute of Pondicherry (FIP)
+* French Institute of Pondicherry (FIP)
+* Cedric Vega for the extension of the plot
 
@@
 
-## S2: Phylogenetic varitions among communities
-
-A phylogenetic tree of the 89 species was built with the R `S.PhyloMaker` function proposed by @qian_updated_2016-1.
-We assessed the phylogenetic signal of each functional trait using the Moran's I statistic.
-Positive Moran's I value indicated convergence of trait values within the phylogeny (i.e.
-trait conversatism) whereas negative Moran's I will showed divergence (i.e.
-trait differentiation).
-
-To investigate phylogenetic variation among communities we calculated $\beta _{st}$ and $\pi _{st}$, for phylogenetic composition in abundance and presence-absence respectively [@hardy_characterizing_2007].
-$\beta _{st}$ integrate only phylogenetic turnover, weighting species by their local abundances; whereas $\pi _{st}$ express phylogenetic turnover, without accounting for species local abundances.
-
-Variation of composition between community pairs with those metrics was translated as a taxonomic, functional and phylogenetic distance, and compared to environmental and spatial distances among community pairs.
-Environmental distances were calculated as geometric distances between communities on the plan composed by the two first axes of environmental variables PCA (see supplementary material S1).
-Correlations with environment and space were assessed based on two Mantel tests successively with taxonomic, functional, and phylogenetic community compositions.
-A firt Mantel statistic assessed the correlation between each metric distance matrix and the environmental distance matrix (E).
-The second partial Mantel statistic assessed the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S), to represent remaining spatial dependency in community composition once the influence of environmental variation is taken into account.
-
-We used null models to test if observed correlation with Mantel statistic differed significantly from expectations under null hypothesis.
-$\beta st$ and $\pi st$ were tested against the nul model *"1s"* suggested by @hardy_testing_2008 which randomizes phylogenetic tree tips.
-
-```{r PhyloMaker, include=FALSE}
-library(phylosignal)
-library(phytools)
-library(phylobase)
-path <- system.file('phylogeny', package = 'Uppangala')
-source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
-phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
-nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
-sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), c('LatinName', 'Genus', 'Family')]
-names(sp) <- c('species', 'genus', 'family')
-results <- invisible(S.PhyloMaker(sp, phylo, nodes)) # Humboldtia_brunonis unmatched
-tree <- results$Scenario.1
-rm(phylo, results, modes, sp, S.PhyloMaker, path)
-```
-
-```{r Phylosignal}
-data <- PFT_sp
-row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
-tips <- row.names(data)[which(is.na(data$WD))]
-phy <- drop.tip(tree, tips)
-data <- data[phy$tip.label,c(2:5,7)]
-p4d <- phylo4d(phy,data)
-phyloS <- phyloSignal(p4d, methods = 'I') 
-phyloS <- data.frame(paste(round(phyloS$stat[,1], 3), 
-                           unlist(lapply(as.list(phyloS$pvalue[,1]), stars, ns = ' '))),
-                     row.names = row.names(phyloS$stat))
-names(phyloS) <- 'Moran\'s I'
-# kable(bold(phyloS))
-rm(data, tips, p4d, phyloS, phy)
-```
-
-```{r BETAst $ PIst}
-# Observed Hardys
-phy <- tree
-phy$tip.label <- Species$SpCode[match(gsub('_', ' ',phy$tip.label), Species$LatinName)]
-tips <- lapply(L, function(s){
-  c(setdiff(phy$tip.label, names(s$abundance)), setdiff(names(s$abundance), phy$tip.label))
-})
-phy <- lapply(tips, function(s){
-  drop.tip(phy, s)
-})
-phy <- lapply(phy, function(s){
-  s$node.label <- NULL
-  return(s)
-})
-spaphy <- mapply(function(Ls, phys){
-  spacodi.calc(sp.plot = t(Ls$abundance), phy = phys, pairwise = T)[c('pairwise.Bst', 'pairwise.PIst')]
-}, Ls = L, phys = phy, SIMPLIFY = F)
-## Mantel values
-spaphy.mantel <- mapply(function(s, Mes, Mss){
-  list(
-    `E` = lapply(s, function(m){
-      vegan::mantel(m, Mes, permutations = 0)$statistic
-      }), 
-    `S | E`= lapply(s, function(m){
-      vegan::mantel.partial(m, Mss, Mes, permutations = 0)$statistic
-      })
-    )
-}, s = spaphy, Mes = Me, Mss = Ms, SIMPLIFY = F)
-
-# Simulated mantel
-nphy <- lapply(phy, function(s){
-  rep(list(s), 999)
-})
-cl <- makeCluster(cores)
-clusterExport(cl, list('nphy', 'L', 'spacodi.calc', 'Ntip', 'resamp.phy', 'branching.times'))
-nphy <- clusterApply(cl, nphy, function(s){
-  lapply(s, resamp.phy)
-})
-stopCluster(cl)
-nspaphy <- mapply(function(s, Ls){
-  lapply(s, function(p){
-  spacodi.calc(sp.plot = t(Ls$abundance), phy = p, pairwise = T)[c('pairwise.Bst', 'pairwise.PIst')]
-  })
-}, s = nphy, Ls = L, SIMPLIFY = F)
-rm(cl, nphy)
-invisible(gc())
-nspaphy.mantel <- mapply(function(s, Mes, Mss){
-  list(
-    `E` =lapply(s, function(x){
-      lapply(x, function(M){
-        vegan::mantel(M, Mes, permutations = 0)$statistic})}), 
-    `S | E` = lapply(s, function(x){
-      lapply(x, function(M){
-        vegan::mantel.partial(M, Mss, Mes, permutations = 0)$statistic
-      })
-    })
-  )
-}, s = nspaphy, Mes = Me, Mss = Ms, SIMPLIFY = F)
-## Non null mantels
-names(nspaphy.mantel) <- c('C', 'U')
-nspaphy.mantel <- lapply(nspaphy.mantel, function(s){
-  lapply(s, function(m){
-    lapply(m, `[`, names(m[[1]]))
-    apply(do.call(rbind, m), 2, as.list)
-  })
-})
-nspaphy.mantel <- lapply(nspaphy.mantel, function(s){
-  lapply(s, function(m){
-    lapply(m, unlist)
-  })
-})
-spaphy.res <- mapply(function(s, ns){
-  mapply(function(m, nm){
-    mapply(function(b, nb){
-      paste(format(b, scientific = T, digits = 3), 
-            Uppangala::stars(min(1-rank(c(b,nb))[1]/(n+1),rank(c(b,nb))[1]/(n+1)), ns = '   '))
-    }, b = m, nb = nm, SIMPLIFY = T)
-  }, m = s, nm = ns, SIMPLIFY = F)
-}, s = spaphy.mantel, ns = nspaphy.mantel, SIMPLIFY = F)
-spaphy.res <- mapply(function(s, str){
-  s <- data.frame(s)
-  s <- cbind(c(str, ' '), c('$U_{st}$', '$\\tau _{st}$'), s)
-  names(s) <- c('S', 'M', 'E', 'S|E')
-  return(s)
-}, s = spaphy.res, str = c('C', 'U'), SIMPLIFY = F)
-spaphy.res <- do.call('rbind', spaphy.res)
-
-# Result
-kable(bold(spaphy.res), row.names = F)
-rm(mtest, spaphy.mantel, nspaphy, nspaphy.mantel, spaphy.res)
-```
-
-While LDMC did not phylogenetic structure (Moran's I of 0 with a $p-value > 0.1$), LA and LT were phylogenetically clustered (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA were weakly clustered (Moran's I of 0.005 and 0.003 with $p-value < 0.1$).
-Phylogenetic variation among communities was neither correlated to environmental variation (Mantel test, $r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor to remaining spatial variation in partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).
-
+## S2: Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in presence-absence
+
+```{r CWM PA}
+kable(bold(CWM.ses.res[c(8:14,22:28),-2]), row.names = F)
+```
+
+Supplementary material S2: Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in presence-absence _We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table 1 for abbreviations.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
 # References

### 2017-10-04 inst/doc/draft/draft2.Rmd
hypervolume update

@@
   bookdown::gitbook
+  # bookdown::pdf_book
+  # word_document
 ---
@@
 
-```{r hypervolume, fig.cap='Functional hypervolume overlap of canopy and understorey species.'}
+```{r hypervolume, eval=FALSE, fig.cap='Functional hypervolume overlap of canopy and understorey species.'}
 bd <- Q %>% 
@@
   hv <- lapply(traits_strata, function(strata) 
-    hypervolume::hypervolume(strata[names(bd)], bandwidth = bd, verbose = F))
-  set <- hypervolume::hypervolume_set(hv$C, hv$U, check_memory = F, verbose = F)
-  hypervolume::hypervolume_sorensen_overlap(set)
+    hypervolume::hypervolume(strata[names(bd)], kde.bandwidth = bd, verbose = F))
+  set <- hypervolume::hypervolume_set(hv$C, hv$U, check.memory = F, verbose = F)
+  hypervolume::hypervolume_overlap_statistics(set)
 }
@@
   xlab('Overlap null distribution')
+summary(overlap$C)
+CalcOverlap(overlap$C, overlap$U)
 ```
@@
 
-```{r hypervolume2}
+```{r hypervolume2, eval=FALSE}
 data <- Q %>% 

### 2017-10-04 inst/doc/draft_mrm.Rmd
hypervolume update

@@
 Table 3: Leven's test of PCA's guilds variance. *Guilds variance inside the principal component analysis (PCA) were compared with Leven's test on PCA first and second axes.*
+
+```{r hypervolume}
+data <- PFT_sp[-which(is.na(PFT_sp$WD)), c('LA', 'WD', 'LT', 'LDMC', 'SLA', 'strata')]
+bd <- estimate_bandwidth(data[-6])
+C <- hypervolume(subset(data, strata == 'C')[-6], bandwidth = bd, name = 'C')
+U <- hypervolume(subset(data, strata == 'U')[-6], bandwidth = bd, name = 'U')
+UC <- new("HypervolumeList", HVList = list(C = C, U = U))
+set <- hypervolume_set(C, U, check_memory = F)
+hypervolume_sorensen_overlap(set)
+get_volume(set)
+```
 

### 2017-11-27 inst/doc/draft2.Rmd
Uppangala 0.1.9038 first draft after PhD beginning (NM1, 2 et 3)

--- 2016-07-18 inst/doc/draft2.Rmd
@@
 ---
-title: 'Uppangala tree functionnal: Article draft'
-author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
-output: pdf_document
+title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
+author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
+date: '`r Sys.Date()`'
+output:
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
+csl: mee.csl
 bibliography: refs.bib
+link-citations: yes
 ---
@@
 ```{r config, message=FALSE, warning=FALSE, include=FALSE}
+rm(list = ls()); invisible(gc())
 library(knitr)
+library(parallel)
+library(Uppangala)
+library(broom)
+library(vegan)
+library(dplyr)
+library(reshape2)
+library(ggplot2)
+library(ggfortify)
+cores <- 16
 opts_chunk$set(
@@
   cache = T, cache.lazy = F)
+cv <- function(x, na.rm = T)
+  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
+path <- '~/Documents/IFP/Uppangala/inst/doc/draft_save/'
+n = 999 # Random repetitions
 ```
@@
 
-Keywords: functionnal trait, tropical evergreen forest, Western Ghats, India...
+## Keywords
+
+functional trait; rainforest; vertical forest strata; Western Ghats; environmental filtering; niche differentiation; ecological guild
 
 # Introduction
+
+Understanding the determinants of species occurrence and co-occurrence is a long standing issue in ecology [@diamond_ecology_1975].
+Tropical rainforests have especially fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
+Niche theory explains species coexistence through niche differences that limit competitive exclusion [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that maintenance of high diversity is possible even for functionally equivalent species, under the influence of stochastic life, death, reproduction and dispersal dynamics [@hubbell_unified_2001].
+Both theories can provide realistic expectations of taxonomic diversity [@chave_comparing_2002] so that inferring ecological processes from taxonomical composition pattern is challenging in community ecology.
+
+Investigating patterns of functional or phylogenetic diversity may reveal signatures of ecological processes that are specific to the niche theory [@mcgill_rebuilding_2006; @westoby_land-plant_2006; @mouquet_ecophylogenetics:_2012].
+Indeed, if taxonomic composition is expected to be spatially structured in both theories, due to the spatial structure of environment or to dispersal limitation, functional or phylogenetic assemblage are expected to be spatially structured only under the niche theory [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012; @parmentier_prevalence_2014].
+
+The variation in functional traits among species is expected to reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007] and expresses fundamental trade-offs along environmental gradients [@mark_westoby_plant_2002].
+For instance, specific leaf area varies along the soil fertility-shade gradient and represents a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@hodgson_is_2011].
+Fundamental trade-offs entail functional traits covariation along basic leaf [@wright_worldwide_2004 but see @osnas_global_2013 and @lloyd_photosynthetically_2013] and wood [@chave_towards_2009] economics spectra.
+The leaf economics spectrum opposes acquisitive to conservative ecological strategies, basically expressed at leaf level.
+The wood economics spectrum distinguishes species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength [@chave_towards_2009].
+@baraloto_decoupled_2010 found decoupled wood and leaf economics of trees in neotropical forest, but @reich_world-wide_2014 conversely posited that both spectra have to be integrated in a comprehensive whole plant economics spectrum [@freschet_evidence_2010] opposing fast growing acquisitive species to slow growing and more conservative ones.
+
+The trait values of coexisting species represent a non-random sample of trait values found in a broader area, du to specific niche-based dynamics in the local community [@lortie_rethinking_2004; @bernard-verdier_community_2012].
+Then the signature of local niche-based dynamics can be gasped by assessing the deviation of local community structure from a larger "species pool", classicaly using null models [@kraft_functional_2008].
+In addition, the nature of functional traits and ecological strategies driving species coexistence can change across scales.
+For instance, @messier_trait_2016 have underlined that the leaf economics spectrum basically expresses changes in ecological strategies along broad-scale environmental gradients, while it is not necessarily a primary driver of local coexistence.
+
+The concept of guilds refer to groups of functionally similar species [@wilson_guilds_1999].
+Ecological guilds can coexist and occupy distinct niches within communities [@vergnon_niches_2009].
+In tropical forests, the assembly of tree communities has been mostly studied irrespectivelly of guilds of canopy and understorey tree [@wright2002plant].
+However, distinctive ecological constraints are expected to drive different resource use strategies and biotic interactions in these guilds, which should translate into different functional composition and response to environmental factors.
+For instance, in Colombian Amazonie, @duque2002different found some contrasted response of the composition of canopy and understorey guilds along environmental gradients.
+
+In this paper, we used a trait-based approach to study separately the assembly of canopy and understorey tree species along topographical and biotic gradients in a large forest plot (10ha) loacted in Western Ghats biodiversity hotspot [@myers_biodiversity_2000], India.
+Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988].
+It thus provides an appropriate context to adress guild-dependent assembly rules and ecological strategies in tropical rainforests.
+Although considered as wet evergreen, the forest undergoes contrasted long dry and intense rainy seasons under the influence of monsoon.
+In this context, we expected water stress and competition for resource to be important drivers of tree functional composition.
+We investigated how both canopy and understorey tree species composition varied with local environmental conditions, using appropriate null models.
+We specifically addressed the following questions:
+
+- Does functional traits variation and covariation reflect basic dimensions of ecological strategy in the rainforest ?
+- Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?
 
@@
 
-The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village.
-The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996.
-The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988].
+The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry in an undisturbed, old-growth wet evergreen Dipterocarp forest [@pascal_structure_1996].
+The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), near the Uppangala village, and thus refers to as "Uppangala plot".
+Annual precipitation averages 5108 mm per year but is highly variable in time with heavy rainfall from June to October (90% of the rainfall during the monsoon) and a severe dry season (monthly rainfall $\leq 45$ mm) lasting 3.8 months [@pelissier_twenty_2011].
+Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996].
 Climate and geology of the 10-ha plot can be considered as homogeneous.
-General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].
-Each tree girth is monitored by a two-year census.
+The topography of the site is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes, so that topography is expected to be a major component of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998].
+Therefore, we hypothesized that topographic variation should influence community assembly and yield changing ecological strategies and functional composition [@condit_spatial_2000].
 
-## Sampling
+## Taxonomic composition and trait sampling
 
-We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates.
-Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals.
-Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot.
-For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013.
-Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy.
-Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.
-Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours.
-For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).
+From 1989 to 2014, the trees $\geq 10$ cm in diameter at breast height (dbh) were identified to species level, tagged, mapped and monitored in an area progressively extended to a contiguous $300 * 330 m^2$ (ca.
+10-ha) plot [@pelissier_twenty_2011].
+The plot included 111 different tree species.
+We collected trait data from 89 over 111 (80.1%) species, which represented 99.0% of tree individuals present in the plot.
+Note that the 32 most abundant species represented more than 80% of tree individuals.
+We sampled five individuals per species for these 32 species, and sampled one to five individuals for the remaining species (in total 267 individuals).
+For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation of @perez-harguindeguy_new_2013.
+Individuals and leaves were selected in consistent light conditions, development stage and sanitary state.
+Dawkins crown exposure index [@dawkins_management_1958] was recorded for each tree sampled to control for any sun-exposure effect.
+Leaf and branch samples were kept in humidified ziplock filled with enriched $CO_{2}$ air inside a black bag until measurement within the next 24 hours.
+For some individuals, we were unable to collect suitable branch sample, i.e.
+with a diameter $\leq 1$ cm (11 missing species over 89 representing 1% of individuals present).
+
+## Trait measurements
 
 Leaf petioles, rachis and petiolules were removed for all measurements.
-Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite].
-Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time.
-Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias.
-Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues.
-Branch wood samples were sink in water to evaluate their volume by water displacement.
-Samples were then dried for 48 hours at 80 degrees celsius and weighted.
+Leaf fresh weights, thickness and area were measured, using a weighing scale (**add detail on brand and precision**), a micrometer (**add detail on brand and precision**) and a bend scanner with ImageJ software [**ref**] respectively.
+Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$.
+Once dried, the leaves were weighted directly after been extracted from the oven to avoid any humidity bias.
+The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues.
+Branch wood samples were immersed in water to evaluate their volume by water displacement.
+Wood samples were then dried for 48 hours at 80 $^\circ C$ and weighted.
 
-## Traits
+We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table \@ref(tab:Traits)).
+We choose those five traits because they proved to be good proxies of the major dimensions of ecological strategies in plants [@mark_westoby_plant_2002; @diaz_global_2015] and were relatively easy to measure on the field.
 
-With leaf and branch wood samples we calculated five functionnal traits: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the wood sample volume in $cm^3$ divided by the wood sample dry weight in $g$ (see table 1).
-Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition.
-For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].
-
-Table 1: Functional traits measured.
-*Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development from @baraloto_decoupled_2010.*
-
-```{r 1 Traits}
-rm(list = ls()); invisible(gc())
-library(Uppangala)
+```{r Traits}
 PFT <- genPFT()
-captions <- genCaptions()
-table <- data.frame(
-  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
-  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
-  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
-            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
-            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
-            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
-            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
-  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
-                         round(sd(PFT$Thick, na.rm = T)), ')'),
-                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
-                         round(sd(PFT$LA, na.rm = T)), ')'),
-                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
-                         round(sd(PFT$LDMC, na.rm = T)), ')'),
-                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
-                         round(sd(PFT$SLA, na.rm = T),2), ')'),
-                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
-                         round(sd(PFT$WD, na.rm = T),2), ')')),
-  unit = c('um', 'mm2', 'mg.g-1', 'm2.kg-1', 'g.cm-3'),
-  strategy = c('Leaf defense vs. investment',
-               'Leaf ressource capture vs.investment',
-               'Leaf defense vs. investment',
-               'Leaf ressource acquisiton vs. defense',
+kable(data.frame(
+  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
+  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
+  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
+  Strategy = c('Leaf defense and investment',
+        'Leaf resource capture vs.investment',
+        'Leaf defense and investment',
+        'Leaf resource acquisiton vs. defense',
         'Stem transport, structure and defense')
-  )
-kable(table,  label="traits", row.names = F)
+), caption = 'Functional traits measured. Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].')
 ```
@@
 
-To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river.
-Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$.
-Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008].
-Distance to the nearest potential river was caluclated under ArcGis (cite).
+Topography was expected to be a major environmental factor influencing the assemby of tree communities [@gimaret-carpentier_sampling_1998].
+We selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed in ArcGis [**ref**].
 
-To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density.
-Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$.
-Both basal area and stem density are calculated with 2013 girth inventory from Uppangala plot on stem with a diameter above 10 centimeters at breast height ($DBH>10$).
-Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$).
-So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).
+In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the abiotic environmental variables (see Supplementary materials [S1: Principal component analysis of topographical variables]).
+We selected four little-correlated variables: curvature representing local water flow paths and soil accumulation areas, wetness highlighting water accumulation areas over the whole study area, southwesterness representing aspect [@bunyan_effect_2015] and slope, which may have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher].
+
+In addition, we selected two biotic factors related to local forest structure: canopy height and basal area.
+Canopy height, i.e.
+the maximum local height of the canopy, was calculated from a 1-m resolution digital canopy model built from a LiDAR (Light Detection And Ranging) dataset acquired in 2013 over the Uppangala plot.
+Basal area was calculated from the girth measurements taken within the 10-ha plot (censuses done in 2013 and 2014) for stems $\geq 10$ cm dbh.
 
@@
 
-### Covariation in traits
+### Guilds
 
-Traits variation was first investigated in three levels with the coefficient of variation ($cv=sd/mean$) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific).
-For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species.
-To test the correlation and covariation among traits we performed a principal component analysis (PCA).
-We investigated peculiar traits covariation highlighted with the PCA with a Spearman's rank correlation test.
+@pascal_wet_1988 defined four vertical strata attained by adult trees: emergent above 39m, canopy trees between 25 and 39m, sub-canopy trees between 15 and 25m, and understorey trees below 15m.
+Despite the vertical hierarchy, steep topography could allow sub-canopy trees receiving more light, especially in slopes, and thus participates to canopy.
+We thus defined two basic guilds called canopy (C), composed of emergent, canopy and sub-canopy species (n = 55), and understorey (U) species (n = 34).
 
-Table 2: Coefficient of variation intra-individual, intra-specific and inter-specific.
-*Mean of coefficient of variation for leaves in all individual (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species.
-See table 1 for abbreviation.*
+### Traits variation and covariation
 
-```{r 2 CV}
-PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
-row.names(PFT_ind) <- PFT_ind$Group.1
-SpT <- unique(PFT[1:2])
-Sp <- SpT$SP
-names(Sp) <- SpT$Tree
-SpCodeT <- unique(PFT[2:3])
-SpCode <- SpCodeT$Sp_Code
-names(SpCode) <- SpCodeT$SP
-CET <- unique(PFT[c(1,4)])
-CE <- CET$CE
-names(CE) <- CET$Tree
-PFT_ind$SP <- Sp[row.names(PFT_ind)]
-PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
-PFT_ind$CE <- CE[row.names(PFT_ind)]
-PFT_ind <- PFT_ind[-c(1:2)]
-cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
-cv$Individual <- unlist(lapply(
-  as.list(row.names(cv)),
-  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
-cv$Species <- unlist(lapply(
-  as.list(row.names(cv)),
-  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
-cv$Community <- unlist(
-  lapply(as.list(row.names(cv)),
-         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
-             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
-cv <- round(cv, 2)
-cv['WD','Individual'] <- ''
-kable(cv)
-```
+We first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among leaves of the same individual (intra-individual), among trees of the same species (intra-specific for species with more than one individual), and among species (inter-specific).
+We then performed a principal component analysis (PCA) of the mean value of traits per species to identify main dimensions of ecological strategies along which traits can covary among guilds.
+In addition we explored guilds functional signature.
+We used guilds functional hypervolume and tested their overlap against 999 nul hypervolume randomizing species in strata.
+Subsequent community-level analyses were based on trait values averaged per species.
 
-Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2).
-Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$).
+### Drivers of community dynamics
 
-### Environment definition
+We subdivided the 10-ha plot into $30*30 m^2$ quadrats leading to 110 local communities.
+We analyzed the functionnal composition of the guilds in the 110 communities.
+Guilds in the communities have a mean number of 28 individuals representing 10 species, with less in understorey than in canopy.
 
-In order to avoid multi-colinearity in future analysis, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement.
-In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling.
-Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits.
-Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*).
-Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).
+We calculated weighted mean (GWM, where G stands for "guild") and weighted variance (GWV) [GWM and GWV, @violle_let_2007] for each guild in each community.
+In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of species $i$ weighted by species relative abundance $p_{i}$:
+$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
+where $S$ is the number of species in the quadrat $j$.
+We also calculated the average trait values based on presence-absence values to assess the influence of less abundant and rare species on community trait composition [@cingolani_filtering_2007].
+GWV of trait $t$ was calculated as :
+$$GWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - GWM_{j})^2 ~ (2)$$
 
-### Community analysis
+**We need a new paragraph to clearly present each null model, its implications and use for subsequent analysis.**
 
-Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats.
-For all communities, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes.
-We used the following formula to compute each community weighted parameter (CWP):
-$$CWP = \sum_{k=1}^{nj} W_{k,j} * P_{k,j} ~ (1)$$ 
-where $nj$ is the number of species sampled in community $j$, $W_{k,j}$ is the weight of species $k$ in community $j$ used to compute CWP (relative abundance, presence-absence, or relative basal area), and $P_{k,j}$ is the parameter of species $k$ in community $j$ (mean or the variance).
-In addition 5 functionnal diversity indices (FD) weighted by abundances were computed: functional richness (FRic), functional evenness (FEve), functional divergence (FDiv) and Rao's quadratic entropy (Q) (@villeger_new_2008, @botta-dukat_raos_2005).
-Spatial variograms were used to investigate our CWM and CWV spatial organization.
-CWM, CWV, and FD spatial autocorrelation (SAC) was also assessed by Moran's I statistic.
+To test for any significant changes in functional composition among guilds, we devised a null model approach.
+A null model (NM1) randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix) [@bernard-verdier_community_2012].
+Note that, spatial correlations in tree abundance are conserved limiting associated type I error.
+We performed 999 randomizations of null model NM1 and we derived a two-tailed test of the proportion of guilds deviating from the null expectation, with adjusted p-values by Benjamini-Hochberg correction **ref**.
+Shifts in GWM values with changing environmental conditions could reflect the shift of species strategies across assemblages [@ackerly_trait-based_2007].
 
-Links between both CWM and CWV with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables.
-Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables.
-A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix).
-Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance.
-Same procedure was also applied to functionnal diversity indices (FD) from our communities.
-Finally spatial autocorrelation (SAC) of residuals was investigated to assess selected models.
+To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (2)$$
+where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of the corresponding null distribution.
+The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$).
+Analyses were conducted on the five functional traits.
 
-### Fourth-corner analysis
-
-In addition of our community analysis by regression, we conducted both RLQ and fourth-corner analyses to test the effect of environments on Uppangala tree functionnal traits following @dray_testing_2008.
-Both matrices R (`Environment*Sites`) with selected features and Q (`Traits*Species`) were reduced by a principal component analysis.
-We used a Monte-Carlo test to compare fourth-corner and RLQ results to 999 nul models built with the model *m6* proposed by @ter_braak_improved_2012.
-Fourth corner analysis was tested with 3 types of L matrices ($Sites*Species$) with species relative abundance, presence absence or relative basal area in each sites; whereas RLQ analysis was only conducted with relative abundance.
-
-### Phylogenetic signal
-
-Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1.
-Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 functionnal traits, the two PCA axes and in the phylogenetic tree.
-Phylogenetic signal structure was assessed by a tree tip distances correlogram.
-In addition, $\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate phylogenetic signal spatial organization with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion.
-Finally phylogenetic association to habitat was also studied by an RLQ analysis with tree tip distances as the Q matrix reduced by a principal coordinate analysis.
-RLQ result with phylogenetic distance matrix was compared to 999 nul models were phylogenetic tree tips were randomized.
-
-All analysis were conducted in the R 3.3.0  free software environment for statistical computing and graphics [@r_core_team_r:_2016] using the pacakge 
-`ade4` [@dray_ade4_2007], 
-`lavaan` [@rosseel_lavaan:_2012], 
-`FD` [@laliberte_distance-based_2010], 
-`gstat` [@pebesma_multivariable_2004], 
-`ape` [@paradis_ape:_2004], 
-`phylosignal` [@keck_phylosignal:_2015] 
-and `spacodiR` [@eastman_spacodir:_2013].
+All analyses were conducted using the R 3.3.0  software [@r_core_team_r:_2016] with packages `ade4` [@dray_ade4_2007], and `vegan` [@oksanen_vegan:_2016].
 
@@
 
-## Traits and environment
+## Traits variation and covariation
 
-The multivariate analysis of trait correlation shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1).
-The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood.
-Whereas the second PCA axis (PCA2) covary mainly with SLA representative of the leaf economic spectrum [LES, @wright_worldwide_2004].
-PCA first axis might be associated with the WD covariation to the wood economic spectrum [WES, @chave_towards_2009].
-Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us).
-Our results integrate less traits (5 against 14), and less species (92 against 758), but still capture the same strategy axes.
+We found an overall increase of functional trait variation from intra-individual to inter-specific levels (see table \@ref(tab:CV)), even though LA showed almost as large variability at intra-individual than at intra-specific levels ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
+given the relatively low intra-individual and intra species variation, compared to inter-species variation, using species trait average values for community analyses is warranted.
 
-Figure 1 shows WD orthogonal to SLA following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$).
-LA is almost opposed to WD and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007.
-Finally, we found the positive relation suggested by @diaz_global_2015 between LDMC and WD (Spearman's $rho = 0.24$, $p-value < 0.001$).
-
-```{r 3 PCA PFT}
-library(vegan)
-
-PFT_ind$LA <- log(PFT_ind$LA)
-PFT_ind$SLA <- log(PFT_ind$SLA)
-PFT_ind$ID <- row.names(PFT_ind)
-row.names(PFT_ind) <- PFT_ind$ID
-PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
-Species <- genSpecies()
-PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
-row.names(PFT_sp) <- PFT_sp$Group.1
-PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
-PFT_sp <- PFT_sp[-c(1:2,4)]
-pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_sp, cor = T)
-
-lambda <- pca$sdev * sqrt(nrow(pca$scores))
-pca.var = apply(pca$score, 2, var)
-PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
-names(PCAscores) <- c('WES', 'LES')
-PCAscores$ID <- row.names(PCAscores)
-plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
-     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
-     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
-     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
-# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
-par(new=T)
-Rot <- t(t(pca$loadings)*lambda)
-XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
-XLIM <- XLIM+(XLIM*0.2)
-plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
-lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
-lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
-legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
-       text.col = c("green", "firebrick"))
-
-arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
-text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
-Hmisc::subplot(function(){
-  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
-          cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(pca.var)), col = 'red') ;
-  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
-         box.col = 'white')},
-  x=grconvertX(c(0.75,1), from='npc'),
-  y=grconvertY(c(0.05,0.30), from='npc')
-)
-axis (3)
-axis (4)
-
-PCAscores <- as.data.frame(pca$scores)[1:2]
-names(PCAscores) <- c('WES', 'LES')
-PCAscores$ID <- row.names(PCAscores)
-PFT_sp$ID <- row.names(PFT_sp)
-PFT_sp <- merge(PFT_sp, PCAscores, all = T)
-row.names(PFT_sp) <- PFT_sp$ID
-PFT_sp <- PFT_sp[-which(names(PFT_sp) == 'ID')]
+```{r CV}
+PFT %>% 
+  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
+  melt(variable.name = 'trait') %>% 
+  group_by(trait) %>% 
+  summarise_each(funs(min(.,na.rm = T), max(., na.rm = T), 
+                      mean(.,na.rm = T), sd(.,na.rm = T))) %>% 
+  left_join(PFT %>% 
+              select(-Sp, -SpCode, -CE, -SHA, -LTD) %>% 
+              melt(id.vars = 'Tree', 
+                   variable.name = 'trait', 
+                   value.name = 'Intra-individuals') %>% 
+              group_by(Tree, trait) %>% 
+              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
+              ungroup() %>% 
+              select(-Tree) %>% 
+              group_by(trait) %>% 
+              summarise_each(funs(mean(., na.rm = T)))) %>% 
+  left_join(PFT %>% 
+              group_by(Tree, Sp, SpCode, CE) %>% 
+              summarise_each(funs(mean)) %>% 
+              ungroup() %>% 
+              select(-Tree, -SpCode, -CE, -SHA, -LTD) %>% 
+              melt(id.vars = 'Sp', 
+                   variable.name = 'trait', value.name = 'Intra-species') %>% 
+              group_by(Sp, trait) %>% 
+              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
+              ungroup() %>% 
+              select(-Sp) %>% 
+              group_by(trait) %>% 
+              summarise_each(funs(mean(., na.rm = T)))) %>% 
+  left_join(PFT %>% 
+              group_by(Tree, Sp, SpCode, CE) %>% 
+              summarise_each(funs(mean)) %>% 
+              ungroup() %>% 
+              select(-Tree, -CE) %>% 
+              group_by(Sp, SpCode) %>% 
+              summarise_each(funs(mean)) %>% 
+              ungroup() %>% 
+              select(-Sp, -SpCode, -SHA, -LTD) %>% 
+              melt(variable.name = 'trait', value.name = 'Inter-species') %>% 
+              group_by(trait) %>% 
+              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T)))) %>% 
+  kable(digits = 2, caption = 'Intra-individual, intra-specific and inter-specific component of trait variation.
+The overall range, mean and standard deviation are provided for each trait.
+The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species.
+See table \\@ref(tab:Traits) for abbreviation.')
 ```
 
-Figure 1: Principal component analysis (PCA) performed on species trait means.
-*Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line).
-See table 1 for abbreviation.*
+The first axis of the principal component  analysis (PCA) of mean species trait values was basically related to wood density and structural leaf trait variations (LT, LA, LDMC), thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood.
+The second axis mostly opposed SLA to LDMC, and to a less extent to wood density.
+However, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between those two traits.
+Finally, hypervolume overlap between the two guilds was not significantly lower than null expectations ($p = 0.165$, see Supplementary materials [S2: Hypervolumes of functional guilds]).
 
-Multi-colinearity was avoided by reducing our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see figure 2).
-Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
-Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables.
-Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary.
-Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)
+```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values.
+First axis, in red, is named wood economic spectrum (WES) ; whereas second axis, in green, is named leaf economic spectrum (LES).
+See table \\@ref(tab:Traits) for abbreviation.'}
+PFT <- PFT %>% 
+  mutate(LA = log(LA)) %>% 
+  mutate(SLA = log(SLA)) %>% 
+  group_by(Tree, Sp, SpCode, CE) %>% 
+  summarise_each(funs(mean(., na.rm = T))) %>% 
+  ungroup() %>% 
+  group_by(Sp, SpCode) %>% 
+  summarise_each(funs(mean(., na.rm = T))) %>% 
+  ungroup() %>% 
+  select(-Tree, -Sp, -CE, -LTD, -SHA)
+row.names(PFT) <- PFT$SpCode
+
+PFT %>% 
+  princomp(~ LA + WD + LT + LDMC + SLA, data = ., cor = T) %>% 
+  autoplot(colour = 'grey', loadings.label.size = 6,
+           loadings.label.colour = 'black',
+           loadings = T, loadings.label = T, loadings.colour = 'black',
+           loadings.label.vjust = 1.2) +
+  coord_equal() +
+  geom_hline(aes(yintercept = 0, linetype = 'Wood'), 
+             col = 'red') +
+  geom_vline(aes(xintercept = 0, linetype = 'Leaf'), 
+             col = 'green') +
+  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
+  scale_linetype_manual(name = "Economics\nSpectrum", values = c(2, 2), 
+                        guide = guide_legend(override.aes = 
+                                            list(color = c("red", "green")))) +
+  theme(legend.position = c(1, 1), 
+       legend.justification = c(1, 1))
+```
 
 ```{r 4 env-com}
-library(raster)
-Env <- genEnv()
-Env$Rivers <- readAll(Env$Rivers)
-Trees <- genTrees()
-Trees$SpCode <- tolower(Trees$SpCode)
-Trees$ID <- row.names(Trees)
-Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
-row.names(Trees) <- Trees$ID
-Trees <- Trees[-which(names(Trees) == 'ID')]
-com <- quadrats()
-Trees$com <- NA
-XY <- data.frame(X = Trees$X, Y = Trees$Y)
-XY <- XY[-which(is.na(XY$X)),] # Removing NA
-coordinates(XY) <- ~ X + Y
-proj4string(XY) <- crs(com)
-Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
-Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
-com@data$id <- paste0('C', com@data$id)
-com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
-names(Deriv)[5] <- 'SW'
-com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
-com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
-com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
-rm(Deriv, XY)
-pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
-lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
-pca.env.var = apply(pca.env$score, 2, var)
-plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
-     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
-     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
-Hmisc::subplot(function(){
-  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
-          cex.names = 0.7, axes = T, names.arg = NA) ;
-  lines(b, bstick(length(pca.env.var)), col = 'red') ;
-  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
-         box.col = 'white')},
-  x=grconvertX(c(0.07,0.37), from='npc'),
-  y=grconvertY(c(0.025,0.30), from='npc')
-)
-par(new=T)
-Rot <- t(t(pca.env$loadings)*lambda)
-XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
-XLIM <- XLIM+(XLIM*0.2)
-plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
-lines(XLIM, c(0,0), lty = 2, lwd = 2)
-lines(c(0,0), XLIM, lty = 2, lwd = 2)
-col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
-arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
-        Rot[,1],Rot[,2],lwd=2,col=col)
-legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
-       text.col = col, cex = 1.2)
-text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 4), rownames(Rot),cex=1.2,col=col)
-axis (3)
-axis (4)
+# LES and WES
+PFT <- PFT %>% 
+  princomp(~ LA + WD + LT + LDMC + SLA, data = ., cor = T) %>% 
+  .$scores %>% 
+  data.frame() %>% 
+  select(Comp.1, Comp.2) %>% 
+  rename(WES = Comp.1) %>% 
+  rename(LES = Comp.2) %>% 
+  mutate(SpCode = row.names(.)) %>% 
+  left_join(PFT)
+# library(raster)
+# Env <- genEnv()
+# Trees <- genTrees()
+# Species <- genSpecies()
+# Trees$ID <- row.names(Trees)
+# Trees <- merge(Trees, PFT_sp)
+# row.names(Trees) <- Trees$ID
+# Trees <- Trees[-which(names(Trees) == 'ID')]
+# com <- quadrats(cs = c(30,30))
+# XY <- Trees[c('x', 'y')]
+# coordinates(XY) <- ~ x + y
+# proj4string(XY) <- crs(com)
+# Trees$com <- (XY %over% com)[,1]
+# com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+# Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/share/applications')
+# names(Deriv)[5] <- 'SW'
+# com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
+# com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+# com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+# BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
+# com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
+# pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
+# R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+# row.names(R) <- com@data$id
+# L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
+# L[is.na(L)] <- 0
+# Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')]
+# row.names(Q) <- PFT_sp$SpCode
+# sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] == 3)
+# Q$strata <- 'C' ; Q$strata[sel] <- 'U'
+# sel <- intersect(row.names(Q), names(L))
+# Q <- Q[sel,]
+# Q$SpCode <- row.names(Q)
+# L <- L[sel]
+# R <- R[row.names(L),]
+# R$community <- row.names(R)
+# Lbin <- L ; Lbin[Lbin > 1]  <- 1
+# rm(Deriv, XY, Env,  sel, BA, com, Trees, Species)
+# save(R, L, Lbin, Q, pca.env, file = file.path(path, 'envcom.Rdata'))
+rm(PFT, PFT_sp, trait)
+load(file.path(path, 'envcom.Rdata'))
 ```
 
-Figure 2: Principal component analysis (PCA) performed on topgraphical variables.
-*Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).
-SW stand for southwesterness.*
-
-## Community analysis
-
-Spatial variograms of CWM and CWV didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$. Functionnal diversity indices show also a plateau around $70 m$.
-
-```{r 5 Com, fig.height=4}
-com <- CWT(Trees, com)
+```{r hypervolume, eval=TRUE, fig.cap='Functional hypervolume overlap of canopy and understorey species.', message=FALSE, warning=FALSE, include=FALSE}
+bd <- Q %>% 
+  filter(!is.na(WES)) %>% 
+  select(-LES, -WES, -SpCode, -strata) %>% 
+  summarise_all(funs(sd)) %>% 
+  unlist()
+CalcOverlap <- function(traits_strata){
+  hv <- lapply(traits_strata, function(strata) 
+    hypervolume::hypervolume(strata[names(bd)], kde.bandwidth = bd, verbose = F))
+  set <- hypervolume::hypervolume_set(hv$C, hv$U, check.memory = F, verbose = F)
+  hypervolume::hypervolume_overlap_statistics(set)
+}
+overlap <- Q %>% 
+  filter(!is.na(WES)) %>% 
+  select(-LES, -WES, -SpCode) %>% 
+  split(., .$strata) %>% 
+  CalcOverlap()
+cl <- makeCluster(cores)
+clusterExport(cl, list('Q', 'CalcOverlap', 'bd'))
+# n_overlap <- parSapply(cl, 1:n, function(i) {
+#   library(dplyr)
+#   Q %>%
+#     filter(!is.na(WES)) %>%
+#     select(-LES, -WES, -SpCode) %>%
+#     mutate(strata = sample(strata)) %>%
+#     split(., .$strata) %>%
+#     CalcOverlap()
+# })
+# stopCluster(cl)
+# save(n_overlap, file = file.path(path, 'overlap.Rdata'))
+load(file.path(path, 'overlap.Rdata'))
+hypervolume_fig <- data.frame(x = n_overlap[1,]) %>% 
+  ggplot(aes(x)) +
+  geom_density() +
+  geom_vline(xintercept = overlap[1],
+             col = 'red') +
+  ggtitle(paste('pvalue =', 
+                round((rank(c(overlap[1], n_overlap[1,]))/(n+1))[1], 3))) +
+  xlab('Overlap null distribution')
 ```
 
-```{r 6 lm & null models}
-reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
-LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
-                                                 SLA.cwm = lm(formula(paste('SLA.cwm', reg))), 
-                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
-                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
-                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
-                                                 LA.cwm = lm(formula(paste('LA.cwm', reg))),
-                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
-                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
-                                                 SLA.cwv = lm(formula(paste('SLA.cwv', reg))), 
-                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
-                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
-                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
-                                                 LA.cwv = lm(formula(paste('LA.cwv', reg))),
-                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
-LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
-# Global F
-Fval <- lapply(LM, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
-Fval <- lapply(Fval, function(x){do.call('rbind.data.frame', x)})
-Fval <- data.frame(Fval)
-names(Fval) <- names(LM)
-# Partial F
-# pFval <- lapply(LM, function(y){lapply(y, function(x){car::Anova(x)['F value']})})
+```{r hypervolume2, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
+data <- Q %>% 
+  filter(!is.na(WES)) %>% 
+  select(-LES, -WES, -SpCode) %>% 
+  split(., .$strata)
+hv <- lapply(data, function(strata) 
+    hypervolume::hypervolume(strata[names(bd)], kde.bandwidth = bd, verbose = F))
+set <- hypervolume::hypervolume_set(hv$C, hv$U, check.memory = F, verbose = F)
+hv <- list(C = set@HVList$Unique_1,
+           U = set@HVList$Unique_2,
+           Intersect = set@HVList$Intersection)
+plot(hypervolume::hypervolume_join(hv$C, hv$U))
+hv <- list(C = data$C[
+  hypervolume::hypervolume_inclusion_test(hv$C, data$C[names(bd)]),],
+           U = data$U[
+  hypervolume::hypervolume_inclusion_test(hv$U, data$U[names(bd)]),],
+           Intersect = do.call('rbind', data)[
+  hypervolume::hypervolume_inclusion_test(hv$Intersect, 
+                                          do.call('rbind', data)[names(bd)]),])
 
-n = 999
-nPFT <- rep(list(PFT_sp), n)
-nPFT <- lapply(nPFT, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
-nPFT <- lapply(nPFT, function(x){x$Sp_Code <- SpCode[row.names(x)] ; return(x)})
-nTrees <- rep(list(Trees[which(names(Trees) %in% c('SpCode', 'com', '2013_girth'))]), n)
-nTrees <- lapply(nTrees, function(x){x$ID <- row.names(x) ; return(x)})
-nTrees <- mapply(function(x,y){merge(x, y, by.x = 'SpCode', by.y = 'Sp_Code')}, x = nTrees, y = nPFT, SIMPLIFY = F)
-rm(nPFT)
-nTrees <- lapply(nTrees, function(x){row.names(x) <- x$ID ; return(x)})
-nTrees <- lapply(nTrees, function(x){x <- x[-which(names(x) == 'ID')] ; return(x)})
-ncom <- lapply(nTrees, function(x){CWT(x, com[[1]])})
-rm(nTrees)
-ncom <- lapply(ncom, `[`, names(ncom[[1]]))
-ncom <- apply(do.call(rbind, ncom), 2, as.list) 
-nLM <- rep(list(LM), n)
-nLM <- lapply(nLM, `[`, names(nLM[[1]]))
-nLM <- apply(do.call(rbind, nLM), 2, as.list)
-ncom <- lapply(ncom, function(c){lapply(c, function(x){x <- rep(list(x), 14) ; return(x)})})
-nLM <- mapply(function(x,y){
-  mapply(function(r,s){
-    mapply(function(p,q){
-      lm(p$call,q)
-    }, p = r, q = s, SIMPLIFY = F)
-  }, r = x, s = y, SIMPLIFY = F)
-}, x = nLM, y = ncom, SIMPLIFY = F)
-rm(ncom) ; gc()
-
-# Global model
-nFval <- lapply(nLM, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
-nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
-nFval <- lapply(nFval, function(f){lapply(f, unlist)})
-
-# nDev <- lapply(nLM, function(lm){lapply(lm, function(y){lapply(y, function(x){deviance(x)})})})
-# nDev <- lapply(nDev, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
-# nDev <- lapply(nDev, function(f){lapply(f, unlist)})
-# invisible(mapply(function(x,y,z){plot(x, y, xlab = 'Fval', ylab = 'Dev', main = z)}, x = nFval$abundance, y = nDev$abundance, z = as.list(names(nFval$abundance))))
-
-pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
-row.names(pval) <- row.names(Fval)
-
-# Partial contribution
-
+hypervolume_table <- kable(do.call('rbind', lapply(hv, function(x)apply(x[names(bd)], 2, mean))),
+      caption = 'means of hypervolumes functional traits')
+rm(hv, set, data, CalcOverlap, bd)
 ```
 
-Table 4: Summary for linear models and null models for traits community weighted means.
-_Linear model results for each trait community weighted means (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model deviance was compared to 999 null model deviances in which traits were randomized among species.
-See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
+## Functional composition of guilds
 
-```{r CWM}
-kable(regression_table(LM, reg, pval, weight = 'abundance', metric = '.cwm'))
-kable(regression_table(LM, reg, pval, weight = 'presence-absence', metric = '.cwm'))
+### Guild weighted means : nul model 1
+
+We used a null model (NM1) which randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix) [@bernard-verdier_community_2012].
+
+The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds (see table \@ref(tab:GWM1)).
+Results in presence-absence are in adequation with observed results in abundance (see **...**).
+Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits, and two major differences were observed between the two guilds.
+The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height affected only understory guild composition.
+
+Focusing on canopy tree species, we found an increase in wood density under south-west orientation, both in abundance and presence-absence.
+We also observed a decrease in leaf area and thickness under the south-west orientation, but only in abundance and presence-absence respectively.
+Leaf area decreased with curvature in both abundance and presence-absence, while wood density and LDMC increased and decreased respectively with curvature in presence-absence only.
+
+For understorey tree species, SLA significantly decreased and wood density and leaf thickness significantly increased with curvature, both in abundance and presence-absence.
+We also detected a significant increase and decrease in wood economic spectrum and leaf area respectively with curvature in abundance.
+SLA in both abundance and presence-absence significantly increased with canopy height, while LDMC and wood density decreased with canopy height in abundance only.
+
+```{r GWM1}
+GWM <- L %>% 
+  t() %>% 
+  data.frame() %>% 
+  mutate(SpCode = row.names(.)) %>% 
+  left_join(Q) %>% 
+  select(-SpCode) %>% 
+  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+       variable.name = 'community', value.name = 'abundance') %>% 
+  group_by(strata, community) %>% 
+  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# nQ <- array(rep(as.matrix(Q), n),
+#             c(dim(Q), n),
+#             list(row.names(Q), names(Q), 1:n))
+# nQ <- array(apply(nQ, 3, function(x){
+#   x[,'SpCode'] <- sample(x[,'SpCode'])
+#   return(x)}), c(dim(Q), n),
+#   list(row.names(Q), names(Q), 1:n))
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('L', 'nQ'))
+# nGWM <- parApply(cl, nQ, 3, function(x){
+#   library(dplyr)
+#   library(reshape2)
+#   L %>%
+#     t() %>%
+#     data.frame() %>%
+#     mutate(SpCode = row.names(.)) %>%
+#     left_join(data.frame(x)) %>%
+#     select(-SpCode) %>%
+#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+#          variable.name = 'community', value.name = 'abundance') %>%
+#     group_by(strata, community) %>%
+#     mutate_each(funs(as.character), -abundance) %>%
+#     mutate_each(funs(as.numeric), -abundance) %>%
+#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# })
+# stopCluster(cl)
+# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
+#               c(dim(GWM[3:9]), n),
+#               list(row.names(GWM), names(GWM[3:9]), 1:n))
+# GWM_ses <- GWM
+# GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
+# rm(cl, nQ, nGWM)
+# save(GWM_ses, file = file.path(path, 'GWM1_ses.Rdata'))
+load(file.path(path, 'GWM1_ses.Rdata'))
+GWM_ses %>% 
+  melt(id.vars = c('strata', 'community')) %>% 
+  rename(trait = variable) %>% 
+  rename(SES = value) %>% 
+  left_join(R) %>%
+  select(-community) %>% 
+  group_by(strata, trait) %>% 
+  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
+          direction = 'both', trace = 0))) %>% 
+  group_by(strata, trait, term) %>%
+  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
+                          stars(p.value, ns = '   '))) %>% 
+  dcast(strata + trait ~ term) %>% 
+  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
+  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
+  filter(trait != 'LES') %>% 
+  filter(trait != 'WES') %>%
+  bold() %>% 
+  kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
 ```
 
-Table 5: Summary for linear models and null models for traits community weighted variances.
-_Linear model results for each trait community weighted variances (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA).
-Model deviance was compared to 999 null model deviances in which traits were randomized among species.
-See table 1 for abbreviations.
-The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._
-
-```{r CWV}
-kable(regression_table(LM, reg, pval, weight = 'abundance', metric = '.cwv'))
-kable(regression_table(LM, reg, pval, weight = 'presence-absence', metric = '.cwv'))
+```{r WD, fig.cap='Exemple of standard effect size (SES) of wood density  (WD) for canopy (C) tree species.
+Each subplot show  standard effect size (SES) of wood density (WD) for canopy (C) tree species along one environmental variable gradient.
+Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
+See table \\@ref(tab:Traits) for abbreviation.'}
+GWM_ses %>% 
+  select(strata, WD, community) %>% 
+  left_join(R) %>% 
+  filter(strata == 'C') %>% 
+  ungroup() %>% 
+  select(-strata, -community) %>% 
+  melt(id.vars = 'WD') %>% 
+  ggplot(aes(value, WD)) +
+  geom_point() +
+  facet_wrap(~ variable, scales = "free") +
+  geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') +
+  geom_hline(yintercept = 0) +
+  xlab('Environmental gradient') +
+  ylab('Standard effect size of wood density')
 ```
 
-## Phylogenetic signal
+### Guild weighted means : nul model 3
 
-```{r PhyloMaker}
-library(phylosignal)
-library(phytools)
-library(phylobase)
+We used a null model (NM3) which randomized species among guilds in the whole plot (see table \@ref(tab:GWM3)).
 
-# Building species list
-l <- PFT_sp
-# setdiff(row.names(l), Species$LatinName)
-row.names(l)[which(row.names(l) == 'Actinodaphne tadulingami')] <- 'Actinodaphne tadulingamii'
-Species['agin',] <- c('agin', "Agrostistachys indica", 'Euphorbiaceae', NA, NA)
-Species['anme',] <- c('anme', "Antidesma menasu", 'Euphorbiaceae', NA, NA)
-Species['anmo',] <- c('anme', "Antidesma montanum", 'Euphorbiaceae', NA, NA)
-row.names(l)[which(row.names(l) == 'Blepharistemma membranifolia')] <- 'Blepharistemma serratum '
-row.names(l)[which(row.names(l) == 'Casearia ovata')] <- 'Casearia ovata '
-row.names(l)[which(row.names(l) == 'Cinnamomum malabathrum')] <- 'Cinnamomum malabatrum'
-row.names(l)[which(row.names(l) == 'Cinnamomum sp1')] <- 'Cinnamomum sp.'
-row.names(l)[which(row.names(l) == 'Dimocarpus longan')] <- 'Dimocarpus longan '
-row.names(l)[which(row.names(l) == 'Drypetes confertiflorus')] <- 'Drypetes confertiflora'
-row.names(l)[which(row.names(l) == 'Glochidion malabaricum')] <- 'Glochidion ellipticum'
-row.names(l)[which(row.names(l) == 'Homalium zeylanicum')] <- 'Homalium zeylanicum '
-row.names(l)[which(row.names(l) == 'Ixora nigricams')] <- 'Ixora nigricans  '
-row.names(l)[which(row.names(l) == 'Knema attenuata')] <- 'Knema attenuata  '
-row.names(l)[which(row.names(l) == 'Leptonychia moacurroides')] <- 'Leptonychia caudata '
-row.names(l)[which(row.names(l) == 'Myristica malabarica')] <- 'Myristica malabarica '
-row.names(l)[which(row.names(l) == 'Olea dioica')] <- 'Olea dioica '
-row.names(l)[which(row.names(l) == 'Reinwardtiodendron anaimalaiense')] <- 'Reinwardtiodendron anamalaiense'
-row.names(l)[which(row.names(l) == 'Strombosia ceylanica')] <- 'Strombosia ceylanica '
-row.names(l)[which(row.names(l) == 'Symplocos cochinensis')] <- 'Symplocos cochinchinensis'
-row.names(l)[which(row.names(l) == 'Syzygium laetum')] <- 'Syzygium laetum '
-row.names(l)[which(row.names(l) == 'Vepris bilocularis')] <- 'Vepris bilocularis '
-sp <- Species[which(Species$LatinName %in% row.names(l)), 2:3]
-sp$genus <- unlist(lapply(strsplit(sp$LatinName, ' ', fixed = T), function(x){x[1]}))
-sp$species <- sub(' ', '_', sp$LatinName, fixed = T)
-sp$family <- sp$Family
-sp <- sp[c('species', 'genus', 'family')]
-sp[which(sp$family == 'Flacourtiaceae'),3] <- 'Salicaceae'
-sp[which(sp$family == 'Steruliaceae'),3] <- 'Malvaceae'
-sp[which(sp$family == 'Cesaliniaceae'),3] <- 'Fabaaceae'
-sp['hepa','family'] <- 'Malvaceae'
-# sp <- sp[-which(row.names(sp) == 'hubr'),]
-# sp <- sp[-which(row.names(sp) %in% levels(droplevels(PFT_sp[which(is.na(PFT_sp$WD)),1]))),]
+We found that **curvature** was responsible for the guilds differenciation for most of the traits.
+Increased curvature resulted in denser lwood and leaves with high SLA in canopy, against lighter wood, low SLA, but bigger and thicker leaves in understorey.
+Some curvature results were congruent with **Wetness** variable.
+**South-West** orientation diminished the leaf area while increasing leaf density for canopy tree species.
+Finally, **Canopy** height and **basal area** affected in opposite ways SLA.
+Increased canopy height and decreased basal area resulted in lower SLA in canopy and increased SLA in understorey.
 
-# Getting phylogeny
-path <- system.file('phylogeny', package = 'Uppangala')
-source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
-phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
-nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
-results <- S.PhyloMaker(sp, phylo, nodes)
-tree <- results$Scenario.1
+```{r GWM3}
+GWM <- L %>% 
+  t() %>% 
+  data.frame() %>% 
+  mutate(SpCode = row.names(.)) %>% 
+  left_join(Q) %>% 
+  select(-SpCode) %>% 
+  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+       variable.name = 'community', value.name = 'abundance') %>% 
+  group_by(strata, community) %>% 
+  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# nQ <- array(rep(as.matrix(Q), n),
+#             c(dim(Q), n),
+#             list(row.names(Q), names(Q), 1:n))
+# nQ <- array(apply(nQ, 3, function(x){
+#   x[,'strata'] <- sample(x[,'strata'])
+#   return(x)}), c(dim(Q), n),
+#   list(row.names(Q), names(Q), 1:n))
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('L', 'nQ'))
+# nGWM <- parApply(cl, nQ, 3, function(x){
+#   library(dplyr)
+#   library(reshape2)
+#   L %>%
+#     t() %>%
+#     data.frame() %>%
+#     mutate(SpCode = row.names(.)) %>%
+#     left_join(data.frame(x)) %>%
+#     select(-SpCode) %>%
+#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+#          variable.name = 'community', value.name = 'abundance') %>%
+#     group_by(strata, community) %>%
+#     mutate_each(funs(as.character), -abundance) %>%
+#     mutate_each(funs(as.numeric), -abundance) %>%
+#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# })
+# stopCluster(cl)
+# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
+#               c(dim(GWM[3:9]), n),
+#               list(row.names(GWM), names(GWM[3:9]), 1:n))
+# GWM_ses <- GWM
+# GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
+# rm(cl, nQ, nGWM)
+# save(GWM_ses, file = file.path(path, 'GWM3_ses.Rdata'))
+load(file.path(path, 'GWM3_ses.Rdata'))
+GWM_ses %>% 
+  melt(id.vars = c('strata', 'community')) %>% 
+  rename(trait = variable) %>% 
+  rename(SES = value) %>% 
+  left_join(R) %>%
+  select(-community) %>% 
+  group_by(strata, trait) %>% 
+  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
+          direction = 'both', trace = 0))) %>% 
+  group_by(strata, trait, term) %>%
+  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
+                          stars(p.value, ns = '   '))) %>% 
+  dcast(strata + trait ~ term) %>% 
+  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
+  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
+  filter(trait != 'LES') %>% 
+  filter(trait != 'WES') %>%
+  bold() %>% 
+  kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
 ```
 
-```{r Phylosignal}
-data <- l
-row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
-tips <- row.names(data)[which(is.na(l$WD))]
-phy <- list(full = drop.tip(tree, tips), `-WD` = tree)
-data <- list(full = data[phy$full$tip.label,c(2:5,7)], `-WD` = data[phy$`-WD`$tip.label,c(3:5,7)])
-p4d <- mapply(function(p,d){phylo4d(p,d)}, p = phy, d = data)
-# lapply(p4d, barplot)
-phyloS <- lapply(p4d, phyloSignal, methods = 'I')
-phyloS <- lapply(phyloS, function(x){cbind(x$stat, x$pvalue)})
-phyloS <- rbind(phyloS$`-WD`, phyloS$full['WD',])
-names(phyloS) <- c('Moran\'s I', 'p-value')
-phyloS$` ` <- lapply(as.list(phyloS$`p-value`), stars)
-phyCor <- mapply(function(d,p){lapply(as.list(names(d)), function(x){phyloCorrelogram(p, trait = as.character(x))})}, d = data, p = p4d)
-phyCor <- c(phyCor$`-WD`, list(phyCor$full[[1]]))
-names(phyCor) <- unlist(lapply(phyCor, function(x){x$trait}))
-lipa <- lapply(p4d, lipaMoran)
-# mapply(function(x,y){barplot.phylo4d(x, bar.col=(y$p.value < 0.05) + 1)}, x = p4d, y = lipa)
-kable(phyloS)
-```
+### Guild weighted variances : nul model 2
 
-## Are communities spatially structured ?
+The second model (NM2) randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012].
 
-```{r Functionnal spatial structure}
-library(spacodiR)
+We found significant relationships between the standardized effect size (SES) of GWVs and the environmental predictors (see table \@ref(tab:GWV3)).
+**Wetness** decreased variance of wood density in canopy and all traits in understorey.
+**Curvature** increased LDMC variance in canopy.
+**Sout-West** orientation results in decreased variance of density of wood and leaf in understorey.
+Finally, **Canopy** hieght diminished variance of leaf are and density in canopy whereas **basal aea** increased variance of SLA in both strata and increased leaf area variance in canopy.
 
-# Traits structure
-Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
-Lm[is.na(Lm)] <- 0
-Lm.l <- list(full = Lm, `-WD` = Lm)
-Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
-row.names(Qm) <- PFT_sp$Sp_Code
-Qm.l <- list(full = Qm, `-WD` = Qm[-3])
-Qm.l <- mapply(function(x,y){x[intersect(row.names(x), names(y)),]}, x = Qm.l, y = Lm.l, SIMPLIFY = F)
-Qm.l$full <- Qm.l$full[-which(is.na(Qm.l$full$WD)),]
-Lm.l <- mapply(function(x,y){y[intersect(row.names(x), names(y))]}, x = Qm.l, y = Lm.l, SIMPLIFY = F)
-spafun <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), sp.traits = x)}, x = Qm.l, y = Lm.l, SIMPLIFY = F)
-spafun <- do.call('rbind.data.frame', lapply(spafun, function(x){x[1:4]}))
-
-Qm.lRD <- lapply(Qm.l, function(x){rep(list(x),999)})
-Qm.lRD <- lapply(Qm.lRD, function(x){lapply(x, function(y){y[sample(row.names(y)),]})})
-Lm.lRD <- lapply(Lm.l, function(x){rep(list(x),999)})
-Lm.lRD <- mapply(function(x,y){mapply(function(r,s){r[row.names(s)]}, r = x, s = y, SIMPLIFY = F)}, x = Lm.lRD, y = Qm.lRD, SIMPLIFY = F)
-spafun.l <- mapply(function(x,y){mapply(function(r,s){spacodi.calc(sp.plot = t(s), sp.traits = r)}, r = x, s = y, SIMPLIFY = F)}, x = Qm.lRD, y = Lm.lRD, SIMPLIFY = F)
-spafun.l <- lapply(spafun.l, function(x){lapply(x, function(y){y[1:4]})})
-spafunRD <- list(Ist = lapply(spafun.l, function(x){unlist(lapply(x, function(y){y$Ist}))}),
-                 Tst = lapply(spafun.l, function(x){unlist(lapply(x, function(y){y$Tst}))}),
-                 Ust = lapply(spafun.l, function(x){unlist(lapply(x, function(y){y$Ust}))}),
-                 TAUst = lapply(spafun.l, function(x){unlist(lapply(x, function(y){y$TAUst}))}))
-pvalues <- mapply(function(obs,rd){
-  mapply(function(o,r){rank(c(o, r))[1]/1000}, o = obs, r = rd)
-  }, obs = lapply(as.list(spafun), as.list), rd = spafunRD)
-row.names(pvalues) <- row.names(spafun)
-
-pvalues <- as.data.frame(pvalues)[3:4]
-pvalues$Ust <-  1- pvalues$Ust
-spafun.res <- format(spafun[3:4], scientific = T, digits = 2)
-pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
-spafun.res$Ust <- paste(spafun.res$Ust, '/', pvalues$Ust)
-spafun.res$TAUst <- paste(spafun.res$TAUst, '/', pvalues$TAUst)
-kable(spafun.res)
-```
-
-```{r Phylogenetic spatial structure}
-for(i in 1:2) phy[[i]]$tip.label <- gsub('_', ' ', phy[[i]]$tip.label, fixed = T)
-for(i in 1:2) phy[[i]]$tip.label <- as.character(l$Sp_Code[match(phy[[i]]$tip.label, row.names(l))])
-Lm <- lapply(phy, function(x){Lm[intersect(names(Lm), x$tip.label)]})
-phyRD <- lapply(phy, function(x){x$node.label <- NULL ; return(x)})
-phyRD <- lapply(phyRD, function(x){rep(list(x), 999)})
-phyRD <- lapply(phyRD, function(x){lapply(x, resamp.phy)})
-LmRD <- lapply(Lm, function(x){rep(list(x), 999)})
-LmRD <- mapply(function(x,y){mapply(function(r,s){r[intersect(names(r), s$tip.label)]}, r = x, s = y, SIMPLIFY = F)}, x = LmRD, y = phyRD, SIMPLIFY = F)
-phyRD <- mapply(function(x,y){mapply(function(r,s){drop.tip(r, setdiff(r$tip.label, names(s)))}, r = x, s= y, SIMPLIFY = F)}, x = phyRD, y = LmRD, SIMPLIFY = F)
-
-spaphy <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), phy = x)}, x = phy, y = Lm.l, SIMPLIFY = F)
-spaphy <- do.call('rbind.data.frame', lapply(spaphy, function(x){x[1:4]}))
-
-spaphy.l <- mapply(function(x,y){mapply(function(r,s){spacodi.calc(t(r), s)}, r = x, s = y, SIMPLIFY = F)},x = LmRD, y = phyRD, SIMPLIFY = F)
-spaphy.l <- lapply(spaphy.l, function(x){lapply(x, function(y){y[1:4]})})
-spaphyRD <- list(Ist = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Ist}))}),
-                 Tst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Tst}))}),
-                 Bst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Bst}))}),
-                 PIst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$PIst}))}))
-pvalues <- mapply(function(obs,rd){
-  mapply(function(o,r){rank(c(o, r))[1]/1000}, o = obs, r = rd)
-  }, obs = lapply(as.list(spaphy), as.list), rd = spaphyRD)
-row.names(pvalues) <- row.names(spaphy)
-
-pvalues <- as.data.frame(pvalues)[3:4]
-pvalues$Bst <-  rep(NaN,2)
-spaphy.res <- format(spaphy[3:4], scientific = T, digits = 2)
-pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
-spaphy.res$Bst <- paste(spaphy.res$Bst, '/', pvalues$Bst)
-spaphy.res$PIst <- paste(spaphy.res$PIst, '/', pvalues$PIst)
-kable(spaphy.res)
-
-# permut <- list(Bst = mapply(function(x,y){spacodi.by.nodes(sp.plot = t(y), x, sp.parm = "Bst", method = "1s")}, x = phy, y = Lm.l, SIMPLIFY = F), PIst = mapply(function(x,y){spacodi.by.nodes(sp.plot = t(y), x, sp.parm = "PIst", method = "1s")}, x = phy, y = Lm.l, SIMPLIFY = F))
-# par(mfrow=c(2,2)) ; invisible(lapply(permut, function(x){lapply(x, spacodi.permutplot)}))
+```{r GWV3}
+GWV <- L %>% 
+  t() %>% 
+  data.frame() %>% 
+  filter(sum(.) > 2) %>% 
+  mutate(SpCode = row.names(.)) %>% 
+  left_join(Q) %>% 
+  select(-SpCode) %>% 
+  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+       variable.name = 'community', value.name = 'abundance') %>% 
+  melt(id.vars = c('strata', 'community', 'abundance'),
+       variable.name = 'trait', value.name = 'trait_value') %>% 
+  left_join(melt(GWM, id.vars = c('strata', 'community'),
+                 variable.name = 'trait', value.name = 'GWM')) %>% 
+  group_by(strata, community, trait) %>% 
+  mutate(GWV = (trait_value - GWM)^2) %>% 
+  select(-trait_value, -GWM) %>% 
+  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
+  dcast(strata + community ~ trait)
+# nL <- array(replicate(n, t(apply(L, 1, function(x){
+#   x[x != 0] <- sample(x[which(x != 0)]) ; return(x)}))),
+#   c(dim(L), n),
+#   list(row.names(L), names(L), 1:n))
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('nL', 'Q', 'GWM'))
+# nGWV <- parApply(cl, nL, 3, function(x){
+#   library(dplyr)
+#   library(reshape2)
+#   x %>%
+#     t() %>%
+#     data.frame() %>%
+#     filter(sum(.) > 2) %>%
+#     mutate(SpCode = row.names(.)) %>%
+#     left_join(Q) %>%
+#     select(-SpCode) %>%
+#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+#          variable.name = 'community', value.name = 'abundance') %>%
+#     melt(id.vars = c('strata', 'community', 'abundance'),
+#          variable.name = 'trait', value.name = 'trait_value') %>%
+#     left_join(melt(GWM, id.vars = c('strata', 'community'),
+#                    variable.name = 'trait', value.name = 'GWM')) %>%
+#     group_by(strata, community, trait) %>%
+#     mutate(GWV = (trait_value - GWM)^2) %>%
+#     select(-trait_value, -GWM) %>%
+#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%
+#     dcast(strata + community ~ trait)
+# })
+# stopCluster(cl)
+# nGWV <- array(unlist(lapply(nGWV, function(x) as.matrix(x[3:9]))),
+#               c(dim(GWV[3:9]), n),
+#               list(row.names(GWV), names(GWV[3:9]), 1:n))
+# GWV_ses <- GWV
+# GWV_ses[3:9] <- (GWV[3:9] - apply(nGWV, 1:2, mean, na.rm = T))/apply(nGWV, 1:2, sd, na.rm = T)
+# rm(cl, nL, nGWV)
+# save(GWV_ses, file = file.path(path, 'GWV_ses.Rdata'))
+load(file.path(path, 'GWV_ses.Rdata'))
+GWV_ses %>% 
+  melt(id.vars = c('strata', 'community')) %>% 
+  rename(trait = variable) %>% 
+  rename(SES = value) %>% 
+  left_join(R) %>%
+  select(-community) %>% 
+  group_by(strata, trait) %>% 
+  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
+          direction = 'both', trace = 0))) %>% 
+  group_by(strata, trait, term) %>%
+  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
+                          stars(p.value, ns = '   '))) %>% 
+  dcast(strata + trait ~ term) %>% 
+  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
+  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
+  filter(trait != 'LES') %>% 
+  filter(trait != 'WES') %>%
+  bold() %>% 
+  kable(caption = 'Linear models of the standardized effect sizes (SES) of GWV for canopy (C) and understorey (U) guilds.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and  number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
 ```
@@
 
+## **"Take home messages" to add :**
+
+- **Pantropical and multiscale WES and LES ?**
+- **Decoupled wood and leaf strategies among canopy and understorey tree species**
+- **WES among canopy tree species**
+- **LES among understorey tree species**
+
+We found back economics spectra evidenced by @wright_worldwide_2004 and @chave_towards_2009 at finer scale than @messier_trait_2016 suggested (figure 1).
+This result advocate for a pantropical and multiscale oganisation of tropical plant strategies in two separated economics spectra.
+Still our results suggest that this two spectra are decoupled between canopy and understorey tree species, whereas existing in both guilds.
+Canopy tree species being the first layer of forest ecosystem interacting with external abiotic environment shows major variation in wood economics spectrum (table 3).
+Indeed wood economic spectrum is related to mechanical resistance, disturbance tolerance and water conductivity efficiency.
+Thus understorey tree will benefit from a mircoclimate **[ref?]** generated by the presence of the canopy tree.
+In this microclimate, leaf economics spectrum is then a major axis of strategy to answer to a changing light environment and soil fertility.
+
+## Wood and leaf economics hold at the local scale
+
+We found an increasing trait variation within individuals, within species, and between species (table 2).
+Our analyses thus reveal that the primary source of trait variation within and among communities is the inter-species variation, justifying the use of species mean traits for the study of tropical tree guild assembly.
+The PCA of functional trait variation among species displayed two main dimensions (figure 1), very similar to the ones found by @baraloto_decoupled_2010 and @fortunel_leaf_2012.
+Those authors interpreted these axes as independent wood and leaf economics spectra.
+We found that SLA and LDMC tend to covary negatively along the leaf economics spectrum, which be seen as a trade-off between productivity, with less carbon investment and higher photosynthesis rates, and persistence, with more leaf defense and investment [@wright_worldwide_2004].
+A second major axis of trait variation opposed wood density to leaf area and thickness.
+It may be interpreted as an axis opposing stress tolerant species, with denser wood, to competitive species, with thicker and larger leaves, conferring a higher efficiency in water conductivity **[ref?]**.
+It may thus be interpreted as the wood economics spectrum firstly introduced by @chave_towards_2009, and then discussed by @baraloto_decoupled_2010.
+Interestingly, @baraloto_decoupled_2010 found an independence between the leaf and wood economics spectra and their wood economics spectrum also showed an opposition between hardwood species having small thin leaves and light wood species having large thick leaves.
+
+Furthermore, our results are similar to the ones found by @fortunel_leaf_2012, where basic PCA axes also represented leaf and wood economics spectra (inverted compare to us).
+Noticeably, our analysis integrated less traits (5 against 14), less species (89 against 758), and considered a set of species found in a local forest-plot while @fortunel_leaf_2012 considered a regional set of species from French Guiana and Peru (i.e.
+the opposite ends of Amazonia).
+Therefore, our results shows that the economics spectra found at a regional scale in the Neotropics are also valid at a far more local scale in an Indian rainforest.
+This result does support the arguments by @messier_trait_2016 and suggests that the two observed spectra underline major ecological strategies in tropical forests, even locally.
+
+## Canopy and understorey species assembly rely on different economics spectra
+
+We found a significanlty higher variance of canopy tree species around the wood economic spectrum (WES) rather than the leaf economic spectrum (LES).
+Similarly understorey tree species seemed to have higher variance around leaf economic spectra (LES) rather than wood economics spectra (WES), besides it was not significant at a level $\alpha = 0.05$ (table 3).
+Those results advocates for decoupled strategies between species of understorey and canopy guilds [@duque2002different].
+Indeed canopy tree species could deploy strategies to endorse the role of first layer against external abiotic environment.
+Those strategies will come through wood economic spectrum (WES) to ensure mechanical resistance, disturbance tolerance and water conductivity efficiency **[biblio ?]**.
+On the other hand, understorey tree species will encounter constraints directly linked to canopy guild limiting light availability.
+Thus they will need to deploy strategies linked to ressource acquisition coming through leaf economics spectrum (LES) **[biblio ?]**.
+
+Furthermore, we analyzed functional composition of $30*30m^2$ guilds from canopy and understorey, as well as the variation of this composition with environment across a 10-ha forest plot.
+We foud that the variations in trait composition along the environmental gradients differ between canopy and understorey species, with a marked variation in traits associated to the leaf economic spectrum (LES) for understorey species and to the wood economics scpectrum (WES) for canopy species.
+In the former case, the curvature and the canopy height both had a significant effect on the LES-based composition of understorey tree species, whereas southwest orientation and curvature both had a significant effect on the WES-based composition of canopy tree species (table 4).
+
+## Environment influence guild-level functionnal composition
+
+We should note that relations found between guilds weighted means (GWM) and environment are globally weak (table 4, most of $R^2 \leq 0.1$, and see supplementary materials S4).
+This low explanatory power is advocating for underlying null process related to neutral theory [@hubbell_unified_2001].
+Null process can effectivelly be responsible for the majority of observed variance in functionnal composition at a such fine scale **[ref?]**.
+Studies highlighting environmental filtering effect often present a higher environmental variation than in our study site [@bernard-verdier_community_2012].
+In parallel, the 10*ha plot scale is really low and partially responsible for the small environmental variation.
+@kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25 ha tropical forest plots including large flat valleys.
+
+Curvature is affecting the variation of leaf economics in understorey species and the variation of wood economics in canopy tree species.
+A locally convex context would accumulate more soil and attract more water, increasing soil fertility and allowing persistence of species investing more in leaf resource acquisition in understorey with higher SLA and lower LDMC, or less in wood ressources in canopy with lighter wood and bigger leaves converging more toward an homogeneous guild composition (lower CWV, unshown result).
+Curvature for some functional traits, as wood density, LDMC and SLA, is also doubled by a similar effect of the wetness variable expressing also the water accumulation in convex context resulting in increased soil fertility.
+Economics spectra are not perfectly divided between canopy and understorey, with wood and leaf respectivelly, for instance LDMC of canopy species also decreased with curvature whereas WD increased in understorey.
+
+The influence of aspect on forest composition is little documented in tropical forests.
+In a similar context that our study site, @gunatilleke_specieshabitat_2006 gave some intuitive insights on the effect of aspect on forest composition: he suggested that aspect through monsoon had a potential effect on rainforests.
+@bunyan_effect_2015 tested the effect of aspect on the occurence of forest fragments in a forest-grassland mosaic and foud that southwest orientation strongly determined the presence of forest in mountain rainforests of Western Ghats.
+They suggested that it could result from monsoon winds.
+In our context, monsoon winds are south-west oriented, and are expected to be a major source of disturbance in forests [@soman_aspects_1990].
+
+The Uppangala plot is situated in a topographically complex area with steep slopes mostly organized along southwest-oriented crests.
+In this context it is diffcult to predict wind flow and orientation inside the 10-ha plot topography.
+We propose two explanation, hardwood could be selected for mechanical resistance in more disturbed souwhtwest oriented area or topography provoke a turbulences on northeast oriented areas responsible of more disturbance and death events in the forest history resulting in lightwood filtering [@larjavaara_rethinking_2010].
+In both case it seems reasonable that canopy tree species are the most influenced by disturbance events linked to monsoon, and may even buffer the effect on understorey species.
+
+Biotic factors, illustrated here by the canopy height and the basal area, also had a significant effect on the LES-based composition of understorey tree species.
+We indeed found higher SLA values and lower LDMC values in understorey assemblages that are located under stands characterized by a tall canopy and a high basal area.
+This results contrasts with the expectation that understorey species invest in higher LDMC for higher leaf lifespan, at the expense of SLA, unde light-limited conditions [@kitajima_tissue-level_2010].
+The Uppangala forest is characterized by a large heterogeneity of canopy height allowing light to penetrate more in lower strata, maybe especially more in high canopy context covering less closely understorey, resulting in those observation on the leaf economics variation of understorey species.
+**Not convincing, better explanation ?**
+
 # *Acknowledgement*
+
+* French Institute of Pondicherry (FIP)
+* Cedric Vega for the extension of the plot
 
@@
 
-*I'm aware that Supplementary materials goes after References (as figure and tables), but it's not possible for the moment with Rmarkdown and knitr.*
+## S1: Principal component analysis of topographical variables
 
-## S1: Structural equation model
-
-```{r SEM}
-library(lavaan)
-m1 <- '
-Abiotic =~ Slope + Curvature + Wetness + SW
-Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
-Biotic =~ BA + Canopy
-Traits ~ Abiotic + Biotic
-'
-m2 <- '
-Abiotic =~ Slope + Curvature + Wetness + SW
-Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
-Biotic =~ BA + Canopy
-Biotic ~ Traits + Abiotic
-'
-fit1 <- sem(m1, data = com$abundance@data)
-fit2 <- sem(m2, data = com$abundance@data)
-semPlot::semPaths(fit1,
-                  what = 'col',
-                  whatLabels = 'std',
-                  style = 'lisrel',
-                  residuals = T,
-                  nCharNodes = 0,
-                  nCharEdges = 0,
-                  groups='manlat',
-                  pastel = T,
-                  ask = F)
+```{r PCAenv, fig.cap='Principal component analysis (PCA) performed on topgraphical variables. SW stand for southwesterness.'}
+pca.env %>% 
+  autoplot(colour = 'grey', loadings.label.size = 6,
+           loadings.label.colour = 'black',
+           loadings = T, loadings.label = T, 
+           loadings.colour = 'black',
+           loadings.label.vjust = 2) +
+  coord_equal() +
+  geom_hline(yintercept = 0, linetype = 'dashed') +
+  geom_vline(xintercept = 0, linetype = 'dashed') +
+  xlab('Axe 1 − 42.52 %') + ylab('Axe 2 − 23.44 %')
 ```
 
-Supplementary material S1: Path diagram for structural equation modelling.
-*Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic.
-Arrow values are indicating standardized parameter estimates*.
+Curvature covaries with profile curvature, plan curvature, distance to potential rivers and the PCA first axis.
+Elevation and wetness are opposed in the principal component analysis and slope and south-westerness does not seem to covary much with other variables **Not precise**.
 
-## S3: Phylogenetic signal among functionnal traits by correlogram
+## S2: Hypervolumes of functional guilds
 
-Supplementary materials S3: Correlogram of phylogenetic signal among the 5 functionnal traits.
-```{r phylosignal plots, eval=FALSE, include=FALSE}
-par(mfrow = c(2,4))
-for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
+```{r hypervfig, fig.cap='Functional hypervolume overlap of canopy and understorey species.'}
+hypervolume_fig
+```
+
+```{r hypervolume_tab}
+hypervolume_table
 ```

### 2018-01-24 inst/doc/draft2.Rmd
Uppangala 0.1.9038 January 2018 Tabebuia run

@@
     toc_float: yes
+  bookdown::pdf_document2:
+    number_sections: false
 csl: mee.csl
@@
 library(ggfortify)
-cores <- 16
+cores <- 3
 opts_chunk$set(
@@
 path <- '~/Documents/IFP/Uppangala/inst/doc/draft_save/'
-n = 999 # Random repetitions
+n = 3 # Random repetitions
 ```
@@
 - Does functional traits variation and covariation reflect basic dimensions of ecological strategy in the rainforest ?
+- Does local functional composition from guilds differ ?
 - Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?
@@
 
-**We need a new paragraph to clearly present each null model, its implications and use for subsequent analysis.**
+To test for any significant changes in functional composition, we devised a null model approach. We thus explored following questions with subsequent hypothesis:
 
-To test for any significant changes in functional composition among guilds, we devised a null model approach.
-A null model (NM1) randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix) [@bernard-verdier_community_2012].
-Note that, spatial correlations in tree abundance are conserved limiting associated type I error.
-We performed 999 randomizations of null model NM1 and we derived a two-tailed test of the proportion of guilds deviating from the null expectation, with adjusted p-values by Benjamini-Hochberg correction **ref**.
-Shifts in GWM values with changing environmental conditions could reflect the shift of species strategies across assemblages [@ackerly_trait-based_2007].
+* __Q1__ Does local functional composition from guilds differ ?
+    * ___H1a__ Environmental filtering differ in selection of local functional composition between canopy and understorey species._
+    * ___H1b__ Environmental filtering strenght differ locally between canopy and understorey species._
+* __Q2__ Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?
+    * ___H2a__ Drivers of environmental filtering on local functional composition differ between canopy and understorey tree species._ 
+    * ___H2b__ Strength of drivers of evironmental filtering differ between canopy and understorey tree species._
+
+To accept or reject hypothesis we used three null models :
+
+* __NM1__ randomized species among guilds. Consequently, NM1 breaks the link between species and guilds while keeping community structure, thus testing for hypothesis __H1a__.
+* __NM2__ randomized trait values among species in the whole plot for each guild (permuting raw in `Trait*Species` matrices of both guilds) [@bernard-verdier_community_2012].
+Consequently, NM2 breaks the link between environement and local functional composition while keeping guilds and species spatial structure, thus testing for hypothesis __H2a__.
+* __NM3__  randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012].
+Consequently NM3, breaks the link between environment and species abundances while keeping guilds, community, and species spatial structure, thus testing for hypothesis __H1b__ and __H2b__.
+
+We performed 999 randomizations of each null model and we derived a two-tailed test of the proportion of guilds deviating from the null expectation, with adjusted p-values by Benjamini-Hochberg correction **ref**.
 
@@
 
-### Guild weighted means : nul model 1
+### *Does local functional composition from guilds differ ?*
 
-We used a null model (NM1) which randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix) [@bernard-verdier_community_2012].
+#### H1a. Environmental filtering differ in selection of local functional composition between canopy and understorey species
 
-The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds (see table \@ref(tab:GWM1)).
-Results in presence-absence are in adequation with observed results in abundance (see **...**).
-Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits, and two major differences were observed between the two guilds.
-The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height affected only understory guild composition.
-
-Focusing on canopy tree species, we found an increase in wood density under south-west orientation, both in abundance and presence-absence.
-We also observed a decrease in leaf area and thickness under the south-west orientation, but only in abundance and presence-absence respectively.
-Leaf area decreased with curvature in both abundance and presence-absence, while wood density and LDMC increased and decreased respectively with curvature in presence-absence only.
-
-For understorey tree species, SLA significantly decreased and wood density and leaf thickness significantly increased with curvature, both in abundance and presence-absence.
-We also detected a significant increase and decrease in wood economic spectrum and leaf area respectively with curvature in abundance.
-SLA in both abundance and presence-absence significantly increased with canopy height, while LDMC and wood density decreased with canopy height in abundance only.
-
-```{r GWM1}
-GWM <- L %>% 
-  t() %>% 
-  data.frame() %>% 
-  mutate(SpCode = row.names(.)) %>% 
-  left_join(Q) %>% 
-  select(-SpCode) %>% 
-  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-       variable.name = 'community', value.name = 'abundance') %>% 
-  group_by(strata, community) %>% 
-  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
-# nQ <- array(rep(as.matrix(Q), n),
-#             c(dim(Q), n),
-#             list(row.names(Q), names(Q), 1:n))
-# nQ <- array(apply(nQ, 3, function(x){
-#   x[,'SpCode'] <- sample(x[,'SpCode'])
-#   return(x)}), c(dim(Q), n),
-#   list(row.names(Q), names(Q), 1:n))
-# cl <- makeCluster(cores)
-# clusterExport(cl, list('L', 'nQ'))
-# nGWM <- parApply(cl, nQ, 3, function(x){
-#   library(dplyr)
-#   library(reshape2)
-#   L %>%
-#     t() %>%
-#     data.frame() %>%
-#     mutate(SpCode = row.names(.)) %>%
-#     left_join(data.frame(x)) %>%
-#     select(-SpCode) %>%
-#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-#          variable.name = 'community', value.name = 'abundance') %>%
-#     group_by(strata, community) %>%
-#     mutate_each(funs(as.character), -abundance) %>%
-#     mutate_each(funs(as.numeric), -abundance) %>%
-#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
-# })
-# stopCluster(cl)
-# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
-#               c(dim(GWM[3:9]), n),
-#               list(row.names(GWM), names(GWM[3:9]), 1:n))
-# GWM_ses <- GWM
-# GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
-# rm(cl, nQ, nGWM)
-# save(GWM_ses, file = file.path(path, 'GWM1_ses.Rdata'))
-load(file.path(path, 'GWM1_ses.Rdata'))
-GWM_ses %>% 
-  melt(id.vars = c('strata', 'community')) %>% 
-  rename(trait = variable) %>% 
-  rename(SES = value) %>% 
-  left_join(R) %>%
-  select(-community) %>% 
-  group_by(strata, trait) %>% 
-  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
-          direction = 'both', trace = 0))) %>% 
-  group_by(strata, trait, term) %>%
-  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
-                          stars(p.value, ns = '   '))) %>% 
-  dcast(strata + trait ~ term) %>% 
-  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
-  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
-  filter(trait != 'LES') %>% 
-  filter(trait != 'WES') %>%
-  bold() %>% 
-  kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
-```
-
-```{r WD, fig.cap='Exemple of standard effect size (SES) of wood density  (WD) for canopy (C) tree species.
-Each subplot show  standard effect size (SES) of wood density (WD) for canopy (C) tree species along one environmental variable gradient.
-Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
-See table \\@ref(tab:Traits) for abbreviation.'}
-GWM_ses %>% 
-  select(strata, WD, community) %>% 
-  left_join(R) %>% 
-  filter(strata == 'C') %>% 
-  ungroup() %>% 
-  select(-strata, -community) %>% 
-  melt(id.vars = 'WD') %>% 
-  ggplot(aes(value, WD)) +
-  geom_point() +
-  facet_wrap(~ variable, scales = "free") +
-  geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') +
-  geom_hline(yintercept = 0) +
-  xlab('Environmental gradient') +
-  ylab('Standard effect size of wood density')
-```
-
-### Guild weighted means : nul model 3
-
-We used a null model (NM3) which randomized species among guilds in the whole plot (see table \@ref(tab:GWM3)).
-
-We found that **curvature** was responsible for the guilds differenciation for most of the traits.
-Increased curvature resulted in denser lwood and leaves with high SLA in canopy, against lighter wood, low SLA, but bigger and thicker leaves in understorey.
-Some curvature results were congruent with **Wetness** variable.
-**South-West** orientation diminished the leaf area while increasing leaf density for canopy tree species.
-Finally, **Canopy** height and **basal area** affected in opposite ways SLA.
-Increased canopy height and decreased basal area resulted in lower SLA in canopy and increased SLA in understorey.
-
-```{r GWM3}
+```{r GWMH1a}
 GWM <- L %>% 
@@
 # rm(cl, nQ, nGWM)
-# save(GWM_ses, file = file.path(path, 'GWM3_ses.Rdata'))
-load(file.path(path, 'GWM3_ses.Rdata'))
+# save(GWM_ses, file = file.path(path, 'GWM_NM1_ses.Rdata'))
+load(file.path(path, 'GWM_NM1_ses.Rdata'))
 GWM_ses %>% 
-  melt(id.vars = c('strata', 'community')) %>% 
-  rename(trait = variable) %>% 
-  rename(SES = value) %>% 
-  left_join(R) %>%
   select(-community) %>% 
-  group_by(strata, trait) %>% 
-  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
-          direction = 'both', trace = 0))) %>% 
-  group_by(strata, trait, term) %>%
-  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
-                          stars(p.value, ns = '   '))) %>% 
-  dcast(strata + trait ~ term) %>% 
-  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
-  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
-  filter(trait != 'LES') %>% 
-  filter(trait != 'WES') %>%
-  bold() %>% 
-  kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+  group_by(strata) %>% 
+  summarise_all(mean, na.rm = T) %>% 
+   kable(caption = 'Mean of the standardized effect size (SES) with null model NM1 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
 See table \\@ref(tab:Traits) for abbreviations.
@@
 
-### Guild weighted variances : nul model 2
+#### H1b. Environmental filtering differ in selection of local functional composition between canopy and understorey species
 
-The second model (NM2) randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012].
-
-We found significant relationships between the standardized effect size (SES) of GWVs and the environmental predictors (see table \@ref(tab:GWV3)).
-**Wetness** decreased variance of wood density in canopy and all traits in understorey.
-**Curvature** increased LDMC variance in canopy.
-**Sout-West** orientation results in decreased variance of density of wood and leaf in understorey.
-Finally, **Canopy** hieght diminished variance of leaf are and density in canopy whereas **basal aea** increased variance of SLA in both strata and increased leaf area variance in canopy.
-
-```{r GWV3}
+```{r GWVH1b}
 GWV <- L %>% 
@@
 # nL <- array(replicate(n, t(apply(L, 1, function(x){
-#   x[x != 0] <- sample(x[which(x != 0)]) ; return(x)}))),
-#   c(dim(L), n),
-#   list(row.names(L), names(L), 1:n))
+#   x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0])
+#   x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0])
+#   return(x)
+#   }))), c(dim(L), n), list(row.names(L), names(L), 1:n))
 # cl <- makeCluster(cores)
@@
 # rm(cl, nL, nGWV)
-# save(GWV_ses, file = file.path(path, 'GWV_ses.Rdata'))
-load(file.path(path, 'GWV_ses.Rdata'))
+# save(GWV_ses, file = file.path(path, 'GWV_NM3_ses.Rdata'))
+load(file.path(path, 'GWV_NM3_ses.Rdata'))
+GWM_ses %>% 
+  select(-community) %>% 
+  group_by(strata) %>% 
+  summarise_all(mean, na.rm = T) %>% 
+   kable(caption = 'Mean of the standardized effect size (SES) with null model NM3 for each trait of GWVs, for canopy (C) and understorey (U) guilds, in abundance.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
+```
+
+### *Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?*
+
+#### H2a. Drivers of environmental filtering on local functional composition differ between canopy and understorey tree species
+
+```{r GWMH2a}
+GWM <- L %>% 
+  t() %>% 
+  data.frame() %>% 
+  mutate(SpCode = row.names(.)) %>% 
+  left_join(Q) %>% 
+  select(-SpCode) %>% 
+  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+       variable.name = 'community', value.name = 'abundance') %>% 
+  group_by(strata, community) %>% 
+  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# nQ <- array(rep(as.matrix(Q), n),
+#             c(dim(Q), n),
+#             list(row.names(Q), names(Q), 1:n))
+# nQ <- array(apply(nQ, 3, function(x){
+#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
+#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
+#   return(x)
+#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('L', 'nQ'))
+# nGWM <- parApply(cl, nQ, 3, function(x){
+#   library(dplyr)
+#   library(reshape2)
+#   L %>%
+#     t() %>%
+#     data.frame() %>%
+#     mutate(SpCode = row.names(.)) %>%
+#     left_join(data.frame(x)) %>%
+#     select(-SpCode) %>%
+#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+#          variable.name = 'community', value.name = 'abundance') %>%
+#     group_by(strata, community) %>%
+#     mutate_each(funs(as.character), -abundance) %>%
+#     mutate_each(funs(as.numeric), -abundance) %>%
+#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# })
+# stopCluster(cl)
+# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
+#               c(dim(GWM[3:9]), n),
+#               list(row.names(GWM), names(GWM[3:9]), 1:n))
+# GWM_ses <- GWM
+# GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
+# rm(cl, nQ, nGWM)
+# save(GWM_ses, file = file.path(path, 'GWM_NM2_ses.Rdata'))
+load(file.path(path, 'GWM_NM2_ses.Rdata'))
+GWM_ses %>% 
+  melt(id.vars = c('strata', 'community')) %>% 
+  rename(trait = variable) %>% 
+  rename(SES = value) %>% 
+  left_join(R) %>%
+  select(-community) %>% 
+  group_by(strata, trait) %>% 
+  do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
+          direction = 'both', trace = 0))) %>% 
+  group_by(strata, trait, term) %>%
+  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
+                          stars(p.value, ns = '   '))) %>% 
+  dcast(strata + trait ~ term) %>% 
+  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
+  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
+  filter(trait != 'LES') %>% 
+  filter(trait != 'WES') %>%
+  bold() %>% 
+  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
+```
+
+```{r WD, fig.cap='Exemple of standard effect size (SES) of wood density  (WD) for canopy (C) tree species.
+Each subplot show  standard effect size (SES) with null model NM2 of wood density (WD) for canopy (C) tree species along one environmental variable gradient.
+Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
+See table \\@ref(tab:Traits) for abbreviation.'}
+GWM_ses %>% 
+  select(strata, WD, community) %>% 
+  left_join(R) %>% 
+  filter(strata == 'C') %>% 
+  ungroup() %>% 
+  select(-strata, -community) %>% 
+  melt(id.vars = 'WD') %>% 
+  ggplot(aes(value, WD)) +
+  geom_point() +
+  facet_wrap(~ variable, scales = "free") +
+  geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') +
+  geom_hline(yintercept = 0) +
+  xlab('Environmental gradient') +
+  ylab('Standard effect size of wood density')
+```
+
+#### H2b. Strength of drivers of evironmental filtering differ between canopy and understorey tree species
+
+```{r GWVH2b}
+GWV <- L %>% 
+  t() %>% 
+  data.frame() %>% 
+  filter(sum(.) > 2) %>% 
+  mutate(SpCode = row.names(.)) %>% 
+  left_join(Q) %>% 
+  select(-SpCode) %>% 
+  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+       variable.name = 'community', value.name = 'abundance') %>% 
+  melt(id.vars = c('strata', 'community', 'abundance'),
+       variable.name = 'trait', value.name = 'trait_value') %>% 
+  left_join(melt(GWM, id.vars = c('strata', 'community'),
+                 variable.name = 'trait', value.name = 'GWM')) %>% 
+  group_by(strata, community, trait) %>% 
+  mutate(GWV = (trait_value - GWM)^2) %>% 
+  select(-trait_value, -GWM) %>% 
+  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
+  dcast(strata + community ~ trait)
+# nL <- array(replicate(n, t(apply(L, 1, function(x){
+#   x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0])
+#   x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0])
+#   return(x)
+#   }))), c(dim(L), n), list(row.names(L), names(L), 1:n))
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('nL', 'Q', 'GWM'))
+# nGWV <- parApply(cl, nL, 3, function(x){
+#   library(dplyr)
+#   library(reshape2)
+#   x %>%
+#     t() %>%
+#     data.frame() %>%
+#     filter(sum(.) > 2) %>%
+#     mutate(SpCode = row.names(.)) %>%
+#     left_join(Q) %>%
+#     select(-SpCode) %>%
+#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+#          variable.name = 'community', value.name = 'abundance') %>%
+#     melt(id.vars = c('strata', 'community', 'abundance'),
+#          variable.name = 'trait', value.name = 'trait_value') %>%
+#     left_join(melt(GWM, id.vars = c('strata', 'community'),
+#                    variable.name = 'trait', value.name = 'GWM')) %>%
+#     group_by(strata, community, trait) %>%
+#     mutate(GWV = (trait_value - GWM)^2) %>%
+#     select(-trait_value, -GWM) %>%
+#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%
+#     dcast(strata + community ~ trait)
+# })
+# stopCluster(cl)
+# nGWV <- array(unlist(lapply(nGWV, function(x) as.matrix(x[3:9]))),
+#               c(dim(GWV[3:9]), n),
+#               list(row.names(GWV), names(GWV[3:9]), 1:n))
+# GWV_ses <- GWV
+# GWV_ses[3:9] <- (GWV[3:9] - apply(nGWV, 1:2, mean, na.rm = T))/apply(nGWV, 1:2, sd, na.rm = T)
+# rm(cl, nL, nGWV)
+# save(GWV_ses, file = file.path(path, 'GWV_NM3_ses.Rdata'))
+load(file.path(path, 'GWV_NM3_ses.Rdata'))
 GWV_ses %>% 
@@
   bold() %>% 
-  kable(caption = 'Linear models of the standardized effect sizes (SES) of GWV for canopy (C) and understorey (U) guilds.
+  kable(caption = 'Linear models of the standardized effect sizes (SES) with null model NM3 of GWV for canopy (C) and understorey (U) guilds.
 We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
@@
 ```
+
+<!-- ### Guild weighted means : nul model 1 -->
+
+<!-- We used a null model (NM1) which randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix) [@bernard-verdier_community_2012]. -->
+
+<!-- The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds (see table \@ref(tab:GWM1)).
+Results in presence-absence are in adequation with observed results in abundance (see **...**).
+Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits, and two major differences were observed between the two guilds.
+The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height affected only understory guild composition.
+-->
+
+<!-- Focusing on canopy tree species, we found an increase in wood density under south-west orientation, both in abundance and presence-absence.
+We also observed a decrease in leaf area and thickness under the south-west orientation, but only in abundance and presence-absence respectively.
+Leaf area decreased with curvature in both abundance and presence-absence, while wood density and LDMC increased and decreased respectively with curvature in presence-absence only.
+-->
+
+<!-- For understorey tree species, SLA significantly decreased and wood density and leaf thickness significantly increased with curvature, both in abundance and presence-absence.
+We also detected a significant increase and decrease in wood economic spectrum and leaf area respectively with curvature in abundance.
+SLA in both abundance and presence-absence significantly increased with canopy height, while LDMC and wood density decreased with canopy height in abundance only.
+-->
+
+<!-- ```{r GWM1} -->
+<!-- GWM <- L %>%  -->
+<!--   t() %>%  -->
+<!--   data.frame() %>%  -->
+<!--   mutate(SpCode = row.names(.)) %>%  -->
+<!--   left_join(Q) %>%  -->
+<!--   select(-SpCode) %>%  -->
+<!--   melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
+<!--        variable.name = 'community', value.name = 'abundance') %>%  -->
+<!--   group_by(strata, community) %>%  -->
+<!--   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) -->
+<!-- # nQ <- array(rep(as.matrix(Q), n), -->
+<!-- #             c(dim(Q), n), -->
+<!-- #             list(row.names(Q), names(Q), 1:n)) -->
+<!-- # nQ <- array(apply(nQ, 3, function(x){ -->
+<!-- #   x[,'SpCode'] <- sample(x[,'SpCode']) -->
+<!-- #   return(x)}), c(dim(Q), n), -->
+<!-- #   list(row.names(Q), names(Q), 1:n)) -->
+<!-- # cl <- makeCluster(cores) -->
+<!-- # clusterExport(cl, list('L', 'nQ')) -->
+<!-- # nGWM <- parApply(cl, nQ, 3, function(x){ -->
+<!-- #   library(dplyr) -->
+<!-- #   library(reshape2) -->
+<!-- #   L %>% -->
+<!-- #     t() %>% -->
+<!-- #     data.frame() %>% -->
+<!-- #     mutate(SpCode = row.names(.)) %>% -->
+<!-- #     left_join(data.frame(x)) %>% -->
+<!-- #     select(-SpCode) %>% -->
+<!-- #     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
+<!-- #          variable.name = 'community', value.name = 'abundance') %>% -->
+<!-- #     group_by(strata, community) %>% -->
+<!-- #     mutate_each(funs(as.character), -abundance) %>% -->
+<!-- #     mutate_each(funs(as.numeric), -abundance) %>% -->
+<!-- #     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) -->
+<!-- # }) -->
+<!-- # stopCluster(cl) -->
+<!-- # nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))), -->
+<!-- #               c(dim(GWM[3:9]), n), -->
+<!-- #               list(row.names(GWM), names(GWM[3:9]), 1:n)) -->
+<!-- # GWM_ses <- GWM -->
+<!-- # GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T) -->
+<!-- # rm(cl, nQ, nGWM) -->
+<!-- # save(GWM_ses, file = file.path(path, 'GWM1_ses.Rdata')) -->
+<!-- load(file.path(path, 'GWM1_ses.Rdata')) -->
+<!-- GWM_ses %>%  -->
+<!--   melt(id.vars = c('strata', 'community')) %>%  -->
+<!--   rename(trait = variable) %>%  -->
+<!--   rename(SES = value) %>%  -->
+<!--   left_join(R) %>% -->
+<!--   select(-community) %>%  -->
+<!--   group_by(strata, trait) %>%  -->
+<!--   do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .),  -->
+<!--           direction = 'both', trace = 0))) %>%  -->
+<!--   group_by(strata, trait, term) %>% -->
+<!--   transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T),  -->
+<!--                           stars(p.value, ns = '   '))) %>%  -->
+<!--   dcast(strata + trait ~ term) %>%  -->
+<!--   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>%  -->
+<!--   mutate_each(funs(replace(., is.na(.), ' '))) %>%  -->
+<!--   filter(trait != 'LES') %>%  -->
+<!--   filter(trait != 'WES') %>% -->
+<!--   bold() %>%  -->
+<!--   kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).') -->
+<!-- ``` -->
+
+<!-- ```{r WD, fig.cap='Exemple of standard effect size (SES) of wood density  (WD) for canopy (C) tree species.
+Each subplot show  standard effect size (SES) of wood density (WD) for canopy (C) tree species along one environmental variable gradient.
+Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
+See table \\@ref(tab:Traits) for abbreviation.'} -->
+<!-- GWM_ses %>%  -->
+<!--   select(strata, WD, community) %>%  -->
+<!--   left_join(R) %>%  -->
+<!--   filter(strata == 'C') %>%  -->
+<!--   ungroup() %>%  -->
+<!--   select(-strata, -community) %>%  -->
+<!--   melt(id.vars = 'WD') %>%  -->
+<!--   ggplot(aes(value, WD)) + -->
+<!--   geom_point() + -->
+<!--   facet_wrap(~ variable, scales = "free") + -->
+<!--   geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') + -->
+<!--   geom_hline(yintercept = 0) + -->
+<!--   xlab('Environmental gradient') + -->
+<!--   ylab('Standard effect size of wood density') -->
+<!-- ``` -->
+
+<!-- ### Guild weighted means : nul model 3 -->
+
+<!-- We used a null model (NM3) which randomized species among guilds in the whole plot (see table \@ref(tab:GWM3)). -->
+
+<!-- We found that **curvature** was responsible for the guilds differenciation for most of the traits.
+Increased curvature resulted in denser lwood and leaves with high SLA in canopy, against lighter wood, low SLA, but bigger and thicker leaves in understorey.
+Some curvature results were congruent with **Wetness** variable.
+**South-West** orientation diminished the leaf area while increasing leaf density for canopy tree species.
+Finally, **Canopy** height and **basal area** affected in opposite ways SLA.
+Increased canopy height and decreased basal area resulted in lower SLA in canopy and increased SLA in understorey.
+-->
+
+<!-- ```{r GWM3} -->
+<!-- GWM <- L %>%  -->
+<!--   t() %>%  -->
+<!--   data.frame() %>%  -->
+<!--   mutate(SpCode = row.names(.)) %>%  -->
+<!--   left_join(Q) %>%  -->
+<!--   select(-SpCode) %>%  -->
+<!--   melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
+<!--        variable.name = 'community', value.name = 'abundance') %>%  -->
+<!--   group_by(strata, community) %>%  -->
+<!--   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) -->
+<!-- # nQ <- array(rep(as.matrix(Q), n), -->
+<!-- #             c(dim(Q), n), -->
+<!-- #             list(row.names(Q), names(Q), 1:n)) -->
+<!-- # nQ <- array(apply(nQ, 3, function(x){ -->
+<!-- #   x[,'strata'] <- sample(x[,'strata']) -->
+<!-- #   return(x)}), c(dim(Q), n), -->
+<!-- #   list(row.names(Q), names(Q), 1:n)) -->
+<!-- # cl <- makeCluster(cores) -->
+<!-- # clusterExport(cl, list('L', 'nQ')) -->
+<!-- # nGWM <- parApply(cl, nQ, 3, function(x){ -->
+<!-- #   library(dplyr) -->
+<!-- #   library(reshape2) -->
+<!-- #   L %>% -->
+<!-- #     t() %>% -->
+<!-- #     data.frame() %>% -->
+<!-- #     mutate(SpCode = row.names(.)) %>% -->
+<!-- #     left_join(data.frame(x)) %>% -->
+<!-- #     select(-SpCode) %>% -->
+<!-- #     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
+<!-- #          variable.name = 'community', value.name = 'abundance') %>% -->
+<!-- #     group_by(strata, community) %>% -->
+<!-- #     mutate_each(funs(as.character), -abundance) %>% -->
+<!-- #     mutate_each(funs(as.numeric), -abundance) %>% -->
+<!-- #     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) -->
+<!-- # }) -->
+<!-- # stopCluster(cl) -->
+<!-- # nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))), -->
+<!-- #               c(dim(GWM[3:9]), n), -->
+<!-- #               list(row.names(GWM), names(GWM[3:9]), 1:n)) -->
+<!-- # GWM_ses <- GWM -->
+<!-- # GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T) -->
+<!-- # rm(cl, nQ, nGWM) -->
+<!-- # save(GWM_ses, file = file.path(path, 'GWM3_ses.Rdata')) -->
+<!-- load(file.path(path, 'GWM3_ses.Rdata')) -->
+<!-- GWM_ses %>%  -->
+<!--   melt(id.vars = c('strata', 'community')) %>%  -->
+<!--   rename(trait = variable) %>%  -->
+<!--   rename(SES = value) %>%  -->
+<!--   left_join(R) %>% -->
+<!--   select(-community) %>%  -->
+<!--   group_by(strata, trait) %>%  -->
+<!--   do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .),  -->
+<!--           direction = 'both', trace = 0))) %>%  -->
+<!--   group_by(strata, trait, term) %>% -->
+<!--   transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T),  -->
+<!--                           stars(p.value, ns = '   '))) %>%  -->
+<!--   dcast(strata + trait ~ term) %>%  -->
+<!--   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>%  -->
+<!--   mutate_each(funs(replace(., is.na(.), ' '))) %>%  -->
+<!--   filter(trait != 'LES') %>%  -->
+<!--   filter(trait != 'WES') %>% -->
+<!--   bold() %>%  -->
+<!--   kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).') -->
+<!-- ``` -->
+
+<!-- ### Guild weighted variances : nul model 2 -->
+
+<!-- The second model (NM2) randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012]. -->
+
+<!-- We found significant relationships between the standardized effect size (SES) of GWVs and the environmental predictors (see table \@ref(tab:GWV3)).
+**Wetness** decreased variance of wood density in canopy and all traits in understorey.
+**Curvature** increased LDMC variance in canopy.
+**Sout-West** orientation results in decreased variance of density of wood and leaf in understorey.
+Finally, **Canopy** hieght diminished variance of leaf are and density in canopy whereas **basal aea** increased variance of SLA in both strata and increased leaf area variance in canopy.
+-->
+
+<!-- ```{r GWV3} -->
+<!-- GWV <- L %>%  -->
+<!--   t() %>%  -->
+<!--   data.frame() %>%  -->
+<!--   filter(sum(.) > 2) %>%  -->
+<!--   mutate(SpCode = row.names(.)) %>%  -->
+<!--   left_join(Q) %>%  -->
+<!--   select(-SpCode) %>%  -->
+<!--   melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
+<!--        variable.name = 'community', value.name = 'abundance') %>%  -->
+<!--   melt(id.vars = c('strata', 'community', 'abundance'), -->
+<!--        variable.name = 'trait', value.name = 'trait_value') %>%  -->
+<!--   left_join(melt(GWM, id.vars = c('strata', 'community'), -->
+<!--                  variable.name = 'trait', value.name = 'GWM')) %>%  -->
+<!--   group_by(strata, community, trait) %>%  -->
+<!--   mutate(GWV = (trait_value - GWM)^2) %>%  -->
+<!--   select(-trait_value, -GWM) %>%  -->
+<!--   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%  -->
+<!--   dcast(strata + community ~ trait) -->
+<!-- # nL <- array(replicate(n, t(apply(L, 1, function(x){ -->
+<!-- #   x[x != 0] <- sample(x[which(x != 0)]) ; return(x)}))), -->
+<!-- #   c(dim(L), n), -->
+<!-- #   list(row.names(L), names(L), 1:n)) -->
+<!-- # cl <- makeCluster(cores) -->
+<!-- # clusterExport(cl, list('nL', 'Q', 'GWM')) -->
+<!-- # nGWV <- parApply(cl, nL, 3, function(x){ -->
+<!-- #   library(dplyr) -->
+<!-- #   library(reshape2) -->
+<!-- #   x %>% -->
+<!-- #     t() %>% -->
+<!-- #     data.frame() %>% -->
+<!-- #     filter(sum(.) > 2) %>% -->
+<!-- #     mutate(SpCode = row.names(.)) %>% -->
+<!-- #     left_join(Q) %>% -->
+<!-- #     select(-SpCode) %>% -->
+<!-- #     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
+<!-- #          variable.name = 'community', value.name = 'abundance') %>% -->
+<!-- #     melt(id.vars = c('strata', 'community', 'abundance'), -->
+<!-- #          variable.name = 'trait', value.name = 'trait_value') %>% -->
+<!-- #     left_join(melt(GWM, id.vars = c('strata', 'community'), -->
+<!-- #                    variable.name = 'trait', value.name = 'GWM')) %>% -->
+<!-- #     group_by(strata, community, trait) %>% -->
+<!-- #     mutate(GWV = (trait_value - GWM)^2) %>% -->
+<!-- #     select(-trait_value, -GWM) %>% -->
+<!-- #     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% -->
+<!-- #     dcast(strata + community ~ trait) -->
+<!-- # }) -->
+<!-- # stopCluster(cl) -->
+<!-- # nGWV <- array(unlist(lapply(nGWV, function(x) as.matrix(x[3:9]))), -->
+<!-- #               c(dim(GWV[3:9]), n), -->
+<!-- #               list(row.names(GWV), names(GWV[3:9]), 1:n)) -->
+<!-- # GWV_ses <- GWV -->
+<!-- # GWV_ses[3:9] <- (GWV[3:9] - apply(nGWV, 1:2, mean, na.rm = T))/apply(nGWV, 1:2, sd, na.rm = T) -->
+<!-- # rm(cl, nL, nGWV) -->
+<!-- # save(GWV_ses, file = file.path(path, 'GWV_ses.Rdata')) -->
+<!-- load(file.path(path, 'GWV_ses.Rdata')) -->
+<!-- GWV_ses %>%  -->
+<!--   melt(id.vars = c('strata', 'community')) %>%  -->
+<!--   rename(trait = variable) %>%  -->
+<!--   rename(SES = value) %>%  -->
+<!--   left_join(R) %>% -->
+<!--   select(-community) %>%  -->
+<!--   group_by(strata, trait) %>%  -->
+<!--   do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .),  -->
+<!--           direction = 'both', trace = 0))) %>%  -->
+<!--   group_by(strata, trait, term) %>% -->
+<!--   transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T),  -->
+<!--                           stars(p.value, ns = '   '))) %>%  -->
+<!--   dcast(strata + trait ~ term) %>%  -->
+<!--   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>%  -->
+<!--   mutate_each(funs(replace(., is.na(.), ' '))) %>%  -->
+<!--   filter(trait != 'LES') %>%  -->
+<!--   filter(trait != 'WES') %>% -->
+<!--   bold() %>%  -->
+<!--   kable(caption = 'Linear models of the standardized effect sizes (SES) of GWV for canopy (C) and understorey (U) guilds.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+See table \\@ref(tab:Traits) for abbreviations.
+Bold and  number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).') -->
+<!-- ``` -->
 

### 2018-01-24 inst/doc/draft2.Rmd
Uppangala 0.1.9040 January 2018 draft results

@@
 library(ggfortify)
-cores <- 3
+cores <- 16
 opts_chunk$set(
@@
 path <- '~/Documents/IFP/Uppangala/inst/doc/draft_save/'
-n = 3 # Random repetitions
+n = 999 # Random repetitions
 ```
@@
 GWM_ses %>% 
-  select(-community) %>% 
+  select(-community, -LES, -WES) %>% 
   group_by(strata) %>% 
-  summarise_all(mean, na.rm = T) %>% 
-   kable(caption = 'Mean of the standardized effect size (SES) with null model NM1 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
+  mutate_all(function(x) abs(x) > 0.95) %>% 
+  summarise_all(function(x) sum(x, na.rm = T)/length(x)*100) %>% 
+  kable(caption = 'Percentage of community significantly (p-value<0.05) different from null expectations with null model NM1 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+See table \\@ref(tab:Traits) for abbreviations.')
 ```
@@
 load(file.path(path, 'GWV_NM3_ses.Rdata'))
-GWM_ses %>% 
-  select(-community) %>% 
+GWV_ses %>% 
+  select(-community, -LES, -WES) %>% 
   group_by(strata) %>% 
-  summarise_all(mean, na.rm = T) %>% 
-   kable(caption = 'Mean of the standardized effect size (SES) with null model NM3 for each trait of GWVs, for canopy (C) and understorey (U) guilds, in abundance.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
+  mutate_all(function(x) abs(x) > 0.95) %>% 
+  summarise_all(function(x) sum(x, na.rm = T)/length(x)*100) %>% 
+  kable(caption = 'Percentage of community significantly (p-value<0.05) different from null expectations with null model NM3 for each trait of GWVs, for canopy (C) and understorey (U) guilds, in abundance.
+See table \\@ref(tab:Traits) for abbreviations.')
 ```
@@
   dcast(strata + trait ~ term) %>% 
+  {if(!("Slope" %in% names(.))) mutate(., Slope = NA)} %>% 
   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 

### 2018-01-24 inst/doc/draft2.Rmd
Uppangala 0.1.9041 January 2018 Tabebuia 2nd run

@@
 #               list(row.names(GWM), names(GWM[3:9]), 1:n))
-# GWM_ses <- GWM
-# GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
-# rm(cl, nQ, nGWM)
-# save(GWM_ses, file = file.path(path, 'GWM_NM1_ses.Rdata'))
-load(file.path(path, 'GWM_NM1_ses.Rdata'))
-GWM_ses %>% 
+# rm(cl, nQ)
+# save(nGWM, file = file.path(path, 'nGWM_NM1.Rdata'))
+load(file.path(path, 'nGWM_NM1.Rdata'))
+GWM_pval <- GWM
+GWM_pval[3:9] <- apply(
+  array(c(as.matrix(GWM[3:9]), nGWM), c(dim(GWM[3:9]), n+1)), 
+  1:2, function(x) rank(x)[1]/(n+1))
+GWM_pval %>% 
   select(-community, -LES, -WES) %>% 
   group_by(strata) %>% 
-  mutate_all(function(x) abs(x) > 0.95) %>% 
+  mutate_all(function(x) findInterval(x, c(0.05, 0.95)) == 1) %>% 
   summarise_all(function(x) sum(x, na.rm = T)/length(x)*100) %>% 
@@
 #               list(row.names(GWV), names(GWV[3:9]), 1:n))
-# GWV_ses <- GWV
-# GWV_ses[3:9] <- (GWV[3:9] - apply(nGWV, 1:2, mean, na.rm = T))/apply(nGWV, 1:2, sd, na.rm = T)
-# rm(cl, nL, nGWV)
-# save(GWV_ses, file = file.path(path, 'GWV_NM3_ses.Rdata'))
-load(file.path(path, 'GWV_NM3_ses.Rdata'))
-GWV_ses %>% 
+# rm(cl, nL)
+# save(nGWV, file = file.path(path, 'nGWV_NM3.Rdata'))
+load(file.path(path, 'nGWV_NM3.Rdata'))
+GWV_pval <- GWV
+GWV_pval[3:9] <- apply(
+  array(c(as.matrix(GWV[3:9]), nGWV), c(dim(GWV[3:9]), n+1)), 
+  1:2, function(x) rank(x)[1]/(n+1))
+GWV_pval %>% 
   select(-community, -LES, -WES) %>% 
@@
 #               list(row.names(GWM), names(GWM[3:9]), 1:n))
-# GWM_ses <- GWM
-# GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
-# rm(cl, nQ, nGWM)
-# save(GWM_ses, file = file.path(path, 'GWM_NM2_ses.Rdata'))
-load(file.path(path, 'GWM_NM2_ses.Rdata'))
+# rm(cl, nQ)
+# save(nGWM, file = file.path(path, 'nGWM_NM2.Rdata'))
+load(file.path(path, 'nGWM_NM2.Rdata'))
+GWM_ses <- GWM
+GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
 GWM_ses %>% 
@@
   dcast(strata + community ~ trait)
-# nL <- array(replicate(n, t(apply(L, 1, function(x){
-#   x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0])
-#   x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0])
-#   return(x)
-#   }))), c(dim(L), n), list(row.names(L), names(L), 1:n))
-# cl <- makeCluster(cores)
-# clusterExport(cl, list('nL', 'Q', 'GWM'))
-# nGWV <- parApply(cl, nL, 3, function(x){
-#   library(dplyr)
-#   library(reshape2)
-#   x %>%
-#     t() %>%
-#     data.frame() %>%
-#     filter(sum(.) > 2) %>%
-#     mutate(SpCode = row.names(.)) %>%
-#     left_join(Q) %>%
-#     select(-SpCode) %>%
-#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-#          variable.name = 'community', value.name = 'abundance') %>%
-#     melt(id.vars = c('strata', 'community', 'abundance'),
-#          variable.name = 'trait', value.name = 'trait_value') %>%
-#     left_join(melt(GWM, id.vars = c('strata', 'community'),
-#                    variable.name = 'trait', value.name = 'GWM')) %>%
-#     group_by(strata, community, trait) %>%
-#     mutate(GWV = (trait_value - GWM)^2) %>%
-#     select(-trait_value, -GWM) %>%
-#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%
-#     dcast(strata + community ~ trait)
-# })
-# stopCluster(cl)
-# nGWV <- array(unlist(lapply(nGWV, function(x) as.matrix(x[3:9]))),
-#               c(dim(GWV[3:9]), n),
-#               list(row.names(GWV), names(GWV[3:9]), 1:n))
-# GWV_ses <- GWV
-# GWV_ses[3:9] <- (GWV[3:9] - apply(nGWV, 1:2, mean, na.rm = T))/apply(nGWV, 1:2, sd, na.rm = T)
-# rm(cl, nL, nGWV)
-# save(GWV_ses, file = file.path(path, 'GWV_NM3_ses.Rdata'))
+load(file.path(path, 'nGWV_NM3.Rdata'))
+GWV_ses <- GWV
+GWV_ses[3:9] <- (GWV[3:9] - apply(nGWV, 1:2, mean, na.rm = T))/apply(nGWV, 1:2, sd, na.rm = T)
 load(file.path(path, 'GWV_NM3_ses.Rdata'))

### 2018-01-25 inst/doc/draft2.Rmd
H1a et H1b correction

@@
 
-```{r GWMH1a}
+```{r GWMH1a, fig.cap="Number of communities significantly (p-value<0.05) different from null expectations with null model NM1 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
+See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
 GWM <- L %>%
@@
   group_by(strata) %>% 
-  mutate_all(function(x) findInterval(x, c(0.05, 0.95)) == 1) %>% 
-  summarise_all(function(x) sum(x, na.rm = T)/length(x)*100) %>% 
-  kable(caption = 'Percentage of community significantly (p-value<0.05) different from null expectations with null model NM1 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-See table \\@ref(tab:Traits) for abbreviations.')
+  mutate_all(function(x) findInterval(x, c(0.05, 0.95)) != 1) %>% 
+  melt(id.vars = "strata", variable.name = "trait") %>% 
+  filter(value == TRUE) %>% 
+  ggplot(aes(x = factor(trait), 
+             fill = factor(strata))) +
+  geom_bar(position = "dodge") +
+  scale_fill_discrete("Guild") +
+  xlab("Functional trait") +
+  ylab("Number of communities") +
+  ylim(c(0,dim(L)[1]))
 ```
@@
 
-```{r GWVH1b}
+```{r GWVH1b, fig.cap="Number of communities significantly (p-value<0.05) different from null expectations with null model NM3 for each trait of GWVs, for canopy (C) and understorey (U) guilds, in abundance.
+See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
 GWV <- L %>%
@@
   group_by(strata) %>% 
-  mutate_all(function(x) abs(x) > 0.95) %>% 
-  summarise_all(function(x) sum(x, na.rm = T)/length(x)*100) %>% 
-  kable(caption = 'Percentage of community significantly (p-value<0.05) different from null expectations with null model NM3 for each trait of GWVs, for canopy (C) and understorey (U) guilds, in abundance.
-See table \\@ref(tab:Traits) for abbreviations.')
+  mutate_all(function(x) findInterval(x, c(0.05, 0.95)) != 1) %>% 
+  melt(id.vars = "strata", variable.name = "trait") %>% 
+  filter(value == TRUE) %>% 
+  ggplot(aes(x = factor(trait), 
+             fill = factor(strata))) +
+  geom_bar(position = "dodge") +
+  scale_fill_discrete("Guild") +
+  xlab("Functional trait") +
+  ylab("Number of communities") +
+  ylim(c(0,dim(L)[1]))
 ```

### 2018-02-13 inst/doc/draft2.Rmd
13/02/18 Draft

@@
 output:
+  bookdown::pdf_document2:
+    includes:
+      before_body: ./template/doc_prefix.tex
+      in_header: ./template/preamble.tex
+    number_sections: false
+    toc: false
   bookdown::html_document2:
@@
     toc_float: yes
-  bookdown::pdf_document2:
-    number_sections: false
+linestretch: 1.5
 csl: mee.csl
@@
         'Stem transport, structure and defense')
-), caption = 'Functional traits measured. Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].')
+), caption = 'Functional traits measured.
+Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].', format = 'pandoc')
 ```
@@
 In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of species $i$ weighted by species relative abundance $p_{i}$:
-$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
+$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~~ (1)$$
 where $S$ is the number of species in the quadrat $j$.
@@
 GWV of trait $t$ was calculated as :
-$$GWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - GWM_{j})^2 ~ (2)$$
-
-To test for any significant changes in functional composition, we devised a null model approach. We thus explored following questions with subsequent hypothesis:
-
-* __Q1__ Does local functional composition from guilds differ ?
-    * ___H1a__ Environmental filtering differ in selection of local functional composition between canopy and understorey species._
-    * ___H1b__ Environmental filtering strenght differ locally between canopy and understorey species._
-* __Q2__ Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?
-    * ___H2a__ Drivers of environmental filtering on local functional composition differ between canopy and understorey tree species._ 
-    * ___H2b__ Strength of drivers of evironmental filtering differ between canopy and understorey tree species._
-
-To accept or reject hypothesis we used three null models :
-
-* __NM1__ randomized species among guilds. Consequently, NM1 breaks the link between species and guilds while keeping community structure, thus testing for hypothesis __H1a__.
-* __NM2__ randomized trait values among species in the whole plot for each guild (permuting raw in `Trait*Species` matrices of both guilds) [@bernard-verdier_community_2012].
-Consequently, NM2 breaks the link between environement and local functional composition while keeping guilds and species spatial structure, thus testing for hypothesis __H2a__.
-* __NM3__  randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012].
-Consequently NM3, breaks the link between environment and species abundances while keeping guilds, community, and species spatial structure, thus testing for hypothesis __H1b__ and __H2b__.
+$$GWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - GWM_{j})^2 ~~ (2)$$
+
+We first tested functional compostion differentiation between guilds within each community with a wilcoxson test, to check if the local functional composition from guilds differed.
+Inside each community we tested the difference between all individuals of each guilds, and condidered the difference to be significative with a p-value inferior to 5%.
+We then summarized communities significativelly differing in their functional traits between understorey and canpoy guilds for each trait.
+
+Then we devised a null model approach to test for any significant changes in functional composition in response to an environmental gradient.
+We thus wanted to explore if the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds.
+We consequently used subsequent hypothesis:
+
+* ___H1__ Drivers of environmental filtering on local functional composition differ between canopy and understorey tree species._ 
+* ___H2__ Strength of drivers of evironmental filtering differ between canopy and understorey tree species._
+
+To accept or reject hypothesis we used two null models :
+
+* __NM1__ randomized trait values among species in the whole plot for each guild (permuting raw in `Trait*Species` matrices of both guilds) [@bernard-verdier_community_2012].
+Consequently, NM1 breaks the link between environement and local functional composition while keeping guilds and species spatial structure, thus testing for hypothesis __H1__.
+* __NM2__  randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012].
+Consequently NM2, breaks the link between environment and species abundances while keeping guilds, community, and species spatial structure, thus testing for hypothesis __H2__.
 
@@
 To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as
-$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (2)$$
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} ~~ (3)$$
 where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of the corresponding null distribution.
@@
 The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species.
-See table \\@ref(tab:Traits) for abbreviation.')
+See table \\@ref(tab:Traits) for abbreviation.', format = 'pandoc')
 ```
@@
 However, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between those two traits.
-Finally, hypervolume overlap between the two guilds was not significantly lower than null expectations ($p = 0.165$, see Supplementary materials [S2: Hypervolumes of functional guilds]).
+Finally, hypervolume overlap between the two guilds was not significantly lower than null expectations ($p = 0.115$, see Supplementary materials [S2: Hypervolumes of functional guilds]).
+Still the overlap was pretty low suggesting a trend of functional differentiation between the two guilds.
 
@@
 
-#### H1a. Environmental filtering differ in selection of local functional composition between canopy and understorey species
+Except for Specific Leaf Area (SLA), most of the traits were not significantly different between guilds in the majority of communities (only 20 to 41% significative for WD, LDMC, LA and LT against 79% for SLA, see figure \@ref(fig:percsign) ).
+But regarding guilds weighted means (GWM), understorey and canopy guilds strongly diverged among communities (see figure \@ref(fig:CWMboxplot) ).
+Canopy guilds have higher SLA, WD and LA but lower LDMC and LT than understorey guilds.
+In other words, canopy guilds have larger, thinner and lighter leaves with increased SLA and denser woods, whereas understorey guilds have smaller, thicker, and more dense leaves with lower SLA and lighter woods.
+
+```{r percsign, fig.cap="Percentage of communities with indivdual traits significantly different between guilds (for canopy (C) and understorey (U) guilds).
+Test realized with an unpaired wilcoxon two sided test, significance was considered for p-values under the level $\\alpha = 0.05$ .See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
+L %>% 
+  mutate(community = row.names(.)) %>% 
+    filter(!(community %in% c("C105", "C11",  "C26",  "C5" ))) %>% 
+  melt(id.vars = "community", 
+       variable.name = "SpCode",
+       value.name = "abundance") %>% 
+  filter(abundance > 0) %>% 
+  tidyr::uncount(abundance) %>% 
+  left_join(Q) %>% 
+  select(- SpCode, -LES, -WES) %>% 
+  melt(id.vars = c("community", "strata"),
+       variable.name = "trait") %>% 
+  group_by(community, trait) %>% 
+  do(wilcoxon = try(wilcox.test(value~strata, data=.)$p.value)) %>% 
+  filter(!inherits(wilcoxon, "try-error")) %>% 
+  mutate(wilcoxon = unlist(wilcoxon)) %>% 
+  group_by(trait) %>% 
+  mutate(test_0.5 = wilcoxon < 0.05) %>%
+  dplyr::summarise(com_sign = sum(test_0.5)/n()*100) %>%
+  ggplot(aes(trait, com_sign)) +
+  geom_bar(stat = "identity") +
+  xlab("Functional trait") +
+  ylab("Percentage of significant communities")
+```
 
 ```{r GWMH1a, fig.cap="Number of communities significantly (p-value<0.05) different from null expectations with null model NM1 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
+See table \\@ref(tab:Traits) for abbreviations.", fig.height=4, include = F}
 GWM <- L %>%
@@
   1:2, function(x) rank(x)[1]/(n+1))
-GWM_pval %>% 
+```
+
+```{r CWMboxplot, fig.cap="Boxplot of guilds weighted means (GWM) among communties for each traits (for canopy (C) and understorey (U) guilds).
+See table \\@ref(tab:Traits) for abbreviations.", fig.height=4, include = T}
+GWM %>% 
   select(-community, -LES, -WES) %>% 
-  group_by(strata) %>% 
-  mutate_all(function(x) findInterval(x, c(0.05, 0.95)) != 1) %>% 
-  melt(id.vars = "strata", variable.name = "trait") %>% 
-  filter(value == TRUE) %>% 
-  ggplot(aes(x = factor(trait), 
-             fill = factor(strata))) +
-  geom_bar(position = "dodge") +
-  scale_fill_discrete("Guild") +
-  xlab("Functional trait") +
-  ylab("Number of communities") +
-  ylim(c(0,dim(L)[1]))
-```
-
-#### H1b. Environmental filtering differ in selection of local functional composition between canopy and understorey species
+  melt(id.vars = "strata", 
+       variable.name = "trait",
+       value.name = "GWM") %>% 
+  ggplot(aes(x = strata, y = GWM, fill = strata)) +
+  geom_boxplot() +
+  facet_wrap(~trait, scales = "free",nrow = 1) +
+  xlab("Guild") + scale_fill_discrete(guide = "none")
+```
 
 ```{r GWVH1b, fig.cap="Number of communities significantly (p-value<0.05) different from null expectations with null model NM3 for each trait of GWVs, for canopy (C) and understorey (U) guilds, in abundance.
-See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
+See table \\@ref(tab:Traits) for abbreviations.", fig.height=4, include = F}
 GWV <- L %>%
@@
 
+```{r CWVboxplot, fig.cap="Caption", fig.height=8, include = F}
+GWV %>% 
+  select(-community, -LES, -WES) %>% 
+  melt(id.vars = "strata", 
+       variable.name = "trait",
+       value.name = "GWV") %>% 
+  ggplot(aes(x = strata, y = GWV, fill = strata)) +
+  geom_boxplot() +
+  facet_wrap(~trait, scales = "free") +
+  ggpubr::stat_compare_means(aes(label = paste("Wilcoxon, p",
+                                               ..p.format..,
+                                               ..p.signif..)))
+```
+
 ### *Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?*
 
-#### H2a. Drivers of environmental filtering on local functional composition differ between canopy and understorey tree species
+The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds (see table \@ref(tab:GWMH2a)).
+Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits (see figure \@ref(fig:WD)), and two major differences were observed between the two guilds.
+The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height and wetness affected mainly understorey guild composition.
+
+Focusing on canopy tree species, we found an increase in wood density under south-west orientation.
+We also observed a decrease in leaf area under the south-west orientation.
+Leaf area decreased with curvature, while wood density and LDMC increased with curvature.
+Finally, heigh canopy was also related to an increased leaf area.
+
+For understorey tree species, SLA significantly decreased and and leaf thickness significantly increased with curvature.
+SLA significantly increased with canopy height, while leaf thickness decreased with canopy height.
+Finally, SLA and wood density significantly decreased and and leaf area and thickness significantly increased with wetness.
 
@@
   mutate_each(funs(replace(., is.na(.), ' '))) %>% 
+  left_join( # Inclusion of R2 but should be split
+    GWM_ses %>% 
+      melt(id.vars = c('strata', 'community')) %>% 
+      rename(trait = variable) %>% 
+      rename(SES = value) %>% 
+      left_join(R) %>%
+      select(-community) %>% 
+      group_by(strata, trait) %>% 
+      do(glance(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
+                     direction = 'both', trace = 0))) %>% 
+      select(r.squared) %>% 
+      mutate(r.squared = round(r.squared, 2)) %>% 
+      rename("$R^2$" = r.squared)
+            )%>% # End of R2
   filter(trait != 'LES') %>% 
@@
 See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
-```
-
-```{r WD, fig.cap='Exemple of standard effect size (SES) of wood density  (WD) for canopy (C) tree species.
-Each subplot show  standard effect size (SES) with null model NM2 of wood density (WD) for canopy (C) tree species along one environmental variable gradient.
+Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
+```
+
+```{r WD, fig.cap='Standard effect size (SES) against curvature, for each trait and both guilds (canopy (C) and understorey (U) guilds).
+Each subplot show  standard effect size (SES) with null model NM12.
 Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
@@
 GWM_ses %>%
-  select(strata, WD, community) %>% 
   left_join(R) %>%
-  filter(strata == 'C') %>% 
   ungroup() %>%
-  select(-strata, -community) %>% 
-  melt(id.vars = 'WD') %>% 
-  ggplot(aes(value, WD)) +
+  select(strata, SLA, WD, LDMC, LA, LT, Curvature) %>%
+  melt(id.vars = c('strata', 'Curvature'),
+       variable.name = 'trait',
+       value.name = 'SES') %>%
+  ggplot(aes(Curvature, SES)) +
   geom_point() +
-  facet_wrap(~ variable, scales = "free") +
+  facet_grid(strata ~ trait, scales = "free") +
   geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') +
-  geom_hline(yintercept = 0) +
-  xlab('Environmental gradient') +
-  ylab('Standard effect size of wood density')
-```
-
-#### H2b. Strength of drivers of evironmental filtering differ between canopy and understorey tree species
+  geom_hline(yintercept = 0)
+```
+
+We found significant relationships between the standardized effect size (SES) of GWVs and the environmental predictors (see table \@ref(tab:GWVH2b)).
+Curvature increased LDMC variance in canopy.
+Wetness decreased variance of SLA and wood density in canopy and leaf area in understorey, but increased variance of wood density in understorey.
+Sout-West orientation results in decreased variance of density of leaf in understorey.
+Finally, Canopy height diminished variance of leaf density in canopy whereas basal aea increased variance of SLA in both strata.
 
@@
 See table \\@ref(tab:Traits) for abbreviations.
-Bold and  number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).')
-```
-
-<!-- ### Guild weighted means : nul model 1 -->
-
-<!-- We used a null model (NM1) which randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix) [@bernard-verdier_community_2012]. -->
-
-<!-- The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds (see table \@ref(tab:GWM1)).
-Results in presence-absence are in adequation with observed results in abundance (see **...**).
-Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits, and two major differences were observed between the two guilds.
-The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height affected only understory guild composition.
--->
-
-<!-- Focusing on canopy tree species, we found an increase in wood density under south-west orientation, both in abundance and presence-absence.
-We also observed a decrease in leaf area and thickness under the south-west orientation, but only in abundance and presence-absence respectively.
-Leaf area decreased with curvature in both abundance and presence-absence, while wood density and LDMC increased and decreased respectively with curvature in presence-absence only.
--->
-
-<!-- For understorey tree species, SLA significantly decreased and wood density and leaf thickness significantly increased with curvature, both in abundance and presence-absence.
-We also detected a significant increase and decrease in wood economic spectrum and leaf area respectively with curvature in abundance.
-SLA in both abundance and presence-absence significantly increased with canopy height, while LDMC and wood density decreased with canopy height in abundance only.
--->
-
-<!-- ```{r GWM1} -->
-<!-- GWM <- L %>%  -->
-<!--   t() %>%  -->
-<!--   data.frame() %>%  -->
-<!--   mutate(SpCode = row.names(.)) %>%  -->
-<!--   left_join(Q) %>%  -->
-<!--   select(-SpCode) %>%  -->
-<!--   melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
-<!--        variable.name = 'community', value.name = 'abundance') %>%  -->
-<!--   group_by(strata, community) %>%  -->
-<!--   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) -->
-<!-- # nQ <- array(rep(as.matrix(Q), n), -->
-<!-- #             c(dim(Q), n), -->
-<!-- #             list(row.names(Q), names(Q), 1:n)) -->
-<!-- # nQ <- array(apply(nQ, 3, function(x){ -->
-<!-- #   x[,'SpCode'] <- sample(x[,'SpCode']) -->
-<!-- #   return(x)}), c(dim(Q), n), -->
-<!-- #   list(row.names(Q), names(Q), 1:n)) -->
-<!-- # cl <- makeCluster(cores) -->
-<!-- # clusterExport(cl, list('L', 'nQ')) -->
-<!-- # nGWM <- parApply(cl, nQ, 3, function(x){ -->
-<!-- #   library(dplyr) -->
-<!-- #   library(reshape2) -->
-<!-- #   L %>% -->
-<!-- #     t() %>% -->
-<!-- #     data.frame() %>% -->
-<!-- #     mutate(SpCode = row.names(.)) %>% -->
-<!-- #     left_join(data.frame(x)) %>% -->
-<!-- #     select(-SpCode) %>% -->
-<!-- #     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
-<!-- #          variable.name = 'community', value.name = 'abundance') %>% -->
-<!-- #     group_by(strata, community) %>% -->
-<!-- #     mutate_each(funs(as.character), -abundance) %>% -->
-<!-- #     mutate_each(funs(as.numeric), -abundance) %>% -->
-<!-- #     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) -->
-<!-- # }) -->
-<!-- # stopCluster(cl) -->
-<!-- # nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))), -->
-<!-- #               c(dim(GWM[3:9]), n), -->
-<!-- #               list(row.names(GWM), names(GWM[3:9]), 1:n)) -->
-<!-- # GWM_ses <- GWM -->
-<!-- # GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T) -->
-<!-- # rm(cl, nQ, nGWM) -->
-<!-- # save(GWM_ses, file = file.path(path, 'GWM1_ses.Rdata')) -->
-<!-- load(file.path(path, 'GWM1_ses.Rdata')) -->
-<!-- GWM_ses %>%  -->
-<!--   melt(id.vars = c('strata', 'community')) %>%  -->
-<!--   rename(trait = variable) %>%  -->
-<!--   rename(SES = value) %>%  -->
-<!--   left_join(R) %>% -->
-<!--   select(-community) %>%  -->
-<!--   group_by(strata, trait) %>%  -->
-<!--   do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .),  -->
-<!--           direction = 'both', trace = 0))) %>%  -->
-<!--   group_by(strata, trait, term) %>% -->
-<!--   transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T),  -->
-<!--                           stars(p.value, ns = '   '))) %>%  -->
-<!--   dcast(strata + trait ~ term) %>%  -->
-<!--   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>%  -->
-<!--   mutate_each(funs(replace(., is.na(.), ' '))) %>%  -->
-<!--   filter(trait != 'LES') %>%  -->
-<!--   filter(trait != 'WES') %>% -->
-<!--   bold() %>%  -->
-<!--   kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).') -->
-<!-- ``` -->
-
-<!-- ```{r WD, fig.cap='Exemple of standard effect size (SES) of wood density  (WD) for canopy (C) tree species.
-Each subplot show  standard effect size (SES) of wood density (WD) for canopy (C) tree species along one environmental variable gradient.
-Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
-See table \\@ref(tab:Traits) for abbreviation.'} -->
-<!-- GWM_ses %>%  -->
-<!--   select(strata, WD, community) %>%  -->
-<!--   left_join(R) %>%  -->
-<!--   filter(strata == 'C') %>%  -->
-<!--   ungroup() %>%  -->
-<!--   select(-strata, -community) %>%  -->
-<!--   melt(id.vars = 'WD') %>%  -->
-<!--   ggplot(aes(value, WD)) + -->
-<!--   geom_point() + -->
-<!--   facet_wrap(~ variable, scales = "free") + -->
-<!--   geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') + -->
-<!--   geom_hline(yintercept = 0) + -->
-<!--   xlab('Environmental gradient') + -->
-<!--   ylab('Standard effect size of wood density') -->
-<!-- ``` -->
-
-<!-- ### Guild weighted means : nul model 3 -->
-
-<!-- We used a null model (NM3) which randomized species among guilds in the whole plot (see table \@ref(tab:GWM3)). -->
-
-<!-- We found that **curvature** was responsible for the guilds differenciation for most of the traits.
-Increased curvature resulted in denser lwood and leaves with high SLA in canopy, against lighter wood, low SLA, but bigger and thicker leaves in understorey.
-Some curvature results were congruent with **Wetness** variable.
-**South-West** orientation diminished the leaf area while increasing leaf density for canopy tree species.
-Finally, **Canopy** height and **basal area** affected in opposite ways SLA.
-Increased canopy height and decreased basal area resulted in lower SLA in canopy and increased SLA in understorey.
--->
-
-<!-- ```{r GWM3} -->
-<!-- GWM <- L %>%  -->
-<!--   t() %>%  -->
-<!--   data.frame() %>%  -->
-<!--   mutate(SpCode = row.names(.)) %>%  -->
-<!--   left_join(Q) %>%  -->
-<!--   select(-SpCode) %>%  -->
-<!--   melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
-<!--        variable.name = 'community', value.name = 'abundance') %>%  -->
-<!--   group_by(strata, community) %>%  -->
-<!--   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) -->
-<!-- # nQ <- array(rep(as.matrix(Q), n), -->
-<!-- #             c(dim(Q), n), -->
-<!-- #             list(row.names(Q), names(Q), 1:n)) -->
-<!-- # nQ <- array(apply(nQ, 3, function(x){ -->
-<!-- #   x[,'strata'] <- sample(x[,'strata']) -->
-<!-- #   return(x)}), c(dim(Q), n), -->
-<!-- #   list(row.names(Q), names(Q), 1:n)) -->
-<!-- # cl <- makeCluster(cores) -->
-<!-- # clusterExport(cl, list('L', 'nQ')) -->
-<!-- # nGWM <- parApply(cl, nQ, 3, function(x){ -->
-<!-- #   library(dplyr) -->
-<!-- #   library(reshape2) -->
-<!-- #   L %>% -->
-<!-- #     t() %>% -->
-<!-- #     data.frame() %>% -->
-<!-- #     mutate(SpCode = row.names(.)) %>% -->
-<!-- #     left_join(data.frame(x)) %>% -->
-<!-- #     select(-SpCode) %>% -->
-<!-- #     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
-<!-- #          variable.name = 'community', value.name = 'abundance') %>% -->
-<!-- #     group_by(strata, community) %>% -->
-<!-- #     mutate_each(funs(as.character), -abundance) %>% -->
-<!-- #     mutate_each(funs(as.numeric), -abundance) %>% -->
-<!-- #     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) -->
-<!-- # }) -->
-<!-- # stopCluster(cl) -->
-<!-- # nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))), -->
-<!-- #               c(dim(GWM[3:9]), n), -->
-<!-- #               list(row.names(GWM), names(GWM[3:9]), 1:n)) -->
-<!-- # GWM_ses <- GWM -->
-<!-- # GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T) -->
-<!-- # rm(cl, nQ, nGWM) -->
-<!-- # save(GWM_ses, file = file.path(path, 'GWM3_ses.Rdata')) -->
-<!-- load(file.path(path, 'GWM3_ses.Rdata')) -->
-<!-- GWM_ses %>%  -->
-<!--   melt(id.vars = c('strata', 'community')) %>%  -->
-<!--   rename(trait = variable) %>%  -->
-<!--   rename(SES = value) %>%  -->
-<!--   left_join(R) %>% -->
-<!--   select(-community) %>%  -->
-<!--   group_by(strata, trait) %>%  -->
-<!--   do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .),  -->
-<!--           direction = 'both', trace = 0))) %>%  -->
-<!--   group_by(strata, trait, term) %>% -->
-<!--   transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T),  -->
-<!--                           stars(p.value, ns = '   '))) %>%  -->
-<!--   dcast(strata + trait ~ term) %>%  -->
-<!--   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>%  -->
-<!--   mutate_each(funs(replace(., is.na(.), ' '))) %>%  -->
-<!--   filter(trait != 'LES') %>%  -->
-<!--   filter(trait != 'WES') %>% -->
-<!--   bold() %>%  -->
-<!--   kable(caption = 'Linear models of the standardized effect size (SES) for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).') -->
-<!-- ``` -->
-
-<!-- ### Guild weighted variances : nul model 2 -->
-
-<!-- The second model (NM2) randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012]. -->
-
-<!-- We found significant relationships between the standardized effect size (SES) of GWVs and the environmental predictors (see table \@ref(tab:GWV3)).
-**Wetness** decreased variance of wood density in canopy and all traits in understorey.
-**Curvature** increased LDMC variance in canopy.
-**Sout-West** orientation results in decreased variance of density of wood and leaf in understorey.
-Finally, **Canopy** hieght diminished variance of leaf are and density in canopy whereas **basal aea** increased variance of SLA in both strata and increased leaf area variance in canopy.
--->
-
-<!-- ```{r GWV3} -->
-<!-- GWV <- L %>%  -->
-<!--   t() %>%  -->
-<!--   data.frame() %>%  -->
-<!--   filter(sum(.) > 2) %>%  -->
-<!--   mutate(SpCode = row.names(.)) %>%  -->
-<!--   left_join(Q) %>%  -->
-<!--   select(-SpCode) %>%  -->
-<!--   melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
-<!--        variable.name = 'community', value.name = 'abundance') %>%  -->
-<!--   melt(id.vars = c('strata', 'community', 'abundance'), -->
-<!--        variable.name = 'trait', value.name = 'trait_value') %>%  -->
-<!--   left_join(melt(GWM, id.vars = c('strata', 'community'), -->
-<!--                  variable.name = 'trait', value.name = 'GWM')) %>%  -->
-<!--   group_by(strata, community, trait) %>%  -->
-<!--   mutate(GWV = (trait_value - GWM)^2) %>%  -->
-<!--   select(-trait_value, -GWM) %>%  -->
-<!--   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%  -->
-<!--   dcast(strata + community ~ trait) -->
-<!-- # nL <- array(replicate(n, t(apply(L, 1, function(x){ -->
-<!-- #   x[x != 0] <- sample(x[which(x != 0)]) ; return(x)}))), -->
-<!-- #   c(dim(L), n), -->
-<!-- #   list(row.names(L), names(L), 1:n)) -->
-<!-- # cl <- makeCluster(cores) -->
-<!-- # clusterExport(cl, list('nL', 'Q', 'GWM')) -->
-<!-- # nGWV <- parApply(cl, nL, 3, function(x){ -->
-<!-- #   library(dplyr) -->
-<!-- #   library(reshape2) -->
-<!-- #   x %>% -->
-<!-- #     t() %>% -->
-<!-- #     data.frame() %>% -->
-<!-- #     filter(sum(.) > 2) %>% -->
-<!-- #     mutate(SpCode = row.names(.)) %>% -->
-<!-- #     left_join(Q) %>% -->
-<!-- #     select(-SpCode) %>% -->
-<!-- #     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'), -->
-<!-- #          variable.name = 'community', value.name = 'abundance') %>% -->
-<!-- #     melt(id.vars = c('strata', 'community', 'abundance'), -->
-<!-- #          variable.name = 'trait', value.name = 'trait_value') %>% -->
-<!-- #     left_join(melt(GWM, id.vars = c('strata', 'community'), -->
-<!-- #                    variable.name = 'trait', value.name = 'GWM')) %>% -->
-<!-- #     group_by(strata, community, trait) %>% -->
-<!-- #     mutate(GWV = (trait_value - GWM)^2) %>% -->
-<!-- #     select(-trait_value, -GWM) %>% -->
-<!-- #     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% -->
-<!-- #     dcast(strata + community ~ trait) -->
-<!-- # }) -->
-<!-- # stopCluster(cl) -->
-<!-- # nGWV <- array(unlist(lapply(nGWV, function(x) as.matrix(x[3:9]))), -->
-<!-- #               c(dim(GWV[3:9]), n), -->
-<!-- #               list(row.names(GWV), names(GWV[3:9]), 1:n)) -->
-<!-- # GWV_ses <- GWV -->
-<!-- # GWV_ses[3:9] <- (GWV[3:9] - apply(nGWV, 1:2, mean, na.rm = T))/apply(nGWV, 1:2, sd, na.rm = T) -->
-<!-- # rm(cl, nL, nGWV) -->
-<!-- # save(GWV_ses, file = file.path(path, 'GWV_ses.Rdata')) -->
-<!-- load(file.path(path, 'GWV_ses.Rdata')) -->
-<!-- GWV_ses %>%  -->
-<!--   melt(id.vars = c('strata', 'community')) %>%  -->
-<!--   rename(trait = variable) %>%  -->
-<!--   rename(SES = value) %>%  -->
-<!--   left_join(R) %>% -->
-<!--   select(-community) %>%  -->
-<!--   group_by(strata, trait) %>%  -->
-<!--   do(tidy(step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .),  -->
-<!--           direction = 'both', trace = 0))) %>%  -->
-<!--   group_by(strata, trait, term) %>% -->
-<!--   transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T),  -->
-<!--                           stars(p.value, ns = '   '))) %>%  -->
-<!--   dcast(strata + trait ~ term) %>%  -->
-<!--   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>%  -->
-<!--   mutate_each(funs(replace(., is.na(.), ' '))) %>%  -->
-<!--   filter(trait != 'LES') %>%  -->
-<!--   filter(trait != 'WES') %>% -->
-<!--   bold() %>%  -->
-<!--   kable(caption = 'Linear models of the standardized effect sizes (SES) of GWV for canopy (C) and understorey (U) guilds.
-We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and  number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).') -->
-<!-- ``` -->
+Bold and  number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
+```
 
@@
 
-## **"Take home messages" to add :**
-
-- **Pantropical and multiscale WES and LES ?**
-- **Decoupled wood and leaf strategies among canopy and understorey tree species**
-- **WES among canopy tree species**
-- **LES among understorey tree species**
-
-We found back economics spectra evidenced by @wright_worldwide_2004 and @chave_towards_2009 at finer scale than @messier_trait_2016 suggested (figure 1).
+<!-- ## **"Take home messages" to add :** -->
+
+<!-- - **Pantropical and multiscale WES and LES ?** -->
+<!-- - **Decoupled wood and leaf strategies among canopy and understorey tree species** -->
+<!-- - **WES among canopy tree species** -->
+<!-- - **LES among understorey tree species** -->
+
+We found back economics spectra evidenced by @wright_worldwide_2004 and @chave_towards_2009 at finer scale than @messier_trait_2016 suggested (figure \@ref(fig:PCA)).
 This result advocate for a pantropical and multiscale oganisation of tropical plant strategies in two separated economics spectra.
 Still our results suggest that this two spectra are decoupled between canopy and understorey tree species, whereas existing in both guilds.
-Canopy tree species being the first layer of forest ecosystem interacting with external abiotic environment shows major variation in wood economics spectrum (table 3).
+Canopy tree species being the first layer of forest ecosystem interacting with external abiotic environment shows major variation in wood economics spectrum (table \@ref(tab:GWMH2a) and \@ref(tab:GWVH2b)).
 Indeed wood economic spectrum is related to mechanical resistance, disturbance tolerance and water conductivity efficiency.
 Thus understorey tree will benefit from a mircoclimate **[ref?]** generated by the presence of the canopy tree.
-In this microclimate, leaf economics spectrum is then a major axis of strategy to answer to a changing light environment and soil fertility.
+In this microclimate, leaf economics spectrum is then a major axis of strategy to answer to a changing light environment and soil fertility (table \@ref(tab:GWMH2a) and \@ref(tab:GWVH2b)).
 
@@
 
-We found an increasing trait variation within individuals, within species, and between species (table 2).
+We found an increasing trait variation within individuals, within species, and between species (table \@ref(tab:CV)).
 Our analyses thus reveal that the primary source of trait variation within and among communities is the inter-species variation, justifying the use of species mean traits for the study of tropical tree guild assembly.
@@
 
-We found a significanlty higher variance of canopy tree species around the wood economic spectrum (WES) rather than the leaf economic spectrum (LES).
-Similarly understorey tree species seemed to have higher variance around leaf economic spectra (LES) rather than wood economics spectra (WES), besides it was not significant at a level $\alpha = 0.05$ (table 3).
+We did not found significant differentiation of functional hypervolumes between canopy and understorey species.
+Still the two hypervolumes were only half overlaped with a p-value of $p = 0.118$ (see [S2: Hypervolumes of functional guilds]).
+
+Furthermore, we analyzed functional composition of $30*30m^2$ guilds from canopy and understorey, as well as the variation of this composition with environment across a 10-ha forest plot.
+We first evidenced guild differenciation in 15 to 80% of communities depending on functional traits (see figure \@ref(fig:percsign)).
+We are above a 5% type one error, and thus we conclude to a global functional differentiation between guilds within communities.
+But the differentiation differ between functional traits.
+Specific Leaf Area (SLA) was differentiated aong guilds in the majority of communities whereas Leaf Thickness (LT) was the less marked.
+This different levels of differentiation among guilds between functional traits could explain the lack of significance with the hypervolume overlap test when considering all functional traits together.
+Anyway, distribution of functional composition among communities shows a strong differentiation between the two guilds for each trait (see figure \@ref(fig:CWMboxplot)).
+Canopy guilds have larger, thinner and lighter leaves with increased SLA and denser woods, whereas understorey guilds have smaller, thicker, and more dense leaves with lower SLA and lighter woods.
+
+In addition, we foud that the variations in trait composition along the environmental gradients differ between canopy and understorey species, with a marked variation in traits associated to the leaf economic spectrum (LES) for understorey species and to the wood economics scpectrum (WES) for canopy species.
+In the former case, the curvature, wetness index, and the canopy height had a significant effect on the LES-based composition of understorey tree species, whereas southwest orientation and curvature both had a significant effect on the WES-based composition of canopy tree species (table \@ref(tab:GWMH2a) and \@ref(tab:GWVH2b)).
+
 Those results advocates for decoupled strategies between species of understorey and canopy guilds [@duque2002different].
@@
 
-Furthermore, we analyzed functional composition of $30*30m^2$ guilds from canopy and understorey, as well as the variation of this composition with environment across a 10-ha forest plot.
-We foud that the variations in trait composition along the environmental gradients differ between canopy and understorey species, with a marked variation in traits associated to the leaf economic spectrum (LES) for understorey species and to the wood economics scpectrum (WES) for canopy species.
-In the former case, the curvature and the canopy height both had a significant effect on the LES-based composition of understorey tree species, whereas southwest orientation and curvature both had a significant effect on the WES-based composition of canopy tree species (table 4).
-
 ## Environment influence guild-level functionnal composition
 
-We should note that relations found between guilds weighted means (GWM) and environment are globally weak (table 4, most of $R^2 \leq 0.1$, and see supplementary materials S4).
+We should note that relations found between guilds weighted means (GWM) and environment are globally weak (table \@ref(tab:GWMH2a), most of $R^2 \leq 0.1$).
 This low explanatory power is advocating for underlying null process related to neutral theory [@hubbell_unified_2001].
@@
 Studies highlighting environmental filtering effect often present a higher environmental variation than in our study site [@bernard-verdier_community_2012].
-In parallel, the 10*ha plot scale is really low and partially responsible for the small environmental variation.
-@kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25 ha tropical forest plots including large flat valleys.
+In parallel, the 10-ha plot scale is really low and partially responsible for the small environmental variation.
+@kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25-ha tropical forest plots including large flat valleys.
 
 Curvature is affecting the variation of leaf economics in understorey species and the variation of wood economics in canopy tree species.
-A locally convex context would accumulate more soil and attract more water, increasing soil fertility and allowing persistence of species investing more in leaf resource acquisition in understorey with higher SLA and lower LDMC, or less in wood ressources in canopy with lighter wood and bigger leaves converging more toward an homogeneous guild composition (lower CWV, unshown result).
-Curvature for some functional traits, as wood density, LDMC and SLA, is also doubled by a similar effect of the wetness variable expressing also the water accumulation in convex context resulting in increased soil fertility.
-Economics spectra are not perfectly divided between canopy and understorey, with wood and leaf respectivelly, for instance LDMC of canopy species also decreased with curvature whereas WD increased in understorey.
+A locally convex context would accumulate more soil and attract more water, increasing soil fertility and allowing persistence of species investing more in leaf resource acquisition in understorey with higher SLA and lower LT, or less in wood ressources in canopy with lighter wood and bigger leaves converging more toward an homogeneous guild composition.
+Curvature effect for understorey species is doubled by the effect of wetness index, expressing also the water accumulation in convex context resulting in increased soil fertility.
+Wet soil will result in lower SLA, lighter wood, and bigger and thicker leaves for understorey species.
+Economics spectra are not perfectly divided between canopy and understorey, with wood and leaf respectivelly, for instance woode density and leaf area of understorey species also responded to wet soils in understorey.
 
@@
 Biotic factors, illustrated here by the canopy height and the basal area, also had a significant effect on the LES-based composition of understorey tree species.
-We indeed found higher SLA values and lower LDMC values in understorey assemblages that are located under stands characterized by a tall canopy and a high basal area.
-This results contrasts with the expectation that understorey species invest in higher LDMC for higher leaf lifespan, at the expense of SLA, unde light-limited conditions [@kitajima_tissue-level_2010].
+We indeed found higher SLA values and lower LT values in understorey assemblages that are located under stands characterized by a tall canopy.
+This results contrasts with the expectation that understorey species invest in higher LDMC for higher leaf lifespan, at the expense of SLA, under light-limited conditions [@kitajima_tissue-level_2010].
 The Uppangala forest is characterized by a large heterogeneity of canopy height allowing light to penetrate more in lower strata, maybe especially more in high canopy context covering less closely understorey, resulting in those observation on the leaf economics variation of understorey species.
-**Not convincing, better explanation ?**
 
@@
 
-```{r hypervfig, fig.cap='Functional hypervolume overlap of canopy and understorey species.'}
+```{r hypervfig, fig.cap='Functional hypervolume overlap of canopy and understorey species.', fig.height=4}
 hypervolume_fig

### 2018-06-07 inst/doc/draft2.Rmd
Uppangala 0.1.9043 June 2018 pre-reorganisation push

@@
 title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
-author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
+author: "Sylvain Schmitt, Maxime Réjou-Méchain, Valérie Raevel, Ayyappan N, Balachandran N., Barathan N., and François Munoz"
 date: '`r Sys.Date()`'
 output:
+  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
 linestretch: 1.5
@@
 library(ggfortify)
+theme_set(bayesplot::theme_default())
 cores <- 16
@@
 
-In this paper, we used a trait-based approach to study separately the assembly of canopy and understorey tree species along topographical and biotic gradients in a large forest plot (10ha) loacted in Western Ghats biodiversity hotspot [@myers_biodiversity_2000], India.
+In this paper, we used a trait-based approach to study separately the assembly of canopy and understorey tree species along topographical and biotic gradients in a large forest plot (10ha) located in Western Ghats biodiversity hotspot [@myers_biodiversity_2000], India.
 Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988].
@@
 We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table \@ref(tab:Traits)).
-We choose those five traits because they proved to be good proxies of the major dimensions of ecological strategies in plants [@mark_westoby_plant_2002; @diaz_global_2015] and were relatively easy to measure on the field.
+We chose those five traits because they proved to be good proxies of the major dimensions of ecological strategies in plants [@mark_westoby_plant_2002; @diaz_global_2015] and were relatively easy to measure on the field.
 
@@
 We found an overall increase of functional trait variation from intra-individual to inter-specific levels (see table \@ref(tab:CV)), even though LA showed almost as large variability at intra-individual than at intra-specific levels ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
-given the relatively low intra-individual and intra species variation, compared to inter-species variation, using species trait average values for community analyses is warranted.
+Given the relatively low intra-individual and intra species variation, compared to inter-species variation, using species trait average values for community analyses is warranted.
 
@@
   princomp(~ LA + WD + LT + LDMC + SLA, data = ., cor = T) %>% 
-  autoplot(colour = 'grey', loadings.label.size = 6,
+  autoplot(colour = 'grey', 
+           loadings.label.size = 6,
            loadings.label.colour = 'black',
@@
 
-For understorey tree species, SLA significantly decreased and and leaf thickness significantly increased with curvature.
+For understorey tree species, SLA significantly decreased and leaf thickness significantly increased with curvature.
 SLA significantly increased with canopy height, while leaf thickness decreased with canopy height.
-Finally, SLA and wood density significantly decreased and and leaf area and thickness significantly increased with wetness.
+Finally, SLA and wood density significantly decreased and leaf area and thickness significantly increased with wetness.
 

### 2018-06-07 inst/doc/draft3.Rmd
Uppangala 0.1.9043 June 2018 pre-reorganisation push

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
author: "Sylvain Schmitt, Maxime Réjou-Méchain, Valérie Raevel, Ayyappan N, Balachandran N., Barathan N., and François Munoz"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
linestretch: 1.5
csl: mee.csl
bibliography: refs.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(parallel)
library(Uppangala)
library(raster)
library(broom)
library(vegan)
library(dplyr)
library(reshape2)
library(ggplot2)
library(ggfortify)
theme_set(bayesplot::theme_default())
cores <- detectCores() - 1
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
path <- '~/Documents/IFP/Uppangala/inst/doc/draft_save/'
n = 999 # Random repetitions
```

# Traits variation

```{r Traits}
kable(data.frame(
  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  Strategy = c('Leaf defense and investment',
        'Leaf resource capture vs.investment',
        'Leaf defense and investment',
        'Leaf resource acquisiton vs. defense',
        'Stem transport, structure and defense')
), caption = 'Functional traits measured. Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].', format = 'pandoc')
```

```{r CV}
genPFT() %>% 
  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
  melt(variable.name = 'trait') %>% 
  group_by(trait) %>% 
  summarise_each(funs(min(.,na.rm = T), max(., na.rm = T), 
                      mean(.,na.rm = T), sd(.,na.rm = T))) %>% 
  left_join(genPFT() %>% 
              select(-Sp, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Tree', 
                   variable.name = 'trait', 
                   value.name = 'Intra-individuals') %>% 
              group_by(Tree, trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Tree) %>% 
              group_by(trait) %>% 
              summarise_each(funs(mean(., na.rm = T)))) %>% 
  left_join(genPFT() %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Sp', 
                   variable.name = 'trait', value.name = 'Intra-species') %>% 
              group_by(Sp, trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Sp) %>% 
              group_by(trait) %>% 
              summarise_each(funs(mean(., na.rm = T)))) %>% 
  left_join(genPFT() %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -CE) %>% 
              group_by(Sp, SpCode) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Sp, -SpCode, -SHA, -LTD) %>% 
              melt(variable.name = 'trait', value.name = 'Inter-species') %>% 
              group_by(trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T)))) %>% 
  kable(digits = 2, caption = 'Intra-individual, intra-specific and inter-specific component of trait variation. The overall range, mean and standard deviation are provided for each trait. The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species. See table \\@ref(tab:Traits) for abbreviation.', format = 'pandoc')
```

```{r Q}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  group_by(Tree, Sp, SpCode, CE) %>% 
  summarise_each(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  group_by(Sp, SpCode) %>% 
  summarise_each(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree, -Sp, -CE, -LTD, -SHA) %>% 
  filter(SpCode != "desi")
row.names(PFT) <- PFT$SpCode
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
Q <- PFT %>% 
  left_join(data.frame(PFT.pca$scores) %>% mutate(SpCode = row.names(.))) %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% 
  left_join(genSpecies()) %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  select(SLA, LDMC, WD, LA, LT, LES, WES, strata, SpCode) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata))
```

```{r L}
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
L <- genTrees() %>% 
  mutate(com = (XY %over% com)[,1]) %>% 
  filter(!is.na(com)) %>% 
  group_by(com, SpCode) %>% 
  summarise(n = n()) %>% 
  dcast(com ~ SpCode) %>% 
  replace(is.na(.), 0)
```

```{r R}
# Env <- genEnv()
# com <- quadrats(cs = c(30,30))
# com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/share/applications')
# names(Deriv)[5] <- 'SW'
# com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
# com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
# com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
# pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
# save(pca.env, file = file.path(path, 'pca.env.Rdata'))
# R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
# row.names(R) <- com@data$id
# R$community <- row.names(R)
# save(R, file = file.path(path, 'R.Rdata'))
load(file.path(path, 'R.Rdata'))
```

# Traits covariation

```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values. First axis, in red, is named wood economic spectrum (WES) ; whereas second axis, in green, is named leaf economic spectrum (LES). See table \\@ref(tab:Traits) for abbreviation.'}
autoplot(PFT.pca, 
         data = Q %>% 
           left_join(L %>% 
                       dplyr::select(-com) %>% 
                       summarise_all(sum) %>% 
                       t() %>% 
                       data.frame() %>% 
                       rename(abundance =  ".") %>% 
                       mutate(SpCode = row.names(.))) %>% 
           filter(!is.na(WES)), 
         colour = "strata",
         size = "abundance",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0, linetype = 'Wood'), 
             col = 'red') +
  geom_vline(aes(xintercept = 0, linetype = 'Leaf'), 
             col = 'green') +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  scale_linetype_manual(name = "Economics\nSpectrum", values = c(2, 2), 
                        guide = guide_legend(override.aes = 
                                               list(color = c("red", "green")))) +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1))
```

```{r varianceTest}
data <- Q %>% select(LES, WES, strata) %>% 
  melt(id.vars = "strata", variable.name = "spectrum")
# Q %>% 
#   select(LES, WES, strata) %>% 
#   melt(id.vars = "strata", variable.name = "spectrum") %>% 
#   group_by(strata) %>% 
#   do(var_test = car::leveneTest(value ~ spectrum, data = .)) %>% 
#   tidy(var_test)
data.frame(
  Variance_LES = c(var(data[data$strata == "C" & data$spectrum == "LES", "value"], na.rm = T),
                   var(data[data$strata == "U" & data$spectrum == "LES", "value"], na.rm = T)),
  Variance_WES = c(var(data[data$strata == "C" & data$spectrum == "WES", "value"], na.rm = T),
                   var(data[data$strata == "U" & data$spectrum == "WES", "value"], na.rm = T)),
  Bartlet_test = c(bartlett.test(value ~ spectrum, data[data$strata == "C",])$p.value,
                   bartlett.test(value ~ spectrum, data[data$strata == "U",])$p.value),
  Leven_test = c(car::leveneTest(value ~ spectrum, data[data$strata == "C",])$`Pr(>F)`[1],
                 car::leveneTest(value ~ spectrum, data[data$strata == "U",])$`Pr(>F)`[1]),
  row.names = c("Canopy", "Understorey")
) %>% kable(col.names = c("Variance of LES variance", "Variance of WES", 
                          "P-value of Bartlet's test", "P-value of Leven's test"),
            caption = "Test of variance differences betweenn guilds and economic spectra")
```

# Guilds hypervolume

```{r hypervolumes}
# library(hypervolume)
# bd <- Q %>% 
#   filter(!is.na(WES)) %>% 
#   select(-LES, -WES, -SpCode, -strata) %>% 
#   summarise_all(funs(sd)) %>% 
#   unlist()
# hv <- hypervolume_set(hypervolume(na.omit(Q[Q$strata == "C",names(bd)]), 
#                                   kde.bandwidth = bd, verbose = F), 
#                       hypervolume(na.omit(Q[Q$strata == "U",names(bd)]), 
#                                   kde.bandwidth = bd, verbose = F), 
#                       check.memory = F, verbose = F)
# plot(hv)
# rm(hv)
```

# Guilds functional differences

```{r percsign, fig.cap="Percentage of communities with indivdual traits significantly different between guilds (for canopy (C) and understorey (U) guilds). Test realized with an unpaired wilcoxon two sided test, significance was considered for p-values under the level $\\alpha = 0.05$ .See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
L %>% 
  rename(community = com) %>% 
  filter(!(community %in% c("C105", "C11",  "C26",  "C5" ))) %>% 
  melt(id.vars = "community", 
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  filter(abundance > 0) %>% 
  tidyr::uncount(abundance) %>% 
  left_join(Q) %>% 
  select(-LES, -WES) %>% 
  melt(id.vars = c("community", "strata", "SpCode"),
       variable.name = "trait") %>% 
  filter(!is.na(value)) %>% 
  group_by(community, trait) %>% 
  do(wilcoxon = wilcox.test(value ~ strata, .)) %>% 
  tidy(wilcoxon) %>% 
  group_by(trait) %>% 
  summarise(n_sign = sum(p.value < 0.05), n = n()) %>% 
  mutate(n_perc = n_sign/n*100)  %>% 
  ggplot(aes(trait, n_perc)) +
  geom_bar(stat = "identity") +
  # coord_flip() +
  xlab("Functional trait") +
  ylab("Percentage of significant communities")
```

```{r GWMboxplot, fig.cap="Boxplot of guilds weighted means (GWM) among communties for each traits (for canopy (C) and understorey (U) guilds). See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com) %>% 
  left_join(Q) %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  select(-community, -LES, -WES) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free", nrow = 1) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(..p.signif..))) 
```

# Environment effect on guilds

```{r GWMses}
GWM <- L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com) %>% 
  left_join(Q) %>% 
  select(-SpCode) %>% 
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>% 
  group_by(strata, community) %>% 
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
nQ <- array(rep(as.matrix(Q), n),
            c(dim(Q), n),
            list(row.names(Q), names(Q), 1:n))
nQ <- array(apply(nQ, 3, function(x){
  x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
  x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
  return(x)
  }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
cl <- makeCluster(cores)
clusterExport(cl, list('L', 'nQ'))
nGWM <- parApply(cl, nQ, 3, function(x){
  library(dplyr)
  library(reshape2)
  L %>%
    melt(id.var = "com",
         variable.name = "SpCode",
         value.name = "abundance") %>%
    dcast(SpCode ~ com) %>%
    left_join(data.frame(x)) %>%
    select(-SpCode) %>%
    melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
         variable.name = 'community', value.name = 'abundance') %>%
    group_by(strata, community) %>%
    mutate_each(funs(as.character), -abundance) %>%
    mutate_each(funs(as.numeric), -abundance) %>%
    summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
})
stopCluster(cl)
nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
              c(dim(GWM[3:9]), n),
              list(row.names(GWM), names(GWM[3:9]), 1:n))
rm(cl, nQ)
save(nGWM, file = file.path(path, 'nGWM.Rdata'))
load(file.path(path, 'nGWM.Rdata'))
GWM_ses <- GWM
GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
```

```{r LM1}
GWM_ses %>% 
  filter(!is.na(strata)) %>% 
  melt(id.vars = c('strata', 'community')) %>% 
  rename(trait = variable) %>% 
  rename(SES = value) %>% 
  left_join(R) %>%
  group_by(strata, trait) %>% 
  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
                     direction = 'both', trace = 0)) %>% 
  tidy(reg) %>% 
  group_by(strata, trait, term) %>%
  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
                          stars(p.value, ns = '   '))) %>% 
  dcast(strata + trait ~ term) %>% 
  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
  left_join( # Inclusion of R2 but should be split
    GWM_ses %>% 
      filter(!is.na(strata)) %>% 
      melt(id.vars = c('strata', 'community')) %>% 
      rename(trait = variable) %>% 
      rename(SES = value) %>% 
      left_join(R) %>%
      group_by(strata, trait) %>% 
      do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
                    direction = 'both', trace = 0)) %>% 
      glance(reg) %>% 
      select(r.squared) %>% 
      mutate(r.squared = round(r.squared, 2)) %>% 
      rename("$R^2$" = r.squared)
  )%>% # End of R2
  filter(trait != 'LES') %>% 
  filter(trait != 'WES') %>%
  bold() %>% 
  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown. See table \\@ref(tab:Traits) for abbreviations. Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
```

```{r LM2}
GWM_ses %>% 
  filter(!is.na(strata)) %>% 
  melt(id.vars = c('strata', 'community')) %>% 
  rename(trait = variable) %>% 
  rename(SES = value) %>% 
  left_join(R) %>%
  group_by(strata, trait) %>% 
  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + BA, .), 
                     direction = 'both', trace = 0)) %>% 
  tidy(reg) %>% 
  group_by(strata, trait, term) %>%
  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
                          stars(p.value, ns = '   '))) %>% 
  dcast(strata + trait ~ term) %>% 
  select(strata, trait, Slope, Curvature, Wetness, SW, BA) %>% 
  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
  left_join( # Inclusion of R2 but should be split
    GWM_ses %>% 
      filter(!is.na(strata)) %>% 
      melt(id.vars = c('strata', 'community')) %>% 
      rename(trait = variable) %>% 
      rename(SES = value) %>% 
      left_join(R) %>%
      group_by(strata, trait) %>% 
      do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + BA, .), 
                    direction = 'both', trace = 0)) %>% 
      glance(reg) %>% 
      select(r.squared) %>% 
      mutate(r.squared = round(r.squared, 2)) %>% 
      rename("$R^2$" = r.squared)
  )%>% # End of R2
  filter(trait != 'LES') %>% 
  filter(trait != 'WES') %>%
  bold() %>% 
  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown. See table \\@ref(tab:Traits) for abbreviations. Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
```

```{r LM3}
GWM_ses %>% 
  filter(!is.na(strata)) %>% 
  melt(id.vars = c('strata', 'community')) %>% 
  rename(trait = variable) %>% 
  rename(SES = value) %>% 
  left_join(R) %>%
  group_by(strata, trait) %>% 
  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                     direction = 'both', trace = 0)) %>% 
  tidy(reg) %>% 
  group_by(strata, trait, term) %>%
  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
                          stars(p.value, ns = '   '))) %>% 
  dcast(strata + trait ~ term) %>% 
  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
  left_join( # Inclusion of R2 but should be split
    GWM_ses %>% 
      filter(!is.na(strata)) %>% 
      melt(id.vars = c('strata', 'community')) %>% 
      rename(trait = variable) %>% 
      rename(SES = value) %>% 
      left_join(R) %>%
      group_by(strata, trait) %>% 
      do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                    direction = 'both', trace = 0)) %>% 
      glance(reg) %>% 
      select(r.squared) %>% 
      mutate(r.squared = round(r.squared, 2)) %>% 
      rename("$R^2$" = r.squared)
  )%>% # End of R2
  filter(trait != 'LES') %>% 
  filter(trait != 'WES') %>%
  bold() %>% 
  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown. See table \\@ref(tab:Traits) for abbreviations. Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
```

```{r WD, fig.cap='Standard effect size (SES) against curvature, for each trait and both guilds (canopy (C) and understorey (U) guilds). Each subplot show  standard effect size (SES) with null model NM12. Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant. See table \\@ref(tab:Traits) for abbreviation.'}
GWM_ses %>%
  filter(!is.na(strata)) %>% 
  left_join(R) %>%
  ungroup() %>%
  select(strata, SLA, WD, LDMC, LA, LT, Curvature) %>%
  melt(id.vars = c('strata', 'Curvature'),
       variable.name = 'trait',
       value.name = 'SES') %>%
  ggplot(aes(Curvature, SES)) +
  geom_point() +
  facet_grid(strata ~ trait, scales = "free") +
  geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') +
  geom_hline(yintercept = 0) 
```

# Environment selection

```{r PCAenv, fig.cap='Principal component analysis (PCA) performed on topgraphical variables. SW stand for southwesterness.'}
load(file.path(path, 'pca.env.Rdata'))
pca.env %>% 
  autoplot(colour = 'grey', loadings.label.size = 6,
           loadings.label.colour = 'black',
           loadings = T, loadings.label = T, 
           loadings.colour = 'black',
           loadings.label.vjust = 2) +
  coord_equal() +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  xlab('Axe 1 − 42.52 %') + ylab('Axe 2 − 23.44 %') 
```

### 2018-06-07 draft.Rmd
Uppangala 0.1.9043 June 2018 June reorganisation

--- 2018-06-07 inst/doc/draft3.Rmd
@@
   sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
-path <- '~/Documents/IFP/Uppangala/inst/doc/draft_save/'
 n = 999 # Random repetitions
@@
 # save(R, file = file.path(path, 'R.Rdata'))
-load(file.path(path, 'R.Rdata'))
+load(file.path("draft_save", 'R.Rdata'))
 ```
@@
 rm(cl, nQ)
-save(nGWM, file = file.path(path, 'nGWM.Rdata'))
-load(file.path(path, 'nGWM.Rdata'))
+save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
+load(file.path("draft_save", 'nGWM.Rdata'))
 GWM_ses <- GWM
@@
 ```{r PCAenv, fig.cap='Principal component analysis (PCA) performed on topgraphical variables. SW stand for southwesterness.'}
-load(file.path(path, 'pca.env.Rdata'))
+load(file.path("draft_save", 'pca.env.Rdata'))
 pca.env %>% 
@@
 ```
+
+# Why not hypervolumes

### 2018-06-08 draft.Rmd
pre-test Maxime

@@
 linestretch: 1.5
-csl: mee.csl
-bibliography: refs.bib
+csl: bibliography/mee.csl
+bibliography: bibliography/refs.bib
 link-citations: yes
@@
 library(parallel)
-library(Uppangala)
 library(raster)
@@
 library(ggfortify)
+sapply(list.files("R", full.names = T), source)
 theme_set(bayesplot::theme_default())
@@
 
-# Traits variation
+# Traits
 
 ```{r Traits}
-kable(data.frame(
+genPFT() %>% 
+  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
+  melt(variable.name = 'Abbreviation') %>% 
+      left_join(data.frame(
     Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
@@
                  'Stem transport, structure and defense')
-), caption = 'Functional traits measured.
-Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].', format = 'pandoc')
-```
-
-```{r CV}
-genPFT() %>% 
-  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
-  melt(variable.name = 'trait') %>% 
-  group_by(trait) %>% 
-  summarise_each(funs(min(.,na.rm = T), max(., na.rm = T), 
-                      mean(.,na.rm = T), sd(.,na.rm = T))) %>% 
+  )) %>% 
+  group_by(Trait, Abbreviation, Unit, Strategy) %>% 
+  summarise_if(is.numeric, funs(min(.,na.rm = T), max(., na.rm = T))) %>% 
+  mutate(range = paste(round(min,2), "-", round(max,2))) %>% 
+  select(-min, -max) %>% 
   left_join(genPFT() %>% 
@@
               melt(id.vars = 'Tree', 
-                   variable.name = 'trait', 
+                   variable.name = 'Abbreviation', 
                    value.name = 'Intra-individuals') %>% 
-              group_by(Tree, trait) %>% 
+              group_by(Tree, Abbreviation) %>% 
               summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
@@
               select(-Tree) %>% 
-              group_by(trait) %>% 
+              group_by(Abbreviation) %>% 
               summarise_each(funs(mean(., na.rm = T)))) %>% 
@@
               melt(id.vars = 'Sp', 
-                   variable.name = 'trait', value.name = 'Intra-species') %>% 
-              group_by(Sp, trait) %>% 
+                   variable.name = 'Abbreviation', value.name = 'Intra-species') %>% 
+              group_by(Sp, Abbreviation) %>% 
               summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
@@
               select(-Sp) %>% 
-              group_by(trait) %>% 
+              group_by(Abbreviation) %>% 
               summarise_each(funs(mean(., na.rm = T)))) %>% 
@@
               select(-Sp, -SpCode, -SHA, -LTD) %>% 
-              melt(variable.name = 'trait', value.name = 'Inter-species') %>% 
-              group_by(trait) %>% 
+              melt(variable.name = 'Abbreviation', value.name = 'Inter-species') %>% 
+              group_by(Abbreviation) %>% 
               summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T)))) %>%
-  kable(digits = 2, caption = 'Intra-individual, intra-specific and inter-specific component of trait variation.
-The overall range, mean and standard deviation are provided for each trait.
-The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species.
-See table \\@ref(tab:Traits) for abbreviation.', format = 'pandoc')
+  mutate(`Intra-individuals` = ifelse(Abbreviation == "WD", NA, `Intra-individuals`)) %>% 
+  kable(digits = 2, format.args = list(big.mark = " "), format = 'pandoc',
+        caption = 'Functional traits measured.
+Traits with their abbreviation, standard unit, associated species strategy, range of values, intra-individual, intra-specific and inter-specific component of variation.
+Species strategy are precised for functioning, survival and reproduction [@baraloto_decoupled_2010].
+The rlast columns represents the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species.')
+
 ```
@@
   ungroup() %>% 
-  select(-Tree, -Sp, -CE, -LTD, -SHA) %>% 
-  filter(SpCode != "desi")
+  select(-Tree, -Sp, -CE, -LTD, -SHA)
 row.names(PFT) <- PFT$SpCode
@@
 
-# Traits covariation
+# PCA
 
 ```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values.
-First axis, in red, is named wood economic spectrum (WES) ; whereas second axis, in green, is named leaf economic spectrum (LES).
+Marginal plots represent the density distribution of species on the two principal component axis for each morphotype.
 See table \\@ref(tab:Traits) for abbreviation.'}
-autoplot(PFT.pca, 
+pca.plot <- autoplot(PFT.pca, 
          data = Q %>% 
@@
   coord_equal() +
-  geom_hline(aes(yintercept = 0, linetype = 'Wood'), 
-             col = 'red') +
-  geom_vline(aes(xintercept = 0, linetype = 'Leaf'), 
-             col = 'green') +
+  geom_hline(aes(yintercept = 0), col = 'lightgrey') +
+  geom_vline(aes(xintercept = 0), col = 'lightgrey') +
   xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
-  scale_linetype_manual(name = "Economics\nSpectrum", values = c(2, 2), 
-                        guide = guide_legend(override.aes = 
-                                               list(color = c("red", "green")))) +
   theme(legend.position = c(1, 1), 
         legend.justification = c(1, 1))
+spectrum.density <-pca.plot$data %>% 
+  select(Comp.1, Comp.2, strata) %>% 
+  rename(WES = Comp.1, LES = Comp.2) %>% 
+  melt(id.vars = "strata", variable.name = "spectrum") %>% 
+  ggplot(aes(value, color = spectrum, fill = spectrum)) +
+  geom_density(alpha = 0.5) + 
+  facet_wrap(~strata, nrow = 2) +
+  coord_flip()
+cowplot::plot_grid(pca.plot, spectrum.density, rel_widths = c(2,1))
 ```
@@
 ```{r varianceTest}
-data <- Q %>% select(LES, WES, strata) %>% 
+data <- pca.plot$data %>% 
+  select(Comp.1, Comp.2, strata) %>% 
+  # rename(WES = Comp.1, LES = Comp.2) %>% 
+  rename(WES = Comp.1, LES = Comp.2) %>% # ! test !
   melt(id.vars = "strata", variable.name = "spectrum")
-# Q %>% 
-#   select(LES, WES, strata) %>% 
-#   melt(id.vars = "strata", variable.name = "spectrum") %>% 
-#   group_by(strata) %>% 
-#   do(var_test = car::leveneTest(value ~ spectrum, data = .)) %>% 
-#   tidy(var_test)
-data.frame(
-  Variance_LES = c(var(data[data$strata == "C" & data$spectrum == "LES", "value"], na.rm = T),
-                   var(data[data$strata == "U" & data$spectrum == "LES", "value"], na.rm = T)),
-  Variance_WES = c(var(data[data$strata == "C" & data$spectrum == "WES", "value"], na.rm = T),
-                   var(data[data$strata == "U" & data$spectrum == "WES", "value"], na.rm = T)),
-  Bartlet_test = c(bartlett.test(value ~ spectrum, data[data$strata == "C",])$p.value,
-                   bartlett.test(value ~ spectrum, data[data$strata == "U",])$p.value),
-  Leven_test = c(car::leveneTest(value ~ spectrum, data[data$strata == "C",])$`Pr(>F)`[1],
-                 car::leveneTest(value ~ spectrum, data[data$strata == "U",])$`Pr(>F)`[1]),
-  row.names = c("Canopy", "Understorey")
-) %>% kable(col.names = c("Variance of LES variance", "Variance of WES", 
-                          "P-value of Bartlet's test", "P-value of Leven's test"),
-            caption = "Test of variance differences betweenn guilds and economic spectra")
+Variance <- list(
+  LES = list(
+    C = round(var(data[data$strata == "C" & data$spectrum == "LES", "value"], na.rm = T), 3),
+    U = round(var(data[data$strata == "U" & data$spectrum == "LES", "value"], na.rm = T), 3)),
+ WES = list(
+   C = round(var(data[data$strata == "C" & data$spectrum == "WES", "value"], na.rm = T), 3),
+   U = round(var(data[data$strata == "U" & data$spectrum == "WES", "value"], na.rm = T) ,3))
+)
+Leven_test <- list(
+  C = round(car::leveneTest(value ~ spectrum, data[data$strata == "C",])$`Pr(>F)`[1], 3),
+  U = round(car::leveneTest(value ~ spectrum, data[data$strata == "U",])$`Pr(>F)`[1], 3)
+)
 ```
 
-# Guilds hypervolume
+Canopy species vary significantly more on the wood than the leaf economic spectrum (`r Variance$LES$C` vs `r Variance$WES$C`, Levene's test p-value=`r Leven_test$C`).
+In parrallel, understorey species show a tendency to vary more on the leaf than the wood economic spectrum (`r Variance$LES$U` vs `r Variance$WES$U`, Levene's test p-value=`r Leven_test$U`).
 
-```{r hypervolumes}
-# library(hypervolume)
-# bd <- Q %>% 
-#   filter(!is.na(WES)) %>% 
-#   select(-LES, -WES, -SpCode, -strata) %>% 
-#   summarise_all(funs(sd)) %>% 
-#   unlist()
-# hv <- hypervolume_set(hypervolume(na.omit(Q[Q$strata == "C",names(bd)]), 
-#                                   kde.bandwidth = bd, verbose = F), 
-#                       hypervolume(na.omit(Q[Q$strata == "U",names(bd)]), 
-#                                   kde.bandwidth = bd, verbose = F), 
-#                       check.memory = F, verbose = F)
-# plot(hv)
-# rm(hv)
-```
-
-# Guilds functional differences
-
-```{r percsign, fig.cap="Percentage of communities with indivdual traits significantly different between guilds (for canopy (C) and understorey (U) guilds).
-Test realized with an unpaired wilcoxon two sided test, significance was considered for p-values under the level $\\alpha = 0.05$ .See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
-L %>% 
-  rename(community = com) %>% 
-  filter(!(community %in% c("C105", "C11",  "C26",  "C5" ))) %>% 
-  melt(id.vars = "community", 
-       variable.name = "SpCode",
-       value.name = "abundance") %>% 
-  filter(abundance > 0) %>% 
-  tidyr::uncount(abundance) %>% 
-  left_join(Q) %>% 
-  select(-LES, -WES) %>% 
-  melt(id.vars = c("community", "strata", "SpCode"),
-       variable.name = "trait") %>% 
-  filter(!is.na(value)) %>% 
-  group_by(community, trait) %>% 
-  do(wilcoxon = wilcox.test(value ~ strata, .)) %>% 
-  tidy(wilcoxon) %>% 
-  group_by(trait) %>% 
-  summarise(n_sign = sum(p.value < 0.05), n = n()) %>% 
-  mutate(n_perc = n_sign/n*100)  %>% 
-  ggplot(aes(trait, n_perc)) +
-  geom_bar(stat = "identity") +
-  # coord_flip() +
-  xlab("Functional trait") +
-  ylab("Percentage of significant communities")
-```
+# Guilds
 
@@
   dcast(SpCode ~ com) %>% 
-  left_join(Q) %>%
+  inner_join(Q) %>%
   select(-SpCode) %>%
@@
 
-# Environment effect on guilds
+# LM
 
@@
   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
-nQ <- array(rep(as.matrix(Q), n),
-            c(dim(Q), n),
-            list(row.names(Q), names(Q), 1:n))
-nQ <- array(apply(nQ, 3, function(x){
-  x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
-  x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
-  return(x)
-  }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
-cl <- makeCluster(cores)
-clusterExport(cl, list('L', 'nQ'))
-nGWM <- parApply(cl, nQ, 3, function(x){
-  library(dplyr)
-  library(reshape2)
-  L %>%
-    melt(id.var = "com",
-         variable.name = "SpCode",
-         value.name = "abundance") %>%
-    dcast(SpCode ~ com) %>%
-    left_join(data.frame(x)) %>%
-    select(-SpCode) %>%
-    melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-         variable.name = 'community', value.name = 'abundance') %>%
-    group_by(strata, community) %>%
-    mutate_each(funs(as.character), -abundance) %>%
-    mutate_each(funs(as.numeric), -abundance) %>%
-    summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
-})
-stopCluster(cl)
-nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
-              c(dim(GWM[3:9]), n),
-              list(row.names(GWM), names(GWM[3:9]), 1:n))
-rm(cl, nQ)
-save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
+# nQ <- array(rep(as.matrix(Q), n),
+#             c(dim(Q), n),
+#             list(row.names(Q), names(Q), 1:n))
+# nQ <- array(apply(nQ, 3, function(x){
+#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
+#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
+#   return(x)
+#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('L', 'nQ'))
+# nGWM <- parApply(cl, nQ, 3, function(x){
+#   library(dplyr)
+#   library(reshape2)
+#   L %>%
+#     melt(id.var = "com",
+#          variable.name = "SpCode",
+#          value.name = "abundance") %>%
+#     dcast(SpCode ~ com) %>%
+#     left_join(data.frame(x)) %>%
+#     select(-SpCode) %>%
+#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+#          variable.name = 'community', value.name = 'abundance') %>%
+#     group_by(strata, community) %>%
+#     mutate_each(funs(as.character), -abundance) %>%
+#     mutate_each(funs(as.numeric), -abundance) %>%
+#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# })
+# stopCluster(cl)
+# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
+#               c(dim(GWM[3:9]), n),
+#               list(row.names(GWM), names(GWM[3:9]), 1:n))
+# rm(cl, nQ)
+# save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
 load(file.path("draft_save", 'nGWM.Rdata'))
@@
 
-```{r LM1}
-GWM_ses %>% 
-  filter(!is.na(strata)) %>% 
-  melt(id.vars = c('strata', 'community')) %>% 
-  rename(trait = variable) %>% 
-  rename(SES = value) %>% 
-  left_join(R) %>%
-  group_by(strata, trait) %>% 
-  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
-                     direction = 'both', trace = 0)) %>% 
-  tidy(reg) %>% 
-  group_by(strata, trait, term) %>%
-  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
-                          stars(p.value, ns = '   '))) %>% 
-  dcast(strata + trait ~ term) %>% 
-  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy, BA) %>% 
-  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
-  left_join( # Inclusion of R2 but should be split
-    GWM_ses %>% 
-      filter(!is.na(strata)) %>% 
-      melt(id.vars = c('strata', 'community')) %>% 
-      rename(trait = variable) %>% 
-      rename(SES = value) %>% 
-      left_join(R) %>%
-      group_by(strata, trait) %>% 
-      do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy + BA, .), 
-                    direction = 'both', trace = 0)) %>% 
-      glance(reg) %>% 
-      select(r.squared) %>% 
-      mutate(r.squared = round(r.squared, 2)) %>% 
-      rename("$R^2$" = r.squared)
-  )%>% # End of R2
-  filter(trait != 'LES') %>% 
-  filter(trait != 'WES') %>%
-  bold() %>% 
-  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
-```
-
-```{r LM2}
-GWM_ses %>% 
-  filter(!is.na(strata)) %>% 
-  melt(id.vars = c('strata', 'community')) %>% 
-  rename(trait = variable) %>% 
-  rename(SES = value) %>% 
-  left_join(R) %>%
-  group_by(strata, trait) %>% 
-  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + BA, .), 
-                     direction = 'both', trace = 0)) %>% 
-  tidy(reg) %>% 
-  group_by(strata, trait, term) %>%
-  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
-                          stars(p.value, ns = '   '))) %>% 
-  dcast(strata + trait ~ term) %>% 
-  select(strata, trait, Slope, Curvature, Wetness, SW, BA) %>% 
-  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
-  left_join( # Inclusion of R2 but should be split
-    GWM_ses %>% 
-      filter(!is.na(strata)) %>% 
-      melt(id.vars = c('strata', 'community')) %>% 
-      rename(trait = variable) %>% 
-      rename(SES = value) %>% 
-      left_join(R) %>%
-      group_by(strata, trait) %>% 
-      do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + BA, .), 
-                    direction = 'both', trace = 0)) %>% 
-      glance(reg) %>% 
-      select(r.squared) %>% 
-      mutate(r.squared = round(r.squared, 2)) %>% 
-      rename("$R^2$" = r.squared)
-  )%>% # End of R2
-  filter(trait != 'LES') %>% 
-  filter(trait != 'WES') %>%
-  bold() %>% 
-  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
-```
-
-```{r LM3}
+```{r LM}
 GWM_ses %>% 
@@
   kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
-We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA), and only the selected predictors are shown.
+We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), and canopy height (Canopy), and only the selected predictors are shown.
 See table \\@ref(tab:Traits) for abbreviations.
@@
 
-```{r WD, fig.cap='Standard effect size (SES) against curvature, for each trait and both guilds (canopy (C) and understorey (U) guilds).
-Each subplot show  standard effect size (SES) with null model NM12.
-Continuous line represents a null SES, whereas dashed line represents limits of the confidence interval above which SES are considered as significant.
-See table \\@ref(tab:Traits) for abbreviation.'}
-GWM_ses %>%
-  filter(!is.na(strata)) %>% 
-  left_join(R) %>%
-  ungroup() %>%
-  select(strata, SLA, WD, LDMC, LA, LT, Curvature) %>%
-  melt(id.vars = c('strata', 'Curvature'),
-       variable.name = 'trait',
-       value.name = 'SES') %>%
-  ggplot(aes(Curvature, SES)) +
-  geom_point() +
-  facet_grid(strata ~ trait, scales = "free") +
-  geom_hline(yintercept = c(0.95, -0.95), linetype = 'dashed') +
-  geom_hline(yintercept = 0) 
-```
-
-# Environment selection
+# Environment
 
@@
 # Why not hypervolumes
+
+```{r hypervolumes}
+# library(hypervolume)
+# bd <- Q %>% 
+#   filter(!is.na(WES)) %>% 
+#   select(-LES, -WES, -SpCode, -strata) %>% 
+#   summarise_all(funs(sd)) %>% 
+#   unlist()
+# hv <- hypervolume_set(hypervolume(na.omit(Q[Q$strata == "C",names(bd)]), 
+#                                   kde.bandwidth = bd, verbose = F), 
+#                       hypervolume(na.omit(Q[Q$strata == "U",names(bd)]), 
+#                                   kde.bandwidth = bd, verbose = F), 
+#                       check.memory = F, verbose = F)
+# plot(hv)
+# rm(hv)
+```
+
+* Outliers make quicly overlapping hypervolumes besides the density stay low in those area
+* Hypervolumes consider uncorrelated axis, which is not true for functional traits, and blur strategy axes highlighted by economic spectra

### 2018-06-08 R/draft.R

@@ -1,200 +1,0 @@
-## Atypic communtiy metrics distributions##
-# CWM.ab <- CWM$abundance
-# nCWM.ab <- nCWM[[1]]
-# nCWM.ab <- lapply(nCWM.ab, `[`, names(nCWM.ab[[1]]))
-# nCWM.ab <- apply(do.call(rbind, nCWM.ab), 2, as.list)
-# nCWM.ab <- lapply(nCWM.ab, function(x){do.call('cbind', x)})
-# nCWM.ab <- lapply(nCWM.ab, data.frame)
-# CWM.ab.ses <- data.frame(mapply(function(x,y){
-#   mapply(function(p,q){
-#     (p - mean(q))/sd(q)
-#   }, p = as.list(x), q = as.list(data.frame(t(y))))
-# }, x = CWM.ab, y = nCWM.ab))
-# extrem_dist <- function(t, ses, l, q, cwm, cwv){
-#   com <- row.names(ses)[c(which(ses[t] == min(ses[t], na.rm = T)), which(ses[t] == max(ses[t], na.rm = T)))]
-#   dist.com <- apply(l[com,], 1, function(x){rep(t(q[t]),x)})
-#   cwm.com <- cwm[[t]][com]
-#   cwv.com <- cwv[[t]][com]
-#   for(c in com){
-#     x <- range(q[t], na.rm = T)
-#     dth <- dnorm(seq(x[1], x[2], 0.01), mean = cwm.com[c], sd = cwv.com[c])
-#     dth <- dth / max(dth)
-#     h <- hist(dist.com[[c]], plot = F)
-#     h$counts <- h$counts / max(h$counts)
-#     plot(h, col = 'lightgrey', main = c, xlab = t, xlim = c(x[1],x[2]), ylim = c(0,max(max(h$counts),max(dth))))
-#     lines(x = seq(x[1], x[2], 0.01), y = dth, col = 'red', lwd = 2)
-#   }
-# }
-#
-# par(mfrow = c(3,2))
-# t = 'Thick'
-# extrem_dist(t, CWM.ab.ses, L$abundance, Q, CWM$abundance, CWV)
-# extrem_dist(t, CWV.ses, L$abundance, Q, CWM$abundance, CWV)
-# extrem_dist(t, CVNND.ses, L$abundance, Q, CWM$abundance, CWV)
-
-## Atypic SES metrics community composition##
-# which(apply(1 - CWV.pval, 2, p.adjust, method = 'fdr') < 0.025, arr.ind = T)
[…]

### 2018-06-08 R/genPFT.R
Uppangala 0.1.9044 Maxime correction source files

@@ -13,7 +13,7 @@ NULL
 #'
 genPFT <- function(){
   # Opening
-  path <- system.file('extdata', 'PFT.csv', package = 'Uppangala')
+  path <- file.path('data', 'PFT.csv')
   PFT <- read.csv(path, stringsAsFactors = F, na.strings = c('#N/A', 'NA'))
   PFT <- PFT[-c(13:14)] # Deleting SD & VD in waiting of measurements
   names(PFT)[3:4] <- c('Sp', 'SpCode')

### 2018-06-08 R/genSpecies.R

@@ -11,7 +11,7 @@
 genSpecies <- function(){
 
   # Opening
-  path <- system.file('extdata', 'Species.csv', package = 'Uppangala')
+  path <- file.path('data', 'Species.csv')
   species <- read.table(path,sep=",",dec=".", stringsAsFactors =F, h = T) # Dans BDD Uppangala
   # Formatting table
   species <- species[c('Sp_code', 'Sp_code_large', 'Latin_name', 'Family', 'Genus', 'Species', 'Pot_strata', 'old_Sp_code')]
@@ -20,7 +20,7 @@ genSpecies <- function(){
   species[duplicated(species$SpCode),1] <- species[duplicated(species$SpCode),8]
   rownames(species) <- species$SpCode
   # Opening strata
-  path <- system.file('extdata', 'UPSP_Species_list.csv', package = 'Uppangala')
+  path <- file.path('data', 'UPSP_Species_list.csv')
   strata <- read.table(path,sep=",",dec=".", stringsAsFactors =F, h = T) # Pelissier et al, 2011
   strata <- strata[c(1:100),]
   row.names(strata) <- strata$SpCode

### 2018-07-09 draft.Rmd
Uppangala 0.1.9047 Bulletproofing code & ToDo.md

@@
 genPFT() %>% 
+  mutate(id = row.names(.)) %>% 
   select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
-  melt(variable.name = 'Abbreviation') %>% 
+  melt(id.vars = "id", variable.name = 'Abbreviation') %>% 
+  mutate(Abbreviation = as.character(Abbreviation)) %>% 
   left_join(data.frame(
@@
                  'Leaf resource acquisiton vs. defense',
-                 'Stem transport, structure and defense')
-  )) %>% 
+                 'Stem transport, structure and defense'),
+    stringsAsFactors = F
+  ),  by = "Abbreviation") %>% 
   group_by(Trait, Abbreviation, Unit, Strategy) %>% 
@@
                    value.name = 'Intra-individuals') %>% 
+              mutate(Abbreviation = as.character(Abbreviation)) %>% 
               group_by(Tree, Abbreviation) %>% 
-              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
+              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
               ungroup() %>% 
@@
               group_by(Abbreviation) %>% 
-              summarise_each(funs(mean(., na.rm = T)))) %>% 
+              summarise_all(funs(mean(., na.rm = T))),  by = "Abbreviation") %>% 
   left_join(genPFT() %>% 
               group_by(Tree, Sp, SpCode, CE) %>% 
-              summarise_each(funs(mean)) %>% 
+              summarise_all(funs(mean)) %>% 
               ungroup() %>% 
@@
                    variable.name = 'Abbreviation', value.name = 'Intra-species') %>% 
+              mutate(Abbreviation = as.character(Abbreviation)) %>% 
               group_by(Sp, Abbreviation) %>% 
-              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
+              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
               ungroup() %>% 
@@
               group_by(Sp, SpCode) %>% 
-              summarise_each(funs(mean)) %>% 
+              summarise_all(funs(mean)) %>% 
               ungroup() %>% 
               select(-Sp, -SpCode, -SHA, -LTD) %>%
-              melt(variable.name = 'Abbreviation', value.name = 'Inter-species') %>% 
+              mutate(id = row.names(.)) %>% 
+              melt(id.vars = "id", 
+                   variable.name = 'Abbreviation', value.name = 'Inter-species') %>% 
+              select(-id) %>% 
+              mutate(Abbreviation = as.character(Abbreviation)) %>% 
               group_by(Abbreviation) %>%
-              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T)))) %>%
+              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))),
+            by = "Abbreviation") %>%
   mutate(`Intra-individuals` = ifelse(Abbreviation == "WD", NA, `Intra-individuals`)) %>% 
@@
 The rlast columns represents the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species.')
-
 ```
@@
   group_by(Tree, Sp, SpCode, CE) %>% 
-  summarise_each(funs(mean(., na.rm = T))) %>% 
+  summarise_all(funs(mean(., na.rm = T))) %>% 
   ungroup() %>% 
+  select(-Tree, -CE) %>% 
   group_by(Sp, SpCode) %>% 
-  summarise_each(funs(mean(., na.rm = T))) %>% 
+  summarise_all(funs(mean(., na.rm = T))) %>% 
   ungroup() %>% 
-  select(-Tree, -Sp, -CE, -LTD, -SHA)
-row.names(PFT) <- PFT$SpCode
+  select(-Sp, -LTD, -SHA)
+suppressWarnings(row.names(PFT) <- PFT$SpCode) # deprecated
 # PCA
@@
 Q <- PFT %>% 
-  left_join(data.frame(PFT.pca$scores) %>% mutate(SpCode = row.names(.))) %>% 
+  mutate(SpCode = as.character(SpCode)) %>% 
+  left_join(data.frame(PFT.pca$scores) %>% 
+              mutate(SpCode = row.names(.)),
+            by = "SpCode") %>% 
   rename(WES = Comp.1, LES = Comp.2) %>% 
-  left_join(genSpecies()) %>% 
+  left_join(genSpecies(), by = "SpCode") %>% 
   mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
@@
   summarise(n = n()) %>% 
-  dcast(com ~ SpCode) %>% 
+  dcast(com ~ SpCode, value.var = "n") %>% 
   replace(is.na(.), 0)
@@
        value.name = "abundance") %>% 
-  dcast(SpCode ~ com) %>% 
-  inner_join(Q) %>%
+  dcast(SpCode ~ com, value.var = "abundance") %>% 
+  mutate(SpCode = as.character(SpCode)) %>% 
+  inner_join(Q, by = "SpCode") %>%
   select(-SpCode) %>%
@@
        value.name = "abundance") %>% 
-  dcast(SpCode ~ com) %>% 
-  left_join(Q) %>% 
+  dcast(SpCode ~ com, value.var = "abundance") %>% 
+  mutate(SpCode = as.character(SpCode)) %>% 
+  left_join(Q, by = "SpCode") %>% 
   select(-SpCode) %>% 
@@
   rename(SES = value) %>% 
-  left_join(R) %>%
+  mutate(community = as.character(community)) %>% 
+  left_join(R, by = "community") %>%
   group_by(strata, trait) %>% 
@@
   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
-  mutate_each(funs(replace(., is.na(.), ' '))) %>% 
+  mutate_all(funs(replace(., is.na(.), ' '))) %>% 
   left_join( # Inclusion of R2 but should be split
@@
       rename(SES = value) %>% 
-      left_join(R) %>%
+      mutate(communtiy = as.character(community)) %>% 
+      left_join(R, by = "community") %>%
       group_by(strata, trait) %>% 
@@
       mutate(r.squared = round(r.squared, 2)) %>% 
-      rename("$R^2$" = r.squared)
+      rename("$R^2$" = r.squared), by = c("strata", "trait")
   )%>% # End of R2

### 2018-07-09 R/genPFT.R

@@ -22,7 +22,7 @@ genPFT <- function(){
   PFT$CE <- as.factor(tolower(PFT$CE))
   # Variables modification
   PFT$CEs <- as.numeric(as.character(substr(PFT$CE, 1, 1)))
-  PFT$LTD <- as.numeric(PFT$LTD)
+  PFT$LTD <- suppressWarnings(as.numeric(PFT$LTD)) # Missing values
   PFT$LDMC = 1000 * PFT$LDMC # from g.g-1 to mg.g-1
   PFT$SLA = PFT$SLA / 1000 # from mm2.g-1 to m2.kg-1
   PFT$HA = PFT$HA / PFT$LA

### 2018-07-10 draft.Rmd
Uppangala 0.1.9048 M&M + Results + Supplementary Materials (updated ToDo)

@@
 title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
-author: "Sylvain Schmitt, Maxime Réjou-Méchain, Valérie Raevel, Ayyappan N, Balachandran N., Barathan N., and François Munoz"
+author: "Sylvain Schmitt, Valérie Raevel, Maxime Réjou-Méchain, Ayyappan N, Balachandran N., Barathan N., and François Munoz"
 date: '`r Sys.Date()`'
 output:
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
-  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     toc: false
+  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
 linestretch: 1.5
@@
 ---
+
+<!-- setup -->
 
@@
 
-# Traits
+<!-- data -->
+
+```{r Q}
+# Sp level PFT
+PFT <- genPFT() %>% 
+  mutate(LA = log(LA)) %>% 
+  mutate(SLA = log(SLA)) %>% 
+  group_by(Tree, Sp, SpCode, CE) %>% 
+  summarise_all(funs(mean(., na.rm = T))) %>% 
+  ungroup() %>% 
+  select(-Tree, -CE) %>% 
+  group_by(Sp, SpCode) %>% 
+  summarise_all(funs(mean(., na.rm = T))) %>% 
+  ungroup() %>% 
+  select(-Sp, -LTD, -SHA)
+suppressWarnings(row.names(PFT) <- PFT$SpCode) # deprecated
+# PCA
+PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
+# Matrix Q
+Q <- PFT %>% 
+  mutate(SpCode = as.character(SpCode)) %>% 
+  left_join(data.frame(PFT.pca$scores) %>% 
+              mutate(SpCode = row.names(.)),
+            by = "SpCode") %>% 
+  rename(WES = Comp.1, LES = Comp.2) %>% 
+  left_join(genSpecies(), by = "SpCode") %>% 
+  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
+  select(SLA, LDMC, WD, LA, LT, LES, WES, strata, SpCode) %>%
+  mutate(strata = ifelse(SpCode == "agin", "U", strata))
+```
+
+```{r L}
+com <- quadrats(cs = c(30,30))
+XY <- genTrees()[c('x', 'y')]
+coordinates(XY) <- ~ x + y
+proj4string(XY) <- crs(com)
+L <- genTrees() %>% 
+  mutate(com = (XY %over% com)[,1]) %>% 
+  filter(!is.na(com)) %>% 
+  group_by(com, SpCode) %>% 
+  summarise(n = n()) %>% 
+  dcast(com ~ SpCode, value.var = "n") %>% 
+  replace(is.na(.), 0)
+```
+
+```{r R}
+# Env <- genEnv()
+# com <- quadrats(cs = c(30,30))
+# com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+# Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/share/applications')
+# names(Deriv)[5] <- 'SW'
+# com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
+# com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+# com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+# BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
+# com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
+# pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
+# save(pca.env, file = file.path(path, 'pca.env.Rdata'))
+# R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+# row.names(R) <- com@data$id
+# R$community <- row.names(R)
+# save(R, file = file.path(path, 'R.Rdata'))
+load(file.path("draft_save", 'R.Rdata'))
+```
+
+```{r GWMses}
+GWM <- L %>%
+  melt(id.var = "com",
+       variable.name = "SpCode",
+       value.name = "abundance") %>% 
+  dcast(SpCode ~ com, value.var = "abundance") %>% 
+  mutate(SpCode = as.character(SpCode)) %>% 
+  left_join(Q, by = "SpCode") %>% 
+  select(-SpCode) %>% 
+  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+       variable.name = 'community', value.name = 'abundance') %>% 
+  group_by(strata, community) %>% 
+  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# nQ <- array(rep(as.matrix(Q), n),
+#             c(dim(Q), n),
+#             list(row.names(Q), names(Q), 1:n))
+# nQ <- array(apply(nQ, 3, function(x){
+#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
+#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
+#   return(x)
+#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
+# cl <- makeCluster(cores)
+# clusterExport(cl, list('L', 'nQ'))
+# nGWM <- parApply(cl, nQ, 3, function(x){
+#   library(dplyr)
+#   library(reshape2)
+#   L %>%
+#     melt(id.var = "com",
+#          variable.name = "SpCode",
+#          value.name = "abundance") %>%
+#     dcast(SpCode ~ com) %>%
+#     left_join(data.frame(x)) %>%
+#     select(-SpCode) %>%
+#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+#          variable.name = 'community', value.name = 'abundance') %>%
+#     group_by(strata, community) %>%
+#     mutate_each(funs(as.character), -abundance) %>%
+#     mutate_each(funs(as.numeric), -abundance) %>%
+#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
+# })
+# stopCluster(cl)
+# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
+#               c(dim(GWM[3:9]), n),
+#               list(row.names(GWM), names(GWM[3:9]), 1:n))
+# rm(cl, nQ)
+# save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
+load(file.path("draft_save", 'nGWM.Rdata'))
+GWM_ses <- GWM
+GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
+```
+
+# Abstract
+
+## Keywords
+
+functional trait; rainforest; vertical forest strata; Western Ghats; environmental filtering; niche differentiation; ecological guild
+
+# Introduction
+
+# Material and methods
+
+## Study site
+
+The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry in an undisturbed, old-growth wet evergreen Dipterocarp forest [@pascal_structure_1996].
+The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), near the Uppangala village, and thus refers to as "Uppangala plot" (see figure \@ref(fig:Uppangala)).
+Annual precipitation averages 5108 mm per year but is highly variable in time with heavy rainfall from June to October (90% of the rainfall during the monsoon) and a severe dry season (monthly rainfall $\leq 45$ mm) lasting 3.8 months [@pelissier_twenty_2011].
+Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996].
+Climate and geology of the 10-ha plot can be considered as homogeneous.
+The topography of the site is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes, so that topography is expected to be a major component of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998].
+Therefore, we hypothesized that topographic variation should influence community assembly and yield changing ecological strategies and functional composition [@condit_spatial_2000].
+
+```{r Uppangala, fig.cap="Canopy model of the 10-ha forest permanent plot in Uppangala, India. Preliminry figure to be improved."}
+knitr::include_graphics("./data/test.png")
+```
+
+## Taxonomic composition and trait sampling
+
+From 1989 to 2014, the trees $\geq 10$ cm in diameter at breast height (dbh) were identified to species level, tagged, mapped and monitored in an area progressively extended to a contiguous $300 * 330 m^2$ (ca.
+10-ha) plot [@pelissier_twenty_2011].
+The plot included 111 different tree species.
+We collected trait data from 89 over 111 (80.1%) species, which represented 99.0% of tree individuals present in the plot.
+Note that the 32 most abundant species represented more than 80% of tree individuals.
+We sampled five individuals per species for these 32 species, and sampled one to five individuals for the remaining species (in total 267 individuals).
+For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation of @perez-harguindeguy_new_2013.
+Individuals and leaves were selected in consistent light conditions, development stage and sanitary state.
+Dawkins crown exposure index [@dawkins_management_1958] was recorded for each tree sampled to control for any sun-exposure effect.
+Leaf and branch samples were kept in humidified ziplock filled with enriched $CO_{2}$ air inside a black bag until measurement within the next 24 hours.
+For some individuals, we were unable to collect suitable branch sample, i.e.
+with a diameter $\leq 1$ cm (11 missing species over 89 representing 1% of individuals present).
+
+## Trait measurements
+
+Leaf petioles, rachis and petiolules were removed for all measurements.
+Leaf fresh weights, thickness and area were measured, using a weighing scale, a micrometer and a bend scanner with ImageJ software [@Schneider2012] respectively.
+Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$.
+Once dried, the leaves were weighted directly after been extracted from the oven to avoid any humidity bias.
+The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues.
+Branch wood samples were immersed in water to evaluate their volume by water displacement.
+Wood samples were then dried for 48 hours at 80 $^\circ C$ and weighted.
+
+We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table \@ref(tab:Traits)).
+We chose those five traits because they proved to be good proxies of the major dimensions of ecological strategies in plants [@mark_westoby_plant_2002; @diaz_global_2015] and were relatively easy to measure on the field.
+
+## Environmental variables
+
+Topography was expected to be a major environmental factor influencing the assemby of tree communities [@gimaret-carpentier_sampling_1998].
+We selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspectindegres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed in ArcGis [@desktop_arcgis_2011].
+
+In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the abiotic environmental variables (see Supplementary materials [S1: Principal component analysis of topographical variables]).
+We selected four little-correlated variables: curvature representing local water flow paths and soil accumulation areas, wetness highlighting water accumulation areas over the whole study area, southwesterness representing aspect [@bunyan_effect_2015] and slope, which may have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher].
+
+In addition, we selected two biotic factors related to local forest structure: canopy height and basal area.
+Canopy height, i.e.
+the maximum local height of the canopy, was calculated from a 1-m resolution digital canopy model built from a LiDAR (Light Detection And Ranging) dataset acquired in 2013 over the Uppangala plot.
+Basal area was calculated from the girth measurements taken within the 10-ha plot (censuses done in 2013 and 2014) for stems $\geq 10$ cm dbh.
+We kept only canopy height because the two variables were highly correlated and canopy height present the advantage to be independant from plot censuses.
+
+## Analyses
+
+### Guilds
+
+@pascal_wet_1988 defined four vertical strata attained by adult trees: emergent above 39m, canopy trees between 25 and 39m, sub-canopy trees between 15 and 25m, and understorey trees below 15m.
+Despite the vertical hierarchy, steep topography could allow sub-canopy trees receiving more light, especially in slopes, and thus participates to canopy.
+We thus defined two basic guilds called canopy (C), composed of emergent, canopy and sub-canopy species (n = `r sum(Q$strata == "C")`), and understorey (U) species (n = `r sum(Q$strata == "U")`).
+
+### Traits variation and covariation
+
+We first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among leaves of the same individual (intra-individual), among trees of the same species (intra-specific for species with more than one individual), and among species (inter-specific).
+We then performed a principal component analysis (PCA) of the mean value of traits per species to identify main dimensions of ecological strategies along which traits can covary among guilds.
+In addition we explored guilds functional organisation around the strategy axes revealed by the PCA.
+We compared guilds functional variance along the two first axis from the PCA with Levene's tests [@Schultz1985].
+Subsequent community-level analyses were based on trait values averaged per species.
+
+### Drivers of community dynamics
+
+We subdivided the 10-ha plot into $30*30 m^2$ quadrats leading to 110 local communities.
+We analyzed the functionnal composition of the guilds in the 110 communities.
+Guilds in the communities have a mean number of 28 individuals representing 10 species, with less in understorey than in canopy.
+
+We calculated weighted mean [GWM, where G stands for "guild", @violle_let_2007] for each guild in each community.
+In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of species $i$ weighted by species relative abundance $p_{i}$:
+$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~~ (1)$$ 
+where $S$ is the number of species in the quadrat $j$.
+
+We devised a null model approach to test for any significant changes in functional composition in response to an environmental gradient.
+We thus wanted to explore if the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds.
+We hypothesized that *drivers of environmental filtering on local functional composition differ between canopy and understorey tree species*.
+To accept or reject the hypothesis we used a null model randomizing trait values among species in the whole plot for each guild (permuting rows in Trait*Species matrices of both guilds) [@bernard-verdier_community_2012].
+Consequently, this nul model breaks the link between environement and local functional composition while keeping guilds and species spatial structure, thus testing for our hypothesis.
+
+We performed 999 randomizations of each null model and we derived a two-tailed test of the proportion of guilds deviating from the null expectation, with adjusted p-values by Benjamini-Hochberg correction [@Benjamini1995].
+
+To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as 
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma {null}} ~~ (3)$$ 
+where $I{obs}$ was the observed metric and $I_{null}$ and $\sigma {null}$ were the mean and standard deviation of the corresponding null distribution.
+The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES{trait} = Slope + Curvature + Wetness + SW + Canopy$).
+Analyses were conducted on the five functional traits.
+
+All analyses were conducted using the R 3.3.0 software [@r_core_team_r:_2016] with packages ade4 [@dray_ade4_2007], and vegan [@oksanen_vegan:_2016].
+
+# Results
+
+## Traits variation and covariation
+
+We found an overall increase of functional trait variation from intra-individual to inter-specific levels (see table \@ref(tab:Traits)), even though LA showed almost as large variability at intra-individual than at intra-specific levels ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$).
+Given the relatively low intra-individual and intra species variation, compared to inter-species variation, using species trait average values for community analyses is warranted.
 
@@
 
-```{r Q}
-# Sp level PFT
-PFT <- genPFT() %>% 
-  mutate(LA = log(LA)) %>% 
-  mutate(SLA = log(SLA)) %>% 
-  group_by(Tree, Sp, SpCode, CE) %>% 
-  summarise_all(funs(mean(., na.rm = T))) %>% 
-  ungroup() %>% 
-  select(-Tree, -CE) %>% 
-  group_by(Sp, SpCode) %>% 
-  summarise_all(funs(mean(., na.rm = T))) %>% 
-  ungroup() %>% 
-  select(-Sp, -LTD, -SHA)
-suppressWarnings(row.names(PFT) <- PFT$SpCode) # deprecated
-# PCA
-PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
-# Matrix Q
-Q <- PFT %>% 
-  mutate(SpCode = as.character(SpCode)) %>% 
-  left_join(data.frame(PFT.pca$scores) %>% 
-              mutate(SpCode = row.names(.)),
-            by = "SpCode") %>% 
-  rename(WES = Comp.1, LES = Comp.2) %>% 
-  left_join(genSpecies(), by = "SpCode") %>% 
-  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
-  select(SLA, LDMC, WD, LA, LT, LES, WES, strata, SpCode) %>%
-  mutate(strata = ifelse(SpCode == "agin", "U", strata))
-```
-
-```{r L}
-com <- quadrats(cs = c(30,30))
-XY <- genTrees()[c('x', 'y')]
-coordinates(XY) <- ~ x + y
-proj4string(XY) <- crs(com)
-L <- genTrees() %>% 
-  mutate(com = (XY %over% com)[,1]) %>% 
-  filter(!is.na(com)) %>% 
-  group_by(com, SpCode) %>% 
-  summarise(n = n()) %>% 
-  dcast(com ~ SpCode, value.var = "n") %>% 
-  replace(is.na(.), 0)
-```
-
-```{r R}
-# Env <- genEnv()
-# com <- quadrats(cs = c(30,30))
-# com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-# Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/share/applications')
-# names(Deriv)[5] <- 'SW'
-# com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
-# com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-# com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
-# BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
-# com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
-# pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
-# save(pca.env, file = file.path(path, 'pca.env.Rdata'))
-# R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
-# row.names(R) <- com@data$id
-# R$community <- row.names(R)
-# save(R, file = file.path(path, 'R.Rdata'))
-load(file.path("draft_save", 'R.Rdata'))
-```
-
-# PCA
-
 ```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values.
@@
   coord_equal() +
-  geom_hline(aes(yintercept = 0), col = 'lightgrey') +
-  geom_vline(aes(xintercept = 0), col = 'lightgrey') +
+  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
+  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
   xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
@@
   facet_wrap(~strata, nrow = 2) +
-  coord_flip()
+  coord_flip() +
+  scale_fill_manual("PCA axis", 
+                    labels = c("WES" = "PCA1", "LES" = "PCA2"),
+                    values = c("green", "purple")) +
+  scale_color_manual("PCA axis", 
+                     labels = c("WES" = "PCA1", "LES" = "PCA2"),
+                     values = c("green", "purple"))
 cowplot::plot_grid(pca.plot, spectrum.density, rel_widths = c(2,1))
@@
 
-Canopy species vary significantly more on the wood than the leaf economic spectrum (`r Variance$LES$C` vs `r Variance$WES$C`, Levene's test p-value=`r Leven_test$C`).
-In parrallel, understorey species show a tendency to vary more on the leaf than the wood economic spectrum (`r Variance$LES$U` vs `r Variance$WES$U`, Levene's test p-value=`r Leven_test$U`).
+The first axis of the principal component analysis (PCA, see figure \@ref(fig:PCA)) of mean species trait values was basically related to wood density and structural leaf trait variations (LT, LA, LDMC), thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood.
+The second axis mostly opposed SLA to LDMC, and to a less extent to wood density.
+However, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between those two traits.
+In addition, canopy species vary significantly more on the PCA first axis than on the second axis (`r Variance$LES$C` vs `r Variance$WES$C`, Levene's test p-value=`r Leven_test$C`).
+In parrallel, understorey species show a tendency to vary more on the PCA second axis than the first axis (`r Variance$LES$U` vs `r Variance$WES$U`, Levene's test p-value=`r Leven_test$U`).
 
-# Guilds
+## Functional composition of guilds
+
+Guilds weighted means (GWM) from understorey and canopy guilds strongly diverged among communities (see figure \@ref(fig:GWMboxplot) ).
+Canopy guilds have higher SLA, WD and LA but lower LDMC and LT than understorey guilds.
+In other words, canopy guilds have larger, thinner and lighter leaves with increased SLA and denser woods, whereas understorey guilds have smaller, thicker, and more dense leaves with lower SLA and lighter woods.
+
+The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds ($R^2$ from 0.04 to 0.5, see table \@ref(tab:LM)).
+Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits, and two major differences were observed between the two guilds.
+The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height and wetness affected mainly understorey guild composition.
+
+Focusing on canopy tree species, we found an increase in wood density under south-west orientation.
+We also observed a decrease in leaf area under the south-west orientation.
+Leaf area decreased with curvature, while wood density and LDMC increased with curvature.
+Finally, heigh canopy was also related to an increased leaf area.
+
+For understorey tree species, SLA significantly decreased and leaf thickness significantly increased with curvature.
+SLA significantly increased with canopy height, while leaf thickness decreased with canopy height.
+Finally, SLA and wood density significantly decreased and leaf area and thickness significantly increased with wetness.
 
@@
   ggpubr::stat_compare_means(aes(label = paste(..p.signif..))) 
-```
-
-# LM
-
-```{r GWMses}
-GWM <- L %>%
-  melt(id.var = "com",
-       variable.name = "SpCode",
-       value.name = "abundance") %>% 
-  dcast(SpCode ~ com, value.var = "abundance") %>% 
-  mutate(SpCode = as.character(SpCode)) %>% 
-  left_join(Q, by = "SpCode") %>% 
-  select(-SpCode) %>% 
-  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-       variable.name = 'community', value.name = 'abundance') %>% 
-  group_by(strata, community) %>% 
-  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
-# nQ <- array(rep(as.matrix(Q), n),
-#             c(dim(Q), n),
-#             list(row.names(Q), names(Q), 1:n))
-# nQ <- array(apply(nQ, 3, function(x){
-#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
-#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
-#   return(x)
-#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
-# cl <- makeCluster(cores)
-# clusterExport(cl, list('L', 'nQ'))
-# nGWM <- parApply(cl, nQ, 3, function(x){
-#   library(dplyr)
-#   library(reshape2)
-#   L %>%
-#     melt(id.var = "com",
-#          variable.name = "SpCode",
-#          value.name = "abundance") %>%
-#     dcast(SpCode ~ com) %>%
-#     left_join(data.frame(x)) %>%
-#     select(-SpCode) %>%
-#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-#          variable.name = 'community', value.name = 'abundance') %>%
-#     group_by(strata, community) %>%
-#     mutate_each(funs(as.character), -abundance) %>%
-#     mutate_each(funs(as.numeric), -abundance) %>%
-#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
-# })
-# stopCluster(cl)
-# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
-#               c(dim(GWM[3:9]), n),
-#               list(row.names(GWM), names(GWM[3:9]), 1:n))
-# rm(cl, nQ)
-# save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
-load(file.path("draft_save", 'nGWM.Rdata'))
-GWM_ses <- GWM
-GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
 ```
@@
 
-# Environment
+# Discussion
+
+# Supplementary materials
+
+## S1: Principal component analysis of topographical variables
+
+We selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspectindegres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed in ArcGis [@desktop_arcgis_2011].
+
+In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA).
 
@@
 
-# Why not hypervolumes
+Plan curvature and profile curvature were obviously highly correlated to curvature (see figure \@ref(fig:PCAenv)).
+We selected curvature to represent local water flow paths and soil accumulation areas.
+Wetness was negatively correlated to elevation, and was kept to represent water accumulation areas over the whole study area.
+Slope was kept apart, besides a correlation to elevation, for its putative effect on forest dynamics and on nutrients/water availability.
+Finally, southwesterness was not in first plan 1-2 from PCA, and was thus kept to represent aspect.
+
+## S2: Why not hypervolumes
+
+**Reminder to be further removed !**
 
@@
 
+We decided to use Levene's test over hypervolume overlap test for several reasons listes below (the chunk included in the Rmarkdown can be used for better vizualisation of the arguments):
+
 * Outliers make quicly overlapping hypervolumes besides the density stay low in those area
 * Hypervolumes consider uncorrelated axis, which is not true for functional traits, and blur strategy axes highlighted by economic spectra
+
+# References

### 2018-07-10 draft.Rmd
Uppangala 0.1.9048 10.07 new version (Intro + Discussion)

@@
 output:
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::word_document2: default
   bookdown::html_document2:
@@
 csl: bibliography/mee.csl
-bibliography: bibliography/refs.bib
+bibliography: ["bibliography/refs.bib", "bibliography/library.bib"] 
 link-citations: yes
@@
 
+<!-- hook -->
+Understanding the determinants of species occurrence and co-occurrence is a long standing issue in ecology [@diamond_ecology_1975].
+Tropical rainforests have especially fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
+<!-- spectrums -->
+The variation in functional traits among species is expected to reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007] and expresses fundamental trade-offs along environmental gradients [@mark_westoby_plant_2002].
+For instance, specific leaf area varies along the soil fertility-shade gradient and represents a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@hodgson_is_2011].
+Fundamental trade-offs entail functional traits covariation along basic leaf [@wright_worldwide_2004 but see @osnas_global_2013 and @lloyd_photosynthetically_2013] and wood [@chave_towards_2009] economics spectra.
+The leaf economics spectrum opposes acquisitive to conservative ecological strategies, basically expressed at leaf level.
+The wood economics spectrum distinguishes species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength [@chave_towards_2009].
+@baraloto_decoupled_2010 found decoupled wood and leaf economics of trees in neotropical forest, but @reich_world-wide_2014 conversely posited that both spectra have to be integrated in a comprehensive whole plant economics spectrum [@freschet_evidence_2010] opposing fast growing acquisitive species to slow growing and more conservative ones.
+
+<!-- Spatial to ecological scale (trans diff guilds) -->
+Nevertheless, the nature of functional traits and ecological strategies driving species coexistence can change across scales.
+For instance, @messier_trait_2016 have underlined that the leaf economics spectrum basically expresses changes in ecological strategies along broad-scale environmental gradients, while it is not necessarily a primary driver of local coexistence [but see @Paine2011].
+Scales can be geographic or ecologic, such as guilds.
+<!-- Guild (diff in resp to env, Pacala & Duque) -->
+The concept of guilds refer to groups of functionally similar species [@wilson_guilds_1999].
+Ecological guilds can coexist and occupy distinct niches within communities [@vergnon_niches_2009].
+In tropical forests, the assembly of tree communities has been mostly studied irrespectivelly of guilds of canopy and understorey tree [@wright2002plant].
+However, distinctive ecological constraints are expected to drive different resource use strategies and biotic interactions in these guilds, which should translate into different functional composition and response to environmental factors.
+For instance, in Colombian Amazonia, @duque2002different found some contrasted response of the composition of canopy and understorey guilds along environmental gradients.
+Consequently, guilds should bear differences of previously underlined functional strategies, so-called economics spectra.
+
+<!-- What we've done -->
+In this paper, we used a trait-based approach to study species strategies between canopy and understorey guilds.
+More specifically, we investigated separately the assembly of canopy and understorey tree species along topographical and biotic gradients in a large forest plot (10ha) located in Western Ghats biodiversity hotspot [@myers_biodiversity_2000], India.
+Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988].
+It thus provides an appropriate context to adress guild-dependent assembly rules and ecological strategies in tropical rainforests.
+Although considered as wet evergreen, the forest undergoes contrasted long dry and intense rainy seasons under the influence of monsoon.
+In this context, we expected water stress and competition for resource to be important drivers of tree functional composition within guilds.
+The trait values of coexisting species within the guilds represent a non-random sample of trait values found in a broader area, du to specific niche-based dynamics in the local community [@lortie_rethinking_2004; @bernard-verdier_community_2012].
+Thus we used the signature of local niche-based dynamics to assess the deviation of local community structure from a larger "species pool", using appopriate null model [@kraft_functional_2008], to study how both canopy and understorey tree species composition varied with local environmental conditions.
+<!-- Questions -->
+We specifically addressed the following questions:
+
+* Does functional traits variation and covariation reflect basic dimensions of ecological strategy at local scale and how are organized canopy and understorey guilds along traits covariation ?
+* Does local functional composition from guilds differ ?
+* Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?
+
 # Material and methods
@@
 To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as 
-$$SES = \frac{(I_{obs} - I_{null})}{\sigma {null}} ~~ (3)$$ 
-where $I{obs}$ was the observed metric and $I_{null}$ and $\sigma {null}$ were the mean and standard deviation of the corresponding null distribution.
+$$SES = \frac{(I_{obs} - I_{null})}{\sigma_{null}} ~~ (3)$$ 
+where $I{obs}$ was the observed metric and $I_{null}$ and $\sigma_{null}$ were the mean and standard deviation of the corresponding null distribution.
 The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES{trait} = Slope + Curvature + Wetness + SW + Canopy$).
@@
 Canopy guilds have higher SLA, WD and LA but lower LDMC and LT than understorey guilds.
-In other words, canopy guilds have larger, thinner and lighter leaves with increased SLA and denser woods, whereas understorey guilds have smaller, thicker, and more dense leaves with lower SLA and lighter woods.
+In other words, canopy guilds have larger, thinner and lighter leaves with increased SLA and denser woods, whereas understorey guilds have smaller, thicker, and denser leaves with lower SLA and lighter woods.
 
@@
 
+<!-- take home messages -->
+We found back economics spectra evidenced by @wright_worldwide_2004 and @chave_towards_2009 at finer scale than @messier_trait_2016 suggested (figure \@ref(fig:PCA)).
+This result advocate for a pantropical and multi-spatial scale oganisation of tropical plant strategies in two separated economics spectra.
+Still our results suggest that this two spectra are decoupled between canopy and understorey tree species, whereas existing in both guilds.
+Canopy tree species being the first layer of forest ecosystem interacting with external abiotic environment shows major variation in wood economics spectrum (table \@ref(tab:LM)).
+Indeed wood economic spectrum is related to mechanical resistance, disturbance tolerance and water conductivity efficiency.
+Thus understorey tree will benefit from a mircoclimate [@Porte2004] generated by the presence of the canopy tree.
+In this microclimate, leaf economics spectrum is then a major axis of strategy to answer to a changing light environment and soil fertility (table \@ref(tab:LM)).
+
+## Wood and leaf economics hold at the local scale
+
+<!-- WIP here -->
+
+We found an increasing trait variation within individuals, within species, and between species (table \@ref(tab:Traits)).
+Our analyses thus reveal that the primary source of trait variation within and among communities is the inter-species variation, justifying the use of species mean traits for the study of tropical tree guild assembly, besides protocol tend to minimise intra-specific variability.
+The PCA of functional trait variation among species displayed two main dimensions (figure \@ref(fig:PCA)), very similar to the ones found by @baraloto_decoupled_2010 and @fortunel_leaf_2012.
+Those authors interpreted these axes as independent wood and leaf economics spectra.
+We found that SLA and LDMC tend to covary negatively along the leaf economics spectrum, which be seen as a trade-off between productivity, with less carbon investment and higher photosynthesis rates, and persistence, with more leaf defense and investment [@wright_worldwide_2004].
+A second major axis of trait variation opposed wood density to leaf area and thickness.
+It may be interpreted as an axis opposing stress tolerant species, with denser wood, to competitive species, with thicker and larger leaves, conferring a higher efficiency in water conductivity [@Santiago2004a].
+It may thus be interpreted as the wood economics spectrum firstly introduced by @chave_towards_2009, and then discussed by @baraloto_decoupled_2010.
+Interestingly, @baraloto_decoupled_2010 found an independence between the leaf and wood economics spectra and their wood economics spectrum also showed an opposition between hardwood species having small thin leaves and light wood species having large thick leaves.
+
+Furthermore, our results are similar to the ones found by @fortunel_leaf_2012, where basic PCA axes also represented leaf and wood economics spectra (inverted compare to us).
+Noticeably, our analysis integrated less traits (5 against 14), less species (89 against 758), and considered a set of species found in a local forest-plot while @fortunel_leaf_2012 considered a regional set of species from French Guiana and Peru (i.e.
+the opposite ends of Amazonia).
+Therefore, our results shows that the economics spectra found at a regional scale in the Neotropics are also valid at a far more local scale in an Indian rainforest.
+This result does support the arguments by @messier_trait_2016 and suggests that the two observed spectra underline major ecological strategies in tropical forests, even locally.
+
+## Canopy and understorey species assembly rely on different economics spectra
+
+We further analyzed guilds distribution along economics spectra evidenced by the principal component analysis (figure \@ref(fig:PCA)).
+We found a weak but significant difference of variation of the canopy species along the wood economics spectrum compared to the leaf economics spectrum, whereas the understorey only showed an unsignificant trend of variation along the leaf more than the wood economics spectrum.
+In addition, we analyzed functional composition of $30*30m^2$ guilds from canopy and understorey, as well as the variation of this composition with environment across a 10-ha forest plot.
+
+Distribution of functional composition among communities shows a strong differentiation between the two guilds for each trait (see figure \@ref(fig:GWMboxplot)).
+Canopy guilds have larger, thinner and lighter leaves with increased SLA and denser woods, whereas understorey guilds have smaller, thicker, and denser leaves with lower SLA and lighter woods.
+
+In addition, we found that the variations in trait composition along the environmental gradients differ between canopy and understorey species, with partially more variation in traits associated to the leaf economic spectrum (LES) for understorey species and to the wood economics scpectrum (WES) for canopy species.
+In the former case, the curvature, wetness index, and the canopy height had a significant effect on the LES-based composition of understorey tree species (but not only), whereas southwest orientation and curvature both had a significant effect on the WES-based composition of canopy tree species (table \@ref(tab:Traits)).
+
+Those results advocates for decoupled strategies between species of understorey and canopy guilds [@duque2002different].
+Indeed canopy tree species could deploy strategies to endorse the role of first layer against external abiotic environment.
+Those strategies will come through wood economic spectrum (WES) to ensure mechanical resistance, disturbance tolerance and water conductivity efficiency.
+Effectivelly, @Poorter2010b found hydraulic traits to be more related to adult stature than light demand.
+They hypothesized that this pattern was related to more drought stress and higher cavitation risk for canopy guild, thus supporting our result.
+On the other hand, understorey tree species will encounter constraints directly linked to canopy guild limiting light availability.
+Thus they will need to deploy strategies linked to ressource acquisition coming through leaf economics spectrum.
+@duque2002different also highlighted increased correlation of understorey species to soil characteristics compared to canopy species in well-drained uplands.
+Furthemore, @Santiago2007a, studying understorey plants compared to trees and lianas, found that understorey leaf-allocation patterns optimize SLA, and hence growth.
+Besides @duque2002different results were significant only for well-drained uplands and @Santiago2007a results concerns plant and not tree understorey, those two studies reinforce the idea that understorey species deploy strategy linked to resources acquisition.
+
+Hence, those results participate to the debate of decoupled economics spectra [@baraloto_decoupled_2010] versus the whole plant economics spectrum [@reich_world-wide_2014].
+Effectivelly, decoupled spectra could be linked to divergent strategy of canopy and understorey species in response to their abiotic and biotic environment.
+And @freschet_evidence_2010 might have lacked to evidence decoupled economic spectra due to an absence of vertical structure in the subartic flora.
+Still, both guilds presented significant effect of the environment on the second spectrum (LES for canopy and WES for understorey), highlighting a need to further investigate this pattern.
+We further tested guilds organization around economics spectra on Amazonian tree species with BRIDGE dataset [@baraloto_decoupled_2010, see supplementary materials X].
+**We found ...**
+
+## Environment influence guild-level functionnal composition
+
+We should note that relations found between guilds weighted means (GWM) and environment are globally weak (table \@ref(tab:LM), most of $R^2 \leq 0.2$ besides SLA model in canopy have an $R^2=0.5$).
+This low explanatory power is advocating for underlying null process related to neutral theory [@hubbell_unified_2001].
+Null process can effectivelly be responsible for the majority of observed variance in functionnal composition at a such fine scale.
+Studies highlighting environmental filtering effect often present a higher environmental variation than in our study site [@bernard-verdier_community_2012].
+In parallel, the 10-ha plot scale is really low and partially responsible for the small environmental variation.
+@kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25-ha tropical forest plots including large flat valleys.
+<!-- NB: We effectivelly discussed with Bruno that Uppangala was really small to be able to derive ecological interpretation (lack of repliaction and statistical power) -->
+
+Curvature is affecting the variation of leaf economics in understorey species and the variation of wood economics in canopy tree species.
+A locally convex context would accumulate more soil and attract more water, increasing soil fertility and allowing persistence of species investing more in leaf resource acquisition in understorey with higher SLA and lower LT, or less in wood ressources in canopy with lighter wood and bigger leaves.
+Curvature effect for understorey species is doubled by the effect of wetness index, expressing also the water accumulation in convex context resulting in increased soil fertility [@ferry2010higher].
+Wet soil will result in lower SLA, and bigger and thicker leaves for understorey species.
+Economics spectra are not perfectly divided between canopy and understorey, with wood and leaf respectivelly, for instance leaf area and thickness of understorey species also responded to wet soils in understorey.
+
+The influence of aspect on forest composition is little documented in tropical forests.
+In a similar context that our study site, @gunatilleke_specieshabitat_2006 gave some intuitive insights on the effect of aspect on forest composition: they suggested that aspect through monsoon had a potential effect on rainforests.
+@bunyan_effect_2015 tested the effect of aspect on the occurence of forest fragments in a forest-grassland mosaic and foud that southwest orientation strongly determined the presence of forest in mountain rainforests of Western Ghats.
+They suggested that it could result from monsoon winds.
+In our context, monsoon winds are south-west oriented, and are expected to be a major source of disturbance in forests [@soman_aspects_1990].
+
+The Uppangala plot is situated in a topographically complex area with steep slopes mostly organized along southwest-oriented crests.
+In this context it is diffcult to predict wind flow and orientation inside the 10-ha plot topography.
+We propose two explanation, hardwood associated to smaller leaves could be selected for mechanical resistance in more disturbed souwhtwest oriented area or topography provoke a turbulences on northeast oriented areas responsible of more disturbance and death events in the forest history resulting in lightwood filtering [@larjavaara_rethinking_2010].
+In both case it seems reasonable that canopy tree species are the most influenced by disturbance events linked to monsoon, and may even buffer the effect on understorey species.
+
+Biotic factors, illustrated here by the canopy height, also had a significant effect on the LES-based composition of understorey tree species.
+We indeed found higher SLA values and lower LT values in understorey assemblages that are located under stands characterized by a tall canopy.
+This results contrasts with the expectation that understorey species invest in higher LDMC for higher leaf lifespan, at the expense of SLA, under light-limited conditions [@kitajima_tissue-level_2010].
+The Uppangala forest is characterized by a large heterogeneity of canopy height allowing light to penetrate more in lower strata, maybe especially more in high canopy context covering less closely understorey, resulting in those observation on the leaf economics variation of understorey species.
+Canopy height also resulted in lighter wood for understorey species, which might parrallel acquisitive species with higher SLA in this context.
+
 # Supplementary materials
@@
 
+## S3: Guilds organization along economics spectra in French Guiana
+
+**In waiting for BRIDGE dataset.**
+
 # References

### 2018-07-13 draft.Rmd
Uppangala 0.1.9050 BRIDGE test

@@
 output:
-  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
+  bookdown::word_document2: default
 linestretch: 1.5
@@
 ```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values.
-Marginal plots represent the density distribution of species on the two principal component axis for each morphotype.
+Marginal plots represent the density distribution of species on the two principal component axis for each strata.
 See table \\@ref(tab:Traits) for abbreviation.'}
@@
 
-**In waiting for BRIDGE dataset.**
+```{r pcaBRIDGE, fig.cap='Principal Component Analysis (PCA) of species mean trait values for Amazonian species (BRIDGE dataset).
+Marginal plots represent the density distribution of species on the two principal component axis for each strata.'}
+dlim = 0.35
+BRIDGE <- readr::read_tsv("./data/species_TROLL_Nino.txt") %>% 
+  mutate(strata = ifelse(dmax > dlim, "C", "U"))
+BRIDGE.pca <- princomp(~ LMA + Nmass + Pmass + wsg, data = BRIDGE, cor = T)
+pca.plot <- autoplot(BRIDGE.pca, 
+         data = BRIDGE,
+         colour = "strata",
+         size = "Freg",
+         loadings.label.size = 6,
+         loadings.label.colour = 'black',
+         loadings = T, loadings.label = T, loadings.colour = 'black',
+         loadings.label.vjust = 1.2) +
+    coord_equal() +
+  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
+  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
+  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
+  theme(legend.position = c(1, 1), 
+        legend.justification = c(1, 1))
+spectrum.density <- pca.plot$data %>% 
+  select(Comp.1, Comp.2, strata) %>% 
+  rename(WES = Comp.1, LES = Comp.2) %>% 
+  melt(id.vars = "strata", variable.name = "spectrum") %>% 
+  ggplot(aes(value, color = spectrum, fill = spectrum)) +
+  geom_density(alpha = 0.5) + 
+  facet_wrap(~strata, nrow = 2) +
+  coord_flip() +
+  scale_fill_manual("PCA axis", 
+                    labels = c("WES" = "PCA1", "LES" = "PCA2"),
+                    values = c("green", "purple")) +
+  scale_color_manual("PCA axis", 
+                     labels = c("WES" = "PCA1", "LES" = "PCA2"),
+                     values = c("green", "purple"))
+data <- pca.plot$data %>% 
+  select(Comp.1, Comp.2, strata) %>% 
+  rename(WES = Comp.1, LES = Comp.2) %>% # ! test !
+  melt(id.vars = "strata", variable.name = "spectrum")
+Variance <- list(
+  LES = list(
+    C = round(var(data[data$strata == "C" & data$spectrum == "LES", "value"], na.rm = T), 3),
+    U = round(var(data[data$strata == "U" & data$spectrum == "LES", "value"], na.rm = T), 3)),
+ WES = list(
+   C = round(var(data[data$strata == "C" & data$spectrum == "WES", "value"], na.rm = T), 3),
+   U = round(var(data[data$strata == "U" & data$spectrum == "WES", "value"], na.rm = T) ,3))
+)
+Leven_test <- list(
+  C = round(car::leveneTest(value ~ spectrum, data[data$strata == "C",])$`Pr(>F)`[1], 3),
+  U = round(car::leveneTest(value ~ spectrum, data[data$strata == "U",])$`Pr(>F)`[1], 3)
+)
+cowplot::plot_grid(pca.plot, spectrum.density, rel_widths = c(2,1))
+```
+
+* Canopy species WES vs LES: `r Variance$LES$C` vs `r Variance$WES$C`, Levene's test p-value=`r Leven_test$C`
+* Understorey species LES vs WES: `r Variance$LES$U` vs `r Variance$WES$U`, Levene's test p-value=`r Leven_test$U`
 

### 2018-07-13 BRIDGE_test/app.R

@@ -1,0 +1,110 @@
+#
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(shinydashboard)
+library(tidyverse)
+library(ggfortify)
+library(reshape2)
+library(cowplot)
+
+# Define UI for application that draws a histogram
+ui <- dashboardPage(
+  dashboardHeader(title = 'BRIDGE'),
+  dashboardSidebar(
+    textOutput("Levene")
+  ),
+  dashboardBody(
+    fluidPage(
+      fluidRow(
+        sliderInput("dlim",
+                    "Strata diameter limit",
+                    min = 0,
+                    max = 100,
+                    value = 35)
+      ),
+      fluidRow(
+        plotOutput("PCA", height = "800px")
+      )
+    )
+  )
+)
+
+
+# Define server logic required to draw a histogram
+server <- function(input, output) {
+  
+  output$Levene <- renderText({
+    BRIDGE <- readr::read_tsv("./species_TROLL_Nino.txt") %>% 
+      mutate(strata = ifelse(dmax > input$dlim/100, "C", "U"))
+    BRIDGE.pca <- princomp(~ LMA + Nmass + Pmass + wsg, data = BRIDGE, cor = T)
+    pca.plot <- autoplot(BRIDGE.pca, data = BRIDGE)
+    data <- pca.plot$data %>%
+      select(Comp.1, Comp.2, strata) %>%
+      rename(WES = Comp.1, LES = Comp.2) %>% # ! test !
+      melt(id.vars = "strata", variable.name = "spectrum")
+    Variance <- list(
+      LES = list(
+        C = round(var(data[data$strata == "C" & data$spectrum == "LES", "value"], na.rm = T), 3),
+        U = round(var(data[data$strata == "U" & data$spectrum == "LES", "value"], na.rm = T), 3)),
+      WES = list(
+        C = round(var(data[data$strata == "C" & data$spectrum == "WES", "value"], na.rm = T), 3),
+        U = round(var(data[data$strata == "U" & data$spectrum == "WES", "value"], na.rm = T) ,3))
+    )
+    Leven_test <- list(
+      C = round(car::leveneTest(value ~ spectrum, data[data$strata == "C",])$`Pr(>F)`[1], 3),
[…]

### 2018-11-06 draft.Rmd
Uppangala 0.1.9051 Guild position test and new map

@@
 ```{r Uppangala, fig.cap="Canopy model of the 10-ha forest permanent plot in Uppangala, India. Preliminry figure to be improved."}
-knitr::include_graphics("./data/test.png")
+knitr::include_graphics("./data/Plot.png")
 ```
@@
 )
+```
+
+```{r positionTest, eval=F}
+data <- pca.plot$data %>% 
+  select(Comp.1, Comp.2, strata) %>% 
+  rename(WES = Comp.1, LES = Comp.2)
+test <- array(NA, c(2,2,n+1))
+test[,,1] <- data %>% 
+  group_by(strata) %>% 
+  summarise_all(mean) %>% 
+  select(-strata) %>% 
+  as.matrix()
+for(i in 1:n)
+  test[,,i+1] <- data %>% 
+  mutate(strata = sample(strata)) %>% 
+  group_by(strata) %>% 
+  summarise_all(mean) %>% 
+  select(-strata) %>% 
+  as.matrix()
+data %>% 
+  group_by(strata) %>% 
+  summarise_all(mean) %>% 
+  cbind(apply(test, 1:2, function(x){
+    p <- rank(x)[1]/(n+1)
+    if(p > 0.5)
+      p <- 1 - p
+    return(p)
+  })) %>% 
+  rename(WES.pval = `1`, LES.pval = `2`)
 ```

### 2018-11-13 draft.Rmd
Uppangala 0.1.9052 Plot map

@@
 ```{r Uppangala, fig.cap="Canopy model of the 10-ha forest permanent plot in Uppangala, India. Preliminry figure to be improved."}
-knitr::include_graphics("./data/Plot.png")
+DCM <- raster("./data/Canopy.tif")
+crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
+DEM <- raster("./data/DEM.tif")
+crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
+limits <- shapefile("./data/Plots/Shapeall10ha.shp")
+DCM <- crop(DCM, limits)
+DEM <- crop(DEM, limits)
+DEM_spdf <- as(DEM, "SpatialPixelsDataFrame")
+DEM_df <- as.data.frame(DEM_spdf)
+colnames(DEM_df) <- c("Elevation", "x", "y")
+DEM_plot <- ggplot() +  
+  geom_tile(data=DEM_df, aes(x=x, y=y, fill=Elevation), alpha=0.8) + 
+  scale_fill_distiller(palette = "Spectral") +
+  coord_equal() +
+  cowplot::theme_map()
+DCM_spdf <- as(DCM, "SpatialPixelsDataFrame")
+DCM_df <- as.data.frame(DCM_spdf)
+colnames(DCM_df) <- c("Height", "x", "y")
+DCM_plot <- ggplot() +  
+  geom_tile(data=DCM_df, aes(x=x, y=y, fill=Height), alpha=0.8) + 
+  coord_equal() +
+  cowplot::theme_map() +
+  scale_fill_distiller(palette = "Spectral")
+ggsave(cowplot::plot_grid(DEM_plot, DCM_plot, labels = c("a) Elevation (m)", "b) Canopy height (m)")), 
+       file = "Plot.png",
+       path = './data/',
+       width = 200, height = 100, unit = 'mm',
+       dpi = 300)
 ```
@@
   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
-  select(-community, -LES, -WES) %>% 
+  # select(-community, -LES, -WES) %>% 
+  # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
+  select(-community) %>% 
   melt(id.vars = "strata", 
@@
   geom_boxplot() +
-  facet_wrap(~trait, scales = "free", nrow = 1) +
+  # facet_wrap(~trait, scales = "free", nrow = 1) +
+  facet_wrap(~trait, scales = "free", nrow = 2) +
   xlab("Guild") + scale_fill_discrete(guide = "none") +

### 2018-11-19 draft.Rmd
Uppangala 0.1.9053 Plot map, CE figure, PCA figure, strata statistics

@@
 limits <- shapefile("./data/Plots/Shapeall10ha.shp")
-DCM <- crop(DCM, limits)
-DEM <- crop(DEM, limits)
-DEM_spdf <- as(DEM, "SpatialPixelsDataFrame")
-DEM_df <- as.data.frame(DEM_spdf)
-colnames(DEM_df) <- c("Elevation", "x", "y")
-DEM_plot <- ggplot() +  
-  geom_tile(data=DEM_df, aes(x=x, y=y, fill=Elevation), alpha=0.8) + 
-  scale_fill_distiller(palette = "Spectral") +
-  coord_equal() +
-  cowplot::theme_map()
-DCM_spdf <- as(DCM, "SpatialPixelsDataFrame")
-DCM_df <- as.data.frame(DCM_spdf)
-colnames(DCM_df) <- c("Height", "x", "y")
-DCM_plot <- ggplot() +  
-  geom_tile(data=DCM_df, aes(x=x, y=y, fill=Height), alpha=0.8) + 
+limits$melt <- 1
+limits <- maptools::unionSpatialPolygons(limits, limits$melt)
+DCM <- crop(DCM, buffer(limits, 30))
+DEM <- crop(DEM, buffer(limits, 30))
+DEM <- as(DEM, "SpatialPixelsDataFrame")
+DEM <- as.data.frame(DEM)
+colnames(DEM) <- c("Elevation", "x", "y")
+DCM <- as(DCM, "SpatialPixelsDataFrame")
+DCM <- as.data.frame(DCM)
+colnames(DCM) <- c("Height", "x", "y")
+ggplot() +  
+  geom_tile(data=DCM, aes(x=x, y=y, fill=Height)) +
+  geom_contour(data=DEM, aes(x=x, y=y, z=Elevation),
+               colour = "black", alpha = 0.6, breaks = seq(300, 500, by = 25)) +
+  directlabels::geom_dl(data=DEM, aes(x=x, y=y, z=Elevation, label = paste0(..level.., "m")),
+                        method = list('last.bumpup', cex = 0.8, hjust = 1),
+                        stat="contour", breaks = seq(300, 500, by = 25)) +
+  geom_polygon(data = fortify(limits),
+                 aes(long, lat, group=group),
+                 colour = "black", lwd = 1.1, alpha = 0) +
   coord_equal() +
   cowplot::theme_map() +
-  scale_fill_distiller(palette = "Spectral")
-ggsave(cowplot::plot_grid(DEM_plot, DCM_plot, labels = c("a) Elevation (m)", "b) Canopy height (m)")), 
-       file = "Plot.png",
+  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
+  ggsave(file = "Plot.png",
          path = './data/',
-       width = 200, height = 100, unit = 'mm',
+         width = 150, height = 200, unit = 'mm',
          dpi = 300)
+rm(DEM, DCM, limits, contour, DEM_df, DCM_df, DEM_spdf, DCM_spdf, DCM_plot, DEM_plot)
 ```
@@
         legend.justification = c(1, 1))
-spectrum.density <-pca.plot$data %>% 
+spectrum.density <- pca.plot$data %>%
   select(Comp.1, Comp.2, strata) %>%
@@
                      values = c("green", "purple"))
-cowplot::plot_grid(pca.plot, spectrum.density, rel_widths = c(2,1))
+spectrum.boxplots <- pca.plot$data %>%
+  select(Comp.1, Comp.2, strata) %>%
+  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
+  melt(id.vars = "strata", variable.name = "spectrum") %>%
+  ggplot(aes(strata, value, fill = strata)) +
+  ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
+  geom_boxplot() +
+  facet_wrap(~spectrum, nrow = 2)
+# cowplot::plot_grid(pca.plot, spectrum.density, rel_widths = c(2,1))
+ggsave(cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1), labels = c("A", "B")), 
+       file = "PCA.png",
+       path = './data/',
+       width = 200, height = 150, unit = 'mm',
+       dpi = 300)
 ```
@@
   # facet_wrap(~trait, scales = "free", nrow = 1) +
+  facet_wrap(~trait, scales = "free", nrow = 2) +
+  xlab("Guild") + scale_fill_discrete(guide = "none") +
+  ggpubr::stat_compare_means(aes(label = paste(..p.signif..))) 
+```
+
+```{r GMboxplot, fig.cap="Boxplot of guilds means (GM) among communties for each traits (for canopy (C) and understorey (U) guilds). See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
+L %>%
+  melt(id.var = "com",
+       variable.name = "SpCode",
+       value.name = "abundance") %>% 
+  mutate(abundance = ifelse(abundance > 0, 1, 0)) %>% 
+  dcast(SpCode ~ com, value.var = "abundance") %>% 
+  mutate(SpCode = as.character(SpCode)) %>% 
+  inner_join(Q, by = "SpCode") %>%
+  select(-SpCode) %>%
+  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
+       variable.name = 'community', value.name = 'abundance') %>%
+  group_by(strata, community) %>%
+  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
+  select(-community) %>% 
+  melt(id.vars = "strata", 
+       variable.name = "trait",
+       value.name = "GM") %>% 
+  ggplot(aes(x = strata, y = GM, fill = strata)) +
+  geom_boxplot() +
   facet_wrap(~trait, scales = "free", nrow = 2) +
@@
 
+## S4 : Dawkins effect on traits
+
+```{r CEeffect}
+# L %>% 
+#   melt(id.vars = "com",
+#        variable.name = "SpCode",
+#        value.name = "individuals") %>% 
+#   filter(individuals > 0) %>% 
+#   left_join(Q %>% 
+#               select(SpCode, strata)) %>% 
+#   filter(!is.na(strata)) %>% 
+#   group_by(com, strata) %>% 
+#   summarise(species = length(unique(SpCode)),
+#             individuals = sum(individuals)) %>% 
+#   ungroup() %>% 
+#   # group_by(strata) %>% 
+#   summarise(species = mean(species), individuals = mean(individuals))
+# genPFT() %>%
+#   mutate(CE = gsub('([[:alpha:]])', "", CE)) %>%
+#   filter(!is.na(CE)) %>%
+#   mutate(LA = log(LA)) %>%
+#   mutate(SLA = log(SLA)) %>%
+#   select(CE, Sp, WD, LT, LA, LDMC, LTD, SLA) %>%
+#   reshape2::melt(id.vars = c("CE", "Sp"), variable.name = "Trait") %>%
+#   group_by(Trait) %>%
+#   do(anova = TukeyHSD(aov(lm(value ~ CE, data = .)))) %>%
+#   tidy(anova) %>%
+#   filter(adj.p.value < 0.05)
+my_comparisons <- list( c("1", "2"), c("2", "3"), c("3", "4"),
+                        c("4", "5"), c("1", "3"), c("3", "5"),
+                        c("1", "5"),  c("2", "4"), c("2", "5"))
+genPFT() %>% 
+  mutate(CE = gsub('([[:alpha:]])', "", CE)) %>% 
+  filter(!is.na(CE)) %>% 
+  select(CE, WD, LT, LA, LDMC, LTD, SLA) %>% 
+  mutate(LA = log(LA)) %>% 
+  mutate(SLA = log(SLA)) %>% 
+  reshape2::melt(id.vars = "CE", variable.name = "Trait") %>% 
+  ggplot(aes(CE, value, fill = CE)) +
+  geom_boxplot() +
+  facet_wrap(~Trait, scales = "free") +
+  ggpubr::stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..)) +
+  ggpubr::stat_compare_means(method = "anova", aes(label = paste("Anova", ..p.signif..))) +
+  ggsave(file = "CE.png",
+         path = './data/',
+         width = 200, height = 200, unit = 'mm',
+         dpi = 300)
+
+genPFT() %>% 
+  select(CE) %>% 
+  mutate(CE = gsub('([[:alpha:]])', "", CE)) %>% 
+  group_by(CE) %>% 
+  summarise(n = n()) %>% 
+  ungroup() %>% 
+  mutate(P = round(n/sum(n)*100))
+```
+
 # References

### 2019-06-27 draft.Rmd
Uppangala 0.1.9054 Francois sharing

@@
 output:
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
 linestretch: 1.5
@@
   theme(legend.position = c(1, 1), 
-        legend.justification = c(1, 1))
-spectrum.density <- pca.plot$data %>%
-  select(Comp.1, Comp.2, strata) %>%
-  rename(WES = Comp.1, LES = Comp.2) %>%
-  melt(id.vars = "strata", variable.name = "spectrum") %>%
-  ggplot(aes(value, color = spectrum, fill = spectrum)) +
-  geom_density(alpha = 0.5) +
-  facet_wrap(~strata, nrow = 2) +
-  coord_flip() +
-  scale_fill_manual("PCA axis",
-                    labels = c("WES" = "PCA1", "LES" = "PCA2"),
-                    values = c("green", "purple")) +
-  scale_color_manual("PCA axis",
-                     labels = c("WES" = "PCA1", "LES" = "PCA2"),
-                     values = c("green", "purple"))
+        legend.justification = c(1, 1)) +
+  scale_color_discrete("Guild")
 spectrum.boxplots <- pca.plot$data %>%
@@
   geom_boxplot() +
-  facet_wrap(~spectrum, nrow = 2)
-# cowplot::plot_grid(pca.plot, spectrum.density, rel_widths = c(2,1))
+  facet_wrap(~spectrum, nrow = 2) +
+  scale_fill_discrete("Guild")
+cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
 ggsave(cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1), labels = c("A", "B")),
@@
 )
+pca.plot$data %>%
+  select(Comp.1, Comp.2, strata) %>%
+  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
+  melt(id.vars = "strata", variable.name = "spectrum") %>% 
+  group_by(spectrum) %>% 
+  do(anova = aov(value ~ strata, data = .)) %>% 
+  broom::tidy(anova)
 ```
@@
   summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
-  # select(-community, -LES, -WES) %>% 
+  select(-community, -LES, -WES) %>%
   # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
-  select(-community) %>% 
+  # select(-community) %>% 
   melt(id.vars = "strata", 
@@
   geom_boxplot() +
-  # facet_wrap(~trait, scales = "free", nrow = 1) +
-  facet_wrap(~trait, scales = "free", nrow = 2) +
+  facet_wrap(~trait, scales = "free", nrow = 1) +
+  # facet_wrap(~trait, scales = "free", nrow = 2) +
   xlab("Guild") + scale_fill_discrete(guide = "none") +
-  ggpubr::stat_compare_means(aes(label = paste(..p.signif..))) 
+  ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3)))) +
+  ggsave(file = "boxplot.png",
+         path = './data/',
+         width = 200, height = 150, unit = 'mm',
+         dpi = 300)
 ```
@@
      direction = 'both', trace = 0)) %>% 
+  # do(reg = lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .),
+  #    direction = 'both', trace = 0) %>%
   tidy(reg) %>%
   group_by(strata, trait, term) %>%
-  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
-                          stars(p.value, ns = '   '))) %>% 
+  # transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
+                          # stars(p.value, ns = '   '))) %>% 
+  transmute(value = format(estimate, scientific = T, digits = 2, trim = T)) %>% 
   dcast(strata + trait ~ term) %>% 
@@
   filter(trait != 'WES') %>%
-  bold() %>% 
+  # bold() %>% 
   kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance.
 We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), and canopy height (Canopy), and only the selected predictors are shown.
-See table \\@ref(tab:Traits) for abbreviations.
-Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
+See table \\@ref(tab:Traits) for abbreviations.', format = 'pandoc')
 ```
@@
 #   filter(adj.p.value < 0.05)
+# genPFT() %>%
+#   mutate(CE = gsub('([[:alpha:]])', "", CE)) %>%
+#   filter(!is.na(CE)) %>%
+#   mutate(LA = log(LA)) %>%
+#   mutate(SLA = log(SLA)) %>%
+#   select(CE, Sp, WD, LT, LA, LDMC, LTD, SLA) %>%
+#   reshape2::melt(id.vars = c("CE", "Sp"), variable.name = "Trait") %>%
+#   mutate(CE = as.numeric(CE)) %>% 
+#   group_by(Trait) %>%
+#   do(lm = lm(value ~ CE + Sp, data = .)) %>%
+#   tidy(lm) %>% 
+#   filter(term == "CE") %>% 
+#   filter(p.value < 0.05)
 my_comparisons <- list( c("1", "2"), c("2", "3"), c("3", "4"),
@@
          dpi = 300)
-
 genPFT() %>% 
@@
   mutate(P = round(n/sum(n)*100))
+genPFT() %>%
+  mutate(CE = as.numeric(gsub('([[:alpha:]])', "", CE))) %>%
+  filter(CE != 5) %>% 
+  filter(!is.na(CE)) %>%
+  ggplot(aes(CE, WD, group = NA)) +
+  geom_boxplot(aes(group = CE), col = "lightgrey") +
+  geom_point(aes(col = Sp), alpha = 0.3) +
+  geom_smooth(method = "lm") +
+  scale_color_discrete(guide = "none")
 ```

### 2019-11-15 Manuscript.Rmd
Uppangala 0.1.9055 Xmas 19 vs

---
title: Environmental filtering favor different functional composition of tree assemblages between canopy and understory guilds within Indian forest communities
date: '`r Sys.Date()`'
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(raster)
library(parallel)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
library(reshape2)
library(ggfortify)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
n = 999 # Random repetitions
sapply(list.files("R", full.names = T), source)
```

```{r data}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Sp, -LTD, -SHA)
suppressWarnings(row.names(PFT) <- PFT$SpCode) # deprecated
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
Q <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(data.frame(PFT.pca$scores) %>% 
              mutate(SpCode = row.names(.)),
            by = "SpCode") %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  select(SLA, LDMC, WD, LA, LT, LES, WES, strata, SpCode) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata))
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
L <- genTrees() %>% 
  mutate(com = (XY %over% com)[,1]) %>% 
  filter(!is.na(com)) %>% 
  group_by(com, SpCode) %>% 
  summarise(n = n()) %>% 
  dcast(com ~ SpCode, value.var = "n") %>% 
  replace(is.na(.), 0)
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWMses
GWM <- L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(Q, by = "SpCode") %>% 
  filter(!is.na(strata)) %>% 
  select(-SpCode) %>% 
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>% 
  group_by(strata, community) %>% 
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# nQ <- array(rep(as.matrix(Q), n),
#             c(dim(Q), n),
#             list(row.names(Q), names(Q), 1:n))
# nQ <- array(apply(nQ, 3, function(x){
#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
#   return(x)
#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
# cl <- makeCluster(detectCores())
# clusterExport(cl, list('L', 'nQ'))
# nGWM <- parApply(cl, nQ, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   L %>%
#     melt(id.var = "com",
#          variable.name = "SpCode",
#          value.name = "abundance") %>%
#     dcast(SpCode ~ com) %>%
#     left_join(data.frame(x)) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     group_by(strata, community) %>%
#     mutate_each(funs(as.character), -abundance) %>%
#     mutate_each(funs(as.numeric), -abundance) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# })
# stopCluster(cl)
# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
#               c(dim(GWM[3:9]), n),
#               list(row.names(GWM), names(GWM[3:9]), 1:n))
# rm(cl, nQ)
# save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
load(file.path("draft_save", 'nGWM.Rdata'))
GWM_ses <- GWM
GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
```

1. Sylvain Schmitt
    - Corresponding author
    - sylvain.m.schmitt@gmail.com
    - orcid.org/0000-0001-7759-7106
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
1. Maxime Réjou-Méchain
    - maxime.rejou@gmail.com
    - orcid.org/0000-0003-2824-267X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR AMAP, IRD, F-34000 Montpellier, France
1. Valérie Raevel
    - vraevel@gmail.com
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Narayanan Ayyappan
    - ayyappan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Natesan Balachandran
    - balachandran.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. N. Barathan
    - barathan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Chandra Shekhar Jha
    - email
    - Forestry and Ecology Group, National Remote Sensing Centre, Hyderabad 500 037, India
1. François Munoz
    - fmunoz@univ-grenoble-alpes.fr
    - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France

\newpage

# Abstract

1. Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities. Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition. Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests, or focusing on canopy trees.
1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India. Combining tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits, we used multivariate approaches and linear regressions to assess the variation of functional assemblages within tropical tree communities between understory and canopy species, and to address the effects of abiotic and biotic environments on these variations.  
1. We found that decoupled wood and leaf economics spectra were present in Indian rainforests and were maintained at local scale. Moreover, we showed that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities. We further showed that the canopy and understory guilds assemblages responded differently across communities to environmental gradients related to soil resource, light availabilities and monsoon disturbances.
1. *Synthesis.* These results illustrate how fine-scale environmental heterogeneity may shape the tree community composition and function. Moreover, our study show guild-specific answer to biotic and abiotic environmental filtering between canopy and understory, highlighting the importance of forest vertical stratification in community assembly processes.

## Keywords

Rainforest; Western Ghats; Economics spectra; Environmental filtering, Vertical strata

\newpage

# Introduction

<!-- hook topo + vertical str -->

Understanding the drivers of hyperdiverse tropical forest tree composition is a long-standing issue in ecology [@Cody1976]. Abiotic environment has been shown to influence tree community composition at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. Specifically, topography can filter different ecological strategies [@kraft_functional_2008] due to variation in water and nutrient availability in tropical forests [@ferry2010higher]. But biotic interactions can also have a major impact on tree community composition [@Uriarte2004]. Indeed, light availability, and thus community vertical structure, represent a major constraint for tree survival and growth in tropical forests [@Stark2015]. Joint study of topography and community vertical structure therefore hold promise  to understand the drivers of hyperdiverse tropical communities [@Jucker2018].

<!-- topo and vert strata, what do we know ? -->

Higher topographic positions, e.g. hilltops, result in decreased water and nutrient availability [@ferry2010higher] filtering for species with conservative strategy of resource use [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017]. In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013]. Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011]. Therefore, understory and canopy guilds can differently respond to environmental gradients. Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests [@wright2002plant], or focusing on canopy trees [@Hubbell2001].  Whereas, investigating community assembly among vertical strata may reveal distinct response to environmental gradients [@Raevel2018], and thus provide a better understanding of community assembly processes.

<!-- FT for community assembly -->

Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], contributing thus to test for community assembly processes. Functional traits have been defined as phenotypic traits that impact fitness indirectly through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007]. Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009]. The leaf economics spectrum opposes acquisitive ecological strategies, with high photosynthetic carbon assimilation, to conservative ones with high investment in leaf defense and durability. For instance, specific leaf area (leaf mass divided by leaf area) varies along soil and light availability gradients, and represents a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016]. The wood economics spectrum opposes slow and fast growing species. However, @reich_world-wide_2014 advocate for a unique plant economics spectrum between fast and slow growing species, evidenced in a subarctic flora [@freschet_evidence_2010]. Functional traits have been used to evidenced community assembly processes at local scale [@fortunel_leaf_2012; @Paine2011; but see @Messier2017], but to our knowledge little is known of the role of tropical forest vertical stratification in community trait-based studies.

<!-- what have we done ? -->

In the present study, we assessed the variation of functional assemblages within tropical tree communities between understory and canopy species, and addressed the effects of abiotic and biotic environments on these variations. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000]. The plot encompasses a large variation of topography [@Antin2013] and vertical structure [@Pascal1988]. Combining tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits, we used multivariate approaches and linear regressions to address the following questions: (i) Do the trait variations among species follow independent leaf and wood economics spectra at local scale ? If so, how are canopy and understory guilds distributed along leaf and wood economics spectra ? (ii) Does functional composition differ between canopy and understory guilds ? and (iii) Does the guild functional composition vary with local environmental heterogeneity, and do the understory and canopy guilds respond differently ? Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017]. We hypothesized functional composition to differ between canopy and understorey species, due to their differences in growth and functional strategies [@Herault2011]. Because this region experiences a long dry season period and is under the influence of intense monsoon rainy seasons with strong wind disturbances, we expected that the local topography is a major driver of tree community functional composition.

# Material and Methods

## Study site

The study was carried out in a nearly 10-ha (300m*330m) permanent plot established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996]. The plot is named after the nearest village, Uppangala, and is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India). Annual precipitation averages 5,100 mm per year but is highly variable throughout the year with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [@Pascal1996]. The geology can be considered as homogeneous within the Uppangala plot [@Ferry1994]. The topography of the site is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig. \@ref(fig:Uppangala)). Topography is thus expected to represent a major component of environmental variation within the forest plot [@Antin2013; @gimaret-carpentier_sampling_1998].

## Taxonomic composition and trait sampling

Since 1989, all trees $\geq$ 10 cm in diameter at breast height (dbh) were identified to species level, tagged, mapped and monitored in an area progressively extended to a contiguous 10-ha plot in 2014 following international standards [@Jairazbhoy2006]. The plot includes 6672 tree individuals belonging to 111 tree species mostly from the Dipterocarpaceae, Myristicaceae and Fabaceae families.

Using the classification of @Pascal1988, we assigned a potential vertical stratum of tree species (vertical stratum reached by most adult trees) to all species from the plot. @Pascal1988’s classification consisted of four tree species categories: emergent species above 39m, canopy species between 25 and 39m, sub-canopy species between 15 and 25m, and understory species below 15m. Because our study site is characterized by a steep topography, sub-canopy species often reach the canopy in the Uppangala plot. We thus defined two vertical strata, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees. This resulted in 54 and 35 species belonging to the canopy and understory guilds, respectively.

We collected trait data from 89 over the 111 (80.1%) species, representing 99.0% of tree individuals in the plot. We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species, depending on leaf accessibility and species abundance (3 individuals on average). For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter close to 1 cm, following recommendation of @Perez-Harguindeguy2013. Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]). Leaf and branch samples were kept in humidified ziplock bags filled with enriched CO2 air in dark until measurement within the next 24 hours. For 11 species (1% of individuals in Uppangala plot), branch samples were too small (1 cm in diameter) to allow measuring branch wood density.

## Trait measurements

We measured five functional traits (Table \@ref(tab:traits)) related to the major dimensions of ecological strategies in plants [@diaz_global_2015; @Wright2002]: leaf thickness (LT, $\mu m$), leaf area (LA, $mm^2$), leaf dry matter content (leaf dry weight divided by the leaf fresh weight, LDMC, $mg.g^{-1}$), specific leaf area (leaf area divided by the leaf dry mass, SLA, $m^2.kg^{-1}$), and branch wood density (wood dry weight divided by its volume, WD, $g.cm^{-3}$). Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weight was measured with an analytical balance (Wensar, KD-H600, 0.01g), leaf fresh thickness with a micrometer and leaf area by  analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012]. Leaves were then kept in herbarium and oven-dried for at least 72 hours at 60°C before measurement of dry weight. The bark and pit of branch samples were removed in order to keep only branch wood without soft tissues. Branch wood samples were immersed in water to evaluate their volume by water displacement. Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of [@Perez-Harguindeguy2013].

## Environmental variables

We used LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy conditions. The LiDAR campaign was conducted in 2013, hence concomitantly with field measurements. Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, which relates to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the effect of dominant monsoon winds [@Bunyan2015]; and (iv) slope, which may have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher]. In addition, using a 1-m resolution LiDAR canopy model (Fig. \@ref(fig:Uppangala)), we used the mean of the top of canopy height to  represent the intensity of local competition for light.

## Statistical Analyses

To assess the degree of trait variation and covariation we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific). We then performed a principal component analysis (PCA) of the average trait values per species to characterize the main dimensions of ecological strategies globally and per vertical stratum.

We then analyzed the functional composition of vertical stratum communities in $30m*30m$ subplots (n=110). Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.

We calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’, [@Raevel2018]) of each trait in each subplot:

$$GWM_j=\sum_{i=1}^Sp_{i,j}*t_i$$
where $p_{i,j}$ is the relative abundance of species $i$ in subplot $j$, $t_i$ the mean trait value of species $i$ and $S$ the number of species in the guild considered.

To test whether the local functional composition of guilds depends on the local topography and canopy structure, we devised a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild, [@bernard-verdier_community_2012]). This null model broke the link between environment and local functional composition while maintaining the spatial structure of guilds and species abundances. Null GWM values generated through 999 randomizations were subsequently used to calculate the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as

$$SES = \frac{GWM_{obs}-\overline{GWM_null}}{\sigma_{null}}$$
where $GWM_{obs}$ is the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ are the mean and standard deviation of the corresponding null distribution, respectively.

Then we assessed how the deviation of guild composition from the null model depended on environmental variables using linear regressions with bidirectional stepwise selection of SES values. Thus, the topographical or canopy structure variables selected with this approach were interpreted as more impacting GWM values than expected by chance. These analyses were performed independently for each of the five functional traits. All analyses were conducted using the R 3.3.0 software [@RCoreTeam2018].

# Results

## Trait variation and covariation

We found an overall increase of functional trait variation from intra-individual to inter-specific levels (Table \@ref(tab:traits)), even though LA showed almost as large variability at intra-individual than at intra-specific level. Thus, trait values averaged at species level grasp most of the functional variability in the species pool considered here.

The first axis of the PCA performed on species trait values (Fig. \@ref(fig:pca)A) was basically related to wood density and structural leaf trait  (LT, LA, LDMC), opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood. The second axis mostly opposed SLA to LDMC. WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that they provide independent information. Canopy and understory species did not show significant differentiation on the first and second PCA axis (Wilcoxon p=0.2 and p=0.3, respectively Fig. \@ref(fig:pca)B).

## Changes in functional composition with environment

At the community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig. \@ref(fig:GWMboxplot)). GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, higher for LA and lower for LDMC and LT than understory tree assemblages. In other words, understory guild assemblages had larger, thinner and lighter leaves with higher SLA and denser woods, compared to canopy guild assemblages.

We detected a significant effect of slope, curvature and wetness on both canopy and understory guilds assemblages, of southwesterness on canopy guild assemblages and of canopy height on understory guild assemblages (Table \@ref(tab:LM)).

# Discussion

Assessing the variation of functional assemblages within tropical tree communities between understory and canopy species, we showed that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities. We further showed that the canopy and understory guilds assemblages responded differently across communities to environmental gradients related to soil resource, light availabilities and monsoon disturbances. While both understory and canopy assemblages responded similarly to water and nutrient access, canopy assemblages responded to monsoon winds with stronger support of the wood economics spectrum and understory assemblages responded to light access with a strong support of the leaf economics spectrum. These results illustrate how fine-scale environmental heterogeneity may shape the tree community composition and function. Moreover our study show guild-specific answer to biotic and abiotic environmental filtering between canopy and understory, highlighting the importance of forest vertical stratification in community assembly processes.

## Decoupled wood and leaf economics spectra at local scale

Despite a reduced number of functional traits, our study support the existence of decoupled wood and leaf economics spectra at local scale in Indian rainforests. We found locally that species trait values varied along two orthogonal main dimensions (Fig. \@ref(fig:pca)). SLA and LDMC negative correlation have already been linked to the leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004], representing a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Baraloto2010; @Evans2001; @hodgson_is_2011]. Whereas, wood density was opposed to leaf area and thickness along the wood economics spectrum [@chave_towards_2009], which contrasted species more tolerant to hydric stress, with denser wood, to more competitive species with thicker and larger leaves and higher efficiency in water conductivity [@Santiago2004a]. To our knowledge, our study is the first to support the two economics spectra in Indian rainforests. In addition, @Baraloto2010 and @Fortunel2012a found a similar structuring of Neotropical trees along these two spectra at the regional scale in Amazonia. Thus, our results indicate that the existence of the two economics spectra can hold at a very local scale (Fig. \@ref(fig:pca)A), although @Messier2017 suggested that the spectrum would not hold at local scale. Finally, our study support the idea of decoupled economics spectra [@Baraloto2010] and not a whole plant economics spectrum [@reich_world-wide_2014].

## Functional differences between canopy and understory guilds

Our results illustrated that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities. Indeed, we found no significant differences between functional traits of species pools from canopy and understory along the two main economics spectra (Fig. \@ref(fig:pca)B). Thus, species pools from canopy and understory overlap in functional space and both span a broad range of ecological strategies. Contrastingly, there was a sharp difference between the average trait values of guild assemblages across subplots (Fig. \@ref(fig:GWMboxplot)). Understory assemblages had larger, thinner and lighter leaves, thus greater SLA, and denser wood, than canopy assemblages, reflecting that understorey assemblages tend to have a more conservative strategy while canopy assemblages have a more competitive and acquisitive strategy. In other words, while understory assemblages favor tolerance to hydric stress and leaf investment and durability, canopy assemblages favor efficiency in water conductivity and resources acquisition. The observed contrast between species pools and guilds assemblages within communities of canopy and understory was solely due to the environmental filtering selecting for more abundant species with different strategies, while the overall range of species strategies broadly overlapped between guilds. Thus, while the guilds were adapting their functional strategy to the environment within communities, they conserved broad ecological variation insuring greater stability in space and time to maintain the guilds functions [@Yachi1999].

## Different responses to abiotic and biotic environment between canopy and understory assemblages

Our results support the hypothesis formulated by @duque2002different that trees experience guild-specific ecological constraints, leading to contrasting functional compositions between canopy and understory assemblages. Indeed, the way assemblages responded to environmental variation changed among guilds (Table \@ref(tab:LM)). While both understory and canopy assemblages responded similarly to slope wetness, and curvature, canopy assemblages responded to southwesterness with stronger support of the wood economics spectrum (higher $R^2$ for WD and LA), and understory assemblages responded to canopy height with a strong support of the SLA representing the leaf economics spectrum ($R^2=0.5$ for SLA model in understory, but WD model have also a good $R^2$). The wood economics spectrum major role in the response to environment of canopy assemblages might be explained by the critical role of mechanical resistance, disturbance tolerance and water conductivity efficiency in the establishment and persistence of canopy trees [@Poorter2010b]. Conversely, understory assemblages are subject to limited light availability due to interception by trees in canopy, resulting in a critical role of resource acquisition constraining functional composition of understory trees along the leaf economics spectrum.

The influence of aspect, through southwesterness, on forest composition is little documented in tropical forests but was expected to be highly context dependent. The Uppangala plot shows large topographic heterogeneity, with steep slopes mostly organized along southwest-oriented crests (Fig. \@ref(fig:Uppangala)). @gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar context in Sri Lanka: they suggested that it can influence the local environment according to the exposure to monsoon winds. In addition, @Bunyan2015 suggested that greater occurrence of forest fragments on southwest orientation in a forest-grassland mosaic of Western Ghats could result from monsoon rains. In our context, monsoon winds are also south-west oriented, and are expected to be a major source of disturbance in forest [@soman_aspects_1990]. The possible consequences are that (i) harder wood and smaller leaves are selected for mechanical resistance in more disturbed southwest oriented area, or (ii) topography entails turbulences on northeast oriented areas responsible of more disturbance and death events in the forest history resulting in lightwood filtering [@larjavaara_rethinking_2010]. In both cases, canopy species tend to be more exposed to disturbance events once adult, which is consistent with the fact that aspect only influenced canopy guild functional assemblages.

In addition, higher canopy significantly impacted the functional composition of understory species assemblages. Taller canopies resulted in more acquisitive strategies with higher SLA and lower LDMC and LT, maybe due to an increased light availability of less covering taller canopy resulting in less limited light-conditions and decreased investment in leaf lifespan [@kitajima_tissue-level_2010]. Putative increased light-conditions of taller canopies, may also have increased understory assemblages competitiveness resulting in lighter woods for relatively faster growth. These remains hypotheses to be thoroughly tested.

Finally, variation in curvature, wetness and slope similarly affected canopy and understory functional assemblages (at the exception of LDMC). Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and retain more water, increasing soil fertility [@ferry2010higher], creating favorable conditions where species assemblages with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA) strategy will outcompete others. By contrast, concave, drier or sloppy areas will be dryer and less fertile favoring species assemblages with more conservative strategies and slower growth.

## Conclusion

> *It does not really bring anything new, should I remove it ?*

Our results show that a broad functional variation among species and assemblages in an Indian rainforest is structured along decoupled leaf and wood economic spectra. We showed a significant environmental determinism in the functional composition of tree communities but also stressed that this determinism is guild-specific. Our results  thus contribute to the debate on the nature and role of economics spectra in the assembly of tropical forests [@Baraloto2010]. They also underline that the functional signature of community assembly should be decomposed to better characterize and understand the different strategies and responses across ecological guilds [@Raevel2018] and across forest vertical structure [@duque2002different]. We thus call for a better acknowledgment of differential guild dynamics within communities and their influence on the global forest functioning.

# Acknowledgements

> *Who financed the project ? I think I'll acknowledge everybody that more or less helped on the field, and we should acknowledge someone for the plot extension right ? Anybody or anything else ?*

# Authors’ contributions

SS, MM, FM and VR conceived the study, designed the methodology and analysed model outputs; SS, FM, VR, NA, NB and NB conducted the field sampling; NA, NB and NB identified species; MM and CJ provided the environmental data; SS, MM, FM and VR led the writing of the manuscript. All the authors contributed to the drafts and gave their final approval for publication.

# Data availability

Functional traits data have been submitted to TRY initiative [@Kattge2011] under the name UppangalaTraits. Environmental variables and spatial positions of individuals were extracted from the Uppangala Station database or **lidar data**, for which access is modulated by **to precise**.

# References

<div id="refs"></div>

\newpage

## Table captions

__Table \@ref(tab:traits)__ Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).

__Table \@ref(tab:LM)__ Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy). Linear model coefficients are reported only for predictors retained after bidirectional stepwise selection, with corresponding adjusted $R^2$. See Table \@ref(tab:traits) for trait abbreviations.

## Figure captions

__Figure \@ref(fig:Uppangala)__ Digital canopy model of the 10-ha forest permanent plot in Uppangala, India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively.

__Figure \@ref(fig:pca)__ Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B). Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test. See Table \@ref(tab:traits) for trait abbreviations.

__Figure \@ref(fig:GWMboxplot)__ Guild weighted mean (GWM) of specific leaf area (SLA), wood density (WD), leaf dry matter content (LDMC), leaf area (LA) and leaf thickness (LT) for canopy (C) and understory (U) guilds. Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines). Asterisks represent significant differences between the two guilds (*** p-value<0.001).

\newpage

<!-- table 1 -->

```{r traits, results="latex"}
t <- genPFT() %>% 
  mutate(id = row.names(.)) %>% 
  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
  melt(id.vars = "id", variable.name = 'Abbreviation') %>% 
  mutate(Abbreviation = as.character(Abbreviation)) %>% 
  left_join(data.frame(
    Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
    Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
    Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
    Strategy = c('Leaf defense and investment',
                 'Leaf resource capture vs. investment',
                 'Leaf defense and investment',
                 'Leaf resource acquisition vs. defense',
                 'Stem transport, structure and defense'),
    stringsAsFactors = F
  ),  by = "Abbreviation") %>% 
  group_by(Trait, Abbreviation, Unit, Strategy) %>% 
  summarise_if(is.numeric, funs(min(.,na.rm = T), max(., na.rm = T))) %>% 
  mutate(range = paste(round(min,2), "-", round(max,2))) %>% 
  select(-min, -max) %>% 
  left_join(genPFT() %>% 
              select(-Sp, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Tree', 
                   variable.name = 'Abbreviation', 
                   value.name = 'Intra-individuals') %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Tree, Abbreviation) %>% 
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Tree) %>% 
              group_by(Abbreviation) %>% 
              summarise_all(funs(mean(., na.rm = T))),  by = "Abbreviation") %>% 
  left_join(genPFT() %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Sp', 
                   variable.name = 'Abbreviation', value.name = 'Intra-species') %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Sp, Abbreviation) %>% 
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Sp) %>% 
              group_by(Abbreviation) %>% 
              summarise_all(funs(mean(., na.rm = T))),  by = "Abbreviation") %>% 
  left_join(genPFT() %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -CE) %>% 
              group_by(Sp, SpCode) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Sp, -SpCode, -SHA, -LTD) %>%
              mutate(id = row.names(.)) %>% 
              melt(id.vars = "id", 
                   variable.name = 'Abbreviation', value.name = 'Inter-species') %>% 
              select(-id) %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Abbreviation) %>%
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))),
            by = "Abbreviation") %>%
  mutate(`Intra-individuals` = ifelse(Abbreviation == "WD", NA, `Intra-individuals`))

if(knitr:::is_latex_output()){
  t %>% 
    ungroup() %>% 
    mutate(Trait = c('Leaf area', 'Leaf dry\nmatter content', 'Leaf\nthickness', 'Specific\nleaf area', 'Wood\ndensity'),
    Strategy = c('Leaf defense\nand investment',
                 'Leaf resource\ncapture vs. investment',
                 'Leaf defense\nand investment',
                 'Leaf resource\nacquisition vs. defense',
                 'Stem transport,\nstructure and defense')) %>% 
    mutate_all(linebreak) %>% 
    kable(caption = 'Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).',
          digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
          col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
                        "\\makecell[l]{Intra-\\\\individuals}", 
                        "\\makecell[l]{Intra-\\\\species}", 
                        "\\makecell[l]{Inter-\\\\species}")) %>%
    kable_styling(full_width = T)
} else {  
  t %>% 
  kable(digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
        caption = 'Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).') %>% 
  kable_styling(full_width = T)
}
```

\newpage

<!-- table 2 -->

```{r LM}
# GWM_ses %>% 
#   filter(!is.na(strata)) %>% 
#   melt(id.vars = c('strata', 'community')) %>% 
#   rename(trait = variable) %>% 
#   rename(SES = value) %>% 
#   mutate(community = as.character(community)) %>% 
#   left_join(R, by = "community") %>%
#   group_by(strata, trait) %>% 
#   do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
#      direction = 'both', trace = 0)) %>% 
#   # do(reg = lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .),
#   #    direction = 'both', trace = 0) %>%
#   tidy(reg) %>%
#   group_by(strata, trait, term) %>%
#   # transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
#                           # stars(p.value, ns = '   '))) %>% 
#   transmute(value = format(estimate, scientific = T, digits = 2, trim = T)) %>% 
#   dcast(strata + trait ~ term) %>% 
#   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
#   mutate_all(funs(replace(., is.na(.), ' '))) %>% 
#   left_join( # Inclusion of R2 but should be split
#     GWM_ses %>% 
#       filter(!is.na(strata)) %>% 
#       melt(id.vars = c('strata', 'community')) %>% 
#       rename(trait = variable) %>% 
#       rename(SES = value) %>% 
#       mutate(communtiy = as.character(community)) %>% 
#       left_join(R, by = "community") %>%
#       group_by(strata, trait) %>% 
#       do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
#                     direction = 'both', trace = 0)) %>% 
#       glance(reg) %>% 
#       select(r.squared) %>% 
#       mutate(r.squared = round(r.squared, 2)) %>% 
#       rename("$R^2$" = r.squared), by = c("strata", "trait")
#   )%>% # End of R2
#   filter(trait != 'LES') %>% 
#   filter(trait != 'WES') %>%
#   # bold() %>% 
# Coefficients are unstable depending on random GWM so I entered the one from the last online version
data.frame(
  Guild = c(rep("C", 5), rep("U", 5)), Trait = rep(c("SLA", "WD", "LDMC", "LA", "LT"), 2),
  Slope = c(-9.5*10^-3, "", "", "", 1.3*10^-2, "", 1.3*10^-2, "", "", ""),
  Curvature = c(-2, 3.8, -3.2, -6, 3, -1.3, 6, 2.9, "", 6.6),
  Wetness = c(-0.16, "",  -0.28, "", "", -0.22, "", "", 0.32, 0.19),
  SW = c("", 0.23, "", -0.19, "", "", "", "", "", ""),
  Canopy = c("", "", "", 1.8*10^-2, "", 2.6*10^-2, -1.5*10^-2, -9*10^-3, "", -1.6*10^-2),
  R2 =c(0.06, 0.11, 0.04, 0.18, 0.05, 0.5, 0.32, 0.06, 0.04, 0.19)
) %>% 
  rename("Adjusted $R^2$" = R2) %>% 
  kable(caption = 'Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy). Linear model coefficients are reported only for predictors retained after bidirectional stepwise selection, with corresponding adjusted $R^2$ . See Table \\@ref(tab:traits) for trait abbreviations.', format = 'pandoc')
```

\newpage

<!-- figure 1 -->

```{r Uppangala, fig.cap="Digital canopy model of the 10-ha forest permanent plot in Uppangala, India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
DCM <- raster("./data/Canopy.tif")
crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
DEM <- raster("./data/DEM.tif")
crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
limits <- shapefile("./data/Plots/Shapeall10ha.shp")
limits$melt <- 1
limits <- maptools::unionSpatialPolygons(limits, limits$melt)
india <- shapefile("./data/India_SHP/INDIA.shp")
DCM <- crop(DCM, buffer(limits, 100))
DEM <- crop(DEM, buffer(limits, 100))
DEM <- as(DEM, "SpatialPixelsDataFrame")
DEM <- as.data.frame(DEM)
colnames(DEM) <- c("Elevation", "x", "y")
DCM <- as(DCM, "SpatialPixelsDataFrame")
DCM <- as.data.frame(DCM)
colnames(DCM) <- c("Height", "x", "y")
ggplot() +  
geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
  geom_contour(data = DEM, aes(x = x, y = y, z = Elevation),
               colour = "black", alpha = 0.5, breaks = seq(300, 500, by = 25)) +
  directlabels::geom_dl(data = DEM, aes(x = x, y = y, z = Elevation, label = paste0(..level.., "m")),
                        method = list('last.bumpup', cex = 0.8, hjust = 1, fontface = 2),
                        stat = "contour", breaks = seq(300, 500, by = 25)) +
  geom_polygon(data = fortify(limits),
               aes(long, lat, group=group),
               colour = "black", lwd = 1.1, alpha = 0) +
  coord_equal() +
  cowplot::theme_map() +
  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
  ggsn::scalebar(dist = 50, dist_unit = "m", x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 20, y.max = max(DCM$y),
                 transform = F, model = "WGS84", location = "bottomleft") +
    annotation_custom(grob = ggplotGrob(ggplot() +  
                                        geom_polygon(data = fortify(india),
                                                     aes(long, lat, group=group),
                                                     colour = "black", alpha = 0) +
                                        geom_point(data = data.frame(long = 75.662778, lat = 12.537500), 
                                                   aes(long, lat), col = "red", size = 3) +
                                        coord_equal() +
                                        cowplot::theme_map() +
                                        theme(panel.border = element_rect(color = "black",size = .5,fill = NA),
                                              plot.background	= element_rect(fill = "white"))), 
                    xmin = min(DCM$x), xmax = min(DCM$x) + 150, 
                    ymin = max(DCM$y) - 150, ymax = max(DCM$y))
  # ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
```

\newpage

<!-- figure 2 -->

```{r pca,  fig.cap='Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B). Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test. See Table  \\@ref(tab:traits) for trait abbreviations.'}
pca.plot <- autoplot(PFT.pca, 
         data = Q %>% 
           left_join(L %>% 
                       dplyr::select(-com) %>% 
                       summarise_all(sum) %>% 
                       t() %>% 
                       data.frame() %>% 
                       rename(abundance =  ".") %>% 
                       mutate(SpCode = row.names(.))) %>% 
           filter(!is.na(WES)), 
         colour = "strata",
         size = "abundance",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1)) +
  scale_color_discrete("Guild")
spectrum.boxplots <- pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>%
  ggplot(aes(strata, value, fill = strata)) +
  ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
  geom_boxplot() +
  facet_wrap(~spectrum, nrow = 2) +
  scale_fill_discrete("Guild")
cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
# ggsave(cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1), labels = c("A", "B")),
#        file = "PCA.png",
#        path = './data/',
#        width = 200, height = 150, unit = 'mm',
#        dpi = 300)
```

\newpage

<!-- figure 3 -->

```{r GWMboxplot, fig.cap="Guild weighted mean (GWM) of specific leaf area (SLA), wood density (WD), leaf dry matter content (LDMC), leaf area (LA) and leaf thickness (LT) for canopy (C) and understory (U) guilds. Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines). Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4}
L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  inner_join(Q, by = "SpCode") %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  select(-community, -LES, -WES) %>%
  # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
  # select(-community) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free", nrow = 1) +
  # facet_wrap(~trait, scales = "free", nrow = 2) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3))))
  # ggsave(file = "boxplot.png",
  #        path = './data/',
  #        width = 200, height = 150, unit = 'mm',
  #        dpi = 300)
```

### 2020-01-14 Manuscript.Rmd
Uppangala 0.1.9056 1.20 Francois review

@@
 ---
-title: Environmental filtering favor different functional composition of tree assemblages between canopy and understory guilds within Indian forest communities
+title: Environment filter different tree functional composition between canopy and understory guilds in India / Different environment filters shape functional composition of canopy and understory guilds in Indian rainforest
 date: '`r Sys.Date()`'
@@
     reference_docx: ./template/template.docx
-  bookdown::html_document2:
-    number_sections: false  
-    toc: true
-    toc_float: yes
   bookdown::pdf_document2:
@@
     toc: false
+  bookdown::html_document2:
+    number_sections: false  
+    toc: true
+    toc_float: yes
 linestretch: 1.5
@@
     - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
-1. Maxime Réjou-Méchain
-    - maxime.rejou@gmail.com
-    - orcid.org/0000-0003-2824-267X
-    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
-    - UMR AMAP, IRD, F-34000 Montpellier, France
 1. Valérie Raevel
@@
     - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
+1. Maxime Réjou-Méchain
+    - maxime.rejou@ird.fr
+    - orcid.org/0000-0003-2824-267X
+    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
+    - AMAP, IRD, CNRS, CIRAD, INRA, Univ Montpellier, Montpellier, France
 1. Narayanan Ayyappan
@@
     - fmunoz@univ-grenoble-alpes.fr
+    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
     - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France
@@
 
-1.
-Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities.
-Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition.
-Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests, or focusing on canopy trees.
-1.
-We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India.
-Combining tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits, we used multivariate approaches and linear regressions to assess the variation of functional assemblages within tropical tree communities between understory and canopy species, and to address the effects of abiotic and biotic environments on these variations.
-1.
-We found that decoupled wood and leaf economics spectra were present in Indian rainforests and were maintained at local scale.
-Moreover, we showed that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities.
-We further showed that the canopy and understory guilds assemblages responded differently across communities to environmental gradients related to soil resource, light availabilities and monsoon disturbances.
-1.
-*Synthesis.* These results illustrate how fine-scale environmental heterogeneity may shape the tree community composition and function.
-Moreover, our study show guild-specific answer to biotic and abiotic environmental filtering between canopy and understory, highlighting the importance of forest vertical stratification in community assembly processes.
+> The abstract is just a first try, maybe ignore it for the moment until the introduction and the discussion is written.
+
+1. Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities. 
+<!-- Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition.
+-->
+Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests.
+We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity, depending on resource availability, hydraulic and biomechanic constraints.
+
+1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. 
+The plot was subdivided into tree communities, each composed of guilds of canopy and understory tree.
+We combined tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits to assess the variation of functional composition of between understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.
+
+1. The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra, even at local scale. 
+Although functional traits did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities.
+Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
+
+1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
+Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.
 
@@
 
-<!-- hook topo + vertical str -->
+Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
+Environmental variation influences tree community composition at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
+Specifically, topography can filter different ecological strategies [@kraft_functional_2008] due to variation in water and nutrient availability in tropical forests [@ferry2010higher; @Blanchard2019] .
+<!-- + Blanchard, Grégoire, François Munoz, Thomas Ibanez, Vanessa Hequet, Hervé Vandrot, Jérémy Girardi, et Philippe Birnbaum.
+2019.
+« Regional rainfall and local topography jointly drive tree community assembly in lowland tropical forests of New Caledonia ».
+Journal of Vegetation Science 0 (ja).
+https://doi.org/10.1111/jvs.12781.
+-->
+At higher topographic positions, e.g.
+hilltops, decreased water and nutrient availability  [@ferry2010higher] is expected to filter species with more conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review].
 
-Understanding the drivers of hyperdiverse tropical forest tree composition is a long-standing issue in ecology [@Cody1976].
-Abiotic environment has been shown to influence tree community composition at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006].
-Specifically, topography can filter different ecological strategies [@kraft_functional_2008] due to variation in water and nutrient availability in tropical forests [@ferry2010higher].
-But biotic interactions can also have a major impact on tree community composition [@Uriarte2004].
-Indeed, light availability, and thus community vertical structure, represent a major constraint for tree survival and growth in tropical forests [@Stark2015].
-Joint study of topography and community vertical structure therefore hold promise  to understand the drivers of hyperdiverse tropical communities [@Jucker2018].
+In addition, light availability decreases downward across vertical forest strata, while hydraulic and mechanical constraints increase upward, so that different survival and growth strategies should be filtered across strata [@Stark2015; @Uriarte2004a].
+Tree species can occupy different vertical strata at adult stage, allowing to classify them into vertical guilds.
+Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
+While the influence of ground-level variation in environmental conditions (soil, topography) on community assembly is now widely recognized, environmental filtering across vertical stratification is little addressed in tropical forests [@wright2002plant].
 
-<!-- topo and vert strata, what do we know ? -->
-
-Higher topographic positions, e.g.
-hilltops, result in decreased water and nutrient availability [@ferry2010higher] filtering for species with conservative strategy of resource use [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017].
-In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013].
+<!-- In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013].
 Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
 Therefore, understory and canopy guilds can differently respond to environmental gradients.
-Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests [@wright2002plant], or focusing on canopy trees [@Hubbell2001].
+Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests [@wright2002plant].
 Whereas, investigating community assembly among vertical strata may reveal distinct response to environmental gradients [@Raevel2018], and thus provide a better understanding of community assembly processes.
-
-<!-- FT for community assembly -->
+-->
 
@@
 Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
-The leaf economics spectrum opposes acquisitive ecological strategies, with high photosynthetic carbon assimilation, to conservative ones with high investment in leaf defense and durability.
-For instance, specific leaf area (leaf mass divided by leaf area) varies along soil and light availability gradients, and represents a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016].
-The wood economics spectrum opposes slow and fast growing species.
-However, @reich_world-wide_2014 advocate for a unique plant economics spectrum between fast and slow growing species, evidenced in a subarctic flora [@freschet_evidence_2010].
-Functional traits have been used to evidenced community assembly processes at local scale [@fortunel_leaf_2012; @Paine2011; but see @Messier2017], but to our knowledge little is known of the role of tropical forest vertical stratification in community trait-based studies.
+The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016].
+Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients.
+The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species.
+Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012].
+Nevertheless, @reich_world-wide_2014 proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].
 
-<!-- what have we done ? -->
+Here we decompose a tree community in spatially circumscribed at ground level into vertical guilds.
+Because they experience different environmental constraints, vertical guilds should show different functional responses to environmental variation between communities [@Raevel2018].
+In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger response of canopy trees to water limitation on hilltops (along the wood economics spectrum), while limited light availability should favor conservative strategies in understorey trees (along the leaf economics spectrum).
+Thus we expected different trait-environment relationships across vertical guilds, depending on the tree species strategies along the fundamental leaf and wood ecnonomics spectra.
+Previous works have shown different response of vertical strata to environmental variation at ground level, in Columbia [@duque2002different] and central Africa [@Ouedraogo2013], but evidences are lacking for Asian tropical forests.
 
-In the present study, we assessed the variation of functional assemblages within tropical tree communities between understory and canopy species, and addressed the effects of abiotic and biotic environments on these variations.
-We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000].
-The plot encompasses a large variation of topography [@Antin2013] and vertical structure [@Pascal1988].
-Combining tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits, we used multivariate approaches and linear regressions to address the following questions: (i) Do the trait variations among species follow independent leaf and wood economics spectra at local scale ?
+We assessed the variation in functional composition of understorey and canoy guilds within tropical tree communities, and examined the effects of abiotic and biotic environments on these variations.
+We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000].
+The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
+Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, we expected hydraulic and mechanic constraints to play a role.
+
+We used tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits, in multivariate approaches and linear regressions to address the following questions: (i) Do the mean trait variations among tree species follow independent leaf and wood economics spectra at local scale ?
 If so, how are canopy and understory guilds distributed along leaf and wood economics spectra ?
-(ii) Does functional composition differ between canopy and understory guilds ?
-and (iii) Does the guild functional composition vary with local environmental heterogeneity, and do the understory and canopy guilds respond differently ?
-Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017].
+(ii) Does functional composition differ between assemblages of canopy and understory guilds ?
+and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently ?
+
+<!-- Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017], with differentiated trait spaces for canopy and understory species.
 We hypothesized functional composition to differ between canopy and understorey species, due to their differences in growth and functional strategies [@Herault2011].
 Because this region experiences a long dry season period and is under the influence of intense monsoon rainy seasons with strong wind disturbances, we expected that the local topography is a major driver of tree community functional composition.
+-->
 
@@
 The plot is named after the nearest village, Uppangala, and is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India).
-Annual precipitation averages 5,100 mm per year but is highly variable throughout the year with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011].
+Annual precipitation averages 5,100 mm per year but is highly variable across seasons with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011].
 Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [@Pascal1996].
-The geology can be considered as homogeneous within the Uppangala plot [@Ferry1994].
-The topography of the site is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig.
+The geologic substrate can be considered homogeneous within the Uppangala plot [@Ferry1994], but the topography is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig.
 \@ref(fig:Uppangala)).
-Topography is thus expected to represent a major component of environmental variation within the forest plot [@Antin2013; @gimaret-carpentier_sampling_1998].
+Topography is thus expected to be a major component of environmental variation within the forest plot [@Antin2013; @gimaret-carpentier_sampling_1998].
 
@@
 
-Since 1989, all trees $\geq$ 10 cm in diameter at breast height (dbh) were identified to species level, tagged, mapped and monitored in an area progressively extended to a contiguous 10-ha plot in 2014 following international standards [@Jairazbhoy2006].
-The plot includes 6672 tree individuals belonging to 111 tree species mostly from the Dipterocarpaceae, Myristicaceae and Fabaceae families.
+All trees $\geq$ 10 cm in diameter at breast height (dbh) have been identified to species level, tagged, mapped and monitored since 1989 following international standards [@Jairazbhoy2006], in an area progressively extended to a contiguous 10-ha plot in 2014.
+The plot includes 6672 tree individuals belonging to 111 species mostly belonging to the Dipterocarpaceae, Myristicaceae and Fabaceae families.
 
-Using the classification of @Pascal1988, we assigned a potential vertical stratum of tree species (vertical stratum reached by most adult trees) to all species from the plot.
-@Pascal1988’s classification consisted of four tree species categories: emergent species above 39m, canopy species between 25 and 39m, sub-canopy species between 15 and 25m, and understory species below 15m.
-Because our study site is characterized by a steep topography, sub-canopy species often reach the canopy in the Uppangala plot.
-We thus defined two vertical strata, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.
-This resulted in 54 and 35 species belonging to the canopy and understory guilds, respectively.
+Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species (vertical stratum reached by most adult trees, with four categories: emergent above 39m, canopy between 25 and 39m, sub-canopy between 15 and 25m, and understory below 15m.
+Because our study site is characterized by a steep topography, sub-canopy species can often reach the canopy in the Uppangala plot.
+We thereby defined two broader categories, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.
 
-We collected trait data from 89 over the 111 (80.1%) species, representing 99.0% of tree individuals in the plot.
+We collected trait data for 89 over the 111 (80.1%) species, representing 99.0% of tree individuals in the plot.
 We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species, depending on leaf accessibility and species abundance (3 individuals on average).
@@
 For 11 species (1% of individuals in Uppangala plot), branch samples were too small (1 cm in diameter) to allow measuring branch wood density.
+Among the selected species, 54 and 35 belonged to the canopy and understory guilds, respectively.
 
@@
 
-We used LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy conditions.
-The LiDAR campaign was conducted in 2013, hence concomitantly with field measurements.
-Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, which relates to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the effect of dominant monsoon winds [@Bunyan2015]; and (iv) slope, which may have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher].
-In addition, using a 1-m resolution LiDAR canopy model (Fig.
-\@ref(fig:Uppangala)), we used the mean of the top of canopy height to  represent the intensity of local competition for light.
+We used LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy structure.
+The LiDAR campaign was conducted in 2013, consistent with field measurements.
+Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, related to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the exposure to dominant monsoon winds [@Bunyan2015]; and (iv) slope, which could have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher].
+In addition, based on a 1-m resolution LiDAR model of canopy (Fig.
+\@ref(fig:Uppangala)), we calculated the top canopy height, as a proxy of decreasing light availability downward and thus of local competition for light.
 
@@
 To assess the degree of trait variation and covariation we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific).
-We then performed a principal component analysis (PCA) of the average trait values per species to characterize the main dimensions of ecological strategies globally and per vertical stratum.
+We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.
 
-We then analyzed the functional composition of vertical stratum communities in $30m*30m$ subplots (n=110).
+We subsampled tree communities in $30m*30m$ subplots (n=110), and characterized the functional composition of vertical guilds within each subplot.
 Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.
 
-We calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’, [@Raevel2018]) of each trait in each subplot:
+We calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’, [@Raevel2018]) for each trait in each subplot:
 
 $$GWM_j=\sum_{i=1}^Sp_{i,j}*t_i$$
-where $p_{i,j}$ is the relative abundance of species $i$ in subplot $j$, $t_i$ the mean trait value of species $i$ and $S$ the number of species in the guild considered.
+where $p_{i,j}$ was the relative abundance of species $i$ in subplot $j$, $t_i$ the mean trait value of species $i$ and $S$ the number of species in the guild considered.
 
-To test whether the local functional composition of guilds depends on the local topography and canopy structure, we devised a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild, [@bernard-verdier_community_2012]).
-This null model broke the link between environment and local functional composition while maintaining the spatial structure of guilds and species abundances.
-Null GWM values generated through 999 randomizations were subsequently used to calculate the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as
+We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild, [@bernard-verdier_community_2012]).
+The null model broke the link between environment and local functional composition while maintaining the spatial distribution of guilds and species abundances.
+We derived 999 randomized GWM values and calculated the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as
 
 $$SES = \frac{GWM_{obs}-\overline{GWM_null}}{\sigma_{null}}$$
-where $GWM_{obs}$ is the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ are the mean and standard deviation of the corresponding null distribution, respectively.
+where $GWM_{obs}$ was the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ were the mean and standard deviation of the null distribution, respectively.
 
-Then we assessed how the deviation of guild composition from the null model depended on environmental variables using linear regressions with bidirectional stepwise selection of SES values.
-Thus, the topographical or canopy structure variables selected with this approach were interpreted as more impacting GWM values than expected by chance.
+We assessed how the deviation of guild composition from the null model depended on environmental variables by performing linear regressions and a bidirectional stepwise selection of predictors.
 These analyses were performed independently for each of the five functional traits.
@@
 
-We found an overall increase of functional trait variation from intra-individual to inter-specific levels (Table \@ref(tab:traits)), even though LA showed almost as large variability at intra-individual than at intra-specific level.
-Thus, trait values averaged at species level grasp most of the functional variability in the species pool considered here.
+We found an overall increase of functional trait variation from intra-individual to inter-specific levels (Table \@ref(tab:traits)), even though LA showed almost as much variability at intra-individual and at intra-specific levels.
+Thus, trait values averaged at species level could grasp most of the functional variability across guilds and communities with varying taxonomic composition.
 
 The first axis of the PCA performed on species trait values (Fig.
-\@ref(fig:pca)A) was basically related to wood density and structural leaf trait  (LT, LA, LDMC), opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood.
+\@ref(fig:pca)A) was mostly related to wood density and structural leaf traits (LT, LA, LDMC), and opposed species with thick, big leaves and lightwood to species with dense leaves and hardwood.
 The second axis mostly opposed SLA to LDMC.
-WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that they provide independent information.
-Canopy and understory species did not show significant differentiation on the first and second PCA axis (Wilcoxon p=0.2 and p=0.3, respectively Fig.
+WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that WD and SLA provide independent information.
+Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p=0.2 and p=0.3, respectively Fig.
 \@ref(fig:pca)B).
@@
 
-At the community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig.
+At community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig.
 \@ref(fig:GWMboxplot)).
@@
 
-We detected a significant effect of slope, curvature and wetness on both canopy and understory guilds assemblages, of southwesterness on canopy guild assemblages and of canopy height on understory guild assemblages (Table \@ref(tab:LM)).
+We detected a significant effect of slope, curvature and wetness on both canopy and understory guilds assemblages, of southwesterness on canopy guilds assemblages and of canopy height on understory guild assemblages (Table \@ref(tab:LM)).
+
+> *Un peu court/abrupt. Préciser les directions des effets observés.* François
 
@@
 
-Assessing the variation of functional assemblages within tropical tree communities between understory and canopy species, we showed that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities.
+We found that the measured functional traits of trees defined two orthogonal axes reflecting the leaf and wood economic spectra of trees at the scale of a 10ha rainforest plot.
+While functional trait values did not differ globally between canopy and understory tree species, we found neat difference between the assemblages of canopy and understory species within subplots.
+Therefore local environmental filtering selected species with diverging functional trait values across vertical guilds within communities.
+In addition the canopy and understory guild assemblages responded differently to environmental variation across communities.
+Soil resource, light availability and monsoon disturbances influenced the functional responses.
+Although understory and canopy assemblages responded similarly to water and nutrient access, functional variation of canopy assemblages was mostly along the wood economics spectrum, in response to aspect and wind exposure, while the functional variation of understory assemblages was mostly along the leaf economics spectrum in response to canopy height and decreasing light availability.
+These results show that environmental conditions drives different functional responses of vertical guilds in a tropical rainforest.
+
+<!-- Assessing the variation of functional assemblages within tropical tree communities between understory and canopy species, we showed that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities.
 We further showed that the canopy and understory guilds assemblages responded differently across communities to environmental gradients related to soil resource, light availabilities and monsoon disturbances.
@@
 Moreover our study show guild-specific answer to biotic and abiotic environmental filtering between canopy and understory, highlighting the importance of forest vertical stratification in community assembly processes.
+-->
 
-## Decoupled wood and leaf economics spectra at local scale
+## Decoupling of wood and leaf economics spectra at local scale
 
-Despite a reduced number of functional traits, our study support the existence of decoupled wood and leaf economics spectra at local scale in Indian rainforests.
+Although based on a limited number of functional traits, our PCA analysis supported two orthogonal main dimensions (Fig. \@ref(fig:pca)). 
+<!-- Despite a reduced number of functional traits, our study support the existence of decoupled wood and leaf economics spectra at local scale in Indian rainforests.
 We found locally that species trait values varied along two orthogonal main dimensions (Fig.
 \@ref(fig:pca)).
-SLA and LDMC negative correlation have already been linked to the leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004], representing a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Baraloto2010; @Evans2001; @hodgson_is_2011].
-Whereas, wood density was opposed to leaf area and thickness along the wood economics spectrum [@chave_towards_2009], which contrasted species more tolerant to hydric stress, with denser wood, to more competitive species with thicker and larger leaves and higher efficiency in water conductivity [@Santiago2004a].
-To our knowledge, our study is the first to support the two economics spectra in Indian rainforests.
-In addition, @Baraloto2010 and @Fortunel2012a found a similar structuring of Neotropical trees along these two spectra at the regional scale in Amazonia.
-Thus, our results indicate that the existence of the two economics spectra can hold at a very local scale (Fig.
-\@ref(fig:pca)A), although @Messier2017 suggested that the spectrum would not hold at local scale.
-Finally, our study support the idea of decoupled economics spectra [@Baraloto2010] and not a whole plant economics spectrum [@reich_world-wide_2014].
+-->
+SLA and LDMC were related in opposing direction to a dimension representing a leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004], 
+<!-- representing a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Baraloto2010; @Evans2001; @hodgson_is_2011]. -->
+while, wood density was opposed to leaf area and thickness along a dimension reflecting a wood economics spectrum [@chave_towards_2009].
+<!-- , which contrasted species more tolerant to hydric stress, with denser wood, to more competitive species with thicker and larger leaves and higher efficiency in water conductivity [@Santiago2004a].
+-->
+<!-- To our knowledge, our study is the first to support the two economics spectra in Indian rainforests. -->
+@Baraloto2010 and @Fortunel2012a also found orthogonal dimensions reflecting wood and leaf economic spectra of Neotropical tree species at regional scale in Amazonia.
+In addition, @Messier2017 suggested that the spectra would not be observed among species co-occuring at a local scale.
+Here (i) we show that the orthogonal spectra do exist across species at the scale of a 10ha plot, and (ii) we provide first evidence of these spectra in Indian rainforests.
+Our results support that economics spectra are decoupled [@Baraloto2010] and not related to a single, whole-plant economics spectrum [@reich_world-wide_2014].
 
@@
 
-Our results illustrated that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities.
-Indeed, we found no significant differences between functional traits of species pools from canopy and understory along the two main economics spectra (Fig.
+Although the positions of canopy and understory species along the economics spectra did not differ globally (Fig.
+\@ref(fig:pca)B), we found significant trait differences between assemblages of canopy and understory species co-occuring in local 30x30m communities (Fig.
+\@ref(fig:GWMboxplot)).
+Therefore, species pools of canopy and understory species broadly overlapped in functional spaces, and spanned a broad range of ecological strategies, but niche differentiation occurred in local communities across canopy and understory assemblages.
+Understory assemblages had larger, thinner and lighter leaves, hence greater SLA, but denser wood, while canopy assemblages showed lighter wood but also smaller leaves.
+The differences did not support a mere differentiation of guild assemblages along economic spectra.
+Light wood suggested greater hydraulic conduction of tall trees [@Santiago2018], despite a conservative leaf strategy in response to (i) tree height increasing xylem water potential [@Koch2004], (ii) canopy light leading to a thicker mesophyll protecting tissues from excessive irradiance [@Chazdon1993], and (iii) a higher risk of herbivory and mechanical damage in canopy [@Sterck1992].
+Drought constraint could be less important for understory trees, allowing greater investment into resource acquisition by leaves to solve higher need for light in more shaded understory [@Chazdon1996], but limited light may result in slower growth [@Herault2011] favoring denser wood.
+These results support the hypothesis that trees experience guild-specific ecological constraints within communities [@duque2002different], leading to a contrasting functional composition between canopy and understory assemblages.
+However the functional variation within guilds was still broad enough to entail overlap of the functional spaces at the scale of a 10ha plot, insuring greater stability in space and time to maintain the guilds functions [@Yachi1999].
+
+<!-- Our results illustrated that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities.
+Contrarily to our expectations, we found no significant differences between functional traits of species pools from canopy and understory along the two main economics spectra (Fig.
 \@ref(fig:pca)B).
@@
 Thus, while the guilds were adapting their functional strategy to the environment within communities, they conserved broad ecological variation insuring greater stability in space and time to maintain the guilds functions [@Yachi1999].
+-->
 
@@
 
-Our results support the hypothesis formulated by @duque2002different that trees experience guild-specific ecological constraints, leading to contrasting functional compositions between canopy and understory assemblages.
-Indeed, the way assemblages responded to environmental variation changed among guilds (Table \@ref(tab:LM)).
-While both understory and canopy assemblages responded similarly to slope wetness, and curvature, canopy assemblages responded to southwesterness with stronger support of the wood economics spectrum (higher $R^2$ for WD and LA), and understory assemblages responded to canopy height with a strong support of the SLA representing the leaf economics spectrum ($R^2=0.5$ for SLA model in understory, but WD model have also a good $R^2$).
-The wood economics spectrum major role in the response to environment of canopy assemblages might be explained by the critical role of mechanical resistance, disturbance tolerance and water conductivity efficiency in the establishment and persistence of canopy trees [@Poorter2010b].
-Conversely, understory assemblages are subject to limited light availability due to interception by trees in canopy, resulting in a critical role of resource acquisition constraining functional composition of understory trees along the leaf economics spectrum.
-
-The influence of aspect, through southwesterness, on forest composition is little documented in tropical forests but was expected to be highly context dependent.
+In addition, the guild assemblages responded differently to environmental variation across subplots (Table \@ref(tab:LM)).
+Canopy assemblages respond to southwesterness in terms of wood economics (greater WD and smaller LA for higher southwesterness), while understory assemblages were insensitive to southwesterness.
+The influence of aspect on forest composition is little documented in tropical forests.
 The Uppangala plot shows large topographic heterogeneity, with steep slopes mostly organized along southwest-oriented crests (Fig.
@@
 @gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar context in Sri Lanka: they suggested that it can influence the local environment according to the exposure to monsoon winds.
-In addition, @Bunyan2015 suggested that greater occurrence of forest fragments on southwest orientation in a forest-grassland mosaic of Western Ghats could result from monsoon rains.
-In our context, monsoon winds are also south-west oriented, and are expected to be a major source of disturbance in forest [@soman_aspects_1990].
-The possible consequences are that (i) harder wood and smaller leaves are selected for mechanical resistance in more disturbed southwest oriented area, or (ii) topography entails turbulences on northeast oriented areas responsible of more disturbance and death events in the forest history resulting in lightwood filtering [@larjavaara_rethinking_2010].
-In both cases, canopy species tend to be more exposed to disturbance events once adult, which is consistent with the fact that aspect only influenced canopy guild functional assemblages.
+In addition, @Bunyan2015 suggested that greater occurrence of forest fragments on southwest orientation in a forest-grassland mosaic of Western Ghats could result from monsoon winds.
+Uppangala, south-west oriented slopes were in fact more exposed to monsoon winds, which could entail greater damages and disturbance [@soman_aspects_1990].
+The consequences could be that (i) mechanical resistance, disturbance tolerance and water conductivity efficiency could be selected in areas more exposed to monsoon wind damages [@Poorter2010b], or (ii) turbulences on northeast oriented areas could yield more disturbance and death events favoring pioneer, lightwood species [@larjavaara_rethinking_2010].
+In both cases, canopy species are more exposed to disturbance than understory trees, hence more prevalent filtering of strategies related to wood economics for canopy assemblages.
 
-In addition, higher canopy significantly impacted the functional composition of understory species assemblages.
-Taller canopies resulted in more acquisitive strategies with higher SLA and lower LDMC and LT, maybe due to an increased light availability of less covering taller canopy resulting in less limited light-conditions and decreased investment in leaf lifespan [@kitajima_tissue-level_2010].
-Putative increased light-conditions of taller canopies, may also have increased understory assemblages competitiveness resulting in lighter woods for relatively faster growth.
-These remains hypotheses to be thoroughly tested.
+Conversely, we found pervasive influence of canopy height on the functional composition of understory assemblages related to leaf economics, namely, more acquisitive strategies were found where the canopy was taller.
+Greater investment into photosynthetic activity, with higher SLA and lower LDMC and LT, was found in understory assemblages under taller canopy.
+A taller canopy could be related to greater vertical differentiation ensuring greater light availability and decreased investment in leaf lifespan in understory [@kitajima_tissue-level_2010], while lesser hydraulic constraints could entail a greater role of competition among understory trees, filtering lighter woods and relatively faster growth.
 
-Finally, variation in curvature, wetness and slope similarly affected canopy and understory functional assemblages (at the exception of LDMC).
+Nonetheless, curvature, wetness and slope similarly affected canopy and understory functional assemblages, except for LDMC.
 Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and retain more water, increasing soil fertility [@ferry2010higher], creating favorable conditions where species assemblages with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA) strategy will outcompete others.
-By contrast, concave, drier or sloppy areas will be dryer and less fertile favoring species assemblages with more conservative strategies and slower growth.
+By contrast, concave, drier or sloppy areas will be drier and less fertile favoring species assemblages with more conservative strategies and slower growth.
 
@@
 
-> *It does not really bring anything new, should I remove it ?*
+> *It does not really bring anything new, should I remove it ?*. Sylvain
+> *A voir selon les pratiques du journal.* François
 

### 2020-01-23 Manuscript.Rmd
Uppangala 0.1.9057 23.20 MRM & VR review

@@
 ---
-title: Environment filter different tree functional composition between canopy and understory guilds in India / Different environment filters shape functional composition of canopy and understory guilds in Indian rainforest
+title: Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest
 date: '`r Sys.Date()`'
@@
     - vraevel@gmail.com
+    - orcid.org/0000-0002-9631-114X
     - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
@@
     - fmunoz@univ-grenoble-alpes.fr
+    - orcid.org/0000-0001-8776-4705
     - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
@@
 
-> The abstract is just a first try, maybe ignore it for the moment until the introduction and the discussion is written.
-
-1. Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities. 
+1.
+Topographic heterogeneity determines changing soil conditions, and environment also changes across vertical strata in rainforests.
+However, the joint influence of these factors on tree functional composition remain little known.
+<!-- Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities.  -->
 <!-- Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition.
 -->
-Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests.
+<!-- Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests. -->
 We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity, depending on resource availability, hydraulic and biomechanic constraints.
-
 1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. 
-The plot was subdivided into tree communities, each composed of guilds of canopy and understory tree.
-We combined tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits to assess the variation of functional composition of between understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.
-
+The plot was subdivided into tree communities, each composed of guilds of canopy and understory trees.
+We combined tree census data, LiDAR-derived topographic and canopy metrics, and functional traits to assess the variation of functional composition in understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.
 1. The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra, even at local scale. 
-Although functional traits did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities.
+Although functional space did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities.
 Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
-
 1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
@@
 
+<!-- > To be checked at the end -->
 Rainforest; Western Ghats; Economics spectra; Environmental filtering, Vertical strata
@@
 Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
-Environmental variation influences tree community composition at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
-Specifically, topography can filter different ecological strategies [@kraft_functional_2008] due to variation in water and nutrient availability in tropical forests [@ferry2010higher; @Blanchard2019] .
-<!-- + Blanchard, Grégoire, François Munoz, Thomas Ibanez, Vanessa Hequet, Hervé Vandrot, Jérémy Girardi, et Philippe Birnbaum.
-2019.
-« Regional rainfall and local topography jointly drive tree community assembly in lowland tropical forests of New Caledonia ».
-Journal of Vegetation Science 0 (ja).
-https://doi.org/10.1111/jvs.12781.
+Environmental variation influences composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
+Specifically, topography is known to filter different ecological strategies [@kraft_functional_2008] due to its impact on water and nutrient availability [@ferry2010higher; @Blanchard2019].
+<!-- At higher topographic positions, e.g., hilltops, decreased water and nutrient availability  [@ferry2010higher] is expected to filter species with more conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review].
 -->
-At higher topographic positions, e.g.
-hilltops, decreased water and nutrient availability  [@ferry2010higher] is expected to filter species with more conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review].
+However, tree species can occupy different vertical strata at adult stage, and thus be classified into vertical guilds (understory and canopy), experimenting contrasted environmental conditions. 
+Indeed, light availability decreases downward across vertical forest strata, while hydraulic and mechanical constraints increase upward, so that different survival and growth strategies should be filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a].
+While the influence of ground-level variation in environmental conditions (soil, topography) on community assembly is now widely recognized, environmental filtering across vertical guilds is little addressed in tropical forests [@wright2002plant].
 
-In addition, light availability decreases downward across vertical forest strata, while hydraulic and mechanical constraints increase upward, so that different survival and growth strategies should be filtered across strata [@Stark2015; @Uriarte2004a].
-Tree species can occupy different vertical strata at adult stage, allowing to classify them into vertical guilds.
-Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
-While the influence of ground-level variation in environmental conditions (soil, topography) on community assembly is now widely recognized, environmental filtering across vertical stratification is little addressed in tropical forests [@wright2002plant].
-
-<!-- In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013].
-Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
-Therefore, understory and canopy guilds can differently respond to environmental gradients.
-Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests [@wright2002plant].
-Whereas, investigating community assembly among vertical strata may reveal distinct response to environmental gradients [@Raevel2018], and thus provide a better understanding of community assembly processes.
--->
+<!-- In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013]. -->
 
@@
 Functional traits have been defined as phenotypic traits that impact fitness indirectly through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007].
-Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
+Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
+However, functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
 The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016].
 Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients.
-The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species.
+The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009].
 Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012].
@@
 
-Here we decompose a tree community in spatially circumscribed at ground level into vertical guilds.
+<!-- Here we decompose a tree community in spatially circumscribed at ground level into vertical guilds.
 Because they experience different environmental constraints, vertical guilds should show different functional responses to environmental variation between communities [@Raevel2018].
-In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger response of canopy trees to water limitation on hilltops (along the wood economics spectrum), while limited light availability should favor conservative strategies in understorey trees (along the leaf economics spectrum).
-Thus we expected different trait-environment relationships across vertical guilds, depending on the tree species strategies along the fundamental leaf and wood ecnonomics spectra.
-Previous works have shown different response of vertical strata to environmental variation at ground level, in Columbia [@duque2002different] and central Africa [@Ouedraogo2013], but evidences are lacking for Asian tropical forests.
+-->
+In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review] of canopy trees to water and nutrient limitation on hilltops [@ferry2010higher],
+ <!-- In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger response of canopy trees to water limitation on hilltops (along the wood economics spectrum), -->
+while limited light availability should favor conservative strategies in understory trees (along the leaf economics spectrum).
+Thus we expected different trait-environment relationships across vertical guilds, depending on tree species strategies along the fundamental leaf and wood economics spectra.
+Previous works have shown different response of vertical strata to environmental variation at ground level, in Columbia [@duque2002different] and central Africa [@Ouedraogo2013].
 
-We assessed the variation in functional composition of understorey and canoy guilds within tropical tree communities, and examined the effects of abiotic and biotic environments on these variations.
+We assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, and examined the effects of abiotic and biotic environments on these variations.
 We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000].
 The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
-Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, we expected hydraulic and mechanic constraints to play a role.
+Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, we expected hydraulic and mechanic constraints to play an important role in tree community assembly.
 
-We used tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits, in multivariate approaches and linear regressions to address the following questions: (i) Do the mean trait variations among tree species follow independent leaf and wood economics spectra at local scale ?
-If so, how are canopy and understory guilds distributed along leaf and wood economics spectra ?
-(ii) Does functional composition differ between assemblages of canopy and understory guilds ?
-and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently ?
+We used tree census data, LiDAR-derived topographic and canopy metrics and measured functional traits, in multivariate approaches and linear regressions to address the following questions: (i) Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale?
+If so, how are canopy and understory guilds distributed along leaf and wood economics spectra?
+(ii) Does functional composition differ between assemblages of canopy and understory guilds?
+and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently?
 
 <!-- Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017], with differentiated trait spaces for canopy and understory species.
-We hypothesized functional composition to differ between canopy and understorey species, due to their differences in growth and functional strategies [@Herault2011].
+We hypothesized functional composition to differ between canopy and understory species, due to their differences in growth and functional strategies [@Herault2011].
 Because this region experiences a long dry season period and is under the influence of intense monsoon rainy seasons with strong wind disturbances, we expected that the local topography is a major driver of tree community functional composition.
@@
 
-The study was carried out in a nearly 10-ha (300m*330m) permanent plot established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996].
+The study was carried out in a nearly 10-ha (300 m*330 m) permanent plot established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996].
 The plot is named after the nearest village, Uppangala, and is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India).
-Annual precipitation averages 5,100 mm per year but is highly variable across seasons with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011].
+Annual precipitation averages 5,100 mm per year but is highly variable across seasons with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall of 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011].
 Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [@Pascal1996].
@@
 All trees $\geq$ 10 cm in diameter at breast height (dbh) have been identified to species level, tagged, mapped and monitored since 1989 following international standards [@Jairazbhoy2006], in an area progressively extended to a contiguous 10-ha plot in 2014.
-The plot includes 6672 tree individuals belonging to 111 species mostly belonging to the Dipterocarpaceae, Myristicaceae and Fabaceae families.
+The plot includes 6,672 tree individuals belonging to 111 species mostly belonging to the Dipterocarpaceae, Myristicaceae and Fabaceae families.
 
-Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species (vertical stratum reached by most adult trees, with four categories: emergent above 39m, canopy between 25 and 39m, sub-canopy between 15 and 25m, and understory below 15m.
+Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species (vertical stratum reached by most adult trees, with four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m.
 Because our study site is characterized by a steep topography, sub-canopy species can often reach the canopy in the Uppangala plot.
@@
 
-We collected trait data for 89 over the 111 (80.1%) species, representing 99.0% of tree individuals in the plot.
-We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species, depending on leaf accessibility and species abundance (3 individuals on average).
-For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter close to 1 cm, following recommendation of @Perez-Harguindeguy2013.
+We collected trait data for 89 over the 111 (80%) species, representing 99% of tree individuals in the plot (54 canopy and 35 understory species).
+We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species (3 individuals on average), depending on leaf accessibility and species abundance.
+For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following recommendation of @Perez-Harguindeguy2013.
 Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]).
-Leaf and branch samples were kept in humidified ziplock bags filled with enriched CO2 air in dark until measurement within the next 24 hours.
-For 11 species (1% of individuals in Uppangala plot), branch samples were too small (1 cm in diameter) to allow measuring branch wood density.
-Among the selected species, 54 and 35 belonged to the canopy and understory guilds, respectively.
+Leaf and branch samples were kept in humidified ziplock bags filled with enriched $CO_2$ air in dark until measurement within the next 24 hours.
+For 11 species (1% of individuals in Uppangala plot), branch samples were too small (< to 1 cm in diameter) to allow measuring branch wood density.
+<!-- Among the selected species, 54 and 35 belonged to the canopy and understory guilds, respectively. -->
 
@@
 Leaf petioles, rachis and petiolules were removed for all measurements.
-Leaf fresh weight was measured with an analytical balance (Wensar, KD-H600, 0.01g), leaf fresh thickness with a micrometer and leaf area by  analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012].
+Leaf fresh weight was measured with an analytical balance at a 0.01 g precision (Wensar, KD-H600), leaf fresh thickness with a micrometer and leaf area by  analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012].
 Leaves were then kept in herbarium and oven-dried for at least 72 hours at 60°C before measurement of dry weight.
@@
 Branch wood samples were immersed in water to evaluate their volume by water displacement.
-Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of [@Perez-Harguindeguy2013].
+Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of @Perez-Harguindeguy2013.
 
@@
 
-We used LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy structure.
+We used airborne LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy structure.
 The LiDAR campaign was conducted in 2013, consistent with field measurements.
@@
 
-To assess the degree of trait variation and covariation we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific).
+To assess the degree of trait variation and covariation, we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific).
 We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.
 
-We subsampled tree communities in $30m*30m$ subplots (n=110), and characterized the functional composition of vertical guilds within each subplot.
+We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot.
 Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.
 
-We calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’, [@Raevel2018]) for each trait in each subplot:
+We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:
 
@@
 
-We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild, [@bernard-verdier_community_2012]).
+We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild; [@bernard-verdier_community_2012]).
 The null model broke the link between environment and local functional composition while maintaining the spatial distribution of guilds and species abundances.
@@
 
-$$SES = \frac{GWM_{obs}-\overline{GWM_null}}{\sigma_{null}}$$
+$$SES = \frac{GWM_{obs}-\overline{GWM_{null}}}{\sigma_{null}}$$
 where $GWM_{obs}$ was the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ were the mean and standard deviation of the null distribution, respectively.
@@
 These analyses were performed independently for each of the five functional traits.
-All analyses were conducted using the R 3.3.0 software [@RCoreTeam2018].
+All analyses were conducted using the R 3.3.0 software [@RCoreTeam2020].
 
@@
 WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that WD and SLA provide independent information.
-Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p=0.2 and p=0.3, respectively Fig.
+Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p=0.2 and p=0.3 respectively, Fig.
 \@ref(fig:pca)B).
@@
 \@ref(fig:GWMboxplot)).
-GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, higher for LA and lower for LDMC and LT than understory tree assemblages.
+GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, lower for LA and higher for LDMC and LT than understory tree assemblages.
 In other words, understory guild assemblages had larger, thinner and lighter leaves with higher SLA and denser woods, compared to canopy guild assemblages.
 
-We detected a significant effect of slope, curvature and wetness on both canopy and understory guilds assemblages, of southwesterness on canopy guilds assemblages and of canopy height on understory guild assemblages (Table \@ref(tab:LM)).
-
-> *Un peu court/abrupt. Préciser les directions des effets observés.* François
+We detected a significant effect of slope, curvature and wetness on both canopy and understory guild assemblages, of southwesterness on canopy guild assemblages and of canopy height on understory guild assemblages (Table \@ref(tab:LM)).
+Wet and convex areas with more slope were associated to dense wood (WD) and thick (LT) leaves with high SLA in assemblages of both canopy and understory guilds. 
+Convex areas showed opposed effect on LDMC, and wet areas opposed effect on LA, between assemblages of canopy and understory guilds.
+Assemblages from canopy guilds under southwest orientation were associated to dense wood (WD) with small leaves (LA).
+Assemblages from understory guilds under short canopy were associated to dense wood (WD) with thick (LT) and dense (LDMC) leaves with low SLA.
 
@@
 
-We found that the measured functional traits of trees defined two orthogonal axes reflecting the leaf and wood economic spectra of trees at the scale of a 10ha rainforest plot.
+Few studies have investigated how the functional characteristics and composition of canopy and understorey guilds vary in tropical forests. 
+We here show that functional trait values do not differ globally between canopy and understory tree species but that contrasting environments between understory and canopy filter species non-randomly, resulting in a markedly different functional composition of canopy and understory guilds within subplots.
+We also showed that the guilds functional assemblages of canopy and understory  respond differently to environmental variation.
+Indeed, although the functional composition of understory and canopy guilds assemblages respond similarly to water and nutrient access, the functional composition of canopy guild communities strongly varies along the wood economics spectrum, in response to wind exposure, and the functional variation of understory assemblages mostly varies along the leaf economics spectrum, in response to  light availability.
+These results show that different environmental gradients simultaneously drives different functional responses among guilds in tropical rainforest.
+
+<!-- We found that the measured functional traits of trees defined two orthogonal axes reflecting the leaf and wood economic spectra of trees at the scale of a 10ha rainforest plot.
 While functional trait values did not differ globally between canopy and understory tree species, we found neat difference between the assemblages of canopy and understory species within subplots.
@@
 These results show that environmental conditions drives different functional responses of vertical guilds in a tropical rainforest.
+-->
 
@@
 
-Although based on a limited number of functional traits, our PCA analysis supported two orthogonal main dimensions (Fig. \@ref(fig:pca)). 
+Functional traits PCA analysis supported two orthogonal main dimensions (Fig. \@ref(fig:pca)). 
 <!-- Despite a reduced number of functional traits, our study support the existence of decoupled wood and leaf economics spectra at local scale in Indian rainforests.
@@
 -->
-SLA and LDMC were related in opposing direction to a dimension representing a leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004], 
+SLA and LDMC were opposed into a dimension representing the leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004], 
 <!-- representing a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Baraloto2010; @Evans2001; @hodgson_is_2011]. -->
-while, wood density was opposed to leaf area and thickness along a dimension reflecting a wood economics spectrum [@chave_towards_2009].
+while, wood density was opposed to leaf area and thickness along a dimension reflecting the wood economics spectrum [@chave_towards_2009].
 <!-- , which contrasted species more tolerant to hydric stress, with denser wood, to more competitive species with thicker and larger leaves and higher efficiency in water conductivity [@Santiago2004a].
@@
 @Baraloto2010 and @Fortunel2012a also found orthogonal dimensions reflecting wood and leaf economic spectra of Neotropical tree species at regional scale in Amazonia.
-In addition, @Messier2017 suggested that the spectra would not be observed among species co-occuring at a local scale.
-Here (i) we show that the orthogonal spectra do exist across species at the scale of a 10ha plot, and (ii) we provide first evidence of these spectra in Indian rainforests.
+In response to these works, @Messier2017 suggested that these two orthogonal spectra reflect large scale difference in tree functional composition and would not be observed among species co-occurring at a local scale.
+<!-- Here we show that the orthogonal spectra do exist across species at the scale of a 10ha plot. -->
+Here we captured  the two orthogonal spectra across species at the scale of a 10 ha plot despite a limited number of traits compared to previous studies.
+<!-- and (ii) we provide first evidence of these spectra in Indian rainforests. -->
 Our results support that economics spectra are decoupled [@Baraloto2010] and not related to a single, whole-plant economics spectrum [@reich_world-wide_2014].
@@
 Although the positions of canopy and understory species along the economics spectra did not differ globally (Fig.
-\@ref(fig:pca)B), we found significant trait differences between assemblages of canopy and understory species co-occuring in local 30x30m communities (Fig.
+\@ref(fig:pca)B), we found significant trait differences between assemblages of canopy and understory species co-occuring in local 30 x 30m communities (Fig.
 \@ref(fig:GWMboxplot)).
-Therefore, species pools of canopy and understory species broadly overlapped in functional spaces, and spanned a broad range of ecological strategies, but niche differentiation occurred in local communities across canopy and understory assemblages.
+Therefore, species pools of canopy and understory species broadly overlap in functional spaces, and both spanned a broad range of ecological strategies (Fig. \@ref(fig:pca)A).
+<!-- However, niche differentiation occurred in local communities across canopy and understory assemblages leading to contrasted functional compositions. -->
+However, environmental niches filtered species in local communities leading to contrasted functional compositions between canopy and understory guild.
 Understory assemblages had larger, thinner and lighter leaves, hence greater SLA, but denser wood, while canopy assemblages showed lighter wood but also smaller leaves. 
-The differences did not support a mere differentiation of guild assemblages along economic spectra.
-Light wood suggested greater hydraulic conduction of tall trees [@Santiago2018], despite a conservative leaf strategy in response to (i) tree height increasing xylem water potential [@Koch2004], (ii) canopy light leading to a thicker mesophyll protecting tissues from excessive irradiance [@Chazdon1993], and (iii) a higher risk of herbivory and mechanical damage in canopy [@Sterck1992].
-Drought constraint could be less important for understory trees, allowing greater investment into resource acquisition by leaves to solve higher need for light in more shaded understory [@Chazdon1996], but limited light may result in slower growth [@Herault2011] favoring denser wood.
-These results support the hypothesis that trees experience guild-specific ecological constraints within communities [@duque2002different], leading to a contrasting functional composition between canopy and understory assemblages.
-However the functional variation within guilds was still broad enough to entail overlap of the functional spaces at the scale of a 10ha plot, insuring greater stability in space and time to maintain the guilds functions [@Yachi1999].
-
-<!-- Our results illustrated that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities.
-Contrarily to our expectations, we found no significant differences between functional traits of species pools from canopy and understory along the two main economics spectra (Fig.
-\@ref(fig:pca)B).
-Thus, species pools from canopy and understory overlap in functional space and both span a broad range of ecological strategies.
-Contrastingly, there was a sharp difference between the average trait values of guild assemblages across subplots (Fig.
-\@ref(fig:GWMboxplot)).
-Understory assemblages had larger, thinner and lighter leaves, thus greater SLA, and denser wood, than canopy assemblages, reflecting that understorey assemblages tend to have a more conservative strategy while canopy assemblages have a more competitive and acquisitive strategy.
-In other words, while understory assemblages favor tolerance to hydric stress and leaf investment and durability, canopy assemblages favor efficiency in water conductivity and resources acquisition.
-The observed contrast between species pools and guilds assemblages within communities of canopy and understory was solely due to the environmental filtering selecting for more abundant species with different strategies, while the overall range of species strategies broadly overlapped between guilds.
-Thus, while the guilds were adapting their functional strategy to the environment within communities, they conserved broad ecological variation insuring greater stability in space and time to maintain the guilds functions [@Yachi1999].
+<!-- The differences did not support a mere differentiation of guild assemblages along economic spectra. -->
+Light wood suggested greater hydraulic conduction of tall trees [@Santiago2018].
+The conservative leaf strategy observed in canopy assemblage may respond to (i) tree height increasing leaf water stress due to gravity [@Koch2004], (ii) canopy light leading to a thicker mesophyll protecting tissues from excessive irradiance [@Chazdon1993], and (iii) a higher risk of herbivory and mechanical damage in canopy [@Sterck1992].
+Drought constraint could be less important for understory trees, allowing greater investment into resource acquisition by leaves to solve the higher need for light in more shaded understory [@Chazdon1996].
+However limited light may result in slower growth [@Herault2011] favoring denser wood. 
+These results support the hypothesis that trees experience guild-specific ecological constraints within communities [@duque2002different], leading to contrasting functional compositions of canopy and understory assemblages.
+<!-- However the functional variation within guilds was still broad enough to entail overlap of the functional spaces at the scale of a 10ha plot, insuring greater stability in space and time to maintain the guilds functions [@Yachi1999].
 -->
@@
 
-In addition, the guild assemblages responded differently to environmental variation across subplots (Table \@ref(tab:LM)).
-Canopy assemblages respond to southwesterness in terms of wood economics (greater WD and smaller LA for higher southwesterness), while understory assemblages were insensitive to southwesterness.
-The influence of aspect on forest composition is little documented in tropical forests.
-The Uppangala plot shows large topographic heterogeneity, with steep slopes mostly organized along southwest-oriented crests (Fig.
-\@ref(fig:Uppangala)).
-@gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar context in Sri Lanka: they suggested that it can influence the local environment according to the exposure to monsoon winds.
-In addition, @Bunyan2015 suggested that greater occurrence of forest fragments on southwest orientation in a forest-grassland mosaic of Western Ghats could result from monsoon winds.
-Uppangala, south-west oriented slopes were in fact more exposed to monsoon winds, which could entail greater damages and disturbance [@soman_aspects_1990].
-The consequences could be that (i) mechanical resistance, disturbance tolerance and water conductivity efficiency could be selected in areas more exposed to monsoon wind damages [@Poorter2010b], or (ii) turbulences on northeast oriented areas could yield more disturbance and death events favoring pioneer, lightwood species [@larjavaara_rethinking_2010].
-In both cases, canopy species are more exposed to disturbance than understory trees, hence more prevalent filtering of strategies related to wood economics for canopy assemblages.
+Our results show that the two guilds assemblages respond differently to environmental variation (Table \@ref(tab:LM)). 
+Contrary to understory assemblages, canopy assemblages respond to southwesterness, which face dominant monsoon winds in our study site, along the wood economics spectrum with a greater WD and smaller LA for higher southwesterness.
+The influence of aspect on tropical forest composition has been little documented.
+@gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar climatic and topographical context in Sri Lanka and also interpreted it as a consequence of a higher exposure to monsoon winds.
+In the Western Ghats of India, @Bunyan2015 found greater occurrence of forest fragments on southwest orientation in a forest-grassland mosaic, probably as a result from monsoon winds. 
+The Uppangala plot is characterized by a large topographic heterogeneity, with steep slopes mostly organized along southwest-oriented crests (Fig. \@ref(fig:Uppangala)).
+South-west oriented slopes are likely to be more exposed to monsoon winds, which could entail greater tree damages and forest disturbance [@soman_aspects_1990].
+The consequences could be that (i) tree strategies associated with mechanical resistance and disturbance tolerance could be selected in areas more exposed to monsoon wind damages [@Poorter2010b], or (ii) turbulences on northeast oriented areas could yield more disturbance and death events favoring pioneer, lightwood species [@larjavaara_rethinking_2010].
+In both cases, canopy species are more exposed to wind disturbances than understory trees and are thus expected to respond more to wind exposure.
 
-Conversely, we found pervasive influence of canopy height on the functional composition of understory assemblages related to leaf economics, namely, more acquisitive strategies were found where the canopy was taller.
-Greater investment into photosynthetic activity, with higher SLA and lower LDMC and LT, was found in understory assemblages under taller canopy.
-A taller canopy could be related to greater vertical differentiation ensuring greater light availability and decreased investment in leaf lifespan in understory [@kitajima_tissue-level_2010], while lesser hydraulic constraints could entail a greater role of competition among understory trees, filtering lighter woods and relatively faster growth.
+Conversely, we found pervasive influence of canopy height on the functional composition of understory assemblages related to leaf economics, with more acquisitive strategies under tall canopies.
+We indeed found a greater investment into photosynthetic activity, with higher SLA and lower LDMC and LT in understory assemblages under taller canopy. 
+Taller canopies tend to bear greater vertical differentiation ensuring greater light availability and decreased investment in leaf lifespan in understory [@kitajima_tissue-level_2010].
+Higher canopies filtered also smaller wood density for understory assemblages.
+Lesser hydraulic constraints  in understory guilds could entail a greater role of competition among the trees, filtering lighter woods and relatively faster growth.
 
 Nonetheless, curvature, wetness and slope similarly affected canopy and understory functional assemblages, except for LDMC.
-Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and retain more water, increasing soil fertility [@ferry2010higher], creating favorable conditions where species assemblages with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA) strategy will outcompete others.
+Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and retain more water, increasing soil fertility [@ferry2010higher].
+It thus create favorable conditions where species with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA) strategies outcompete others.
 By contrast, concave, drier or sloppy areas will be drier and less fertile favoring species assemblages with more conservative strategies and slower growth.
@@
 
-> *It does not really bring anything new, should I remove it ?*. Sylvain
-> *A voir selon les pratiques du journal.* François
-
-Our results show that a broad functional variation among species and assemblages in an Indian rainforest is structured along decoupled leaf and wood economic spectra.
-We showed a significant environmental determinism in the functional composition of tree communities but also stressed that this determinism is guild-specific.
+Our results show that functional variations among species and assemblages preferentially vary along decoupled leaf and wood economic spectra in an Indian rainforest.
+We found a significant environmental determinism in the functional composition of tree communities but also stressed that this determinism is guild-specific.
 Our results  thus contribute to the debate on the nature and role of economics spectra in the assembly of tropical forests [@Baraloto2010].
@@
 
-> *Who financed the project ? I think I'll acknowledge everybody that more or less helped on the field, and we should acknowledge someone for the plot extension right ? Anybody or anything else ?*
+We would like to thank Cedric Vega for extending Uppangala plot to 10 ha.
+We also want to acknowledge the data cleaning operated by Arianne Tanguy.
+And we are thankful to Claire Fortunel for her comments.
 
@@
 
-SS, MM, FM and VR conceived the study, designed the methodology and analysed model outputs; SS, FM, VR, NA, NB and NB conducted the field sampling; NA, NB and NB identified species; MM and CJ provided the environmental data; SS, MM, FM and VR led the writing of the manuscript.
+SS, MRM, FM and VR conceived the study, designed the methodology and analysed model outputs; SS, FM, VR, NA, NB and NB conducted the field sampling; NA, NB and NB identified species; MRM and CSJ provided the environmental data; SS, MRM, FM and VR led the writing of the manuscript.
 All the authors contributed to the drafts and gave their final approval for publication.
@@
 
-Functional traits data have been submitted to TRY initiative [@Kattge2011] under the name UppangalaTraits.
+Functional traits data have been submitted to TRY initiative [@Kattge2019] under the name UppangalaTraits.
 Environmental variables and spatial positions of individuals were extracted from the Uppangala Station database or **lidar data**, for which access is modulated by **to precise**.

### 2020-04-02 Manuscript.Rmd
20200402

@@
   ungroup() %>% 
-  select(-Sp, -LTD, -SHA)
-suppressWarnings(row.names(PFT) <- PFT$SpCode) # deprecated
+  select(-LTD, -SHA, -Sp)
 # PCA
@@
 # Matrix Q
-Q <- PFT %>% 
+PFT <- PFT %>% 
   mutate(SpCode = as.character(SpCode)) %>% 
-  left_join(data.frame(PFT.pca$scores) %>% 
-              mutate(SpCode = row.names(.)),
-            by = "SpCode") %>% 
-  rename(WES = Comp.1, LES = Comp.2) %>% 
   left_join(genSpecies(), by = "SpCode") %>% 
   mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
-  select(SLA, LDMC, WD, LA, LT, LES, WES, strata, SpCode) %>%
-  mutate(strata = ifelse(SpCode == "agin", "U", strata))
+  mutate(strata = ifelse(SpCode == "agin", "U", strata)) %>% 
+  dplyr::select(strata, SpCode, LA, WD, LT, LDMC, SLA) %>% 
+  reshape2::melt(id.vars = c("strata", "SpCode"), 
+                 variable.name = "trait", value.name = "traitvalue")
 # Matrix L
@@
 proj4string(XY) <- crs(com)
-L <- genTrees() %>% 
-  mutate(com = (XY %over% com)[,1]) %>% 
-  filter(!is.na(com)) %>% 
-  group_by(com, SpCode) %>% 
-  summarise(n = n()) %>% 
-  dcast(com ~ SpCode, value.var = "n") %>% 
-  replace(is.na(.), 0)
+trees <- genTrees() %>% 
+  mutate(community = (XY %over% com)[,1]) %>% 
+  filter(!is.na(community)) %>% 
+  group_by(community, SpCode) %>% 
+  summarise(Abundance = n()) %>% 
+  filter(community != "C99")
 # Matrix R
 load(file.path("draft_save", 'R.Rdata'))
-# GWMses
-GWM <- L %>%
-  melt(id.var = "com",
-       variable.name = "SpCode",
-       value.name = "abundance") %>% 
-  dcast(SpCode ~ com, value.var = "abundance") %>% 
-  mutate(SpCode = as.character(SpCode)) %>% 
-  left_join(Q, by = "SpCode") %>% 
+# GWM
+GWM <- trees %>% 
+  left_join(PFT) %>% 
   filter(!is.na(strata)) %>% 
-  select(-SpCode) %>% 
-  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-       variable.name = 'community', value.name = 'abundance') %>% 
-  group_by(strata, community) %>% 
-  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
-# nQ <- array(rep(as.matrix(Q), n),
-#             c(dim(Q), n),
-#             list(row.names(Q), names(Q), 1:n))
-# nQ <- array(apply(nQ, 3, function(x){
-#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
-#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
-#   return(x)
-#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
-# cl <- makeCluster(detectCores())
-# clusterExport(cl, list('L', 'nQ'))
-# nGWM <- parApply(cl, nQ, 3, function(x){
-#   library(dplyr)
-#   library(reshape2)
-#   L %>%
-#     melt(id.var = "com",
-#          variable.name = "SpCode",
-#          value.name = "abundance") %>%
-#     dcast(SpCode ~ com) %>%
-#     left_join(data.frame(x)) %>%
-#     select(-SpCode) %>%
-#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-#          variable.name = 'community', value.name = 'abundance') %>%
-#     group_by(strata, community) %>%
-#     mutate_each(funs(as.character), -abundance) %>%
-#     mutate_each(funs(as.numeric), -abundance) %>%
-#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
-# })
-# stopCluster(cl)
-# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
-#               c(dim(GWM[3:9]), n),
-#               list(row.names(GWM), names(GWM[3:9]), 1:n))
-# rm(cl, nQ)
-# save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
-load(file.path("draft_save", 'nGWM.Rdata'))
-GWM_ses <- GWM
-GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
+  group_by(strata, community, trait) %>% 
+  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
+# GWM plot
+left_join(GWM, R) %>% 
+  reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
+  filter(variable != "BA") %>% 
+  ggplot(aes(value, GWM, col = strata)) +
+  facet_grid(trait ~ variable, scales = "free") +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  theme(legend.position = "bottom", axis.title = element_blank())
+# nGWM
+nGWM <- lapply(1:n, function(n) 
+  left_join(trees, 
+            group_by(PFT, strata) %>% 
+              mutate(SpCode = sample(SpCode)),
+            by = "SpCode") %>% 
+    filter(!is.na(strata)) %>% 
+    group_by(strata, community, trait) %>% 
+    summarise(nGWM = weighted.mean(traitvalue, Abundance, na.rm = T)) %>% 
+    mutate(NullModel = n)) %>% 
+  bind_rows()
+# nGWM plot
+ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
+  geom_density(alpha = .5) +
+  facet_wrap(~ strata + trait, scales = "free") +
+  scale_color_discrete(guide = "none")
+# SES
+SES <- GWM %>% 
+  left_join(group_by(nGWM, strata, community, trait) %>% 
+              summarise(nGWMmean = mean(nGWM, na.rm = T),
+                        nGWMsd = sd(nGWM, na.rm = T))) %>% 
+  mutate(SES = (GWM - nGWMmean)/nGWMsd) %>% 
+  dplyr::select(-GWM, -nGWMmean, -nGWMsd) %>% 
+  left_join(R)
+# SES plot
+reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
+  filter(variable != "BA") %>% 
+  ggplot(aes(value, SES, col = strata)) +
+  facet_grid(trait ~ variable, scales = "free") +
+  geom_point() +
+  geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
+  geom_smooth(method = "lm") +
+  theme(legend.position = "bottom", axis.title = element_blank())
 ```
@@
 ```{r LM}
-# GWM_ses %>% 
-#   filter(!is.na(strata)) %>% 
-#   melt(id.vars = c('strata', 'community')) %>% 
-#   rename(trait = variable) %>% 
-#   rename(SES = value) %>% 
-#   mutate(community = as.character(community)) %>% 
-#   left_join(R, by = "community") %>%
-#   group_by(strata, trait) %>% 
-#   do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
-#      direction = 'both', trace = 0)) %>% 
-#   # do(reg = lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .),
-#   #    direction = 'both', trace = 0) %>%
-#   tidy(reg) %>%
-#   group_by(strata, trait, term) %>%
-#   # transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
-#                           # stars(p.value, ns = '   '))) %>% 
-#   transmute(value = format(estimate, scientific = T, digits = 2, trim = T)) %>% 
-#   dcast(strata + trait ~ term) %>% 
-#   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
-#   mutate_all(funs(replace(., is.na(.), ' '))) %>% 
-#   left_join( # Inclusion of R2 but should be split
-#     GWM_ses %>% 
-#       filter(!is.na(strata)) %>% 
-#       melt(id.vars = c('strata', 'community')) %>% 
-#       rename(trait = variable) %>% 
-#       rename(SES = value) %>% 
-#       mutate(communtiy = as.character(community)) %>% 
-#       left_join(R, by = "community") %>%
-#       group_by(strata, trait) %>% 
-#       do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
-#                     direction = 'both', trace = 0)) %>% 
-#       glance(reg) %>% 
-#       select(r.squared) %>% 
-#       mutate(r.squared = round(r.squared, 2)) %>% 
-#       rename("$R^2$" = r.squared), by = c("strata", "trait")
-#   )%>% # End of R2
-#   filter(trait != 'LES') %>% 
-#   filter(trait != 'WES') %>%
-#   # bold() %>% 
-# Coefficients are unstable depending on random GWM so I entered the one from the last online version
-data.frame(
-  Guild = c(rep("C", 5), rep("U", 5)), Trait = rep(c("SLA", "WD", "LDMC", "LA", "LT"), 2),
-  Slope = c(-9.5*10^-3, "", "", "", 1.3*10^-2, "", 1.3*10^-2, "", "", ""),
-  Curvature = c(-2, 3.8, -3.2, -6, 3, -1.3, 6, 2.9, "", 6.6),
-  Wetness = c(-0.16, "",  -0.28, "", "", -0.22, "", "", 0.32, 0.19),
-  SW = c("", 0.23, "", -0.19, "", "", "", "", "", ""),
-  Canopy = c("", "", "", 1.8*10^-2, "", 2.6*10^-2, -1.5*10^-2, -9*10^-3, "", -1.6*10^-2),
-  R2 =c(0.06, 0.11, 0.04, 0.18, 0.05, 0.5, 0.32, 0.06, 0.04, 0.19)
-) %>% 
-  rename("Adjusted $R^2$" = R2) %>% 
+group_by(SES, strata, trait) %>% 
+  mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), scale, center = F) %>%
+  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
+                direction = 'both', trace = 0)) %>%
+  tidy(reg) %>% 
+  filter(term != "(Intercept)") %>% 
+  mutate(estimate = round(estimate, 2)) %>% 
+  group_by(strata, trait, term) %>%
+  dcast(strata + trait ~ term, value.var = "estimate") %>%
+  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>%
+    mutate_all(funs(replace(., is.na(.), ' '))) %>%
+  left_join(group_by(SES, strata, trait) %>% 
+              mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), 
+                        scale, center = F) %>%
+              do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
+                            direction = 'both', trace = 0)) %>%
+              glance(reg) %>%
+              dplyr::select(r.squared) %>%
+              mutate(r.squared = round(r.squared, 2)), 
+            by = c("strata", "trait")) %>%
+            rename("$R^2$" = r.squared, Guild = strata, Trait = trait) %>%  
     kable(caption = 'Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy).

### 2020-04-02 Manuscript.Rmd
20200402 without extreme canopy removal

@@
 output:
-  bookdown::word_document2:
-    reference_docx: ./template/template.docx
+  bookdown::html_document2:
+    number_sections: false  
+    toc: true
+    toc_float: yes
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::html_document2:
-    number_sections: false  
-    toc: true
-    toc_float: yes
+  bookdown::word_document2:
+    reference_docx: ./template/template.docx
 linestretch: 1.5
@@
                fig.height = 8, fig.width = 8,
-               cache = T, cache.lazy = F)
+               cache = F, cache.lazy = F)
 cv <- function(x, na.rm = T)
@@
   group_by(community, SpCode) %>% 
-  summarise(Abundance = n()) %>% 
-  filter(community != "C99")
+  summarise(Abundance = n())
 # Matrix R
@@
   left_join(trees, 
-            group_by(PFT, strata) %>% 
+            group_by(PFT, trait, strata) %>% 
               mutate(SpCode = sample(SpCode)),
@@
 Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test.
-See Table  \\@ref(tab:traits) for trait abbreviations.'}
+See Table  \\@ref(tab:traits) for trait abbreviations.', eval=F}
 pca.plot <- autoplot(PFT.pca, 
@@
 cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
-# ggsave(cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1), labels = c("A", "B")),
-#        file = "PCA.png",
-#        path = './data/',
-#        width = 200, height = 150, unit = 'mm',
-#        dpi = 300)
 ```
@@
 Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines).
-Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4}
+Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4, eval=F}
 L %>%

### 2020-10-13 Manuscript.Rmd
Uppangala 0.1.9058 13.10.20 Reboot

@@
 ---
-title: Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest
+title: Topography and canopy height differently shape the functional composition of canopy and understory guilds in an Indian rainforest
 date: '`r Sys.Date()`'
 output:
-  bookdown::html_document2:
-    number_sections: false  
-    toc: true
-    toc_float: yes
+  bookdown::word_document2:
+    reference_docx: ./template/template.docx
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::word_document2:
-    reference_docx: ./template/template.docx
+  bookdown::html_document2:
+    number_sections: false  
+    toc: true
+    toc_float: yes
 linestretch: 1.5
@@
 library(ggfortify)
+library(rstan)
+options(mc.cores = 4)
+rstan_options(auto_write = T)
+library(bayesplot)
 theme_set(bayesplot::theme_default())
@@
                fig.height = 8, fig.width = 8,
-               cache = F, cache.lazy = F)
+               cache = T, cache.lazy = F)
 cv <- function(x, na.rm = T)
@@
 
-```{r data}
+```{r data, include = F}
 # Sp level PFT
@@
 # nGWM plot
-ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
-  geom_density(alpha = .5) +
-  facet_wrap(~ strata + trait, scales = "free") +
-  scale_color_discrete(guide = "none")
+# ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
+#   geom_density(alpha = .5) +
+#   facet_wrap(~ strata + trait, scales = "free") +
+#   scale_color_discrete(guide = "none")
 # SES
@@
 # SES plot
-reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
-  filter(variable != "BA") %>% 
-  ggplot(aes(value, SES, col = strata)) +
-  facet_grid(trait ~ variable, scales = "free") +
+# reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
+#   filter(variable != "BA") %>% 
+#   ggplot(aes(value, SES, col = strata)) +
+#   facet_grid(trait ~ variable, scales = "free") +
+#   geom_point() +
+#   geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
+#   geom_smooth(method = "lm") +
+#   theme(legend.position = "bottom", axis.title = element_blank())
+```
+
+```{r, eval=F}
+filter(trees, community %in% filter(R, Canopy < 20, Slope > 30)$community) %>% 
+  group_by(community, SpCode) %>% 
+  summarise(Abundance = sum(Abundance)) %>% 
+  left_join(filter(PFT, trait == "WD")) %>% 
+  filter(strata == "U") %>% 
+  arrange(community, desc(traitvalue)) %>% 
+  kable(caption = "Community with low canopy (<20m) and strong slope (>30)")
+filter(trees, community %in% filter(R, Slope < 15)$community) %>% 
+  group_by(community, SpCode) %>% 
+  summarise(Abundance = sum(Abundance)) %>% 
+  left_join(filter(PFT, trait == "WD")) %>% 
+  filter(strata == "U") %>% 
+  arrange(community, desc(traitvalue)) %>% 
+  kable(caption = "Community with low slope (<15)")
+filter(PFT, strata == "U", trait == "WD") %>% 
+  ggplot(aes(traitvalue)) +
+  geom_histogram() +
+  ggtitle("WD distribution among understory species.")
+left_join(trees, PFT) %>% 
+  filter(!is.na(strata)) %>% 
+  group_by(strata, community) %>% 
+  summarise(Abundance = sum(Abundance)) %>% 
+  left_join(R) %>% 
+  ggplot(aes(Slope, Abundance)) +
   geom_point() +
-  geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
-  geom_smooth(method = "lm") +
-  theme(legend.position = "bottom", axis.title = element_blank())
+  facet_wrap(~ strata, scales = "free") +
+  ggtitle("Number of trees per community along slope gradient.")
+left_join(trees, filter(PFT, trait == "WD")) %>% 
+  filter(strata == "U") %>% 
+  filter(!is.na(traitvalue)) %>% 
+  mutate(WDclass = cut(traitvalue, 
+                       breaks = c(0, 0.5, 0.7, 1), 
+                       labels = c("low", "mid", "high"))) %>% 
+  group_by(WDclass, community) %>% 
+  summarise(Abundance = sum(Abundance)) %>% 
+  left_join(R) %>% 
+  ggplot(aes(Slope, Abundance)) +
+  geom_point() +
+  facet_wrap(~ WDclass, scales = "free", nrow = 2) +
+  ggtitle("Number of trees per community along slope gradient",
+          "depending on WD category (low < 0.5 and high > 0.7).")
 ```
@@
 1. Chandra Shekhar Jha
-    - email
-    - Forestry and Ecology Group, National Remote Sensing Centre, Hyderabad 500 037, India
+    - chandra.s.jha@gmail.com
+    - National Remote Sensing Centre, Indian Space Research Organisation, Hyderabad 500 037, India 
 1. François Munoz
@@
 
+# Changes
+
+## 13 octobre 2020
+
+* I changed the title.
+* I removed the bidirectional stepwise selection of predictors as it was unstable.
+* I used bayesian inference for the linear regression (minor).
+* I replaced table 2 by a figure of posteriors distribution and significance (seemed clearer).
+* I removed the ideas of LES and WES between strata.
+That's unclear and difficult to defend.
+WD model in understorey showed a $R^2=0.3$, which is huge and removed the idea that the WES mainly affects canopy.
+Similarly, SLA and LDMC models in canopy had low $R^2$ but showed significant effects of wetness, curvature and slope.
+* I kept Maxime’s hypothesis for canopy height effect in understory.
+I don’t have a better hypothesis.
+I am therefore to keep it calling for further testing.
+**Maybe we should remove canopy height to simplify the message?**
+* I am not a fan of the conclusion. I kept it for the moment but I didn’t feel it brought much more than the take home message; So I’m hesitating to simply remove it.
+
 # Abstract
@@
 1.
-Topographic heterogeneity determines changing soil conditions, and environment also changes across vertical strata in rainforests.
-However, the joint influence of these factors on tree functional composition remain little known.
-<!-- Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities.  -->
-<!-- Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition.
--->
-<!-- Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests. -->
-We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity, depending on resource availability, hydraulic and biomechanic constraints.
-1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. 
+Topographic heterogeneity determines changing soil conditions, while the height of the canopy modulates the microclimate in the forest.
+But the environmental conditions encountered by trees also change across the vertical strata of tropical forests.
+However, the joint influence of topography and canopy height with vertical strata on tree functional composition remain little known.
+We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity and canopy height, depending on resource availability, hydraulic and biomechanical constraints.
+1.
+We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India.
 The plot was subdivided into tree communities, each composed of guilds of canopy and understory trees.
 We combined tree census data, LiDAR-derived topographic and canopy metrics, and functional traits to assess the variation of functional composition in understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.
-1. The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra, even at local scale. 
+1.
+The variation of functional traits across sampled tree species followed locally decoupled wood and leaf economics spectra.
 Although functional space did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities.
 Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
-1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
+1.
+*Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
 Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.
@@
 
-<!-- > To be checked at the end -->
 Rainforest; Western Ghats; Economics spectra; Environmental filtering, Vertical strata
@@
 Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
-Environmental variation influences composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
-Specifically, topography is known to filter different ecological strategies [@kraft_functional_2008] due to its impact on water and nutrient availability [@ferry2010higher; @Blanchard2019].
-<!-- At higher topographic positions, e.g., hilltops, decreased water and nutrient availability  [@ferry2010higher] is expected to filter species with more conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review].
--->
-However, tree species can occupy different vertical strata at adult stage, and thus be classified into vertical guilds (understory and canopy), experimenting contrasted environmental conditions. 
-Indeed, light availability decreases downward across vertical forest strata, while hydraulic and mechanical constraints increase upward, so that different survival and growth strategies should be filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a].
-While the influence of ground-level variation in environmental conditions (soil, topography) on community assembly is now widely recognized, environmental filtering across vertical guilds is little addressed in tropical forests [@wright2002plant].
+Environmental variation influences the composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003] 
+and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
+Topography is known to filter different ecological strategies [@kraft_functional_2008; @Schmitt2020]
+due to its impact on water and nutrient availability [@Blanchard2019; @ferry2010higher].
+In addition, canopy height modulates forest microclimates due to topography [@Jucker2018].
+But, tree species  occupy different vertical strata at adult stage, and can thus be classified into different vertical guilds (understory and canopy),
+experimenting contrasted environmental conditions. 
+Indeed, light availability decreases downward across vertical forest strata [@chazdon1984],
+while hydraulic and mechanical constraints increase upward [@Koch2004], 
+so that different survival and growth strategies are filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a]. 
+While functional strategies of tree species have been documented independently across topography, canopy height and vertical guilds,
+little is known on their interactions,
+particularly the effect of topography and canopy height on species assemblages between vertical guilds has not been addressed [@wright2002plant].
 
-<!-- In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013]. -->
-
-Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], contributing thus to test for community assembly processes.
-Functional traits have been defined as phenotypic traits that impact fitness indirectly through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007].
-Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
-However, functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
+Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], 
+contributing thus to test for community assembly processes. 
+Functional traits have been defined as phenotypic traits that impact fitness through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007].
+Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018
+along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
 The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016].
@@
 Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012]. 
-Nevertheless, @reich_world-wide_2014 proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].
+Nevertheless, @reich_world-wide_2014
+proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].
+The concept of plant economics spectrum may be highly relevant for understanding the variation of strategy between vertical guilds as understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
 
-<!-- Here we decompose a tree community in spatially circumscribed at ground level into vertical guilds.
-Because they experience different environmental constraints, vertical guilds should show different functional responses to environmental variation between communities [@Raevel2018].
--->
-In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review] of canopy trees to water and nutrient limitation on hilltops [@ferry2010higher],
- <!-- In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger response of canopy trees to water limitation on hilltops (along the wood economics spectrum), -->
-while limited light availability should favor conservative strategies in understory trees (along the leaf economics spectrum).
-Thus we expected different trait-environment relationships across vertical guilds, depending on tree species strategies along the fundamental leaf and wood economics spectra.
-Previous works have shown different response of vertical strata to environmental variation at ground level, in Columbia [@duque2002different] and central Africa [@Ouedraogo2013].
+In the context of seasonal rainforests, we hypothesized different trait-environment relationships across vertical guilds, 
+with varying strategies along the fundamental leaf and wood economics spectra. 
+More specifically, because understory trees are strongly light-limited and are less subject to hydraulic and mechanical constraints than canopy trees, 
+we expect that  understory and canopy species tend to maximize their difference along the leaf and wood economic spectra, respectively. 
+More generally, we expect that understory species and assemblages have more conservative strategies than canopy ones and that the lower water and nutrient availability found on hilltops [@ferry2010higher]
+entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; @Schmitt2020]. 
+Previous works have shown different response of vertical strata to environmental variation at ground level, with increased edaphic constraint on understory compared to canopy  for Columbian drained uplands [@duque2002different].
 
-We assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, and examined the effects of abiotic and biotic environments on these variations.
+Here, we assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, 
+and examined the effects of topography and canopy height on these variations. 
 We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000].
 The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
-Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, we expected hydraulic and mechanic constraints to play an important role in tree community assembly.
+Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, 
+we expected hydraulic and mechanical constraints to play an important role in tree community assembly.
 
-We used tree census data, LiDAR-derived topographic and canopy metrics and measured functional traits, in multivariate approaches and linear regressions to address the following questions: (i) Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale?
+We used tree census data, LiDAR-derived topographic and canopy metrics and measured functional traits, 
+in multivariate approaches and linear regressions to address the following questions: 
+(i) Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale? 
 If so, how are canopy and understory guilds distributed along leaf and wood economics spectra? 
@@
 and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently?
-
-<!-- Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017], with differentiated trait spaces for canopy and understory species.
-We hypothesized functional composition to differ between canopy and understory species, due to their differences in growth and functional strategies [@Herault2011].
-Because this region experiences a long dry season period and is under the influence of intense monsoon rainy seasons with strong wind disturbances, we expected that the local topography is a major driver of tree community functional composition.
--->
 
@@
 
+```{r Uppangala, fig.cap="Digital canopy model of the 10-ha forest permanent plot in Uppangala, India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
+DCM <- raster("./data/Canopy.tif")
+crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
+DEM <- raster("./data/DEM.tif")
+crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
+limits <- shapefile("./data/Plots/Shapeall10ha.shp")
+limits$melt <- 1
+limits <- maptools::unionSpatialPolygons(limits, limits$melt)
+india <- shapefile("./data/India_SHP/INDIA.shp")
+DCM <- crop(DCM, buffer(limits, 100))
+DEM <- crop(DEM, buffer(limits, 100))
+DEM <- as(DEM, "SpatialPixelsDataFrame")
+DEM <- as.data.frame(DEM)
+colnames(DEM) <- c("Elevation", "x", "y")
+DCM <- as(DCM, "SpatialPixelsDataFrame")
+DCM <- as.data.frame(DCM)
+colnames(DCM) <- c("Height", "x", "y")
+ggplot() +  
+geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
+  geom_contour(data = DEM, aes(x = x, y = y, z = Elevation),
+               colour = "black", alpha = 0.5, breaks = seq(300, 500, by = 25)) +
+  directlabels::geom_dl(data = DEM, aes(x = x, y = y, z = Elevation, label = paste0(..level.., "m")),
+                        method = list('last.bumpup', cex = 0.8, hjust = 1, fontface = 2),
+                        stat = "contour", breaks = seq(300, 500, by = 25)) +
+  geom_polygon(data = fortify(limits),
+               aes(long, lat, group=group),
+               colour = "black", lwd = 1.1, alpha = 0) +
+  coord_equal() +
+  cowplot::theme_map() +
+  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
+  ggsn::scalebar(dist = 50, dist_unit = "m", x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 20, y.max = max(DCM$y),
+                 transform = F, model = "WGS84", location = "bottomleft") +
+    annotation_custom(grob = ggplotGrob(ggplot() +  
+                                        geom_polygon(data = fortify(india),
+                                                     aes(long, lat, group=group),
+                                                     colour = "black", alpha = 0) +
+                                        geom_point(data = data.frame(long = 75.662778, lat = 12.537500), 
+                                                   aes(long, lat), col = "red", size = 3) +
+                                        coord_equal() +
+                                        cowplot::theme_map() +
+                                        theme(panel.border = element_rect(color = "black",size = .5,fill = NA),
+                                              plot.background	= element_rect(fill = "white"))), 
+                    xmin = min(DCM$x), xmax = min(DCM$x) + 150, 
+                    ymin = max(DCM$y) - 150, ymax = max(DCM$y))
+  # ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
+```
+
 ## Taxonomic composition and trait sampling
@@
 
-Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species (vertical stratum reached by most adult trees, with four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m.
-Because our study site is characterized by a steep topography, sub-canopy species can often reach the canopy in the Uppangala plot.
-We thereby defined two broader categories, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.
+Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species corresponding to the vertical stratum reached by most adult trees.
+The strata included four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m. 
+Because our study site is characterized by a steep topography (Fig. \@ref(fig:Uppangala)), 
+sub-canopy species can often reach the canopy in the Uppangala plot. 
+We thereby defined two broader categories, namely, canopy guild (C) including emergent, 
+canopy and sub-canopy species and understory guild (U) composed of understory trees.
 
@@
 We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species (3 individuals on average), depending on leaf accessibility and species abundance.
-For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following recommendation of @Perez-Harguindeguy2013.
+For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following the recommendation of @Perez-Harguindeguy2013.
 Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]).
@@
 For 11 species (1% of individuals in Uppangala plot), branch samples were too small (< to 1 cm in diameter) to allow measuring branch wood density.
-<!-- Among the selected species, 54 and 35 belonged to the canopy and understory guilds, respectively. -->
 
@@
 Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of @Perez-Harguindeguy2013.
-
-## Environmental variables
-
-We used airborne LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy structure.
-The LiDAR campaign was conducted in 2013, consistent with field measurements.
-Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, related to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the exposure to dominant monsoon winds [@Bunyan2015]; and (iv) slope, which could have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher].
-In addition, based on a 1-m resolution LiDAR model of canopy (Fig.
-\@ref(fig:Uppangala)), we calculated the top canopy height, as a proxy of decreasing light availability downward and thus of local competition for light.
-
-## Statistical Analyses
-
-To assess the degree of trait variation and covariation, we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific).
-We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.
-
-We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot.
-Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.
-
-We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:
-
-$$GWM_j=\sum_{i=1}^Sp_{i,j}*t_i$$
-where $p_{i,j}$ was the relative abundance of species $i$ in subplot $j$, $t_i$ the mean trait value of species $i$ and $S$ the number of species in the guild considered.
-
-We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild; [@bernard-verdier_community_2012]).
-The null model broke the link between environment and local functional composition while maintaining the spatial distribution of guilds and species abundances.
-We derived 999 randomized GWM values and calculated the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as
-
-$$SES = \frac{GWM_{obs}-\overline{GWM_{null}}}{\sigma_{null}}$$
-where $GWM_{obs}$ was the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ were the mean and standard deviation of the null distribution, respectively.
-
-We assessed how the deviation of guild composition from the null model depended on environmental variables by performing linear regressions and a bidirectional stepwise selection of predictors.
-These analyses were performed independently for each of the five functional traits.
-All analyses were conducted using the R 3.3.0 software [@RCoreTeam2020].
-
-# Results
-
-## Trait variation and covariation
-
-We found an overall increase of functional trait variation from intra-individual to inter-specific levels (Table \@ref(tab:traits)), even though LA showed almost as much variability at intra-individual and at intra-specific levels.
-Thus, trait values averaged at species level could grasp most of the functional variability across guilds and communities with varying taxonomic composition.
-
-The first axis of the PCA performed on species trait values (Fig.
-\@ref(fig:pca)A) was mostly related to wood density and structural leaf traits (LT, LA, LDMC), and opposed species with thick, big leaves and lightwood to species with dense leaves and hardwood.
-The second axis mostly opposed SLA to LDMC.
-WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that WD and SLA provide independent information.
-Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p=0.2 and p=0.3 respectively, Fig.
-\@ref(fig:pca)B).
-
-## Changes in functional composition with environment
-
-At community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig.
-\@ref(fig:GWMboxplot)).
-GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, lower for LA and higher for LDMC and LT than understory tree assemblages.
-In other words, understory guild assemblages had larger, thinner and lighter leaves with higher SLA and denser woods, compared to canopy guild assemblages.
-
-We detected a significant effect of slope, curvature and wetness on both canopy and understory guild assemblages, of southwesterness on canopy guild assemblages and of canopy height on understory guild assemblages (Table \@ref(tab:LM)).
-Wet and convex areas with more slope were associated to dense wood (WD) and thick (LT) leaves with high SLA in assemblages of both canopy and understory guilds. 
-Convex areas showed opposed effect on LDMC, and wet areas opposed effect on LA, between assemblages of canopy and understory guilds.
-Assemblages from canopy guilds under southwest orientation were associated to dense wood (WD) with small leaves (LA).
-Assemblages from understory guilds under short canopy were associated to dense wood (WD) with thick (LT) and dense (LDMC) leaves with low SLA.
-
-# Discussion
-
-Few studies have investigated how the functional characteristics and composition of canopy and understorey guilds vary in tropical forests. 
-We here show that functional trait values do not differ globally between canopy and understory tree species but that contrasting environments between understory and canopy filter species non-randomly, resulting in a markedly different functional composition of canopy and understory guilds within subplots.
-We also showed that the guilds functional assemblages of canopy and understory  respond differently to environmental variation.
-Indeed, although the functional composition of understory and canopy guilds assemblages respond similarly to water and nutrient access, the functional composition of canopy guild communities strongly varies along the wood economics spectrum, in response to wind exposure, and the functional variation of understory assemblages mostly varies along the leaf economics spectrum, in response to  light availability.
-These results show that different environmental gradients simultaneously drives different functional responses among guilds in tropical rainforest.
-
-<!-- We found that the measured functional traits of trees defined two orthogonal axes reflecting the leaf and wood economic spectra of trees at the scale of a 10ha rainforest plot.
-While functional trait values did not differ globally between canopy and understory tree species, we found neat difference between the assemblages of canopy and understory species within subplots.
-Therefore local environmental filtering selected species with diverging functional trait values across vertical guilds within communities.
-In addition the canopy and understory guild assemblages responded differently to environmental variation across communities.
-Soil resource, light availability and monsoon disturbances influenced the functional responses.
-Although understory and canopy assemblages responded similarly to water and nutrient access, functional variation of canopy assemblages was mostly along the wood economics spectrum, in response to aspect and wind exposure, while the functional variation of understory assemblages was mostly along the leaf economics spectrum in response to canopy height and decreasing light availability.
-These results show that environmental conditions drives different functional responses of vertical guilds in a tropical rainforest.
--->
-
-<!-- Assessing the variation of functional assemblages within tropical tree communities between understory and canopy species, we showed that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities.
-We further showed that the canopy and understory guilds assemblages responded differently across communities to environmental gradients related to soil resource, light availabilities and monsoon disturbances.
-While both understory and canopy assemblages responded similarly to water and nutrient access, canopy assemblages responded to monsoon winds with stronger support of the wood economics spectrum and understory assemblages responded to light access with a strong support of the leaf economics spectrum.
-These results illustrate how fine-scale environmental heterogeneity may shape the tree community composition and function.
-Moreover our study show guild-specific answer to biotic and abiotic environmental filtering between canopy and understory, highlighting the importance of forest vertical stratification in community assembly processes.
--->
-
-## Decoupling of wood and leaf economics spectra at local scale
-
-Functional traits PCA analysis supported two orthogonal main dimensions (Fig. \@ref(fig:pca)). 
-<!-- Despite a reduced number of functional traits, our study support the existence of decoupled wood and leaf economics spectra at local scale in Indian rainforests.
-We found locally that species trait values varied along two orthogonal main dimensions (Fig.
-\@ref(fig:pca)).
--->
-SLA and LDMC were opposed into a dimension representing the leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004], 
-<!-- representing a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Baraloto2010; @Evans2001; @hodgson_is_2011]. -->
-while, wood density was opposed to leaf area and thickness along a dimension reflecting the wood economics spectrum [@chave_towards_2009].
-<!-- , which contrasted species more tolerant to hydric stress, with denser wood, to more competitive species with thicker and larger leaves and higher efficiency in water conductivity [@Santiago2004a].
--->
-<!-- To our knowledge, our study is the first to support the two economics spectra in Indian rainforests. -->
-@Baraloto2010 and @Fortunel2012a also found orthogonal dimensions reflecting wood and leaf economic spectra of Neotropical tree species at regional scale in Amazonia.
-In response to these works, @Messier2017 suggested that these two orthogonal spectra reflect large scale difference in tree functional composition and would not be observed among species co-occurring at a local scale.
-<!-- Here we show that the orthogonal spectra do exist across species at the scale of a 10ha plot. -->
-Here we captured  the two orthogonal spectra across species at the scale of a 10 ha plot despite a limited number of traits compared to previous studies.
-<!-- and (ii) we provide first evidence of these spectra in Indian rainforests. -->
-Our results support that economics spectra are decoupled [@Baraloto2010] and not related to a single, whole-plant economics spectrum [@reich_world-wide_2014].
-
-## Functional differences between canopy and understory guilds
-
-Although the positions of canopy and understory species along the economics spectra did not differ globally (Fig.
-\@ref(fig:pca)B), we found significant trait differences between assemblages of canopy and understory species co-occuring in local 30 x 30m communities (Fig.
-\@ref(fig:GWMboxplot)).
-Therefore, species pools of canopy and understory species broadly overlap in functional spaces, and both spanned a broad range of ecological strategies (Fig. \@ref(fig:pca)A).
-<!-- However, niche differentiation occurred in local communities across canopy and understory assemblages leading to contrasted functional compositions. -->
-However, environmental niches filtered species in local communities leading to contrasted functional compositions between canopy and understory guild.
-Understory assemblages had larger, thinner and lighter leaves, hence greater SLA, but denser wood, while canopy assemblages showed lighter wood but also smaller leaves. 
-<!-- The differences did not support a mere differentiation of guild assemblages along economic spectra. -->
-Light wood suggested greater hydraulic conduction of tall trees [@Santiago2018].
-The conservative leaf strategy observed in canopy assemblage may respond to (i) tree height increasing leaf water stress due to gravity [@Koch2004], (ii) canopy light leading to a thicker mesophyll protecting tissues from excessive irradiance [@Chazdon1993], and (iii) a higher risk of herbivory and mechanical damage in canopy [@Sterck1992].
-Drought constraint could be less important for understory trees, allowing greater investment into resource acquisition by leaves to solve the higher need for light in more shaded understory [@Chazdon1996].
-However limited light may result in slower growth [@Herault2011] favoring denser wood. 
-These results support the hypothesis that trees experience guild-specific ecological constraints within communities [@duque2002different], leading to contrasting functional compositions of canopy and understory assemblages.
-<!-- However the functional variation within guilds was still broad enough to entail overlap of the functional spaces at the scale of a 10ha plot, insuring greater stability in space and time to maintain the guilds functions [@Yachi1999].
--->
-
-## Different responses to abiotic and biotic environment between canopy and understory assemblages
-
-Our results show that the two guilds assemblages respond differently to environmental variation (Table \@ref(tab:LM)). 
-Contrary to understory assemblages, canopy assemblages respond to southwesterness, which face dominant monsoon winds in our study site, along the wood economics spectrum with a greater WD and smaller LA for higher southwesterness.
-The influence of aspect on tropical forest composition has been little documented.
-@gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar climatic and topographical context in Sri Lanka and also interpreted it as a consequence of a higher exposure to monsoon winds.
-In the Western Ghats of India, @Bunyan2015 found greater occurrence of forest fragments on southwest orientation in a forest-grassland mosaic, probably as a result from monsoon winds. 
-The Uppangala plot is characterized by a large topographic heterogeneity, with steep slopes mostly organized along southwest-oriented crests (Fig. \@ref(fig:Uppangala)).
-South-west oriented slopes are likely to be more exposed to monsoon winds, which could entail greater tree damages and forest disturbance [@soman_aspects_1990].
-The consequences could be that (i) tree strategies associated with mechanical resistance and disturbance tolerance could be selected in areas more exposed to monsoon wind damages [@Poorter2010b], or (ii) turbulences on northeast oriented areas could yield more disturbance and death events favoring pioneer, lightwood species [@larjavaara_rethinking_2010].
-In both cases, canopy species are more exposed to wind disturbances than understory trees and are thus expected to respond more to wind exposure.
-
-Conversely, we found pervasive influence of canopy height on the functional composition of understory assemblages related to leaf economics, with more acquisitive strategies under tall canopies.
-We indeed found a greater investment into photosynthetic activity, with higher SLA and lower LDMC and LT in understory assemblages under taller canopy. 
-Taller canopies tend to bear greater vertical differentiation ensuring greater light availability and decreased investment in leaf lifespan in understory [@kitajima_tissue-level_2010].
-Higher canopies filtered also smaller wood density for understory assemblages.
-Lesser hydraulic constraints  in understory guilds could entail a greater role of competition among the trees, filtering lighter woods and relatively faster growth.
-
-Nonetheless, curvature, wetness and slope similarly affected canopy and understory functional assemblages, except for LDMC.
-Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and retain more water, increasing soil fertility [@ferry2010higher].
-It thus create favorable conditions where species with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA) strategies outcompete others.
-By contrast, concave, drier or sloppy areas will be drier and less fertile favoring species assemblages with more conservative strategies and slower growth.
-
-## Conclusion
-
-Our results show that functional variations among species and assemblages preferentially vary along decoupled leaf and wood economic spectra in an Indian rainforest.
-We found a significant environmental determinism in the functional composition of tree communities but also stressed that this determinism is guild-specific.
-Our results  thus contribute to the debate on the nature and role of economics spectra in the assembly of tropical forests [@Baraloto2010].
-They also underline that the functional signature of community assembly should be decomposed to better characterize and understand the different strategies and responses across ecological guilds [@Raevel2018] and across forest vertical structure [@duque2002different].
-We thus call for a better acknowledgment of differential guild dynamics within communities and their influence on the global forest functioning.
-
-# Acknowledgements
-
-We would like to thank Cedric Vega for extending Uppangala plot to 10 ha.
-We also want to acknowledge the data cleaning operated by Arianne Tanguy.
-And we are thankful to Claire Fortunel for her comments.
-
-# Authors’ contributions
-
-SS, MRM, FM and VR conceived the study, designed the methodology and analysed model outputs; SS, FM, VR, NA, NB and NB conducted the field sampling; NA, NB and NB identified species; MRM and CSJ provided the environmental data; SS, MRM, FM and VR led the writing of the manuscript.
-All the authors contributed to the drafts and gave their final approval for publication.
-
-# Data availability
-
-Functional traits data have been submitted to TRY initiative [@Kattge2019] under the name UppangalaTraits.
-Environmental variables and spatial positions of individuals were extracted from the Uppangala Station database or **lidar data**, for which access is modulated by **to precise**.
-
-# References
-
-<div id="refs"></div>
-
-\newpage
-
-## Table captions
-
-__Table \@ref(tab:traits)__ Functional traits measured.
-Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).
-
-__Table \@ref(tab:LM)__ Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy).
-Linear model coefficients are reported only for predictors retained after bidirectional stepwise selection, with corresponding adjusted $R^2$.
-See Table \@ref(tab:traits) for trait abbreviations.
-
-## Figure captions
-
-__Figure \@ref(fig:Uppangala)__ Digital canopy model of the 10-ha forest permanent plot in Uppangala, India.
-Thin and thick black lines represent the elevation contour lines and plot limits, respectively.
-
-__Figure \@ref(fig:pca)__ Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B).
-Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test.
-See Table \@ref(tab:traits) for trait abbreviations.
-
-__Figure \@ref(fig:GWMboxplot)__ Guild weighted mean (GWM) of specific leaf area (SLA), wood density (WD), leaf dry matter content (LDMC), leaf area (LA) and leaf thickness (LT) for canopy (C) and understory (U) guilds.
-Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines).
-Asterisks represent significant differences between the two guilds (*** p-value<0.001).
-
-\newpage
-
-<!-- table 1 -->
 
@@
   kable(digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
+                  col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
+                        "Intra-individuals", "Intra-species", "Inter-species"),
         caption = 'Functional traits measured.
@@
 
-\newpage
+## Environmental variables
 
-<!-- table 2 -->
+We used airborne LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy structure.
+The LiDAR campaign was conducted in 2013, consistent with field measurements.
+Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, related to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the exposure to dominant monsoon winds [@Bunyan2015]; and (iv) slope, which could have both a strong effect on forest dynamics and on the availability of nutrients and water [@ferry2010higher].
+In addition, based on a 1-m resolution LiDAR model of canopy (Fig.
+\@ref(fig:Uppangala)), we calculated the top canopy height, as a proxy of decreasing light availability downward and thus of local competition for light.
 
-```{r LM}
-group_by(SES, strata, trait) %>% 
-  mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), scale, center = F) %>%
-  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
-                direction = 'both', trace = 0)) %>%
-  tidy(reg) %>% 
-  filter(term != "(Intercept)") %>% 
-  mutate(estimate = round(estimate, 2)) %>% 
-  group_by(strata, trait, term) %>%
-  dcast(strata + trait ~ term, value.var = "estimate") %>%
-  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>%
-    mutate_all(funs(replace(., is.na(.), ' '))) %>%
-  left_join(group_by(SES, strata, trait) %>% 
-              mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), 
-                        scale, center = F) %>%
-              do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
-                            direction = 'both', trace = 0)) %>%
-              glance(reg) %>%
-              dplyr::select(r.squared) %>%
-              mutate(r.squared = round(r.squared, 2)), 
-            by = c("strata", "trait")) %>%
-            rename("$R^2$" = r.squared, Guild = strata, Trait = trait) %>%  
-    kable(caption = 'Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy).
-Linear model coefficients are reported only for predictors retained after bidirectional stepwise selection, with corresponding adjusted $R^2$ .
-See Table \\@ref(tab:traits) for trait abbreviations.', format = 'pandoc')
-```
+## Analyses
 
-\newpage
+To assess the degree of trait variation and covariation, we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific).
+We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.
 
-<!-- figure 1 -->
+We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot.
+Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.
+We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:
 
-```{r Uppangala, fig.cap="Digital canopy model of the 10-ha forest permanent plot in Uppangala, India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
-DCM <- raster("./data/Canopy.tif")
-crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
-DEM <- raster("./data/DEM.tif")
-crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
-limits <- shapefile("./data/Plots/Shapeall10ha.shp")
-limits$melt <- 1
-limits <- maptools::unionSpatialPolygons(limits, limits$melt)
-india <- shapefile("./data/India_SHP/INDIA.shp")
-DCM <- crop(DCM, buffer(limits, 100))
-DEM <- crop(DEM, buffer(limits, 100))
-DEM <- as(DEM, "SpatialPixelsDataFrame")
-DEM <- as.data.frame(DEM)
-colnames(DEM) <- c("Elevation", "x", "y")
-DCM <- as(DCM, "SpatialPixelsDataFrame")
-DCM <- as.data.frame(DCM)
-colnames(DCM) <- c("Height", "x", "y")
-ggplot() +  
-geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
-  geom_contour(data = DEM, aes(x = x, y = y, z = Elevation),
-               colour = "black", alpha = 0.5, breaks = seq(300, 500, by = 25)) +
-  directlabels::geom_dl(data = DEM, aes(x = x, y = y, z = Elevation, label = paste0(..level.., "m")),
-                        method = list('last.bumpup', cex = 0.8, hjust = 1, fontface = 2),
-                        stat = "contour", breaks = seq(300, 500, by = 25)) +
-  geom_polygon(data = fortify(limits),
-               aes(long, lat, group=group),
-               colour = "black", lwd = 1.1, alpha = 0) +
-  coord_equal() +
-  cowplot::theme_map() +
-  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
-  ggsn::scalebar(dist = 50, dist_unit = "m", x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 20, y.max = max(DCM$y),
-                 transform = F, model = "WGS84", location = "bottomleft") +
-    annotation_custom(grob = ggplotGrob(ggplot() +  
-                                        geom_polygon(data = fortify(india),
-                                                     aes(long, lat, group=group),
-                                                     colour = "black", alpha = 0) +
-                                        geom_point(data = data.frame(long = 75.662778, lat = 12.537500), 
-                                                   aes(long, lat), col = "red", size = 3) +
-                                        coord_equal() +
-                                        cowplot::theme_map() +
-                                        theme(panel.border = element_rect(color = "black",size = .5,fill = NA),
-                                              plot.background	= element_rect(fill = "white"))), 
-                    xmin = min(DCM$x), xmax = min(DCM$x) + 150, 
-                    ymin = max(DCM$y) - 150, ymax = max(DCM$y))
-  # ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
-```
+$$GWM_j=\sum_{i=1}^Sp_{i,j}*t_i$$
+where $p_{i,j}$ was the relative abundance of species $i$ in subplot $j$, $t_i$ the mean trait value of species $i$ and $S$ the number of species in the guild considered.
 
-\newpage
+We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in $Trait \times Species$ matrices of each guild; [@bernard-verdier_community_2012]).
+The null model broke the link between environment and local functional composition while maintaining the spatial distribution of guilds and species abundances.
+We derived 999 randomized GWM values and calculated the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as
 
-<!-- figure 2 -->
+$$SES = \frac{GWM_{obs}-\overline{GWM_{null}}}{\sigma_{null}}$$
+where $GWM_{obs}$ was the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ were the mean and standard deviation of the null distribution, respectively.
+
+We assessed how the deviation of guild composition from the null model depended on environmental variables by performing linear regressions.
+Predictors were all normalised to ease model inference and enable their comparison. 
+The analyses were performed independently for each of the five functional traits.
+A Bayesian method was used to infer parameters using stan language [@Carpenter2017] and rstan package [@Team2018] in the R environment [@RCoreTeam2020].
+
+# Results
+
+## Trait variation and covariation
+
+We found an overall increase of functional trait variation from intra-individual to inter-specific levels (Table \@ref(tab:traits)), even though LA showed almost as much variability at intra-individual and at intra-specific levels.
+Thus, trait values averaged at species level could grasp most of the functional variability across guilds and communities with varying taxonomic composition.
+
+The first axis of the PCA performed on species trait values (Fig.
+\@ref(fig:pca)A) was mostly related to wood density and structural leaf traits (LT, LA, LDMC), and opposed species with thick, big leaves and lightwood to species with dense leaves and hardwood.
+The second axis mostly opposed SLA to LDMC.
+WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that WD and SLA provide independent information.
+Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p>0.2, Fig.
+\@ref(fig:pca)B).
 
@@
 Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test.
-See Table  \\@ref(tab:traits) for trait abbreviations.', eval=F}
+See Table  \\@ref(tab:traits) for trait abbreviations.'}
 pca.plot <- autoplot(PFT.pca, 
-         data = Q %>% 
-           left_join(L %>% 
-                       dplyr::select(-com) %>% 
-                       summarise_all(sum) %>% 
-                       t() %>% 
-                       data.frame() %>% 
-                       rename(abundance =  ".") %>% 
-                       mutate(SpCode = row.names(.))) %>% 
-           filter(!is.na(WES)), 
+         data = na.omit(reshape2::dcast(PFT, strata + SpCode ~ trait, value.var = "traitvalue")),
          colour = "strata",
-         size = "abundance",
          loadings.label.size = 6,
@@
   ggplot(aes(strata, value, fill = strata)) +
-  ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
+  # ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
   geom_boxplot() +
@@
 
-\newpage
+## Changes in functional composition with environment
 
-<!-- figure 3 -->
+At community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig.
+\@ref(fig:GWMboxplot)).
+GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, lower for LA and higher for LDMC and LT than understory tree assemblages.
+In other words, understory guild assemblages had larger, thinner and lighter leaves with higher SLA and denser woods, compared to canopy guild assemblages.
 
@@
 Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines).
-Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4, eval=F}
-L %>%
-  melt(id.var = "com",
-       variable.name = "SpCode",
-       value.name = "abundance") %>% 
-  dcast(SpCode ~ com, value.var = "abundance") %>% 
-  mutate(SpCode = as.character(SpCode)) %>% 
-  inner_join(Q, by = "SpCode") %>%
-  select(-SpCode) %>%
-  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
-       variable.name = 'community', value.name = 'abundance') %>%
-  group_by(strata, community) %>%
-  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
-  select(-community, -LES, -WES) %>%
-  # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
-  # select(-community) %>% 
-  melt(id.vars = "strata", 
-       variable.name = "trait",
-       value.name = "GWM") %>% 
-  ggplot(aes(x = strata, y = GWM, fill = strata)) +
+Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4}
+ggplot(GWM, aes(x = strata, y = GWM, fill = strata)) +
   geom_boxplot() +
   facet_wrap(~trait, scales = "free", nrow = 1) +
-  # facet_wrap(~trait, scales = "free", nrow = 2) +
   xlab("Guild") + scale_fill_discrete(guide = "none") +
   ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3))))
-  # ggsave(file = "boxplot.png",
-  #        path = './data/',
-  #        width = 200, height = 150, unit = 'mm',
-  #        dpi = 300)
 ```
+
+We detected a significant effect of slope, curvature and wetness on both canopy and understory guild assemblages, 
+of southwesterness on canopy guild assemblages 
+and of canopy height mostly on understory guild assemblages (Fig. \@ref(fig:LMres)).
+Curvature appeared as the most influential factor (7 significant effects out of 10 tested), 
+with a negative effect on SLA and a positive effect on LT in both canopy and understory assemblages. 
+In addition, curvature had a positive effect on WD and LDMC only in the understory assemblages,
+and a negative effect on LA only in the canopy assemblages. 
+Similarly to curvature, wetness had a negative effect on SLA in both canopy and understory assemblages,
+and a positive effect on LT in the understory assemblages.
+Still, wetness had a negative effect on LDMC in canopy assemblages.
+As curvature, slope had a negative effect on SLA and a positive effect on LT in canopy assemblages. 
+and a positive effect on WD in the understory assemblages.
+Southwesterness only influenced canopy assemblages,
+with a positive effect on WD and a negative effect on LA. 
+Canopy height mainly influenced understory assemblages,
+with a positive effect on SLA and a negative effect on WD, LDMC and LT.
+Still, canopy height  had a positive effect on LA in canopy assemblages.
+
+```{stan LMmodel, output.var="lm"}
+data {
+  int<lower=1> N ;
+  int<lower=1> K ;
+  vector[N] SES ;
+  matrix[N,K] X ;
+}
+parameters {
+  real alpha ;
+  vector [K] beta ;
+  real<lower=0> sigma ;
+}
+model {
+  SES ~ normal(alpha + X * beta, sigma) ;
+}
+generated quantities {
+  real R2 = 1 - dot_self(SES - alpha - X * beta)/dot_self(SES - mean(SES)) ;
+}
+```
+
+```{r LMres, fig.cap="Effects of slope, curvature, wetness, southwesterness, and canopy height on  standardized effect size values of functional traits for canopy and understory guilds.
+The effects of predictors was estimated as the posterior distribution of the slope parameters using Bayesian inference.
+Circles represent the mean estimate, thick lines the 50% confidence interval, and thin lines the 90% confidence interval, and colour the corresponding predictor of the posterior distributions.
+Transparency represents significance, with transparent non-significant posteriors and opaque significant posteriors (p < 0.05).
+See Table \\@ref(tab:traits) for trait abbreviations."}
+mdata <- group_by(SES, strata, trait) %>% 
+  mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), scale) %>% 
+  group_by(strata, trait) %>% 
+  do(mdata = list(N = nrow(.), K = 5, SES = .$SES, X = as.matrix(.[c("Slope", "Curvature", "Wetness", "SW", "Canopy")])))
+fits <- lapply(mdata$mdata, function(m) sampling(lm, data = m))
+names(fits) <- paste0(mdata$strata, "_", mdata$trait)
+R2 <- lapply(fits, mcmc_intervals_data, regex_pars = "R2") %>% 
+  bind_rows(.id = "model") %>% 
+  separate(model, c("Guild", "Trait"), "_") %>% 
+  mutate(R2 = m) %>% 
+    mutate(Guild = recode(Guild, 
+                           "U" = "Understory",
+                           "C" = "Canopy")) %>% 
+  dplyr::select(Guild, Trait, R2) %>% 
+  mutate(Variable = "Canopy height", Significance = 1) %>% 
+  mutate(R2 = paste0("R²=", round(R2, 2)))
+lapply(fits, mcmc_intervals_data, regex_pars = "beta") %>% 
+  bind_rows(.id = "model") %>% 
+  separate(model, c("Guild", "Trait"), "_") %>% 
+  mutate(Variable = recode(parameter, 
+                           "beta[1]" = "Slope",
+                           "beta[2]" = "Curvature",
+                           "beta[3]" = "Wetness",
+                           "beta[4]" = "Southwesterness",
+                           "beta[5]" = "Canopy height")) %>% 
+    mutate(Guild = recode(Guild, 
+                           "U" = "Understory",
+                           "C" = "Canopy")) %>% 
+  mutate(Significance = 0) %>% 
+  mutate(Significance = ifelse(m > 0 & ll > 0, 1, Significance)) %>% 
+  mutate(Significance = ifelse(m < 0 & hh < 0, 1, Significance)) %>% 
+  ggplot(aes(x = Variable, xend = Variable,
+             color = Variable, fill = Variable,
+             alpha = as.factor(Significance))) +
+  geom_hline(yintercept = 0, color = "gray90", size = 1) +
+  geom_hline(yintercept = c(-0.25, 0.25), 
+             color = "gray90", size = 0.5, lty = "dashed") +
+  geom_point(aes(y = m), shape = 21, size = 3) +
+  geom_segment(aes(y = ll, yend = hh),
+               size = 1, show.legend = F) +
+  geom_segment(aes(y = l, yend = h), size = 2) +
+  geom_text(y = -0.35, aes(label = R2), col = "black", data = R2) +
+  coord_flip() +
+  facet_grid(Trait ~ Guild, labeller = label_parsed, 
+              scales = "free_y") +
+  xaxis_title(F) +
+  ylab("Standardised effect size") +
+  scale_color_discrete(guide = "none") +
+  scale_fill_discrete(guide = "none") +
+  scale_alpha_manual(guide = "none", values = c(0.3, 1))
+```
+
+# Discussion
+
+Few studies have investigated how the functional characteristics and composition of canopy and understorey guilds vary in tropical forests. 
+We here show that functional trait values do not differ globally between canopy and understory tree species but that the contrasting environments between understory and canopy filter species non-randomly,
+resulting in a markedly different functional composition of canopy and understory guilds within subplots. 
+We found that topography and canopy height significantly and differently impacts both canopy and understory guild assemblages.
+Indeed, the functional composition of understory and canopy guilds assemblages respond similarly to water and nutrient access
+but the functional composition of canopy guild communities strongly varies in response to wind exposure, 
+and the functional variation of understory assemblages varies in response to light availability.
+These results show that different environmental gradients simultaneously drive different functional responses among guilds in tropical rainforest.
+
+## Decoupling of wood and leaf economics spectra at local scale
+
+The PCA performed on species functional traits supported two orthogonal main dimensions (Fig. \@ref(fig:pca)).
+SLA and LDMC were opposed into a dimension representing the leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004],
+while, wood density was opposed to leaf area and thickness along a dimension previously interpreted as the wood economics spectrum  [@chave_towards_2009].
+@Baraloto2010 and @Fortunel2012a also found orthogonal dimensions reflecting wood and leaf economic spectra of Neotropical tree species at regional scale in Amazonia.
+In response to these works, @Messier2017 suggested that these two orthogonal spectra reflect large scale differences in tree functional composition and would not be observed among species co-occurring at a local scale.
+Here we captured the two orthogonal spectra across species at the scale of a 10 ha plot despite a limited number of traits compared to previous studies.  
+Our results support that economics spectra are decoupled in tropical forests [@Baraloto2010] and not related to a single, whole-plant economics spectrum [@reich_world-wide_2014].
+
+## Functional differences between canopy and understory guilds
+
+We found that species pools of canopy and understory species broadly overlap in functional spaces, 
+and both spanned a broad range of ecological strategies (Fig. \@ref(fig:pca)). 
+However, we found significant trait differences between assemblages of canopy and understory species co-occuring in local 30 x 30m communities (Fig. \@ref(fig:GWMboxplot)).
+Environmental conditions thus filtered species in local communities according to their trait, 
+leading to highly contrasted functional compositions between canopy and understory guilds. 
+Understory assemblages had larger, thinner and lighter leaves, hence greater SLA, 
+and denser wood than canopy guild assemblages.
+Greater hydraulic conduction is associated with lighter wood [@Santiago2018],
+which is especially important for canopy species. 
+The conservative leaf strategy observed in canopy assemblages may originate from two mechanisms. 
+First, leaf water stress is expected to increase with tree height due to gravity [@Koch2004]
+and to a higher water evaporation associated with sun and wind exposure [@Chazdon1993]. 
+Second, a higher risk of herbivory exists in the canopy [@Sterck1992],
+which may select for species resistant to herbivory with a more conservative strategy. 
+Drought constraint could be less important for understory trees, 
+allowing greater investment into resource acquisition by leaves to counterbalance the higher need for light in more shaded understory [@Chazdon1996]. 
+However limited light may result in slower growth [@Herault2011] favoring denser wood. 
+These results support the hypothesis that trees experience guild-specific ecological constraints within communities [@duque2002different], 
+leading to contrasting functional compositions of canopy and understory assemblages.
+
+## Different responses to abiotic and biotic environment between canopy and understory assemblages
+
+Our results show that the two guilds assemblages respond differently to some environmental gradients (Fig. \@ref(fig:LMres)). 
+Curvature, wetness and slope similarly affected canopy and understory functional assemblages, except for LDMC. 
+Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and retain more water, increasing soil fertility [@ferry2010higher].
+It thus creates favorable conditions where species with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA) strategies outcompete others. 
+By contrast, concave, drier or sloppy areas will be drier and less fertile favoring species assemblages with more conservative strategies and slower growth.
+
+Contrary to understory assemblages, canopy assemblages respond to southwesterness, 
+which face dominant monsoon winds in our study site, 
+along the wood economics spectrum with a greater WD and smaller LA for higher southwesterness. 
+The influence of aspect on tropical forest composition has been little documented. 
+@gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar climatic and topographical context in Sri Lanka and also interpreted it as a consequence of a higher exposure to monsoon winds.
+The Uppangala plot is characterized by a large topographic heterogeneity,
+with steep slopes mostly organized along southwest-oriented crests (Fig. \@ref(fig:Uppangala)).
+South-west oriented slopes are likely to be more exposed to monsoon winds,
+which could entail greater tree damages and forest disturbance [@soman_aspects_1990]. 
+The consequences could be that 
+(i) tree strategies associated with mechanical resistance and disturbance tolerance could be selected in areas more exposed to monsoon wind damages [@Poorter2010b],
+or (ii) turbulences on northeast oriented areas could yield more disturbance and death events favoring pioneer, lightwood species [@larjavaara_rethinking_2010]. 
+In both cases, canopy species are more exposed to wind disturbances than understory trees and their functional composition is thus expected to respond more to wind exposure.
+
+Conversely, we found pervasive influence of canopy height on the functional composition of understory assemblages related to leaf and wood economics, with increased acquisitive and growth strategies under tall canopies.
+We indeed found a greater investment into photosynthetic activity and growth, with higher SLA and lower LDMC, LT and WD, in understory assemblages under taller canopy. 
+We expected canopy height to represent decreased light-availability contrasting with the observed functional strategies expected under increased access to light [@kitajima_tissue-level_2010]. 
+As canopy height is as dynamic as individual recruitment of current species assemblages,
+we hypothesized taller communities to result from more recent gaps with an increased abundance of acquisitive and fast growing post-pioneer species in understory.
+This hypothesis needs to be thoroughly tested with growth or canopy time series.
+
+## Conclusion
+
+Our results show that functional variations among species and assemblages vary along the decoupled economic spectra of leaves and wood in an Indian rainforest.
+We found a significant effect of topography and canopy height on the functional composition of tree communities, 
+but also pointed out that these effects may be guild specific.
+Our results argue in favour of decomposing the functional signature of community assemblage
+in order to better characterize and understand the different strategies and responses of ecological guilds [@Raevel2018]
+and the vertical structure of forests [@duque2002different]. 
+We therefore call for a better recognition of the different dynamics of guilds within communities and their influence on the global functioning of forests.
+
+# Acknowledgements
+
+We would like to thank Cedric Vega for extending Uppangala plot to 10 ha.
+We also want to acknowledge the data cleaning operated by Arianne Tanguy.
+And we are thankful to Claire Fortunel for her comments.
+
+# Authors’ contributions
+
+SS, MRM, FM and VR conceived the study, designed the methodology and analysed model outputs; SS, FM, VR, NA, NB and NB conducted the field sampling; NA, NB and NB identified species; MRM and CSJ provided the environmental data; SS, MRM, FM and VR led the writing of the manuscript.
+All the authors contributed to the drafts and gave their final approval for publication.
+
+# Data availability
+
+Functional traits data have been submitted to TRY initiative [@Kattge2019] under the name UppangalaTraits. 
+Environmental variables and spatial positions of individuals were extracted from the Uppangala Station database.
+Access to LiDAR data is restricted by the Indian government and should be the object of a convention (contact: chandra.s.jha@gmail.com).
+
+# References

### 2020-11-09 Manuscript.Rmd
JVAS formatting

@@
 ---
-title: Topography and canopy height differently shape the functional composition of canopy and understory guilds in an Indian rainforest
-date: '`r Sys.Date()`'
+title: Contrasting influences of topography and canopy height on the functional assembly of canopy and understory tree guilds in an Indian rainforest
 output:
@@
 linestretch: 1.5
-csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
+csl: /home/sylvain/Documents/Bibliography/csl/journal-of-vegetation-science.csl
 bibliography: /home/sylvain/Documents/Bibliography/library.bib
@@
     - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
-    - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France
+    - Laboratoire Interdisciplinaire de Physique, Univ. Grenoble Alpes, FR-38000 Grenoble, France
+
+**Running title:** Functional assembly of canopy and understory tree
+
+# Funding
+
+No funding to declare. **Comment: ?**
 
@@
 
-# Changes
-
-## 13 octobre 2020
-
-* I changed the title.
-* I removed the bidirectional stepwise selection of predictors as it was unstable.
-* I used bayesian inference for the linear regression (minor).
-* I replaced table 2 by a figure of posteriors distribution and significance (seemed clearer).
-* I removed the ideas of LES and WES between strata.
-That's unclear and difficult to defend.
-WD model in understorey showed a $R^2=0.3$, which is huge and removed the idea that the WES mainly affects canopy.
-Similarly, SLA and LDMC models in canopy had low $R^2$ but showed significant effects of wetness, curvature and slope.
-* I kept Maxime’s hypothesis for canopy height effect in understory.
-I don’t have a better hypothesis.
-I am therefore to keep it calling for further testing.
-**Maybe we should remove canopy height to simplify the message?**
-* I am not a fan of the conclusion. I kept it for the moment but I didn’t feel it brought much more than the take home message; So I’m hesitating to simply remove it.
-
 # Abstract
 
-1.
-Topographic heterogeneity determines changing soil conditions, while the height of the canopy modulates the microclimate in the forest.
-But the environmental conditions encountered by trees also change across the vertical strata of tropical forests.
-However, the joint influence of topography and canopy height with vertical strata on tree functional composition remain little known.
-We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity and canopy height, depending on resource availability, hydraulic and biomechanical constraints.
-1.
-We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India.
-The plot was subdivided into tree communities, each composed of guilds of canopy and understory trees.
-We combined tree census data, LiDAR-derived topographic and canopy metrics, and functional traits to assess the variation of functional composition in understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.
-1.
-The variation of functional traits across sampled tree species followed locally decoupled wood and leaf economics spectra.
-Although functional space did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities.
+* *Questions.* Topographic heterogeneity determines changing soil conditions, while canopy height modulates the microclimate in forest ecosystems.
+In addition, tree species form multiple vertical layers undergoing different light, biomechanic, and hydraulic constraints.
+However, how topography and canopy height influence the functional composition of vertical strata remain little addressed.
+We hypothesized that leaf and wood functional strategies differ between canopy and understory tree assemblages depending on local environmental heterogeneity.
+* *Location.* An undisturbed, old-growth wet evergreen Dipterocarp forest plot located in the Kadamakal Reserve Forest, Karnataka State, India.
+* *Methods.* We measured leaf and wood traits of 89 tree species representing 99% of all individuals in a 10-ha rainforest plot in the Western Ghats, South India.
+We subdivided the plot into subplots representing varying topographic conditions.
+We assigned tree species to guilds of canopy and understory trees.
+We combined tree census data, LiDAR-derived topographic and canopy metrics to assess the variation of the weighted means of functional trait values in understory and canopy guilds with canopy height and topography.
+* *Results.* The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra at the local scale.
+Although functional trait values did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities.
 Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
-1.
-*Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
+* *Conclusions.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
 Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.
@@
 Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976].
-Environmental variation influences the composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003] 
-and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
-Topography is known to filter different ecological strategies [@kraft_functional_2008; @Schmitt2020]
-due to its impact on water and nutrient availability [@Blanchard2019; @ferry2010higher].
-In addition, canopy height modulates forest microclimates due to topography [@Jucker2018].
-But, tree species  occupy different vertical strata at adult stage, and can thus be classified into different vertical guilds (understory and canopy),
-experimenting contrasted environmental conditions. 
-Indeed, light availability decreases downward across vertical forest strata [@chazdon1984],
-while hydraulic and mechanical constraints increase upward [@Koch2004], 
-so that different survival and growth strategies are filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a]. 
-While functional strategies of tree species have been documented independently across topography, canopy height and vertical guilds,
-little is known on their interactions,
-particularly the effect of topography and canopy height on species assemblages between vertical guilds has not been addressed [@wright2002plant].
+Environmental variation influences the composition of tree communities at regional [@Rejou-Mechain2008; @Tuomisto2003] and local [@Ferry2007; @gunatilleke_specieshabitat_2006] scales.
+Topography influences water and nutrient availability [@Blanchard2019; @ferry2010higher], in interaction with canopy height [@Jucker2018], and thereby filters different ecological strategies [@kraft_functional_2008; @Schmitt2020].
+In addition,  tree species form different vertical strata at the adult stage allowing to distinguish guilds of understory and canopy trees.
+These vertical guilds should undergo contrasting environmental conditions: light availability decreases downward across vertical forest strata [@chazdon1984], while hydraulic and mechanical constraints increase upward [@Koch2004].
+Therefore, we expect different survival and growth strategies to be filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a].
+While much emphasis has been put on how ecological strategies of tree species vary with topography or canopy height in forests, far fewer works have addressed whether and how these variations depend on vertical guilds [@wright2002plant].
 
-Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], 
-contributing thus to test for community assembly processes. 
 Functional traits have been defined as phenotypic traits that impact fitness through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007].
-Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018
-along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
+Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
 The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016].
-Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients. 
-The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009].
-Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012]. 
-Nevertheless, @reich_world-wide_2014
-proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].
-The concept of plant economics spectrum may be highly relevant for understanding the variation of strategy between vertical guilds as understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
+Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients, with more acquisitive strategies with decreasing resources availability [@hodgson_is_2011].
+The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009].
+Whether and how trait variations along the leaf and wood economics spectra are decoupled across tropical tree species remain debated [@Baraloto2010; @fortunel_leaf_2012; but see @reich_world-wide_2014; @freschet_evidence_2010].
 
-In the context of seasonal rainforests, we hypothesized different trait-environment relationships across vertical guilds, 
-with varying strategies along the fundamental leaf and wood economics spectra. 
-More specifically, because understory trees are strongly light-limited and are less subject to hydraulic and mechanical constraints than canopy trees, 
-we expect that  understory and canopy species tend to maximize their difference along the leaf and wood economic spectra, respectively. 
-More generally, we expect that understory species and assemblages have more conservative strategies than canopy ones and that the lower water and nutrient availability found on hilltops [@ferry2010higher]
-entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; @Schmitt2020]. 
-Previous works have shown different response of vertical strata to environmental variation at ground level, with increased edaphic constraint on understory compared to canopy  for Columbian drained uplands [@duque2002different].
+Here, we measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot.
+We categorized tree species as belonging to canopy and understory guilds at the adult stage.
+First, we analyzed the variation of functional traits across species and between guilds.
+Second, we combined trait information with tree census data, LiDAR-derived topographic and canopy metrics to characterize the functional composition of tree assemblages and their variation with topography [@Antin2013] and vertical structure [@Pascal1988].
+We used multivariate approaches and linear regressions to address the following questions:
 
-Here, we assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, 
-and examined the effects of topography and canopy height on these variations. 
-We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000].
-The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
-Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, 
-we expected hydraulic and mechanical constraints to play an important role in tree community assembly.
+* Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale?
+If so, how are canopy and understory guilds distributed along leaf and wood economics spectra?
+* Does functional composition differ between assemblages of canopy and understory guilds? 
+* Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently?
 
-We used tree census data, LiDAR-derived topographic and canopy metrics and measured functional traits, 
-in multivariate approaches and linear regressions to address the following questions: 
-(i) Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale? 
-If so, how are canopy and understory guilds distributed along leaf and wood economics spectra? 
-(ii) Does functional composition differ between assemblages of canopy and understory guilds? 
-and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently?
+We hypothesized different trait-environment relationships across vertical guilds, with varying strategies along the leaf and wood economics spectra.
+First we expected understory species to be strongly light-limited but less subject to hydraulic and mechanical constraints than canopy species, so that understory species should have more conservative strategy [@herault_functional_2011] but less drought tolerance than canopy species .
+Second, because of the long dry season period and influence of intense monsoon during rainy seasons with strong wind disturbances, we expected topography and canopy height to influence particularly resource use strategies in our plot.
+Lower water and nutrient availability found on hilltops [@ferry2010higher] should entail more conservative strategy [@Cunningham2007; @kraft_functional_2008;@Schmitt2020], while acquisitive strategies are expected in bottom-lands.
 
@@
 
-The study was carried out in a nearly 10-ha (300 m*330 m) permanent plot established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996].
-The plot is named after the nearest village, Uppangala, and is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India).
-Annual precipitation averages 5,100 mm per year but is highly variable across seasons with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall of 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011].
-Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [@Pascal1996].
-The geologic substrate can be considered homogeneous within the Uppangala plot [@Ferry1994], but the topography is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig.
+The nearly 10-ha (300 m*330 m) permanent Uppangala plot was established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996] in the Kadamakal Reserve Forest (Karnataka State, India).
+Annual precipitation averages 5,100 mm per year but is highly variable across seasons with most rainfall (90%) occurring during the monsoon from June to October,  while a severe dry season (monthly rainfall of 45 mm) lasts 3.8 months on average [@pelissier_twenty_2011].
+Mean annual temperatureis about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [data from a station at 15 km; @Pascal1996].
+The geologic substrate is homogeneous within the plot [@Ferry1994], but the topography is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig.
 \@ref(fig:Uppangala)).
@@
 
-```{r Uppangala, fig.cap="Digital canopy model of the 10-ha forest permanent plot in Uppangala, India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
+```{r Uppangala, fig.cap="Digital canopy model built from LiDAR data of the 10-ha forest permanent Uppangala plot , India.
+Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
 DCM <- raster("./data/Canopy.tif")
@@
 
-## Taxonomic composition and trait sampling
+## Census data and vertical guilds
 
@@
 
-Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species corresponding to the vertical stratum reached by most adult trees.
+Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species corresponding to the vertical stratum reached by the adult trees.
 The strata included four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m.
-Because our study site is characterized by a steep topography (Fig. \@ref(fig:Uppangala)), 
-sub-canopy species can often reach the canopy in the Uppangala plot. 
-We thereby defined two broader categories, namely, canopy guild (C) including emergent, 
-canopy and sub-canopy species and understory guild (U) composed of understory trees.
+Because our study site is characterized by a steep topography (Fig.
+\@ref(fig:Uppangala)), sub-canopy species can often reach the canopy in the Uppangala plot.
+We thereby defined two broader categories, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.
+
+## Functional trait measurements
 
@@
 For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following the recommendation of @Perez-Harguindeguy2013.
-Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]).
+Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; @Dawkins1958) **comment: beware double space**.
 Leaf and branch samples were kept in humidified ziplock bags filled with enriched $CO_2$ air in dark until measurement within the next 24 hours.
@@
 
-## Trait measurements
-
-We measured five functional traits (Table \@ref(tab:traits)) related to the major dimensions of ecological strategies in plants [@diaz_global_2015; @Wright2002]: leaf thickness (LT, $\mu m$), leaf area (LA, $mm^2$), leaf dry matter content (leaf dry weight divided by the leaf fresh weight, LDMC, $mg.g^{-1}$), specific leaf area (leaf area divided by the leaf dry mass, SLA, $m^2.kg^{-1}$), and branch wood density (wood dry weight divided by its volume, WD, $g.cm^{-3}$).
+We measured five functional traits (Table \@ref(tab:traits)) related to the major dimensions of ecological strategies in plants [@diaz_global_2015; @Wright2002] following the protocol of @Perez-Harguindeguy2013:
+leaf thickness (LT, $\mu m$), leaf area (LA, $mm^2$), leaf dry matter content (leaf dry weight divided by the leaf fresh weight, LDMC, $mg.g^{-1}$), specific leaf area (leaf area divided by the leaf dry mass, SLA, $m^2.kg^{-1}$), and branch wood density (wood dry weight divided by its volume, WD, $g.cm^{-3}$).
 Leaf petioles, rachis and petiolules were removed for all measurements.
-Leaf fresh weight was measured with an analytical balance at a 0.01 g precision (Wensar, KD-H600), leaf fresh thickness with a micrometer and leaf area by  analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012].
+Leaf fresh weight was measured with an digital balance at a 0.01 g precision (Wensar, KD-H600), leaf fresh thickness measured with a micrometer and leaf area by analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012].
 Leaves were then kept in herbarium and oven-dried for at least 72 hours at 60°C before measurement of dry weight.
-The bark and pit of branch samples were removed in order to keep only branch wood without soft tissues.
+We removed the bark and pit of branch samples in order to keep only branch wood without soft tissues.
 Branch wood samples were immersed in water to evaluate their volume by water displacement.
-Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of @Perez-Harguindeguy2013.
+We dried wood samples for 48 hours at 80°C before weighing.
 
@@
     mutate_all(linebreak) %>% 
-    kable(caption = 'Functional traits measured.
-Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).',
+    kable(caption = "Functional traits measured, with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).",
           digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
@@
                         "Intra-individuals", "Intra-species", "Inter-species"),
-        caption = 'Functional traits measured.
-Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).') %>%
+        caption = "Functional traits measured, with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).") %>%
   kable_styling(full_width = T)
@@
 The LiDAR campaign was conducted in 2013, consistent with field measurements.
-Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, related to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the exposure to dominant monsoon winds [@Bunyan2015]; and (iv) slope, which could have both a strong effect on forest dynamics and on the availability of nutrients and water [@ferry2010higher].
+Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, related to local water and soil accumulation; (ii) wetness index, accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), measured as $cos(aspect~in~degree+225)$, to account for the exposure to dominant monsoon winds from south-west [@Bunyan2015]; and (iv) slope, which could have both a strong effect on forest dynamics and on the availability of nutrients and water [@ferry2010higher].
 In addition, based on a 1-m resolution LiDAR model of canopy (Fig.
@@
 To assess the degree of trait variation and covariation, we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific).
-We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.
+We then performed a principal component analysis (PCA) of the average trait values per species to characterize the main dimensions of ecological strategies.
+We compared trait values between species according to the different vertical guilds.
 
-We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot.
-Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.
-We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:
+We subsampled tree communities in $30~m*30~m$ subplots (n=110), and characterized the functional composition of vertical guilds within each subplot.
+Each subplot included 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.
+We calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) **comment: beware double space** for each trait in each subplot:
 
@@
 
-We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in $Trait \times Species$ matrices of each guild; [@bernard-verdier_community_2012]).
-The null model broke the link between environment and local functional composition while maintaining the spatial distribution of guilds and species abundances.
-We derived 999 randomized GWM values and calculated the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as
+We explored the variation in functional composition according to local topography and canopy height for each guild.
+We designed a null model randomizing trait values among species within each guild (permuting rows in $Trait \times Species$ matrices of each guild; [@bernard-verdier_community_2012]) **comment: beware double space**.
+The null model thus broke the link between environment and local functional composition, while maintaining the spatial distribution of guilds and species abundances.
+We derived 999 randomized GWM values and calculated the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution, as
 
@@
 
-We assessed how the deviation of guild composition from the null model depended on environmental variables by performing linear regressions.
-Predictors were all normalised to ease model inference and enable their comparison. 
-The analyses were performed independently for each of the five functional traits.
-A Bayesian method was used to infer parameters using stan language [@Carpenter2017] and rstan package [@Team2018] in the R environment [@RCoreTeam2020].
+We analyzed the variation in functional composition, in terms of SES value, according to local topography and canopy height for each guild, by performing linear regressions.
+Predictors were normalised to ease model inference and enable their comparison.
+We performed analyses independently for each of the five functional traits, and performed Bayesian inference of parameter values using stan language [@Carpenter2017] and rstan package [@Team2018] in the R environment [@RCoreTeam2020].
 
@@
 
-We found an overall increase of functional trait variation from intra-individual to inter-specific levels (Table \@ref(tab:traits)), even though LA showed almost as much variability at intra-individual and at intra-specific levels.
-Thus, trait values averaged at species level could grasp most of the functional variability across guilds and communities with varying taxonomic composition.
+We found an overall increase of functional trait variation from intra-individual to inter-specific levels (Table \@ref(tab:traits)), even though LA showed almost same variability at intra-individual and intra-specific levels.
+Therefore, trait values averaged at species level could grasp most of the functional variability across tree individuals.
 
-The first axis of the PCA performed on species trait values (Fig.
+The first axis of the PCA of species trait values (Fig.
 \@ref(fig:pca)A) was mostly related to wood density and structural leaf traits (LT, LA, LDMC), and opposed species with thick, big leaves and lightwood to species with dense leaves and hardwood.
 The second axis mostly opposed SLA to LDMC.
-WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that WD and SLA provide independent information.
-Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p>0.2, Fig.
+WD was orthogonal to SLA in the plan 1-2 of the PCA, suggesting that WD and SLA provided independent information.
+Canopy and understory species did not show significant differentiation along the first and second PCA axes (Wilcoxon p>0.2, Fig.
 \@ref(fig:pca)B).
@@
 Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test.
-See Table  \\@ref(tab:traits) for trait abbreviations.'}
+See trait abbreviations in Table \\@ref(tab:traits).'}
 pca.plot <- autoplot(PFT.pca, 
@@
 
-At community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig.
+At community level, the guild weighted mean (GWM) of all traits significantly differed in subplots between understory and canopy guild assemblages (paired Student p<0.001,  Fig.
 \@ref(fig:GWMboxplot)).
-GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, lower for LA and higher for LDMC and LT than understory tree assemblages.
+We found lower GWM values of canopy assemblages for SLA and WD with almost no overlap and, to a lesser extent, for LA but higher for LDMC and LT than understory tree assemblages.
 In other words, understory guild assemblages had larger, thinner and lighter leaves with higher SLA and denser woods, compared to canopy guild assemblages.
 
-```{r GWMboxplot, fig.cap="Guild weighted mean (GWM) of specific leaf area (SLA), wood density (WD), leaf dry matter content (LDMC), leaf area (LA) and leaf thickness (LT) for canopy (C) and understory (U) guilds.
+```{r GWMboxplot, fig.cap="Guild weighted mean (GWM) of leaf area (LA) , wood density (WD), leaf thickness (LT), leaf dry matter content (LDMC) and  specific leaf area (SLA) for canopy (C) and understory (U) guilds.
 Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines).
-Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4}
+Asterisks represent significant differences between the two guilds (*** paired Student p-value<0.001).", fig.height=4}
 ggplot(GWM, aes(x = strata, y = GWM, fill = strata)) +
@@
 
-We detected a significant effect of slope, curvature and wetness on both canopy and understory guild assemblages, 
-of southwesterness on canopy guild assemblages 
-and of canopy height mostly on understory guild assemblages (Fig. \@ref(fig:LMres)).
-Curvature appeared as the most influential factor (7 significant effects out of 10 tested), 
-with a negative effect on SLA and a positive effect on LT in both canopy and understory assemblages. 
-In addition, curvature had a positive effect on WD and LDMC only in the understory assemblages,
-and a negative effect on LA only in the canopy assemblages. 
-Similarly to curvature, wetness had a negative effect on SLA in both canopy and understory assemblages,
-and a positive effect on LT in the understory assemblages.
+Slope, curvature and wetness significantly influenced SES values of canopy and understory guild assemblages.
+Southwesterness influenced SES of canopy guild assemblages, while canopy height mostly played on SES of understory guild assemblages (Fig.
+\@ref(fig:LMres)).
+Overall, curvature was the most influential factor (7 significant effects out of 10 tested), with a negative effect on SLA and a positive effect on LT for both canopy and understory assemblages.
+In addition, curvature positively influenced SES of WD and LDMC only in the understory assemblages, and negatively influenced SES of LA only in the canopy assemblages.
+Wetness had a negative effect on SLA in both canopy and understory assemblages, and a positive effect on LT in understory assemblages.
 Still, wetness had a negative effect on LDMC in canopy assemblages.
-As curvature, slope had a negative effect on SLA and a positive effect on LT in canopy assemblages. 
+Slope had a negative effect on SLA and a positive effect on LT in canopy assemblages.
 and a positive effect on WD in the understory assemblages.
-Southwesterness only influenced canopy assemblages,
-with a positive effect on WD and a negative effect on LA. 
-Canopy height mainly influenced understory assemblages,
-with a positive effect on SLA and a negative effect on WD, LDMC and LT.
+Southwesterness only influenced canopy assemblages, with a positive effect on WD and a negative effect on LA.
+Canopy height mainly influenced understory assemblages, with a positive effect on SLA and a negative effect on WD, LDMC and LT.
 Still, canopy height had a positive effect on LA in canopy assemblages.
 
-```{stan LMmodel, output.var="lm"}
+```{stan LMmodel, output.var="lm", cache=F, eval=F}
 data {
@@
 
-```{r LMres, fig.cap="Effects of slope, curvature, wetness, southwesterness, and canopy height on  standardized effect size values of functional traits for canopy and understory guilds.
-The effects of predictors was estimated as the posterior distribution of the slope parameters using Bayesian inference.
-Circles represent the mean estimate, thick lines the 50% confidence interval, and thin lines the 90% confidence interval, and colour the corresponding predictor of the posterior distributions.
-Transparency represents significance, with transparent non-significant posteriors and opaque significant posteriors (p < 0.05).
-See Table \\@ref(tab:traits) for trait abbreviations."}
+```{r LMres, fig.cap="Effects of slope, curvature, wetness, southwesterness, and canopy height on standardized effect size (SES) values of functional traits (display on the left)  for canopy and understory guilds.
+The effects of predictors (shown with different colors) was estimated as the posterior distribution of the slope parameters.
+Circles represent the mean estimate, thick lines the 50% credible interval, and thin lines the 90% credible interval.
+Non-significant posteriors are transparent and significant posteriors (p < 0.05) are opaque .
+See trait abbreviations in Table \\@ref(tab:traits)."}
 mdata <- group_by(SES, strata, trait) %>% 
@@
 
-Few studies have investigated how the functional characteristics and composition of canopy and understorey guilds vary in tropical forests. 
-We here show that functional trait values do not differ globally between canopy and understory tree species but that the contrasting environments between understory and canopy filter species non-randomly,
-resulting in a markedly different functional composition of canopy and understory guilds within subplots. 
-We found that topography and canopy height significantly and differently impacts both canopy and understory guild assemblages.
-Indeed, the functional composition of understory and canopy guilds assemblages respond similarly to water and nutrient access
-but the functional composition of canopy guild communities strongly varies in response to wind exposure, 
-and the functional variation of understory assemblages varies in response to light availability.
-These results show that different environmental gradients simultaneously drive different functional responses among guilds in tropical rainforest.
+Few studies have compared the functional composition of canopy and understorey guilds in tropical forests and their variation along topography and with canopy height.
+Here we found contrasted results among species and among assemblages.
+On one hand, we found broad overlap in average functional trait values between species belonging to canopy and understory guilds.
+On the other hand, the functional composition of assemblages in subplots broadly differed between understory and canopy guilds.
+Furthermore, the guild assemblages responded differently to environmental variation across subplots.
+Functional composition of understory and canopy guild assemblages responded similarly to topographic wetness, but only the canopy guild responded to wind exposure, while functional composition of understory assemblages varied in response to canopy height.
+These results underline that multiple environmental factors play and determine different functional responses of vertical guilds in a tropical rainforest.
+**Comments on last sentence: A bit blur to me, Sylvain.
+Oui, on pourrait probablement se passer de cette phrase, François.
+Je partage cet avis.
+Valérie.**
 
@@
 
-The PCA performed on species functional traits supported two orthogonal main dimensions (Fig. \@ref(fig:pca)).
-SLA and LDMC were opposed into a dimension representing the leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004],
-while, wood density was opposed to leaf area and thickness along a dimension previously interpreted as the wood economics spectrum  [@chave_towards_2009].
-@Baraloto2010 and @Fortunel2012a also found orthogonal dimensions reflecting wood and leaf economic spectra of Neotropical tree species at regional scale in Amazonia.
-In response to these works, @Messier2017 suggested that these two orthogonal spectra reflect large scale differences in tree functional composition and would not be observed among species co-occurring at a local scale.
-Here we captured the two orthogonal spectra across species at the scale of a 10 ha plot despite a limited number of traits compared to previous studies.  
-Our results support that economics spectra are decoupled in tropical forests [@Baraloto2010] and not related to a single, whole-plant economics spectrum [@reich_world-wide_2014].
+The PCA performed on species functional traits supported two orthogonal main dimensions (Fig.
+\@ref(fig:pca)).
+SLA and LDMC were opposed along a dimension representing the leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004], while wood density was opposed to leaf area and thickness along a dimension previously interpreted as the wood economics spectrum [@chave_towards_2009].
+@Baraloto2010 and @Fortunel2012a also found orthogonal dimensions reflecting wood and leaf economic spectra of Neotropical tree species at a regional scale in Amazonia.
+Conversely, @Messier2017 advocated that these orthogonal spectra reflect large scale differences in tree functional strategies, and that the decoupling should not hold among species co-occurring at a local scale.
+Here, however, we captured the two orthogonal spectra across species at the scale of a 10 ha plot despite a limited number of traits compared to previous studies.
+Our results thus support that economics spectra are decoupled in tropical forests [@Baraloto2010] and do not form a single, whole-plant economics spectrum [@reich_world-wide_2014].
 
@@
 
-We found that species pools of canopy and understory species broadly overlap in functional spaces, 
-and both spanned a broad range of ecological strategies (Fig. \@ref(fig:pca)). 
-However, we found significant trait differences between assemblages of canopy and understory species co-occuring in local 30 x 30m communities (Fig. \@ref(fig:GWMboxplot)).
-Environmental conditions thus filtered species in local communities according to their trait, 
-leading to highly contrasted functional compositions between canopy and understory guilds. 
-Understory assemblages had larger, thinner and lighter leaves, hence greater SLA, 
-and denser wood than canopy guild assemblages.
-Greater hydraulic conduction is associated with lighter wood [@Santiago2018],
-which is especially important for canopy species. 
-The conservative leaf strategy observed in canopy assemblages may originate from two mechanisms. 
-First, leaf water stress is expected to increase with tree height due to gravity [@Koch2004]
-and to a higher water evaporation associated with sun and wind exposure [@Chazdon1993]. 
-Second, a higher risk of herbivory exists in the canopy [@Sterck1992],
-which may select for species resistant to herbivory with a more conservative strategy. 
-Drought constraint could be less important for understory trees, 
-allowing greater investment into resource acquisition by leaves to counterbalance the higher need for light in more shaded understory [@Chazdon1996]. 
-However limited light may result in slower growth [@Herault2011] favoring denser wood. 
-These results support the hypothesis that trees experience guild-specific ecological constraints within communities [@duque2002different], 
-leading to contrasting functional compositions of canopy and understory assemblages.
+We found that the species pools of canopy and understory species broadly overlapped in functional space, and both spanned a broad range of ecological strategies (Fig.
+\@ref(fig:pca)).
+Conversely, the functional composition of assemblages within 30 x 30m subplots differed between canopy and understory species (Fig.
+\@ref(fig:GWMboxplot)).
+Environmental conditions filtered species in assemblages according to their functional traits, leading to contrasted functional compositions between canopy and understory guilds.
+Understory assemblages had larger, thinner and lighter leaves, hence greater SLA, and denser wood than canopy guild assemblages.
+Greater hydraulic conduction is associated with lighter wood [@Santiago2018], which is especially important for canopy species.
+The conservative leaf strategy observed in assemblages of canopy species could originate from two mechanisms.
+First, leaf water stress is expected to increase with tree height due to gravity [@Koch2004] and to higher water evaporation associated with sun and wind exposure [@Chazdon1993].
+Second, herbivory can be more intense in the canopy [@Sterck1992], selecting species resistant to herbivory with a more conservative strategy.
+Conversely, drought constraint could be less important for understory trees, allowing greater investment into resource acquisition by leaves, counterbalancing the higher need for light in shaded understory [@Chazdon1996].
+However, limited light can also  entail slower growth [@Herault2011], favoring denser wood.
+These results support the hypothesis that guild-specific ecological constraints determine tree community assembly in a forest ecosystem [@duque2002different], leading to contrasting functional compositions of the assemblages of canopy and understory species.
 
-## Different responses to abiotic and biotic environment between canopy and understory assemblages
+## Different abiotic and biotic predictors of functional composition between canopy and understory assemblages **coments: I see what you mean but I d'ont like this title.
+I would instead introduce directly topography and canopy height instead of predictors.
+Sylvain.
+Your suggestion? Francois.**
 
-Our results show that the two guilds assemblages respond differently to some environmental gradients (Fig. \@ref(fig:LMres)). 
-Curvature, wetness and slope similarly affected canopy and understory functional assemblages, except for LDMC. 
-Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and retain more water, increasing soil fertility [@ferry2010higher].
-It thus creates favorable conditions where species with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA) strategies outcompete others. 
-By contrast, concave, drier or sloppy areas will be drier and less fertile favoring species assemblages with more conservative strategies and slower growth.
+We found that the assemblages of canopy and understory species responded differently to environmental predictors (Fig.
+\@ref(fig:LMres)).
+Curvature, wetness and slope similarly affected the assemblages of canopy and understory species, except for LDMC.
+Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and to retain more water, increasing soil fertility [@ferry2010higher].
+These conditions are suitable for species with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA).
+By contrast, concave, drier or sloppy areas should be drier and less fertile, then filtering species with more conservative strategies and slower growth.
 
-Contrary to understory assemblages, canopy assemblages respond to southwesterness, 
-which face dominant monsoon winds in our study site, 
-along the wood economics spectrum with a greater WD and smaller LA for higher southwesterness. 
+In contrast to understory assemblages, functional composition of canopy trees depended on southwesterness, along the wood economics spectrum, with a greater WD and smaller LA for higher southwesterness.
 The influence of aspect on tropical forest composition has been little documented.
-@gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar climatic and topographical context in Sri Lanka and also interpreted it as a consequence of a higher exposure to monsoon winds.
-The Uppangala plot is characterized by a large topographic heterogeneity,
-with steep slopes mostly organized along southwest-oriented crests (Fig. \@ref(fig:Uppangala)).
-South-west oriented slopes are likely to be more exposed to monsoon winds,
-which could entail greater tree damages and forest disturbance [@soman_aspects_1990]. 
-The consequences could be that 
-(i) tree strategies associated with mechanical resistance and disturbance tolerance could be selected in areas more exposed to monsoon wind damages [@Poorter2010b],
-or (ii) turbulences on northeast oriented areas could yield more disturbance and death events favoring pioneer, lightwood species [@larjavaara_rethinking_2010]. 
-In both cases, canopy species are more exposed to wind disturbances than understory trees and their functional composition is thus expected to respond more to wind exposure.
+@gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar climatic and topographical context in Sri Lanka and interpreted it as a consequence of a higher exposure to monsoon winds.
+The Uppangala plot is characterized by a large topographic heterogeneity, with steep slopes mostly organized along southwest-oriented crests (Fig.
+\@ref(fig:Uppangala)).
+South-west oriented slopes are likely to be more exposed to monsoon winds, which could entail greater tree damages and forest disturbance [@soman_aspects_1990].
+The consequences could be that (i) tree strategies associated with mechanical resistance and disturbance tolerance could be selected in areas more exposed to monsoon wind damages [@Poorter2010b], and/or (ii) turbulences on northeast oriented areas could yield more disturbance and death events favoring pioneer, lightwood species [@larjavaara_rethinking_2010].
+In both cases, canopy species are more exposed to wind disturbances than understory trees, and their functional composition is thus expected to be more responsive to wind exposure.
 
-Conversely, we found pervasive influence of canopy height on the functional composition of understory assemblages related to leaf and wood economics, with increased acquisitive and growth strategies under tall canopies.
-We indeed found a greater investment into photosynthetic activity and growth, with higher SLA and lower LDMC, LT and WD, in understory assemblages under taller canopy. 
-We expected canopy height to represent decreased light-availability contrasting with the observed functional strategies expected under increased access to light [@kitajima_tissue-level_2010]. 
-As canopy height is as dynamic as individual recruitment of current species assemblages,
-we hypothesized taller communities to result from more recent gaps with an increased abundance of acquisitive and fast growing post-pioneer species in understory.
+Conversely, we found pervasive influence of canopy height on the functional composition of understory assemblages for both leaf and wood economics, with increased acquisitive and growth strategies under taller canopies.
+Higher SLA and lower LDMC, LT and WD, suggested  greater investment into photosynthetic activity and growth in assemblages of understory trees under taller canopy.
+We expected canopy height to represent decreased light availability contrasting with the observed functional strategies expected under increased access to light [@kitajima_tissue-level_2010].
+As canopy height is as dynamic as individual recruitment of current species assemblages, we hypothesized taller communities to result from more recent gaps with an increased abundance of acquisitive and fast growing post-pioneer species in understory.
 This hypothesis needs to be thoroughly tested with growth or canopy time series.
+**Comments from We expected canopy: Dans la mesure où les très grands arbres sont des émergents, on peut supposer que la canopée est moins "closed" horizontalement (on en avait déjà parlé).
+François.
+Je partage l'avis de François.
+Valérie.
+Formulation/signification peu claires.
+François.
+Pas sûr de comprendre ce que tu veux dire.
+François.**
 
@@
 Our results show that functional variations among species and assemblages vary along the decoupled economic spectra of leaves and wood in an Indian rainforest.
-We found a significant effect of topography and canopy height on the functional composition of tree communities, 
-but also pointed out that these effects may be guild specific.
-Our results argue in favour of decomposing the functional signature of community assemblage
-in order to better characterize and understand the different strategies and responses of ecological guilds [@Raevel2018]
-and the vertical structure of forests [@duque2002different]. 
-We therefore call for a better recognition of the different dynamics of guilds within communities and their influence on the global functioning of forests.
+We found a significant effect of topography and canopy height on the functional composition of tree communities, but also pointed out that these effects change with vertical guilds.
+Our results suggest that forest community assembly is better understood as a combination of different, guild-specific assembly rules [@Raevel2018], representing different environmental constraints across vertical forest strata [@duque2002different].
+Therefore, we call for designing trait-based studies at the level of ecological guilds coexisting at a given place [@Laroche2020].
+We expect that decomposing guild dynamics should contribute to better understanding and predict global functioning of forests.
 
@@
 Access to LiDAR data is restricted by the Indian government and should be the object of a convention (contact: chandra.s.jha@gmail.com).
+**Comment: Demander explicitement a Jha si ca lui convient quand tu enverras a tous les co-auteurs.**
 

### 2021-08-26 Manuscript.Rmd
Final version accepted in JVS

@@
 library(bayesplot)
+library(ggmap)
 theme_set(bayesplot::theme_default())
@@
                cache = T, cache.lazy = F)
-cv <- function(x, na.rm = T)
-  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
 n = 999 # Random repetitions
 sapply(list.files("R", full.names = T), source)
+```
+
+```{r, eval=FALSE}
+Env <- genEnv()
+Trees <- genTrees()
+Trees$ID <- row.names(Trees)
+# Trees <- merge(Trees, PFT_sp)
+row.names(Trees) <- Trees$ID
+Trees <- Trees[-which(names(Trees) == 'ID')]
+com <- quadrats(cs = c(30,30))
+XY <- Trees[c('x', 'y')]
+coordinates(XY) <- ~ x + y
+proj4string(XY) <- crs(com)
+Trees$com <- (XY %over% com)[,1]
+com@data$Elevation <- raster::extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/bin/')
+names(Deriv)[5] <- 'SW'
+com@data <- cbind(com@data, raster::extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
+com@data$Wetness <- raster::extract(DEMderiv(Env$DEM, 'wetness', path = '/usr/bin/'),
+                                    as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+com@data$Canopy <- raster::extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
+BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
+com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
+R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
+row.names(R) <- com@data$id
+R <- rownames_to_column(R, "community")
+save(R, file = file.path("draft_save", 'R2.Rdata'))
 ```
@@
   group_by(community, SpCode) %>% 
-  summarise(Abundance = n())
+  dplyr::summarise(Abundance = n())
 # Matrix R
-load(file.path("draft_save", 'R.Rdata'))
+# load(file.path("draft_save", 'R.Rdata'))
+load(file.path("draft_save", 'R2.Rdata'))
 # GWM
@@
 # GWM plot
-left_join(GWM, R) %>% 
-  reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
-  filter(variable != "BA") %>% 
-  ggplot(aes(value, GWM, col = strata)) +
-  facet_grid(trait ~ variable, scales = "free") +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  theme(legend.position = "bottom", axis.title = element_blank())
+# left_join(GWM, R) %>% 
+#   reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
+#   filter(variable != "BA") %>% 
+#   ggplot(aes(value, GWM, col = strata)) +
+#   facet_grid(trait ~ variable, scales = "free") +
+#   geom_point() +
+#   geom_smooth(method = "lm") +
+#   theme(legend.position = "bottom", axis.title = element_blank())
 # nGWM
@@
 #   theme(legend.position = "bottom", axis.title = element_blank())
+
+```
+
+```{r, eval=FALSE}
+trees %>% 
+  left_join(PFT) %>% 
+  filter(!is.na(strata)) %>% 
+  dplyr::group_by(strata, community, trait) %>% 
+  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
 ```
@@
           "depending on WD category (low < 0.5 and high > 0.7).")
+
+trees %>% 
+  group_by(community, SpCode) %>% 
+  summarise(Abundance = sum(Abundance)) %>% 
+  left_join(select(PFT,SpCode, strata)) %>% 
+  unique() %>% 
+  filter(!is.na(strata)) %>% 
+  group_by(community, strata) %>% 
+  summarise(N = sum(Abundance), S = length(unique(SpCode))) %>% 
+  group_by(strata) %>% 
+  # summarise(N = sd(N), S = sd(S))
+  summarise(N = mean(N), Nsd = sd(N),
+            S = mean(S), Ssd = sd(S))
+  # group_by(strata, community) %>% summarise(N = sum(Abundance), S = length(unique(SpCode))) %>% group_by(strata) %>% summarise(N = mean(N), S = mean(S))
+```
+
+```{r, eval=F}
+data <- genPFT() %>% 
+  mutate(LA = log(LA)) %>% 
+  mutate(SLA = log(SLA)) %>% 
+  select(-CE) %>% 
+  mutate(LMA = 1/SLA, LT.LDMC = LT*LDMC) %>% 
+  group_by(Tree, Sp, SpCode) %>% 
+  summarise_all(mean, na.rm = T)
+summary(lm(LMA ~ LT.LDMC, data))
+ggplot(data, aes(LT.LDMC, LMA)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  ggtitle("LMA = 0.30 + 0.17 LT x LDMC",
+          "r2 = 0.48; n = 267; P < 0.001") +
+  xlab("LT x LDMC")
+
 ```
@@
 colnames(DCM) <- c("Height", "x", "y")
-ggplot() +  
+d <- 12
+lng <- 75.662778
+lat <- 12.537500
+bbox <- c(left = lng-d, bottom = lat-d, right = lng+d, top = lat+d)
+map <- get_stamenmap(bbox, zoom = 5)
+g <- ggplot() +  
   geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
@@
   coord_equal() +
-  cowplot::theme_map() +
+  theme(axis.line = element_blank(), legend.position = c(0.9, 0.5), axis.text = element_text(face = "bold"),
+        legend.background = element_rect(fill = "white", colour = NA)) +
   scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
@@
                  transform = F, model = "WGS84", location = "bottomleft") +
-    annotation_custom(grob = ggplotGrob(ggplot() +  
-                                        geom_polygon(data = fortify(india),
-                                                     aes(long, lat, group=group),
-                                                     colour = "black", alpha = 0) +
-                                        geom_point(data = data.frame(long = 75.662778, lat = 12.537500), 
-                                                   aes(long, lat), col = "red", size = 3) +
-                                        coord_equal() +
-                                        cowplot::theme_map() +
-                                        theme(panel.border = element_rect(color = "black",size = .5,fill = NA),
-                                              plot.background	= element_rect(fill = "white"))), 
-                    xmin = min(DCM$x), xmax = min(DCM$x) + 150, 
-                    ymin = max(DCM$y) - 150, ymax = max(DCM$y))
+  ggsn::north(symbol = 10, x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 30, y.max = max(DCM$y)-10) +
+  xlab("Longitude") + ylab("Latitude") +
+  scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0)) +
+  annotation_custom(grob = ggplotGrob(
+    ggmap(map) +
+      geom_point(x = lng, y = lat, colour = "red", size = 4) +
+      theme(axis.title = element_blank(), axis.line = element_blank(), 
+            axis.ticks = element_blank(), axis.text = element_blank(),
+            panel.border = element_rect(fill = NA))
+  ),
+  xmin = min(DCM$x), xmax = min(DCM$x) + 200,
+  ymin = max(DCM$y) - 200, ymax = max(DCM$y))
 # ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
+ggsave(g, filename = "figures/Figure_1.tiff", device = "tiff", dpi = 600, bg = "white", width = 8, height = 8, compression = "lzw")
 ```
@@
 See trait abbreviations in Table \\@ref(tab:traits).'}
-pca.plot <- autoplot(PFT.pca, 
+# pca.plot <- autoplot(PFT.pca, 
+#          data = na.omit(reshape2::dcast(PFT, strata + SpCode ~ trait, value.var = "traitvalue")),
+#          colour = "strata",
+#          loadings.label.size = 6,
+#          loadings.label.colour = 'black',
+#          loadings = T, loadings.label = T, loadings.colour = 'black',
+#          loadings.label.vjust = 1.2) +
+#   coord_equal() +
+#   geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
+#   geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
+#   xlab('Axis 1') + ylab('Axis 2') +
+#   theme(legend.position = c(1, 1), 
+#         legend.justification = c(1, 1)) +
+#   scale_color_discrete("Guild")
+# spectrum.boxplots <- pca.plot$data %>%
+  # select(Comp.1, Comp.2, strata) %>%
+  # rename(`Axis 1` = Comp.1, `Axis 2` = Comp.2) %>%
+  # melt(id.vars = "strata", variable.name = "spectrum") %>%
+  # ggplot(aes(strata, value, fill = strata)) +
+  # # ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
+  # geom_boxplot() +
+  # facet_wrap(~spectrum, nrow = 2) +
+  # scale_fill_discrete("Guild")
+g <- autoplot(PFT.pca, 
          data = na.omit(reshape2::dcast(PFT, strata + SpCode ~ trait, value.var = "traitvalue")),
@@
   geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
-  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
+  stat_ellipse(aes(group = strata, col = strata), linetype = "dashed") +
+  xlab('Axis 1') + ylab('Axis 2') +
   theme(legend.position = c(1, 1), 
-        legend.justification = c(1, 1)) +
+        legend.justification = c(1, 1),
+        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
   scale_color_discrete("Guild")
-spectrum.boxplots <- pca.plot$data %>%
-  select(Comp.1, Comp.2, strata) %>%
-  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
-  melt(id.vars = "strata", variable.name = "spectrum") %>%
-  ggplot(aes(strata, value, fill = strata)) +
-  # ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
-  geom_boxplot() +
-  facet_wrap(~spectrum, nrow = 2) +
-  scale_fill_discrete("Guild")
-cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
+ggsave(g, filename = "figures/Figure_2.tiff", device = "tiff", dpi = 600, bg = "white", width = 8, height = 8, compression = "lzw")
 ```
@@
 Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines).
-Asterisks represent significant differences between the two guilds (*** paired Student p-value<0.001).", fig.height=4}
-ggplot(GWM, aes(x = strata, y = GWM, fill = strata)) +
+Asterisks represent significant differences between the two guilds (*** paired Student p-value<0.001).", fig.height=4, fig.width=10}
+g <- GWM %>% 
+  mutate(trait = recode(trait, 
+                        "SLA" = "A", 
+                        "LDMC" = "B",
+                        "LT" = "C", 
+                        "LA" = "D",
+                        "WD" = "E")) %>% 
+  mutate(trait = factor(trait, levels = c("A", "B", "C", "D", "E"))) %>%
+  ggplot(aes(x = strata, y = GWM, fill = strata)) +
   geom_boxplot() +
-  facet_wrap(~trait, scales = "free", nrow = 1) +
-  xlab("Guild") + scale_fill_discrete(guide = "none") +
-  ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3))))
+  geom_text(aes(label = label), label.size = 0.2,
+            data.frame(
+              strata = "C",
+              trait = LETTERS[1:5],
+              label = c("A SLA", "B LDMC", "C LT", "D LA", "E WD"),
+              GWM = c(2.8, 435, 28, 10, 0.7)
+            )) +
+  facet_wrap(~trait, scales = "free", nrow = 1,
+             strip.position = "left", 
+             labeller = as_labeller(c(A = "SLA (m²/kg)", B = "LDMC (mg/g)",
+                                      C = "LT (µm)", D = "LA (mm²)", E = "WD (g/cm³)") ) )  +
+  ylab(NULL) +
+  theme(strip.background = element_blank(),
+        strip.placement = "outside") +
+  xlab("Guild") + scale_fill_discrete(guide = "none")
+ggsave(g, filename = "figures/Figure_3.tiff", device = "tiff", dpi = 600, bg = "white", width = 10, height = 4, compression = "lzw")
 ```
@@
 See trait abbreviations in Table \\@ref(tab:traits)."}
-mdata <- group_by(SES, strata, trait) %>% 
-  mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), scale) %>% 
-  group_by(strata, trait) %>% 
-  do(mdata = list(N = nrow(.), K = 5, SES = .$SES, X = as.matrix(.[c("Slope", "Curvature", "Wetness", "SW", "Canopy")])))
-fits <- lapply(mdata$mdata, function(m) sampling(lm, data = m))
-names(fits) <- paste0(mdata$strata, "_", mdata$trait)
+# mdata <- group_by(SES, strata, trait) %>%
+#   mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), scale) %>%
+#   group_by(strata, trait) %>%
+#   do(mdata = list(N = nrow(.), K = 5, SES = .$SES, X = as.matrix(.[c("Slope", "Curvature", "Wetness", "SW", "Canopy")])))
+# fits <- lapply(mdata$mdata, function(m) sampling(lm, data = m))
+# names(fits) <- paste0(mdata$strata, "_", mdata$trait)
+# save(fits, file = "data/fits.Rdata")
+load(file = "data/fits.Rdata")
 R2 <- lapply(fits, mcmc_intervals_data, regex_pars = "R2") %>% 
@@
   dplyr::select(Guild, Trait, R2) %>% 
-  mutate(Variable = "Canopy height", Significance = 1) %>% 
-  mutate(R2 = paste0("R²=", round(R2, 2)))
-lapply(fits, mcmc_intervals_data, regex_pars = "beta") %>% 
+  mutate(Variable = "Canopy height", Significance = "significant") %>% 
+  mutate(R2 = paste0("R²=", round(R2, 2))) %>% 
+   mutate(label = paste(Trait, "-", Guild)) %>% 
+  mutate(label = recode(label,
+                        "LA - Canopy" = "A LA - Canopy",
+                        "LA - Understory" = "B LA - Understory",
+                        "LDMC - Canopy" = "C LDMC - Canopy",
+                        "LDMC - Understory" = "D LDMC - Understory",
+                        "LT - Canopy" = "E LT - Canopy",
+                        "LT - Understory" = "F LT - Understory",
+                        "SLA - Canopy" = "G SLA - Canopy",
+                        "SLA - Understory" = "H SLA - Understory",
+                        "WD - Canopy" = "I WD - Canopy",
+                        "WD - Understory" = "J WD - Understory"
+                        )) 
+t <- lapply(fits, mcmc_intervals_data, regex_pars = "beta") %>% 
   bind_rows(.id = "model") %>% 
@@
                            "C" = "Canopy")) %>% 
-  mutate(Significance = 0) %>% 
-  mutate(Significance = ifelse(m > 0 & ll > 0, 1, Significance)) %>% 
-  mutate(Significance = ifelse(m < 0 & hh < 0, 1, Significance)) %>% 
+  mutate(Significance = "non significant") %>% 
+  mutate(Significance = ifelse(m > 0 & ll > 0, "significant", Significance)) %>% 
+  mutate(Significance = ifelse(m < 0 & hh < 0, "significant", Significance)) %>% 
+  mutate(Var2 = ifelse(Significance == "significant", as.character(Variable), "A"))
+
+g <- t %>% 
+  mutate(label = paste(Trait, "-", Guild)) %>% 
+  mutate(label = recode(label,
+                        "LA - Canopy" = "A LA - Canopy",
+                        "LA - Understory" = "B LA - Understory",
+                        "LDMC - Canopy" = "C LDMC - Canopy",
+                        "LDMC - Understory" = "D LDMC - Understory",
+                        "LT - Canopy" = "E LT - Canopy",
+                        "LT - Understory" = "F LT - Understory",
+                        "SLA - Canopy" = "G SLA - Canopy",
+                        "SLA - Understory" = "H SLA - Understory",
+                        "WD - Canopy" = "I WD - Canopy",
+                        "WD - Understory" = "J WD - Understory"
+                        )) %>% 
   ggplot(aes(x = Variable, xend = Variable,
-             color = Variable, fill = Variable,
-             alpha = as.factor(Significance))) +
+             color = Var2)) +
   geom_hline(yintercept = 0, color = "gray90", size = 1) +
@@
              color = "gray90", size = 0.5, lty = "dashed") +
-  geom_point(aes(y = m), shape = 21, size = 3) +
+  geom_point(aes(y = m, fill = Var2), shape = 21, size = 3) +
   geom_segment(aes(y = ll, yend = hh),
@@
   coord_flip() +
-  facet_grid(Trait ~ Guild, labeller = label_parsed, 
-              scales = "free_y") +
+    facet_wrap(~ label,
+              scales = "free_y", ncol = 2
+              ) +
   xaxis_title(F) +
   ylab("Standardised effect size") +
-  scale_color_discrete(guide = "none") +
-  scale_fill_discrete(guide = "none") +
-  scale_alpha_manual(guide = "none", values = c(0.3, 1))
+  scale_color_manual(guide = "none", values = c("grey", RColorBrewer::brewer.pal(5, "Set1"))) +
+  scale_fill_manual(guide = "none", values = c("grey", RColorBrewer::brewer.pal(5, "Set1"))) +
+  theme(legend.position = "bottom", strip.text.x = element_text(hjust = -0.01))
+ggsave(g, filename = "figures/Figure_4.tiff", device = "tiff", dpi = 600, bg = "white", width = 10, height = 4, compression = "lzw")
 ```

### 2021-08-26 R/topo_deriv.R

@@ -87,7 +87,7 @@ DEMderiv = function(DEM, choices = c('slope', 'curvature', 'plancurvature', 'pro
       Deriv <- stack(Deriv, Aspect)
     }
     if('cosaspect' %in% choices){
-      Cosaspect <- cos((Aspect + dir) / 180 * pi)
+      Cosaspect <- sin((Aspect + dir) / 180 * pi)
       names(Cosaspect) <- 'Cosaspect'
       Deriv <- stack(Deriv, Cosaspect)
     }
